### 2012-09-05 sociology_paper.tex
first commit

%\documentclass{article}
\documentclass[10pt]{article}
\usepackage{times}
%\usepackage{natbib}
%\usepackage{multicol}
\RequirePackage{natbib}
\usepackage{amsmath, amssymb, fullpage, amsthm, array, algorithm2e,graphicx,asa}
%\usepackage[dvips]{graphics}

%\usepackage{hyperref} % for hyper reference

\graphicspath{{images/}}

% \usepackage{pifont} % this package is used to print check mark \checkmark
% \linespread{1.6} % factor 1.6 = double space

%\usepackage{setspace, url}
%\doublespacing

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-0.4in}
\setlength{\textheight}{9in}
\evensidemargin 
\oddsidemargin

\newtheorem{thm}{Theorem}[section]
\newtheorem{dfn}{Definition}[section]
\newtheorem{cor}{Corollary}[thm]
\newtheorem{con}{Conjecture}[thm]
\newtheorem{lemma}[thm]{Lemma}

%\topmargin -0.10in   % when making pdf
%\textheight 9.15in  % when making pdf

\pdfminorversion=4 % as instructed by JASA file upload

\begin{document}

% Article top matter
\title{Impact of Demographics and Skills of the Observer on Visual Statistical Inference }
\author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
\thanks{Mahbubul Majumder is a PhD student (e-mail: mahbub72@gmail.com) , Heike Hofmann is an Associate  Professor and Dianne Cook is a Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210. This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
\date{\vspace{-.5in}}
%\date{\today}  %\today is replaced with the current date
\maketitle

\begin {abstract}  
Visual Inference is dependent on the careful evaluation of the lineups by individual observers. Each individual is different in their cognitive phycology and judiciousness which can affect the power of visual inference. To estimate the power of visual inference this can be controlled by getting evaluations from multiple people from a diverse population. But the other factor that may affect the power of visual inference is the demographics and skills of the observer. In this paper we examine this in details. The simulation experiments suggest that individual skill as well as demographics are very significant for the power of visual inference.

{\bf Keywords: \sf statistical graphics, lineup, non-parametric test, cognitive phycology, visualization, exploratory data analysis} 
\end {abstract}

%\begin{multicols}{2}
%\twocolumn

\section{Introduction}

The fundamental concept of visual statistical inference is introduced by \citet{buja:2009} and later these concepts are validated by \citet{majumder:2012}.

\subsection{Visual Inference} A short introduction of visual inferential procedure.
\subsection{Estimation of Power of Visual Inference} Discussion of the methods for estimating the power of visual inference \citep{majumder:2012}.

\section{Factors Affecting the Power of Visual Inference} A brief description of the factors that may affect the power of visual inference.

\subsection{Choice of Visual Test Statistic} Discuss the effective feature of visual test statistics \citep{heike:2012}. May present some results from \textit{Distance measure} of visual test statistics \citep{niladri:2012}.
\subsection{Question that Human Observer Faces}
\subsection{Observer Personality}
\subsection{Visual Perception of Human Eye} Discuss the results we have on eye tracking experiment \citep{zhao:2012}.
\subsection{Demographics of Observer}
\subsection{Skills of Observer}

\section{Simulation Experiment}

\subsection{Experiment Setup}
\subsection{Diversity of the Experimental Subjects}

\section{Results}
\subsection{Overview of the Data}
\subsection{Learning Trend of the Observer}

\section{Conclusion}

\bibliographystyle{asa}
\bibliography{references}

\end{document}

%\checkmark

Statistical graphics nourish the discovery process in data analysis by revealing unexpected things,  finding structure that was not previously anticipated,  or orthogonally by contrasting prevailing hypotheses. The area of graphics is often associated with exploratory data analysis, which was pioneered by \cite{tukey:eda} and, is particularly pertinent in today's data-rich world where discovery during data mining has become an important activity. Graphics are also used in many places where numerical summaries simply do not suffice: model checking, diagnosis, and in the communication of findings.

Statistical graphics research and development is agreeably established now, and has been cemented by several new developments in recent years. Early studies on evaluating how well statistical plots are perceived and read by the human eye \citep{cleveland:1984}, have been repeated and expanded \citep{simkin:1987,spence:1991, heer:2010} with findings supporting the original results. This body of work provides a framework for evaluating new statistical graphics.  In a complementary direction, new research on formalizing statistical graphics with language characteristics makes it easier to abstractly define, compare and contrast data plots. \cite{wilkinson:1999} developed a grammar of graphics that is enhanced by \cite{hadley:2009}. These methods provide a mechanism to abstract the way data is mapped to graphical form. Finally, technology advances make it simple and easy for everyone to draw plots of data, and particularly the existence of software systems, such as R \citep{R}, enable making beautiful data graphics that can be tightly coupled with statistical modeling.

However, measuring the strength of patterns seen in plots, and differences in individual perceptual ability, is something that is difficult and perhaps handicaps graphics  use among statisticians, where measuring probabilities is of primary importance. This has also been addressed in recent research. \citet{buja:2009} proposes a protocol that allows the testing of discoveries made from statistical graphics. This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis. One of the protocols, the lineup, places the actual data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different. Figure \ref{fig:test_category} shows an example lineup. Which plot do you think is the most different from the others? (The position of the actual data plot is provided in Section \ref{sec:category}.) Wrapped in a process that mirrors conventional inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis. The null hypothesis typically derives from the task at hand, or the type of plot being made. The alternative encompasses all possible antitheses, all types of patterns that might be detected in the actual data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time. The probability of rejection can be quantified, along with Type I, and Type II error, and $p$-value and power can be defined and estimated.

\begin{figure}[htp]
%\begin{figurehere}
   \centering
       \includegraphics[width=0.95\textwidth]{lineup_category.pdf}
       \caption{Lineup plot ($m=20$) using side-by-side boxplots for testing $H_0: \beta_k=0$. One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true. Which plot is the most different from the others, in the sense that there is the largest shift or location difference between the boxplots? (The position of the actual data plot is provided in Section \ref{sec:category}.)}
       \label{fig:test_category}
\end{figure}
%\end{figurehere}

The protocol has only been informally tested until now. In the work described in this paper, the lineup protocol is compared head to head with the equivalent conventional test. Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model. This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario for visual inference. The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test. The thought is though, that the conventional setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark. Testing is done based on a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups. The results are compared with those of the conventional test.

%, analysis requires plots of data. For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics. Even though we have established mathematical procedures to obtain various statistics, we need to support the results by also producing the relevant plots.

%The scientific foundation of graphical methods for data analysis is well established by \cite{cleveland:1984}. In recent years we have seen several major advances in statistical graphics. Modern computing systems like R and SAS ease the production of high quality statistical plots. A grammar of graphics introduced by \cite{wilkinson:1999} presents a structured way to generate specific graphics from data and define connections between disparate types of plots.  \cite{hadley:2009} has implemented a revised version of the grammar of graphics in R, in the package {\tt ggplot2}.   \blue {Graham Wills has implemented the grammar in SPSS, but I'm not sure how to cite this.}

%\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics. This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis.

%In this paper we compare the lineup protocol to conventional statistical inference, specifically applied to the regression framework. We also present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.

The paper is organized as follows. Section \ref{sec:visual_test} defines terms as used in visual inference, and describes how to estimate the important quantities from experimental data. The effect of the lineup size and number of observers on the power of the test is discussed in Section \ref{sec:size}. Section \ref{sec:regression} focuses on the application of  visual inference to linear models.  Section \ref{sec:simulation} describes  three  user studies based on simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and  Section \ref{sec:results} presents an analysis of the resulting data. %The last section describes the next steps in this research.

\section{Definitions and Explanations for Visual Statistical Inference} \label{sec:visual_test}

Table \ref{tbl:compare} illustrates the lineup protocol in relation to conventional hypothesis testing. Both start from the same place, the same set of hypotheses. The conventional test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error. In the lineup protocol, the test statistic is a plot of the data. Here, side-by-side boxplots are used, because the variable of interest is categorical and takes just two values. In conventional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true. If it is extreme on this scale then the null hypothesis is rejected. In contrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution. If the actual data plot is selected as the most different, then this results in rejection of the null hypothesis.

\begin{table*}[hbtp]
\caption{Comparison of visual inference with conventional inference.}
\centering 
\begin{tabular}{llll} 
\hline \hline
  & Conventional Inference &  Lineup Protocol \\ %[0.5ex] % inserts table %heading 
\hline
  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta > 0$& $H_0: \beta=0$ vs $H_1: \beta > 0$\\
 & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\

 Test statistic & $T(y)=\frac{\hat{\beta}}{se(\hat{\beta})}$ & $T(y)=$ \begin{minipage}[h]{1cm} \begin{center} \scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} \\

 & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\

 Sampling Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.32}{\includegraphics{lineup_category_small.pdf}} \end{center} \end{minipage} \\
 & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
 Reject $H_0$ if & actual $T$ is extreme & actual plot is identifiable \\
\hline 
\end{tabular}
\label{tbl:compare}
\end{table*}

%Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.

%Power of a plot and $p$-value have been derived in Buja and Wickham as ...

%for this paper, we want to place the parametric inference of a regression setting into the framework of visual testing:

%This section outlines the concepts of visual inference in comparison to the procedures of conventional statistical inference. Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.

In general, we define $\theta$ to be a population parameter of interest, with $\theta \in \Theta$, the parameter space. Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$. A test statistic, $T(y)$, is a function that maps the sample into a numerical summary, that can be used to test the null hypothesis. The hypothesis test maps the test statistic into \{0, 1\}, based on whether $T(y)$ falls into the acceptance region, or the rejection region, respectively. $T(y)$ is assessed relative to null values of this statistic $T(y_0)$ which would be possible values of $T$ if $\theta \in \Theta$.  %Similarly, for any test statistic, we have a rejection region $R$ and, complementary to that, $R^c$,  that partitions the space of the test statistic, which is usually a subspace of the real numbers. 
For visual inference, unlike in the conventional hypothesis test, the statistic is not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$. When the alternative hypothesis is true, it is expected that the plot of the actual data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true. We will call a plot with this property a {\it visual statistic} for $\theta$. More formally,

\begin{dfn} \label{dfn:test}
A \textbf{visual test statistic}, $T(.)$, is a function of a sample that produces a plot. $T(y)$ maps the actual data to the plot, and we call this the \textbf{(actual) data plot}, and $T(y_0)$ maps a sample drawn from the null distribution into the same plot form. These type of plots are called \textbf{null plots}. 
\end{dfn}

\noindent Ideally, the visual test statistic is defined and constructed using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility. \\
\noindent
The visual test statistic is compared with values $T(y_0)$ using a lineup, which is defined as:

\begin{dfn}\label{dfn:lplot}
A \textbf{lineup} is a layout of $m$ randomly placed visual statistics, consisting of 
\begin{itemize}\itemsep-3pt
\item $m-1$ statistics, $T(y_0)$, simulated from the model specified by $H_0$  (null plots) and 
\item the test statistic, $T(y)$, produced by plotting the actual data, possibly arising from $H_1$.
\end{itemize}
\end{dfn}

\noindent The $(m-1)$ null plots are % can be 
members of the sampling distribution of the test statistic assuming that the null hypothesis is true. If $H_1$ is true, we expect this to be reflected as a feature in the test statistic, i.e. the plot of the data, that makes it visually distinguishable from the null plots. A careful visual inspection of the lineup by %\strikethrough{an} 
independent observers follows;  observers are asked to point out the plot most different from the lineup. If the test statistic is identified in the lineup, this is considered as evidence against the null hypothesis. 
%If it is not identified the conclusion is to not reject the null hypothesis.

This leads us to a definition for the $p$-value of a lineup: under the null hypothesis, each observer has a $1/m$ chance of picking the test statistic from the lineup.  For $K$ independent observers, let $X$ be  the number of observers picking the test statistic from the lineup. Under the null hypothesis $X \sim \text{Binom}_{K, 1/m}$,  therefore:

\begin{dfn}\label{dfn:pvalue}
The $p$-value of a lineup of size $m$ evaluated by $K$ observers is  given as

\begin{equation}\label{binom}
P(X \ge x) = 1 - \text{Binom}_{{K, 1/m}} (x-1) = \sum_{i=x}^{K} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i}
\end{equation}

\noindent with $X$ defined as above, and $x$ is the number of observers selecting the actual data plot. 
\end{dfn}
\noindent Note that for $x=0$ the $p$-value becomes, mathematically, equal to 1. It might make more sense from a practical point of view to think of the $p$-value as
being larger than $P(X \ge 1)$ in this situation. If we increase either $m$ or $K$, we would be able to determine the value at a higher precision.

Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$. It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the actual data plot.

\begin{table}[htp]
\caption{Possible $p$-values for different numbers of observers, $K$, for fixed size $m = 20$ lineups.}
\begin{center}
\scalebox{.9}{
\begin{tabular}{|rrr|rrr|rrr|rrr|rrr|}\hline
$K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\ \hline %\cline{1-3}\cline{4-6}\cline{7-9}\cline{10-12}\cline{17-19}
1 &  1 & 0.0500 & 2 &  1 & 0.0975 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
&&& 2 &  2 & 0.0025 & 3 & 2 & 0.0073 & 4 & 2 & 0.0140 & 5 & 2 & 0.0226 \\
&&& &&& 3 & 3 & 0.0001 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
&&&  &    &           &   &        && 4 & 4 & $< 0.0001$ & 5 & 4 & $< 0.0001$ \\\hline
\end{tabular}
}
\end{center}
\label{pvalue}
\end{table}

%\begin{dfn}\label{dfn:test}
%\red{why do we need the $\theta$? - it's not defined and not explained. usually we would assume that lineups are non-parametric}
%The \textbf{visual test}, $V_{\theta}$, is defined to be 
%\begin{itemize}\itemsep-3pt
%\item \textbf{Reject} $H_0$ if `enough' observers select the actual data plot in the lineup of $m$ plots, and
%\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
%\end{itemize}
%where `enough' observers is defined as some threshold $x_{1-\alpha}$, for which $P(X \ge x_{1-\alpha}) < \alpha$, the level of significance.
%\end{dfn}

%\blue{To avoid confusion with one tailed $(1-\alpha)$ vs two tailed $(1-\alpha/2)$ notation, can we just use $x_{\alpha}$? }

%\red{alternative definition of visual test where $type-I$ error $\alpha$ is attached.}

\begin{dfn}\label{dfn:test}
The \textbf{visual test}, $V_{\theta}$  of  size $m$ and significance level $\alpha$, is defined to  
\begin{itemize}\itemsep-3pt
\item \textbf{Reject} $H_0$ if out of $K$ observers at least $x_{\alpha}$ correctly identify the actual data plot in the lineup, and
\item \textbf{Fail to reject} $H_0$  otherwise. 
\end{itemize}
where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
\end{dfn}

\noindent
Associated with any test is the risk of  Type I or II errors, which for visual inference are defined as follows:

\begin{dfn}\label{dfn:error}
The \textbf{Type I error} associated with visual test $V_{\theta}$ is the probability of rejecting $H_0$ when it is true; the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.\\ \noindent
The \textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
\end{dfn}

%\green{Oh, we are using observed and observer, which might be confusing. May need to change observed data plot to something else. Will think about this. Will use ``actual'' for now.} \hh{I think that change is done}

\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$. %, with an exact $x_{\alpha}$. 
For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers. 
Type II error is harder to calculate, as is usually the case. In visual inference, individual abilities need to be incorporated to calculate Type II error. %This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance. 
Here, we need to estimate the probability that an observer sees the actual data plot as different, when it really is different. This involves understanding the individual's visual skills. Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i \ (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise. Let $p_i$ be the probability that individual $i$ picks out the actual data plot from the lineup. If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.

If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p_i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$. This makes the distribution of $X$ a Poisson-Binomial by definition (see \citet{butler93} for details). Ways to estimate $p_i$ will be  discussed in the following sections.
%one of the main concerns for the remainder of the paper. One way we can estimate this is fitting a generalized mixed effect model as described in Section \ref{sec:model}.

%\green{WHY??? And how does this relate to the logistic model below?} 
%This leads to the definition of the power of the visual test, as follows:

%The lineup plot can be evaluated by one or more individuals. When a single individual identifies the observed graph in the lineup plot we report a $p$-value of at most $1/m$, otherwise the $p$-value is at least $1-\frac 1m$.

 %Notice that when $N=1$, this $p$-value is $\frac1m$.

%\subsection{Power}

%Next, we will develop power theoretically and then relate it to the empirical results.
%We will first theoretically define and  derive the power of  a visual test in the setting of a parametric hypothesis test, i.e. we assume a test  of $H_0: \theta = 0$ vs $H_1: \theta \neq 0$ for some real valued parameter vector $\theta$. In a next step we will then relate the theoretical results to our empirical findings.

%\red{How are $\Theta_0$ and $\Theta^c_0$ defined? - why do we need them? wouldn't it be better to re-formulate power in terms of acceptance region, not in terms of the parameter space? - $1/m$ also doesn't look right for the power - I think that's just looking at one observer.}
%\begin{dfn} \label{dfn:power}
%The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
%    \begin{equation*}
%      \text{Power}_V(\theta)= 
%        \begin{cases} 
%              \frac1m & \text{if $\theta \in \Theta_0$,} \\
%              Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
%        \end{cases}
%    \end{equation*}
%\end{dfn}
%
%
%\red{Alternative Power definition:}
\begin{dfn} \label{dfn:power}
The \textbf{power} of a visual test, $V_{\theta}$, is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
    \begin{equation*}
      \text{Power}_V(\theta)= Pr(\text{Reject } H_0 \mid \theta) 
%      = P(X \ge x_{\alpha}) =
%        \begin{cases} 
%               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true}.\\
%              1 - F_{X, \theta} (x_{\alpha} - 1) & \text{ if } H_1 \text{ is true}.
%        \end{cases}
    \end{equation*}
\end{dfn}

\noindent An important difference between conventional and visual testing is that lineups will depend on observers evaluation. Thus, the number of observers affects the estimation of power, where $X$ is the number of observers (out of $K$) who identify the actual data plot from the lineup, and power is estimated by
\[
\widehat{Power}_{V} (\theta) = {Power}_{V, K} (\theta) = 1 - F_{X, \theta} (x_{\alpha} - 1),
\] 
Here $F_{X, \theta}$ is the distribution of $X$ and $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. Note that the distribution  $F_X$ depends on which hypothesis is true:
under the null hypothesis, $X \sim Binom_{K, 1/m}$, leading to:
\[
Power_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
\]
%where $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
If the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$. Assessing an individual's skill to identify the actual data plot will require that an individual evaluates multiple lineups.

%\noindent Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.

%\red{there may be a flow in this definition. Due to the discrete nature of binomial (specially $x_{\alpha}$) this definition would only give a conservative lower limit of the actual power. I feel like we should not define power using binomial.}

Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in deciding on the choice of plot to use. Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose. This is akin to different choices of statistics in conventional hypothesis testing, for example, mean, median, or trimmed mean. One is typically better than another. For two different visual test statistics of the same actual data, one is considered to be better, if $T(y)$ is more easily distinguishable to the observer. Power is typically used to measure this characteristic of a test. %Power is the probability of rejecting $H_0$ when it is not true.  We can assess power therefore both empirically based on experimental data and through theoretical considerations.

\section{Effect of Observer Skills and Lineup Size}~\label{sec:size}

%\green{Notation changes $i$ to $j$, and $p$ to $\pi$. Can someone please decide which to use, and make the appropriate changes? How does the ``Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$...'' below relate to the ``let $X_i$ be a binary random variable...'' above? It looks like this is a repeat. One needs to go once notation is clarified.}

%\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate actual data. \\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for $p$-value.  \\ Should I change everything accordingly? I see that we have been careful not to use $\beta$ for power! }

\subsection{Subject-specific abilities}~\label{sec:model}
%Calculating the power  is intricately related to the probability to identify the actual data plot from the lineup.

Suppose each of $K$ independent observers gives evaluations on multiple lineups, and responses are considered to be binary random variable, $X_{\ell i} \sim  Binom_{1, p_{\ell i}}$, where $X_{\ell i} = 1$, if subject $i$ correctly identifies the actual data plot on lineup $\ell$, $1 \le \ell \le L$, and 0 otherwise. A mixed effects logistic regression model is used for $P(X_{\ell i} = 1) =  p_{\ell i} = E(X_{\ell i})$, accommodating both for different abilities of observers as well as differences in the difficulty of lineups.

%Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the actual data plot from lineup $\ell$, and 0 otherwise.
The model can be  fit as:% $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
\begin{equation} \label{eqn:mixed}
g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
\end{equation}

\noindent where $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
$W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters. Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well lineup-specific elements, e.g. effect size or difficulty level.
%
$Z_{\ell i}$,  $1 \le i \le K$, $1 \le \ell \le L$,  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$; and 
$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$. $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups. 
Note that $\tau_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.

\noindent Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leads to the estimate of the subject and the lineup specific probability of successful evaluation as 
\begin{equation} \label{eqn:mixed_power}
\hat p_{\ell i} =  g^{-1}(W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}).
\end{equation}

%\red{This model is  described in more detail in equation \eqref{eqn:mixed}.}

%\noindent For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$. When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$. This leads us to obtain the probability 
%\begin{equation}\label{eqn:pvalue}
%Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} \left(\frac{1}{m}\right)^k\left(1-\frac 1m\right)^{(K-k)}}
%\end{equation}
%where $u$ is the observed number of successful evaluations. We use this probability to  define the $p$-value for visual test $V_{\theta}$ as follows:
%\begin{dfn} \label{dfn:pval}
%The $p$-value associated with the decision of a visual test $V_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
%where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$ as given in eqn (\ref{eqn:pvalue}). Since $u$ is a discrete value with boundaries 0 and $K$,
% we can also re-formulate the $p$-value in terms of upper and lower boundaries:
% \begin{equation*}
% \begin{array}{rcll}
% P(U \ge u+1) \le &p\text{-value}& (\le 1), & \text{if }u < K, \text{and in particular for } u=0\\
%( 0 \le) &p\text{-value}& \le P(U \ge u-1), & \text{if }u > 1, \text{and in particular for } u=K.\\ 
% \end{array}
% \end{equation*}
% }

%\red{HH: Tried to re-write it to answer the question below.
%Why is the $p$ value not just $Pr(U \ge u)$ ? (We need to modify this a bit with respect to the rejection region $R$, but there should not be a difference in terms of whether we reject or accept. }
%
%    \begin{equation*}
%        \begin{cases} 
%              p\text{-value} \le Pr(U \ge u)    & \text{when reject $H_0$,} \\
%              p\text{-value} > 1-Pr(U \ge u)  & \text{when fail to reject $H_0$}
%        \end{cases}
%    \end{equation*}
%\end{dfn}

%\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}\\
%\red{For the following paragraph, we wanted to get rid of the absolute value of $t$ and instead look at $t \in R$ and $t \in R^c$.}

%\noindent
\subsection{Lineup size, $m$}

The finite number $m-1$ of representatives of the null distribution used  as comparison against the test statistic, is a major difference between visual inference and the conventional testing. The choice of $m$ has an obvious impact on the test.

The following properties can only be derived for the situation of a fully parameterized  simulation study, as conducted in this paper. They allow for a direct comparison of lineup tests against the conventional counterparts, and also to identify properties relevant for a quality assessment of lineups when they are used in practical settings. Two assumptions are critical:
\begin{enumerate} \itemsep 0in
\item  the plot setup is structured in a way that makes it possible for an observer to identify a deviation from the null hypothesis,
\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
\end{enumerate}
Evidence in support of the second assumption will be seen in the data from the study discussed in Section~\ref{sec:simulation}, the degree to which the first assumption is fulfilled is reflected by the power of a lineup. The better suited a design is for a particular task, the higher its power will be.

In order to compare the power of conventional and visual tests side-by-side, it is necessary to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.

\begin{lemma}~\label{lemma}
Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the conventional test statistic. Suppose the associated test statistic in a conventional test is observed as  $t_{obs}$ with $p$-value $p_D$.

The probability of picking the data plot from a lineup depends on the size $m$ of the lineup and the strength of the signal in the data plot. 
Under the above assumptions, the probability is expressed as:
\[
P(p_D < p_0) =  E\left[ (1 - p_D)^{m-1}\right]
\]
where $p_D$ is the $p$-value associated with the data in the test statistic, and $p_0$ is the minimum of all $p$-values in the data going into null plots.
\end{lemma}

\begin{proof}
%We will now make use of properties of the data going into the lineup and assume that the properties are reflected in the lineup display:

 By definition

$$p_D=Pr\left(|t| \ge t_{obs} \mid H_0\right)=1-F_{|t|}(t_{obs}) \ \ \Rightarrow \ \  |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$

%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the conventional test statistic.
%, with non-centrality parameter $\delta$. 
\noindent Then the distribution function of the $p$-value, $p_D$, under $H_0$, is uniform, since:  
\begin{eqnarray}\label{dist_p}
F_{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
  &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
  &=& 1-Pr\left(|t_{obs}| \le F_{|t|}^{-1}(1-p)\right) \nonumber \\
  &=&%\left\{ \begin{array}{ll}
          1-F_{|t|}\left( F_{|t|}^{-1}(1-p)\right)=p \mbox{ ; under $H_0$} 
 %         1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
%       \end{array} \right.     
\end{eqnarray}

%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$. As noted by \cite{Ruppert:2007}, under $H_1$ the density of $p_D$ $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.

Let $p_{0,i}$, $i=1, ..., m-1$ denote the  $p$-values associated with data corresponding to the $m-1$ null plots. Since this data is generated consistently with the null hypothesis,  the $p$-values are independent and  follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$. The minimum $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$ then follows a Beta distribution with shape parameters 1 and $m-1$, and corresponding distribution function 
\[
F_{p_0} (x) =  1- (1- x) ^{m-1}  \text{ for } x \in [0,1].
\]

\noindent Thus
\begin{eqnarray*}
P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt   \\
&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt  \\
&=&  E\left[ (1 - p_D)^{m-1}\right].
\end{eqnarray*}

%\noindent This proves the statement above. 
%We further see:
%\[
%E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
%\]

% If the density of $p$-value is very right skewed, the expectation term would be large. The distribution function would be highly right skewed when there would be strong signal in the data plot.Thats where the distribution function of $p$-value under alternative comes in. Should we add that in Equation \ref{dist_p}? )},
%conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
%}
%\blue{This shows that the lineup size $m$ has an effect on the power of the visual test: it is the $m-1$th moment of the distribution of $1-p_D$.

%
%
%
%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
%\begin{eqnarray*}
%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
%\end{eqnarray*}
%which, in particular, is independent of $p_{0,i}$ for all $i$.
%
%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value. Further we will assume that all observers  have the same ability in identifying the data plot.
%
%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
%
%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
%\[
%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
%\]
\end{proof}

%\red{pull the results for $P(p_D < p_{i,0})$ and $P(p_D < p_0)$ to the front, including the Figure on $1/m$ versus power. These results are independent of the regression setting - they don't make any assumptions in terms of regression.}

The above lemma allows two immediate conclusions for the use of lineups.
The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$, since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the actual data plot decreases with $m$. %This may seem counterintuitive, but it makes sense. 
%With fewer null plots the actual data plot will stand out, when the alternative hypothesis is true. 
Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$ -- if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and less affected by an increase in $m$. This can also be seen in Figure \ref{fig:pval_power}, which illustrates lemma \ref{lemma}.   Figure \ref{fig:pval_power} shows the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance. It can also be seen that for a $p$-value, $p_D$, of about 0.15 for the data plot, the signal in the plot is so weak, that it cannot be distinguished from null plots in a lineup of size $m=20$. % This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
%also seen in our experiments as shown in Figure \ref{fig:pval_pcorrect}.

%Even under $H_1$ we expect some observers to pick a null plot with a probability that depends on the strength of the signal in our data plot. 
%Reversely, we will now make use of the number of observers who do not pick the data plot in a lineup to infer the strength of the signal in the test statistic.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=3in]{images/powerplot.pdf} 
   \caption{Probability that the data plot has the smallest probability in a lineup of size $m$. With increasing $p$-value the probability drops -- when it reaches $1/m$ a horizontal line is drawn to emphasize insufficient sensitivity of the test due to the lineup size. 
   %Comparison of probabilities to pick the actual data plot for different size lineups, $m$, and different values for data plot strength, $p_D$. The dashed  line, and grey 95\% pointwise confidence band, shows the probability of picking the data plot, by chance, if $H_0$ is true. }
}
   \label{fig:pval_power}
\end{figure}

%Now assume that the same lineup plot is evaluated by $K$ independent observers. %W.l.o.g.\green{Can we get rid of cliches like WLOG?} 
%Let the last plot be the data plot and plots 1 through $m-1$ the null plots. Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.  Then  $(N_1, N_2, ..., N_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$ and $\sum_{k=1}^{m} n_k = K$. We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$. 
%
%For the distribution of $N_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
%we get
%\[
%N_m\sim \left \{ 
%\begin{array}{ll}
%B_{K, \hh{E\left[ (1 - p_D)^{m-1}\right]}} & \text { under } H_1\\
%B_{K, 1/m} & \text { under } H_0
%\end{array}
%\right .
%\]
%
%By matching the expected values, we therefore get an expression  for the $p$-value $p_D$ in the data 
%%under $H_1$  %% no such thing aas $p$-value under H_1 - even though it's just an unfortunate choice of word.
%as:
%\begin{equation}\label{eqn:power_estimate}
%%{E[p_D]} = 1- \left(\frac{E[N_m]}{K}\right)^{1/(m-1)}.
%\hh{1 - E[N_m]/K = (m-1)E[p_D] + O(E[p_D^2])}
%\end{equation}
%
%Note that the left hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers. While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$. 
%\hh{Also note, that this equation yields an estimate of $p_D$ that is bounded by $1/(m-1)$. We should therefore not make the decision of significance depend on this estimate, but only use this equation after the fact. 
%Both of these topics are further highlighted in the application section.}

%Picking out the data plot is still a measure of how different the data plot is compared to the null plots, and therefore is a test statistic of the visual test.

%This allows us to compute this value independently of the parametric setting. We therefore define the {\it signal strength} of a  visual test as the following:
%
%\begin{dfn} \label{dfn:signal}
%The {\it signal strength} $\sigma_T$ of a visual test is defined as the probability to identify the data plot.
%\end{dfn}
%
%In a lineup of size $m$, evaluated by $K$ independent observers,  we can estimate the signal strength of the test as
%\[
%\sigma_T^{m-1} = \frac{n_m}{K}.
%\]
%
%In a parametric setting, signal strength of a test coincides with the power of the test.

%Theoretical power for the regression parameters is derived in the next section.
%\subsection{Estimating Power from Lineups}
%
%We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}. Suppose, we have responses from $K$ independent observers with $u$ identifications of the actual plot. This gives an estimated power of
%\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}

%From model \ref{eqn:mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
%the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
%%Thus we have the overall probability of success ($\pi$ or power) 
%for specified sample size ($n$) and  variance ($\sigma$)  as 
%
%\begin{equation}\label{eqn:power} 
%Power(\beta) = \pi=Pr(Y=1|\beta, n, \sigma) 
%\end{equation}
%
%The main covariate factors that determine the power of the test are sample size ($n$), $\sigma$ and the parameter $\beta$. combining these we obtain one single covariate effect denoted by $E$ and defined as $E = \beta/ \sigma*\sqrt{n}$. Using equation \eqref{eqn:mixed}, we then have the power as
%
%\begin{equation}\label{eqn:effect} 
%Power(E) = \pi=Pr(Y=1| E) 
%\end{equation}

\section{Application to Linear Models} \label{sec:regression}

To make these concepts more concrete consider how this would operate in the linear models setting. Consider a linear regression model 
\begin{equation}\label{multi} Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i 
\end{equation}
where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariates ($X_j, j=1,..,p$) can be continuous or discrete.

In this setting, there are many established graphics that are used to evaluate and diagnose the fit of a regression model (e.g. \citet{cook:99}). Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and commonly used statistical plots that might be used as corresponding visual test statistics.
%These plots should be familiar, because they are all in common use. 
For example, to examine the effect of variable $X_j$ on $Y$, we would plot residual obtained from fitting model without $X_j$ against $X_j$ or for single covariate we may plot $Y$ against $X_j$ (cases 1-4). To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5). For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable respectively.

%Note that some charts can be associated with several different testing situations. In our paper, we pick the examples for situation 2 and 3 as candidates.

\begin{table*}[hbtp] 
\caption{Visual test statistics for testing hypotheses related to the model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i  $ } 
\centering 
\begin{tabular}{m{0.5cm}m{3cm}m{2cm}m{3cm}m{5.5cm}} 
\hline\hline 
Case & Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
\hline 
1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}\end{minipage} & Scatter plot with least square line overlaid. For lineup plot, we simulate data from fitted null model. \\ 
2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{3cm}   \scalebox{0.4}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Residual vs $X_k$ plots. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{minipage} & Box plot of residuals grouped by category of $X_k$. For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$. \\
4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_interection.pdf}}  \end{minipage} & Scatter plot with least square lines of each category overlaid. For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space 
5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_nonlinear.pdf}}  \end{minipage} & Residual vs predictor plots with loess smoother overlaid. For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$. \\ 
6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{3cm} \scalebox{0.4}{\includegraphics{stat_sigma_box.pdf}}\end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$. For lineup plot, we simulate data from standard normal. \\ 				
7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression. For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$. \\ 
8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_goodness_simple.pdf}}  \end{minipage} & Histogram of the response data. For lineup plot, we simulate data from fitted model. \\[1ex] % [1ex] adds vertical space 
9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\			
\hline 
\end{tabular} 
\label{tbl:stat_multiple} 
\end{table*}

Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$. If the responses for the two levels of the categorical variable $X_k$ in the model are different, the  residuals from fitting the null model should show a significant difference between the two groups. For a visual test, we draw boxplots of the residuals conditioned on the two levels of $X_k$. If $\beta_k\ne 0$ the boxplots should show a vertical displacement.

%A lineup including the test statistic  for a binary $X_k$ is shown in Figure \ref{fig:test_category}. The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma}^2)$. The test statistic, the plot containing the actual data, is randomly placed among these null plots. If the test statistic is identifiable the null hypothesis is rejected. % with a $p$-value of at most 0.05.

%\subsection{Expected power}

% Di: This is specific for the regression model
%Consider the hypothesis test $H_0: \beta_k=0$ against $H_1: \beta_k \ne 0$ in model \ref{multi}.
%%Now consider estimating the power of the visual test. 
%%Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$. For the regression slope parameter $\beta$ suppose the associated test statistic in conventional test be $t_{obs}$ with $p$-value $p_D$. By definition we have 
%%
%%$$p_D=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
%%
%%\red{HH: the following paragraph should probably move up, but I'm not sure how to change it appropriately. Mahbub, could you take care of that?}
%%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
%%\begin{eqnarray}\label{dist_p}
%%F{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
%%  &=& 1-Pr(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
%%  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
%%  &=&\left\{ \begin{array}{ll}
%%          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
%%          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
%%       \end{array} \right.     
%%\end{eqnarray}
%%
%%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$. As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.  
%In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$. Suppose  $p_0=\min_i(p_{0,i})$. Thus $p_0 \sim \text{Beta}(1,m-1)$.

The conventional test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis, if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.  It forms the benchmark upon which we evaluate the visual test. To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.  This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the conventional $p$-value as details given in Lemma \ref{lemma}. Thus the expected power of the visual test in this scenario is

\begin{equation}\label{power_exp} 
   Power(\beta)=Pr(p_{D} < p_0)  \quad \text{for}  \quad \beta \ne 0
\end{equation}
Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}. Notice that the expected power of the visual test is %almost as good as 
close to the power of the conventional test.

%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
%\begin{thm}\label{thm:power}
% equation \eqref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
%\end{thm}
% 
%{ \em proof}
%
%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
%\begin{eqnarray*}
%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
%
%Thus expected power in equation \eqref{power_exp} be
%\begin{eqnarray*}
% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
%  &=& E(1-p_D)^{m-1}   \\
%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_D}(0) \quad \text{as} \quad m \rightarrow \infty
%\end{eqnarray*}
%
%
\begin{figure}[hbtp]
   \centering
       \scalebox{.6}{\includegraphics{power_expected.pdf}}
       \caption{Expected power of visual test  and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
       \label{fig:power_expected}
\end{figure}

% \pagebreak

% -------- Heike's writeup added here

%\subsection{Some Power Considerations}

%\subsection{Expected number of choices}
%Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
%\[
%E[X] = (m-1)(1-E[p_D]).
%\]
%With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
%This should allow us to infer some of the signal strength $p_D$ as
%\begin{equation} \label{plot_signal}
%E[p_D] = 1 - E[X]/(m-1),
%\end{equation}
%i.e. by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.
%

\section{Human Subjects Experiments with Simulated Data} \label{sec:simulation}

Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression scenario. The first two are conventional settings where we would not expect the lineup protocol to do better than the conventional test, but we hope it to perform favorably. The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test. (Data and lineups used in the experiments are available in the supplementary material.)

After many small pilot studies with local personnel, it was clear that some care was needed to set up the human subjects experiments. It was best for a observer or a subject to see a block of 10 lineups with varying difficulty, with a reasonable number of ``easy'' lineups. The explanations about each experiment (below) includes an explanation of how the lineups were sampled and provided to the subjects, for each of the three experiments.

Participants for all the experiments were recruited through Amazon's online web service, Mechanical Turk  \citep{turk}. A summary of the data obtained for all three experiments are shown in Table \ref{tbl:summary}. Participants were asked to select the plot they think best matched the question given, provide a reason for their choice, and say how confident they are in their choice. Gender, age, education and geographic location of each participant are also collected. %The data was collected through the web. 
For each of the experiments one of the lineups was used as a test plot (easy plot) which everyone should get correct, so that a measure of the quality of the subjects effort could be made.

Note that no participant was shown the same lineup twice.
%\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.} \blue{We do have some data where we notice multiple responses for a lineup by a subject. This happens when subject accidentally use browser's ``back'' button and same data is saved twice. We made sure in the data cleaning process that we do not include more than one response on a lineup by a subject. For this we only keep the very first response.}

\subsection{Discrete covariate}\label{sec:category}

The experiment is designed to study the ability of human subjects to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}). Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$). The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical conventional test. Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian. Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}. Three replicates at each level were generated. These produced 60 different ``actual data sets'', and thus, 60 different lineups. To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates. The actual data plot was generated by making  side-by-side boxplots of the residuals (Table \ref{tbl:stat_multiple}, case 3). The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and  plotted in the same way. The actual data plot was randomly placed among these null data plots to produce the lineup. Figure \ref{fig:test_category} is an example of one of these lineups. The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.

%{\bf Steps to obtain lineup}
%\begin{enumerate}
%	\item For a given parameter setting obtain actual data set
%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
%	\item Generate actual data plot using the procedure as described in table \ref{tbl:stat_multiple}
%	\item Generate 19 null (residual) data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
%\end{enumerate}

The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power. Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}. Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$. 
For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05. Thus we have $ME =1.96 \sqrt{ \gamma(1-\gamma) / n_{\gamma} } \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$  
% \hh{we're not even close to $n_\gamma$ with ten lineups .... we need more in the order of 400. Let's not emphasize this so much.}

Each subject viewed at least 10 lineups with the option to evaluate more.
%and we allowed them to provide more evaluations if they wanted. 
Depending on the parameter combinations we group the lineups in different difficulty levels  as easy, medium, hard and mixed (actual numbers are given in the supplementary material). For each difficulty level a specific number of lineups was randomly picked  for evaluation. This number is chosen so that total number of evaluations for each lineup for that group exceed the threshold $n_{\gamma}$. To satisfy this plan we needed to recruit at least 300 subjects.

\begin{table}[hbtp]
\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for each of the simulation experiments.} % title name of the table
\centering
\scalebox{.9}{
\begin{tabular}{c c c c c}
% creating 10 columns
\hline\hline

% inserting double-line
& & \multicolumn{3}{c}{Slope ($\beta$)} \\
 \cline{3-5}

Sample size  & Error SD  &  \multicolumn{1}{c} {Experiment 1}  & \multicolumn{1}{c} {Experiment 2}  & \multicolumn{1}{c} {Experiment 3} \\
 ($n$) &   ($\sigma$) & Discrete covariate & Continuous covariate & Contaminated data 
\\ [0.5ex]
\hline
% inserts single-line

% Entering 1st row
&  5 & 0, 1,  3, 5, 8  & 0.25, 0.75, 1.25, 1.75, 2.75 & 0.1, 0.4, 0.75, 1.25, 1.5, 2.25\\[-1ex]
\raisebox{1.5ex}{100} &12
& 1, 3, 8, 10, 16  & 0.5, 1.5, 3.5, 4.5, 6 &\\[1ex]

% Entering 2rd row
&  5 & 0, 1, 2, 3, 5  & 0.1, 0.4, 0.7, 1, 1.5&\\[-1ex]
\raisebox{1.5ex}{300} & 12
& 1, 3, 5, 7, 10  & 0, 0.8, 1.75, 2.3, 3.5 &\\[1ex]
% [1ex] adds vertical space
\hline
% inserts single-line
\end{tabular}
}
\label{tbl:experiment_params}
\end{table}

%For a given proportion $p$ we want to have margin of error (ME) to be 0.1. Thus we have $ME =1.96 \sqrt{ \frac 1 n p(1-p)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_p \geq \frac{p(1-p)}{(0.05/1.96)^2}.$$ Now from the theoretical (UMP) power curve for each parameter combinations shown in Table \ref{tbl:experiment_params} we know the power(say $p$) and thus we can estimate the number of evaluations required for that specific combinations of parameter. We collect data such that each of the parameter combinations in the Table \ref{tbl:experiment_params} has approximately the estimated sample size that we calculated here.

%\green{Did you in fact bin like this???}\red{For each parameter combination a set of 1000 actual data are generated and their associated p-values were computed. This produces an empirical distribution of p-value under alternative hypothesis. Finally, three data sets that have 16th, 50th and 84th quantile of the p-values were taken. Similar procedure was used to generate 19 null plots based on the empirical distribution of p-value under null hypothesis}. For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets. A 1000 sets were generated for each parameter combination and the conventional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated. The 3 replicates were drawn from each of three blocks of $p$-values: (0.0-$q_{33}$), ($q_{33}$-$q_{66}$), ($q_{66}$-1) where $q_i$ is the $i$th percentile in the distribution of the $p$-values. Additional control was applied to the 19 null plots. Because the distribution of these $p$-values should follow a Uniform(0,1) distribution, data sets were binned on this range by $p$-value, and a data set was randomly selected from each bin.

\subsection{Continuous covariate}  \label{sec:continuous}

This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case 2. Data is simulated with two sample sizes ($n=100, 300$), two standard deviations of the error ($\sigma=5, 12$), a variety of slopes ($\beta$), as given in Table  \ref{tbl:experiment_params}. We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$. For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups. For some parameter combinations we have more than three replications, because additional lineups were produced for values of the parameters that exhibited large variation in the initial data.

The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid. To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data. 
To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters. For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest). 
Figure \ref{fig:test_continuous} shows an example lineup for this experiment from difficulty level 4.  This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5. The actual data plot location is 11.

For each combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups. Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0. The order that the lineups are shown to participants is randomized. Each subject saw a substantial number of difficult lineups.

%\hh{While this allows us a better estimation for the power in small slopes, }
%%The intention was noble, to get a better handle on the power of the lineups for small slopes, 
%but practically this made the study very tiring for the participants. %It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. 
%\green{perhaps should go into so much detail about this, or keep this for the conclusions section.}

%{\bf Steps to obtain lineup}
%\begin{enumerate}
%	\item For a given parameter setting obtain actual data set
%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
%	\item Generate actual data plot (a scatter plot of $Y$ vs $X_1$ a least square line overlaid)
%	\item Generate 19 null data sets from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and obtain 19 null plots
%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
%\end{enumerate}

%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.}

%{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation. For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in Table \ref{tbl:experiment_params}. For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4. The 10 lineups are chosen as follows;

%\begin{enumerate}
%	\item For one combination of  sample size ($n$) and error SD ($\sigma$) randomly select five lineups with one from each difficulty level
%	\item For another combination of sample size ($n$) and error SD ($\sigma$) select four lineups with one from each difficulty level 0 through 3
%	\item Finally select another lineup with difficulty level 0
%\end{enumerate}

\begin{figure}[p]
%\begin{figurehere}
   \centering
       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous. One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true. Which plot is the most different from the others, in the sense that there is the steepest slope? (The position of the actual data plot is provided in Section \ref{sec:continuous}.)}
       \label{fig:test_continuous}
\end{figure}

\subsection{Contaminated data} \label{sec:contamination}

The first two simulation experiments use data generated under a normal error model, satisfying the conditions for conventional test procedures. In these situations there exists a test, and there would, in general, be no need to use visual inference. The simulation is conducted with the hopes that the visual test procedure,  will at least compare favorably with the conventional test -- without any ambition of performing equally well.
 This third simulation is closer to the mark for the purpose of visual inference. The assumptions for the conventional test are violated, by contaminating the data. The contamination makes the estimated slopes effectively 0, yet the true value of slope parameter is not.  The data is generated from the following model:

\[
  Y_i = \left\{
  \begin{array}{l l}
    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=1, ...,n_c\\
  \end{array} \right.
\]
where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$. $n_c$ is the size of the contaminated data. For the experiment we consider $n=100$ and $n_c=15$ producing actual data with $115$ points. Further, $\alpha=0$, $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$. %Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.  
A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data. This experiment  follows the test in Table \ref{tbl:stat_multiple}, case 2. The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null  plots are scatterplots of null data generated by plotting simulated residuals from $N(0, {\hat{\sigma}}^2$)  against $X_1$.

Experiment three consists of a total of 30 lineups, made up of five replicates for each of the six slopes as shown in Table \ref{tbl:experiment_params}. We use the slope directly as a measure of difficulty, with difficulty = 0 for the largest slope and difficulty = 5 for the smallest slope.  Subjects were exposed to a total of ten lineups, with two lineups from each of the difficulty levels 0 through 3, and one  lineup each from levels 4 and 5.

%{\bf Steps to obtain lineup}
%\begin{enumerate}
%	\item For a given parameter setting obtain actual data set
%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
%	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
%	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
%\end{enumerate}

An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? The actual data plot location is 7.

%\begin{figure}[hbt]
%\begin{figurehere}
%   \centering
%       \scalebox{0.6}{\includegraphics{contaminated_data.pdf}}
%       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant. The conventional t-test $p$-value for the slope is 0.67 for this data set but visually we can see a convincing evidence of positive slope. 
%}
%       \label{fig:cont_dat}
%\end{figure}

\begin{figure}[p]
%\begin{figurehere}
   \centering
       \includegraphics[width=0.95\textwidth]{lineup_contaminated.pdf}
       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous but the inclusion of some contamination with the data spoils the normality assumption of error structure. One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true. Which plot is the most different from the others, in the sense that there is the steepest slope? (The position of the actual data plot is provided in Section \ref{sec:contamination}.)}
       \label{fig:test_contaminated}
\end{figure}

%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects. Show a sample lineup, probably the most difficult, instead of the single plot.}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
%       \label{fig:power_observed}
%\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
%       \label{fig:power_loess1}
%\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
%       \label{fig:power_loess2}
%\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
%       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in Table \ref{tbl:experiment_params}.}
%       \label{fig:power_diff_exp}
%\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3. }
%       \label{fig:power_observed_exp3}
%\end{figure*}

\section{Results} \label{sec:results}

\subsection{Data Cleaning}

% In this process of collecting data some participants did not provide demographic information as we see in Table \ref{tbl:summary} that male and female participants do not add up to total participants. It is not clear why they did not provide the information. Now the concern arises whether we should keep their responses in the study. It is also possible that participants were not serious enough to provide feedback. 
%For each of the experiments we had a test plot (easy plot) for which they should have a correct evaluation without much effort.  This could be a criteria to determine their attentiveness in the study. For this we consider six different data cleaning criteria as shown below;

%{\bf Screening scenarios:}
Amazon Mechanical Turk workers are paid for their efforts, not substantial, but on the scale of the minimum wage of the USA. Some workers will try to maximize their earnings for minimum effort, which can affect the results from the data. For example, some workers may simply randomly pick a plot, without actively examining the plots in the lineup. For the purpose of identifying these participants and cleaning the data, we take action based on a subject's answer to one of the very easy lineups that everybody was exposed to:
if the subject failed to identify the actual data plot, we remove all of this subject's data from the analysis. If the answer on the lineup is correct, we remove this lineup from the analysis, but keep all of the remaining answers.
Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.

%
%one of these was picked at random as the filtering lineup, and action was taken depending on whether they answered correctly or incorrectly. The data of participants who did not answer correctly was removed. 
%For the subjects, who correctly answered the filtering lineup, all of their data was kept, except for that particular lineup. That is, for each block of 10 lineups, only the data from 9 lineups was used for each subject.

%Amazon Mechanical Turk data suffers from the problem that some participants are trying to 'game' the system, because they are being paid for their efforts, not much by the standards of the US minimum wage, but enough to encourage honest efforts.  in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.  In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis. 
%There are six suggested approach to clean the data. 
%They are
%\begin{enumerate}
%\item {\bf include all participants} and their evaluations
%\item exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
%\item exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
%\item include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing
%\item include participants' records, if at least {\bf 50\% of all very easy lineups are correct} 
%\item use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records. If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
%\end{enumerate}

%\begin{table}[hbtp]
%\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
%Note that in some of the lines the number of male and female participants does not add up to the total number of participants. This is due to missing demographic information. \hh{Do we need this table or could we make this a visual comparison? For me, there's too many numbers with too many digits.}}
%\begin{center}
%\begin{tabular}{rrrrr|rrrr|rrrr}
%  \hline
%Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
% \cline{2-5}  \cline{6-9}   \cline{10-13}
%criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
%  \hline
%1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
%  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
%  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
%  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
%  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
%  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
%   \hline
%\end{tabular}
%\end{center}
%\label{tbl:summary}
%\end{table}

\begin{table}[hbtp]
\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets. Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information. }%\blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
\begin{center}
\begin{tabular}{ccrrcc}
   \hline \hline
 Experiment & Subject & Male & Female & Responses &Lineup\\ 
    \hline
1 & 239 & 121 & 107 & 2249 &  60 \\ 
  2 & 351 & 185 & 164 & 3636 &  70 \\ 
  3 & 155 & 103 &  52 & 1511 &  29 \\  
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}

\subsection{Model fitting}
 For each parameter combination, the effect is computed as

\[
E=\sqrt {n} \cdot \beta/\sigma.
\]
\noindent
The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept, i.e. 
 $W_{\ell i} = E_{\ell i}$. Instead of fitting an intercept, we make use of a fixed  offset of $\log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$. Different skill levels of subjects are accounted for  by allowing subject-specific random  slopes for  effect ($E$).

For experiment 3 we do fit intercepts: both a fixed  and subject-specific random effects, since forcing power
%While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power 
to be fixed at 0.05 for $E=0$ is not justified by the data. % (Figure \ref{fig:power_loess_effect}). *** You cannot reference the figure before you explain it!!!

%For all the three experiments, each subject saw a lineup just once. This does not allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.

%The subject-specific probability, $p_i$, of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}. 
For computation we use package {\tt lme4} \citep{lme4:2011} and software R 2.15.0 \citep{R}. $p$-value calculations are based on asymptotic normality.

Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation. The fixed effects estimates indicate that for all experiments  the proportion of correct responses increases as the effect increases. This effect is less pronounced  for experiment 3. The subject-specific variability is smaller for experiment 1, and relatively large for experiment 3. %The larger variability in the response data for experiment 3% as we see in Figure \ref{fig:power_loess_effect}  *** You cannot reference the figure before you explain it!!!
%contributes to the random effect variance estimate be much larger compared to those of other experiments. For the same reason we see the standard error for intercept estimate is much higher as well. Also notice the slope estimate for experiment three is lowest as well. It is because for small parameter values the power is much higher for experiment 3.  % \green{What do we learn about the models from these parameter estimates? Explain this here too.}

%\green{This section needs to have sufficient detail of the model fitting, output and diagnostics to feed into the next few sections discussing the findings.}

\begin{table}[hbtp]
\caption{Parameter estimates of model in Equation \ref{eqn:mixed}. Estimates are highly significant with $p$-value $<$  0.0001 for all three experiment data.}
\begin{center}
\begin{tabular}{cr@{.}lcc}
  \hline \hline
 &  \multicolumn{3}{c} {Fixed effect}  & Random effect\\
 \cline{2-4}
 Experiment & \multicolumn{2}{l}{Estimate}  &Std. error & Variance\\
  \hline
  1 & 0&39 & 0.0094 & 0.0080 \\ 
  2 & 1&21 &  0.0197 &  0.0443 \\ 
  3 & 0&59 (Intercept)  &   0.1668 & 1.9917\\ 
     & 0&21 (Slope)    &  0.0511     &  0.0245\\ 
     &-0&78 (correlation) & & \\
   \hline
\end{tabular}
\end{center}
\label{tbl:model_par}
\end{table}

\subsection{Power comparison}

 Figure \ref{fig:power_loess_effect} shows an overview of estimated power  against effect for the three experiments.  Responses from each experiment are summarized by effect size and represented as dots,  with size indicating the number of responses.
A loess  fit to the data gives an estimate of  the observed power for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands \citep{buja}. For comparison, the dashed lines show the corresponding power curves of the conventional tests. Not surprisingly, the observed power for experiments 1 and 2 is lower than the power of the conventional test.  It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and comes quite close in experiment 2, where scatterplots are used as the test statistic. 
Note that for effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2, making the fixed offset a reasonable assumption.

\begin{figure}[hbtp]
   \centering
       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
       \caption{Power in comparison to effect for the three experiments. Points indicate subject responses, with size indicating count. Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup. The loess curve (continuous line) estimates the observed power, and surrounding bands show simultaneous bootstrap confidence band.  Conventional test power is drawn as a dashed line. For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data. Power for contaminated data is shown by cross mark.}
       \label{fig:power_loess_effect}
\end{figure}

Results for experiment 3 are quite different. This is the situation where we expect to see the potential of visual inference, and indeed we do: the power of visual inference is always high, and much higher than the conventional test at small effect sizes. There is no actual conventional power in this situation, because assumptions are violated.
The dashed line shows conventional power based on uncontaminated data, whereas the cross marks show effective power based on the coefficient estimated from the contaminated data.
% It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model. The  conventional power of true effect of the contaminated data is computed anyway and presented in figure \ref{fig:power_loess_effect} by cross marks. For this we obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated. 
% 
%Note that the power structure is completely broken for contaminated data. %For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3

Results of experiment 3 are curious insofar, as  power of the visual test is  high independent of effect size. However, these results are based on correct identification of the actual data plot, regardless of reason. Although subjects were asked to select the plot that exhibited the highest association between the two variables, they might have cued in on the cluster of contaminated data. This will be explored further in Section \ref{sec:TypeIII}.

%(scatter plot) is better than what we observed for experiment 1 (box plot). For experiment 3, observed power is always above the conventional test power since for contaminated data we can visually see a trend while conventional test fail to address that.

%The power curves obtained by fitting model \ref{eqn:mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}. This shows how the screening criteria may affect the results. The corresponding power curves of conventional test are also shown. Notice that no mater what criteria we apply the result does not change much for experiment 2. Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.

%
%\begin{figure}[hbtp]
%   \centering
%       \scalebox{0.70}{\includegraphics{power_mixed.pdf}}
%       \caption{The overall power estimated from model \ref{eqn:mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
%       \label{fig:power_mixed}
%\end{figure}

%Each participant was shown a sequence of 10 lineup plots.  In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.  The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using equation \eqref{eqn:power_estimate} along with 95\% confidence interval calculated using Fisher's exact method.

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
%       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
%       \caption{Observed power shown by loess smoother and the power of conventional test for different effect size $E$ as per experiment 1 and 2.}
%       \label{fig:power_effect1}
%\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
%       \caption{Subject specific power obtained from equation \eqref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
%       \label{fig:power_effect_subject1}
%\end{figure*}

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3. Notice that for experiment 3, power does not seem to depend on the parameter $\beta$. \blue{Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met. Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}}
%       \label{fig:power_loess3}
%\end{figure*}

%\begin{figure}[hbtp]
%   \centering
%       \scalebox{0.40}{\includegraphics{power_model.pdf}}
%       \caption{Estimated power curve from equation \eqref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.  The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
%       \label{fig:power_model}
%\end{figure}

%We fit model \eqref{eqn:mixed} to the survey data obtained from the simulation experiment. The estimated overall power curve obtained from equation \eqref{eqn:power} is shown in Figure \ref{fig:power_model}. Model \ref{eqn:mixed} also gives the subject specific power curves shown in Figure \ref{fig:power_subject}. The plot includes 20 randomly selected subject-specific power curves. Notice that the power curve estimated for one subject is above the UMP test power curve.

%\begin{figure}[hbtp]
%   \centering
%       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
%       \caption{Estimated subject specific power curve from model \ref{eqn:mixed} for sample size $n$ = 100 and $\sigma$ = 12.  The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
%       \label{fig:power_subject}
%\end{figure}

\subsection{Subject-specific variation}

%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
Figure  \ref{fig:power_mixed_subject} shows power curves for both the overall experiment and subject-specific variations. The thick continuous line shows overall estimated power, the thinner lines correspond to subject-specific power curves. For comparison, the  dashed lines show power curves of the conventional test. Subject-specific power is quite different between the three experiments. In experiment 1 subjects performed similarly, and substantially worse than the conventional test. In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects. In experiment 3 there is the most subject-specific variation. Some subjects performed substantially better than the conventional test, and on average the visual test was better.

\begin{figure}[hbtp]
   \centering
       \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
       \caption{Subject-specific  power estimated from model \ref{eqn:mixed} are shown. The corresponding power curve for conventional test (dashed line) is shown for comparison. The overall average power curve estimated from the model is shown (light blue).}
       \label{fig:power_mixed_subject}
\end{figure}

\subsection{Estimating the $p$-value in the real world}

%\green{third experiment doesn't really have a p-value, does it? Need to explain this. Why are there still so many small p-values in experiment 3?} \blue{$p$-values are computed fitting the full model to the contaminated data anyway. So many small p-values are coming from the fact that contamination makes the slope negatively significant. For example, for some slope the contamination reverses the slope (positive to negative) and negative slope got significant. It is like the effect of influential outlier.}

In the real setting, where visual inference is to be useful, there will be no conventional tests $p$-values. Assessing the strength of perceived structure is a critical component of visual inference. In experiments 1 and 2, there is a $p$-value associated with the actual data plot in each lineup. As the $p$-value increases the proportion of correct responses falls (Figure \ref{fig:pval_pcorrect}), which is evidence of direct association between proportion of correct responses and conventional test $p$-values. For  $p$-values  larger than 0.15 it is very uncommon for subjects to correctly identify the actual data plot in the lineup.

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{p_val_prop_correct.pdf}}
       \caption{Proportion of correct responses decreases rapidly with the increase of p-value. After p-value exceeds 0.15 it is very unlikely to identify the actual plot. The theoretical justification of this is shown in Figure \ref{fig:pval_power}. }
       \label{fig:pval_pcorrect}
\end{figure*}

From the experimental data the visual $p$-values are estimated based on Definition \ref{dfn:pvalue}. Figure \ref{fig:pval_definition} displays resulting estimates for each lineup against the conventional $p$-value.  The pattern of visual $p$-values is interesting: for small $p$-values the visual estimates tend to be very small, while lineups with larger $p$-values result in very large visual estimate, 
% If there is a signal in the data plot, visual $p$-value has a tendency to take a smaller value (small enough to reject $H_0$) and if not it takes a larger value (large enough to not reject $H_0$) 
% 
 giving a clear indication to reject $H_0$ or not. This is why we do not see lot of visual $p$-values  between 0.05 and 0.8 especially for experiment 2. This guides the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values making the decision whether to reject or not harder.
 % for which researchers get confused whether to reject or 'not reject' $H_0$. 
 For visual tests this is not common.

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . Values are shown on square root scale. }
    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
       \label{fig:pval_definition}
\end{figure*}

%We also see the same pattern from the loess smoother fitted through the points in Figure \ref{fig:pval_definition}. 
%The sigmoid shaped loess curves indicate a clear separation of visual $p$-values. For experiment 1, most of the visual $p$-values are above 0.4 when conventional $p$-values are above 0.2. This is more evident for experiment 2 where most of the visual $p$-values are above 0.8.

For experiment 3, we see that  the visual $p$ values are very small no matter what the conventional $p$-values are. This is expected since conventional test fails to reject $H_0$ where we should reject, whereas the visual test performs well.
% and the visual test does the job nicely.

%With multiple subjects it is possible to estimate the visual $p$-value, and a measure of plot signal strength. The visual $p$-value is calculated from the binomial formula (Equation \ref{binom}). Assuming $H_o$ is true, the probability of picking the actual data plot by chance is $1/m$, and the visual $p$-value is the probability that $x$ or more subjects select that plot. Figure \ref{fig:pval_definition} displays this visual $p$-value for each lineup against the conventional $p$value.

%Similary, the strength of the signal, is calculated using the proportion of subjects picking actual data plot. It is closely related to the power of the test. From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $. This gives the signal strength estimate as  
%\[
%\hat p_D= (1- {u}/{K})/({m-1})
%\]
%where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations. Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values, on a square root scale. For experiment 3, technically, there are no conventional $p$-values, but the $p$-values are computed in the same manner as experiment 2, assuming that the data satifies the assumptions. The blue line is a loess fit to the data, and the grey line indicates equality of the two axes. The relationship between the conventional $p$-value and the plot signal strength is very strong, particularly when the values are small. As the conventional $p$-value increases, the estimated signal strength increases -- subjects find it harder to detect the actual data plot at larger $p$-values. The rise is very sharp for experiment 1, which indicates that subjects have a harder time detecting the actual data plot for even small $p$-values.

%\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
% \blue{Are we going to keep both figures \ref{fig:pval_definition} and \ref{fig:pval_plot_signal}?}
%
%\green{Explain why the term ``plot signal'' is introduced}

%, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot. For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} \blue{(should I bring this figure back?)}  displays scatterplots of these two variables for each experiment. As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values. When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup. We observed this for both experiment 1 and 2. For experiment 3, the scenario is different, because there are two structures, the association between variables, and the cluster of contaminated points.% as in this case people might have observed something other than the effect $E$.

%Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.

%Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$. This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a conventional sense. Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.36, 0.79, and -0.48 based on $n_1 = 44, n_2 = 45$, and $n_3= 26$ lineups, respectively.  }  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}. But it is true for experiment 2, right?. }
%is equivalent to conventional test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
%and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.5}{\includegraphics{p_val_plot_signal.pdf}}
%       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$). The solid line is a 45 degree line which represents the equity shown in Equation \ref{eqn:power_estimate}. p-values are shown on square root scale scale.}
%    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
%       \label{fig:pval_plot_signal}
%\end{figure*}

\subsection{Do people tend to pick the lowest $p$-value?}

 One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal. In experiments 1 and 2, this corresponds to the plot with the smallest $p$-value. This can be examined for the data collected, to see if this assumption is, indeed, reasonable. %In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the subject would pick the plot in the lineup that had the lowest $p$-value. In this section we take a look at the picks that subjects made, to see if this is, indeed, true.

Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2} show the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively. In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the count of the number of subjects who selected this plot is displayed vertically. The dot represents this point, and the lines connect vertically with 0 count. Columns are ordered according to effect size as defined in section 6.2; rows show replicates for the same combination of parameters.

%Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest. Columns correspond to the other levels of the experiment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate. (Empty cells indicate no lineup shown with this combination.)

Red indicates the plot with the lowest $p$-value in the lineup. Blue indicates the plot of the actual data when it is different from that with the lowest $p$-value. In both experiments  people tended to select the plot with the lowest $p$-value with an increase in effect size.
 %This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small. 
 The results are clearer for experiment 2, that used a continuous covariate. But even when subjects did not pick the plot with the lowest $p$-value they tended to oscillate their choices between the several low $p$-value plots. So for most subjects, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.

\begin{figure*}[hbtp]
   \centering
      \includegraphics[width=0.95\textwidth]{p_val_log_counts-a.pdf}
       \includegraphics[width=0.95\textwidth]{p_val_log_counts-b.pdf}
       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 60 individual lineups for experiment 1. Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value. The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.}
       \label{fig:P-val_log}
\end{figure*}

\begin{figure*}[hbtp]
   \centering
       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-a.pdf}       
       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-b.pdf}
       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 70 individual lineups for experiment 2. Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value. 
Columns are ordered according to effect size, with rows showing replicates of the same parameter combination on top of each other.
       Empty cells indicate combination of parameters with fewer than five replicates. 
       Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.
       }
       \label{fig:P-val_log2}
\end{figure*}

There are some noticeable exceptions to this rule. In experiment 1, when $\beta=0, n=100, \sigma=5, rep=1$ people overwhelmingly chose a plot with much larger $p$-value, similarly, for 
parameters $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value. 
For several of these exceptions, along with several easy lineups, a follow up experiment was conducted using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
%(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)

%Experiment 3 does not have $p$-values associated with each plot, so the assumption can only be tested with the first two experiments. Practically, this is not an assumption that is explicitly made, because there will not be a $p$-value available. Rather it is assumed that the subject has the ability to select the plot with the most extreme structure. A few of the lineups indicated that occasionally people cue on different patterns, for example, in experiment 1 the lineup $\beta=0, n=100, \sigma=5, rep=1$ (first row, fourth column) many subjects chose the same plot, that had a higher $p$-value. Some of these anomalous lineups were selected for follow-up studies using an eye-tracker to investigate what people were looking at.

%Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in Table \ref{tbl:pval_assumption} for both experiment 1 and 2. The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}. We see the same pattern in Figure \ref{fig:P-val_rank} where the median number of people who pick the data plot with minimum $p$-value (rank 1) are much higher than other ranks.

%\begin{table}[hbtp]
%\caption{Number of lineups for which most people pick the plot with minimum $p$-value. The data used is from screening criteria 1. } 
%\begin{center}
%\begin{tabular}{cccc}
%  \hline
%screening & Experiment & Total &  minimum $p$-value\\ 
%criteria & & lineups & picked by the most \\
%  \hline
%   & 1 &  60 &  37 \\ [-1ex]
%    \raisebox{1.5ex}{1} & 2 &  70 &  60 \\ 
%\hline
%     & 1 &  60 &  38 \\  [-1ex]
%    \raisebox{1.5ex}{2} & 2 &  70 &  61 \\ 
%\hline
%     & 1 &  60 &  38 \\  [-1ex]
%   \raisebox{1.5ex} {3} & 2 &  70 &  60 \\ 
%\hline
%     & 1 &  60 &  41 \\  [-1ex]
%    \raisebox{1.5ex}{4} & 2 &  70 &  60 \\ 
%\hline
%     & 1 &  60 &  39 \\  [-1ex]
%    \raisebox{1.5ex}{5} & 2 &  70 &  60 \\ 
%\hline
%     & 1 &  60 &  40 \\  [-1ex]
%    \raisebox{1.5ex}{6} & 2 &  70 &  60 \\ 
%   \hline
%\end{tabular}
%\end{center}
%\end{table}

%\begin{table}[hbtp]
%\caption{Number of lineups for which most people pick the plot with minimum $p$-value.  } 
%\begin{center}
%\begin{tabular}{ccc} \hline
%screening & \multicolumn{2}{c} {lineups picked by minimum $p$-value} \\
% \cline{2-3}
%criteria & Experiment 1 & Experiment 2 \\ 
%  \hline
%1 & 37 & 60 \\ 
%2 & 38 & 61 \\  
%3 & 38 & 60 \\ 
%4 & 41 & 60 \\ 
%5 & 39 & 60 \\
%6 & 40 & 60 \\ 
%\hline
%Total lineups & 60 & 70 \\  
%\hline
%\end{tabular}
%\end{center}
%\label{tbl:pval_assumption} 
%\end{table}

%\begin{figure*}[hbtp]
%   \centering
%       \includegraphics[width=0.95\textwidth]{p_val_rank_counts12.pdf}
%       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1). This indicates that most of the subjects pick the plot with minimum p-value (rank 1). There are subjects choosing the lineups with highest p-value(rank 20) but we can attribute this behavior to the type-I error. Can we?  \green{Maybe don't need this plot}}
%       \label{fig:P-val_rank}
%\end{figure*}

\subsection{How much do null plots affect the choice?}

 Visual inference falls into the same framework as randomization tests, where the statistics from the data are compared with those from null data. 
 But unlike these approaches visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution. How this small set of null plots influences the subjects choice is important for understanding the reliability of visual inference. If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice. Measuring the difference, generally, between  plots is almost impossible. However,  in this controlled setting we can use $p$-values of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.

We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2}). What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure. If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot. Figure \ref{fig:pval_difference} investigates this. The difference between the $p$-value of the actual data is compared with the lowest from the null plots. This is plotted horizontally, and the proportion correct is plotted vertically. Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots. In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, but only just. This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups. Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure. Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot. When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).

%\blue{In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value. If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an subject will pick this null plot instead of the actual data plot. Figure \ref{fig:pval_difference} investigates this for the data collected in experiments 1 and 2. The plots display the proportion of subjects who choose the actual data plot against the $p$-value of the actual data minus the lowest seen in the null data. A small negative value of the difference means that the $p$-value of the actual data is smaller than that of the null data. If there is a big negative difference we'd expect to see that the subjects would more often make the correct choice of plot. When there is a large positive difference, subjects fail to pick the correct plot. If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices. That is why we see a big range of proportion correct where difference is close to zero in Figure \ref{fig:pval_difference}.}

%\green{Something is definitely not right! The code looks like it does actual data p-value - min(null $p$-value). But how can you get a difference of 0.8 from this??} \blue{It is possible theoretically! Suppose data plot $p$-value is larger than 0.8. Under null hypothesis, $p$-value $\sim U(0,1)$. Thus min(null $p$-value) could be close to zero by chance producing this large difference. And this has happened in both experiment 1 and 2.}

 %\green{Is this really pct correct, or 1-pct correct? \blue{This is proportion correct}. When there is a big diff in p-value almost no-one gets it right! Maybe it is just a scale issue?} \blue{In fact this is what we expect. If there is large difference that means data plot $p$-value is large while minimum $p$-value is very small. This should lead people to pick the wrong plot. Our minimum $p$-value assumption basically says this fact, right?}. We see that when there is a big difference between the $p$-values subjects choose the actual data plot more often.

%Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct? Thanks!

In practice, the $p$-value is not going to be a valid way to compare plots. Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed. This is investigated in \citet{niladri:2012}. Here, numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.

\begin{figure}[hbtp]
   \centering
       \scalebox{0.70}{\includegraphics{pval_difference.pdf}}
       \caption{Investigating how the strength of structure in the null plots affects the choice of the actual data plot. Difference between the data plot's $p$-value and the  smallest $p$ value of a null plot vs proportion correct. Negative values indicate the $p$-value of the actual data plot are smaller than those of all of the null plots. Close to zero shows a wide range in the proportion correct, suggesting that when at least one null plot has structure almost as strong as the actual data plot, subjects had a difficult time in making their choice.}
       \label{fig:pval_difference}
\end{figure}

\subsection{Type III error}~\label{sec:TypeIII}

A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}. This occurs when the null hypothesis is correctly rejected but for the wrong reason. Experiment 3 is prone to this type of error. Observers were asked to identify the plot that had the largest slope, either positive or negative. But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see any trend. For the human eye this cluster of points is  as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope. This would be considered to be a Type III error because it is not what was requested, but  still gives the correct answer. This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart. Here, in the controlled but simple setting, it is of interest to see what observers cued to in the actual data plot.

Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size. Teasing this out of the results is possible by looking at the reasons observers gave for their choices. Observers were provided with four possible reasons to use for their choice:

\begin{enumerate} \itemsep 0in
\item Most different plot
\item Visible trend
\item Clustering visible
\item Other
\end{enumerate}

\noindent with the possibility to use more than one. The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice. Reasons 1 or 3 would be indicative of Type III error. Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices. If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason, or at least, when they use this reason they overwhelmingly correctly choose the actual data plot. This is not what we see. At left, are the reasons subjects gave for their choices --- 123 means that they gave all three reasons. The horizontal axis shows proportion of times that subjects correctly chose the actual data plot, and the reasons are sorted from most accurate to least accurate. The size of the point corresponds to the number of subjects putting this as the reason. Subjects that chose all three reasons almost always chose the actual data plot. This was followed by using 1 and 3, and then 1 and 4. The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted. ``Visible trend'' comes in third. This is strong evidence that for many subjects even though they are correctly choosing the data plot, often they are cueing to other structure in the plot than the trend, making a Type III error.

%It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2. Figure \ref{fig:choice_reason} examines this. For observers who chose the actual data plot, the proportions for the three reasons are shown. It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third. \green{What about other??} \blue{Not many people chose other as we see in Figure \ref{fig:choice_reason}.}

%This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e. the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}

%\green{Not sure that this goes here, but something needs to be said about Type III.}

%\blue{We also see this feature in figure \ref{fig:choice_reason}}.
%\hh{Mahbub, it's not directly fig \ref{fig:choice_reason} - it's actually the lineup corresponding to the large outlier at (0.2,  0.6) in the left panel of fig \ref{fig:choice_reason}.}

\begin{figure}[hbtp]
   \centering
%       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
%       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
       \caption{Exploring the reasons for plot choices for experiment 3, to examine the occurrence of Type III error. The vertical axis shows the reasons. At left, all subjects choices are shown, and reason 123 means all three reasons are used. At right, if the subject used a reason, regardless if they also used more than this reason, they are counted. The horizontal axis shows the proportion of times the subjects correctly chose the actual data plot. Size of the point corresponds to the number of subjects using that reason.}
       \label{fig:choice_reason}
\end{figure}

For visual inference, making a Type III, is not actually a problem. It is only a possibility in this experiment because we are working with known structure. In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis. This does show the importance of the qualitative answers that observers give for their choices.

\section{Conclusions}

%\begin{itemize} \itemsep 0in
%\item Summarize the key pieces of the paper again, especially the major findings

It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics. This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that. The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable. The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.

Power of visual tests can be estimated in different ways, which was explained in this paper. If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.  Interestingly, the theoretical power of visual test is close to that of conventional test in the conditions tested in this paper.

%\item Explain caveats, nuances, advantages with the work

One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.  A good test statistic is expected to reveal structure if the null hypothesis is not true. Small differences in plot choices such as color, orientation, and mapping of data values to graphical elements of the plot can make a difference. This is related to research in a grammar of graphics \citep{wilkinson:1999,hadley:2009}, which helps to ensure good practices. %This is the part of our ongoing and future research.

The question that is posed to observers can affect the ability to evaluate lineups. In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing. In practice, the question posed will be more general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.

% when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario. If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study. This has to be done very carefully, as any suggestions for reasons also lead the observer  what patterns to look out for.}
% By delivering an appropriate question may reduce the type III error in an appropriate scenario.}

Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure a more reliable estimate of the $p$-value for the test, and more confidence in the decision.

Experimental subjects in the simulation study came from various level of education, both genders and many different ages. Many subjects had little background in statistics, and were not necessarily aware of statistical plots. Yet, the results were good. We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking. A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it  may make statistics fun to learn.

The performance of subjects was quite varied, but consistent. No restrictions were placed on Turk workers, in terms of abilities. There were clearly some subjects who performed very badly, but it was very interesting to see that there were some super-observers, people who detected the actual data plot at a rate better than that of the power of the best conventional test. 
 It would be interesting to see how well trained subjects might perform. Prior to the Turk experiments, we conducted pilot studies using local graphics experts and obtained good results, indicating that training in data visualization might be helpful for visual inference. Future work might explore this.

%\item Describe the next steps of the work, and possible ideas for future extensions

When there is no conventional test available, visual testing procedures can be used due to its few assumptions and non-parametric nature. We have successfully used the procedures informally for local projects. It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance. The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance. If the null hypothesis is rejected, it is generally that we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection. Follow-up questions on the reasons provide qualitative insight. Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.

%Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$). The direction of future research would be to study that and apply the methods in those areas.

%The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods. We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.  The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where conventional tests cannot be used. For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots. Future experiments will be conducted to compare other regression parameters as described in Table \ref{tbl:stat_multiple} and assess sensitivity of power to modeling conditions.

%\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice. Along the lines of the Army tests that Susan brought to the working group. \blue{While recruiting subjects from the Amazon Mechanical Turk web site we did not put any restriction on the skill levels of the subjects. Thus anyone from any background could participate the experiment. It would be interesting to see how well the trained subjects performs. Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects. One of our future work would be to examine this in details. For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show. After being trained they may have to pass the qualification test before participating the experiment. Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.}.

%Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}

%In practice there is no correct answer. In these experiments we controlled the effects that observers saw, so there was a correct choice in each case. In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it. And subjects would provide open-ended answers for their choice of plot.

%What are your suggestions for future experiments, and next steps of the research?

%\paragraph{Acknowledgement:}
%This work was funded in part by National Science Foundation grant DMS 1007697.

%\bibliographystyle{plain}
%\bibliographystyle{plainnat}
\bibliographystyle{asa}
%\bibliographystyle{ieeetr}
\bibliography{references}

%\end{multicols}

\end{document}  %End of document.

\section*{Appendix}

\begin{table}[hbtp]
\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{sec:category}.} 
\centering
\begin{tabular}{c c c c  c c c}
\hline\hline
Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
\hline
easy&100& 5&8 & 1& 12 & 1\\
&100&12&16 &1&&\\
&300& 5&5 &1&&\\
&300&12&10 &1 &&\\
\hline
medium&100& 5&3 &203 & 9 &2\\
&300& 5&2,3 & 97, 1&&\\
\hline
hard&100&12&3,8,10 & 277, 126, 23& 18 &6\\
&300& 5&1 & 371 &&\\
&300&12&3,5& 375,74 &&\\
\hline
mixed&100& 5&1,5,0& 214, 2,73 & 21 &1\\
&100&12&1& 100& &\\
&300& 5&0 & 73&&\\
&300&12&7,1& 2, 152&&\\
\hline
Total &&&&&60&10\\
\hline
\end{tabular}
\label{tbl:dist_lineup1}
\end{table}

%\begin {abstract} 
%Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently, there were no formal visual methods in place for determining statistical significance of findings. This changed when Buja et al. (2009) conceptually introduced the lineup protocol for formal tests of visual findings. In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
%\end {abstract}

\end{document}

### 2012-09-05 sociology_paper.tex
trim the code

@@
 \RequirePackage{natbib}
-\usepackage{amsmath, amssymb, fullpage, amsthm, array, algorithm2e,graphicx,asa}
+\usepackage{amsmath, amssymb, fullpage, amsthm, array,,graphicx,asa}
 %\usepackage[dvips]{graphics}
@@
 \end{document}
-
-%\checkmark
-
-Statistical graphics nourish the discovery process in data analysis by revealing unexpected things,  finding structure that was not previously anticipated,  or orthogonally by contrasting prevailing hypotheses.
-The area of graphics is often associated with exploratory data analysis, which was pioneered by \cite{tukey:eda} and, is particularly pertinent in today's data-rich world where discovery during data mining has become an important activity.
-Graphics are also used in many places where numerical summaries simply do not suffice: model checking, diagnosis, and in the communication of findings.
-
-Statistical graphics research and development is agreeably established now, and has been cemented by several new developments in recent years.
-Early studies on evaluating how well statistical plots are perceived and read by the human eye \citep{cleveland:1984}, have been repeated and expanded \citep{simkin:1987,spence:1991, heer:2010} with findings supporting the original results.
-This body of work provides a framework for evaluating new statistical graphics.
-In a complementary direction, new research on formalizing statistical graphics with language characteristics makes it easier to abstractly define, compare and contrast data plots.
-\cite{wilkinson:1999} developed a grammar of graphics that is enhanced by \cite{hadley:2009}.
-These methods provide a mechanism to abstract the way data is mapped to graphical form.
-Finally, technology advances make it simple and easy for everyone to draw plots of data, and particularly the existence of software systems, such as R \citep{R}, enable making beautiful data graphics that can be tightly coupled with statistical modeling.
-
-However, measuring the strength of patterns seen in plots, and differences in individual perceptual ability, is something that is difficult and perhaps handicaps graphics  use among statisticians, where measuring probabilities is of primary importance.
-This has also been addressed in recent research.
-\citet{buja:2009} proposes a protocol that allows the testing of discoveries made from statistical graphics.
-This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis.
-One of the protocols, the lineup, places the actual data plot among  a page of plots of null data, and asks a human judge to pick the plot that is different.
-Figure \ref{fig:test_category} shows an example lineup.
-Which plot do you think is the most different from the others?
-(The position of the actual data plot is provided in Section \ref{sec:category}.) Wrapped in a process that mirrors conventional inference, where there is an explicit, a priori, null hypothesis, picking the plot of the data from the null plots represents a rejection of that null hypothesis.
-The null hypothesis typically derives from the task at hand, or the type of plot being made.
-The alternative encompasses all possible antitheses, all types of patterns that might be detected in the actual data plot, accounting for all possible deviations from the null without the requirement to specify these ahead of time.
-The probability of rejection can be quantified, along with Type I, and Type II error, and $p$-value and power can be defined and estimated.
-
-\begin{figure}[htp]
-%\begin{figurehere}
-   \centering
-       \includegraphics[width=0.95\textwidth]{lineup_category.pdf}
-       \caption{Lineup plot ($m=20$) using side-by-side boxplots for testing $H_0: \beta_k=0$.
-One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the largest shift or location difference between the boxplots?
-(The position of the actual data plot is provided in Section \ref{sec:category}.)}
-       \label{fig:test_category}
-\end{figure}
-%\end{figurehere}
-
-The protocol has only been informally tested until now.
-In the work described in this paper, the lineup protocol is compared head to head with the equivalent conventional test.
-Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model.
-This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario for visual inference.
-The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test.
-The thought is though, that the conventional setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark.
-Testing is done based on a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups.
-The results are compared with those of the conventional test.
-
-%, analysis requires plots of data.
-For exploratory data analysis, statistical graphics play an invaluable role in model checking and diagnostics.
-Even though we have established mathematical procedures to obtain various statistics, we need to support the results by also producing the relevant plots.
-
-%The scientific foundation of graphical methods for data analysis is well established by \cite{cleveland:1984}.
-In recent years we have seen several major advances in statistical graphics.
-Modern computing systems like R and SAS ease the production of high quality statistical plots.
-A grammar of graphics introduced by \cite{wilkinson:1999} presents a structured way to generate specific graphics from data and define connections between disparate types of plots.
-\cite{hadley:2009} has implemented a revised version of the grammar of graphics in R, in the package {\tt ggplot2}.
-\blue {Graham Wills has implemented the grammar in SPSS, but I'm not sure how to cite this.}
-
-%\citet{buja:2009}, following from \cite{gelman:2004}, proposed two protocols that allow the testing of discoveries made from statistical graphics.
-This work represents a major advance for graphics, because it bridges the gulf between conventional statistical inference procedures and exploratory data analysis.
-
-%In this paper we compare the lineup protocol to conventional statistical inference, specifically applied to the regression framework.
-We also present results of a human-subject study assessing the performance of individuals on lineup plots \citep{buja:2009} for testing significance of regression parameters.
-
-The paper is organized as follows.
-Section \ref{sec:visual_test} defines terms as used in visual inference, and describes how to estimate the important quantities from experimental data.
-The effect of the lineup size and number of observers on the power of the test is discussed in Section \ref{sec:size}.
-Section \ref{sec:regression} focuses on the application of  visual inference to linear models.
-Section \ref{sec:simulation} describes  three  user studies based on simulation experiments conducted to compare the power of the lineup protocol with the equivalent conventional test and  Section \ref{sec:results} presents an analysis of the resulting data.
-%The last section describes the next steps in this research.
-
-\section{Definitions and Explanations for Visual Statistical Inference} \label{sec:visual_test}
-
-Table \ref{tbl:compare} illustrates the lineup protocol in relation to conventional hypothesis testing.
-Both start from the same place, the same set of hypotheses.
-The conventional test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error.
-In the lineup protocol, the test statistic is a plot of the data.
-Here, side-by-side boxplots are used, because the variable of interest is categorical and takes just two values.
-In conventional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true.
-If it is extreme on this scale then the null hypothesis is rejected.
-In contrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution.
-If the actual data plot is selected as the most different, then this results in rejection of the null hypothesis.
-
-\begin{table*}[hbtp]
-\caption{Comparison of visual inference with conventional inference.}
-\centering 
-\begin{tabular}{llll} 
-\hline \hline
-  & Conventional Inference &  Lineup Protocol \\ %[0.5ex] % inserts table %heading 
-\hline
-  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta > 0$& $H_0: \beta=0$ vs $H_1: \beta > 0$\\
- & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
-
- Test statistic & $T(y)=\frac{\hat{\beta}}{se(\hat{\beta})}$ & $T(y)=$ \begin{minipage}[h]{1cm} \begin{center} \scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} \\
-
- & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
-
- Sampling Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.32}{\includegraphics{lineup_category_small.pdf}} \end{center} \end{minipage} \\
- & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
- Reject $H_0$ if & actual $T$ is extreme & actual plot is identifiable \\
-\hline 
-\end{tabular}
-\label{tbl:compare}
-\end{table*}
-
-%Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.
-
-%Power of a plot and $p$-value have been derived in Buja and Wickham as ...
-
-%for this paper, we want to place the parametric inference of a regression setting into the framework of visual testing:
-
-%This section outlines the concepts of visual inference in comparison to the procedures of conventional statistical inference.
-Table \ref{tbl:compare} (derived from \citet{buja:2009}) gives a summarized overview of this comparison.
-
-In general, we define $\theta$ to be a population parameter of interest, with $\theta \in \Theta$, the parameter space.
-Any null hypothesis $H_0$ then partitions the parameter space into $\Theta_0$ and $\Theta_0^c$, with $H_0: \theta \in \Theta_0$ versus $H_1: \theta \in \Theta_0^c$.
-A test statistic, $T(y)$, is a function that maps the sample into a numerical summary, that can be used to test the null hypothesis.
-The hypothesis test maps the test statistic into \{0, 1\}, based on whether $T(y)$ falls into the acceptance region, or the rejection region, respectively.
-$T(y)$ is assessed relative to null values of this statistic $T(y_0)$ which would be possible values of $T$ if $\theta \in \Theta$.
-%Similarly, for any test statistic, we have a rejection region $R$ and, complementary to that, $R^c$,  that partitions the space of the test statistic, which is usually a subspace of the real numbers.
-For visual inference, unlike in the conventional hypothesis test, the statistic is not a single value, but a graphical representation of the data chosen to display the strength of the parameter of interest, $\theta$.
-When the alternative hypothesis is true, it is expected that the plot of the actual data, the test statistic, will have visible feature(s) consistent with $\theta \in \Theta_0^c$, and that visual artifacts will not distinguish the test statistic as different when $H_1$ is not true.
-We will call a plot with this property a {\it visual statistic} for $\theta$.
-More formally,
-
-\begin{dfn} \label{dfn:test}
-A \textbf{visual test statistic}, $T(.)$, is a function of a sample that produces a plot.
-$T(y)$ maps the actual data to the plot, and we call this the \textbf{(actual) data plot}, and $T(y_0)$ maps a sample drawn from the null distribution into the same plot form.
-These type of plots are called \textbf{null plots}.
-\end{dfn}
-
-\noindent Ideally, the visual test statistic is defined and constructed using the grammar of graphics \citep{wilkinson:1999,hadley:2009}, consisting of type, and specification of aesthetics, necessary for complete reproducibility.
-\\
-\noindent
-The visual test statistic is compared with values $T(y_0)$ using a lineup, which is defined as:
-
-\begin{dfn}\label{dfn:lplot}
-A \textbf{lineup} is a layout of $m$ randomly placed visual statistics, consisting of 
-\begin{itemize}\itemsep-3pt
-\item $m-1$ statistics, $T(y_0)$, simulated from the model specified by $H_0$  (null plots) and 
-\item the test statistic, $T(y)$, produced by plotting the actual data, possibly arising from $H_1$.
-\end{itemize}
-\end{dfn}
-
-\noindent The $(m-1)$ null plots are % can be 
-members of the sampling distribution of the test statistic assuming that the null hypothesis is true.
-If $H_1$ is true, we expect this to be reflected as a feature in the test statistic, i.e.
-the plot of the data, that makes it visually distinguishable from the null plots.
-A careful visual inspection of the lineup by %\strikethrough{an}
-independent observers follows;  observers are asked to point out the plot most different from the lineup.
-If the test statistic is identified in the lineup, this is considered as evidence against the null hypothesis.
-%If it is not identified the conclusion is to not reject the null hypothesis.
-
-This leads us to a definition for the $p$-value of a lineup: under the null hypothesis, each observer has a $1/m$ chance of picking the test statistic from the lineup.
-For $K$ independent observers, let $X$ be  the number of observers picking the test statistic from the lineup.
-Under the null hypothesis $X \sim \text{Binom}_{K, 1/m}$,  therefore:
-
-\begin{dfn}\label{dfn:pvalue}
-The $p$-value of a lineup of size $m$ evaluated by $K$ observers is  given as
-
-\begin{equation}\label{binom}
-P(X \ge x) = 1 - \text{Binom}_{{K, 1/m}} (x-1) = \sum_{i=x}^{K} {K \choose i} \left(\frac{1}{m}\right)^{i} \left(\frac{m-1}{m}\right)^{K-i}
-\end{equation}
-
-\noindent with $X$ defined as above, and $x$ is the number of observers selecting the actual data plot. 
-\end{dfn}
-\noindent Note that for $x=0$ the $p$-value becomes, mathematically, equal to 1. It might make more sense from a practical point of view to think of the $p$-value as
-being larger than $P(X \ge 1)$ in this situation. If we increase either $m$ or $K$, we would be able to determine the value at a higher precision.
-
-Table \ref{pvalue} shows $p$-values for different numbers of observers for lineups of size $m = 20$.
-It can be seen that, if the null hypothesis is true, it is unlikely that multiple observers would pick the actual data plot.
-
-\begin{table}[htp]
-\caption{Possible $p$-values for different numbers of observers, $K$, for fixed size $m = 20$ lineups.}
-\begin{center}
-\scalebox{.9}{
-\begin{tabular}{|rrr|rrr|rrr|rrr|rrr|}\hline
-$K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$ &  $x$ & $p$-value & $K$  & $x$ & $p$-value & $K$  & $x$ & $p$-value\\ \hline %\cline{1-3}\cline{4-6}\cline{7-9}\cline{10-12}\cline{17-19}
-1 &  1 & 0.0500 & 2 &  1 & 0.0975 & 3 & 1 & 0.1426 & 4 & 1 & 0.1855 & 5 & 1 & 0.2262 \\%\cline{1-3}
-&&& 2 &  2 & 0.0025 & 3 & 2 & 0.0073 & 4 & 2 & 0.0140 & 5 & 2 & 0.0226 \\
-&&& &&& 3 & 3 & 0.0001 & 4 & 3 & 0.0005 & 5 & 3 & 0.0012 \\%\cline{1-6}
-&&&  &    &           &   &        && 4 & 4 & $< 0.0001$ & 5 & 4 & $< 0.0001$ \\\hline
-\end{tabular}
-}
-\end{center}
-\label{pvalue}
-\end{table}
-
-%\begin{dfn}\label{dfn:test}
-%\red{why do we need the $\theta$? - it's not defined and not explained. usually we would assume that lineups are non-parametric}
-%The \textbf{visual test}, $V_{\theta}$, is defined to be 
-%\begin{itemize}\itemsep-3pt
-%\item \textbf{Reject} $H_0$ if `enough' observers select the actual data plot in the lineup of $m$ plots, and
-%\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
-%\end{itemize}
-%where `enough' observers is defined as some threshold $x_{1-\alpha}$, for which $P(X \ge x_{1-\alpha}) < \alpha$, the level of significance.
-%\end{dfn}
-
-%\blue{To avoid confusion with one tailed $(1-\alpha)$ vs two tailed $(1-\alpha/2)$ notation, can we just use $x_{\alpha}$? }
-
-%\red{alternative definition of visual test where $type-I$ error $\alpha$ is attached.}
-
-\begin{dfn}\label{dfn:test}
-The \textbf{visual test}, $V_{\theta}$  of  size $m$ and significance level $\alpha$, is defined to  
-\begin{itemize}\itemsep-3pt
-\item \textbf{Reject} $H_0$ if out of $K$ observers at least $x_{\alpha}$ correctly identify the actual data plot in the lineup, and
-\item \textbf{Fail to reject} $H_0$  otherwise. 
-\end{itemize}
-where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
-\end{dfn}
-
-\noindent
-Associated with any test is the risk of  Type I or II errors, which for visual inference are defined as follows:
-
-\begin{dfn}\label{dfn:error}
-The \textbf{Type I error} associated with visual test $V_{\theta}$ is the probability of rejecting $H_0$ when it is true; the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.\\ \noindent
-The \textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
-\end{dfn}
-
-%\green{Oh, we are using observed and observer, which might be confusing.
-May need to change observed data plot to something else.
-Will think about this.
-Will use ``actual'' for now.} \hh{I think that change is done}
-
-\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$. %, with an exact $x_{\alpha}$. 
-For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers. 
-Type II error is harder to calculate, as is usually the case.
-In visual inference, individual abilities need to be incorporated to calculate Type II error.
-%This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance.
-Here, we need to estimate the probability that an observer sees the actual data plot as different, when it really is different.
-This involves understanding the individual's visual skills.
-Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i \ (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise.
-Let $p_i$ be the probability that individual $i$ picks out the actual data plot from the lineup.
-If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.
-
-If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p_i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$.
-This makes the distribution of $X$ a Poisson-Binomial by definition (see \citet{butler93} for details).
-Ways to estimate $p_i$ will be  discussed in the following sections.
-%one of the main concerns for the remainder of the paper. One way we can estimate this is fitting a generalized mixed effect model as described in Section \ref{sec:model}.
-
-%\green{WHY??? And how does this relate to the logistic model below?} 
-%This leads to the definition of the power of the visual test, as follows:
-
-%The lineup plot can be evaluated by one or more individuals.
-When a single individual identifies the observed graph in the lineup plot we report a $p$-value of at most $1/m$, otherwise the $p$-value is at least $1-\frac 1m$.
-
- %Notice that when $N=1$, this $p$-value is $\frac1m$.
-
-%\subsection{Power}
-
-%Next, we will develop power theoretically and then relate it to the empirical results.
-%We will first theoretically define and  derive the power of  a visual test in the setting of a parametric hypothesis test, i.e.
-we assume a test  of $H_0: \theta = 0$ vs $H_1: \theta \neq 0$ for some real valued parameter vector $\theta$.
-In a next step we will then relate the theoretical results to our empirical findings.
-
-%\red{How are $\Theta_0$ and $\Theta^c_0$ defined?
-- why do we need them?
-wouldn't it be better to re-formulate power in terms of acceptance region, not in terms of the parameter space?
-- $1/m$ also doesn't look right for the power - I think that's just looking at one observer.}
-%\begin{dfn} \label{dfn:power}
-%The \textbf{power} of a visual test $V_{\theta}$ is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
-%    \begin{equation*}
-%      \text{Power}_V(\theta)= 
-%        \begin{cases} 
-%              \frac1m & \text{if $\theta \in \Theta_0$,} \\
-%              Pr(\text{Reject } H_0) &\text{if $\theta \in \Theta^c_0$.}
-%        \end{cases}
-%    \end{equation*}
-%\end{dfn}
-%
-%
-%\red{Alternative Power definition:}
-\begin{dfn} \label{dfn:power}
-The \textbf{power} of a visual test, $V_{\theta}$, is defined as the probability to reject the null hypothesis for a given parameter value $\theta$:
-    \begin{equation*}
-      \text{Power}_V(\theta)= Pr(\text{Reject } H_0 \mid \theta) 
-%      = P(X \ge x_{\alpha}) =
-%        \begin{cases} 
-%               1 - \text{Binom}_{K, 1/m}(x_{\alpha} - 1) & \text{ if } H_0 \text{ is true}.\\
-%              1 - F_{X, \theta} (x_{\alpha} - 1) & \text{ if } H_1 \text{ is true}.
-%        \end{cases}
-    \end{equation*}
-\end{dfn}
-
-\noindent An important difference between conventional and visual testing is that lineups will depend on observers evaluation.
-Thus, the number of observers affects the estimation of power, where $X$ is the number of observers (out of $K$) who identify the actual data plot from the lineup, and power is estimated by
-\[
-\widehat{Power}_{V} (\theta) = {Power}_{V, K} (\theta) = 1 - F_{X, \theta} (x_{\alpha} - 1),
-\] 
-Here $F_{X, \theta}$ is the distribution of $X$ and $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. Note that the distribution  $F_X$ depends on which hypothesis is true:
-under the null hypothesis, $X \sim Binom_{K, 1/m}$, leading to:
-\[
-Power_V(\theta, K)= 1 - Binom_{K, 1/m} (x_\alpha - 1),
-\]
-%where $x_\alpha$ is such that $P(X \ge x_{\alpha}) \le \alpha$. 
-If the alternative hypothesis is true, with a fixed parameter value $\theta$, we can assume that an individual's probability to identify the data plot depends on the parameter value, and $X_i \sim Binom_{1, p_i(\theta)}$.
-Assessing an individual's skill to identify the actual data plot will require that an individual evaluates multiple lineups.
-
-%\noindent Note that $F_{X, \theta}$ is the distribution of $X$ under the alternative with given parameter value $\theta$.
-
-%\red{there may be a flow in this definition.
-Due to the discrete nature of binomial (specially $x_{\alpha}$) this definition would only give a conservative lower limit of the actual power.
-I feel like we should not define power using binomial.}
-
-Power is an important consideration in deciding which test to use for solving a problem. %In conventional testing, new test procedures are examined for their power under a variety of scenarios. 
-Here we use it to compare the performance of the visual test with the conventional test, but in practice for visual inference it will mostly be important in deciding on the choice of plot to use.
-Analysts typically have a choice of plots to make, and a myriad of possible options such as reference grids, for any particular purpose.
-This is akin to different choices of statistics in conventional hypothesis testing, for example, mean, median, or trimmed mean.
-One is typically better than another.
-For two different visual test statistics of the same actual data, one is considered to be better, if $T(y)$ is more easily distinguishable to the observer.
-Power is typically used to measure this characteristic of a test.
-%Power is the probability of rejecting $H_0$ when it is not true.
-We can assess power therefore both empirically based on experimental data and through theoretical considerations.
-
-\section{Effect of Observer Skills and Lineup Size}~\label{sec:size}
-
-%\green{Notation changes $i$ to $j$, and $p$ to $\pi$.
-Can someone please decide which to use, and make the appropriate changes?
-How does the ``Let $X_{j\ell} \sim B_{1, \pi_{j\ell}}$...'' below relate to the ``let $X_i$ be a binary random variable...'' above?
-It looks like this is a repeat.
-One needs to go once notation is clarified.}
-
-%\red{Here is what my suggestion for notation \\  Use $B$ for logistic model and $\beta$ for linear model \\ Use $X$ to indicate logistic model covariate matrix and $U$ to denote binomial variable \\ $Y$ to indicate actual data.
-\\ Use $\pi_i$ instead of $p_i$ since $p_i$ is used for $p$-value.
-\\ Should I change everything accordingly?
-I see that we have been careful not to use $\beta$ for power!
-}
-
-\subsection{Subject-specific abilities}~\label{sec:model}
-%Calculating the power  is intricately related to the probability to identify the actual data plot from the lineup.
-
-Suppose each of $K$ independent observers gives evaluations on multiple lineups, and responses are considered to be binary random variable, $X_{\ell i} \sim  Binom_{1, p_{\ell i}}$, where $X_{\ell i} = 1$, if subject $i$ correctly identifies the actual data plot on lineup $\ell$, $1 \le \ell \le L$, and 0 otherwise.
-A mixed effects logistic regression model is used for $P(X_{\ell i} = 1) =  p_{\ell i} = E(X_{\ell i})$, accommodating both for different abilities of observers as well as differences in the difficulty of lineups.
-
-%Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the actual data plot from lineup $\ell$, and 0 otherwise.
-The model can be  fit as:% $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
-\begin{equation} \label{eqn:mixed}
-g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
-\end{equation}
-
-\noindent where $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-$W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters.
-Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well lineup-specific elements, e.g.
-effect size or difficulty level.
-%
-$Z_{\ell i}$,  $1 \le i \le K$, $1 \le \ell \le L$,  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$; and 
-$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$.
-$\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
-Note that $\tau_{\ell i }$ usually only includes a partial interaction; for a full interaction of subjects' skills and lineup-specific difficulty we would need replicates of the same subject evaluating the same lineup, which in practice is not feasible without loosing independence.
-
-\noindent Using the inverse {\it logit} link function, $g^{-1}(.)$, from  Equation \ref{eqn:mixed} leads to the estimate of the subject and the lineup specific probability of successful evaluation as 
-\begin{equation} \label{eqn:mixed_power}
-\hat p_{\ell i} =  g^{-1}(W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}).
-\end{equation}
-
-%\red{This model is  described in more detail in equation \eqref{eqn:mixed}.}
-
-%\noindent For any single observer, the power of a visual test has a lower limit of $\frac1m$ since the probability that an observer will randomly pick the test statistic under $H_0$ is $\frac1m$.
-When $K$ individuals evaluate a lineup plot independently, the number of successful evaluations, $U \sim \text{Binom} (K,1/m)$  under $H_0$.
-This leads us to obtain the probability
-%\begin{equation}\label{eqn:pvalue}
-%Pr(U \ge u)= \sum_{k \ge u}^K {{K \choose k} \left(\frac{1}{m}\right)^k\left(1-\frac 1m\right)^{(K-k)}}
-%\end{equation}
-%where $u$ is the observed number of successful evaluations. We use this probability to  define the $p$-value for visual test $V_{\theta}$ as follows:
-%\begin{dfn} \label{dfn:pval}
-%The $p$-value associated with the decision of a visual test $V_{\theta}$ is given \blue{as the probability to observe a value at least as extreme as the test statistic $u$, 
-%where $u$ is the number of successful evaluations of the lineup, which leads us to a $p$-value of $P(U \ge u)$ as given in eqn (\ref{eqn:pvalue}).
-Since $u$ is a discrete value with boundaries 0 and $K$,
-% we can also re-formulate the $p$-value in terms of upper and lower boundaries:
-% \begin{equation*}
-% \begin{array}{rcll}
-% P(U \ge u+1) \le &p\text{-value}& (\le 1), & \text{if }u < K, \text{and in particular for } u=0\\
-%( 0 \le) &p\text{-value}& \le P(U \ge u-1), & \text{if }u > 1, \text{and in particular for } u=K.\\ 
-% \end{array}
-% \end{equation*}
-% }
-
-%\red{HH: Tried to re-write it to answer the question below.
-%Why is the $p$ value not just $Pr(U \ge u)$ ? (We need to modify this a bit with respect to the rejection region $R$, but there should not be a difference in terms of whether we reject or accept. }
-%
-%    \begin{equation*}
-%        \begin{cases} 
-%              p\text{-value} \le Pr(U \ge u)    & \text{when reject $H_0$,} \\
-%              p\text{-value} > 1-Pr(U \ge u)  & \text{when fail to reject $H_0$}
-%        \end{cases}
-%    \end{equation*}
-%\end{dfn}
-
-%\red{HH: the following paragraph needs some adjustment in terms of t and $\theta$ - could you try to take of that, Mahbub?}\\
-%\red{For the following paragraph, we wanted to get rid of the absolute value of $t$ and instead look at $t \in R$ and $t \in R^c$.}
-
-%\noindent
-\subsection{Lineup size, $m$}
-
-The finite number $m-1$ of representatives of the null distribution used  as comparison against the test statistic, is a major difference between visual inference and the conventional testing.
-The choice of $m$ has an obvious impact on the test.
-
-The following properties can only be derived for the situation of a fully parameterized  simulation study, as conducted in this paper.
-They allow for a direct comparison of lineup tests against the conventional counterparts, and also to identify properties relevant for a quality assessment of lineups when they are used in practical settings.
-Two assumptions are critical:
-\begin{enumerate} \itemsep 0in
-\item  the plot setup is structured in a way that makes it possible for an observer to identify a deviation from the null hypothesis,
-\item an observer is able to identify the plot with the strongest `signal' (or deviation from $H_0$)  from a lineup.
-\end{enumerate}
-Evidence in support of the second assumption will be seen in the data from the study discussed in Section~\ref{sec:simulation}, the degree to which the first assumption is fulfilled is reflected by the power of a lineup.
-The better suited a design is for a particular task, the higher its power will be.
-
-In order to compare the power of conventional and visual tests side-by-side, it is necessary to assume, that we are in the controlled environment of a simulation with tests corresponding to a known parameter value $\theta \in R$ and associated distribution function $F_t$ of the test statistic.
-
-\begin{lemma}~\label{lemma}
-Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$, the conventional test statistic.
-Suppose the associated test statistic in a conventional test is observed as  $t_{obs}$ with $p$-value $p_D$.
-
-The probability of picking the data plot from a lineup depends on the size $m$ of the lineup and the strength of the signal in the data plot. 
-Under the above assumptions, the probability is expressed as:
-\[
-P(p_D < p_0) =  E\left[ (1 - p_D)^{m-1}\right]
-\]
-where $p_D$ is the $p$-value associated with the data in the test statistic, and $p_0$ is the minimum of all $p$-values in the data going into null plots.
-\end{lemma}
-
-\begin{proof}
-%We will now make use of properties of the data going into the lineup and assume that the properties are reflected in the lineup display:
-
- By definition
-
-$$p_D=Pr\left(|t| \ge t_{obs} \mid H_0\right)=1-F_{|t|}(t_{obs}) \ \ \Rightarrow \ \  |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
-
-%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$, the conventional test statistic.
-%, with non-centrality parameter $\delta$. 
-\noindent Then the distribution function of the $p$-value, $p_D$, under $H_0$, is uniform, since:  
-\begin{eqnarray}\label{dist_p}
-F_{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
-  &=& 1-Pr\left(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) \right) \nonumber \\
-  &=& 1-Pr\left(|t_{obs}| \le F_{|t|}^{-1}(1-p)\right) \nonumber \\
-  &=&%\left\{ \begin{array}{ll}
-          1-F_{|t|}\left( F_{|t|}^{-1}(1-p)\right)=p \mbox{ ; under $H_0$} 
- %         1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
-%       \end{array} \right.     
-\end{eqnarray}
-
-%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density of $p_D$ $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.
-
-Let $p_{0,i}$, $i=1, ..., m-1$ denote the  $p$-values associated with data corresponding to the $m-1$ null plots.
-Since this data is generated consistently with the null hypothesis,  the $p$-values are independent and  follow a standard Uniform distribution, $p_{i,0} \sim U[0,1], i= 1, ..., m-1$.
-The minimum $p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}$ then follows a Beta distribution with shape parameters 1 and $m-1$, and corresponding distribution function
-\[
-F_{p_0} (x) =  1- (1- x) ^{m-1}  \text{ for } x \in [0,1].
-\]
-
-\noindent Thus
-\begin{eqnarray*}
-P(p_D < p_{0}) &=& 1 - P(p_{0} \le p_D) = 1- \int_0^1  P(p_{0} \le p_D \mid p_D=t) f_{p_D}(t) dt   \\
-&=& 1 - \int_0^1 F_{p_{0}}(t) f_{p_D}(t) dt = 1 - \int_0^1 f_{p_D}(t) dt + \int_0^1 (1-t)^{m-1} f_{p_D}(t) dt  \\
-&=&  E\left[ (1 - p_D)^{m-1}\right].
-\end{eqnarray*}
-
-%\noindent This proves the statement above. 
-%We further see:
-%\[
-%E\left[ (1 - p_D)^{m-1}\right] =  \sum_{k=0}^{m-1} {m-1 \choose k} (-1)^k E[p_D^k] = 1-(m-1)E[p_D] + O(E[p_D^2]).
-%\]
-
-% If the density of $p$-value is very right skewed, the expectation term would be large.
-The distribution function would be highly right skewed when there would be strong signal in the data plot.Thats where the distribution function of $p$-value under alternative comes in.
-Should we add that in Equation \ref{dist_p}?
-)},
-%conversely, as $m$, the number of choices given in the lineup, goes up the probability to pick the data plot  goes down. 
-%}
-%\blue{This shows that the lineup size $m$ has an effect on the power of the visual test: it is the $m-1$th moment of the distribution of $1-p_D$.
-
-%
-%
-%
-%Let us think of a lineup as a head-to-head comparison of the test statistic and $m-1$ null plots. 
-%%Let the data plot have a $p$-value of $p_D$, and the null plots $p$ values of $p_{0, i}$ with $1 \le i \le m-1$.
-%We know that for each comparison the probability that the data, on which the plot is based, has the smaller $p$ value is 
-%\begin{eqnarray*}
-%P(p_D < p_{0,i}) &=& 1 - P(p_{i,0} \le p_D) = 1- \int_0^1  P(p_{i,0} \le p_D \mid p_D=t) f_{p_D}(t) dt =  \\
-%&=& 1 - \int_0^1 F_{p_{0,i}}(t) f_{p_D}(t) dt = 1 - \int_0^1 t f_{p_D}(t) dt = 1 - E[p_D],
-%\end{eqnarray*}
-%which, in particular, is independent of $p_{0,i}$ for all $i$.
-%
-%Let us make the assumption that an observer is able to identify  the chart corresponding to the data with the smallest $p$-value.
-Further we will assume that all observers  have the same ability in identifying the data plot.
-%
-%With that, we define $Z$  as the number of null plots in a lineup, for which the $p$-value $p_{0,i}$ is smaller than $p_D$.  
-%
-%Then $Z \sim B_{m-1, E[p_D]}$, and the probability that an observer will pick the data plot in a given lineup is 
-%\[
-%P(Z=0) = \left(1 - E[p_D] \right)^{m-1} = P(p_D \le p_0), \ \ \ \text{ where } p_0 = \min_{1 \le i \le m-1}  \ p_{0,i}.
-%\]
-\end{proof}
-
-%\red{pull the results for $P(p_D < p_{i,0})$ and $P(p_D < p_0)$ to the front, including the Figure on $1/m$ versus power.
-These results are independent of the regression setting - they don't make any assumptions in terms of regression.}
-
-The above lemma allows two immediate conclusions for the use of lineups.
-The probability that the observer correctly identifies the data plot is closely connected to the size of the lineup $m$, since the right hand side of the above equation decreases for larger $m$, the probability of correctly identifying the actual data plot decreases with $m$.
-%This may seem counterintuitive, but it makes sense.
-%With fewer null plots the actual data plot will stand out, when the alternative hypothesis is true. 
-Further we see, that the rate of this decrease depends strongly on the distribution of $p_D$ -- if the density of $p_D$ is very right skewed, the  expectation term on right hand side will be large and less affected by an increase in $m$.
-This can also be seen in Figure \ref{fig:pval_power}, which illustrates lemma \ref{lemma}.
-Figure \ref{fig:pval_power} shows the probability of picking the actual data plot for lineups of different size: as $m$ increases we have an increased probability to observe a more highly structured null plot by chance.
-It can also be seen that for a $p$-value, $p_D$, of about 0.15 for the data plot, the signal in the plot is so weak, that it cannot be distinguished from null plots in a lineup of size $m=20$.
-% This pattern is confirmed by our findings described in Section~\ref{sec:simulation}.
-%also seen in our experiments as shown in Figure \ref{fig:pval_pcorrect}.
-
-%Even under $H_1$ we expect some observers to pick a null plot with a probability that depends on the strength of the signal in our data plot. 
-%Reversely, we will now make use of the number of observers who do not pick the data plot in a lineup to infer the strength of the signal in the test statistic.
-
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=3in]{images/powerplot.pdf} 
-   \caption{Probability that the data plot has the smallest probability in a lineup of size $m$.
-With increasing $p$-value the probability drops -- when it reaches $1/m$ a horizontal line is drawn to emphasize insufficient sensitivity of the test due to the lineup size.
-   %Comparison of probabilities to pick the actual data plot for different size lineups, $m$, and different values for data plot strength, $p_D$.
-The dashed  line, and grey 95\% pointwise confidence band, shows the probability of picking the data plot, by chance, if $H_0$ is true.
-}
-}
-   \label{fig:pval_power}
-\end{figure}
-
-%Now assume that the same lineup plot is evaluated by $K$ independent observers. %W.l.o.g.\green{Can we get rid of cliches like WLOG?} 
-%Let the last plot be the data plot and plots 1 through $m-1$ the null plots.
-Let $n_i$ be the number of observers who picked plot $i$ as the plot with the structure least consistent with the null hypothesis, corresponding to random variable $N_i$.
-Then  $(N_1, N_2, ..., N_m) \sim$ Mult$_{\pi_1, ..., \pi_m}$ with $\sum_i \pi_i = 1$ and $\sum_{k=1}^{m} n_k = K$.
-We can estimate $\pi_i$ as $\widehat{\pi_i} = n_i/K$.
-%
-%For the distribution of $N_m$, %the number of times out of $K$ that observers picked the last plot to show the structure least consistent with the null hypothesis, 
-%we get
-%\[
-%N_m\sim \left \{ 
-%\begin{array}{ll}
-%B_{K, \hh{E\left[ (1 - p_D)^{m-1}\right]}} & \text { under } H_1\\
-%B_{K, 1/m} & \text { under } H_0
-%\end{array}
-%\right .
-%\]
-%
-%By matching the expected values, we therefore get an expression  for the $p$-value $p_D$ in the data 
-%%under $H_1$  %% no such thing aas $p$-value under H_1 - even though it's just an unfortunate choice of word.
-%as:
-%\begin{equation}\label{eqn:power_estimate}
-%%{E[p_D]} = 1- \left(\frac{E[N_m]}{K}\right)^{1/(m-1)}.
-%\hh{1 - E[N_m]/K = (m-1)E[p_D] + O(E[p_D^2])}
-%\end{equation}
-%
-%Note that the left hand side in Equation \ref{eqn:power_estimate} is independent of the parameter $\theta$, but just based on the lineup evaluation by independent observers.
-While this allows us to compute this value independently of the parametric setting, it also means that observers might pick out the data plot for a reason not related to the value of $\theta$.
-%\hh{Also note, that this equation yields an estimate of $p_D$ that is bounded by $1/(m-1)$.
-We should therefore not make the decision of significance depend on this estimate, but only use this equation after the fact.
-%Both of these topics are further highlighted in the application section.}
-
-%Picking out the data plot is still a measure of how different the data plot is compared to the null plots, and therefore is a test statistic of the visual test.
-
-%This allows us to compute this value independently of the parametric setting. We therefore define the {\it signal strength} of a  visual test as the following:
-%
-%\begin{dfn} \label{dfn:signal}
-%The {\it signal strength} $\sigma_T$ of a visual test is defined as the probability to identify the data plot.
-%\end{dfn}
-%
-%In a lineup of size $m$, evaluated by $K$ independent observers,  we can estimate the signal strength of the test as
-%\[
-%\sigma_T^{m-1} = \frac{n_m}{K}.
-%\]
-%
-%In a parametric setting, signal strength of a test coincides with the power of the test.
-
-%Theoretical power for the regression parameters is derived in the next section.
-%\subsection{Estimating Power from Lineups}
-%
-%We estimate the empirical power from responses on a specific lineup plot  generated with known values of sample size ($n$), variance ($\sigma^2$) and regression parameter ($\beta$) in model \ref{multi}.
-Suppose, we have responses from $K$ independent observers with $u$ identifications of the actual plot.
-This gives an estimated power of
-%\begin{equation}\label{power_est} Power(\beta)=\frac{u}{K} \hspace{1cm} 0 \le u \le K \end{equation}
-
-%From model \ref{eqn:mixed} we obtain %$\pi=Pr(Y=1|X_1, X_2, ..., X_p)$, 
-%the power of the underlying testing procedure as a population average %by ignoring the random effect part. 
-%%Thus we have the overall probability of success ($\pi$ or power) 
-%for specified sample size ($n$) and  variance ($\sigma$)  as 
-%
-%\begin{equation}\label{eqn:power} 
-%Power(\beta) = \pi=Pr(Y=1|\beta, n, \sigma) 
-%\end{equation}
-%
-%The main covariate factors that determine the power of the test are sample size ($n$), $\sigma$ and the parameter $\beta$.
-combining these we obtain one single covariate effect denoted by $E$ and defined as $E = \beta/ \sigma*\sqrt{n}$.
-Using equation \eqref{eqn:mixed}, we then have the power as
-%
-%\begin{equation}\label{eqn:effect} 
-%Power(E) = \pi=Pr(Y=1| E) 
-%\end{equation}
-
-\section{Application to Linear Models} \label{sec:regression}
-
-To make these concepts more concrete consider how this would operate in the linear models setting. Consider a linear regression model 
-\begin{equation}\label{multi} Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i 
-\end{equation}
-where $\epsilon_i \stackrel{iid}{ \sim } N(0,\sigma^2)$, $i=1,2, .., n$. The covariates ($X_j, j=1,..,p$) can be continuous or discrete.
-
-In this setting, there are many established graphics that are used to evaluate and diagnose the fit of a regression model (e.g.
-\citet{cook:99}).
-Table \ref{tbl:stat_multiple} lists several common hypotheses related to the regression setting, and commonly used statistical plots that might be used as corresponding visual test statistics.
-%These plots should be familiar, because they are all in common use. 
-For example, to examine the effect of variable $X_j$ on $Y$, we would plot residual obtained from fitting model without $X_j$ against $X_j$ or for single covariate we may plot $Y$ against $X_j$ (cases 1-4).
-To assess whether the linear assumption is appropriate we would draw a plot of residuals against fitted values (case 5).
-For the purposes of comparing visual against conventional inference, we focus on cases 2 and 3, with a continuous and categorical explanatory variable respectively.
-
-%Note that some charts can be associated with several different testing situations. In our paper, we pick the examples for situation 2 and 3 as candidates.
-
-\begin{table*}[hbtp] 
-\caption{Visual test statistics for testing hypotheses related to the model $Y_i = \beta_0 + \beta_1 X_{i1} + \beta_1 X_{i2} + \beta_3 X_{i1}X_{i2} + ... + \epsilon_i  $ } 
-\centering 
-\begin{tabular}{m{0.5cm}m{3cm}m{2cm}m{3cm}m{5.5cm}} 
-\hline\hline 
-Case & Null Hypothesis & Statistic & Test Statistic & Description \\ [0.5ex] % inserts table %heading 
-\hline 
-1 & $H_0: \beta_0=0$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}\end{minipage} & Scatter plot with least square line overlaid.
-For lineup plot, we simulate data from fitted null model.
-\\
-2 & $H_0: \beta_k=0$ & Residual plot & \begin{minipage}[t]{3cm}   \scalebox{0.4}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Residual vs $X_k$ plots.
-For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
-\\
-3 & $H_0: \beta_k=0$ (for categorical $X_k$) & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{minipage} & Box plot of residuals grouped by category of $X_k$.
-For lineup plot, we simulate data from normal with mean 0 variance $\hat{\sigma}^2$.
-\\
-4 & $H_0: \beta_k=0$ (interaction with categorical $X_k$) & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_interection.pdf}}  \end{minipage} & Scatter plot with least square lines of each category overlaid.
-For lineup plot, we simulate data from fitted null model.\\[1ex] % [1ex] adds vertical space
-5 & $H_0: X$ Linear & Residual Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_nonlinear.pdf}}  \end{minipage} & Residual vs predictor plots with loess smoother overlaid.
-For lineup plot, we simulate residual data from normal with mean 0 variance $\hat{\sigma}^2$.
-\\
-6 & $H_0: \sigma^2=\sigma^2_0$ & Box plot & \begin{minipage}[t]{3cm} \scalebox{0.4}{\includegraphics{stat_sigma_box.pdf}}\end{minipage} & Box plot of standardized residual divided by  $\sigma^2_0$.
-For lineup plot, we simulate data from standard normal.
-\\
-7 & $H_0: \rho_{X,Y|Z}=\rho$ & Scatter Plot & \begin{minipage}[t]{3cm}  	\scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot of Residuals obtained by fitting partial regression.
-For lineup plot, we simulate data (mean 0 and variance 1) with specific correlation $\rho$.
-\\
-8 & $H_0:$ Model Fits & Histogram & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_goodness_simple.pdf}}  \end{minipage} & Histogram of the response data.
-For lineup plot, we simulate data from fitted model.
-\\[1ex] % [1ex] adds vertical space
-9 & Special case $p=1$ $H_0: \rho_{X,Y} =\rho$ & Scatter plot & \begin{minipage}[t]{3cm}  \scalebox{0.4}{\includegraphics{stat_intercept.pdf}}  \end{minipage} & Scatter plot with least square line overlaid For lineup plot, we simulate data with correlation $\rho$.\\
-\hline 
-\end{tabular} 
-\label{tbl:stat_multiple} 
-\end{table*}
-
-Suppose $X_k$ is a categorical variable with two levels, and we test the hypothesis $H_0:\beta_k=0$ vs $H_1: \beta_k \ne 0$.
-If the responses for the two levels of the categorical variable $X_k$ in the model are different, the  residuals from fitting the null model should show a significant difference between the two groups.
-For a visual test, we draw boxplots of the residuals conditioned on the two levels of $X_k$.
-If $\beta_k\ne 0$ the boxplots should show a vertical displacement.
-
-%A lineup including the test statistic  for a binary $X_k$ is shown in Figure \ref{fig:test_category}.
-The 19 null plots are generated  by simulating residuals from $N(0,\hat{\sigma}^2)$.
-The test statistic, the plot containing the actual data, is randomly placed among these null plots.
-If the test statistic is identifiable the null hypothesis is rejected.
-% with a $p$-value of at most 0.05.
-
-%\subsection{Expected power}
-
-% Di: This is specific for the regression model
-%Consider the hypothesis test $H_0: \beta_k=0$ against $H_1: \beta_k \ne 0$ in model \ref{multi}.
-%%Now consider estimating the power of the visual test. 
-%%Suppose $F_{|t|}(.)$ is the distribution function of an absolute value of $t$.
-For the regression slope parameter $\beta$ suppose the associated test statistic in conventional test be $t_{obs}$ with $p$-value $p_D$.
-By definition we have
-%%
-%%$$p_D=Pr(|t| \ge t_{obs}| H_0)=1-F_{|t|}(t_{obs}) \Rightarrow |t_{obs}|=F_{|t|}^{-1}(1-p_D)$$
-%%
-%%\red{HH: the following paragraph should probably move up, but I'm not sure how to change it appropriately. Mahbub, could you take care of that?}
-%%Now suppose $F_{|t|}(.;\delta)$ denotes the distribution function of an absolute value of $t$ distribution with non-centrality parameter $\delta$. Thus the distribution function of $p-value$ be  
-%%\begin{eqnarray}\label{dist_p}
-%%F{p_D}(p) &=& Pr(p_D \le p)=1-Pr(1-p_D \le 1-p) \nonumber \\
-%%  &=& 1-Pr(F_{|t|}^{-1}(1-p_D) \le F_{|t|}^{-1}(1-p) ) \nonumber \\
-%%  &=& 1-Pr(|t_{obs}| \le F_{|t|}^{-1}(1-p)) \nonumber \\
-%%  &=&\left\{ \begin{array}{ll}
-%%          1-F_{|t|}( F_{|t|}^{-1}(1-p))=p &\mbox{ ; under $H_0$} \\
-%%          1-F_{|t|}( F_{|t|}^{-1}(1-p); \delta) &\mbox{ ; under $H_1$} 
-%%       \end{array} \right.     
-%%\end{eqnarray}
-%%
-%%Thus the density of $p_D$ is Uniform(0,1)  under $H_0$.
-As noted by \cite{Ruppert:2007}, under $H_1$ the density $$f_{p_D}(p_D; \delta)= \frac{f_{|t|}(F_{|t|}^{-1}(1-p_D);\delta)}{f_{|t|}(F_{|t|}^{-1}(1-p_D))}$$ derived from equation \eqref{dist_p}  is a right skewed distribution.
-%In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$.
-Suppose  $p_0=\min_i(p_{0,i})$.
-Thus $p_0 \sim \text{Beta}(1,m-1)$.
-
-The conventional test in this scenario uses $T= \hat{\beta_k}/ se(\hat{\beta_k})$ and rejects the null hypothesis, if $T$ is extreme on the scale of a $t$ distribution with $n-p$ degrees of freedom.
-It forms the benchmark upon which we evaluate the visual test.
-To calculate what we might expect for the power of the visual test, under perfect conditions, first assume that the observer is able to pick the plot with the smallest $p$-value from a lineup plot.
-This leads to the decision to reject $H_0$ when $p_{D} < p_0$, where $p_{D}$ is the conventional $p$-value as details given in Lemma \ref{lemma}.
-Thus the expected power of the visual test in this scenario is
-
-\begin{equation}\label{power_exp} 
-   Power(\beta)=Pr(p_{D} < p_0)  \quad \text{for}  \quad \beta \ne 0
-\end{equation}
-Figure \ref{fig:power_expected} shows the power of the conventional test in comparison to the expected power of the visual test obtained from Equation \ref{power_exp}.
-Notice that the expected power of the visual test is %almost as good as
-close to the power of the conventional test.
-
-%Theorem \ref{thm:power} helps to compute the expected power of visual inference.
-%\begin{thm}\label{thm:power}
-% equation \eqref{power_exp}  yields expected power of visual inference as Power($\beta$) = $E(1-p_D)^{m-1}$ .
-%\end{thm}
-% 
-%{ \em proof}
-%
-%Since $p_0 \sim \text{Beta}(1,m-1)$,  for $t \in (0,1)$ we have the distribution function of $p_0$ be
-%\begin{eqnarray*}
-%F{p_0}(t) &=& (m-1) \int_0^t{(1-p_0)^{(m-2)}dp_0} \\
-%  &=&  -(m-1) \int_1^{1-t}{x^{(m-2)}dx}\\
-%  &=& \left[ x^{m-1}\right]^1_{1-t}\\
-%  &=&1-(1-t)^{m-1} \rightarrow 1 \quad \text{as} \quad m \rightarrow \infty
-%\end{eqnarray*}
-%
-%% verified this computation from http://www.wolframalpha.com using command int (m-1)*(1-p)^(m-2) dp, p=0,t
-%
-%Thus expected power in equation \eqref{power_exp} be
-%\begin{eqnarray*}
-% Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
-%  &=& 1- \int_{0}^{1}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
-%  &=& 1-  \int_{0}^{1}{F{p_0}(t) f_{p_{B} }(t)dt } \\
-%  &=& E(1-p_D)^{m-1}   \\
-%%  & \rightarrow &1-  \int_{0}^{1}{ f_{p_{B} }(t)dt }  = F_{p_D}(0) \quad \text{as} \quad m \rightarrow \infty
-%\end{eqnarray*}
-%
-%
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{.6}{\includegraphics{power_expected.pdf}}
-       \caption{Expected power of visual test  and the power of conventional test for sample size $n =100$, $\sigma = 12$ and $m$=20. }
-       \label{fig:power_expected}
-\end{figure}
-
-% \pagebreak
-
-% -------- Heike's writeup added here
-
-%\subsection{Some Power Considerations}
-
-%\subsection{Expected number of choices}
-%Since $X$ has a binomial distribution, we can have a look at the number of plots among the null plots that we would expect to be picked over the data plot:
-%\[
-%E[X] = (m-1)(1-E[p_D]).
-%\]
-%With an increase of the lineup size $m$ the number of null plots with a potentially stronger signal than the data plot grows linearly.
-%This should allow us to infer some of the signal strength $p_D$ as
-%\begin{equation} \label{plot_signal}
-%E[p_D] = 1 - E[X]/(m-1),
-%\end{equation}
-%i.e.
-by averaging over the number of plots with  a stronger signal than the data plot we can evaluate signal strength of the plot even in the case where the data plot is not being picked by the observer.
-%
-
-\section{Human Subjects Experiments with Simulated Data} \label{sec:simulation}
-
-Three experiments were conducted to evaluate the effectiveness of the lineup protocol relative to the equivalent test statistic used in the regression scenario.
-The first two are conventional settings where we would not expect the lineup protocol to do better than the conventional test, but we hope it to perform favorably.
-The third experiment is a scenario where assumptions required for the conventional test are violated, and we would expect the lineup protocol to outperform the conventional test.
-(Data and lineups used in the experiments are available in the supplementary material.)
-
-After many small pilot studies with local personnel, it was clear that some care was needed to set up the human subjects experiments.
-It was best for a observer or a subject to see a block of 10 lineups with varying difficulty, with a reasonable number of ``easy'' lineups.
-The explanations about each experiment (below) includes an explanation of how the lineups were sampled and provided to the subjects, for each of the three experiments.
-
-Participants for all the experiments were recruited through Amazon's online web service, Mechanical Turk  \citep{turk}.
-A summary of the data obtained for all three experiments are shown in Table \ref{tbl:summary}.
-Participants were asked to select the plot they think best matched the question given, provide a reason for their choice, and say how confident they are in their choice.
-Gender, age, education and geographic location of each participant are also collected.
-%The data was collected through the web.
-For each of the experiments one of the lineups was used as a test plot (easy plot) which everyone should get correct, so that a measure of the quality of the subjects effort could be made.
-
-Note that no participant was shown the same lineup twice.
-%\green{Something probably needs to be said about subjects who do multiple tests, what checks are done.} \blue{We do have some data where we notice multiple responses for a lineup by a subject.
-This happens when subject accidentally use browser's ``back'' button and same data is saved twice.
-We made sure in the data cleaning process that we do not include more than one response on a lineup by a subject.
-For this we only keep the very first response.}
-
-\subsection{Discrete covariate}\label{sec:category}
-
-The experiment is designed to study the ability of human subjects to detect the effect of a single categorical variable $X_2$ (corresponding to parameter $\beta_2$) in a two variable ($p=2$) regression model (Equation \ref{multi}).
-Data is simulated a range of values of $\beta_2~ (=0, 1, 3, 5, 7, 8, 10, 16)$, two different sample sizes ($n=100, 300$) and two standard deviations of the error ($\sigma=5, 12$).
-The range of $\beta_2$ values was chosen so that estimates of the power would produce reasonably continuous power curves, comparable to that calculated for the theoretical conventional test.
-Values were fixed for other regression parameters, $\beta_0 = 5$,  $\beta_1=15$, and the values for $X_1$ were randomly generated from a Poisson $(\lambda=30)$ distribution, which is almost Gaussian.
-Data sets were generated with parameter values shown in Table \ref{tbl:experiment_params}.
-Three replicates at each level were generated.
-These produced 60 different ``actual data sets'', and thus, 60 different lineups.
-To produce the lineups, the null model was fit to each actual data set, to obtain residuals and parameter estimates.
-The actual data plot was generated by making  side-by-side boxplots of the residuals (Table \ref{tbl:stat_multiple}, case 3).
-The 19 null data sets were generated by simulating from  $N(0, {\hat{\sigma}}^2)$, and  plotted in the same way.
-The actual data plot was randomly placed among these null data plots to produce the lineup.
-Figure \ref{fig:test_category} is an example of one of these lineups.
-The actual data plot is number 11 which was generated for $n$=300, $\beta_2$=15 and $\sigma$=12.
-
-%{\bf Steps to obtain lineup}
-%\begin{enumerate}
-%	\item For a given parameter setting obtain actual data set
-%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-%	\item Generate actual data plot using the procedure as described in table \ref{tbl:stat_multiple}
-%	\item Generate 19 null (residual) data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
-%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-%\end{enumerate}
-
-The number of evaluations required for each lineup to provide reasonable estimates of the power is determined by the variance of the estimated power.
-Suppose $\gamma$ denotes the conventional test power for each parameter combination shown in Table \ref{tbl:experiment_params}.
-Since the expected power of visual inference is very close to the power of conventional test (Figure \ref{fig:power_expected}) the proportion of successful evaluations of lineups should give a rough estimate for  $\gamma$.
-For a given proportion $\gamma$ it is desired to have a margin of error (ME) less than or equal to 0.05.
-Thus we have $ME =1.96 \sqrt{ \gamma(1-\gamma) / n_{\gamma} } \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_{\gamma} \geq \frac{\gamma(1-\gamma)}{(0.05/1.96)^2}.$$
-% \hh{we're not even close to $n_\gamma$ with ten lineups .... we need more in the order of 400. Let's not emphasize this so much.}
-
-Each subject viewed at least 10 lineups with the option to evaluate more.
-%and we allowed them to provide more evaluations if they wanted. 
-Depending on the parameter combinations we group the lineups in different difficulty levels  as easy, medium, hard and mixed (actual numbers are given in the supplementary material).
-For each difficulty level a specific number of lineups was randomly picked  for evaluation.
-This number is chosen so that total number of evaluations for each lineup for that group exceed the threshold $n_{\gamma}$.
-To satisfy this plan we needed to recruit at least 300 subjects.
-
-\begin{table}[hbtp]
-\caption{Combination of parameter values, $\beta_2$,  $n$ and $\sigma$, used for each of the simulation experiments.} % title name of the table
-\centering
-\scalebox{.9}{
-\begin{tabular}{c c c c c}
-% creating 10 columns
-\hline\hline
-
-% inserting double-line
-& & \multicolumn{3}{c}{Slope ($\beta$)} \\
- \cline{3-5}
-
-Sample size  & Error SD  &  \multicolumn{1}{c} {Experiment 1}  & \multicolumn{1}{c} {Experiment 2}  & \multicolumn{1}{c} {Experiment 3} \\
- ($n$) &   ($\sigma$) & Discrete covariate & Continuous covariate & Contaminated data 
-\\ [0.5ex]
-\hline
-% inserts single-line
-
-% Entering 1st row
-&  5 & 0, 1,  3, 5, 8  & 0.25, 0.75, 1.25, 1.75, 2.75 & 0.1, 0.4, 0.75, 1.25, 1.5, 2.25\\[-1ex]
-\raisebox{1.5ex}{100} &12
-& 1, 3, 8, 10, 16  & 0.5, 1.5, 3.5, 4.5, 6 &\\[1ex]
-
-% Entering 2rd row
-&  5 & 0, 1, 2, 3, 5  & 0.1, 0.4, 0.7, 1, 1.5&\\[-1ex]
-\raisebox{1.5ex}{300} & 12
-& 1, 3, 5, 7, 10  & 0, 0.8, 1.75, 2.3, 3.5 &\\[1ex]
-% [1ex] adds vertical space
-\hline
-% inserts single-line
-\end{tabular}
-}
-\label{tbl:experiment_params}
-\end{table}
-
-%For a given proportion $p$ we want to have margin of error (ME) to be 0.1.
-Thus we have $ME =1.96 \sqrt{ \frac 1 n p(1-p)} \le 0.05$ which gives us the estimation of minimum number of evaluations $$n_p \geq \frac{p(1-p)}{(0.05/1.96)^2}.$$ Now from the theoretical (UMP) power curve for each parameter combinations shown in Table \ref{tbl:experiment_params} we know the power(say $p$) and thus we can estimate the number of evaluations required for that specific combinations of parameter.
-We collect data such that each of the parameter combinations in the Table \ref{tbl:experiment_params} has approximately the estimated sample size that we calculated here.
-
-%\green{Did you in fact bin like this???}\red{For each parameter combination a set of 1000 actual data are generated and their associated p-values were computed.
-This produces an empirical distribution of p-value under alternative hypothesis.
-Finally, three data sets that have 16th, 50th and 84th quantile of the p-values were taken.
-Similar procedure was used to generate 19 null plots based on the empirical distribution of p-value under null hypothesis}.
-For added control, to ensure a signal in the simulated actual data a blocking structure was used to filter data sets.
-A 1000 sets were generated for each parameter combination and the conventional $t$-statistic and $p$-value associated with $H_0: \beta_2=0$ were calculated.
-The 3 replicates were drawn from each of three blocks of $p$-values: (0.0-$q_{33}$), ($q_{33}$-$q_{66}$), ($q_{66}$-1) where $q_i$ is the $i$th percentile in the distribution of the $p$-values.
-Additional control was applied to the 19 null plots.
-Because the distribution of these $p$-values should follow a Uniform(0,1) distribution, data sets were binned on this range by $p$-value, and a data set was randomly selected from each bin.
-
-\subsection{Continuous covariate}  \label{sec:continuous}
-
-This experiment is very similar to the previous one, except that there is a single continuous covariate and no second covariate (Equation \ref{multi} with $p=1$), following the test in Table \ref{tbl:stat_multiple}, case 2.
-Data is simulated with two sample sizes ($n=100, 300$), two standard deviations of the error ($\sigma=5, 12$), a variety of slopes ($\beta$), as given in Table  \ref{tbl:experiment_params}.
-We arbitrarily fix $\beta_0 = 6$ and values for $X_1$ are simulated from $N(0,1)$.
-For each combination of parameters, three different actual data sets are produced, yielding a total of 65 lineups.
-For some parameter combinations we have more than three replications, because additional lineups were produced for values of the parameters that exhibited large variation in the initial data.
-
-The actual data plot is generated by making a scatterplot of $Y$ vs $X_1$ with the least squares regression line overlaid.
-To produce the null plots in the lineup null data was simulated from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and plotted using the same scatterplot method as the actual data.
-To select 10 lineups for a subject, each combination of sample size ($n$) and error SD ($\sigma$) is given a difficulty value based on the slope ($\beta$) parameters.
-For the smallest slopes the difficulty is 4 (hardest) and for the largest slopes the difficulty is 0 (easiest).
-Figure \ref{fig:test_continuous} shows an example lineup for this experiment from difficulty level 4.
-This lineup is generated using sample size ($n$)=100, slope($\beta$)=1.25 and and error SD($\sigma$)=5.
-The actual data plot location is 11.
-
-For each combination of sample size and standard deviation, the subject sees a randomly selected lineup of each difficulty level, all the slopes, giving 5 lineups.
-Another four lineups are chosen from a second choice of sample size and standard deviation, with difficulty levels 0-3, and the last lineup shown was a randomly selected lineup with difficulty 0.
-The order that the lineups are shown to participants is randomized.
-Each subject saw a substantial number of difficult lineups.
-
-%\hh{While this allows us a better estimation for the power in small slopes, }
-%%The intention was noble, to get a better handle on the power of the lineups for small slopes, 
-%but practically this made the study very tiring for the participants. %It would be better to provide the subjects with a larger selection of easier lineups and just a few difficult ones. 
-%\green{perhaps should go into so much detail about this, or keep this for the conclusions section.}
-
-%{\bf Steps to obtain lineup}
-%\begin{enumerate}
-%	\item For a given parameter setting obtain actual data set
-%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-%	\item Generate actual data plot (a scatter plot of $Y$ vs $X_1$ a least square line overlaid)
-%	\item Generate 19 null data sets from $N(X \hat{\beta}, {\hat{\sigma}}^2$) and obtain 19 null plots
-%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-%\end{enumerate}
-
-%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.}
-
-%{\bf Presenting lineups for evaluation} Each subject got at least 10 lineups for evaluation.
-For each combination of sample size ($n$) and error SD ($\sigma$) we define the difficulty of the lineup based on the five slope ($\beta$) parameters we choose as shown in Table \ref{tbl:experiment_params}.
-For the smallest slope parameter the difficulty is 0 and for the largest parameter the difficulty is 4.
-The 10 lineups are chosen as follows;
-
-%\begin{enumerate}
-%	\item For one combination of  sample size ($n$) and error SD ($\sigma$) randomly select five lineups with one from each difficulty level
-%	\item For another combination of sample size ($n$) and error SD ($\sigma$) select four lineups with one from each difficulty level 0 through 3
-%	\item Finally select another lineup with difficulty level 0
-%\end{enumerate}
-
-\begin{figure}[p]
-%\begin{figurehere}
-   \centering
-       \includegraphics[width=0.95\textwidth]{lineup_continuous.pdf}
-       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous.
-One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the steepest slope?
-(The position of the actual data plot is provided in Section \ref{sec:continuous}.)}
-       \label{fig:test_continuous}
-\end{figure}
-
-\subsection{Contaminated data} \label{sec:contamination}
-
-The first two simulation experiments use data generated under a normal error model, satisfying the conditions for conventional test procedures.
-In these situations there exists a test, and there would, in general, be no need to use visual inference.
-The simulation is conducted with the hopes that the visual test procedure,  will at least compare favorably with the conventional test -- without any ambition of performing equally well.
- This third simulation is closer to the mark for the purpose of visual inference.
-The assumptions for the conventional test are violated, by contaminating the data.
-The contamination makes the estimated slopes effectively 0, yet the true value of slope parameter is not.
-The data is generated from the following model:
-
-\[
-  Y_i = \left\{
-  \begin{array}{l l}
-    \alpha+\beta X_i + \epsilon_i  & \quad  X_i \sim N(0,1) \quad  i =1,...,n\\
-    \lambda+ \eta_i & \quad X_i \sim N(\mu,1/3) \quad  i=1, ...,n_c\\
-  \end{array} \right.
-\]
-where $\epsilon_i \stackrel{iid}\sim N(0,\sigma)$, $\eta_i \stackrel{iid}\sim N(0,\sigma/3)$ and $\mu = -1.75$.
-$n_c$ is the size of the contaminated data.
-For the experiment we consider $n=100$ and $n_c=15$ producing actual data with $115$ points.
-Further, $\alpha=0$, $\lambda=10$, and $\sigma$ is chosen to be approximately 3.5, so that error standard deviation across both groups of the data is $5$.
-%Figure \ref{fig:cont_dat} shows an example of contaminated data, generated by this model for $\beta=5$.
-A linear model (Equation \ref{multi} with $p=1$  and intercept $\beta_0=0$) is fit to the contaminated data.
-This experiment  follows the test in Table \ref{tbl:stat_multiple}, case 2.
-The actual data plot shows a scatterplot of the residuals vs $X_1$, and the null  plots are scatterplots of null data generated by plotting simulated residuals from $N(0, {\hat{\sigma}}^2$)  against $X_1$.
-
-Experiment three consists of a total of 30 lineups, made up of five replicates for each of the six slopes as shown in Table \ref{tbl:experiment_params}.
-We use the slope directly as a measure of difficulty, with difficulty = 0 for the largest slope and difficulty = 5 for the smallest slope.
-Subjects were exposed to a total of ten lineups, with two lineups from each of the difficulty levels 0 through 3, and one  lineup each from levels 4 and 5.
-
-%{\bf Steps to obtain lineup}
-%\begin{enumerate}
-%	\item For a given parameter setting obtain actual data set
-%	\item Fit the null model to the actual data and obtain residuals and other parameter estimates ($\hat{\beta}$ and $\hat{\sigma}$)
-%	\item Generate actual data plot (a scatter plot of Residual vs $X_1$)
-%	\item Generate 19 null residual data sets from $N(0, {\hat{\sigma}}^2$) and obtain 19 null plots
-%	\item Finally randomly place 19 null plots and one actual plot in a layout (4X5) to get a lineup 
-%\end{enumerate}
-
-An example lineup for slope $\beta=0.4$ is shown in Figure \ref{fig:test_contaminated}.  Can you pick which plot is different? The actual data plot location is 7.
-
-%\begin{figure}[hbt]
-%\begin{figurehere}
-%   \centering
-%       \scalebox{0.6}{\includegraphics{contaminated_data.pdf}}
-%       \caption{Scatterplot of a simulated contaminated data set generated such that a simple linear model produce a slope almost negligible, ie, if we fit model (Equation \ref{multi}) with $p=1$  to this data, the resulting estimate of $\beta_1$ would not be statistically significant.
-The conventional t-test $p$-value for the slope is 0.67 for this data set but visually we can see a convincing evidence of positive slope.
-%}
-%       \label{fig:cont_dat}
-%\end{figure}
-
-\begin{figure}[p]
-%\begin{figurehere}
-   \centering
-       \includegraphics[width=0.95\textwidth]{lineup_contaminated.pdf}
-       \caption{Lineup plot ($m=20$) using scatter plots for testing $H_0: \beta_k=0$ where covariate $X_k$ is continuous but the inclusion of some contamination with the data spoils the normality assumption of error structure.
-One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true.
-Which plot is the most different from the others, in the sense that there is the steepest slope?
-(The position of the actual data plot is provided in Section \ref{sec:contamination}.)}
-       \label{fig:test_contaminated}
-\end{figure}
-
-%\green{Fill in details of lineups, probably need a figure with an example lineup, and how lineups were assigned to subjects.
-Show a sample lineup, probably the most difficult, instead of the single plot.}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.6}{\includegraphics{power_observed.pdf}}
-%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-%       \label{fig:power_observed}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.7}{\includegraphics{power_loess_exp1.pdf}}
-%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 1.}
-%       \label{fig:power_loess1}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.7}{\includegraphics{power_loess_exp2.pdf}}
-%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of UMP test for sample size $n= 100,300$ and $\sigma = 12,5$ as per experiment 2.}
-%       \label{fig:power_loess2}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.7}{\includegraphics{power_diff_exp.pdf}}
-%       \caption{Difference of Observed power from UMP test power obtained in three experiments for sample size $n= 100,300$ and $\sigma = 12,5$ and values of slope parameter $\beta$ as shown in Table \ref{tbl:experiment_params}.}
-%       \label{fig:power_diff_exp}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.6}{\includegraphics{power_observed_exp3.pdf}}
-%       \caption{Observed power of visual test from equation \eqref{eqn:power_estimate} with pointwise 95\% confidence limits and the power of UMP test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-}
-%       \label{fig:power_observed_exp3}
-%\end{figure*}
-
-\section{Results} \label{sec:results}
-
-\subsection{Data Cleaning}
-
-% In this process of collecting data some participants did not provide demographic information as we see in Table \ref{tbl:summary} that male and female participants do not add up to total participants.
-It is not clear why they did not provide the information.
-Now the concern arises whether we should keep their responses in the study.
-It is also possible that participants were not serious enough to provide feedback.
-%For each of the experiments we had a test plot (easy plot) for which they should have a correct evaluation without much effort.
-This could be a criteria to determine their attentiveness in the study.
-For this we consider six different data cleaning criteria as shown below;
-
-%{\bf Screening scenarios:}
-Amazon Mechanical Turk workers are paid for their efforts, not substantial, but on the scale of the minimum wage of the USA.
-Some workers will try to maximize their earnings for minimum effort, which can affect the results from the data.
-For example, some workers may simply randomly pick a plot, without actively examining the plots in the lineup.
-For the purpose of identifying these participants and cleaning the data, we take action based on a subject's answer to one of the very easy lineups that everybody was exposed to:
-if the subject failed to identify the actual data plot, we remove all of this subject's data from the analysis.
-If the answer on the lineup is correct, we remove this lineup from the analysis, but keep all of the remaining answers.
-Table \ref{tbl:summary} tabulates the number of subjects, genders and lineups evaluated after applying the data screening procedure.
-
-%
-%one of these was picked at random as the filtering lineup, and action was taken depending on whether they answered correctly or incorrectly.
-The data of participants who did not answer correctly was removed.
-%For the subjects, who correctly answered the filtering lineup, all of their data was kept, except for that particular lineup.
-That is, for each block of 10 lineups, only the data from 9 lineups was used for each subject.
-
-%Amazon Mechanical Turk data suffers from the problem that some participants are trying to 'game' the system, because they are being paid for their efforts, not much by the standards of the US minimum wage, but enough to encourage honest efforts.
-in our situation, these are participants who did not put in a best effort to identify the data plot, but just randomly picked a plot to maximize their 'winnings'.
-In order to avoid using those results, we are comparing six different screening methods for excluded participants from the data analysis.
-%There are six suggested approach to clean the data. 
-%They are
-%\begin{enumerate}
-%\item {\bf include all participants} and their evaluations
-%\item exclude all participants' evaluations, who did {\bf not} share their {\bf demographic information} (age, gender education level -- all three pieces of information are either missing or all present).
-%\item exclude participants' records, if {\bf none of the evaluations}  correctly identified the data plot -- every participant was shown a range of `easy' lineups.
-%\item include participants' records, if   {\bf at least 20 percent} of the evaluations are {\bf correct}  -- based on ten evaluations per participants, two correct evaluations are significant evidence against a person just guessing
-%\item include participants' records, if at least {\bf 50\% of all very easy lineups are correct} 
-%\item use easy lineups as {\bf reference charts}: sample one easy lineup from a person's records.
-If that lineup is evaluated {\bf correctly}, include all (other) lineups of that person, otherwise exclude all lineup evaluations by this participant.
-%\end{enumerate}
-
-%\begin{table}[hbtp]
-%\caption{Number of unique subjects and their total feedbacks after applying six screening process on all three experimental data sets. 
-%Note that in some of the lines the number of male and female participants does not add up to the total number of participants.
-This is due to missing demographic information.
-\hh{Do we need this table or could we make this a visual comparison?
-For me, there's too many numbers with too many digits.}}
-%\begin{center}
-%\begin{tabular}{rrrrr|rrrr|rrrr}
-%  \hline
-%Screening &  \multicolumn{4}{c} {Experiment 1}  & \multicolumn{4}{c} {Experiment 2}  & \multicolumn{4}{c} {Experiment 3} \\
-% \cline{2-5}  \cline{6-9}   \cline{10-13}
-%criteria & Subj & Male & Fem & Total & Subj & Male & Fem & Total & Subj & Male & Fem & Total \\ 
-%  \hline
-%1 & 520 & 264 & 201 & 4914 & 390 & 203 & 182 & 4603 & 257 & 166 &  84 & 2830 \\ 
-%  2 & 465 & 264 & 201 & 4624 & 385 & 203 & 182 & 4558 & 250 & 166 &  84 & 2782 \\ 
-%  3 & 384 & 200 & 165 & 4187 & 384 & 199 & 181 & 4573 & 227 & 147 &  78 & 2648 \\ 
-%  4 & 318 & 168 & 135 & 3395 & 378 & 195 & 179 & 4492 & 205 & 133 &  72 & 2393 \\ 
-%  5 & 277 & 139 & 124 & 3008 & 374 & 193 & 177 & 4451 & 185 & 122 &  63 & 2121 \\ 
-%  6 & 239 & 121 & 107 & 2317 & 351 & 185 & 164 & 3858 & 156 & 106 &  50 & 1681 \\ 
-%   \hline
-%\end{tabular}
-%\end{center}
-%\label{tbl:summary}
-%\end{table}
-
-\begin{table}[hbtp]
-\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets.
-Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information.
-}%\blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
-\begin{center}
-\begin{tabular}{ccrrcc}
-   \hline \hline
- Experiment & Subject & Male & Female & Responses &Lineup\\ 
-    \hline
-1 & 239 & 121 & 107 & 2249 &  60 \\ 
-  2 & 351 & 185 & 164 & 3636 &  70 \\ 
-  3 & 155 & 103 &  52 & 1511 &  29 \\  
-   \hline
-\end{tabular}
-\end{center}
-\label{tbl:summary}
-\end{table}
-
-\subsection{Model fitting}
- For each parameter combination, the effect is computed as
-
-\[
-E=\sqrt {n} \cdot \beta/\sigma.
-\]
-\noindent
-The model in Equation \ref{eqn:mixed} is fit using $E$ as the only fixed effect covariate without intercept, i.e. 
- $W_{\ell i} = E_{\ell i}$.
-Instead of fitting an intercept, we make use of a fixed  offset of $\log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$.
-Different skill levels of subjects are accounted for  by allowing subject-specific random  slopes for  effect ($E$).
-
-For experiment 3 we do fit intercepts: both a fixed  and subject-specific random effects, since forcing power
-%While for experiment 1 and 2 we don't allow intercept but offset, for experiment 3 an intercept is estimated to allow the power curves to go above 0.05 when $E=0$. This is because forcing power 
-to be fixed at 0.05 for $E=0$ is not justified by the data. % (Figure \ref{fig:power_loess_effect}). *** You cannot reference the figure before you explain it!!!
-
-%For all the three experiments, each subject saw a lineup just once.
-This does not allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i} = \tau_i$.
-
-%The subject-specific probability, $p_i$, of successful evaluation of a lineup is then estimated using Equation \ref{eqn:mixed_power}. 
-For computation we use package {\tt lme4} \citep{lme4:2011} and software R 2.15.0 \citep{R}. $p$-value calculations are based on asymptotic normality.
-
-Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.
-The fixed effects estimates indicate that for all experiments  the proportion of correct responses increases as the effect increases.
-This effect is less pronounced  for experiment 3.
-The subject-specific variability is smaller for experiment 1, and relatively large for experiment 3.
-%The larger variability in the response data for experiment 3% as we see in Figure \ref{fig:power_loess_effect}  *** You cannot reference the figure before you explain it!!!
-%contributes to the random effect variance estimate be much larger compared to those of other experiments.
-For the same reason we see the standard error for intercept estimate is much higher as well.
-Also notice the slope estimate for experiment three is lowest as well.
-It is because for small parameter values the power is much higher for experiment 3.
-% \green{What do we learn about the models from these parameter estimates?
-Explain this here too.}
-
-%\green{This section needs to have sufficient detail of the model fitting, output and diagnostics to feed into the next few sections discussing the findings.}
-
-\begin{table}[hbtp]
-\caption{Parameter estimates of model in Equation \ref{eqn:mixed}. Estimates are highly significant with $p$-value $<$  0.0001 for all three experiment data.}
-\begin{center}
-\begin{tabular}{cr@{.}lcc}
-  \hline \hline
- &  \multicolumn{3}{c} {Fixed effect}  & Random effect\\
- \cline{2-4}
- Experiment & \multicolumn{2}{l}{Estimate}  &Std. error & Variance\\
-  \hline
-  1 & 0&39 & 0.0094 & 0.0080 \\ 
-  2 & 1&21 &  0.0197 &  0.0443 \\ 
-  3 & 0&59 (Intercept)  &   0.1668 & 1.9917\\ 
-     & 0&21 (Slope)    &  0.0511     &  0.0245\\ 
-     &-0&78 (correlation) & & \\
-   \hline
-\end{tabular}
-\end{center}
-\label{tbl:model_par}
-\end{table}
-
-\subsection{Power comparison}
-
- Figure \ref{fig:power_loess_effect} shows an overview of estimated power  against effect for the three experiments.
-Responses from each experiment are summarized by effect size and represented as dots,  with size indicating the number of responses.
-A loess  fit to the data gives an estimate of  the observed power for different effect sizes, with grey bands indicating simultaneous bootstrap confidence bands \citep{buja}.
-For comparison, the dashed lines show the corresponding power curves of the conventional tests.
-Not surprisingly, the observed power for experiments 1 and 2 is lower than the power of the conventional test.
-It is encouraging to see that visual inference mirrors the power vs effect relationship of conventional testing, and comes quite close in experiment 2, where scatterplots are used as the test statistic.
-Note that for effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2, making the fixed offset a reasonable assumption.
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.55}{\includegraphics{power_loess_effect.pdf}}
-       \caption{Power in comparison to effect for the three experiments.
-Points indicate subject responses, with size indicating count.
-Responses are 1 and 0 depending on the success or failure respectively to identify the actual plot in the lineup.
-The loess curve (continuous line) estimates the observed power, and surrounding bands show simultaneous bootstrap confidence band.
-Conventional test power is drawn as a dashed line.
-For experiment 3, conventional power is based on the slopes of the linear part of the non-contaminated data.
-Power for contaminated data is shown by cross mark.}
-       \label{fig:power_loess_effect}
-\end{figure}
-
-Results for experiment 3 are quite different.
-This is the situation where we expect to see the potential of visual inference, and indeed we do: the power of visual inference is always high, and much higher than the conventional test at small effect sizes.
-There is no actual conventional power in this situation, because assumptions are violated.
-The dashed line shows conventional power based on uncontaminated data, whereas the cross marks show effective power based on the coefficient estimated from the contaminated data.
-% It is computed assuming that there is no contamination, using the non-contaminated part of the simulation model.
-The  conventional power of true effect of the contaminated data is computed anyway and presented in figure \ref{fig:power_loess_effect} by cross marks.
-For this we obtain the estimated slope for the contaminated data and compute the estimated effect for which conventional power is calculated.
-% 
-%Note that the power structure is completely broken for contaminated data. %For smaller original effect we get huge power and that explains why we see so many small $p$-values for experiment 3
-
-Results of experiment 3 are curious insofar, as  power of the visual test is  high independent of effect size.
-However, these results are based on correct identification of the actual data plot, regardless of reason.
-Although subjects were asked to select the plot that exhibited the highest association between the two variables, they might have cued in on the cluster of contaminated data.
-This will be explored further in Section \ref{sec:TypeIII}.
-
-%(scatter plot) is better than what we observed for experiment 1 (box plot).
-For experiment 3, observed power is always above the conventional test power since for contaminated data we can visually see a trend while conventional test fail to address that.
-
-%The power curves obtained by fitting model \ref{eqn:mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}.
-This shows how the screening criteria may affect the results.
-The corresponding power curves of conventional test are also shown.
-Notice that no mater what criteria we apply the result does not change much for experiment 2.
-Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.
-
-%
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.70}{\includegraphics{power_mixed.pdf}}
-%       \caption{The overall power estimated from model \ref{eqn:mixed} are shown. The corresponding power curve for conventional test is shown for comparison.}
-%       \label{fig:power_mixed}
-%\end{figure}
-
-%Each participant was shown a sequence of 10 lineup plots.
-In total, 3629 lineups were evaluated by 324 people coming from many different locations across the globe.
-The results of the experiment are summarized in Figure \ref{fig:power_observed} which shows the observed power from the survey data calculated using equation \eqref{eqn:power_estimate} along with 95\% confidence interval calculated using Fisher's exact method.
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
-%       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
-%       \caption{Observed power shown by loess smoother and the power of conventional test for different effect size $E$ as per experiment 1 and 2.}
-%       \label{fig:power_effect1}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp1.pdf}}
-%       \scalebox{0.45}{\includegraphics{effect_power_subject_exp2.pdf}}
-%       \caption{Subject specific power obtained from equation \eqref{eqn:effect} and the power of UMP test for different effect size $E$ as per experiment 1 and 2.}
-%       \label{fig:power_effect_subject1}
-%\end{figure*}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
-%       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3.
-Notice that for experiment 3, power does not seem to depend on the parameter $\beta$.
-\blue{Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met.
-Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}}
-%       \label{fig:power_loess3}
-%\end{figure*}
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.40}{\includegraphics{power_model.pdf}}
-%       \caption{Estimated power curve from equation \eqref{eqn:power} along with 95\% confidence interval for sample size $n$ = 100 and $\sigma$ = 12.
-The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
-%       \label{fig:power_model}
-%\end{figure}
-
-%We fit model \eqref{eqn:mixed} to the survey data obtained from the simulation experiment.
-The estimated overall power curve obtained from equation \eqref{eqn:power} is shown in Figure \ref{fig:power_model}.
-Model \ref{eqn:mixed} also gives the subject specific power curves shown in Figure \ref{fig:power_subject}.
-The plot includes 20 randomly selected subject-specific power curves.
-Notice that the power curve estimated for one subject is above the UMP test power curve.
-
-%\begin{figure}[hbtp]
-%   \centering
-%       \scalebox{0.45}{\includegraphics{power_subject.pdf}}
-%       \caption{Estimated subject specific power curve from model \ref{eqn:mixed} for sample size $n$ = 100 and $\sigma$ = 12.
-The corresponding power curve for Uniformly Most powerful (UMP) test is shown for comparison.}
-%       \label{fig:power_subject}
-%\end{figure}
-
-\subsection{Subject-specific variation}
-
-%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{eqn:mixed}).   
-Figure  \ref{fig:power_mixed_subject} shows power curves for both the overall experiment and subject-specific variations.
-The thick continuous line shows overall estimated power, the thinner lines correspond to subject-specific power curves.
-For comparison, the  dashed lines show power curves of the conventional test.
-Subject-specific power is quite different between the three experiments.
-In experiment 1 subjects performed similarly, and substantially worse than the conventional test.
-In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects.
-In experiment 3 there is the most subject-specific variation.
-Some subjects performed substantially better than the conventional test, and on average the visual test was better.
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
-       \caption{Subject-specific  power estimated from model \ref{eqn:mixed} are shown.
-The corresponding power curve for conventional test (dashed line) is shown for comparison.
-The overall average power curve estimated from the model is shown (light blue).}
-       \label{fig:power_mixed_subject}
-\end{figure}
-
-\subsection{Estimating the $p$-value in the real world}
-
-%\green{third experiment doesn't really have a p-value, does it?
-Need to explain this.
-Why are there still so many small p-values in experiment 3?} \blue{$p$-values are computed fitting the full model to the contaminated data anyway.
-So many small p-values are coming from the fact that contamination makes the slope negatively significant.
-For example, for some slope the contamination reverses the slope (positive to negative) and negative slope got significant.
-It is like the effect of influential outlier.}
-
-In the real setting, where visual inference is to be useful, there will be no conventional tests $p$-values.
-Assessing the strength of perceived structure is a critical component of visual inference.
-In experiments 1 and 2, there is a $p$-value associated with the actual data plot in each lineup.
-As the $p$-value increases the proportion of correct responses falls (Figure \ref{fig:pval_pcorrect}), which is evidence of direct association between proportion of correct responses and conventional test $p$-values.
-For  $p$-values  larger than 0.15 it is very uncommon for subjects to correctly identify the actual data plot in the lineup.
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{p_val_prop_correct.pdf}}
-       \caption{Proportion of correct responses decreases rapidly with the increase of p-value.
-After p-value exceeds 0.15 it is very unlikely to identify the actual plot.
-The theoretical justification of this is shown in Figure \ref{fig:pval_power}.
-}
-       \label{fig:pval_pcorrect}
-\end{figure*}
-
-From the experimental data the visual $p$-values are estimated based on Definition \ref{dfn:pvalue}.
-Figure \ref{fig:pval_definition} displays resulting estimates for each lineup against the conventional $p$-value.
-The pattern of visual $p$-values is interesting: for small $p$-values the visual estimates tend to be very small, while lineups with larger $p$-values result in very large visual estimate,
-% If there is a signal in the data plot, visual $p$-value has a tendency to take a smaller value (small enough to reject $H_0$) and if not it takes a larger value (large enough to not reject $H_0$) 
-% 
- giving a clear indication to reject $H_0$ or not.
-This is why we do not see lot of visual $p$-values  between 0.05 and 0.8 especially for experiment 2.
-This guides the researcher to make decision confidently while conventional $p$-values take some marginal $p$-values making the decision whether to reject or not harder.
- % for which researchers get confused whether to reject or 'not reject' $H_0$. 
- For visual tests this is not common.
-
-\begin{figure*}[hbtp]
-   \centering
-       \scalebox{0.6}{\includegraphics{p_val_definition.pdf}}
-       \caption{Conventional test $p$-value ($p_D$) vs visual $p$-value obtained from the definition . Values are shown on square root scale. }
-    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
-       \label{fig:pval_definition}
-\end{figure*}
-
-%We also see the same pattern from the loess smoother fitted through the points in Figure \ref{fig:pval_definition}. 
-%The sigmoid shaped loess curves indicate a clear separation of visual $p$-values.
-For experiment 1, most of the visual $p$-values are above 0.4 when conventional $p$-values are above 0.2.
-This is more evident for experiment 2 where most of the visual $p$-values are above 0.8.
-
-For experiment 3, we see that  the visual $p$ values are very small no matter what the conventional $p$-values are.
-This is expected since conventional test fails to reject $H_0$ where we should reject, whereas the visual test performs well.
-% and the visual test does the job nicely.
-
-%With multiple subjects it is possible to estimate the visual $p$-value, and a measure of plot signal strength.
-The visual $p$-value is calculated from the binomial formula (Equation \ref{binom}).
-Assuming $H_o$ is true, the probability of picking the actual data plot by chance is $1/m$, and the visual $p$-value is the probability that $x$ or more subjects select that plot.
-Figure \ref{fig:pval_definition} displays this visual $p$-value for each lineup against the conventional $p$value.
-
-%Similary, the strength of the signal, is calculated using the proportion of subjects picking actual data plot.
-It is closely related to the power of the test.
-From Equation \ref{eqn:power_estimate}  we can write $1 - E[N_m]/K \approx (m-1)E[p_D] $.
-This gives the signal strength estimate as
-%\[
-%\hat p_D= (1- {u}/{K})/({m-1})
-%\]
-%where $K$ is the total number of independent evaluations on a lineup and $u$ is the number of successful evaluations.
-Figure \ref{fig:pval_plot_signal} shows the plot signal strengths in relation with conventional test $p$-values, on a square root scale.
-For experiment 3, technically, there are no conventional $p$-values, but the $p$-values are computed in the same manner as experiment 2, assuming that the data satifies the assumptions.
-The blue line is a loess fit to the data, and the grey line indicates equality of the two axes.
-The relationship between the conventional $p$-value and the plot signal strength is very strong, particularly when the values are small.
-As the conventional $p$-value increases, the estimated signal strength increases -- subjects find it harder to detect the actual data plot at larger $p$-values.
-The rise is very sharp for experiment 1, which indicates that subjects have a harder time detecting the actual data plot for even small $p$-values.
-
-%\green{Where is Figure \ref{fig:pval_definition} referred to and discussed?}
-% \blue{Are we going to keep both figures \ref{fig:pval_definition} and \ref{fig:pval_plot_signal}?}
-%
-%\green{Explain why the term ``plot signal'' is introduced}
-
-%, to assess the signal strength we examine the proportion of correct picks, with multiple users and the $p$-value of the actual data plot.
-For experiment 3 the \green{How is p-value calculated here?} Figure \ref{fig:pval_pcorrect} \blue{(should I bring this figure back?)}  displays scatterplots of these two variables for each experiment.
-As the $p$-value increases the proportion of correct responses falls, evidence of direct association between proportion of correct responses and conventional test $p$-values.
-When $p$-value is larger than 0.15, it is almost impossible to correctly identify the observed plot in the lineup.
-We observed this for both experiment 1 and 2.
-For experiment 3, the scenario is different, because there are two structures, the association between variables, and the cluster of contaminated points.% as in this case people might have observed something other than the effect $E$.
-
-%Yet, it is possible to obtain an estimate the strength of the signal in the actual data plot, when there are multiple evaluations, using Lemma \ref{lemma}.
-
-%Note that $\hat{p}_D$ is bounded within $(0, 1/(m-1))$.
-This estimate has to be interpreted carefully, since the choice of a large lineup size, $m$, will make the value look `significant' in a conventional sense.
-Values close to the upper bound of $1/(m-1)$ should therefore be interpreted as `at least as high', rather than their actual numeric value.
-%\hh{For $p$-values below $1/(m-1)$ visual $p$-values are very similar to actual $p$-values, the correlation for those values is: 0.36, 0.79, and -0.48 based on $n_1 = 44, n_2 = 45$, and $n_3= 26$ lineups, respectively.
-}  \blue{This is not happening as we see in figure \ref{fig:pval_plot_signal}.
-But it is true for experiment 2, right?.
-}
-%is equivalent to conventional test $p$-value when $p$-value is less than or equal to $1/(m-1)$.  
-%and beyond that the plot signal strength levels off on a point bigger than 0.05. Thus if plot signal strength is less than 0.05, we may decide to reject the null hypothesis.
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \scalebox{0.5}{\includegraphics{p_val_plot_signal.pdf}}
-%       \caption{Conventional test $p$-value ($p_D$) vs plot signal strength ($\hat p_D$).
-The solid line is a 45 degree line which represents the equity shown in Equation \ref{eqn:power_estimate}.
-p-values are shown on square root scale scale.}
-%    %   \hh{Mahbub, another change: could you change the y label to 'estimate of visual p-value $\hat{p_D}$ ?}
-%       \label{fig:pval_plot_signal}
-%\end{figure*}
-
-\subsection{Do people tend to pick the lowest $p$-value?}
-
- One assumption made in order to calculate the effect of lineup size in the calculations of visual $p$-value and signal strength was that subjects would tend to pick the plot in the lineup that had the strongest signal.
-In experiments 1 and 2, this corresponds to the plot with the smallest $p$-value.
-This can be examined for the data collected, to see if this assumption is, indeed, reasonable.
-%In Section \ref{sec:size} in order to make comparisons with conventional tests an assumption was made that the subject would pick the plot in the lineup that had the lowest $p$-value.
-In this section we take a look at the picks that subjects made, to see if this is, indeed, true.
-
-Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2} show the selection counts and $p$-values of all plots for each lineup in experiments 1 and 2, respectively.
-In each of these plots, the log of $p$-value for each plot in the lineup is plotted horizontally and the count of the number of subjects who selected this plot is displayed vertically.
-The dot represents this point, and the lines connect vertically with 0 count.
-Columns are ordered according to effect size as defined in section 6.2; rows show replicates for the same combination of parameters.
-
-%Each row corresponds to a different $\beta$ value ordered top to bottom from lowest to highest.
-Columns correspond to the other levels of the experiment, sample size $(n)$, error standard deviation $(\sigma)$ and replicate.
-(Empty cells indicate no lineup shown with this combination.)
-
-Red indicates the plot with the lowest $p$-value in the lineup.
-Blue indicates the plot of the actual data when it is different from that with the lowest $p$-value.
-In both experiments  people tended to select the plot with the lowest $p$-value with an increase in effect size.
- %This is less so the case when $\beta$ is small, when standard deviation is large and sample size is small. 
- The results are clearer for experiment 2, that used a continuous covariate.
-But even when subjects did not pick the plot with the lowest $p$-value they tended to oscillate their choices between the several low $p$-value plots.
-So for most subjects, the assumption that they pick the plot with the smallest $p$-value would appear to be reasonable, and the actual power of the visual test should be close to the expected power.
-
-\begin{figure*}[hbtp]
-   \centering
-      \includegraphics[width=0.95\textwidth]{p_val_log_counts-a.pdf}
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts-b.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 60 individual lineups for experiment 1.
-Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
-The columns of both top and bottom panel are ordered according to effect size, rows within panels show replicates for each parameter combination.}
-       \label{fig:P-val_log}
-\end{figure*}
-
-\begin{figure*}[hbtp]
-   \centering
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-a.pdf}       
-       \includegraphics[width=0.95\textwidth]{p_val_log_counts2-b.pdf}
-       \caption{Relative frequency of plot picks compared to other plots in the lineup plotted against the $p$-value (on log$_{10}$ scale) of each plot for all of the 70 individual lineups for experiment 2.
-Red indicates the plot with the lowest $p$-value, and blue indicates the actual data plot, when it is different from that with the lowest $p$-value.
-Columns are ordered according to effect size, with rows showing replicates of the same parameter combination on top of each other.
-       Empty cells indicate combination of parameters with fewer than five replicates. 
-       Highest counts tend to be the plot in the lineup having the lowest $p$-value, even more so than for experiment 1.
-       }
-       \label{fig:P-val_log2}
-\end{figure*}
-
-There are some noticeable exceptions to this rule. In experiment 1, when $\beta=0, n=100, \sigma=5, rep=1$ people overwhelmingly chose a plot with much larger $p$-value, similarly, for 
-parameters $\beta=5, n=300, \sigma=12, rep=3$, people tended to pick the plot with the second smallest $p$-value. 
-For several of these exceptions, along with several easy lineups, a follow up experiment was conducted using an eye-tracker to examine which patterns or features participants are cueing on in making their choices \citep{zhao:2012}.
-%(Lineups chosen for the eye-tracker experiment were in experiment 1, (1) $n=100, \beta=3, \sigma=5, rep=1$ (2) $n=100, \beta=5, \sigma=5, rep=1$ (3) $n=100, \beta=10, \sigma=12, rep=1$ and (4) $n=300, \beta=5, \sigma=5, rep=1$, and in experiment 2, (5) $n=100, \beta=3.5, \sigma=12, rep=1$ (6) $n=100, \beta=4.5, \sigma=12, rep=2$ (7) $n=300, \beta=1, \sigma=5, rep=1$ and (8) $n=300, \beta=1, \sigma=5, rep=2$.)
-
-%Experiment 3 does not have $p$-values associated with each plot, so the assumption can only be tested with the first two experiments.
-Practically, this is not an assumption that is explicitly made, because there will not be a $p$-value available.
-Rather it is assumed that the subject has the ability to select the plot with the most extreme structure.
-A few of the lineups indicated that occasionally people cue on different patterns, for example, in experiment 1 the lineup $\beta=0, n=100, \sigma=5, rep=1$ (first row, fourth column) many subjects chose the same plot, that had a higher $p$-value.
-Some of these anomalous lineups were selected for follow-up studies using an eye-tracker to investigate what people were looking at.
-
-%Number of lineups for which most of the subjects picked the data plot based on the minimum $p$-values are shown in Table \ref{tbl:pval_assumption} for both experiment 1 and 2.
-The $p$-values of lineups which were not chosen based on minimum $p$-value are also very small (close to minimum) as we see in both Figures \ref{fig:P-val_log} and \ref{fig:P-val_log2}.
-We see the same pattern in Figure \ref{fig:P-val_rank} where the median number of people who pick the data plot with minimum $p$-value (rank 1) are much higher than other ranks.
-
-%\begin{table}[hbtp]
-%\caption{Number of lineups for which most people pick the plot with minimum $p$-value. The data used is from screening criteria 1. } 
-%\begin{center}
-%\begin{tabular}{cccc}
-%  \hline
-%screening & Experiment & Total &  minimum $p$-value\\ 
-%criteria & & lineups & picked by the most \\
-%  \hline
-%   & 1 &  60 &  37 \\ [-1ex]
-%    \raisebox{1.5ex}{1} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  38 \\  [-1ex]
-%    \raisebox{1.5ex}{2} & 2 &  70 &  61 \\ 
-%\hline
-%     & 1 &  60 &  38 \\  [-1ex]
-%   \raisebox{1.5ex} {3} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  41 \\  [-1ex]
-%    \raisebox{1.5ex}{4} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  39 \\  [-1ex]
-%    \raisebox{1.5ex}{5} & 2 &  70 &  60 \\ 
-%\hline
-%     & 1 &  60 &  40 \\  [-1ex]
-%    \raisebox{1.5ex}{6} & 2 &  70 &  60 \\ 
-%   \hline
-%\end{tabular}
-%\end{center}
-%\end{table}
-
-%\begin{table}[hbtp]
-%\caption{Number of lineups for which most people pick the plot with minimum $p$-value.  } 
-%\begin{center}
-%\begin{tabular}{ccc} \hline
-%screening & \multicolumn{2}{c} {lineups picked by minimum $p$-value} \\
-% \cline{2-3}
-%criteria & Experiment 1 & Experiment 2 \\ 
-%  \hline
-%1 & 37 & 60 \\ 
-%2 & 38 & 61 \\  
-%3 & 38 & 60 \\ 
-%4 & 41 & 60 \\ 
-%5 & 39 & 60 \\
-%6 & 40 & 60 \\ 
-%\hline
-%Total lineups & 60 & 70 \\  
-%\hline
-%\end{tabular}
-%\end{center}
-%\label{tbl:pval_assumption} 
-%\end{table}
-
-%\begin{figure*}[hbtp]
-%   \centering
-%       \includegraphics[width=0.95\textwidth]{p_val_rank_counts12.pdf}
-%       \caption{Rank of p-values in a lineup of size 20 and the number of subjects choosing that plot in the lineup of experiment 1 and 2 with raw data (screening criteria 1).
-This indicates that most of the subjects pick the plot with minimum p-value (rank 1).
-There are subjects choosing the lineups with highest p-value(rank 20) but we can attribute this behavior to the type-I error.
-Can we?
-\green{Maybe don't need this plot}}
-%       \label{fig:P-val_rank}
-%\end{figure*}
-
-\subsection{How much do null plots affect the choice?}
-
- Visual inference falls into the same framework as randomization tests, where the statistics from the data are compared with those from null data. 
- But unlike these approaches visual inference is constrained to make the comparison with just a few draws $(m-1)$ from the null distribution.
-How this small set of null plots influences the subjects choice is important for understanding the reliability of visual inference.
-If the actual data plot is very different from the null plots, then the null plots should not have much influence on the choice.
-Measuring the difference, generally, between  plots is almost impossible.
-However,  in this controlled setting we can use $p$-values of the test statistic calculated on the data used in each plot as a proxy for similarity of structure between the plots.
-
-We have seen that the subjects tend to pick the plot in the lineup that has the lowest $p$-value (Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2}).
-What we are also interested in  is how this pick is affected by the distribution  of $p$-values of other plots in the lineup, particularly the $p$-value of the null plot with the strongest structure.
-If there is a null plot with a small $p$-value, or one close to that of the actual data plot, we would expect that subjects have a harder time detecting the actual data plot.
-Figure \ref{fig:pval_difference} investigates this.
-The difference between the $p$-value of the actual data is compared with the lowest from the null plots.
-This is plotted horizontally, and the proportion correct is plotted vertically.
-Negative values indicate lineups where the actual data plot had a smaller $p$-value than the minimum of the null plots.
-In experiment 1 (boxplots) there were a lot of lineups where the actual data plot had the smallest $p$-value, but only just.
-This caused quite some confusion for subjects, as seen because the variability in the proportion correct is huge for these lineups.
-Similarly large variability in correctness can be seen in the results of experiment 2 (scatterplots) except that the greater range of of differences in $p$-values shows the strength of subject's ability to pick the plot with most structure.
-Figures \ref{fig:P-val_log}, \ref{fig:P-val_log2} shed some more light on this story: when there is a big difference between the $p$-values (eg experiment 1, $\beta>7$) the subjects as one force chose the same plot.
-When there is less difference the distribution of counts is much more evenly spread between plots (eg experiment 1, $\beta=1$).
-
-%\blue{In the previous section we noticed that subjects tend to pick the plot with the lowest $p$-value.
-If a null plot has data that, by chance because the sample is a random draw from the null, has a smaller $p$-value than the actual data, we would expect that there is a strong possibility that an subject will pick this null plot instead of the actual data plot.
-Figure \ref{fig:pval_difference} investigates this for the data collected in experiments 1 and 2.
-The plots display the proportion of subjects who choose the actual data plot against the $p$-value of the actual data minus the lowest seen in the null data.
-A small negative value of the difference means that the $p$-value of the actual data is smaller than that of the null data.
-If there is a big negative difference we'd expect to see that the subjects would more often make the correct choice of plot.
-When there is a large positive difference, subjects fail to pick the correct plot.
-If however, there is little difference between the $p$-values then we'd expect to see more confusion in the choices.
-That is why we see a big range of proportion correct where difference is close to zero in Figure \ref{fig:pval_difference}.}
-
-%\green{Something is definitely not right!
-The code looks like it does actual data p-value - min(null $p$-value).
-But how can you get a difference of 0.8 from this??} \blue{It is possible theoretically!
-Suppose data plot $p$-value is larger than 0.8.
-Under null hypothesis, $p$-value $\sim U(0,1)$.
-Thus min(null $p$-value) could be close to zero by chance producing this large difference.
-And this has happened in both experiment 1 and 2.}
-
- %\green{Is this really pct correct, or 1-pct correct?
-\blue{This is proportion correct}.
-When there is a big diff in p-value almost no-one gets it right!
-Maybe it is just a scale issue?} \blue{In fact this is what we expect.
-If there is large difference that means data plot $p$-value is large while minimum $p$-value is very small.
-This should lead people to pick the wrong plot.
-Our minimum $p$-value assumption basically says this fact, right?}.
-We see that when there is a big difference between the $p$-values subjects choose the actual data plot more often.
-
-%Mahbub, could you have a look at the lineups, and compute the difference between the data plot's $p$-value and the  smallest $p$ value of a null plot, and plot that in a scatterplot against proportion correct?
-Thanks!
-
-In practice, the $p$-value is not going to be a valid way to compare plots.
-Rather metrics that can measure how graphical elements from one plot to another are perceived similarly are needed.
-This is investigated in \citet{niladri:2012}.
-Here, numerical measures of the similarity between plots are proposed to provide quality metrics for lineups.
-
-\begin{figure}[hbtp]
-   \centering
-       \scalebox{0.70}{\includegraphics{pval_difference.pdf}}
-       \caption{Investigating how the strength of structure in the null plots affects the choice of the actual data plot.
-Difference between the data plot's $p$-value and the  smallest $p$ value of a null plot vs proportion correct.
-Negative values indicate the $p$-value of the actual data plot are smaller than those of all of the null plots.
-Close to zero shows a wide range in the proportion correct, suggesting that when at least one null plot has structure almost as strong as the actual data plot, subjects had a difficult time in making their choice.}
-       \label{fig:pval_difference}
-\end{figure}
-
-\subsection{Type III error}~\label{sec:TypeIII}
-
-A little known error amongst statisticians is what was coined as Type III error in \citet{mosteller:48}.
-This occurs when the null hypothesis is correctly rejected but for the wrong reason.
-Experiment 3 is prone to this type of error.
-Observers were asked to identify the plot that had the largest slope, either positive or negative.
-But the actual data plot had a cluster of points, the contamination that forced the conventional test to fail to see any trend.
-For the human eye this cluster of points is  as visible as the association between the remaining points, leading the observer to detect the actual data plot by looking for the cluster instead of the slope.
-This would be considered to be a Type III error because it is not what was requested, but  still gives the correct answer.
-This is not a problem with the lineup protocol, because we want observers to react to anything that sets the the actual data plot apart.
-Here, in the controlled but simple setting, it is of interest to see what observers cued to in the actual data plot.
-
-Figure \ref{fig:power_loess_effect} indicates that Type III error might be occurring in experiment 3: correct identification of the actual data plot is not positively associated with effect size.
-Teasing this out of the results is possible by looking at the reasons observers gave for their choices.
-Observers were provided with four possible reasons to use for their choice:
-
-\begin{enumerate} \itemsep 0in
-\item Most different plot
-\item Visible trend
-\item Clustering visible
-\item Other
-\end{enumerate}
-
-\noindent with the possibility to use more than one.
-The task requested subjects to identify the plot that had the largest slope, which would correspond to choosing ``visible trend'' (2) as the reason for their choice.
-Reasons 1 or 3 would be indicative of Type III error.
-Figure \ref{fig:choice_reason} explores the reasons subjects gave for their choices.
-If there were no Type III errors committed, we would expect that people overwhelmingly using ``visible trend'' as their reason, or at least, when they use this reason they overwhelmingly correctly choose the actual data plot.
-This is not what we see.
-At left, are the reasons subjects gave for their choices --- 123 means that they gave all three reasons.
-The horizontal axis shows proportion of times that subjects correctly chose the actual data plot, and the reasons are sorted from most accurate to least accurate.
-The size of the point corresponds to the number of subjects putting this as the reason.
-Subjects that chose all three reasons almost always chose the actual data plot.
-This was followed by using 1 and 3, and then 1 and 4.
-The most common reasons given were reasons 1-3 individually, and the accuracy for these reasons ranged from 75\% for reason ``most different plot'' to 60\% for ``visible trend''.
-At right is a simplified view, containing just the four possible reasons -- if the subject chose one of these, regardless if they also chose another reason it is counted.
-``Visible trend'' comes in third.
-This is strong evidence that for many subjects even though they are correctly choosing the data plot, often they are cueing to other structure in the plot than the trend, making a Type III error.
-
-%It is interesting to calculate the proportion of observers who choose the correct plot, but use reasons 1 or 3, relative to reason 2.
-Figure \ref{fig:choice_reason} examines this.
-For observers who chose the actual data plot, the proportions for the three reasons are shown.
-It can be seen that the most common reason was ``most different plot'' (1), followed by clustering visible (3), with ``visible trend'' coming in third.
-\green{What about other??} \blue{Not many people chose other as we see in Figure \ref{fig:choice_reason}.}
-
-%This might be an indication that participants were not considering slope as distinguishing feature for the data, but the apparent clustering, i.e.
-the null hypothesis is being rejected, but not for the reason of the slope -- which the experimental setup was asking for --  leading to a type III error of correctly rejecting the null hypothesis for the wrong reason', as defined by \citet{mosteller:48}
-
-%\green{Not sure that this goes here, but something needs to be said about Type III.}
-
-%\blue{We also see this feature in figure \ref{fig:choice_reason}}.
-%\hh{Mahbub, it's not directly fig \ref{fig:choice_reason} - it's actually the lineup corresponding to the large outlier at (0.2,  0.6) in the left panel of fig \ref{fig:choice_reason}.}
-
-\begin{figure}[hbtp]
-   \centering
-%       \scalebox{0.60}{\includegraphics{choice_reason.pdf}}
-%       \scalebox{0.60}{\includegraphics{choice_reason_count.pdf}}
-       \includegraphics[width=0.45\textwidth]{choice_reason2.pdf}
-       \includegraphics[width=0.45\textwidth]{choice_reason3.pdf}
-       \caption{Exploring the reasons for plot choices for experiment 3, to examine the occurrence of Type III error.
-The vertical axis shows the reasons.
-At left, all subjects choices are shown, and reason 123 means all three reasons are used.
-At right, if the subject used a reason, regardless if they also used more than this reason, they are counted.
-The horizontal axis shows the proportion of times the subjects correctly chose the actual data plot.
-Size of the point corresponds to the number of subjects using that reason.}
-       \label{fig:choice_reason}
-\end{figure}
-
-For visual inference, making a Type III, is not actually a problem.
-It is only a possibility in this experiment because we are working with known structure.
-In the real setting, we are excited to see observers detecting the actual data plot, and curious about how they detect it, with all possible reasons encapsulated in the alternative hypothesis.
-This does show the importance of the qualitative answers that observers give for their choices.
-
-\section{Conclusions}
-
-%\begin{itemize} \itemsep 0in
-%\item Summarize the key pieces of the paper again, especially the major findings
-
-It is widely accepted that statistical graphics has important use in exploratory data analysis and model diagnostics.
-This paper has demonstrated that statistical graphics can be used in statistical inference and validates the lineup protocol proposed by \citet{buja:2009} as a formal framework to accomplish that.
-The simulation study showed that even in the situation where the conventional test has its assumptions met, visual inference is comparable.
-The results also suggest that when the underlying assumptions of a conventional test do not hold, visual inference will perform well.
-
-Power of visual tests can be estimated in different ways, which was explained in this paper.
-If it is assumed that observers tend to pick the plot with the strongest signal, which was consistent with the experimental data, the power of visual test can be obtained theoretically.
-Interestingly, the theoretical power of visual test is close to that of conventional test in the conditions tested in this paper.
-
-%\item Explain caveats, nuances, advantages with the work
-
-One of the critical parts in visual testing is appropriate selection of a statistic for the hypothesis of interest -- that means an appropriate choice of plot for the analytical purpose.
-A good test statistic is expected to reveal structure if the null hypothesis is not true.
-Small differences in plot choices such as color, orientation, and mapping of data values to graphical elements of the plot can make a difference.
-This is related to research in a grammar of graphics \citep{wilkinson:1999,hadley:2009}, which helps to ensure good practices.
-%This is the part of our ongoing and future research.
-
-The question that is posed to observers can affect the ability to evaluate lineups.
-In these experiments subjects were asked very specific questions, because the purpose was to compare with very specific conventional testing.
-In practice, the question posed will be more general to allow observers to choose plots with any remarkable features, which enables the data mining discovery process.
-
-% when evaluating the lineup: this question can be lined up with the alternative hypothesis of a classical testing scenario.
-If we are interested in the reason why participants picked a plot as the most different, we need to collect and evaluate this information in the study.
-This has to be done very carefully, as any suggestions for reasons also lead the observer  what patterns to look out for.}
-% By delivering an appropriate question may reduce the type III error in an appropriate scenario.}
-
-Even though a single observer can evaluate a lineup, as suggested in the initial lineup protocol, this work recommends having multiple observers to ensure a more reliable estimate of the $p$-value for the test, and more confidence in the decision.
-
-Experimental subjects in the simulation study came from various level of education, both genders and many different ages.
-Many subjects had little background in statistics, and were not necessarily aware of statistical plots.
-Yet, the results were good.
-We expect that visual inference could be a useful tool for high school students to learn about randomness as a primer to statistical thinking.
-A major advantage of visual statistical inference is that it is intuitive, as illustrated by the performance of subjects with limited education, and it  may make statistics fun to learn.
-
-The performance of subjects was quite varied, but consistent.
-No restrictions were placed on Turk workers, in terms of abilities.
-There were clearly some subjects who performed very badly, but it was very interesting to see that there were some super-observers, people who detected the actual data plot at a rate better than that of the power of the best conventional test.
- It would be interesting to see how well trained subjects might perform.
-Prior to the Turk experiments, we conducted pilot studies using local graphics experts and obtained good results, indicating that training in data visualization might be helpful for visual inference.
-Future work might explore this.
-
-%\item Describe the next steps of the work, and possible ideas for future extensions
-
-When there is no conventional test available, visual testing procedures can be used due to its few assumptions and non-parametric nature.
-We have successfully used the procedures informally for local projects.
-It is hoped that this proves to be extremely valuable in data mining applications, and exploratory analyses, where there are no existing gauges of statistical significance.
-The lineup protocol operates similarly to general tests, such as two-sided, analysis of variance $F$-tests, and Wilks $\Lambda$ for multivariate analysis of variance.
-If the null hypothesis is rejected, it is generally that we can say that ``there is something there'' but not specifically what it is in the data that triggers the rejection.
-Follow-up questions on the reasons provide qualitative insight.
-Multiple comparisons are often done to refine and understand the results of the general conventional testing, and perhaps some similar approaches might be developed for visual inference.
-
-%Such a scenario could be the inference with small sample size with large number of covariates, data mining findings or even in genetic research where many covariates restrict the available inference very conservative for simultaneous testing (required $p$-value is very small to be able to reject $H_0$).
-The direction of future research would be to study that and apply the methods in those areas.
-
-%The purpose of this paper has been to examine the effectiveness of visual inference methods in direct comparison to existing inference methods.
-We need to be clear that this is not the purpose of visual inference generally: visual methods should not be seen as competitors to conventional inference.
-The purpose here, is to  establish properties and  efficacy of visual testing procedures in order to use them in situations where conventional tests cannot be used.
-For this experiment the effect of $\beta_2$ was examined using side-by-side boxplots.
-Future experiments will be conducted to compare other regression parameters as described in Table \ref{tbl:stat_multiple} and assess sensitivity of power to modeling conditions.
-
-%\green{Mention the possibility of a visual skills test, to select the workers that are better qualified, to implement the lineup protocol in practice.
-Along the lines of the Army tests that Susan brought to the working group.
-\blue{While recruiting subjects from the Amazon Mechanical Turk web site we did not put any restriction on the skill levels of the subjects.
-Thus anyone from any background could participate the experiment.
-It would be interesting to see how well the trained subjects performs.
-Many pilot studies with advanced knowledge of statistical graphics as well as statistics indicates that the performance is very good for trained subjects.
-One of our future work would be to examine this in details.
-For this we are planning to trained Turk subjects with specific graphics and educate them what pattern a specific graphics may show.
-After being trained they may have to pass the qualification test before participating the experiment.
-Even though we did not do this, it is very common to Turk subjects to go through this process for various other tasks they do in the Turk web site.}.
-
-%Discuss the purpose of randomly generating data null plots, rather than selectively sampling from the null distribution.}
-
-%In practice there is no correct answer.
-In these experiments we controlled the effects that observers saw, so there was a correct choice in each case.
-In the real setting, there is not correct choice, just whether the actual data plot stands apart from the rest, so that the observer can find it.
-And subjects would provide open-ended answers for their choice of plot.
-
-%What are your suggestions for future experiments, and next steps of the research?
-
-%\paragraph{Acknowledgement:}
-%This work was funded in part by National Science Foundation grant DMS 1007697.
-
-%\bibliographystyle{plain}
-%\bibliographystyle{plainnat}
-\bibliographystyle{asa}
-%\bibliographystyle{ieeetr}
-\bibliography{references}
-
-%\end{multicols}
-
-\end{document}  %End of document.
-
-\section*{Appendix}
-
-\begin{table}[hbtp]
-\caption{Selection of 10 lineups for a person to evaluate in the simulation experiment with discrete covariate described in section \ref{sec:category}.} 
-\centering
-\begin{tabular}{c c c c  c c c}
-\hline\hline
-Difficulty& \multicolumn{3}{c}{parameter combination}& Number of evaluations &Total number  & number of lineups\\
-level & $n$ & $\sigma$ & $\beta$ &required($n_{\gamma}$) & of lineups & randomly shown \\
-\hline
-easy&100& 5&8 & 1& 12 & 1\\
-&100&12&16 &1&&\\
-&300& 5&5 &1&&\\
-&300&12&10 &1 &&\\
-\hline
-medium&100& 5&3 &203 & 9 &2\\
-&300& 5&2,3 & 97, 1&&\\
-\hline
-hard&100&12&3,8,10 & 277, 126, 23& 18 &6\\
-&300& 5&1 & 371 &&\\
-&300&12&3,5& 375,74 &&\\
-\hline
-mixed&100& 5&1,5,0& 214, 2,73 & 21 &1\\
-&100&12&1& 100& &\\
-&300& 5&0 & 73&&\\
-&300&12&7,1& 2, 152&&\\
-\hline
-Total &&&&&60&10\\
-\hline
-\end{tabular}
-\label{tbl:dist_lineup1}
-\end{table}
-
-%\begin {abstract} 
-%Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis.
-Until recently, there were no formal visual methods in place for determining statistical significance of findings.
-This changed when Buja et al.
-(2009) conceptually introduced the lineup protocol for formal tests of visual findings.
-In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models.
-A human subjects experiment is conducted using simulated data to provide controlled conditions.
-Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
-%\end {abstract}
-
-\end{document}

### 2012-09-07 R/sociology_paper.R
Learning trend plot added

@@ -1,11 +1,11 @@
-# The complete R code for JASA Paper
+# The complete R code for Sociology Paper
 
 library(ggplot2)
 library(plyr)
 library(reshape)
-library(xtable)
 library(lme4)
-library(stringr)
+#library(lubridate)
+
 
 # getting the data and some common functions
 
@@ -37,1068 +37,156 @@ clean_data <- function(raw_dat){
   return(subset(cleaned_dat,indx_dup))
 }
 
-indx1 <- c(3,6,7,8,9,17,18,20)
-indx2 <- c(4,7,8,9,10,17,18,20)
-
-dat1 <- clean_data(raw.dat1)[,-indx1]
-dat2 <- clean_data(raw.dat2)[,-indx1]
-dat3 <- clean_data(raw.dat3)[,-indx2]
+dat1 <- clean_data(raw.dat1)
+dat2 <- clean_data(raw.dat2)
+dat3 <- clean_data(raw.dat3)
 
 
-# saving cleaned data as suplementary materials
-write.csv(dat1, file='../submission_initial/suplementary/data/data_turk1.csv',row.names=F)
-write.csv(dat2, file='../submission_initial/suplementary/data/data_turk2.csv',row.names=F)
-write.csv(dat3, file='../submission_initial/suplementary/data/data_turk3.csv',row.names=F)
+# Examining learning trend while giving feedback
 
-
-
-
-# =================== mathematical test t distribution  ===================
-
-t <- seq(-4, 4, by = 0.1)
-density_t <- dt(t, df = 15)
-qplot(t, density_t, geom = "line") + geom_vline(xintercept = 2.5) + ylab("Density")
-ggsave("../images/stat_mathematical_test.pdf",height=2,width=2.2)
-
-
-# ===== computation of visual p-values for diffrent combination of K and x 
-
-get_pval <- function(m,K,x){
-	p <- pbinom(size=K,prob=1/m,q=x)
-	px <- dbinom(size=K,prob=1/m,x=x)
-	pval <- 1-p + px
-	return(pval)
+get_trend <- function(dat){
+  return(
+    ddply(dat,.(id), summarize,
+          attempt = 1:length(start_time),
+          pic_id = pic_id[order(start_time)],
[…]

### 2012-09-07 sociology_paper.tex
Learning trend plot added
model parameter estimates added

@@
 \subsection{Overview of the Data}
-\subsection{Learning Trend of the Observer}
+\subsection{Learning Trend of the Observer} Since each subject is shown multiple lineups, they have a chance for self learning and do better on lineups shown later in the sequence.
+We examine this by fitting generalized mixed effect model presented in \cite{majumder:2012} with covariates attempt number and $p$-value of the lineup.
+The $p$-value of the lineup measures some sort of difficulty and including this should adjust for the difficulty level of the lineup while estimating the probability of correct responses in each attempt.
+Figure \ref{fig:learning_trend} shows the fitted subject wise learning pattern as well as overall learning pattern for each of the experiments.
+The parameter estimates od the model are shown in Table \ref{tbl:model_par}.
+
+%\begin{table}[hbtp]
+%\caption{Parameter estimates of mixed model. Estimates are highly significant with $p$-value $<$  0.0001 for all three experiment data.}
+%\begin{center}
+%\begin{tabular}{cr@{.}lcc}
+%  \hline \hline
+% &  \multicolumn{3}{c} {Fixed effect}  & Random effect\\
+% \cline{2-4}
+% Experiment & \multicolumn{2}{l}{Estimate}  &Std. error & Variance\\
+%  \hline
+%  1 & 0&39 & 0.0094 & 0.0080 \\ 
+%  2 & 1&21 &  0.0197 &  0.0443 \\ 
+%  3 & 0&59 (Intercept)  &   0.1668 & 1.9917\\ 
+%     & 0&21 (Slope)    &  0.0511     &  0.0245\\ 
+%     &-0&78 (correlation) & & \\
+%   \hline
+%\end{tabular}
+%\end{center}
+%\label{tbl:model_par}
+%\end{table}
+
+% latex table generated in R 2.15.0 by xtable 1.7-0 package
+% Fri Sep  7 10:28:47 2012
+\begin{table}[hbtp]
+\caption{Fixed effect parameter estimates of mixed model.
+Note that attempt is not significant for experiment 2.
+The continuous covariate lineup difficulty was measured by the $p$-value of the actual plot in the lineup.}
+\begin{center}
+\begin{tabular}{llrrrrl}
+  \hline
+& Parameters & Estimate & Std..Error & z.value & P-value  & \\ 
+  \hline
+\multicolumn{2}{l}{\bf{Experiment 1} } &&&& &\\
+&(Intercept) & -0.30 & 0.10 & -2.89 & 0.00 & ***\\ 
+ & attempt & 0.08 & 0.02 & 4.21 & 0.00 & ***\\ 
+ & lineup difficulty & -11.32 & 1.01 & -11.19 & 0.00 & ***\\ 
+\multicolumn{2}{l}{\bf{Experiment 2} } &&&&& \\
+ & (Intercept) & 2.36 & 0.15 & 16.09 & 0.00 & ***\\ 
+ & attempt & 0.01 & 0.03 & 0.35 & 0.72 \\ 
+ & lineup difficulty & -55.03 & 2.20 & -25.00 & 0.00 & ***\\ 
+\multicolumn{2}{l}{\bf{Experiment 3} } &&&& \\
+ & (Intercept) & 0.25 & 0.16 & 1.58 & 0.11 & .\\ 
+ & attempt & 0.10 & 0.03 & 3.34 & 0.00 & ***\\ 
+ & lineup difficulty & 3.00 & 0.56 & 5.36 & 0.00 & ***\\ 
+   \hline
+\end{tabular}
+\end{center}
+\label{tbl:model_par}
+\end{table}
+
+% \footnote{Signif. codes: 0 Ô***Õ 0.001 Ô**Õ 0.01 Ô*Õ 0.05 Ô.Õ 0.1}
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=6in]{learning_trend.pdf} 
+   \caption{Probability of correct response for different attempts by each subjects for a plot with $p$-value=0.05 .
+The overall probability increases with attempts indicating a learning trend of the observers for experiments 1 and 3.
+For experiment 2 the overall trend trend is not statistically significant.
+Subject to subject variability is seen in learning pattern.}
+   \label{fig:learning_trend}
+\end{figure}
 

### 2012-09-07 R/sociology_paper.R

@@ -4,6 +4,7 @@ library(ggplot2)
 library(plyr)
 library(reshape)
 library(lme4)
+library(xtable)
 #library(lubridate)
 
 
@@ -144,10 +145,15 @@ fit_model <- function(dat){
   return(summary(fit))
 }
 
-fit_model(subset(dtrend, Experiment =="1"))
-fit_model(subset(dtrend, Experiment =="2"))
-fit_model(subset(dtrend, Experiment =="3"))
+f1 <- fit_model(dtrend1)
+f2 <- fit_model(dtrend2)
+f3 <- fit_model(dtrend3)
 
+model.out <- rbind(round(f1@coefs,2), round(f2@coefs,2), round(f3@coefs,2))
+parameters <- rownames(model.out)
+print(xtable(data.frame(parameters,model.out)), include.rownames=FALSE)
+
+print(xtable(pval_sm), include.rownames=FALSE)
 
 pred.mixed <- function(X, subject=0,fit) {
   eta <- fixef(fit)[1] + X * fixef(fit)[2] + 0.05 * fixef(fit)[3] + subject*X

### 2012-09-07 sociology_paper.tex
abstract updated
table caption updated

@@
 The simulation experiments suggest that individual skill as well as demographics are very significant for the power of visual inference.
+Moreover, the data shows the evidence of learning or getting skilled while sequentially evaluating multiple lineups.
 
@@
    \includegraphics[width=6in]{learning_trend.pdf} 
-   \caption{Probability of correct response for different attempts by each subjects for a plot with $p$-value=0.05 .
+   \caption{Probability of correct response for different attempts by each subjects for a plot with lineup difficulty ($p$-value of actual plot) = 0.05 .
 The overall probability increases with attempts indicating a learning trend of the observers for experiments 1 and 3.

### 2012-09-07 sociology_paper.tex
outline updated

@@
 May present some results from \textit{Distance measure} of visual test statistics \citep{niladri:2012}.
-\subsection{Question that Human Observer Faces}
+\subsection{Question that Human Observer Faces} Discuss the importance of the question the observer needs to answer while evaluating the lineup.
 \subsection{Observer Personality}
 \subsection{Visual Perception of Human Eye} Discuss the results we have on eye tracking experiment \citep{zhao:2012}.
-\subsection{Demographics of Observer}
-\subsection{Skills of Observer}
+\subsection{Demographics of Observer}  Age, gender and location of the observer.
+\subsection{Skills of Observer} Knowledge about statistical plot as well as education level.
 

### 2012-09-07 sociology_paper.tex
ouline description added

@@
 \RequirePackage{natbib}
-\usepackage{amsmath, amssymb, fullpage, amsthm, array,,graphicx,asa}
+\usepackage{amsmath, amssymb, fullpage, amsthm, array,,graphicx,asa, url}
 %\usepackage[dvips]{graphics}
@@
 
-%\usepackage{setspace, url}
+%\usepackage{setspace}
 %\doublespacing
@@
 \subsection{Choice of Visual Test Statistic} Discuss the effective feature of visual test statistics \citep{heike:2012}.
-May present some results from \textit{Distance measure} of visual test statistics \citep{niladri:2012}.
+Present some related results from \textit{Distance measure} of visual test statistics \citep{niladri:2012}.
 \subsection{Question that Human Observer Faces} Discuss the importance of the question the observer needs to answer while evaluating the lineup.
@@
 
-\subsection{Experiment Setup}
-\subsection{Diversity of the Experimental Subjects}
+\subsection{Experiment Setup} Description of the simulation experiment giving focus on demographics and skills of experimental subjects.
+\subsection{Diversity of the Experimental Subjects} Discuss why it would be important and how we accomplish this using Amazon Mechanical Turk \cite{turk} web site.
 
@@
 \begin{table}[hbtp]
-\caption{Fixed effect parameter estimates of mixed model.
+\caption{Fixed effect parameter estimates of generalized mixed model.
 Note that attempt is not significant for experiment 2.
-The continuous covariate lineup difficulty was measured by the $p$-value of the actual plot in the lineup.}
+The continuous covariate, lineup difficulty, was measured by the $p$-value of the actual plot in the lineup.}
 \begin{center}

### 2013-01-17 R/sociology_paper.R
update

@@ -43,24 +43,71 @@ dat2 <- clean_data(raw.dat2)
 dat3 <- clean_data(raw.dat3)
 
 
+# Examining performance based on plot location in lineup
+
+dat <- dat2
+
+loc.performance <- ddply(dat, .(plot_location),summarize,
+                         per_correct=mean(response)*100,
+                         n=length(response),
+                         p_value = mean(p_value))
+
+model <- as.formula(response ~ factor(plot_location) + (plot_location|id) + (1|pic_id))
+model <- as.formula(response ~ p_value + factor(plot_location) + (plot_location|id) )
+fit <- lmer(model,family="binomial",data=dat)
+summary(fit)
+
+X <- diag(rep(1,(length(unique(dat$plot_location))-1)))
+eta0 <- fixef(fit)[1]+ 0.01* fixef(fit)[2]
+eta <- c(eta0, eta0 + X %*% fixef(fit)[-(1:2)]) 
+loc.performance$prob_correct <- round(exp(eta)/(1+exp(eta)), 5)
+
+no_loc <- (1:20) [!(1:20 %in% loc.performance$plot_location)]
+loc.dat <- rbind(loc.performance, 
+                 cbind(plot_location=no_loc,per_correct=0,
+                       n=0,p_value=0,prob_correct=0))
+
+ggplot(loc.dat, aes(1,prob_correct))+
+  geom_bar(stat="identity", fill="darkgreen")+
+  facet_wrap(~plot_location, ncol=5) + 
+  ylab("Probability of correct evaluation") +
+  xlab("True plot location in the lineup") +
+  theme(axis.text.x = element_blank())
+
+ggsave("../images/location_effect2.pdf")
+
+
+qplot(p_value, per_correct, data = loc.performance, geom="point")
+
+qplot(plot_location,per_correct, geom="point", data = loc.performance)+
+  facet_wrap(~plot_location, ncol=5)
+
+qplot(x=1,y=per_correct, geom="bar", binwidth=1, data = loc.performance)+
+  facet_wrap(~plot_location, ncol=5)
+
+qplot(1,1, geom="bar", binwidth=1,data = loc.performance, fill=per_correct)+
+  facet_wrap(~plot_location, ncol=5)+ scale_x_continuous()+ scale_y_continuous()
+  + opts(axis.ticks=theme_blank())
+
+
+
+
 # Examining learning trend while giving feedback
 
[…]

### 2013-02-26 sociology_paper.tex
update
discussion of visual test statistic

@@
 
+\tableofcontents
+
 % Article top matter
@@
 
-\subsection{Choice of Visual Test Statistic} Discuss the effective feature of visual test statistics \citep{heike:2012}.
-Present some related results from \textit{Distance measure} of visual test statistics \citep{niladri:2012}.
+\subsection{Choice of Visual Test Statistic} Visual test statistics as defined in \citep{majumder:2012} serve two main purposes.
+One is to display a prominent pattern that should direct the human observers to select the actual plot in the lineup when the null hypothesis is not true.
+The another feature is that it should not show any strange pattern when null hypothesis is true, i.e., it should show similar pattern that null plots may have.
+Thus a visual test statistic is highly associated with the hypothesis.
+To achieve these purposes it is very important to decide which plot type and plot features should be adopted.
+As discussed in \citep{majumder:2012} following grammar of graphics developed  by \citet{wilkinson:1999} and later improved by \citet{hadley:2009} can give the task a standard control.
+
+Some of the effective features of visual test statistics are discussed in \citep{heike:2012} including plot type, color and shape of the plots.
+It is also observed that a scatterplot may do a better job than a box plot when using as a visual test statistic for regression parameters \citep{majumder:2012} .
+\citep{niladri:2012} presents some \textit{distance measure} to determine how a plot may be different from another.
+
+Plots to do: for the same signal strength show the proportion correct for scatterplot and boxplot. eg., qplot(effect, ump-visual power, color= plot-type)
+
 \subsection{Question that Human Observer Faces} Discuss the importance of the question the observer needs to answer while evaluating the lineup.
@@
 \subsection{Experiment Setup} Description of the simulation experiment giving focus on demographics and skills of experimental subjects.
-\subsection{Diversity of the Experimental Subjects} Discuss why it would be important and how we accomplish this using Amazon Mechanical Turk \cite{turk} web site.
+\subsection{Estimation of sample size}  How the sample size is estimated
+\subsection{Data collection methods}  Description of the simulation experiment giving focus on demographics and skills of experimental subjects.
 
@@
 \subsection{Overview of the Data}
+\subsection{Diversity of the Experimental Subjects} Amazon Mechanical Turk \cite{turk} web site is a good source for recruiting people to evaluation lineups.
+It is a source of diverse workers who can give us feedback on visual inspection.
+This is the kind of task where human are better than computers.
+Figure \ref{} shows the location from where we received data.
+Notice that we received data from around the world and we observed diversity of location in the collected data.
+We notice diversity in not only location of the observers but also their gender, age groups and education levels.
+Table \ref{} shows that we have data from almost similar size of male and female population.
+
 \subsection{Learning Trend of the Observer} Since each subject is shown multiple lineups, they have a chance for self learning and do better on lineups shown later in the sequence.

### 2013-02-26 sociology_paper.tex
Lineup question section updated

@@
 
-\subsection{Question that Human Observer Faces} Discuss the importance of the question the observer needs to answer while evaluating the lineup.
+\subsection{Question that Human Observer Faces} When a observer is prepared to evaluate a lineup, there needs to be a guideline what exactly should be looked for in a lineup.
+The researcher knows about the hypothesis but the observer does not know the underlying details of the lineup.
+Thus the researcher needs to ask a question observer needs to answer while evaluating the lineup.
+This question should provide the observer a little clue so that the answer reflects the hypothesized patterns in the actual plot.
+For example Table \ref{tbl:visual_stat} shows the questions asked for the simulation experiments done by  \citet{majumder:2012}.
+Notice that for case 1 if the observer can identify the actual plot in the lineup that should indicate that the plot chosen has the most vertical difference between groups A and B which is exactly what the researchers intend to examine.
+Similarly for case 2, a correct evaluation would indicate that the slope is different than the slope that may show up just from randomness.
+
+\begin{table*}[hbtp] 
+\caption{Visual test statistics and question asked to the observers to answer while evaluating a lineup } 
+\centering 
+\begin{tabular}{m{0.5cm}m{2cm}m{3cm}m{7.5cm}} 
+\hline\hline 
+Case & Statistic & Test Statistic & Lineup question \\ [0.5ex] % inserts table %heading 
+\hline 
+1  & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{minipage} & Which set of box plots shows biggest vertical difference 
+between group A and B? \\
+2 &  Scatter plot & \begin{minipage}[t]{3cm}   \scalebox{0.45}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Of the scatter plots below which one shows data that has steepest slope? \\
+
+\hline 
+\end{tabular} 
+\label{tbl:visual_stat} 
+\end{table*}
+
+These questions are very crucial for the power of visual test.
+They help observer think in a controlled direction.
+Notice that there may be unnecessary patterns in the actual data plot which may not necessarily indicate the existence of the significant signal in the plot.
+These question help observer not to be misguided by those patterns.
+To review further on this \citet{majumder:2012} have also asked why the observers choose a specific plot.
+
 \subsection{Observer Personality}
@@
 
-\section{Simulation Experiment}
+\section{Simulation Experiment for Learning Effect}
 

### 2013-02-28 sociology_paper.tex
visual perception of eye updated

@@
 
-\subsection{Question that Human Observer Faces} When a observer is prepared to evaluate a lineup, there needs to be a guideline what exactly should be looked for in a lineup.
+\subsection{Question that Human Observer Answers} When a observer is prepared to evaluate a lineup, there needs to be a guideline what exactly should be looked for in a lineup.
 The researcher knows about the hypothesis but the observer does not know the underlying details of the lineup.
@@
 
-\subsection{Observer Personality}
-\subsection{Visual Perception of Human Eye} Discuss the results we have on eye tracking experiment \citep{zhao:2012}.
+\subsection{Observer Personality} Each person is different from others in some way.
+For example, in a controlled experimental study \citet{zhao:2012}  notice that some people spend a lot of time to decide no matter whether the lineup is difficult or easy while some simply glance at lineups to make a decision.
+This influences the response of the observer and we see the subject specific variation in the power of visual test \citep{majumder:2012}.
+
+\subsection{Visual Perception of Human Eye} Different people observe the lineup in different ways.
+With the help of eye-tracking equipment \citet{zhao:2012} track the observers' eyes to see how they go through the plots in a lineup to come to their answers.
+The result suggest that people have particular methods of reading lineups.
+Some people read lineups from left to write direction while some read from upward to downward.
+Some people start looking from the center of the lineup while others start from the top left corner.
+In the earlier phase of the exercise, the observer tend to scan the plots and then start comparing plots to make a final decision.
+Beside right-left or up-down directions observer show some diagonal movement too.
+
+Given this pattern of human eye tracks, it may effect the performance of an observer depending on where the actual plot is placed in the lineup.
+If the actual plot is on the top left corner and the observer starts from that point it may be easier to detect the actual plot earlier in the exercise and the observer could get plenty of time to make comparison.
+Thus the position of the actual plot in a lineup has some impact.
+
 \subsection{Demographics of Observer}  Age, gender and location of the observer.

### 2013-02-28 R/sociology_paper.R

@@ -43,6 +43,12 @@ dat2 <- clean_data(raw.dat2)
 dat3 <- clean_data(raw.dat3)
 
 
+# Summary statistics of the data
+
+with(dat3, table(age))
+
+
+
 # Examining performance based on plot location in lineup
 
 dat <- dat2
@@ -265,5 +271,30 @@ pic_dat[1,c("pic_id","pic_name","plot_location","p_value")]
 xtabs(data=pic_dat, ~choice_reason+response_no)
 
 
+# ---------------------  description of coded variables -----------
+
+## gender 1 = male 
+#         2 = female
+## age 1 = below 18 
+#      2 = 18-25 
+#      3 = 26-30 
+#      4 = 31-35 
+#      5 = 36-40
+#      6 = 41-45
+#      7 = 46-50
+#      8 = 51-55
+#      9 = 56-60
+#     10 = above 60
+## academic_study 1 = High school or less
+#                 2 = Some under graduate courses 
+#                 3 = Under graduate degree
+#                 4 = Some graduate courses
+#                 5 = Graduate degree
+## conf_level 1 = most, 5 = least
+## choice_reason 1 = Big vertical difference 
+#                2 = Medians are different
+#                3 = Outliers
+#                4 = others
+## unit(time_taken) = second 
 
 

### 2013-05-21 sociology_paper.tex
location effect discussion added

@@
 \subsection{Data collection methods}  Description of the simulation experiment giving focus on demographics and skills of experimental subjects.
+\subsection{Amazon Mechanical Turk}
+
+Amazon Mechanical Turk \cite{turk} is an online work place where people from around the world can perform simple task to get paid.
+Usually tasks are very simple and no specialized training is required.
+Being a human is the main requirement.
+The tasks are designed such that it does not take much time to complete.
+Humans are still better than computers in performing these types of tasks.
+The the amount of money paid for each task is very small as well.
+Figure \ref{fig:amazon_task} displays an example of such a task.
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=5in]{amazon_task.pdf} 
+   \caption{An example of amazon mechanical turk task.
+Tasks are usually very simple and human are still better than computer in performing these types of tasks.
+With each task, simple instructions are given for workers to follow.
+The workers first accept the task before submitting their response.}
+   \label{fig:amazon_task}
+\end{figure}
 
@@
 Figure \ref{fig:learning_trend} shows the fitted subject wise learning pattern as well as overall learning pattern for each of the experiments.
-The parameter estimates od the model are shown in Table \ref{tbl:model_par}.
+The parameter estimates of the model are shown in Table \ref{tbl:model_par}.
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=4in]{learning_trend.pdf} 
+   \caption{Mean absolute residuals vs attempt obtained by adjusting for lineup and subject specific variation.
+The trend over attempt does not show any learning trend since the slopes are not statistically significant.}
+   \label{fig:learning_trend}
+\end{figure}
+
+\subsection{Location Effect of Data Plot in the Lineup}
+
+We set up a Turk experiment to examine whether there is any difference of performance based on the location of actual data plot in the lineup.
+For this five locations of a lineup were randomly chosen to put the same actual data plot.
+The locations are 2,9,12,16,20 for Interaction effect and 1,8,12,17,20 for Genotype effect.
+For a lineup with specific data plot location, five sets of null plots are used giving 5 lineups for each location.
+In total we have 25 lineups for Interaction effect and 25 lineups for Genotype effect.
+These 50 lineups are then evaluated by 100 people recruited from Amazon Mechanical Turk.
+Each person evaluated 3 lineups one from Interaction, one from Genotype and the third one was a test lineup.
+The test lineup is used to process the payment and examine the quality of the data and the responses of the test plot is not added to our analysis.
+
+The proportion of correct responses by the turk observers are shown in Figure \ref{fig:location_effect}.
+Notice that the data plot is same for each location but we see some variability of performance based on different null plots.
+This may happen if some null plots appear to be more similar to the actual plot while the others are not making some lineups harder than others even though the actual data plot is same.
+This pattern is evident in the Figure \ref{fig:location_effect} as we see proportion correct for null plot 5 is consistently above the null plot 1 for each of the locations.
+But this variability could be controlled at some extent by collecting more data for each location.
+Since the variability is larger when we have small number of responses as we see in the figure.
+For example, we have only one response for null plot 3 and 5 for Interaction effect which should lead to the extreme proportion of correct since one response would either be correct or wrong.
+
+\begin{figure}[htbp] 
+   \centering
+    \includegraphics[width=6.5in]{proportion_nulls_guide.pdf} 
+   \caption{Location of data plot in the lineup and proportion correct for both Interaction and Genotype effect.
+Each colored line represents a null set and the size of the dots represents number of responses.
+The overall average proportions are shown by dashed line.
+The actual data plot locations are shaded grey on the top panels to demonstrate their relative positions on a lineup.}
+   \label{fig:location_effect}
+\end{figure}
+
+It is also interesting to check whether the proportion correct differs if the actual plot is on the outer boundary of the lineup or inside the outer boundary.
+As we see in Figure \ref{fig:location_effect}, location 9, 12 are inside for Interaction effect and location 8, 12 are inside for Genotype.
+It does not appear to have any difference whether the actual plot is inside or outer border of the lineup.
+
+The overall proportion of correct response does not seem to vary much for different location indicating the non-existence of location effect.
+Since the actual data plot is same for each of the null sets of plot, the response data constitutes a multivariate response.
+Thus, to examine if the difference in proportion correct among the locations is statistically significant we fit a one way multi variate analysis of variance (MANOVA) model to the data.
+
+ Suppose $Y=(Y_1,Y_2, ...
+, Y_p)$ is a vector of random variable with dimension $p=5$ representing the response for five null sets.
+let $Y_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., 5$.
+We fit the following MANOVA model
+\begin{equation}\label{manova}
+Y_{ij} = \mu_{i} + \epsilon_{ij}
+\end{equation}
+where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$.
+
+Model \eqref{manova} enables to test if the mean vectors are similar for different locations.
+To fit the model we use anova function of stats package of \cite{R:2012}.
+The results are shown in Table \ref{tbl:manova}.
+The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05 and we failed to reject the hypothesis that there is no difference in location.
+
+\begin{table}[hbtp]
+\caption{The results obtained by fitting MANOVA Model \eqref{manova}.}
+\begin{center}
+\begin{tabular}{ccccccccc}
+  \hline \hline
+ Location & && & \multicolumn{3}{c} {Degrees of Freedom}  & F test \\
+ \cline{5-7}
+ Effect & DF & Pillai & Approx. F & Numerator& Denumerator &Residual & p value\\
+  \hline
+  Interaction& 3&1.4783 &0.7772&15&12 & 6 & 0.6821 \\ 
+  Genotype &4&1.7796 &1.1221&20&28 & 8 &  0.3824 \\ 
+   \hline
+\end{tabular}
+\end{center}
+\label{tbl:manova}
+\end{table}
+
+We also fit Model \eqref{manova} with only two locations based on outer or inner plot location.
+By outer location we mean the data plot location which is on the border area of the lineup.
+For example inner locations of a lineup would be 7,8,9,12,13,14.
+It turns out that inner and outer locations are also not significantly different.
 
@@
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=6in]{learning_trend.pdf} 
-   \caption{Probability of correct response for different attempts by each subjects for a plot with lineup difficulty ($p$-value of actual plot) = 0.05 .
+%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+%   \centering
+%   \includegraphics[width=4in]{learning_trend.pdf} 
+%   \caption{Probability of correct response for different attempts by each subjects for a plot with lineup difficulty ($p$-value of actual plot) = 0.05 .
 The overall probability increases with attempts indicating a learning trend of the observers for experiments 1 and 3.
-For experiment 2 the overall trend trend is not statistically significant.
+For experiment 2 the overall trend is not statistically significant.
 Subject to subject variability is seen in learning pattern.}
-   \label{fig:learning_trend}
-\end{figure}
+%   \label{fig:learning_trend}
+%\end{figure}
 

### 2013-05-21 R/sociology_paper.R

@@ -5,6 +5,7 @@ library(plyr)
 library(reshape)
 library(lme4)
 library(xtable)
+library(grid)
 #library(lubridate)
 
 
@@ -13,6 +14,10 @@ library(xtable)
 raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
 raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
 raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
+raw.dat4 <- read.csv("../data/raw_data_turk4.csv")
+raw.dat5 <- read.csv("../data/raw_data_turk5.csv")
+raw.dat6 <- read.csv("../data/raw_data_turk6.csv")
+raw.dat7 <- read.csv("../data/raw_data_turk7.csv")
 
 source("calculate_ump_power.R") # functions to compute power
 
@@ -38,14 +43,20 @@ clean_data <- function(raw_dat){
   return(subset(cleaned_dat,indx_dup))
 }
 
+remove_duplicated <- function(dat){
+  indx_dup <- with(dat, !duplicated(data.frame(id,pic_id))) #duplication index
+  # now returning cleaned data removing duplication
+  return(subset(dat,indx_dup))
+}
+
 dat1 <- clean_data(raw.dat1)
 dat2 <- clean_data(raw.dat2)
 dat3 <- clean_data(raw.dat3)
-dat4 <- read.csv("../data/raw_data_turk4.csv")
+dat4 <- remove_duplicated(raw.dat4)
 dat4$p_value=0
-dat5 <- read.csv("../data/raw_data_turk5.csv")
-dat6 <- read.csv("../data/raw_data_turk6.csv")
-dat7 <- read.csv("../data/raw_data_turk7.csv")
+dat5 <- remove_duplicated(raw.dat5)
+dat6 <- remove_duplicated(raw.dat6)
+dat7 <- remove_duplicated(raw.dat7)
 
 
 # Summary statistics of the data
@@ -54,7 +65,7 @@ with(dat3, table(age))
 
 
 
-# Examining performance based on plot location in lineup
+# Examining performance based on plot location effect in lineup
 
 dat <- dat2
 
@@ -170,9 +181,9 @@ dtrend <- rbind(data.frame(dtrend1, Experiment = "1"),
 qplot(data=dtrend, attempt, response, geom="point", colour=Experiment) +
   stat_smooth(method="loess")
 
[…]

### 2013-05-22 sociology_paper.tex
choice reason wordle plot added

@@
 This is the kind of task where human are better than computers.
-Figure \ref{} shows the location from where we received data.
+Figure \ref{fig:turker_location} shows the location from where we received data.
 Notice that we received data from around the world and we observed diversity of location in the collected data.
@@
 Table \ref{} shows that we have data from almost similar size of male and female population.
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=4in]{turker_country.pdf} 
+   \includegraphics[width=4in]{turker_location.pdf}    
+   \caption{Words used to explain the reasons for selection of data plot in a lineup. Larger font indicate more people choosing that word. Different color is used just to separate the words.}
+   \label{fig:turker_location}
+\end{figure}
+
+\subsection{How people pick the data plot} In the experiment setup we have a question for the turk observers to choose the reason for their selection of a particular data plot.
+It is shown by \cite{majumder:2012} that these choice reasons reveal the way people picks the data plot in the lineup.
+To investigate this further we added free text input option for their choice reason instead of some fixed reasons to select from.
+In this experiment people could write whatever they think their reasons for choice are.
+Figure \ref{fig:wordle} shows the words used to explain their reasons for choice.
+The most common words used to explain their choice are points and green which indicates two important feature of a plot.
+One is the indicator of plot aesthetics and the other is the color.
+Spread, steepest, line and apart are some other important words used frequently.
+Spread and apart are indicative of variability in the data.
+Steepest and line indicates some sort of systematic pattern in the data.
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=3.5in]{reasoning_words.pdf} 
+   \caption{Words used to explain the reasons for selection of data plot in a lineup. Larger font indicate more people choosing that word. Different color is used just to separate the words.}
+   \label{fig:wordle}
+\end{figure}
+
+Figure \ref {fig:wordle} aslo shows some insight about peoples way of reading a plot.
+We notice that variability in the data, color and aesthetics used to generate plots and existence of any systematic pattern in the plot are some of the important features revealed from the figure.
+These features are commonly used by human brain to examine and compare plots.
+Notice that these features may be specific to this particular experiment.
+There could be different other features people would use to evaluate a lineup depending on the situation.
+But these words we have a general idea how people may think to evaluate a lineup.
 

### 2013-05-22 sociology_paper.tex
amazon turk study explained

@@
 The researcher knows about the hypothesis but the observer does not know the underlying details of the lineup.
-Thus the researcher needs to ask a question observer needs to answer while evaluating the lineup.
+Thus the researcher needs to ask a question observer is to answer while evaluating the lineup.
 This question should provide the observer a little clue so that the answer reflects the hypothesized patterns in the actual plot.
@@
 
-Amazon Mechanical Turk \cite{turk} is an online work place where people from around the world can perform simple task to get paid.
+Amazon Mechanical Turk \cite{turk} or MTurk is an online work place where people from around the world can perform some tasks and get paid.
 Usually tasks are very simple and no specialized training is required.
 Being a human is the main requirement.
+Tasks are designed for anyone to do but some tasks may require that workers satisfy some skill level depending on the recruiters need.
 The tasks are designed such that it does not take much time to complete.
@@
 The the amount of money paid for each task is very small as well.
-Figure \ref{fig:amazon_task} displays an example of such a task.
+Figure \ref{fig:amazon_task} displays an example of such a task taken from their web site.
+In this task a worker just has to select from a pool of options and submit the task.
+The instructions are usually accompany the task and they are very simple and easy to follow.
 
@@
    \caption{An example of amazon mechanical turk task.
-Tasks are usually very simple and human are still better than computer in performing these types of tasks.
+Tasks are usually very simple and designed for human evaluations.
 With each task, simple instructions are given for workers to follow.
@@
 
+It is very fast, cheap and reliable to recruit people from MTurk.
+Thats why it is getting very popular among the researchers who perform human subject experiment.
+The another benefit is that a very diverse pool of subjects can be recruited which is otherwise very hard to obtain for a study.
+The researchers can easily filter the workers based on their experimental design, such as recruiting people only from a specific geographical location or a group of people who satisfy certain criteria etc.
+The recruiter can decide who they pay or not.
+Workers have to satisfy the task requirement to ensure payment.
+But at the end it is the recruiter who has the final say.
+Usually recruiters pay promptly after the task has been done properly and thats why MTurk is very popular among the online job seeker.
+At any point of time thousands of tasks are available for the thousands of workers around the world.
+
+Because of its convenience it is getting popular for scientific research study.
+In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time, cost and provided more convenience.
+\cite{majumder:2012} recruited people from MTurk for their simulation study in estimating the power of visual statistical inference.
+They have done numerous pilot studies in lab before doing actual MTurk study and found similar results.
+\cite{mason:2012} explains numerous features of MTurk and describes how it can be used as part of human behavioral study.
+
 \section{Results}
@@
 It is a source of diverse workers who can give us feedback on visual inspection.
-This is the kind of task where human are better than computers.
 Figure \ref{fig:turker_location} shows the location from where we received data.
@@
    \centering
-   \includegraphics[width=4in]{turker_country.pdf} 
-   \includegraphics[width=4in]{turker_location.pdf}    
-   \caption{Words used to explain the reasons for selection of data plot in a lineup. Larger font indicate more people choosing that word. Different color is used just to separate the words.}
+   \includegraphics[width=4.5in]{turker_country.pdf} 
+   \includegraphics[width=4.5in]{turker_location.pdf}    
+   \caption{Location of the Amazon Mechanical Turk workers participating our study. Most of the people are coming from India and United States even though there are people from around the world.}
    \label{fig:turker_location}
@@
    \includegraphics[width=3.5in]{reasoning_words.pdf} 
-   \caption{Words used to explain the reasons for selection of data plot in a lineup. Larger font indicate more people choosing that word. Different color is used just to separate the words.}
+   \caption{Words used to explain the reasons for selection of data plot in a lineup show what features of a lineup may influence a non-statistician to examine it.
+Larger font indicate more people choosing that word.
+Different color is used just to separate the words.}
    \label{fig:wordle}
@@
 
-Figure \ref {fig:wordle} aslo shows some insight about peoples way of reading a plot.
+Figure \ref {fig:wordle} also shows some insight about peoples way of reading a plot.
 We notice that variability in the data, color and aesthetics used to generate plots and existence of any systematic pattern in the plot are some of the important features revealed from the figure.
@@
 
+Turk workers are not necessarily trained on statistics or aware of specific terms used in statistics or statistical plots.
+Thus it is also interesting to note that how they explain things that have specific definitions and meaning.
+Notice in Figure \ref {fig:wordle} that some keywords like spread, apart should be analogous to larger variability while together,  close may be for indicating smaller variability.
+
 \subsection{Learning Trend of the Observer} Since each subject is shown multiple lineups, they have a chance for self learning and do better on lineups shown later in the sequence.
-We examine this by fitting generalized mixed effect model presented in \cite{majumder:2012} with covariates attempt number and $p$-value of the lineup.
+They may learn from their previous mistake or they may learn about the features of the plots being evaluated as they progress.
+We examine this by fitting generalized mixed effect model presented in \cite{majumder:2012} with lineups and subjects as random effect and no fixed effect.
+This should give us the estimation of proportion correct responses adjusted for lineup difficulties and individual skills.
+The resulting residual should show some trend if plotted against attempt.
+Figure \ref{fig:learning_trend} shows such plots for three experiments naming experiment 5 through 6.
+For experiment 5 we see a downward trend in absolute residual as the attempt progress.
+This may indicate a learning trend but this trend is not statistically significant.
+We see upward trend for experiment 6 and 7 but again they are not statistically significant too.
+
+%covariates attempt number and $p$-value of the lineup.
 The $p$-value of the lineup measures some sort of difficulty and including this should adjust for the difficulty level of the lineup while estimating the probability of correct responses in each attempt.
@@
    \centering
-   \includegraphics[width=4in]{learning_trend.pdf} 
+   \includegraphics[width=6.5in]{learning_trend.pdf} 
    \caption{Mean absolute residuals vs attempt obtained by adjusting for lineup and subject specific variation.
-The trend over attempt does not show any learning trend since the slopes are not statistically significant.}
+For experiment 5, we see a downward trend which may indicate some learning as attempt progress, but none of the slopes in these three experiments are statistically significant.}
    \label{fig:learning_trend}
@@
 
-% latex table generated in R 2.15.0 by xtable 1.7-0 package
-% Fri Sep  7 10:28:47 2012
-\begin{table}[hbtp]
-\caption{Fixed effect parameter estimates of generalized mixed model.
+%% latex table generated in R 2.15.0 by xtable 1.7-0 package
+%% Fri Sep  7 10:28:47 2012
+%\begin{table}[hbtp]
+%\caption{Fixed effect parameter estimates of generalized mixed model.
 Note that attempt is not significant for experiment 2.
 The continuous covariate, lineup difficulty, was measured by the $p$-value of the actual plot in the lineup.}
-\begin{center}
-\begin{tabular}{llrrrrl}
-  \hline
-& Parameters & Estimate & Std..Error & z.value & P-value  & \\ 
-  \hline
-\multicolumn{2}{l}{\bf{Experiment 1} } &&&& &\\
-&(Intercept) & -0.30 & 0.10 & -2.89 & 0.00 & ***\\ 
- & attempt & 0.08 & 0.02 & 4.21 & 0.00 & ***\\ 
- & lineup difficulty & -11.32 & 1.01 & -11.19 & 0.00 & ***\\ 
-\multicolumn{2}{l}{\bf{Experiment 2} } &&&&& \\
- & (Intercept) & 2.36 & 0.15 & 16.09 & 0.00 & ***\\ 
- & attempt & 0.01 & 0.03 & 0.35 & 0.72 \\ 
- & lineup difficulty & -55.03 & 2.20 & -25.00 & 0.00 & ***\\ 
-\multicolumn{2}{l}{\bf{Experiment 3} } &&&& \\
- & (Intercept) & 0.25 & 0.16 & 1.58 & 0.11 & .\\ 
- & attempt & 0.10 & 0.03 & 3.34 & 0.00 & ***\\ 
- & lineup difficulty & 3.00 & 0.56 & 5.36 & 0.00 & ***\\ 
-   \hline
-\end{tabular}
-\end{center}
-\label{tbl:model_par}
-\end{table}
+%\begin{center}
+%\begin{tabular}{llrrrrl}
+%  \hline
+%& Parameters & Estimate & Std..Error & z.value & P-value  & \\ 
+%  \hline
+%\multicolumn{2}{l}{\bf{Experiment 1} } &&&& &\\
+%&(Intercept) & -0.30 & 0.10 & -2.89 & 0.00 & ***\\ 
+% & attempt & 0.08 & 0.02 & 4.21 & 0.00 & ***\\ 
+% & lineup difficulty & -11.32 & 1.01 & -11.19 & 0.00 & ***\\ 
+%\multicolumn{2}{l}{\bf{Experiment 2} } &&&&& \\
+% & (Intercept) & 2.36 & 0.15 & 16.09 & 0.00 & ***\\ 
+% & attempt & 0.01 & 0.03 & 0.35 & 0.72 \\ 
+% & lineup difficulty & -55.03 & 2.20 & -25.00 & 0.00 & ***\\ 
+%\multicolumn{2}{l}{\bf{Experiment 3} } &&&& \\
+% & (Intercept) & 0.25 & 0.16 & 1.58 & 0.11 & .\\ 
+% & attempt & 0.10 & 0.03 & 3.34 & 0.00 & ***\\ 
+% & lineup difficulty & 3.00 & 0.56 & 5.36 & 0.00 & ***\\ 
+%   \hline
+%\end{tabular}
+%\end{center}
+%\label{tbl:model_par}
+%\end{table}
 

### 2013-05-22 R/sociology_paper.R

@@ -148,6 +148,7 @@ ggsave(file="../images/pval_difference_plot_region.pdf", height=4, width=8)
 # getting those three unusual lineups for experiment 1
 subset(pval.diff, prop_correct <.25 & plot_region == "Centre" & pval_diff <0)
 
+# =====================================================
 # Examining learning trend while giving feedback
 
 get_trend <- function(dat){
@@ -283,11 +284,21 @@ ggplot(trend.dat, aes(attempt,resid, colour=experiment))+
 ddt <- ddply(trend.dat,.(attempt, experiment), summarise,
              mean_resid = mean(resid))
 
-qplot(attempt,mean_resid, data= ddt, colour=experiment) +
-  geom_smooth(method="lm") + ylab("Mean absolute residual") +
-  scale_x_continuous(breaks = seq(2,10,by=2))
+qplot(attempt,mean_resid, data= ddt) + geom_point(size=2.5) +
+  geom_smooth(method="lm", se=F) + ylab("Mean absolute residual") +
+  facet_wrap(~experiment, scales="free_y") +
+  scale_x_continuous(breaks = seq(2,10,by=2)) 
 
-ggsave("../images/learning_trend.pdf", width=6, height = 4)
+ggsave("../images/learning_trend.pdf", width=10.5, height = 3.5)
+
+
+# Checking if the trend shown in the plot is significant or not
+# For all experiments 5,6,7 slope is not statistically significant
+
+fit1 <- lm(resid ~ attempt, data=subset(trend.dat, experiment=="experiment 5"))
+fit2 <- lm(resid ~ attempt, data=subset(trend.dat, experiment=="experiment 6"))
+fit3 <- lm(resid ~ attempt, data=subset(trend.dat, experiment=="experiment 7"))
+
 
 dd <- subset(dtrend4, id %in% id[attempt > 9])
 fit <- lmer(model,family="binomial",data=dd)
@@ -355,7 +366,7 @@ anova(fit2, fit1)
 fit3 <- lmer(response ~ factor(attempt) + (attempt|id) + (1|pic_id), family="binomial", data=dt)
 anova(fit3,fit2)
 
-
+# =======================================================
 # some analysis of turk9 data and location effect of the plot
 
 # keep only three responses data. Any responses after the first three are ignored.

### 2013-05-22 sociology_paper.tex
abstract updated

@@
 The simulation experiments suggest that individual skill as well as demographics are very significant for the power of visual inference.
-Moreover, the data shows the evidence of learning or getting skilled while sequentially evaluating multiple lineups.
+We examine the evidence of learning or getting skilled while sequentially evaluating multiple lineups.
+The experimental data suggest that there is no learning effect.
+The location of actual plot in the lineups does not have any significant effect as well.
 
@@
 They have done numerous pilot studies in lab before doing actual MTurk study and found similar results.
-\cite{mason:2012} explains numerous features of MTurk and describes how it can be used as part of human behavioral study.
+\cite{mason:2012} explains various features of MTurk and describes how it can be used as part of human behavioral study.
 
@@
 
-\subsection{Learning Trend of the Observer} Since each subject is shown multiple lineups, they have a chance for self learning and do better on lineups shown later in the sequence.
+\subsection{Learning Trend of the Observer} Each subject participating the experiment has seen multiple number of lineups for evaluations.
+Suppose $K$ be the number of lineups evaluated by a subject and $A_k$ be the $kth$ attempt, $k=1,2, ..., K$.
+Since each subject is shown multiple lineups, they have a chance for self learning and do better on lineups shown later in the sequence.
 They may learn from their previous mistake or they may learn about the features of the plots being evaluated as they progress.
-We examine this by fitting generalized mixed effect model presented in \cite{majumder:2012} with lineups and subjects as random effect and no fixed effect.
-This should give us the estimation of proportion correct responses adjusted for lineup difficulties and individual skills.
-The resulting residual should show some trend if plotted against attempt.
+Thus if there is any learning trend that should be apparent in proportion correct over $A_k$.
+But to obtain this we have to adjust for lineup difficulty and individual performances.
+
+To accomplish this we fit generalized mixed effect model presented in \cite{majumder:2012} with lineups and subjects as random effect and no fixed effect.
+This gives the estimation of proportion correct responses  and we obtain residuals of the fitted model.
+The residuals give variations in proportion correct adjusted for lineup difficulties and individual skills.
+If there is any learning trend over different attempt $A_k$, the residuals should have that information since $A_k$ was not included as a covariate in the model.
+At this point we focus on absolute residuals over different attempts to study the deviations due to attempts.
+An smaller absolute residual at attempt $A_k$ should indicate learning from all other attempts less than $k$.
+
+A plot of absolute residuals against the attempts $A_k$ should reveal the trend if there is any.
 Figure \ref{fig:learning_trend} shows such plots for three experiments naming experiment 5 through 6.
@@
 This may indicate a learning trend but this trend is not statistically significant.
-We see upward trend for experiment 6 and 7 but again they are not statistically significant too.
+We see upward trend for experiment 6 and 7.
+This could be attributed to the exhaustion of subjects' attempt after some earlier evaluations are made.
+But again these upward trends are not statistically significant as well.
 

### 2013-05-23 R/sociology_paper.R
canged R codes for version 3

@@ -2,7 +2,7 @@
 
 library(ggplot2)
 library(plyr)
-library(reshape)
+library(reshape2)
 library(lme4)
 library(xtable)
 library(grid)
@@ -104,12 +104,12 @@ qplot(p_value, per_correct, data = loc.performance, geom="point")
 qplot(plot_location,per_correct, geom="point", data = loc.performance)+
   facet_wrap(~plot_location, ncol=5)
 
-qplot(x=1,y=per_correct, geom="bar", binwidth=1, data = loc.performance)+
+qplot(x=1,y=per_correct, geom="bar", binwidth=1, data = loc.performance, stat="identity")+
   facet_wrap(~plot_location, ncol=5)
 
-qplot(1,1, geom="bar", binwidth=1,data = loc.performance, fill=per_correct)+
-  facet_wrap(~plot_location, ncol=5)+ scale_x_continuous()+ scale_y_continuous()
-  + opts(axis.ticks=theme_blank())
+qplot(1,1, geom="bar", binwidth=1,data = loc.performance, fill=per_correct, stat="identity")+
+  facet_wrap(~plot_location, ncol=5)+ scale_x_continuous()+ scale_y_continuous() + 
+  theme(axis.ticks=element_blank())
 
 # difference with minimum p-value and location of actual plot
 
@@ -192,15 +192,17 @@ qplot(data=dtrend1, attempt, response, geom="point", colour=id) +
 
 get_smooth_loess <- function(dtrend){
   smooth_fitted =NULL
-  for (i in unique(dtrend1$id)){
+  for (i in unique(dtrend$id)){
     #i=221
-    dat <- subset(dtrend1, id==i)
-    max_attempt <- max(dat$attempt)
-    attempt  <- seq(1,max_attempt, by=.1) 
-    fit <- loess(response~attempt, data=dat)
-    fitted <- predict(fit, attempt)
-    fdat <- data.frame(id=i, attempt, fitted)
-    smooth_fitted <- rbind(smooth_fitted, fdat)
+    dat <- subset(dtrend, id==i)
+    if(nrow(dat)>9){
+      max_attempt <- max(dat$attempt)
+      attempt  <- seq(1,max_attempt, by=.1) 
+      fit <- loess(response~attempt, data=dat)
+      fitted <- predict(fit, attempt)
+      fdat <- data.frame(id=i, attempt, fitted)
+      smooth_fitted <- rbind(smooth_fitted, fdat)
[…]

### 2013-05-23 sociology_paper.tex
Age and study figures added

@@
    \label{fig:turker_location}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=3.5in]{country_study_bar.pdf} 
+   \caption{Countrywise distribution of education level of the MTurk workers participating the experiments.
+The spike in graduate degree in India may be due to their consideration of undergraduate degree to be graduate degree as well.}
+   \label{fig:country_study}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=3.5in]{age_study_bar.pdf} 
+   \caption{Countrywise distribution of age levels of the MTurk workers participating the experiments.}
+   \label{fig:age_study}
 \end{figure}

### 2013-05-25 sociology_paper.tex
demograpic information figure added

--- 2013-05-22 sociology_paper.tex
@@
 
+%\begin{figure}[htbp] 
+%   \centering
+%   \includegraphics[width=3.5in]{age_study_country_bar.pdf} 
+%   \caption{Countrywise distribution of education level of the MTurk workers participating the experiments.
+The spike in graduate degree in India may be due to their consideration of undergraduate degree to be graduate degree as well.}
+%   \label{fig:country_study}
+%\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=6.5in]{demographic_info.pdf} 
+   \caption{Countrywise distribution of age levels of the MTurk workers participating the experiments.}
+   \label{fig:age_study}
+\end{figure}
+
 \subsection{How people pick the data plot} In the experiment setup we have a question for the turk observers to choose the reason for their selection of a particular data plot.
@@
    \includegraphics[width=3.5in]{reasoning_words.pdf} 
-   \caption{Words used to explain the reasons for selection of data plot in a lineup show what features of a lineup may influence a non-statistician to examine it.
+   \caption{Words used to explain the reasons for selection of data plot in a lineup show what features of a lineup may help a non-statistician to evaluate it.
 Larger font indicate more people choosing that word.

### 2013-05-25 R/sociology_paper.R

@@ -8,8 +8,9 @@ library(xtable)
 library(grid)
 #library(lubridate)
 
-
-# getting the data and some common functions
+# ===================================================
+# getting all the data and some common functions
+# ---------------------------------------------------
 
 raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
 raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
@@ -18,11 +19,18 @@ raw.dat4 <- read.csv("../data/raw_data_turk4.csv")
 raw.dat5 <- read.csv("../data/raw_data_turk5.csv")
 raw.dat6 <- read.csv("../data/raw_data_turk6.csv")
 raw.dat7 <- read.csv("../data/raw_data_turk7.csv")
+raw.dat8 <- read.csv("../data/raw_data_turk8.csv")
+raw.dat9 <- read.csv("../data/raw_data_turk9.csv")
+raw.dat10 <- read.csv("../data/raw_data_turk10.csv")
+
+ip.details <- read.csv("../data/ip_details.csv")
 
 source("calculate_ump_power.R") # functions to compute power
 
+# ========================================================================
 # cleaning the data based on criteria 6 and removing the duplicated data
 # (id,pic_id) should be unique since no person gets a lineup more than once.
+# --------------------------------------------------------------------------
 clean_data <- function(raw_dat){
   easy_dat <- subset(raw_dat, p_value < 0.0002)
   if (raw_dat$experiment[1]=='turk3') easy_dat <- subset(raw_dat, difficulty==0)
@@ -57,11 +65,99 @@ dat4$p_value=0
 dat5 <- remove_duplicated(raw.dat5)
 dat6 <- remove_duplicated(raw.dat6)
 dat7 <- remove_duplicated(raw.dat7)
+dat8 <- remove_duplicated(raw.dat8)
+dat10 <- remove_duplicated(raw.dat10)
+
+# Cleaning up experiment 9 data
+# keep only three responses data. Any responses after the first three are ignored.
+# Also data with less than three responses are discarded considering incomplete.
+
+indx_unique <- with(raw.dat9, !duplicated(data.frame(id,pic_id))) #unique data index
+uniq.dat9 <- raw.dat9[indx_unique,] # removing duplicated data from raw.dat if any
+uniq.dat9 <- ddply(uniq.dat9,.(id), transform,
+                   response_order = rank(start_time), 
+                   tot_response=length(response) )
+dat9 <- subset(uniq.dat9, response_order < 4 & tot_response > 2) # filering applied
[…]

### 2013-05-27 sociology_paper.tex
demographic information added

@@
 The fundamental concept of visual statistical inference is introduced by \citet{buja:2009} and later these concepts are validated by \citet{majumder:2012}.
+
+%\section{Kodu} Tui ekta chor. Tore kilai.
 
@@
 Because of its convenience it is getting popular for scientific research study.
-In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time, cost and provided more convenience.
+In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time and cost while provided more convenience.
 \cite{majumder:2012} recruited people from MTurk for their simulation study in estimating the power of visual statistical inference.
@@
    \centering
-   \includegraphics[width=6.5in]{demographic_info.pdf} 
-   \caption{Countrywise distribution of age levels of the MTurk workers participating the experiments.}
+   \includegraphics[width=6in]{demographic_info.pdf} 
+   \caption{Countrywise distribution of age and academic levels of the MTurk workers participating the experiments shows the diversity of the subjects in all the demographic aspect.
+Almost equal number of male and female subjects participated the online experiments.}
    \label{fig:age_study}
@@
    \centering
+   \includegraphics[width=6.5in]{learning_trend_time.pdf} 
+   \caption{Mean residuals of time taken vs attempt obtained by adjusting for lineup and subject specific variation.
+For all the experiment the downward slopes indicate that MTurk workers take less time as they progress through their attempts.}
+   \label{fig:learning_trend_time}
+\end{figure}
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
    \includegraphics[width=6.5in]{learning_trend.pdf} 
-   \caption{Mean absolute residuals vs attempt obtained by adjusting for lineup and subject specific variation.
+   \caption{ Mean residuals of proportion correct vs attempt obtained by adjusting for lineup and subject specific variation.
 For experiment 5, we see a downward trend which may indicate some learning as attempt progress, but none of the slopes in these three experiments are statistically significant.}
@@
 
+\begin{table}[hbtp]
+\caption{Demographic information of the subjects participated the MTurk experiments.}
+\centering
+\begin{tabular}{rlrrr}
+  \hline
+ & levels & subsjects & avg\_time & response \\ 
+  \hline
+Gender & Male & 1348 & 48.51 & 13493 \\ 
+   & Female & 991 & 43.75 & 10564 \\ 
+\hline
+  Education & High school or less & 193 & 37.21 & 2241 \\ 
+   & Some under graduate courses & 418 & 42.84 & 4070 \\ 
+   & Under graduate degree & 584 & 44.29 & 5775 \\ 
+   & Some graduate courses & 245 & 43.43 & 2460 \\ 
+   & Graduate degree & 902 & 52.18 & 9511 \\ 
+\hline
+  Age & 18-25 & 740 & 42.97 & 7311 \\ 
+   & 26-30 & 547 & 46.27 & 5585 \\ 
+   & 31-35 & 376 & 44.27 & 3923 \\ 
+   & 36-40 & 257 & 55.03 & 2714 \\ 
+   & 41-45 & 141 & 43.90 & 1519 \\ 
+   & 46-50 &  95 & 49.29 & 1003 \\ 
+   & 51-55 &  83 & 48.67 & 867 \\ 
+   & 56-60 &  64 & 59.73 & 678 \\ 
+   & above 60 &  38 & 48.67 & 457 \\ 
+   \hline
+\end{tabular}\label{tbl:demographics}
+\end{table}
+
 %\begin{table}[hbtp]

### 2013-05-27 R/sociology_paper.R

@@ -9,7 +9,7 @@ library(grid)
 #library(lubridate)
 
 # ===================================================
-# getting all the data and some common functions
+# Loading the data and some common functions
 # ---------------------------------------------------
 
 raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
@@ -79,6 +79,8 @@ uniq.dat9 <- ddply(uniq.dat9,.(id), transform,
                    tot_response=length(response) )
 dat9 <- subset(uniq.dat9, response_order < 4 & tot_response > 2) # filering applied
 dat9$plot_type <- factor(dat9$difficulty, labels=c("Interaction", "Genotype", "Filter"))
+
+# dat9 response_all is true if at least one of the multiple selection is correct
 dat9$response_all <- apply(subset(dat9, select=c(plot_location,response_no)), 1,
                   function(x){
                     plot <- as.numeric(x[1])
@@ -86,22 +88,20 @@ dat9$response_all <- apply(subset(dat9, select=c(plot_location,response_no)), 1,
                     return(plot %in% res_no)
                   })
 
-# write.csv(dat9$choice_reason[dat9$response_all==T], file="../data/dat9_reasoning_correct.csv")
-# write.csv(dat9$choice_reason[dat9$response_all==F], file="../data/dat9_reasoning_wrong.csv")
-
 
 # ==============================================================
 # Summary statistics of demographic data
 # --------------------------------------------------------------
 
-# Function to obtain real ip from proxy and local
+# Function to obtain real ip from proxy and local part of ip
+# First ip of two ip's is the real ip, 2nd one is local
+# in case where no proxy is used, the ip is real ip
 get_real_ip <- function(ip) {
   ip_vector <- strsplit(as.character(ip),split=',')
   return(ip_vector[[1]][1])
 }
 
 # merging demographic data from all the experiment together
-# and geogrphical location using ip address
 demographics <- NULL
 for (i in 1:10){
   di <- subset(get(paste("dat",i, sep="")), age > 1,
@@ -111,29 +111,43 @@ for (i in 1:10){
   di$academic_study[di$academic_study==0] <- NA
   di$experiment = paste("experiment_",i, sep="")
   di$id = paste("exp",i,"_",di$id, sep="")
[…]

### 2013-05-27 sociology_paper.tex
turker time of participation plot added

@@
 The simulation experiments suggest that individual skill as well as demographics are very significant for the power of visual inference.
-We examine the evidence of learning or getting skilled while sequentially evaluating multiple lineups.
-The experimental data suggest that there is no learning effect.
+We examine the evidence of learning or getting skilled while sequentially evaluating multiple lineups which enables them evaluate a lineup faster in the later ateempts.
+The experimental data suggest that there is no learning effect in terms of correct response.
 The location of actual plot in the lineups does not have any significant effect as well.
@@
 We notice diversity in not only location of the observers but also their gender, age groups and education levels.
-Table \ref{} shows that we have data from almost similar size of male and female population.
+Table \ref{tbl:demographics} shows that number of male and female participants are almost same.
 
@@
 
-%\begin{figure}[htbp] 
-%   \centering
-%   \includegraphics[width=3.5in]{age_study_country_bar.pdf} 
-%   \caption{Countrywise distribution of education level of the MTurk workers participating the experiments.
-The spike in graduate degree in India may be due to their consideration of undergraduate degree to be graduate degree as well.}
-%   \label{fig:country_study}
-%\end{figure}
+The large number of participants are from age group 18 to 35.
+Interestingly we see many participants from older age groups as well.
+Specially for united states almost all the age groups show uniform participations after age 30 as in Figure \ref{fig:demographic_info}.
+Notice that fewer people participated from India beyond age 40 compared to united states.
+Even thoug total participations from india is larger compared to United States, they are mostly young people who are capable of using internet and computer.
 
@@
 Almost equal number of male and female subjects participated the online experiments.}
-   \label{fig:age_study}
+   \label{fig:demographic_info}
 \end{figure}
@@
 
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=4.5in]{participation_time.pdf} 
+   \caption{Time of the day when the participants work. Experiment 1 shows MTurk workers participated the experiments around the clock. Other experiments did not take a whole day to finish.}
+   \label{fig:participation_time}
+\end{figure}
+
 \subsection{Learning Trend of the Observer} Each subject participating the experiment has seen multiple number of lineups for evaluations.
 Suppose $K$ be the number of lineups evaluated by a subject and $A_k$ be the $kth$ attempt, $k=1,2, ..., K$.
-Since each subject is shown multiple lineups, they have a chance for self learning and do better on lineups shown later in the sequence.
+Subjects may have a chance for self learning and do better on lineups shown later in the sequence.
 They may learn from their previous mistake or they may learn about the features of the plots being evaluated as they progress.
@@
 If there is any learning trend over different attempt $A_k$, the residuals should have that information since $A_k$ was not included as a covariate in the model.
-At this point we focus on absolute residuals over different attempts to study the deviations due to attempts.
+%At this point we focus on absolute residuals over different attempts to study the deviations due to attempts.
 An smaller absolute residual at attempt $A_k$ should indicate learning from all other attempts less than $k$.
 
-A plot of absolute residuals against the attempts $A_k$ should reveal the trend if there is any.
-Figure \ref{fig:learning_trend} shows such plots for three experiments naming experiment 5 through 6.
-For experiment 5 we see a downward trend in absolute residual as the attempt progress.
-This may indicate a learning trend but this trend is not statistically significant.
-We see upward trend for experiment 6 and 7.
+A plot of residuals against the attempts $A_k$ should reveal the trend if there is any.
+Figure \ref{fig:learning_trend} shows such plots for three experiments naming experiment 5, 6 and 7.
+For all the experiments we see some increasing trend of the residuals.
 This could be attributed to the exhaustion of subjects' attempt after some earlier evaluations are made.
-But again these upward trends are not statistically significant as well.
+But none of apparent trend are statistically significant.
 
@@
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=6.5in]{learning_trend.pdf} 
+   \caption{ Mean residuals of proportion correct vs attempt obtained by adjusting for lineup and subject specific variation.
+For experiment 5, we see a downward trend which may indicate some learning as attempt progress, but none of the slopes in these three experiments are statistically significant.}
+   \label{fig:learning_trend}
+\end{figure}
+
+Learning trend is not only related to the proportion correct responses.
+It may be possible that people learn to answer fast.
+We examine this by the time taken for each evaluation.
+We fit a mixed effect linear model with time taken considering lineup and subjects as random effect.
+The resulting residuals are plotted in Figure \ref{fig:learning_trend_time}.
+It appears that in the later attempts observers take less time to evaluate a lineup.
+Even though it does not tell much about improving their performance over attempts, but it does indicate the improvement of their evaluation skill.
+Adding this information with what we observe in Figure \ref{fig:learning_trend}, we can say that the observers' improved skills could not contribute to their performance.
+It only help them finish the task faster in later attempts.
+
+\begin{figure}[htbp] 
    \centering
@@
    \label{fig:learning_trend_time}
-\end{figure}
-
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=6.5in]{learning_trend.pdf} 
-   \caption{ Mean residuals of proportion correct vs attempt obtained by adjusting for lineup and subject specific variation.
-For experiment 5, we see a downward trend which may indicate some learning as attempt progress, but none of the slopes in these three experiments are statistically significant.}
-   \label{fig:learning_trend}
 \end{figure}
@@
 
-\begin{table}[hbtp]
-\caption{Demographic information of the subjects participated the MTurk experiments.}
-\centering
-\begin{tabular}{rlrrr}
-  \hline
- & levels & subsjects & avg\_time & response \\ 
-  \hline
-Gender & Male & 1348 & 48.51 & 13493 \\ 
-   & Female & 991 & 43.75 & 10564 \\ 
-\hline
-  Education & High school or less & 193 & 37.21 & 2241 \\ 
-   & Some under graduate courses & 418 & 42.84 & 4070 \\ 
-   & Under graduate degree & 584 & 44.29 & 5775 \\ 
-   & Some graduate courses & 245 & 43.43 & 2460 \\ 
-   & Graduate degree & 902 & 52.18 & 9511 \\ 
-\hline
-  Age & 18-25 & 740 & 42.97 & 7311 \\ 
-   & 26-30 & 547 & 46.27 & 5585 \\ 
-   & 31-35 & 376 & 44.27 & 3923 \\ 
-   & 36-40 & 257 & 55.03 & 2714 \\ 
-   & 41-45 & 141 & 43.90 & 1519 \\ 
-   & 46-50 &  95 & 49.29 & 1003 \\ 
-   & 51-55 &  83 & 48.67 & 867 \\ 
-   & 56-60 &  64 & 59.73 & 678 \\ 
-   & above 60 &  38 & 48.67 & 457 \\ 
-   \hline
-\end{tabular}\label{tbl:demographics}
-\end{table}
-
-%\begin{table}[hbtp]
-%\caption{Parameter estimates of mixed model. Estimates are highly significant with $p$-value $<$  0.0001 for all three experiment data.}
-%\begin{center}
-%\begin{tabular}{cr@{.}lcc}
-%  \hline \hline
-% &  \multicolumn{3}{c} {Fixed effect}  & Random effect\\
-% \cline{2-4}
-% Experiment & \multicolumn{2}{l}{Estimate}  &Std. error & Variance\\
-%  \hline
-%  1 & 0&39 & 0.0094 & 0.0080 \\ 
-%  2 & 1&21 &  0.0197 &  0.0443 \\ 
-%  3 & 0&59 (Intercept)  &   0.1668 & 1.9917\\ 
-%     & 0&21 (Slope)    &  0.0511     &  0.0245\\ 
-%     &-0&78 (correlation) & & \\
-%   \hline
-%\end{tabular}
-%\end{center}
-%\label{tbl:model_par}
-%\end{table}
-
 %% latex table generated in R 2.15.0 by xtable 1.7-0 package
@@
 
-%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-%   \centering
-%   \includegraphics[width=4in]{learning_trend.pdf} 
-%   \caption{Probability of correct response for different attempts by each subjects for a plot with lineup difficulty ($p$-value of actual plot) = 0.05 .
-The overall probability increases with attempts indicating a learning trend of the observers for experiments 1 and 3.
-For experiment 2 the overall trend is not statistically significant.
-Subject to subject variability is seen in learning pattern.}
-%   \label{fig:learning_trend}
-%\end{figure}
+\section{Conclusion}
 
-\section{Conclusion}
+\appendix
+
+\begin{table}[hbtp]
+\caption{Demographic information of the subjects participated the MTurk experiments.}
+\centering
+\begin{tabular}{rlrrr}
+  \hline
+ & levels & subsjects & avg\_time & response \\ 
+  \hline
+Gender & Male & 1348 & 48.51 & 13493 \\ 
+   & Female & 991 & 43.75 & 10564 \\ 
+\hline
+  Education & High school or less & 193 & 37.21 & 2241 \\ 
+   & Some under graduate courses & 418 & 42.84 & 4070 \\ 
+   & Under graduate degree & 584 & 44.29 & 5775 \\ 
+   & Some graduate courses & 245 & 43.43 & 2460 \\ 
+   & Graduate degree & 902 & 52.18 & 9511 \\ 
+\hline
+  Age & 18-25 & 740 & 42.97 & 7311 \\ 
+   & 26-30 & 547 & 46.27 & 5585 \\ 
+   & 31-35 & 376 & 44.27 & 3923 \\ 
+   & 36-40 & 257 & 55.03 & 2714 \\ 
+   & 41-45 & 141 & 43.90 & 1519 \\ 
+   & 46-50 &  95 & 49.29 & 1003 \\ 
+   & 51-55 &  83 & 48.67 & 867 \\ 
+   & 56-60 &  64 & 59.73 & 678 \\ 
+   & above 60 &  38 & 48.67 & 457 \\ 
+   \hline
+\end{tabular}\label{tbl:demographics}
+\end{table}
 

### 2013-05-27 R/sociology_paper.R

@@ -6,7 +6,7 @@ library(reshape2)
 library(lme4)
 library(xtable)
 library(grid)
-#library(lubridate)
+library(lubridate)
 
 # ===================================================
 # Loading the data and some common functions
@@ -169,6 +169,15 @@ print(p2 ,vp = viewport(layout.pos.row = 1, layout.pos.col = 2))
 dev.off()
 
 
+# Time of the work
+turker$hours <- as.factor(hour(as.POSIXlt(turker$start_time)))
+turker$exp <- factor(turker$experiment, levels=names(table(turker$experiment))[order(c(1,10,2:9))])
+qplot(hours, data=turker) +
+  facet_grid(exp~.,scales="free_y") +
+  xlab("Hour of the day") + ylab("Number of subjects")
+ggsave("../images/participation_time.pdf", width=8, height=10)
+
+
 # getting demographic summary table
 get_summary <- function(dat, var){
   res <- ddply(dat,c(var), summarize,
@@ -188,6 +197,7 @@ xtable(sdat[,-1])
 
 qplot(lbls, avg_time, geom="bar", stat="identity", data=sdat) + coord_flip()
   facet_wrap(var~.)
+
 
 
 # ==============================================================

### 2013-05-27 sociology_paper.tex
Figures on Amazon Task statisitcs added

@@
 \subsection{Experiment Setup} Description of the simulation experiment giving focus on demographics and skills of experimental subjects.
+
+\begin{table}[hbtp]
+\caption{Amazon mechanical turk experiments and their properties. Duration in hours per 100 tasks show the popularity of some tasks compared to others.}
+\centering
+\begin{tabular}{rlrrrrrrr}
+  \hline
+& Experiment& \multicolumn{2}{c}{ Total Task}& Average & \multicolumn{2}{c}{Duration (hour)} & Payment & Pay rate\\
+\cline{3-4} \cline{6-7}
+Serial & description & submitted & rejected & time(min) & Actual & 100 task& \$/task & \$/hour\\ 
+  \hline
+1 & Boxplot & 406 & 106 & 10.68 & 146.48 & 36.08 & 0.50 & 2.81 \\ 
+  2 & Scatterplot & 359 &   9 & 10.80 & 42.68 & 11.89 & 1.00 & 5.58 \\ 
+  3 & Contaminated plot & 219 &  19 & 13.53 & 126.17 & 57.61 & 1.00 & 2.22 \\ 
+  4 & Polar vs Cartesian & 110 &  10 & 20.65 & 11.65 & 10.59 & 1.00 & 2.91 \\ 
+  5 & Hist vs density & 234 &  37 & 17.85 & 41.57 & 17.76 & 1.00 & 3.36 \\ 
+  6 & Violine vs boxplot & 417 &  17 & 17.95 & 105.87 & 25.39 & 1.00 & 3.34 \\ 
+  7 & Group separation & 106 &   6 & 16.13 & 5.15 & 4.86 & 1.00 & 3.72 \\ 
+  8 & Sine Illusion & 101 &   1 & 16.52 & 78.38 & 77.60 & 1.00 & 3.63 \\ 
+  9 & Gene expression & 103 &   3 & 12.47 & 11.27 & 10.94 & 0.50 & 2.41 \\ 
+  10 & Test normality & 406 &   6 & 22.70 & 74.35 & 18.31 & 1.00 & 2.64 \\ 
+   \hline
+\end{tabular}
+\label{tbl:mturk}
+\end{table}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=3in]{rejected_task.pdf}
+      \includegraphics[width=3in]{task_duration.pdf} 
+   \caption{Percentage of rejected tasks and duration in hour per 100 tasks for each of the 10 experiments.
+Most of the tasks got rejected for box plot experiment.
+Even though the sine illusion experiment took longest to finish the rejection rate is lowest for this experiment.}
+   \label{fig:task_duration}
+\end{figure}
+
 \subsection{Estimation of sample size}  How the sample size is estimated
@@
 The the amount of money paid for each task is very small as well.
-Figure \ref{fig:amazon_task} displays an example of such a task taken from their web site.
-In this task a worker just has to select from a pool of options and submit the task.
-The instructions are usually accompany the task and they are very simple and easy to follow.
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=5in]{amazon_task.pdf} 
-   \caption{An example of amazon mechanical turk task.
-Tasks are usually very simple and designed for human evaluations.
-With each task, simple instructions are given for workers to follow.
-The workers first accept the task before submitting their response.}
-   \label{fig:amazon_task}
-\end{figure}
 

### 2013-05-27 R/sociology_paper.R

@@ -24,6 +24,7 @@ raw.dat9 <- read.csv("../data/raw_data_turk9.csv")
 raw.dat10 <- read.csv("../data/raw_data_turk10.csv")
 
 ip.details <- read.csv("../data/ip_details.csv")
+turk.summary <- read.csv("../data/turk_summary.csv")
 
 source("calculate_ump_power.R") # functions to compute power
 
@@ -197,6 +198,32 @@ xtable(sdat[,-1])
 
 qplot(lbls, avg_time, geom="bar", stat="identity", data=sdat) + coord_flip()
   facet_wrap(var~.)
+
+# Turk experiment summary including payments and duration
+turk.summary$duration_hour <- round(turk.summary$duration/3600,2)
+turk.summary$avg_minute <- round(turk.summary$time_per_task/60,2)
+turk.summary$duration_rate <- with(turk.summary, round(duration_hour*100/submitted,2))
+turk.summary$percent_rejected <- with(turk.summary, round(rejected*100/submitted,2))
+xtable(subset(turk.summary,
+              select=c(plot,submitted, rejected,avg_minute,
+                       duration_hour,duration_rate,pay_per_task,pay_rate)))
+
+  
+
+ggplot(turk.summary, aes(x = factor(plot, levels=plot[order(duration_rate)]), y=duration_rate))+
+  geom_bar(stat="identity") + ylab("Duration in hour per 100 task") + 
+  xlab("Experiment") + coord_flip() 
+ggsave("../images/task_duration.pdf", width=6, height=4)
+
+
+ggplot(turk.summary, aes(x = factor(plot, levels=plot[order(percent_rejected)]), y=percent_rejected))+
+  geom_bar(stat="identity") + xlab("Experiment") +
+  coord_flip() + ylab("Percentage of rejected task")
+ggsave("../images/rejected_task.pdf", width=6, height=4)
+
+
+
+
 
 
 

### 2013-05-29 sociology_paper.tex
Introduction updated

@@
 \begin {abstract}  
-Visual Inference is dependent on the careful evaluation of the lineups by individual observers.
+Visual Statistical Inference is dependent on the careful evaluation of the lineups by individual observers.
 Each individual is different in their cognitive phycology and judiciousness which can affect the power of visual inference.
@@
 The simulation experiments suggest that individual skill as well as demographics are very significant for the power of visual inference.
-We examine the evidence of learning or getting skilled while sequentially evaluating multiple lineups which enables them evaluate a lineup faster in the later ateempts.
+We examine the evidence of learning or getting skilled while sequentially evaluating multiple lineups which enables them evaluate a lineup faster in the later attempts.
 The experimental data suggest that there is no learning effect in terms of correct response.
@@
 
-The fundamental concept of visual statistical inference is introduced by \citet{buja:2009} and later these concepts are validated by \citet{majumder:2012}.
+Statistical graphics can be used to test the significance of findings during the exploratory data analysis.
+The method is now termed as visual statistical inference which is discussed in the following section.
 
-%\section{Kodu} Tui ekta chor. Tore kilai.
+\subsection{Visual Statistical Inference} The fundamental concept of visual statistical inference is introduced by \citet{buja:2009}.
+In their paper, they proposed lineup protocol to discover significant pattern in the data.
+Later these concepts are further developed by defining necessary terminologies and validating using a head to head comparison with conventional inference by \citet{majumder:2013}.
+The power of the visual test is obtained and it is observed that the power is as good as conventional test while in some scenarios, the power exceeds the conventional test power.
 
-\subsection{Visual Inference} A short introduction of visual inferential procedure.
-\subsection{Estimation of Power of Visual Inference} Discussion of the methods for estimating the power of visual inference \citep{majumder:2012}.
+The test statistic used in visual inference is a plot of the data.
+This plot is then placed on a layout of plot called lineup.
+The rest of the plots in a lineup are generated from the model specified by the null hypothesis.
+An observer is then asked to evaluate the lineup.
+If the observer can detect the observed plot in the lineup, the hypothesis is rejected.
+Figure \ref{fig:lineup_geno} displays a lineup of 20 plots where one of the plots is observed data.
+The rest of the 19 plots are generated from the assumption that there is no difference between two groups of data.
+For this, the group structure of the data is broken by random permutation of the groups.
+If the null hypothesis is true, the observed data should look similar to the rest of the plots in the lineup making it difficult to detect the observed plot.
+On the other hand if the null hypothesis is not true, observed data should be different from the rest of the plots in the lineup.
 
-\section{Factors Affecting the Power of Visual Inference} A brief description of the factors that may affect the power of visual inference.
+A lineup can be evaluated by a single person or multiple observers.
+Based on the observer feedback the $p$-value is computed \citep{majumder:2013} and finally a decision is made about the hypothesis being tested.
 
-\subsection{Choice of Visual Test Statistic} Visual test statistics as defined in \citep{majumder:2012} serve two main purposes.
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=6.5in]{plot_turk9_geno_4_3.png} 
+   \caption{A lineup of 20 plots. One of the plots is observed data and rest of the 19 plots are obtained by randomly permuting the group structure. Can you identify the observed data?}
+   \label{fig:lineup_geno}
+\end{figure}
+
+\subsection{Performance of the observer} Observer who evaluates a lineup do not necessarily be aware of the data underlying a lineup.
+In response to a specific question the observer is to make a choice based on what he or she sees in the lineup.
+The choice could be right or wrong depending on whether the data plot is selected or not.
+Under null hypothesis, all the null plots look similar to what the observed data plot appears to be.
+This would make it harder to detect the observed data plot in the lineup.
+It is not expected that an observer would be able to detect the data plot in this scenario.
+But since there are limited number of plots in a lineup which is 20 in this case, there is a 1/20 chance that the observer would pick the actual plot.
+This proportion is associated with the type-I error probability.
+On the other hand if null hypothesis is not true, the observed plot would be different than the null plots making it easier to be distinguished.
+When multiple observers evaluate a lineup, the proportion of correct response can be used to estimate the power of the visual test \citep{majumder:2013}.
+
+For a lineup with fixed difficulty, the proportion correct responses may vary for different observers.
+If an observe evaluates multiple lineups, the proportion of correct evaluations by the observer could be used as a measure of performance by an observer.
+The another way of measuring the performance of the observer could be the time taken by an observer to evaluate a lineup.
+Since performance would also depend on the lineup difficulty it is essential to adjust for this before estimating the observer performance.
+
+Some of the factors that may affect the performance of the observer are discussed in Section \ref{sec:factor_performance}.
+In this paper we focus on estimating the learning effect of the observer in terms of both proportion correct and time taken to evaluate a lineup through the attempts made during the evaluation of multiple plots.
+We also estimate the location effect of the observed data plot in a lineup.
+The other factors are controlled in simulation experiments so that learning effect and location effect can be estimated.
+
+This paper is organized as follows; Section \ref{sec:factor_performance} describes some of the factors that may affect the performance of the observer.
+Section \ref{sec:exp_design} describes the design of the simulation experiments and methods to estimate the learning trend of the observer and location effect of the observed plot in a lineup.
+Section \ref{sec:result_socio} presents the analysis of the experimental data and the results.
+
+\section{Factors Affecting the performance of the observer} \label{sec:factor_performance}
+
+A brief description of the factors that may affect the power of visual inference.
+
+\subsection{Signal in the data}
+
+The power can be estimated from the proportion of correct evaluations. Thus power could be used as a measure of performance of the observer.
+
+\subsection{Choice of Visual Test Statistic} Visual test statistics as defined in \citep{majumder:2013} serve two main purposes.
 One is to display a prominent pattern that should direct the human observers to select the actual plot in the lineup when the null hypothesis is not true.
@@
 To achieve these purposes it is very important to decide which plot type and plot features should be adopted.
-As discussed in \citep{majumder:2012} following grammar of graphics developed  by \citet{wilkinson:1999} and later improved by \citet{hadley:2009} can give the task a standard control.
+As discussed in \citep{majumder:2013} following grammar of graphics developed  by \citet{wilkinson:1999} and later improved by \citet{hadley:2009} can give the task a standard control.
 
 Some of the effective features of visual test statistics are discussed in \citep{heike:2012} including plot type, color and shape of the plots.
-It is also observed that a scatterplot may do a better job than a box plot when using as a visual test statistic for regression parameters \citep{majumder:2012} .
+It is also observed that a scatterplot may do a better job than a box plot when using as a visual test statistic for regression parameters \citep{majumder:2013} .
 \citep{niladri:2012} presents some \textit{distance measure} to determine how a plot may be different from another.
@@
 This question should provide the observer a little clue so that the answer reflects the hypothesized patterns in the actual plot.
-For example Table \ref{tbl:visual_stat} shows the questions asked for the simulation experiments done by  \citet{majumder:2012}.
+For example Table \ref{tbl:visual_stat} shows the questions asked for the simulation experiments done by  \citet{majumder:2013}.
 Notice that for case 1 if the observer can identify the actual plot in the lineup that should indicate that the plot chosen has the most vertical difference between groups A and B which is exactly what the researchers intend to examine.
@@
 These question help observer not to be misguided by those patterns.
-To review further on this \citet{majumder:2012} have also asked why the observers choose a specific plot.
+To review further on this \citet{majumder:2013} have also asked why the observers choose a specific plot.
 
-\subsection{Observer Personality} Each person is different from others in some way.
-For example, in a controlled experimental study \citet{zhao:2012}  notice that some people spend a lot of time to decide no matter whether the lineup is difficult or easy while some simply glance at lineups to make a decision.
-This influences the response of the observer and we see the subject specific variation in the power of visual test \citep{majumder:2012}.
+%\subsection{Observer Personality}
 
@@
 \subsection{Demographics of Observer}  Age, gender and location of the observer.
-\subsection{Skills of Observer} Knowledge about statistical plot as well as education level.
 
-\section{Simulation Experiment for Learning Effect}
+\subsection{Individual Performance of the Observer} Each person is different from others in some way.
+For example, in a controlled experimental study \citet{zhao:2012}  notice that some people spend a lot of time to decide no matter whether the lineup is difficult or easy while some simply glance at lineups to make a decision.
+This influences the response of the observer and we see the subject specific variation in the power of visual test \citep{majumder:2013}.
+
+\section{Simulation Experiment and Methods}\label{sec:exp_design}
+The performance of the observer while evaluating a lineup would depend on the factors described in Section \ref{sec:facotr_performance}.
+In this section we present the experimental design to control some  of them so that learning effect of observer and location effect of a plot in the lineup can be estimated.
 
@@
       \includegraphics[width=3in]{task_duration.pdf} 
-   \caption{Percentage of rejected tasks and duration in hour per 100 tasks for each of the 10 experiments.
+   \caption{Percentage of rejected tasks and duration of each experiment in hour per 100 tasks for each of the 10 experiments.
 Most of the tasks got rejected for box plot experiment.
@@
 
-\subsection{Estimation of sample size}  How the sample size is estimated
+%\subsection{Estimation of sample size}  How the sample size is estimated
 \subsection{Data collection methods}  Description of the simulation experiment giving focus on demographics and skills of experimental subjects.
@@
 In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time and cost while provided more convenience.
-\cite{majumder:2012} recruited people from MTurk for their simulation study in estimating the power of visual statistical inference.
+\cite{majumder:2013} recruited people from MTurk for their simulation study in estimating the power of visual statistical inference.
 They have done numerous pilot studies in lab before doing actual MTurk study and found similar results.
@@
 
-\section{Results}
+\subsection{Model to Estimate Learning Trend}
+
+\subsection{Model to Estimate Location Effect}
+
+\section{Results}\label{sec:result_socio}
 \subsection{Overview of the Data}
@@
 \subsection{How people pick the data plot} In the experiment setup we have a question for the turk observers to choose the reason for their selection of a particular data plot.
-It is shown by \cite{majumder:2012} that these choice reasons reveal the way people picks the data plot in the lineup.
+It is shown by \cite{majumder:2013} that these choice reasons reveal the way people picks the data plot in the lineup.
 To investigate this further we added free text input option for their choice reason instead of some fixed reasons to select from.
@@
 
-To accomplish this we fit generalized mixed effect model presented in \cite{majumder:2012} with lineups and subjects as random effect and no fixed effect.
+To accomplish this we fit generalized mixed effect model presented in \cite{majumder:2013} with lineups and subjects as random effect and no fixed effect.
 This gives the estimation of proportion correct responses  and we obtain residuals of the fitted model.

### 2013-05-29 R/sociology_paper.R

@@ -234,6 +234,10 @@ ggsave("../images/rejected_task.pdf", width=6, height=4)
 # write.csv(dat9$choice_reason[dat9$response_all==T], file="../data/dat9_reasoning_correct.csv")
 # write.csv(dat9$choice_reason[dat9$response_all==F], file="../data/dat9_reasoning_wrong.csv")
 
+reason.correct <- tolower(as.character(dat9$choice_reason[dat9$response_all==T]))
+gregexpr("\\W+", reason.correct)
+
+
 # =====================================================
 # Examining learning trend while giving feedback
 

### 2013-05-30 sociology_paper.tex
conclusion added

@@
 An observer is then asked to evaluate the lineup.
-If the observer can detect the observed plot in the lineup, the hypothesis is rejected.
+If the observer can detect the actual data plot in the lineup, the hypothesis is rejected.
 Figure \ref{fig:lineup_geno} displays a lineup of 20 plots where one of the plots is observed data.
@@
 For a lineup with fixed difficulty, the proportion correct responses may vary for different observers.
-If an observe evaluates multiple lineups, the proportion of correct evaluations by the observer could be used as a measure of performance by an observer.
+If an observer evaluates multiple lineups, the proportion of correct evaluations by the observer could be used as a measure of performance by an observer.
 The another way of measuring the performance of the observer could be the time taken by an observer to evaluate a lineup.
-Since performance would also depend on the lineup difficulty it is essential to adjust for this before estimating the observer performance.
+Since performance would also depend on the lineup difficulty it is essential to adjust for lineup difficulty before estimating the performance of the observer.
 
@@
 
-The power can be estimated from the proportion of correct evaluations. Thus power could be used as a measure of performance of the observer.
+The visual test statistic is chosen so that it displays a specific patter in case the null hypothesis not true.
+Thus the most important factor that help an observer to correctly evaluate a lineup is the presence of any detectable signal in the data.
+On the other hand, if the null hypothesis is not true, visual test statistic should not display any distinguishable pattern.
+In fact, some of the null plots may appear to be the most different plot in a lineup influencing the observer to chose a plot different than actual data plot.
+This is an elegant feature of the lineup.
+It force the observer to chose a wrong plot when the null hypothesis should not be rejected.
 
@@
 \section{Simulation Experiment and Methods}\label{sec:exp_design}
-The performance of the observer while evaluating a lineup would depend on the factors described in Section \ref{sec:facotr_performance}.
+The performance of the observer while evaluating a lineup would depend on the factors described in Section \ref{sec:factor_performance}.
 In this section we present the experimental design to control some  of them so that learning effect of observer and location effect of a plot in the lineup can be estimated.
@@
 \subsection{Experiment Setup} Description of the simulation experiment giving focus on demographics and skills of experimental subjects.
-
-\begin{table}[hbtp]
-\caption{Amazon mechanical turk experiments and their properties. Duration in hours per 100 tasks show the popularity of some tasks compared to others.}
-\centering
-\begin{tabular}{rlrrrrrrr}
-  \hline
-& Experiment& \multicolumn{2}{c}{ Total Task}& Average & \multicolumn{2}{c}{Duration (hour)} & Payment & Pay rate\\
-\cline{3-4} \cline{6-7}
-Serial & description & submitted & rejected & time(min) & Actual & 100 task& \$/task & \$/hour\\ 
-  \hline
-1 & Boxplot & 406 & 106 & 10.68 & 146.48 & 36.08 & 0.50 & 2.81 \\ 
-  2 & Scatterplot & 359 &   9 & 10.80 & 42.68 & 11.89 & 1.00 & 5.58 \\ 
-  3 & Contaminated plot & 219 &  19 & 13.53 & 126.17 & 57.61 & 1.00 & 2.22 \\ 
-  4 & Polar vs Cartesian & 110 &  10 & 20.65 & 11.65 & 10.59 & 1.00 & 2.91 \\ 
-  5 & Hist vs density & 234 &  37 & 17.85 & 41.57 & 17.76 & 1.00 & 3.36 \\ 
-  6 & Violine vs boxplot & 417 &  17 & 17.95 & 105.87 & 25.39 & 1.00 & 3.34 \\ 
-  7 & Group separation & 106 &   6 & 16.13 & 5.15 & 4.86 & 1.00 & 3.72 \\ 
-  8 & Sine Illusion & 101 &   1 & 16.52 & 78.38 & 77.60 & 1.00 & 3.63 \\ 
-  9 & Gene expression & 103 &   3 & 12.47 & 11.27 & 10.94 & 0.50 & 2.41 \\ 
-  10 & Test normality & 406 &   6 & 22.70 & 74.35 & 18.31 & 1.00 & 2.64 \\ 
-   \hline
-\end{tabular}
-\label{tbl:mturk}
-\end{table}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=3in]{rejected_task.pdf}
-      \includegraphics[width=3in]{task_duration.pdf} 
-   \caption{Percentage of rejected tasks and duration of each experiment in hour per 100 tasks for each of the 10 experiments.
-Most of the tasks got rejected for box plot experiment.
-Even though the sine illusion experiment took longest to finish the rejection rate is lowest for this experiment.}
-   \label{fig:task_duration}
-\end{figure}
 
@@
 
+As per the design of the experiment for learning trend we have two different response variables.
+One is observer feedback on a lineup which is a binary variable.
+The other variable is the time taken for each evaluation which is a continuous variable.
+Suppose random variable $Y_{\ell, i}$ denotes the response from subject $ i=1,2, ..., K$ on lineup $\ell=1,2, ..., L$.
+We model the mean $\mu_{\ell, i} = E(Y_{\ell, i})$ with a generalized liner mixed effect model
+
+\begin{equation} \label{eqn:trend}
+g( \mu_{\ell, i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
+\end{equation}
+
+\noindent where $g(.)$ denotes the link function.
+$W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters.
+Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well as the attempt made by the observer.
+$Z_{\ell i}$,  $1 \le i \le K$, $1 \le \ell \le L$,  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$; and
+$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$.
+$\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
+
+The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend} leads to the estimate of the subject and the lineup specific probability of successful evaluation by a single observer as 
+\begin{equation} \label{eqn:trend_power}
+\hat p_{\ell i} =  g^{-1}(W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}).
+\end{equation}
+
+For $Y_{l,i}$ being binary response we use {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ and $\hat p_{\ell i}$ gives the estimate of proportion correct by subject $i$ with lineup $\ell$.
+For continuous response  $Y_{l,i}$ we use gamma link function and $\hat p_{\ell i}$ gives the estimate of mean time taken by subject $i$ for lineup $\ell$.
+
 \subsection{Model to Estimate Location Effect}
@@
 \section{Results}\label{sec:result_socio}
+
 \subsection{Overview of the Data}
+
+\begin{table}[hbtp]
+\caption{Demographic information of the subjects participated the MTurk experiments.}
+\centering
+\begin{tabular}{rlrrr}
+  \hline
+ & levels & subsjects & avg\_time & response \\ 
+  \hline
+Gender & Male & 1348 & 48.51 & 13493 \\ 
+   & Female & 991 & 43.75 & 10564 \\ 
+\hline
+  Education & High school or less & 193 & 37.21 & 2241 \\ 
+   & Some under graduate courses & 418 & 42.84 & 4070 \\ 
+   & Under graduate degree & 584 & 44.29 & 5775 \\ 
+   & Some graduate courses & 245 & 43.43 & 2460 \\ 
+   & Graduate degree & 902 & 52.18 & 9511 \\ 
+\hline
+  Age & 18-25 & 740 & 42.97 & 7311 \\ 
+   & 26-30 & 547 & 46.27 & 5585 \\ 
+   & 31-35 & 376 & 44.27 & 3923 \\ 
+   & 36-40 & 257 & 55.03 & 2714 \\ 
+   & 41-45 & 141 & 43.90 & 1519 \\ 
+   & 46-50 &  95 & 49.29 & 1003 \\ 
+   & 51-55 &  83 & 48.67 & 867 \\ 
+   & 56-60 &  64 & 59.73 & 678 \\ 
+   & above 60 &  38 & 48.67 & 457 \\ 
+\hline   
+  Country  & United States & 1087 & 39.64 & 10769 \\  
+  & India & 980 & 52.63 & 10227 \\ 
+  & Rest of the world & 254 & 46.86 & 2819 \\
+
+   \hline
+\end{tabular}\label{tbl:demographics}
+\end{table}
+
 \subsection{Diversity of the Experimental Subjects} Amazon Mechanical Turk \cite{turk} web site is a good source for recruiting people to evaluation lineups.
@@
 
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=6in]{demographic_info.pdf} 
-   \caption{Countrywise distribution of age and academic levels of the MTurk workers participating the experiments shows the diversity of the subjects in all the demographic aspect.
-Almost equal number of male and female subjects participated the online experiments.}
-   \label{fig:demographic_info}
-\end{figure}
-
-\subsection{How people pick the data plot} In the experiment setup we have a question for the turk observers to choose the reason for their selection of a particular data plot.
-It is shown by \cite{majumder:2013} that these choice reasons reveal the way people picks the data plot in the lineup.
-To investigate this further we added free text input option for their choice reason instead of some fixed reasons to select from.
-In this experiment people could write whatever they think their reasons for choice are.
-Figure \ref{fig:wordle} shows the words used to explain their reasons for choice.
-The most common words used to explain their choice are points and green which indicates two important feature of a plot.
-One is the indicator of plot aesthetics and the other is the color.
-Spread, steepest, line and apart are some other important words used frequently.
-Spread and apart are indicative of variability in the data.
-Steepest and line indicates some sort of systematic pattern in the data.
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=3.5in]{reasoning_words.pdf} 
-   \caption{Words used to explain the reasons for selection of data plot in a lineup show what features of a lineup may help a non-statistician to evaluate it.
-Larger font indicate more people choosing that word.
-Different color is used just to separate the words.}
-   \label{fig:wordle}
-\end{figure}
-
-Figure \ref {fig:wordle} also shows some insight about peoples way of reading a plot.
-We notice that variability in the data, color and aesthetics used to generate plots and existence of any systematic pattern in the plot are some of the important features revealed from the figure.
-These features are commonly used by human brain to examine and compare plots.
-Notice that these features may be specific to this particular experiment.
-There could be different other features people would use to evaluate a lineup depending on the situation.
-But these words we have a general idea how people may think to evaluate a lineup.
-
-Turk workers are not necessarily trained on statistics or aware of specific terms used in statistics or statistical plots.
-Thus it is also interesting to note that how they explain things that have specific definitions and meaning.
-Notice in Figure \ref {fig:wordle} that some keywords like spread, apart should be analogous to larger variability while together,  close may be for indicating smaller variability.
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=4.5in]{participation_time.pdf} 
-   \caption{Time of the day when the participants work. Experiment 1 shows MTurk workers participated the experiments around the clock. Other experiments did not take a whole day to finish.}
-   \label{fig:participation_time}
-\end{figure}
-
 \subsection{Learning Trend of the Observer} Each subject participating the experiment has seen multiple number of lineups for evaluations.
@@
 But none of apparent trend are statistically significant.
-
-%covariates attempt number and $p$-value of the lineup.
-The $p$-value of the lineup measures some sort of difficulty and including this should adjust for the difficulty level of the lineup while estimating the probability of correct responses in each attempt.
-Figure \ref{fig:learning_trend} shows the fitted subject wise learning pattern as well as overall learning pattern for each of the experiments.
-The parameter estimates of the model are shown in Table \ref{tbl:model_par}.
 
@@
 
-\appendix
+With controlled simulation experiments, it is observed that the performance of human observers does not increase through successive attempts while evaluating multiple lineups.
+Even though the performance does not significantly improve, their skill in evaluating the lineup in shorter time gets improved over successive evaluations.
+When an observer evaluates multiple lineups, the earlier evaluations take significantly longer than the later evaluations.
+This result is very important as it suggests that the power estimated for visual inference using human subject experiment is robust and may not change if those participants are allowed to give feedback again.
+This investigation suggest that the skilled person may only do it faster.
+
+The another simulation experiment reveals that the effect of location of actual data plot in the lineup is not significant.
+This is important as the visual statistical inference procedure suggest that the data plot be placed at random anywhere in the lineup.
+This paper suggest that any random place in a lineup is as good as other places in the lineup.
+
+More experiment may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of statistical graphics.
+The numerous pilot studies done with the more trained participants suggest that power of visual statical inference is higher with skilled observer.
+
+This investigation allowed each observer to evaluate 10 lineups assuming that it would not cause fatigue or disinterest of the observers towards the experiment based on the pilot studies.
+The future research may involve doing more experiments to check if fewer or more than 10 lineups have any significant impact on performance of the observer.
+Also a lineup with fewer or larger than 10 plots may show different results.
+If the size of the lineup is much higher than 20, it is intuitive that there may be location effect of the data plot in the lineup.
+
+\section{Some figures and tables (for review)}
+
+\subsection{How people pick the data plot} In the experiment setup we have a question for the turk observers to choose the reason for their selection of a particular data plot.
+It is shown by \cite{majumder:2013} that these choice reasons reveal the way people picks the data plot in the lineup.
+To investigate this further we added free text input option for their choice reason instead of some fixed reasons to select from.
+In this experiment people could write whatever they think their reasons for choice are.
+Figure \ref{fig:wordle} shows the words used to explain their reasons for choice.
+The most common words used to explain their choice are points and green which indicates two important feature of a plot.
+One is the indicator of plot aesthetics and the other is the color.
+Spread, steepest, line and apart are some other important words used frequently.
+Spread and apart are indicative of variability in the data.
+Steepest and line indicates some sort of systematic pattern in the data.
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=3.5in]{reasoning_words.pdf} 
+   \caption{Words used to explain the reasons for selection of data plot in a lineup show what features of a lineup may help a non-statistician to evaluate it.
+Larger font indicate more people choosing that word.
+Different color is used just to separate the words.}
+   \label{fig:wordle}
+\end{figure}
+
+Figure \ref {fig:wordle} also shows some insight about peoples way of reading a plot.
+We notice that variability in the data, color and aesthetics used to generate plots and existence of any systematic pattern in the plot are some of the important features revealed from the figure.
+These features are commonly used by human brain to examine and compare plots.
+Notice that these features may be specific to this particular experiment.
+There could be different other features people would use to evaluate a lineup depending on the situation.
+But these words we have a general idea how people may think to evaluate a lineup.
+
+Turk workers are not necessarily trained on statistics or aware of specific terms used in statistics or statistical plots.
+Thus it is also interesting to note that how they explain things that have specific definitions and meaning.
+Notice in Figure \ref {fig:wordle} that some keywords like spread, apart should be analogous to larger variability while together,  close may be for indicating smaller variability.
 
 \begin{table}[hbtp]
-\caption{Demographic information of the subjects participated the MTurk experiments.}
+\caption{Amazon mechanical turk experiments and their properties. Duration in hours per 100 tasks show the popularity of some tasks compared to others.}
 \centering
-\begin{tabular}{rlrrr}
+\begin{tabular}{rlrrrrrrr}
   \hline
- & levels & subsjects & avg\_time & response \\ 
+& Experiment& \multicolumn{2}{c}{ Total Task}& Average & \multicolumn{2}{c}{Duration (hour)} & Payment & Pay rate\\
+\cline{3-4} \cline{6-7}
+Serial & description & submitted & rejected & time(min) & Actual & 100 task& \$/task & \$/hour\\ 
   \hline
-Gender & Male & 1348 & 48.51 & 13493 \\ 
-   & Female & 991 & 43.75 & 10564 \\ 
+1 & Boxplot & 406 & 106 & 10.68 & 146.48 & 36.08 & 0.50 & 2.81 \\ 
+  2 & Scatterplot & 359 &   9 & 10.80 & 42.68 & 11.89 & 1.00 & 5.58 \\ 
+  3 & Contaminated plot & 219 &  19 & 13.53 & 126.17 & 57.61 & 1.00 & 2.22 \\ 
+  4 & Polar vs Cartesian & 110 &  10 & 20.65 & 11.65 & 10.59 & 1.00 & 2.91 \\ 
+  5 & Hist vs density & 234 &  37 & 17.85 & 41.57 & 17.76 & 1.00 & 3.36 \\ 
+  6 & Violine vs boxplot & 417 &  17 & 17.95 & 105.87 & 25.39 & 1.00 & 3.34 \\ 
+  7 & Group separation & 106 &   6 & 16.13 & 5.15 & 4.86 & 1.00 & 3.72 \\ 
+  8 & Sine Illusion & 101 &   1 & 16.52 & 78.38 & 77.60 & 1.00 & 3.63 \\ 
+  9 & Gene expression & 103 &   3 & 12.47 & 11.27 & 10.94 & 0.50 & 2.41 \\ 
+  10 & Test normality & 406 &   6 & 22.70 & 74.35 & 18.31 & 1.00 & 2.64 \\ 
    \hline
-  Education & High school or less & 193 & 37.21 & 2241 \\ 
-   & Some under graduate courses & 418 & 42.84 & 4070 \\ 
-   & Under graduate degree & 584 & 44.29 & 5775 \\ 
-   & Some graduate courses & 245 & 43.43 & 2460 \\ 
-   & Graduate degree & 902 & 52.18 & 9511 \\ 
-\hline
-  Age & 18-25 & 740 & 42.97 & 7311 \\ 
-   & 26-30 & 547 & 46.27 & 5585 \\ 
-   & 31-35 & 376 & 44.27 & 3923 \\ 
-   & 36-40 & 257 & 55.03 & 2714 \\ 
-   & 41-45 & 141 & 43.90 & 1519 \\ 
-   & 46-50 &  95 & 49.29 & 1003 \\ 
-   & 51-55 &  83 & 48.67 & 867 \\ 
-   & 56-60 &  64 & 59.73 & 678 \\ 
-   & above 60 &  38 & 48.67 & 457 \\ 
-   \hline
-\end{tabular}\label{tbl:demographics}
+\end{tabular}
+\label{tbl:mturk}
 \end{table}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=3in]{rejected_task.pdf}
+      \includegraphics[width=3in]{task_duration.pdf} 
+   \caption{Percentage of rejected tasks and duration of each experiment in hour per 100 tasks for each of the 10 experiments.
+Most of the tasks got rejected for box plot experiment.
+Even though the sine illusion experiment took longest to finish the rejection rate is lowest for this experiment.}
+   \label{fig:task_duration}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=6in]{demographic_info.pdf} 
+   \caption{Countrywise distribution of age and academic levels of the MTurk workers participating the experiments shows the diversity of the subjects in all the demographic aspect.
+Almost equal number of male and female subjects participated the online experiments.}
+   \label{fig:demographic_info}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=6in]{turker_location_experiment.pdf} 
+   \caption{World maps showing where the participants are coming from for all the 10 experiments.}
+   \label{fig:turker_location_experiment}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=4.5in]{participation_time.pdf} 
+   \caption{Time of the day when the participants work. Experiment 1 shows MTurk workers participated the experiments around the clock. Other experiments did not take a whole day to finish.}
+   \label{fig:participation_time}
+\end{figure}
 

### 2013-05-30 R/sociology_paper.R

@@ -127,20 +127,21 @@ for (i in 1:10){
 # obtained from www.ipaddressapi.com using real ip address
 # based on ip, it gets longitude, latitude, country, city etc.
 demographics <- merge(demographics,ip.details,all.x=TRUE, by="ip_address")
+demographics$country <- as.character(demographics$country_name)
+demographics$country[(demographics$country_code != "IN") & (demographics$country_code != "US")] <- "Rest of the world"
+demographics$country[demographics$country=="Namibia"] <- "Rest of the world"
+
+
+
 
 # getting unique participants for plotting purpose only
 # Since each participant has multiple responses in demographics data
 turker <- demographics[!duplicated(demographics$id),]
-turker$country <- as.character(turker$country_name)
-turker$country[(turker$country_code != "IN") & (turker$country_code != "US")] <- "Rest of the world"
-turker$country[turker$country=="Namibia"] <- "Rest of the world"
 turker$study <- turker$academic_study
 turker$study[turker$age_level=="36-40"] <- c("High school or less (1)","Some under grad course (2)",
                                              "Under graduate degree (3)", "Some graduate courses (4)",
                                              "Graduate degree (5)")[turker$study[turker$age_level=="36-40"]]
-
-
-
+turker$exp <- factor(turker$experiment, levels=names(table(turker$experiment))[order(c(1,10,2:9))])
 levels(turker$age_level)[7:9] <- 7
 levels(turker$age_level)[1:7] <- c("18-25","26-30","31-35","36-40",
                                 "41-45", "46-50", "above 50")
@@ -172,7 +173,6 @@ dev.off()
 
 # Time of the work
 turker$hours <- as.factor(hour(as.POSIXlt(turker$start_time)))
-turker$exp <- factor(turker$experiment, levels=names(table(turker$experiment))[order(c(1,10,2:9))])
 qplot(hours, data=turker) +
   facet_grid(exp~.,scales="free_y") +
   xlab("Hour of the day") + ylab("Number of subjects")
@@ -192,8 +192,9 @@ get_summary <- function(dat, var){
 sg <- get_summary(demographics, "gender")
 se <- get_summary(demographics, "degree")
 sa <- get_summary(demographics, "age_level")
+sl <- get_summary(demographics, "country")
 
-sdat <- rbind(sg,se,sa)
+sdat <- rbind(sg,se,sa,sl)
[…]

### 2013-05-30 sociology_paper.tex
brief description of experimental design

@@
 Later these concepts are further developed by defining necessary terminologies and validating using a head to head comparison with conventional inference by \citet{majumder:2013}.
-The power of the visual test is obtained and it is observed that the power is as good as conventional test while in some scenarios, the power exceeds the conventional test power.
+The power of the visual test is obtained and it is observed that the power is as good as that of conventional test while in some scenarios, the power exceeds the conventional test power.
 
-The test statistic used in visual inference is a plot of the data.
-This plot is then placed on a layout of plot called lineup.
+The test statistic used in visual inference is a plot of the observed data.
+This plot, called actual data plot, is then placed on a layout of plot called lineup.
 The rest of the plots in a lineup are generated from the model specified by the null hypothesis.
@@
 This paper is organized as follows; Section \ref{sec:factor_performance} describes some of the factors that may affect the performance of the observer.
-Section \ref{sec:exp_design} describes the design of the simulation experiments and methods to estimate the learning trend of the observer and location effect of the observed plot in a lineup.
+Section \ref{sec:exp_design} describes the design of the simulation experiments and methods to estimate the learning trend of the observer and location effect of the actual plot in a lineup.
 Section \ref{sec:result_socio} presents the analysis of the experimental data and the results.
@@
 
-\subsection{Choice of Visual Test Statistic} Visual test statistics as defined in \citep{majumder:2013} serve two main purposes.
-One is to display a prominent pattern that should direct the human observers to select the actual plot in the lineup when the null hypothesis is not true.
-The another feature is that it should not show any strange pattern when null hypothesis is true, i.e., it should show similar pattern that null plots may have.
-Thus a visual test statistic is highly associated with the hypothesis.
-To achieve these purposes it is very important to decide which plot type and plot features should be adopted.
-As discussed in \citep{majumder:2013} following grammar of graphics developed  by \citet{wilkinson:1999} and later improved by \citet{hadley:2009} can give the task a standard control.
-
+\subsection{Choice of Visual Test Statistic}  The visual test statistic should be highly associated with the hypothesis under investigation.
+To achieve this purposes it is very important to decide which plot type and plot features should be adopted.
+In a linear regression setting, the visual test statistics are presented in \citep{majumder:2013} for some common hypothesis testing.
+It is also observed that a scatterplot may do a better job than a box plot when using as a visual test statistic for regression parameters.
 Some of the effective features of visual test statistics are discussed in \citep{heike:2012} including plot type, color and shape of the plots.
-It is also observed that a scatterplot may do a better job than a box plot when using as a visual test statistic for regression parameters \citep{majumder:2013} .
-\citep{niladri:2012} presents some \textit{distance measure} to determine how a plot may be different from another.
+\cite{niladri:2012} presents some \textit{distance measure} to determine how a plot may be different from each other.
 
@@
 
-\subsection{Question that Human Observer Answers} When a observer is prepared to evaluate a lineup, there needs to be a guideline what exactly should be looked for in a lineup.
-The researcher knows about the hypothesis but the observer does not know the underlying details of the lineup.
-Thus the researcher needs to ask a question observer is to answer while evaluating the lineup.
+\subsection{Question that Human Observer Answers} The researcher knows about the underlying hypothesis but the observer does not necessarily know the underlying details of the lineup.
+So, the researcher needs to ask a question observer is to answer while evaluating the lineup.
 This question should provide the observer a little clue so that the answer reflects the hypothesized patterns in the actual plot.
@@
 If the actual plot is on the top left corner and the observer starts from that point it may be easier to detect the actual plot earlier in the exercise and the observer could get plenty of time to make comparison.
-Thus the position of the actual plot in a lineup has some impact.
+Thus the position of the actual plot in a lineup may have some impact.
 
-\subsection{Demographics of Observer}  Age, gender and location of the observer.
+\subsection{Demographics of Observer}  Age, gender and geographical location of the observer.
 
@@
 For example, in a controlled experimental study \citet{zhao:2012}  notice that some people spend a lot of time to decide no matter whether the lineup is difficult or easy while some simply glance at lineups to make a decision.
-This influences the response of the observer and we see the subject specific variation in the power of visual test \citep{majumder:2013}.
+This influences the response of the observer.
+Also subject specific variation in the power of visual test is observed in \citep{majumder:2013}.
 
@@
 
-\subsection{Experiment Setup} Description of the simulation experiment giving focus on demographics and skills of experimental subjects.
+\subsection{Experiment Setup}  Multiple simulation experiments are designed to investigate the learning trend of the observer and the location effect of the actual data in the lineup.
+These are explained in tis section.
 
-%\subsection{Estimation of sample size}  How the sample size is estimated
-\subsection{Data collection methods}  Description of the simulation experiment giving focus on demographics and skills of experimental subjects.
-\subsection{Amazon Mechanical Turk}
+\paragraph{Design for Learning Trend:} Learning trend of an observer can be observed in terms of performance over successive feedback when multiple lineups are shown for evaluation.
+Each subject is shown a total of 10 lineups.
+The lineups are not necessarily with the same difficulty levels.
+But the order of lineups is randomized.
+The responses of the lineups are recorded by attempt 1 through 10.
+Attempt 1 means the response is for the first lineup and attempt 10 means the response is for the 10th lineup.
+The goal is to estimate whether performance of the observer improves from attempt 1 to attempt 10.
 
-Amazon Mechanical Turk \cite{turk} or MTurk is an online work place where people from around the world can perform some tasks and get paid.
+\paragraph{Design for Location effect:} The experiment is set up based on the gene expression data with two groups.
+The groups are defined based on the presence or absence of iron sufficiency.
+The null hypothesis is that there is no difference between this two groups.
+Null plots are obtained by randomizing the group structure of the data.
+The actual plot is then randomly placed in five different locations.
+For each location, five different sets of null plots are used to produce five different lineups.
+This produces 25 different lineups.
+One such lineup is shown in Figure \ref{fig:lineup_geno}.Similar procedure is applied to obtain 25 lineups to test the interaction between Genotype and Empty Vector (EV).
+
+Each observer is shown three lineups.
+First lineups shown is a randomly selected Interaction lineup, then one from randomly selected Genotype lineup.
+Finally a test lineup is shown to screen out unusual evaluation.
+The subject is not aware which one is the test plot but it is informed in the starting of the task about having a test plot.
+The test plot is very easy and everyone should detect the data plot.
+The MTurk worker is paid based on whether he or she could correctly detect the data plot.
+But the data on test plot are excluded from the analysis.
+
+\subsection{Data collection methods}  The human observers who evaluate the experimental lineups are recruited through Amazon Mechanical Turk \cite{turk} or MTurk Web site.
+It is an online work place where people from around the world can perform some tasks and get paid.
 Usually tasks are very simple and no specialized training is required.
@@
 \cite{mason:2012} explains various features of MTurk and describes how it can be used as part of human behavioral study.
+
+We design and develop a web site which enables to display the lineups to the observers as per experimental design.
+The MTurk workers are redirect to the web site and the data are collected and automatically stored into a local server.
+Demographic informations such as age group, gender and education levels were collected.
+The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback is received.
+The location of the observer is determined by the ip address of the observer.
 
@@
 
+Since the actual data plot is same for each of the null sets of plot, the response data constitutes a multivariate response.
+Thus, to examine if the difference in proportion correct among the locations is statistically significant we fit a one way multi variate analysis of variance (MANOVA) model to the data.
+
+Suppose $Y=(Y_1,Y_2, ...
+, Y_p)$ is a vector of random variable with dimension $p=5$ representing the response for five null sets.
+let $Y_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., 5$.
+We fit the following MANOVA model
+\begin{equation}\label{manova}
+Y_{ij} = \mu_{i} + \epsilon_{ij}
+\end{equation}
+where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$.
+
 \section{Results}\label{sec:result_socio}
 
-\subsection{Overview of the Data}
+\subsection{Overview of the Data} Figure \ref{fig:turker_location} shows the location from where we received data.
+Notice that we received data from around the world and we observed diversity of location in the collected data.
+We notice diversity in not only location of the observers but also their gender, age groups and education levels.
+Table \ref{tbl:demographics} shows that number of male and female participants are almost same.
 
@@
 
-\subsection{Diversity of the Experimental Subjects} Amazon Mechanical Turk \cite{turk} web site is a good source for recruiting people to evaluation lineups.
-It is a source of diverse workers who can give us feedback on visual inspection.
-Figure \ref{fig:turker_location} shows the location from where we received data.
-Notice that we received data from around the world and we observed diversity of location in the collected data.
-We notice diversity in not only location of the observers but also their gender, age groups and education levels.
-Table \ref{tbl:demographics} shows that number of male and female participants are almost same.
-
 \begin{figure}[htbp] 
@@
 
-To accomplish this we fit generalized mixed effect model presented in \cite{majumder:2013} with lineups and subjects as random effect and no fixed effect.
+To accomplish this we fit generalized mixed effect model \eqref{eqn:trend} with lineups and subjects as random effect and no fixed effect.
 This gives the estimation of proportion correct responses  and we obtain residuals of the fitted model.
@@
 The overall proportion of correct response does not seem to vary much for different location indicating the non-existence of location effect.
-Since the actual data plot is same for each of the null sets of plot, the response data constitutes a multivariate response.
-Thus, to examine if the difference in proportion correct among the locations is statistically significant we fit a one way multi variate analysis of variance (MANOVA) model to the data.
-
- Suppose $Y=(Y_1,Y_2, ...
-, Y_p)$ is a vector of random variable with dimension $p=5$ representing the response for five null sets.
-let $Y_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., 5$.
-We fit the following MANOVA model
-\begin{equation}\label{manova}
-Y_{ij} = \mu_{i} + \epsilon_{ij}
-\end{equation}
-where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$.
-
-Model \eqref{manova} enables to test if the mean vectors are similar for different locations.
+Model \eqref{manova} is used to test if the mean vectors are similar for different locations.
 To fit the model we use anova function of stats package of \cite{R:2012}.
@@
 With controlled simulation experiments, it is observed that the performance of human observers does not increase through successive attempts while evaluating multiple lineups.
-Even though the performance does not significantly improve, their skill in evaluating the lineup in shorter time gets improved over successive evaluations.
+But their skill in evaluating the lineup in shorter time gets improved over successive evaluations.
 When an observer evaluates multiple lineups, the earlier evaluations take significantly longer than the later evaluations.
@@
 The another simulation experiment reveals that the effect of location of actual data plot in the lineup is not significant.
-This is important as the visual statistical inference procedure suggest that the data plot be placed at random anywhere in the lineup.
-This paper suggest that any random place in a lineup is as good as other places in the lineup.
+This is important as the visual statistical inference procedure suggests that the data plot be placed at random anywhere in the lineup.
+This paper suggests that any random place in a lineup is as good as other places in the lineup.
 
-More experiment may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of statistical graphics.
+More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of statistical graphics.
 The numerous pilot studies done with the more trained participants suggest that power of visual statical inference is higher with skilled observer.

### 2013-05-30 sociology_paper.tex
link for gamma response updated

@@
 
-For $Y_{l,i}$ being binary response we use {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ and $\hat p_{\ell i}$ gives the estimate of proportion correct by subject $i$ with lineup $\ell$.
-For continuous response  $Y_{l,i}$ we use gamma link function and $\hat p_{\ell i}$ gives the estimate of mean time taken by subject $i$ for lineup $\ell$.
+For $Y_{l,i}$ being binary response we use {\it logit} link function $g(\mu)=g_1(\mu)=\log(\mu) - \log(1-\mu); 0 \le \mu \le 1$ and $\hat p_{\ell i}$ gives the estimate of proportion correct by subject $i$ with lineup $\ell$.
+The response time taken is a positive random variable with a right skewed distribution.
+For this the usual convention is to model a gamma distribution with time.
+Thus we used inverse link function $g(\mu)=g_2(\mu)=1/\mu;  \mu >0$ for gamma response $Y_{l,i}$ and $\hat p_{\ell i}$ gives the estimate of mean time taken by subject $i$ for lineup $\ell$.
 

### 2013-05-30 R/sociology_paper.R

@@ -244,7 +244,7 @@ p<- ggplot() +
   geom_point(aes(longitude,latitude, colour=factor("A")), data=turker, alpha=.6) +
   facet_wrap(~exp, ncol=2)+ theme_bw()+
   theme(legend.position="none", axis.text = element_blank(), axis.title=element_blank()) 
-ggsave(p,"../images/turker_location_experiment.pdf", width=8, height=10)
+ggsave("../images/turker_location_experiment.pdf", width=8, height=10)
 
 # country-wise participants counts
 qplot(factor(country, levels=names(table(country))[order(table(country))]), 
@@ -319,6 +319,14 @@ qplot(attempt,mean_resid, data= dmt) + geom_point(size=2.5) +
   scale_x_continuous(breaks = seq(2,10,by=2)) 
 
 ggsave("../images/learning_trend_time.pdf", width=10.5, height = 3.5)
+
+
+# Fitting generalized mixed effect model with time taken
+# Link function is inverse link (1/mu) since response is gamma
+
+qplot(time_taken, geom="histogram", data=subset(dd, time_taken<500), binwidth=10)
+
+qplot(time_taken, geom="density", data=subset(dpt, time_taken<300), color=factor(experiment))
 
 
 # Checking if the performance increases with attempts

### 2013-06-02 R/polls.r
data and code for intro example

@@ -1,0 +1,155 @@
+# url <- "http://freedomslighthouse.net/2012-presidential-election-electoral-vote-map/"
+# library(XML)
+# tabs <- readHTMLTable(url)
+# polls <- tabs[[3]]
+# 
+# write.csv(polls, "polls.csv", row.names=F)
+polls <- read.csv("data/polls.csv")
+names(polls) <- c("State", "Electoral.Votes", "Results2008", "Poll2012", "Results2012")
+
+library(plyr)
+plist <- strsplit(as.character(polls$Poll2012), " ", fixed=TRUE)
+p2012 <- ldply(plist, function(x) c(x,rep(NA, 5))[1:5])
+names(p2012) <- paste(c("Candidate","Poll","Major", "Diss", "Minor"), "2012", sep="")
+p2012$Poll2012 <- as.numeric(gsub("+","", as.character(p2012$Poll2012)))
+p2012$Democrat <- p2012$Candidate2012=="Obama"
+
+rlist <- strsplit(as.character(polls$Results2008), " ", fixed=TRUE)
+r2008 <- ldply(rlist, function(x) c(x,rep(NA, 2))[1:2])
+names(r2008) <- c("Candidate08", "Result08")
+r2008$Result08 <- as.numeric(gsub("+|%","", as.character(r2008$Result08)))
+r2008$Democrat <- r2008$Candidate08=="Obama"
+
+r12list <- strsplit(as.character(polls$Results2012), " ", fixed=TRUE)
+r2012 <- ldply(r12list, function(x) c(x,rep(NA, 2))[1:5])
+names(r2012) <- paste(c("Candidate","Result","Major", "Diss", "Minor"), "2012", sep="")
+r2012$Result2012 <- as.numeric(gsub("+","", as.character(r2012$Result2012)))
+r2012$Democrat <- r2012$Candidate=="Obama"
+
+simulatePoll <- function(x, sd=3, backup) {
+  suppressWarnings(res <- rnorm(length(x), mean=x, sd=sd))
+  idx <- which(is.na(res))
+  res[idx] <- rnorm(length(idx), mean=backup[idx], sd=sd)
+  res
+}
+
+simDemocrat <- function(margin, party, backup) {
+  democrat <- party 
+  idx <- which(is.na(democrat))
+  democrat[idx] <- backup[idx]
+  idx <- which(margin < 0)
+  democrat[idx] <- !democrat[idx]
+  
+  democrat
+}
+
+res <- replicate(1000, {
[…]

### 2013-06-02 sociology_paper.tex
irb and nsf note
tower example

@@
 Later these concepts are further developed by defining necessary terminologies and validating using a head to head comparison with conventional inference by \citet{majumder:2013}.
-The power of the visual test is obtained and it is observed that the power is as good as that of conventional test while in some scenarios, the power exceeds the conventional test power.
+The power of the visual test is obtained and it is observed that the power is as good as that of conventional test while in some scenarios, visual power exceeds that of conventional tests.
 
@@
 If the observer can detect the actual data plot in the lineup, the hypothesis is rejected.
-Figure \ref{fig:lineup_geno} displays a lineup of 20 plots where one of the plots is observed data.
-The rest of the 19 plots are generated from the assumption that there is no difference between two groups of data.
+Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are rendered from data generated under a null model.
+Which of these towers is the most extreme?
+Which side is higher?
+In an experiment XXX of XXX observers picked plot \# XXX as the most extreme plot from the lineup, and all agreed that the blue side was higher.
+This gives us a visual $p$-value of XXX, indicating that we have enough evidence to reject the null hypothesis. 
+What does this mean, though?
+For that, we need to know the context of the data and we need to have more information about the generation of the null plots: in this example, we are investigating, whether the results from the 2012 US presidential election is consistent with polling results.
+As a null hypothesis we  assume that  the election results are consistent with polling results.
+The polling results give us a null model to generate data.
+Each poll gives us a margin between candidates.
+This is what we use as the mean of a normal distribution.
+In the absence of information of error for a poll, we will assume a standard deviation of 3 percentage points for all of the polls.
+While this is a severe simplification, the resulting margin of error of $\pm 6$ percentage points is realistic, and could be easily replaced if additional information on a poll became available.
+A null data set is then generate as a set of draws from normal distributions based on the latest poll results.
+Each panel in figure~\ref{fig:tower} shows an `electoral building' \citet{nytimes}: each state is represented by a rectangle.
+The width of a rectangle is determined by the difference between (polling) results, its height is given by the number of electoral votes, and fill color shows party affiliation (blue for Democrat, and red for Republican).
+Rectangles are placed on top of each other, separated by affiliation and sorted by margin between polls.
+A party needs 270 electoral votes or more to win the presidential election.
+
+%The rest of the 19 plots are generated from the assumption that there is no difference between two groups of data.
 For this, the group structure of the data is broken by random permutation of the groups.
+
 If the null hypothesis is true, the observed data should look similar to the rest of the plots in the lineup making it difficult to detect the observed plot.
-On the other hand if the null hypothesis is not true, observed data should be different from the rest of the plots in the lineup.
+On the other hand, if the null hypothesis is not true, the observed data should be different from the rest of the plots in the lineup and observers should be able to pick the data plot out from the rest.
 
@@
 
+%\begin{figure}[htbp] 
+%   \centering
+%   \includegraphics[width=6.5in]{plot_turk9_geno_4_3.png} 
+%   \caption{A lineup of 20 plots. One of the plots is observed data and rest of the 19 plots are obtained by randomly permuting the group structure. Can you identify the observed data?}
+%   \label{fig:lineup_geno}
+%\end{figure}
+
 \begin{figure}[htbp] 
    \centering
-   \includegraphics[width=6.5in]{plot_turk9_geno_4_3.png} 
-   \caption{A lineup of 20 plots. One of the plots is observed data and rest of the 19 plots are obtained by randomly permuting the group structure. Can you identify the observed data?}
-   \label{fig:lineup_geno}
+   \includegraphics[width=0.9\textwidth]{tower.pdf} 
+   \caption{ \label{fig:tower} Which of these towers is the most extreme? Which side is higher?}
 \end{figure}
@@
 
+%------------------------------------------------------------------------------------
+\paragraph{Acknowledgments}
+%------------------------------------------------------------------------------------
+This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the internal review board IRB 10-347.
+
 \section{Conclusion}

### 2013-06-02 R/polls.r
fixed blue left, red right

@@ -4,7 +4,7 @@
 # polls <- tabs[[3]]
 # 
 # write.csv(polls, "polls.csv", row.names=F)
-polls <- read.csv("data/polls.csv")
+polls <- read.csv("../data/polls.csv")
 names(polls) <- c("State", "Electoral.Votes", "Results2008", "Poll2012", "Results2012")
 
 library(plyr)
@@ -72,7 +72,7 @@ location <- p3$sample[nrow(p3)]
 p2 <- p3
 p2 <- p2[order(p2$Sim.Margin, decreasing=TRUE),]
 p2 <- ddply(p2, .(.n, Sim.Democrat), transform, tower=cumsum(Electoral.Votes[order(Sim.Margin, decreasing=TRUE)]))
-p2$diff <- with(p2, Sim.Margin*c(-1,1)[as.numeric(Sim.Democrat)+1])
+p2$diff <- with(p2, Sim.Margin*c(1,-1)[as.numeric(Sim.Democrat)+1])
 
 write.csv(p2, file=sprintf("electoral/electoral-%s-%s.csv", null, location))
 }
@@ -139,16 +139,16 @@ for (fname in files) {
 
 
 
-ggplot(aes(x=diff, y=tower, colour = factor(Sim.Democrat)), data=p2) + 
+ggplot(aes(x=diff, y=tower, colour = factor(Sim.Democrat)), data=dframe) + 
 #  geom_point() +
-  geom_segment(aes(x=diff, xend=diff, y=0, yend=tower, colour=Sim.Democrat), size=1) +
+  geom_segment(aes(x=diff, xend=diff, y=0, yend=tower, colour=Sim.Democrat), size=0) +
   scale_colour_manual(values=c("red", "blue"), guide="none") + 
   scale_fill_manual(values=c("red", "blue"), guide="none") + 
   geom_hline(yintercept=270, colour="grey70") + 
-  ylab("Electoral Votes")+xlab("Republican       ---          Democrat") + 
+  ylab("Electoral Votes")+xlab("Democrat      ---          Republican") + 
   scale_x_continuous(breaks=c(-50,0,50), labels=c("50", "0", "50"), limits=c(-90,90)) + 
   geom_rect(aes(xmin=0, xmax=diff, ymin=0, ymax=tower, fill=Sim.Democrat), size=0) +
-  geom_segment(aes(x=0,xend=diff, y=tower,yend=tower), colour="grey80", size=0.5) + 
[…]

### 2013-06-02 R/polls.r
small update to tower and code

@@ -26,7 +26,7 @@ names(r2012) <- paste(c("Candidate","Result","Major", "Diss", "Minor"), "2012", 
 r2012$Result2012 <- as.numeric(gsub("+","", as.character(r2012$Result2012)))
 r2012$Democrat <- r2012$Candidate=="Obama"
 
-simulatePoll <- function(x, sd=3, backup) {
+simulatePoll <- function(x, sd=2.5, backup) {
   suppressWarnings(res <- rnorm(length(x), mean=x, sd=sd))
   idx <- which(is.na(res))
   res[idx] <- rnorm(length(idx), mean=backup[idx], sd=sd)
@@ -138,7 +138,8 @@ for (fname in files) {
 
 
 
-
+fname <- "electoral-4-2.csv"
+  dframe <- read.csv(sprintf("../electoral/%s", fname))
 ggplot(aes(x=diff, y=tower, colour = factor(Sim.Democrat)), data=dframe) + 
 #  geom_point() +
   geom_segment(aes(x=diff, xend=diff, y=0, yend=tower, colour=Sim.Democrat), size=0) +

### 2013-06-02 sociology_paper.tex
factors affecting the peroformance updated

--- 2013-05-30 sociology_paper.tex
@@
 % Article top matter
-\title{Impact of Demographics and Skills of the Observer on Visual Statistical Inference }
+\title{Human Factors Influencing Visual Statistical Inference }
 \author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is a PhD student (e-mail: mahbub72@gmail.com) , Heike Hofmann is an Associate  Professor and Dianne Cook is a Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
+\thanks{Mahbubul Majumder is a PhD student (e-mail: mahbub72@gmail.com), Heike Hofmann is an Associate  Professor and Dianne Cook is a Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
 This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
@@
 
-\subsection{Visual Perception of Human Eye} Different people observe the lineup in different ways.
+%\subsection{Visual Perception of Human Eye} Different people observe the lineup in different ways.
 With the help of eye-tracking equipment \citet{zhao:2012} track the observers' eyes to see how they go through the plots in a lineup to come to their answers.
@@
 Beside right-left or up-down directions observer show some diagonal movement too.
-
-Given this pattern of human eye tracks, it may effect the performance of an observer depending on where the actual plot is placed in the lineup.
+%
+%Given this pattern of human eye tracks, it may effect the performance of an observer depending on where the actual plot is placed in the lineup.
 If the actual plot is on the top left corner and the observer starts from that point it may be easier to detect the actual plot earlier in the exercise and the observer could get plenty of time to make comparison.
@@
 \subsection{Demographics of Observer}  Age, gender and geographical location of the observer.
+
+\subsection{Learning Trend of the Observer} When an observer evaluates a lineup he or she may learn something from the experience.
+If the same observer is shown another lineup the learning from previous evaluation may help.
+The more evaluations an observer makes the more skillful the observer may become.
+This learning trend may or may not be significant overall but it could influence the performance of the observer in some way.
+This paper investigates the learning trend with multiple simulation experiments.
+
+\subsection{Location of Actual Plot in the Lineup} The actual data plot is placed in a random spot in a lineup.
+While evaluating the lineup some people may start looking from some specific part of the lineup.
+With the help of eye-tracking equipment \citet{zhao:2012} track the observers' eyes to see how they go through the plots in a lineup to come to their answers.
+The results suggest that people have particular methods of reading lineups.
+Some people read lineups from left to write direction while some read from upward to downward.
+Some people start looking from the center of the lineup while others start from the top left corner.
+In the earlier phase of the exercise, the observer tend to scan the plots and then start comparing plots to make a final decision.
+Beside right-left or up-down directions observer show some diagonal movement too.
+
+Given a specific pattern of eye movement of the observer in examining the lineup, the location of actual plot in the lineup may have some effect.
+Those who start exploring from the left top corner may get first glance of the actual plot if it is on that location.
+This should give the observer more time to compare it with the rest of the null plots.
+Those who start looking from the center may then scan towards right or left direction and thus don't have the chance to scan the null plots continuously as they could if they would start from a corner.
+Thus they should scan the center plot over and over again while scanning the left or right side plots.
+If the data plot is in the center location it may have multiple chance to be examined and hence be identified.
+
+It may be possible that some part of the lineup is little explored or not scanned by the observer if he or she feels that the actual plot is found before even seriously scanning the whole lineup.
+In those scenarios the location where the observer first start scanning is important.
+Placing the actual data plot in that location may yield different results.
+
+\subsection{Selection of Null Plots} A lineup becomes difficult to evaluate if one of the null plots appears to be very similar to the actual plot.
+When null hypothesis is true it is a common scenario.
+With alternative hypothesis being true,  it may happen if  we compare actual plot with many null plots.
+But we have only specific number of null plots in a lineup.
+So, a different set of null plots may yield some variation in the difficulty of the lineup with the same actual plot.
 
@@
 This produces 25 different lineups.
-One such lineup is shown in Figure \ref{fig:lineup_geno}.Similar procedure is applied to obtain 25 lineups to test the interaction between Genotype and Empty Vector (EV).
+One such lineup is shown in Figure \ref{fig:lineup_geno}.
+Similar procedure is applied to obtain 25 lineups to test the interaction between Genotype and Empty Vector (EV).
 
@@
 
-\subsection{Learning Trend of the Observer} Each subject participating the experiment has seen multiple number of lineups for evaluations.
+\subsection{Learning Trend} Each subject participating the experiment has seen multiple number of lineups for evaluations.
 Suppose $K$ be the number of lineups evaluated by a subject and $A_k$ be the $kth$ attempt, $k=1,2, ..., K$.
@@
 
-\subsection{Location Effect of Data Plot in the Lineup}
+\subsection{Location Effect}
 
@@
 
-The another simulation experiment reveals that the effect of location of actual data plot in the lineup is not significant.
+The simulation experiment reveals that the effect of location of actual data plot in the lineup is not significant.
 This is important as the visual statistical inference procedure suggests that the data plot be placed at random anywhere in the lineup.
+Even though there are variations on the performance depending on actual plot locations, it is not statistically significant to be of any concern.
 This paper suggests that any random place in a lineup is as good as other places in the lineup.
@@
 More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of statistical graphics.
-The numerous pilot studies done with the more trained participants suggest that power of visual statical inference is higher with skilled observer.
+The numerous pilot studies done with the more trained participants suggest that power of visual statical inference may be higher with skilled observer.
+Thus the variations observed with individual performances could be attributed to individual abilities.
 
@@
 The future research may involve doing more experiments to check if fewer or more than 10 lineups have any significant impact on performance of the observer.
-Also a lineup with fewer or larger than 10 plots may show different results.
+
+Also, a lineup with fewer or larger than 20 plots may yield different results.
 If the size of the lineup is much higher than 20, it is intuitive that there may be location effect of the data plot in the lineup.
+It is because, the observer may get tiered of scanning and make decision based on the partial scanning of the lineup.
+On the other hand fewer than 20 plots would give observer compare the actual plot with fewer null plots give more chance of picking actual plot as an error not as an actual plot.
+These are the some of the issues that needs to investigated.
 

### 2013-06-02 R/sociology_paper.R

@@ -184,10 +184,13 @@ get_summary <- function(dat, var){
   res <- ddply(dat,c(var), summarize,
                subsjects = length(unique(id)),
                avg_time = round(mean(time_taken),2),
-               response = length(response)             
+               response = length(response),
+               prop_correct = mean(response)
   )
   return(data.frame(var=var,lbls=res[,1], res[,-1]))
 }
+
+
 
 sg <- get_summary(demographics, "gender")
 se <- get_summary(demographics, "degree")
@@ -324,9 +327,8 @@ ggsave("../images/learning_trend_time.pdf", width=10.5, height = 3.5)
 # Fitting generalized mixed effect model with time taken
 # Link function is inverse link (1/mu) since response is gamma
 
-qplot(time_taken, geom="histogram", data=subset(dd, time_taken<500), binwidth=10)
-
-qplot(time_taken, geom="density", data=subset(dpt, time_taken<300), color=factor(experiment))
+qplot(time_taken, geom="density", data=subset(dpt, time_taken<300), color=factor(experiment)) +
+  scale_colour_hue(name="Experiment")
 
 
 # Checking if the performance increases with attempts

### 2013-06-02 sociology_paper.tex
true data
sd is 2.5

--- 2013-06-02 sociology_paper.tex
@@
 Which side is higher?
-In an experiment XXX of XXX observers picked plot \# XXX as the most extreme plot from the lineup, and all agreed that the blue side was higher.
+In an experiment XXX of XXX observers picked plot \# $3^2+4$ as the most extreme plot from the lineup, and all agreed that the blue side was higher.
 This gives us a visual $p$-value of XXX, indicating that we have enough evidence to reject the null hypothesis. 
@@
 This is what we use as the mean of a normal distribution.
-In the absence of information of error for a poll, we will assume a standard deviation of 3 percentage points for all of the polls.
-While this is a severe simplification, the resulting margin of error of $\pm 6$ percentage points is realistic, and could be easily replaced if additional information on a poll became available.
+In the absence of information of error for a poll, we will assume a standard deviation of 2.5 percentage points for all of the polls.
+While this is a severe simplification, the resulting margin of error of $\pm 5$ percentage points is realistic, and could be easily replaced if additional information on a poll became available.
 A null data set is then generate as a set of draws from normal distributions based on the latest poll results.

### 2013-06-02 sociology_paper.tex

--- 2013-06-02 sociology_paper.tex
@@
 If the observer can detect the actual data plot in the lineup, the hypothesis is rejected.
-Figure \ref{fig:lineup_geno} displays a lineup of 20 plots where one of the plots is observed data.
-The rest of the 19 plots are generated from the assumption that there is no difference between two groups of data.
+Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are rendered from data generated under a null model.
+Which of these towers is the most extreme?
+Which side is higher?
+In an experiment XXX of XXX observers picked plot \# $3^2+4$ as the most extreme plot from the lineup, and all agreed that the blue side was higher.
+This gives us a visual $p$-value of XXX, indicating that we have enough evidence to reject the null hypothesis. 
+What does this mean, though?
+For that, we need to know the context of the data and we need to have more information about the generation of the null plots: in this example, we are investigating, whether the results from the 2012 US presidential election is consistent with polling results.
+As a null hypothesis we  assume that  the election results are consistent with polling results.
+The polling results give us a null model to generate data.
+Each poll gives us a margin between candidates.
+This is what we use as the mean of a normal distribution.
+In the absence of information of error for a poll, we will assume a standard deviation of 3 percentage points for all of the polls.
+While this is a severe simplification, the resulting margin of error of $\pm 6$ percentage points is realistic, and could be easily replaced if additional information on a poll became available.
+A null data set is then generate as a set of draws from normal distributions based on the latest poll results.
+Each panel in figure~\ref{fig:tower} shows an `electoral building' \citet{nytimes}: each state is represented by a rectangle.
+The width of a rectangle is determined by the difference between (polling) results, its height is given by the number of electoral votes, and fill color shows party affiliation (blue for Democrat, and red for Republican).
+Rectangles are placed on top of each other, separated by affiliation and sorted by margin between polls.
+A party needs 270 electoral votes or more to win the presidential election.
+
+%The rest of the 19 plots are generated from the assumption that there is no difference between two groups of data.
 For this, the group structure of the data is broken by random permutation of the groups.
+
 If the null hypothesis is true, the observed data should look similar to the rest of the plots in the lineup making it difficult to detect the observed plot.
-On the other hand if the null hypothesis is not true, observed data should be different from the rest of the plots in the lineup.
+On the other hand, if the null hypothesis is not true, the observed data should be different from the rest of the plots in the lineup and observers should be able to pick the data plot out from the rest.
 

### 2013-06-05 sociology_paper.tex
added demographic information

@@
 
-\subsection{Demographics of Observer}  Age, gender and geographical location of the observer.
+\subsection{Demographics of Observer}  Some of the demographics of the observer such as age, gender, education level and geographical location may have effect on how an observer would examine the lineup.
+This may produce some variability in the performance of the observer.
+For example, a high school student may not respond same as a well educated person while evaluating a lineup.
+Also, to meet the Institutional Regulatory Board (IRB) requirement is is necessary to exclude any subject less than 18 years of old to participate human subject experiment.
+Thus we intend to only focus on the performance of observer with age 18 years or older.
+But still there may be variation in different age groups.
+Recruiting subjects with a well variety of age groups may be necessary to have control over this variability.
+
+Education level has some relation with age.
+Younger observer may not have higher degree.
+Thus some of the variability in performance for different education level may be confounded with age.
+On the other hand, it could be possible to have undergrad degree with higher age.
+Specially for any geographical location where not many certain gender group have higher education, this may occur.
+Thus the effect of all these demographical factors may be confounded at certain level.
 
@@
 So, a different set of null plots may yield some variation in the difficulty of the lineup with the same actual plot.
+This may affect the performance of the observer while evaluating a lineup.
 
@@
 \begin{table}[hbtp]
-\caption{Demographic information of the subjects participated the MTurk experiments.}
+\caption{Demographic information of the subjects participated the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
 \centering
@@
    \centering
+   \includegraphics[width=6.5in]{age_gender_within_country_bar.pdf} 
+   \caption{Countrywise distribution of age and academic levels of the MTurk workers participating the experiments shows the diversity of the subjects in all the demographic aspect.
+Male and female participants differ in India specially for agelevel 18-25.
+For United States number of participants are similar beyond age 40 while few number of participants coming from India beyond that age.}
+   \label{fig:gender_country}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
    \includegraphics[width=6in]{turker_location_experiment.pdf} 

### 2013-06-05 R/sociology_paper.R

@@ -151,6 +151,7 @@ levels(turker$age_level)[1:7] <- c("18-25","26-30","31-35","36-40",
 #   scale_x_discrete( aes(breaks=academic_study,labels=as.charater(study)))
 
 
+# Plotting age, gender and country together
 p1 <- qplot(degree, facets=age_level~gender_level, data=turker) +
   coord_flip() + xlab("Academic Study") + labs(title="Gender")
 
@@ -169,6 +170,12 @@ pushViewport(viewport(layout = grid.layout(1, 2)))
 print(p1, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))     
 print(p2 ,vp = viewport(layout.pos.row = 1, layout.pos.col = 2)) 
 dev.off()
+
+# Potting Education and Gender within country by age
+qplot(degree, data=turker[complete.cases(turker),]) +
+  facet_grid(age_level~country+gender_level) + coord_flip() + xlab("Academic Study") 
+
+ggsave("../images/age_gender_within_country_bar.pdf", width=10, height=8)
 
 
 # Time of the work

### 2013-06-05 sociology_paper.tex
appendix for lineup results

@@
 
+\clearpage
+\section{Electoral Building Lineups and Results}
+Figures~\ref{fig:elect.1} through~\ref{fig:elect.5} show the five lineups shown to (different) Amazon Turk workers in an experiment.
+They all are created as described in the introduction of the paper.
+In order to not bias observers, no context information was given about how these plots were created or what data they were displaying.
+This also made it necessary to slightly more stylize the display.
+
+% latex table generated in R 3.0.1 by xtable 1.7-1 package
+% Wed Jun  5 17:39:09 2013
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ lineup & \#1 & \#2 & \#3 & \#4 & \#5 \\ 
+  \hline
+  \# correct/ \#evaluation &  12/72 & 11/66 & 5/74 & 14/72 & 19/57 \\ 
+$p$-value & 0.00023 & 0.00041 & 0.31 & 1.2e-05 &  1.9e-11 \\ 
+data panel & $3\cdot4+1$ & $2^4+1$ & $4^2+2 $ & $12+\sqrt{25}$ & $2^3-7$ \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+% latex table generated in R 3.0.1 by xtable 1.7-1 package
+% Wed Jun  5 18:00:55 2013
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrrrrrrrrrrrrrrrr}
+  \hline
+& \multicolumn{10}{l}{panel chosen}\\
+Lineup & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\ 
+  \hline
+\#  1 & 2 & 2 & 0 & 10 & 2 & 2 & 6 &  23 & 1 & 1 & 0 & 1 & \bf 12 & 3 & 3 & 1 & 0 & 1 & 1 & 1 \\ 
+\#  2 & 0 & 16 & 1 & 1 & 5 & 1 & 0 & 8 & 0 & 2 & 0 & 0 & 0 & 4 & 2 & 1 & \bf 11 & 1 & 0 & 13 \\ 
+\#  3 & 7 & 26 & 0 & 2 & 0 & 5 & 3 & 0 & 2 & 1 & 0 & 4 & 0 & 0 & 2 & 0 & 9 & \bf 5 & 0 & 6 \\ 
+\#  4 & 0 & 0 & 0 & 2 & 0 & 0 & 0 & 3 & 1 & 10 & 2 & 18 & 1 & 0 & 4 & 2 &\bf 14 & 0 & 13 & 0 \\ 
+\# 5 &\bf 19 & 1 & 4 & 1 & 0 & 1 & 0 & 12 & 0 & 0 & 0 & 4 & 1 & 0 & 0 & 12 & 1 & 1 & 0 & 0 \\ 
+   \hline
+\end{tabular}
+\caption{\label{tab:response}Overview of all choices by observers for each of the lineups.
+The correct choice is bolded.
+In most lineups there are null plots that were picked more often by observers, but the actual result is among the plots being picked most often, indicating that there is some indication that the election result is not completely consistent with the polls.}
+\end{table}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=0.8\linewidth]{electoral-5-13.pdf} 
+   \caption{Lineup \#1 - electoral building. Which building looks the most different from the other buildings? }
+   \label{fig:elect.1}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=0.8\linewidth]{electoral-2-17.pdf} 
+   \caption{Lineup \#2 - electoral building. Which building looks the most different from the other buildings? }
+   \label{fig:elect.2}
+\end{figure}
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=0.8\linewidth]{electoral-3-18.pdf} 
+   \caption{Lineup \#3 - electoral building. Which building looks the most different from the other buildings? }
+   \label{fig:elect.3}
+\end{figure}
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=0.8\linewidth]{electoral-4-17.pdf} 
+   \caption{Lineup \#4 - electoral building. Which building looks the most different from the other buildings? }
+   \label{fig:elect.4}
+\end{figure}
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=0.8\linewidth]{electoral-1-1.pdf} 
+   \caption{Lineup \#5 - electoral building. Which building looks the most different from the other buildings? }
+   \label{fig:elect.5}
+\end{figure}
+
 \bibliographystyle{asa}

### 2013-06-06 sociology_paper.tex
added numbers to the example

@@
 Which side is higher?
-In an experiment XXX of XXX observers picked plot \# $3^2+4$ as the most extreme plot from the lineup, and all agreed that the blue side was higher.
-This gives us a visual $p$-value of XXX, indicating that we have enough evidence to reject the null hypothesis. 
+In an experiment 12 of 72 observers picked plot \# $3^2+4$ as the most extreme plot from the lineup, based on the highest amount of asymmetry (36\%), a trend (26\%) or `outliers' (16\%).
+This gives us a visual $p$-value of 0.00023, indicating that we have enough evidence to reject the null hypothesis (in another four repetitions we found similar results, giving an overall $p$-value of less that $10^{-17}$ based on 62 data picks out of 341 evaluation).
 What does this mean, though?

### 2013-06-06 sociology_paper.tex
model formula

@@
 %\doublespacing
+
+\usepackage{color}
+\newcommand{\hh}[1]{{\color{magenta} #1}}
 
@@
 The other variable is the time taken for each evaluation which is a continuous variable.
-Suppose random variable $Y_{\ell, i}$ denotes the response from subject $ i=1,2, ..., K$ on lineup $\ell=1,2, ..., L$.
+Suppose random variable $Y_{\ell, i}$ denotes the response from subject $ i=1,2, ..., I$ on lineup $\ell=1,2, ..., L$.
 We model the mean $\mu_{\ell, i} = E(Y_{\ell, i})$ with a generalized liner mixed effect model
@@
 
-\noindent where $g(.)$ denotes the link function.
+\hh{
+Let $\pi_{ijk}$ be the probability that  observer $i$ picks the data panel from lineup $j$ in his/her $k$th attempt. We model this in a mixed effects model of the form
+\begin{equation} \label{eqn:trend}
+g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
+\end{equation}
+where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k$th attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$.
+$u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$. $u_i$ is a random intercept, describing a basic subject-specific ability. We assume $u_i \sim N(0, \sigma_u^2)$. 
+$a_i$ is a random slope capturing an individuals' specific learning effect over the course of $K$ attempts. We assume $a_i \sim N(0, \sigma_a^2)$. 
+For $\ell_j$ we again assume a normal distribution, $N(0, \sigma_\ell^2)$. $\ell_j$ is a random intercept predicting lineup difficulty level.
+}
+
+\noindent  $g(.)$ denotes the link function.
 $W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters.

### 2013-06-06 sociology_paper.tex
result table updated

--- 2013-06-05 sociology_paper.tex
@@
    \centering
-   \includegraphics[width=0.9\textwidth]{tower.pdf} 
+   \includegraphics[width=0.99\textwidth]{tower.pdf} 
    \caption{ \label{fig:tower} Which of these towers is the most extreme? Which side is higher?}
@@
   \hline
- & levels & subsjects & avg\_time & response \\ 
+ Factor & levels & subsjects & avg\_time & response \\ 
   \hline
@@
 
-\clearpage
-\section{Electoral Building Lineups and Results}
-Figures~\ref{fig:elect.1} through~\ref{fig:elect.5} show the five lineups shown to (different) Amazon Turk workers in an experiment.
-They all are created as described in the introduction of the paper.
-In order to not bias observers, no context information was given about how these plots were created or what data they were displaying.
-This also made it necessary to slightly more stylize the display.
-
-% latex table generated in R 3.0.1 by xtable 1.7-1 package
-% Wed Jun  5 17:39:09 2013
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrr}
-  \hline
- lineup & \#1 & \#2 & \#3 & \#4 & \#5 \\ 
-  \hline
-  \# correct/ \#evaluation &  12/72 & 11/66 & 5/74 & 14/72 & 19/57 \\ 
-$p$-value & 0.00023 & 0.00041 & 0.31 & 1.2e-05 &  1.9e-11 \\ 
-data panel & $3\cdot4+1$ & $2^4+1$ & $4^2+2 $ & $12+\sqrt{25}$ & $2^3-7$ \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-% latex table generated in R 3.0.1 by xtable 1.7-1 package
-% Wed Jun  5 18:00:55 2013
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrrrrrrrrrrrrrrr}
-  \hline
-& \multicolumn{10}{l}{panel chosen}\\
-Lineup & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\ 
-  \hline
-\#  1 & 2 & 2 & 0 & 10 & 2 & 2 & 6 &  23 & 1 & 1 & 0 & 1 & \bf 12 & 3 & 3 & 1 & 0 & 1 & 1 & 1 \\ 
-\#  2 & 0 & 16 & 1 & 1 & 5 & 1 & 0 & 8 & 0 & 2 & 0 & 0 & 0 & 4 & 2 & 1 & \bf 11 & 1 & 0 & 13 \\ 
-\#  3 & 7 & 26 & 0 & 2 & 0 & 5 & 3 & 0 & 2 & 1 & 0 & 4 & 0 & 0 & 2 & 0 & 9 & \bf 5 & 0 & 6 \\ 
-\#  4 & 0 & 0 & 0 & 2 & 0 & 0 & 0 & 3 & 1 & 10 & 2 & 18 & 1 & 0 & 4 & 2 &\bf 14 & 0 & 13 & 0 \\ 
-\# 5 &\bf 19 & 1 & 4 & 1 & 0 & 1 & 0 & 12 & 0 & 0 & 0 & 4 & 1 & 0 & 0 & 12 & 1 & 1 & 0 & 0 \\ 
-   \hline
-\end{tabular}
-\caption{\label{tab:response}Overview of all choices by observers for each of the lineups.
-The correct choice is bolded.
-In most lineups there are null plots that were picked more often by observers, but the actual result is among the plots being picked most often, indicating that there is some indication that the election result is not completely consistent with the polls.}
-\end{table}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.8\linewidth]{electoral-5-13.pdf} 
-   \caption{Lineup \#1 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.1}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.8\linewidth]{electoral-2-17.pdf} 
-   \caption{Lineup \#2 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.2}
-\end{figure}
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.8\linewidth]{electoral-3-18.pdf} 
-   \caption{Lineup \#3 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.3}
-\end{figure}
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.8\linewidth]{electoral-4-17.pdf} 
-   \caption{Lineup \#4 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.4}
-\end{figure}
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.8\linewidth]{electoral-1-1.pdf} 
-   \caption{Lineup \#5 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.5}
-\end{figure}
-
 \bibliographystyle{asa}

### 2013-06-06 sociology_paper.tex
updated learning trend model

--- 2013-06-06 sociology_paper.tex
@@
 
-As per the design of the experiment for learning trend we have two different response variables.
+The feedback provided by each observer on a lineup is a binary response variable.
+Suppose $Y_{ijk}$ denotes the response from observer $i$ on a  lineup $j$ in his/her $k$th attempt.
+$Y_{ijk}=1$ if the response is correct otherwise  $Y_{ijk}=0$.
+
+%As per the design of the experiment for learning trend we have two different response variables.
 One is observer feedback on a lineup which is a binary variable.
@@
 We model the mean $\mu_{\ell, i} = E(Y_{\ell, i})$ with a generalized liner mixed effect model
+%
+%\begin{equation} \label{eqn:trend}
+%g( \mu_{\ell, i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
+%\end{equation}
 
-\begin{equation} \label{eqn:trend}
-g( \mu_{\ell, i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
-\end{equation}
-
-\hh{
-Let $\pi_{ijk}$ be the probability that  observer $i$ picks the data panel from lineup $j$ in his/her $k$th attempt. We model this in a mixed effects model of the form
-\begin{equation} \label{eqn:trend}
+%\hh{
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the data panel from lineup $j$ in his/her $k$th attempt. We model this in a generalized mixed effects model of the form
+\begin{equation} \label{eqn:trend_response}
 g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
@@
 $a_i$ is a random slope capturing an individuals' specific learning effect over the course of $K$ attempts. We assume $a_i \sim N(0, \sigma_a^2)$. 
-For $\ell_j$ we again assume a normal distribution, $N(0, \sigma_\ell^2)$. $\ell_j$ is a random intercept predicting lineup difficulty level.
-}
+For $\ell_j$ we again assume a normal distribution, $N(0, \sigma_\ell^2)$.
+$\ell_j$ is a random intercept predicting lineup difficulty level.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=g_1(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ .
 
-\noindent  $g(.)$ denotes the link function.
-$W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters.
+%$W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters.
 Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well as the attempt made by the observer.
 $Z_{\ell i}$,  $1 \le i \le K$, $1 \le \ell \le L$,  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$; and
-$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$.
+%$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$.
 $\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
 
-The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend} leads to the estimate of the subject and the lineup specific probability of successful evaluation by a single observer as 
+The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $kth$ attempt by a single observer as
 \begin{equation} \label{eqn:trend_power}
-\hat p_{\ell i} =  g^{-1}(W_{\ell i} \hat {\delta} +  Z_{\ell i}  \hat {\tau}_{\ell i}).
+\hat p_{ijk} =  g^{-1}(\hat{\mu} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
 \end{equation}
 
-For $Y_{l,i}$ being binary response we use {\it logit} link function $g(\mu)=g_1(\mu)=\log(\mu) - \log(1-\mu); 0 \le \mu \le 1$ and $\hat p_{\ell i}$ gives the estimate of proportion correct by subject $i$ with lineup $\ell$.
+The learning of each observer over a course of $K$ evaluations may be observed as the improvement of the time taken to evaluate a lineup in the later attempts.
+Suppose $Z_{ijk}$ denotes the time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
+Let $\mu_{ijk}=  E(Z_{ijk})$ be the average time taken by  observer $i$ to pick the data panel from lineup $j$ in his/her $k$th attempt.
+We model this in a generalized mixed effects model of the form
+\begin{equation} \label{eqn:trend_time}
+g_1( \mu_{ijk} )= \mu_1 + \alpha_k + u_i +  a_{i} k + \ell_j,  
+\end{equation}
+where the link function $g_1(\mu)=1/\mu;  \mu >0$ and $\mu_1$ represents the inverse of overall average time taken to evaluate a lineup.
+
+The inverse link function, $g_1^{-1}(.)$, from  Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $kth$ attempt by a single observer as
+\begin{equation} \label{eqn:trend_power}
+\hat \mu_{ijk} =  g_1^{-1}(\hat{\mu_1} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
+\end{equation}
+
+%For $Y_{l,i}$ being binary response we use {\it logit} link function $g(\mu)=g_1(\mu)=\log(\mu) - \log(1-\mu); 0 \le \mu \le 1$ and $\hat p_{\ell i}$ gives the estimate of proportion correct by subject $i$ with lineup $\ell$.
 The response time taken is a positive random variable with a right skewed distribution.

### 2013-06-06 sociology_paper.tex
time of the work plot updated

@@
 
-\subsection{Performance of the observer} Observer who evaluates a lineup do not necessarily be aware of the data underlying a lineup.
+\subsection{Performance of the observer} Observer who evaluates a lineup do not necessarily be aware of the data that constitue a lineup.
 In response to a specific question the observer is to make a choice based on what he or she sees in the lineup.
@@
 %\hh{
-Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the data panel from lineup $j$ in his/her $k$th attempt. We model this in a generalized mixed effects model of the form
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in his/her $k$th attempt.
+We model this in a generalized mixed effects model of the form
 \begin{equation} \label{eqn:trend_response}
@@
 
-To accomplish this we fit generalized mixed effect model \eqref{eqn:trend} with lineups and subjects as random effect and no fixed effect.
+To accomplish this we fit generalized mixed effect model \eqref{eqn:trend_response} with lineups and subjects as random effect and no fixed effect.
 This gives the estimation of proportion correct responses  and we obtain residuals of the fitted model.
@@
    \includegraphics[width=6.5in]{learning_trend.pdf} 
-   \caption{ Mean residuals of proportion correct vs attempt obtained by adjusting for lineup and subject specific variation.
-For experiment 5, we see a downward trend which may indicate some learning as attempt progress, but none of the slopes in these three experiments are statistically significant.}
+   \caption{ Mean residuals of proportion correct obtained from the model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
+A linear regression line through the points are shown.
+None of the slopes in these three experiments are statistically significant.}
    \label{fig:learning_trend}
@@
    \includegraphics[width=6.5in]{learning_trend_time.pdf} 
-   \caption{Mean residuals of time taken vs attempt obtained by adjusting for lineup and subject specific variation.
-For all the experiment the downward slopes indicate that MTurk workers take less time as they progress through their attempts.}
+   \caption{Mean residuals of time taken obtained from model \eqref{eqn:trend_time} fitted without covariate attempt plotted against attempt.
+Linear regression lines are fitted through the points.
+For all the experiments the significantly downward slopes indicate that MTurk workers take less time as they progress through their attempts.}
    \label{fig:learning_trend_time}
@@
    \centering
-   \includegraphics[width=4.5in]{participation_time.pdf} 
-   \caption{Time of the day when the participants work. Experiment 1 shows MTurk workers participated the experiments around the clock. Other experiments did not take a whole day to finish.}
+   \includegraphics[width=6.5in]{participation_time.pdf} 
+   \caption{Time of the day when the participants work (central time).
+Experiment 1 shows MTurk workers participated the experiments around the clock.
+Other experiments did not take a whole day to finish.
+For experiment 3 most of the participants are from India because of timing.
+No matter when the experiment is started, subjects from India shows participations.
+For United States, subjects participated if the experiment is not in the mid night, except for experiment 6.}
    \label{fig:participation_time}
@@
 % Wed Jun  5 17:39:09 2013
-\begin{table}[ht]
+\begin{table}[hbtp]
 \centering
@@
 % Wed Jun  5 18:00:55 2013
-\begin{table}[ht]
+\begin{table}[hbtp]
 \centering
+\caption{\label{tbl:response}Overview of all choices by observers for each of the lineups.
+The correct choice is bolded.
+In most lineups there are null plots that were picked more often by observers, but the actual result is among the plots being picked most often, indicating that there is some indication that the election result is not completely consistent with the polls.}
 \begin{tabular}{rrrrrrrrrrrrrrrrrrrrr}
@@
 \end{tabular}
-\caption{\label{tab:response}Overview of all choices by observers for each of the lineups.
-The correct choice is bolded.
-In most lineups there are null plots that were picked more often by observers, but the actual result is among the plots being picked most often, indicating that there is some indication that the election result is not completely consistent with the polls.}
 \end{table}

### 2013-06-06 R/sociology_paper.R

@@ -180,10 +180,10 @@ ggsave("../images/age_gender_within_country_bar.pdf", width=10, height=8)
 
 # Time of the work
 turker$hours <- as.factor(hour(as.POSIXlt(turker$start_time)))
-qplot(hours, data=turker) +
-  facet_grid(exp~.,scales="free_y") +
+qplot(hours, data=subset(turker, complete.cases(turker))) +
+  facet_grid(exp~country,scales="free_y") +
   xlab("Hour of the day") + ylab("Number of subjects")
-ggsave("../images/participation_time.pdf", width=8, height=10)
+ggsave("../images/participation_time.pdf", width=10, height=12)
 
 
 # getting demographic summary table

### 2013-06-06 sociology_paper.tex
time taken model - slightly reformulated

@@
 \RequirePackage{natbib}
+\usepackage[colorlinks=true, citecolor=blue, linkcolor=blue]{hyperref}
 \usepackage{amsmath, amssymb, fullpage, amsthm, array,,graphicx,asa, url}
@@
 We model this in a generalized mixed effects model of the form
+\hh{
+\begin{equation} \label{eqn:trend_time}
+g( \mu_{ijk} )= \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j,  
+\end{equation}
+where $g$ is now the inverse link function $g(\mu) = 1/\mu$ for positive values of $\mu$.
+$\mu$ represents the inverse of the overall average time taken by an observer to evaluate a lineup.
+$\alpha$ is the average change in time taken for each additional attempt.
+$\alpha_1$ is an offset in time taken for the first effect.
+All other effects are random effects: as before, $u_i$ is a subject-specific intercept representing  individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$.
+$a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$.
+We assume $a_i \sim N(0, \sigma_a^2)$.
+$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$.
+}
 \begin{equation} \label{eqn:trend_time}

### 2013-06-06 sociology_paper.tex
revised abstract

@@
 \begin {abstract}  
-Visual Statistical Inference is dependent on the careful evaluation of the lineups by individual observers.
-Each individual is different in their cognitive phycology and judiciousness which can affect the power of visual inference.
-To estimate the power of visual inference this can be controlled by getting evaluations from multiple people from a diverse population.
-But the other factor that may affect the power of visual inference is the demographics and skills of the observer.
-In this paper we examine this in details.
-The simulation experiments suggest that individual skill as well as demographics are very significant for the power of visual inference.
-We examine the evidence of learning or getting skilled while sequentially evaluating multiple lineups which enables them evaluate a lineup faster in the later attempts.
-The experimental data suggest that there is no learning effect in terms of correct response.
-The location of actual plot in the lineups does not have any significant effect as well.
+Visual statistical inference is a way to determine significance of patterns found while exploring data.
+It is dependent on the evaluation of a lineup, of a data plot among a sample of null plots, by human observers.
+Each individual is different in their cognitive psychology and judiciousness, which can affect the visual inference.
+The usual way to estimate the effectiveness of a statistical test is its power.
+The estimate of power of a lineup can be controlled by combining evaluations from multiple observers.
+Factors that may also affect the power of visual inference are the observers' demographics, visual skills, and experience, the sample of null plots taken from the null distribution, the position of the data plot in the lineup, and the signal strength in the data.
+This paper examine these factors.
+Results from multiple visual inference studies are examined to provide an assessment of these.
+The experiments suggest that individual skills vary substantially,  but demographics do not have a huge effect.
+There is evidence that a learning effect exists that observers get better and faster with repeated evaluations which suggests a potential for using the lineup protocol for training people to read statistical graphics.
+The placement of actual data plot in the lineups does not affect the inference.
 
-{\bf Keywords: \sf statistical graphics, lineup, non-parametric test, cognitive phycology, visualization, exploratory data analysis} 
+{\bf Keywords: \sf statistical graphics, non-parametric test, cognitive phycology, data visualization, exploratory data analysis, data mining, visual analytics} 
 \end {abstract}

### 2013-06-06 sociology_paper.tex
one more tweak to abstract
performance plots by demographics added

@@
 This paper examine these factors.
-Results from multiple visual inference studies are examined to provide an assessment of these.
+Results from multiple visual inference studies using Amazon's Mechanical Turk are examined to provide an assessment of these.
 The experiments suggest that individual skills vary substantially,  but demographics do not have a huge effect.
@@
    \centering
+   \includegraphics[width=6.5in]{age_gender_within_country_time.pdf} 
+   \caption{Countrywise average time taken for different age and academic levels of the MTurk workers participating the experiments shows that the demographic factors may not have effect on time taken.
+}
+   \label{fig:gender_country}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=6.5in]{age_gender_within_country_correct.pdf} 
+   \caption{Countrywise percentage of correct responses for different age and academic levels of the MTurk workers participating the experiments shows that the demographic factors may not have effect on the percentage of correct responses.
+}
+   \label{fig:gender_country}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
    \includegraphics[width=6in]{turker_location_experiment.pdf} 

### 2013-06-06 R/sociology_paper.R

@@ -177,6 +177,19 @@ qplot(degree, data=turker[complete.cases(turker),]) +
 
 ggsave("../images/age_gender_within_country_bar.pdf", width=10, height=8)
 
+# Potting average time taken by Education and Gender within country vs age
+pdat <- ddply(demographics, .(country,gender_level, age_level,degree), summarize,
+              per_correct = mean(response),
+              avg_time = mean(time_taken))
+qplot(degree, avg_time, geom="bar", stat="identity", data=pdat[complete.cases(pdat),]) +
+  facet_grid(age_level~country+gender_level) + coord_flip() + xlab("Academic Study") 
+
+ggsave("../images/age_gender_within_country_correct.pdf", width=10, height=8)
+
+qplot(degree, per_correct, geom="bar", stat="identity", data=pdat[complete.cases(pdat),]) +
+  facet_grid(age_level~country+gender_level) + coord_flip() + xlab("Academic Study") 
+
+ggsave("../images/age_gender_within_country_time.pdf", width=10, height=8)
 
 # Time of the work
 turker$hours <- as.factor(hour(as.POSIXlt(turker$start_time)))

### 2013-06-06 sociology_paper.tex
start of intro revised

--- 2013-06-06 sociology_paper.tex
@@
 
-Statistical graphics can be used to test the significance of findings during the exploratory data analysis.
-The method is now termed as visual statistical inference which is discussed in the following section.
+The lineup protocol introduced in \citet{buja:2009} can be used to test the significance of findings during the exploratory data analysis.
+The methodology is a part of what is called visual statistical inference.
+These concepts have been developed further by \citet{majumder:2013} who refined the terminology and validated the lineup protocol with a head to head comparison with conventional inference.
+One of the major contributions of  \citet{majumder:2013}  is to define the power of the visual test and how to estimate the power for a particular lineup.
+It was observed that the power can be as good or better than that of a conventional test in some scenarios.
 
-\subsection{Visual Statistical Inference} The fundamental concept of visual statistical inference is introduced by \citet{buja:2009}.
-In their paper, they proposed lineup protocol to discover significant pattern in the data.
-Later these concepts are further developed by defining necessary terminologies and validating using a head to head comparison with conventional inference by \citet{majumder:2013}.
-The power of the visual test is obtained and it is observed that the power is as good as that of conventional test while in some scenarios, visual power exceeds that of conventional tests.
+In visual inference, the test statistic is a plot of the observed data.
+To create a lineup, this plot, called actual data plot, is placed on a layout of null plots.
+The null plots  are generated from the model specified by a null hypothesis, essentially describing what the plot might look like if the data had no structure.
+An observer is asked to evaluate the lineup.
+If the  actual data plot is detected by the observer, the null hypothesis is rejected.
+This means that the structure in the actual data plot has significant structure, a pattern that is not simply due to randomness.
 
-The test statistic used in visual inference is a plot of the observed data.
-This plot, called actual data plot, is then placed on a layout of plot called lineup.
-The rest of the plots in a lineup are generated from the model specified by the null hypothesis.
-An observer is then asked to evaluate the lineup.
-If the observer can detect the actual data plot in the lineup, the hypothesis is rejected.
+\subsection{Visual inference example}
  Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are rendered from data generated under a null model.

### 2013-06-06 sociology_paper.tex
revising intro

@@
 In visual inference, the test statistic is a plot of the observed data.
-To create a lineup, this plot, called actual data plot, is placed on a layout of null plots.
+To create a lineup, this plot, called actual data plot, is placed in a layout of null plots.
 The null plots  are generated from the model specified by a null hypothesis, essentially describing what the plot might look like if the data had no structure.
@@
 This means that the structure in the actual data plot has significant structure, a pattern that is not simply due to randomness.
+Combining the choices of multiple observers provides more stability in the estimation of significance.
 
 \subsection{Visual inference example}
+
 Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are rendered from data generated under a null model.
@@
 
-\subsection{Performance of the observer} Observer who evaluates a lineup do not necessarily be aware of the data that constitue a lineup.
+\subsection{Observer task} Observer who evaluates a lineup do not necessarily be aware of the data that constitue a lineup.
 In response to a specific question the observer is to make a choice based on what he or she sees in the lineup.
@@
 Section \ref{sec:result_socio} presents the analysis of the experimental data and the results.
+
+\subsection{Experiments pooled}
+
+\begin{table*}[hbtp] 
+\caption{Visual test statistics and question asked to the observers to answer while evaluating a lineup } 
+\centering 
+\begin{tabular}{m{0.5cm}m{2cm}m{3cm}m{7.5cm}} 
+\hline\hline 
+Case & Statistic & Test Statistic & Lineup question \\ [0.5ex] % inserts table %heading 
+\hline 
+1  & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{minipage} & Which set of box plots shows biggest vertical difference 
+between group A and B? \\
+2 &  Scatter plot & \begin{minipage}[t]{3cm}   \scalebox{0.45}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Of the scatter plots below which one shows data that has steepest slope? \\
+
+\hline 
+\end{tabular} 
+\label{tbl:visual_stat} 
+\end{table*}
 
@@
 Similarly for case 2, a correct evaluation would indicate that the slope is different than the slope that may show up just from randomness.
-
-\begin{table*}[hbtp] 
-\caption{Visual test statistics and question asked to the observers to answer while evaluating a lineup } 
-\centering 
-\begin{tabular}{m{0.5cm}m{2cm}m{3cm}m{7.5cm}} 
-\hline\hline 
-Case & Statistic & Test Statistic & Lineup question \\ [0.5ex] % inserts table %heading 
-\hline 
-1  & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{minipage} & Which set of box plots shows biggest vertical difference 
-between group A and B? \\
-2 &  Scatter plot & \begin{minipage}[t]{3cm}   \scalebox{0.45}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Of the scatter plots below which one shows data that has steepest slope? \\
-
-\hline 
-\end{tabular} 
-\label{tbl:visual_stat} 
-\end{table*}
 

### 2013-06-06 R/sociology_paper.R
figure placement changed

@@ -306,7 +306,7 @@ get_trend <- function(dat){
     ))
 }
 
-# merging all data for trend analysis
+# merging all data for learning trend analysis
 dtrend <-NULL
 for (i in 1:7){
   di <- get(paste("dat",i, sep=""))
@@ -386,6 +386,7 @@ fit1 <- lm(resid ~ attempt, data=subset(trend.dat, experiment==5))
 fit2 <- lm(resid ~ attempt, data=subset(trend.dat, experiment==6))
 fit3 <- lm(resid ~ attempt, data=subset(trend.dat, experiment==7))
 
+# fitting models with response (percent correct)
 
 dt <- subset(dtrend, experiment==5)
 fit0 <- lmer(response ~ attempt + p_value + (attempt -1|id), family="binomial", data=dt)
@@ -394,7 +395,22 @@ fit1 <- lmer(response ~ attempt + (attempt|id) , family="binomial", data=dt)
 fit2 <- lmer(response ~ attempt + (attempt|id) + (1|pic_id), family="binomial", data=dt)
 anova(fit2, fit1)
 fit3 <- lmer(response ~ factor(attempt) + (attempt|id) + (1|pic_id), family="binomial", data=dt)
+
 anova(fit3,fit2)
+
+
+
+# Fitting gamma mixed effect models with time taken
+
+qplot(time_taken, data=dt, geom="histogram", binwidth=10)
+
+# f0 <- lmer(time_taken ~ factor(attempt) + (attempt|id) + (1|pic_id), family=Gamma(link="inverse"), data=dt)
+
+f0 <- lmer(time_taken ~ factor(attempt) + (attempt|id) + (1|pic_id), family=Gamma(link="log"), data=dt)
+
+
+
+
 
 # ==============================================================
 # Analysis of location effect of the plot using turk 9 data

### 2013-06-07 R/sociology_paper.R
plot for demographic factor main effect added

@@ -106,7 +106,7 @@ get_real_ip <- function(ip) {
 demographics <- NULL
 for (i in 1:10){
   di <- subset(get(paste("dat",i, sep="")), age > 1,
-                 select = c(id, response, start_time, time_taken, 
+                 select = c(id, response, start_time, time_taken, pic_name,
                             ip_address, gender, academic_study, age))
   di$age[di$age==0] <- NA # for experiment 1, age=0 means NA
   di$academic_study[di$academic_study==0] <- NA
@@ -130,6 +130,7 @@ demographics <- merge(demographics,ip.details,all.x=TRUE, by="ip_address")
 demographics$country <- as.character(demographics$country_name)
 demographics$country[(demographics$country_code != "IN") & (demographics$country_code != "US")] <- "Rest of the world"
 demographics$country[demographics$country=="Namibia"] <- "Rest of the world"
+demographics$country[demographics$country==NA] <- "Rest of the world"
 
 
 
@@ -184,12 +185,38 @@ pdat <- ddply(demographics, .(country,gender_level, age_level,degree), summarize
 qplot(degree, avg_time, geom="bar", stat="identity", data=pdat[complete.cases(pdat),]) +
   facet_grid(age_level~country+gender_level) + coord_flip() + xlab("Academic Study") 
 
-ggsave("../images/age_gender_within_country_correct.pdf", width=10, height=8)
+ggsave("../images/age_gender_within_country_time.pdf", width=10, height=8)
 
 qplot(degree, per_correct, geom="bar", stat="identity", data=pdat[complete.cases(pdat),]) +
   facet_grid(age_level~country+gender_level) + coord_flip() + xlab("Academic Study") 
 
-ggsave("../images/age_gender_within_country_time.pdf", width=10, height=8)
+ggsave("../images/age_gender_within_country_correct.pdf", width=10, height=8)
+
+# Demographic factor main effect for percent correct and average time aggregated by lineup
+gdat <- ddply(demographics, .(gender_level, pic_name), summarize,
+              prop_correct = mean(response),
+              avg_time = mean(time_taken))
+get_effect <- function(dat, var){
+  res <- ddply(dat,c(var), summarize,
+               log_avg_time = round(log(mean(time_taken)),2),
[…]

### 2013-06-07 sociology_paper.tex
figure placement changed
plot for demographic factor main effect added
subject-wise residual time taken

@@
 
+\begin{figure}[hbtp] 
+   \centering
+   \includegraphics[width=0.99\textwidth]{tower.pdf} 
+   \caption{ \label{fig:tower} Which of these towers is the most extreme? Which side is higher?}
+\end{figure}
+
 %The rest of the 19 plots are generated from the assumption that there is no difference between two groups of data.
@@
 %\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.99\textwidth]{tower.pdf} 
-   \caption{ \label{fig:tower} Which of these towers is the most extreme? Which side is higher?}
-\end{figure}
 
@@
    \includegraphics[width=6.5in]{learning_trend_time.pdf}    
-   \caption{Mean residuals of time taken obtained from model \eqref{eqn:trend_time} fitted without covariate attempt plotted against attempt.
+   \includegraphics[width=6.5in]{learning_trend_time_subject.pdf}    
+   \caption{ Mean residuals of log(time taken) obtained from model \eqref{eqn:trend_time} fitted without covariate attempt plotted against attempt (above).
+Subject-wise residuals of log(time taken) vs attempt (below).
 Linear regression lines are fitted through the points.
@@
    \centering
+   \includegraphics[width=6.5in]{demographic_effect.pdf} 
+   \caption{Boxplots of average time taken in log scale and proportion correct responses of all the lineups plotted for each demographic factor variables.
+None of the demographic factors appear to be statistically significant except India effect for proportion correct.
+Effect of graduate degree in proportion correct is also significant but is actually the India effect since, Most of the graduate degree holders are coming from India.}
+   \label{fig:gender_country}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
    \includegraphics[width=6.5in]{age_gender_within_country_time.pdf} 

### 2013-06-07 R/sociology_paper.R

@@ -131,8 +131,8 @@ demographics$country <- as.character(demographics$country_name)
 demographics$country[(demographics$country_code != "IN") & (demographics$country_code != "US")] <- "Rest of the world"
 demographics$country[demographics$country=="Namibia"] <- "Rest of the world"
 demographics$country[demographics$country==NA] <- "Rest of the world"
-
-
+levels(demographics$age_level)[7:9] <- 7
+levels(demographics$age_level)[1:7] <- c(levels(turker$age_level)[1:6], "above 50")
 
 
 # getting unique participants for plotting purpose only
@@ -143,9 +143,8 @@ turker$study[turker$age_level=="36-40"] <- c("High school or less (1)","Some und
                                              "Under graduate degree (3)", "Some graduate courses (4)",
                                              "Graduate degree (5)")[turker$study[turker$age_level=="36-40"]]
 turker$exp <- factor(turker$experiment, levels=names(table(turker$experiment))[order(c(1,10,2:9))])
-levels(turker$age_level)[7:9] <- 7
-levels(turker$age_level)[1:7] <- c("18-25","26-30","31-35","36-40",
-                                "41-45", "46-50", "above 50")
+#levels(turker$age_level)[7:9] <- 7
+#levels(turker$age_level)[1:7] <- c(levels(turker$age_level)[1:6], "above 50")
 
 # qplot(factor(academic_study), facets=age_level~gender_level, data=turker) +
 #   coord_flip() + xlab("Academic Study") + labs(title="Gender") +
@@ -215,10 +214,19 @@ mdat <- melt(rbind(gdat,edat,cdat,adat), id=c("variable_level", "pic_name", "var
 qplot(variable_level, value, geom="boxplot", data=mdat[complete.cases(mdat),]) +
   facet_grid(variable~variable_name, scales="free") +
   theme(axis.text.x=element_text(angle=90, hjust=1))
-
 ggsave("../images/demographic_effect.pdf", width=8, height=8)
 
-# Time of the work
+# testing significance of demographic factor main effects
+
+ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+ (1|id)+(1|pic_name), data=demographics)
+summary(ft)
+
[…]

### 2013-06-07 sociology_paper.tex
finished with intro

@@
 \newcommand{\hh}[1]{{\color{magenta} #1}}
+\newcommand{\dc}[1]{{\color{green} #1}}
 
@@
 There is evidence that a learning effect exists that observers get better and faster with repeated evaluations which suggests a potential for using the lineup protocol for training people to read statistical graphics.
-The placement of actual data plot in the lineups does not affect the inference.
+The placement in the lineup does not affect the inference.
 
@@
 In visual inference, the test statistic is a plot of the observed data.
-To create a lineup, this plot, called actual data plot, is placed in a layout of null plots.
+To create a lineup, this plot, called the actual data plot, is placed in a layout of null plots.
 The null plots  are generated from the model specified by a null hypothesis, essentially describing what the plot might look like if the data had no structure.
@@
 
-\subsection{Visual inference example}
+%\subsection{Visual inference example}
 
 Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are rendered from data generated under a null model.
-Which of these towers is the most extreme?
-Which side is higher?
-In an experiment 12 of 72 observers picked plot \# $3^2+4$ as the most extreme plot from the lineup, based on the highest amount of asymmetry (36\%), a trend (26\%) or `outliers' (16\%).
-This gives us a visual $p$-value of 0.00023, indicating that we have enough evidence to reject the null hypothesis (in another four repetitions we found similar results, giving an overall $p$-value of less that $10^{-17}$ based on 62 data picks out of 341 evaluation).
+Which one of the 20 plots is the most different from the others?
+When asked this question, 12 of 72 observers picked the same plot (with number equal to the result of $3^2+4$), with reasons given being 'asymmetry' (36\%), 'trend' (26\%) or `outliers' (16\%).
+The corresponding $p$-value is 0.00023, indicating sufficient evidence to reject the null hypothesis.
 What does this mean, though?
-For that, we need to know the context of the data and we need to have more information about the generation of the null plots: in this example, we are investigating, whether the results from the 2012 US presidential election is consistent with polling results.
-As a null hypothesis we  assume that  the election results are consistent with polling results.
-The polling results give us a null model to generate data.
-Each poll gives us a margin between candidates.
-This is what we use as the mean of a normal distribution.
-In the absence of information of error for a poll, we will assume a standard deviation of 2.5 percentage points for all of the polls.
-While this is a severe simplification, the resulting margin of error of $\pm 5$ percentage points is realistic, and could be easily replaced if additional information on a poll became available.
-A null data set is then generate as a set of draws from normal distributions based on the latest poll results.
-Each panel in figure~\ref{fig:tower} shows an `electoral building' \citet{nytimes}: each state is represented by a rectangle.
-The width of a rectangle is determined by the difference between (polling) results, its height is given by the number of electoral votes, and fill color shows party affiliation (blue for Democrat, and red for Republican).
-Rectangles are placed on top of each other, separated by affiliation and sorted by margin between polls.
-A party needs 270 electoral votes or more to win the presidential election.
+For that, we need to know the context of the data and we need to have more information about the generation of the null plots.
+This example investigates the results from the 2012 US presidential election in comparison to the poll results just prior to the election.
+(Although this example is more simplistic than most of the tests conducted to date, it will serve the purpose of illustrating the lineup protocol.) The data is looking at the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
+Each panel in Figure~\ref{fig:tower} shows an `electoral building' \citet{nytimes} where each state in the union is represented by a rectangle.
+This difference is plotted horizontally, and the height of each box corresponds to the state's electoral votes.
+Color indicates party affiliation.
+The null hypothesis is ``that the election results were consistent with the polls''.
+The polling results provides the null model from which data is simulated.
+Because each poll has a margin of error, this is used to simulate different scenarios that might have resulted on election day, if the polls were on target.
+A null data set is then generate as a set of draws from normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, qpproximating a margin of error of 5\%.
+These samples are plotted as electoral buildings, and the plot made with the results from the election is placed randomly among them in a lineup of size $5\times 4$.
+If the null hypothesis is true then the actual data plot should look just like any of the other plots, and not be identified by an observer.
+%A party needs 270 electoral votes or more to win the presidential election.
 
@@
    \includegraphics[width=0.99\textwidth]{tower.pdf} 
-   \caption{ \label{fig:tower} Which of these towers is the most extreme? Which side is higher?}
+   \caption{ \label{fig:tower} \dc{Which one of the plots is the most from the others? Maybe this question will have people picking 4, because it is the only one with red higher.}}
 \end{figure}
@@
 
-If the null hypothesis is true, the observed data should look similar to the rest of the plots in the lineup making it difficult to detect the observed plot.
+%If the null hypothesis is true, the observed data should look similar to the rest of the plots in the lineup making it difficult to detect the observed plot.
 On the other hand, if the null hypothesis is not true, the observed data should be different from the rest of the plots in the lineup and observers should be able to pick the data plot out from the rest.
@@
 A lineup can be evaluated by a single person or multiple observers.
-Based on the observer feedback the $p$-value is computed \citep{majumder:2013} and finally a decision is made about the hypothesis being tested.
+A \dc{multinomial} distribution is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
+Observers should not be aware of the data that constitutes a lineup, and should not have seen the actual data plot before seeing the lineup.
+This is the reason that in the election example, above, the scenario was explained after the lineup question, in the text.
+
+The question that is asked of the observer should be as general as possible, effectively asking the observer to pick the plot that is different, and allowing them to provide their reasons for seeing their pick as different.
+In some of the studies, the ones described in  \cite{majumder:2013} very specific questions were asked, because the experiments were being conducted to compare results from the lineup protocol with those of conventional tests.
+In those experiments, structure in the data was strictly controlled in the simulation process, which allowed for specific questions to be asked.
+In the election example, observers were asked ``which plot is the most different?''.
+The type of plot, showing two (modified) stacked bar charts in different colors should suggest to the observer that the interesting feature is most difference between the two heights.
+Most observers got this, but it is possible that some observers might pick plot 4, where the red tower is slightly above the blue as the most different because it is the only plot with this feature.
+So a better question may have been ``Which plot shows the biggest height difference between the two towers?'' except that this tailors the inference to a specific feature which does not match the null hypothesis of interest.
 
@@
 
-\subsection{Observer task} Observer who evaluates a lineup do not necessarily be aware of the data that constitue a lineup.
-In response to a specific question the observer is to make a choice based on what he or she sees in the lineup.
-The choice could be right or wrong depending on whether the data plot is selected or not.
-Under null hypothesis, all the null plots look similar to what the observed data plot appears to be.
-This would make it harder to detect the observed data plot in the lineup.
+%\subsection{Observer task}
+
+For any evaluation the observer may or may not identify the actual data plot.
+Under the null hypothesis, the actual data plot should look similar to the null plots making it harder to detect.
 It is not expected that an observer would be able to detect the data plot in this scenario.
-But since there are limited number of plots in a lineup which is 20 in this case, there is a 1/20 chance that the observer would pick the actual plot.
-This proportion is associated with the type-I error probability.
-On the other hand if null hypothesis is not true, the observed plot would be different than the null plots making it easier to be distinguished.
-When multiple observers evaluate a lineup, the proportion of correct response can be used to estimate the power of the visual test \citep{majumder:2013}.
+But since there are limited number of plots in a lineup, which is 20 in the election example, there is a 1/20 chance that the observer would pick the actual plot.
+This proportion is associated with the Type~I error of the test.
+On the other hand, if null hypothesis is not true, the observed plot should look different from the null plots, making it easier to be detected.
+This is the definition of the power of the test.
+When multiple observers evaluate a lineup, the proportion of correct response can be used to estimate the power.
+The ability of individual observers can vary, and examining this is the purpose of this paper.
 
-For a lineup with fixed difficulty, the proportion correct responses may vary for different observers.
+%For a lineup with fixed difficulty, the proportion correct responses may vary for different observers.
 If an observer evaluates multiple lineups, the proportion of correct evaluations by the observer could be used as a measure of performance by an observer.
@@
 
-Some of the factors that may affect the performance of the observer are discussed in Section \ref{sec:factor_performance}.
+%Some of the factors that may affect the performance of the observer are discussed in Section \ref{sec:factor_performance}.
 In this paper we focus on estimating the learning effect of the observer in terms of both proportion correct and time taken to evaluate a lineup through the attempts made during the evaluation of multiple plots.
@@
 
-This paper is organized as follows; Section \ref{sec:factor_performance} describes some of the factors that may affect the performance of the observer.
-Section \ref{sec:exp_design} describes the design of the simulation experiments and methods to estimate the learning trend of the observer and location effect of the actual plot in a lineup.
-Section \ref{sec:result_socio} presents the analysis of the experimental data and the results.
+There have been \dc{ten} experiments conducted using Amazon's Mechanical Turk~\cite{turk} that are being used to evaluate the the effect of the observer.
+Table \ref{tbl:visual_stat} summarizes these.
+Experiments XXX are used to evaluate the effects of demographic factors whethere there is a learning trend.
+The design of experiment XXX enabled effect of position in the lineup, and the sample of nulls, to be evaluated.
+Section \ref{sec:factor_performance} describes the factors that may affect the performance of the observer.
+Section \ref{sec:exp_design} describes the methods used to assess the effects, and Section \ref{sec:result_socio} describes the results.
 
-\subsection{Experiments pooled}
+%\subsection{Experiments pooled}
 

### 2013-06-07 sociology_paper.tex
updating experimental table
experimental table updated

@@
 Table \ref{tbl:visual_stat} summarizes these.
-Experiments XXX are used to evaluate the effects of demographic factors whethere there is a learning trend.
-The design of experiment XXX enabled effect of position in the lineup, and the sample of nulls, to be evaluated.
+All of these experiments are used to evaluate the effects of demographic factors.
+Experiments 5,6 and 7 are used to study learning trend of the observer.
+The design of experiment 9 enabled effect of position in the lineup, and the sample of nulls, to be evaluated.
 Section \ref{sec:factor_performance} describes the factors that may affect the performance of the observer.
@@
 \centering 
-\begin{tabular}{m{0.5cm}m{2cm}m{3cm}m{7.5cm}} 
+\begin{tabular}{m{.5cm}m{3cm}m{3cm}m{6.5cm}} 
 \hline\hline 
-Case & Statistic & Test Statistic & Lineup question \\ [0.5ex] % inserts table %heading 
+Serial & Experiment & Test Statistic & Lineup question \\ [0.5ex] % inserts table %heading 
 \hline 
-1  & Box plot & \begin{minipage}[t]{3cm} 	\scalebox{0.45}{\includegraphics{stat_category.pdf}} \end{minipage} & Which set of box plots shows biggest vertical difference 
+1  & Box plot & \begin{minipage}[t]{3cm} \begin{center}	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
 between group A and B? \\
-2 &  Scatter plot & \begin{minipage}[t]{3cm}   \scalebox{0.45}{\includegraphics{stat_beta_k.pdf}}\end{minipage} & Of the scatter plots below which one shows data that has steepest slope? \\
+2 &  Scatter plot & \begin{minipage}[t]{3cm}  \begin{center} \scalebox{0.45}{\includegraphics{stat_beta_k.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+\\
+
+  3 & Contaminated plot &\begin{minipage}[t]{3cm} \begin{center} \scalebox{0.45}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+\\  \\
+  4 & Polar vs Cartesian & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different? \\ 
+  5 & Hist vs density & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?\\
+  6 & Violine vs boxplot & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?\\
+  7 & Group separation & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?\\
+  8 & Sine Illusion & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?\\
+  9 & Gene expression &\begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?\\
+  10 & Test normality & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?\\
 
@@
 
-Plots to do: for the same signal strength show the proportion correct for scatterplot and boxplot. eg., qplot(effect, ump-visual power, color= plot-type)
+%Plots to do: for the same signal strength show the proportion correct for scatterplot and boxplot. eg., qplot(effect, ump-visual power, color= plot-type)
 

### 2013-06-07 sociology_paper.tex
model update

@@
 The learning of each observer over a course of $K$ evaluations may be observed as the improvement of the time taken to evaluate a lineup in the later attempts.
-Suppose $Z_{ijk}$ denotes the time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
-Let $\mu_{ijk}=  E(Z_{ijk})$ be the average time taken by  observer $i$ to pick the data panel from lineup $j$ in his/her $k$th attempt.
-We model this in a generalized mixed effects model of the form
+Suppose $Z_{ijk}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
+Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k$th attempt.
+We model this in a mixed effects model of the form
 \hh{
 \begin{equation} \label{eqn:trend_time}
-g( \mu_{ijk} )= \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j,  
+Z_{ijk} = \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
 \end{equation}
-where $g$ is now the inverse link function $g(\mu) = 1/\mu$ for positive values of $\mu$.
-$\mu$ represents the inverse of the overall average time taken by an observer to evaluate a lineup.
-$\alpha$ is the average change in time taken for each additional attempt.
-$\alpha_1$ is an offset in time taken for the first effect.
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
+$\alpha$ is the average change in log time taken for each additional attempt.
+$\alpha_1$ is an offset in log time taken for the first attempt.
 All other effects are random effects: as before, $u_i$ is a subject-specific intercept representing  individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$.
@@
 We assume $a_i \sim N(0, \sigma_a^2)$.
-$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$.
+$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
 }
-\begin{equation} \label{eqn:trend_time}
-g_1( \mu_{ijk} )= \mu_1 + \alpha_k + u_i +  a_{i} k + \ell_j,  
-\end{equation}
-where the link function $g_1(\mu)=1/\mu;  \mu >0$ and $\mu_1$ represents the inverse of overall average time taken to evaluate a lineup.
 
-The inverse link function, $g_1^{-1}(.)$, from  Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $kth$ attempt by a single observer as
+%\begin{equation} \label{eqn:trend_time}
+%Z_{ijk} = \mu_1 + \alpha_k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
+%\end{equation}
+%where the link function $g_1(\mu)=1/\mu;  \mu >0$ and $\mu_1$ represents the inverse of overall average time taken to evaluate a lineup.
+
+Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $kth$ attempt by a single observer as 
 \begin{equation} \label{eqn:trend_power}
-\hat \mu_{ijk} =  g_1^{-1}(\hat{\mu_1} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
+\hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
 \end{equation}

### 2013-06-07 sociology_paper.tex
election result more figures

@@
 
+\begin{figure}[hbtp] 
+   \centering
+   \includegraphics[width=0.99\textwidth]{electoral-5-13.pdf} 
+   \caption{ \label{fig:elect-1} \dc{Which one of the plots is the most different from the others?
+Maybe this question will have people picking 4, because it is the only one where the red side is higher.}}
+\end{figure}
+
 Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are rendered from data generated under a null model.
@@
 The corresponding $p$-value is 0.00023, indicating sufficient evidence to reject the null hypothesis.
+, indicating that we have enough evidence to reject the null hypothesis.
 What does this mean, though?
+For that, we need to know the context of the data and we need to have more information about the generation of the null plots.
 For that, we need to know the context of the data and we need to have more information about the generation of the null plots.
@@
 (Although this example is more simplistic than most of the tests conducted to date, it will serve the purpose of illustrating the lineup protocol.) The data is looking at the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
-Each panel in Figure~\ref{fig:tower} shows an `electoral building' \citet{nytimes} where each state in the union is represented by a rectangle.
+Each panel in Figure~\ref{fig:elect-1} shows an `electoral building' \citet{nytimes} where each state in the union is represented by a rectangle.
 This difference is plotted horizontally, and the height of each box corresponds to the state's electoral votes.
@@
 Because each poll has a margin of error, this is used to simulate different scenarios that might have resulted on election day, if the polls were on target.
-A null data set is then generate as a set of draws from normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, qpproximating a margin of error of 5\%.
+A null data set is then generate as a set of draws from normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
 These samples are plotted as electoral buildings, and the plot made with the results from the election is placed randomly among them in a lineup of size $5\times 4$.
 If the null hypothesis is true then the actual data plot should look just like any of the other plots, and not be identified by an observer.
+ \hh{Figure~\ref{fig:tower} shows a plot of the electoral building with added context information and labels.}
  %A party needs 270 electoral votes or more to win the presidential election. 
+%\begin{figure}[htbp] 
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{electoral-5-13.pdf} 
+%   \caption{Which plot looks the most different from the other plots?  Lineup \#1 - electoral building; another set of four lineups can be found in figs.~{fig:elect.2} through~{fig:elect.5}}
+%   \label{fig:elect.1}
+%\end{figure}
 
@@
    \centering
-   \includegraphics[width=0.99\textwidth]{tower.pdf} 
-   \caption{ \label{fig:tower} \dc{Which one of the plots is the most from the others? Maybe this question will have people picking 4, because it is the only one with red higher.}}
+   \includegraphics[width=0.49\textwidth]{tower.pdf} \hfill
+   \includegraphics[width=0.50\textwidth]{freedom-sim.pdf} 
+   \caption{ \label{fig:tower} Electoral building plot of the results of the 2012 U.S.
+Presidential Election (left).
+On the right a histogram of 10,000 simulations from the same model.
+The $p$-value of observing results as extreme as in the  2012 U.S.
+election based on the bootstrap is 0.0533 (with Bootstrap standard error of 0.002).}
 \end{figure}
@@
 So a better question may have been ``Which plot shows the biggest height difference between the two towers?'' except that this tailors the inference to a specific feature which does not match the null hypothesis of interest.
+%=======
+%A lineup can be evaluated by a single person or multiple observers.
+Based on the feedback of $K$ observers a $p$-value is computed with the help of a binomial random variable $X$ for the number of observers picking the data plot.
+Under the null hypothesis is $X$ is  distributed according to $B_{K, 1/20}$ \citep{majumder:2013}.
+The $p$-value for a lineup is then given as $P(X \ge x)$.
+%%This finally a decision is made about the hypothesis being tested.
+%>>>>>>> example
 
@@
    \centering
-   \includegraphics[width=0.8\linewidth]{electoral-5-13.pdf} 
-   \caption{Lineup \#1 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.1}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
    \includegraphics[width=0.8\linewidth]{electoral-2-17.pdf} 

### 2013-06-07 sociology_paper.tex
figures for RCP averages
experimental setup updated

--- 2013-06-07 sociology_paper.tex
@@
 \begin{table*}[hbtp] 
-\caption{Visual test statistics and question asked to the observers to answer while evaluating a lineup } 
 \centering 
+\caption{Visual test statistics used in 10 different simulation experiments. The observers are asked different questions to answer while evaluating a lineup} 
 \begin{tabular}{m{.5cm}m{3cm}m{3cm}m{6.5cm}} 
@@
 \\  \\
-  4 & Polar vs Cartesian & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different? \\ 
-  5 & Hist vs density & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?\\
+  4 & Polar vs Cartesian & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different? \\ 
+  5 & Hist vs density & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?\\
   6 & Violine vs boxplot & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?\\
@@
 The performance of the observer while evaluating a lineup would depend on the factors described in Section \ref{sec:factor_performance}.
-In this section we present the experimental design to control some  of them so that learning effect of observer and location effect of a plot in the lineup can be estimated.
+This section presents the experimental setup, data collection methods and data analysis plans with models to asses the influence of those demographic factors on visual statistical inference.
 
-\subsection{Experiment Setup}  Multiple simulation experiments are designed to investigate the learning trend of the observer and the location effect of the actual data in the lineup.
-These are explained in tis section.
+\subsection{Experiment Setup}  In addition to the evaluation response of the lineup, all the 10 experiments shown in Table \ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
 
-\paragraph{Design for Learning Trend:} Learning trend of an observer can be observed in terms of performance over successive feedback when multiple lineups are shown for evaluation.
-Each subject is shown a total of 10 lineups.
+\begin{enumerate} \itemsep 0in
+\item Age level
+\item Gender
+\item Academic education level
+\item Geographical location
+\end{enumerate}
+
+Instead of collecting exact age, 9 levels of age were collected.
+They are 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, 51-55, 56-60, above 60.
+There were five levels of academic information.
+They are (1) High school or less , (2) Some under grad course, (3) Under graduate degree, (4) Some graduate courses,  (5) Graduate degree.
+Geographical locations were collected using the true public ip addresses of the participants' computer.
+This provides latitude and longitude of the ip address as well as the city and country information.
+
+%Multiple simulation experiments are designed to investigate the learning trend of the observer and the location effect of the actual data in the lineup. These are explained in tis section.
+
+\paragraph{Learning Trend:} Learning trend of an observer can be observed in terms of performance over successive feedback when multiple lineups are shown for evaluation.
+Experiments 5, 6 and 7 were used for this.
+Each subject was shown a total of 10 lineups randomly selected from a pool of lineups.
 The lineups are not necessarily with the same difficulty levels.
@@
 The responses of the lineups are recorded by attempt 1 through 10.
-Attempt 1 means the response is for the first lineup and attempt 10 means the response is for the 10th lineup.
+Attempt 1 means the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
 The goal is to estimate whether performance of the observer improves from attempt 1 to attempt 10.
 
-\paragraph{Design for Location effect:} The experiment is set up based on the gene expression data with two groups.
+\paragraph{Design for Location effect:} The experiment was set up based on the gene expression data with two groups.
 The groups are defined based on the presence or absence of iron sufficiency.
 The null hypothesis is that there is no difference between this two groups.
-Null plots are obtained by randomizing the group structure of the data.
-The actual plot is then randomly placed in five different locations.
-For each location, five different sets of null plots are used to produce five different lineups.
-This produces 25 different lineups.
-One such lineup is shown in Figure \ref{fig:lineup_geno}.
-Similar procedure is applied to obtain 25 lineups to test the interaction between Genotype and Empty Vector (EV).
+Null plots were obtained by randomizing the group structure of the data.
+The actual plot was randomly placed in five different locations.
+For each location, five different sets of null plots were used to produce five different lineups.
+This produced 25 different lineups.
+Similar procedure was applied to obtain 25 lineups to test the interaction between Genotype and Empty Vector (EV).
 
-Each observer is shown three lineups.
-First lineups shown is a randomly selected Interaction lineup, then one from randomly selected Genotype lineup.
-Finally a test lineup is shown to screen out unusual evaluation.
-The subject is not aware which one is the test plot but it is informed in the starting of the task about having a test plot.
-The test plot is very easy and everyone should detect the data plot.
-The MTurk worker is paid based on whether he or she could correctly detect the data plot.
+%One such lineup is shown in Figure \ref{fig:lineup_geno}.
+
+Each observer was shown three lineups.
+First lineups shown is a randomly selected Interaction lineup, one from randomly selected Genotype lineup.
+Finally a test lineup was shown to screen out unusual evaluation.
+The subjects were not aware which one is the test plot but it was informed in the starting of the task about having a test plot.
+The test plot was very easy and everyone should detect the data plot.
+The MTurk worker was paid based on whether he or she could correctly detect the data plot.
 But the data on test plot are excluded from the analysis.
@@
 Even thoug total participations from india is larger compared to United States, they are mostly young people who are capable of using internet and computer.
+
+\subsection{Demographic Factors}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=6.5in]{demographic_effect.pdf} 
+   \caption{Boxplots of average time taken in log scale and proportion correct responses of all the lineups plotted for each demographic factor variables.
+None of the demographic factors appear to be statistically significant except India effect for proportion correct.
+Effect of graduate degree in proportion correct is also significant but is actually the India effect since, Most of the graduate degree holders are coming from India.}
+   \label{fig:gender_country}
+\end{figure}
 
@@
    \centering
-   \includegraphics[width=6.5in]{demographic_effect.pdf} 
-   \caption{Boxplots of average time taken in log scale and proportion correct responses of all the lineups plotted for each demographic factor variables.
-None of the demographic factors appear to be statistically significant except India effect for proportion correct.
-Effect of graduate degree in proportion correct is also significant but is actually the India effect since, Most of the graduate degree holders are coming from India.}
-   \label{fig:gender_country}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
    \includegraphics[width=6.5in]{age_gender_within_country_time.pdf} 

### 2013-06-07 sociology_paper.tex

@@
    \centering
-   \includegraphics[width=0.99\textwidth]{tower.pdf} 
-   \caption{ \label{fig:tower} \dc{Which one of the plots is the most from the others? Maybe this question will have people picking 4, because it is the only one with red higher.}}
+   \begin{minipage}[c]{0.49\textwidth}
+   \includegraphics[width=\textwidth]{tower.pdf}
+   \end{minipage}
+    \hfill
+   \begin{minipage}[c]{0.45\textwidth}
+   Freedom's Lighthouse Averages \\
+      \includegraphics[width=\textwidth]{freedom-sim.pdf} 
+   RealClearPolitics Averages\\
+      \includegraphics[width=\textwidth]{rcp-sim.pdf} 
+   \end{minipage}
+   \caption{ \label{fig:tower} Electoral building plot of the results of the 2012 U.S.
+Presidential Election (left).
+On the right a histogram of 10,000 simulations from the same model.
+The $p$-value of observing results as extreme as in the  2012 U.S.
+election based on the bootstrap is 0.0533 (with Bootstrap standard error of 0.002).}
 \end{figure}

### 2013-06-07 sociology_paper.tex
bootstrap explanation

@@
 Presidential Election (left).
-On the right a histogram of 10,000 simulations from the same model.
-The $p$-value of observing results as extreme as in the  2012 U.S.
-election based on the bootstrap is 0.0533 (with Bootstrap standard error of 0.002).}
+On the right two histograms of 10,000 simulations each based on polling averages from two different sources.
+For the histogram on the top, the $p$-value of observing results as extreme as  the  2012 U.S.
+election results based on the bootstrap is 0.0533 (with Bootstrap standard error of 0.002), making the election results almost significantly different from the polls.
+There is no indication of any inconsistency between polls and election results based on the bootstrap simulation below.
+The lineups are based on the top source.
+}
 \end{figure}

### 2013-06-07 sociology_paper.tex
edited experiment setup

--- 2013-06-07 sociology_paper.tex
@@
 The performance of the observer while evaluating a lineup would depend on the factors described in Section \ref{sec:factor_performance}.
-This section presents the experimental setup, data collection methods and data analysis plans with models to asses the influence of those demographic factors on visual statistical inference.
+This section presents the simulation experimental setup, data collection methods and data analysis plans with models to asses the influence of those factors on visual statistical inference.
 
-\subsection{Experiment Setup}  In addition to the evaluation response of the lineup, all the 10 experiments shown in Table \ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
+\subsection{Experiment Setup}  In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table \ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
 

### 2013-06-07 R/sociology_paper.R
dots showing averages added to demographic factor plot

@@ -213,6 +213,7 @@ mdat <- melt(rbind(gdat,edat,cdat,adat), id=c("variable_level", "pic_name", "var
 
 qplot(variable_level, value, geom="boxplot", data=mdat[complete.cases(mdat),]) +
   facet_grid(variable~variable_name, scales="free") +
+  stat_summary(fun.y=mean, geom="point") + xlab("Levels of demographic factors")+
   theme(axis.text.x=element_text(angle=90, hjust=1))
 ggsave("../images/demographic_effect.pdf", width=8, height=8)
 

### 2013-06-07 sociology_paper.tex
dots showing averages added to demographic factor plot
learning trend on time plot updated

@@
    \caption{Boxplots of average time taken in log scale and proportion correct responses of all the lineups plotted for each demographic factor variables.
+The dots inside the box represents mean.
 None of the demographic factors appear to be statistically significant except India effect for proportion correct.
@@
    \caption{ Mean residuals of proportion correct obtained from the model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
-A linear regression line through the points are shown.
+Linear regression lines fitted through the points are shown.
 None of the slopes in these three experiments are statistically significant.}
@@
    \includegraphics[width=6.5in]{learning_trend_time.pdf}    
-   \includegraphics[width=6.5in]{learning_trend_time_subject.pdf}    
-   \caption{ Mean residuals of log(time taken) obtained from model \eqref{eqn:trend_time} fitted without covariate attempt plotted against attempt (above).
-Subject-wise residuals of log(time taken) vs attempt (below).
-Linear regression lines are fitted through the points.
-For all the experiments the significantly downward slopes indicate that MTurk workers take less time as they progress through their attempts.}
+%   \includegraphics[width=6.5in]{learning_trend_time_subject.pdf}    
+   \caption{ Mean residuals of log(time taken) obtained from model \eqref{eqn:trend_time} fitted without covariate attempt plotted against attempt.
+Linear regression lines are fitted through the points excluding the first attempt.
+For all the three experiments the downward slopes are statistically significant which indicates that MTurk workers take less time as they progress through their attempts.}
    \label{fig:learning_trend_time}

### 2013-06-07 R/sociology_paper.R

@@ -389,6 +389,12 @@ qplot(attempt,mean_resid, data= dmt) + geom_point(size=2.5) +
   facet_wrap(~experiment, scales="free_y") +
   scale_x_continuous(breaks = seq(2,10,by=2)) 
 
+ggplot() + 
+  geom_point(aes(attempt,mean_resid), data= dmt) +
+  geom_smooth(aes(attempt,mean_resid), data= subset(dmt, attempt>1), method="lm", se=F) +
+  facet_wrap(~experiment, scales="free_y") + ylab("Residual log(time taken)") +
+  scale_x_continuous(breaks = seq(2,10,by=2))
+
 ggsave("../images/learning_trend_time.pdf", width=10.5, height = 3.5)
 
 

### 2013-06-08 sociology_paper.tex
Results of Model for time taken added

@@
   5 & Hist vs density & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?\\
-  6 & Violine vs boxplot & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?\\
+  6 & Violin vs boxplot & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?\\
   7 & Group separation & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?\\
@@
 
-\begin{enumerate} \itemsep 0in
+\begin{enumerate} \leftmargin 5cm  \itemsep 0in
 \item Age level
@@
 We model this in a mixed effects model of the form
-\hh{
+%\hh{
 \begin{equation} \label{eqn:trend_time}
@@
 $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-}
+%}
 
@@
 
-\subsection{Demographic Factors}
+\subsection{Demographic Factors} Proportions of correct responses and the logarithm of average time taken for each lineup are computed and their distributions are shown in Figure \ref{fig:demographic_effect} for different demographic factor variables.
 
@@
    \centering
-   \includegraphics[width=6.5in]{demographic_effect.pdf} 
+   \includegraphics[width=4in]{demographic_effect.pdf} 
    \caption{Boxplots of average time taken in log scale and proportion correct responses of all the lineups plotted for each demographic factor variables.
@@
 Effect of graduate degree in proportion correct is also significant but is actually the India effect since, Most of the graduate degree holders are coming from India.}
-   \label{fig:gender_country}
+   \label{fig:demographic_effect}
 \end{figure}
 
-\subsection{Learning Trend} Each subject participating the experiment has seen multiple number of lineups for evaluations.
+\subsection{Learning Trend} Model \eqref{eqn:trend_time} is fitted to the data from experiment 5, 6, and 7.
+The results are shown in Table \ref{tbl:model_result_time}.
+The model is fitted using R function $lmer()$ from package $lme4$.
+To obtain the $p$-values of fixed effect parameters, normal approximation is used $Z$ scores are computed accordingly.
+
+%Each subject participating the experiment has seen multiple number of lineups for evaluations.
 Suppose $K$ be the number of lineups evaluated by a subject and $A_k$ be the $kth$ attempt, $k=1,2, ..., K$.
@@
 But to obtain this we have to adjust for lineup difficulty and individual performances.
+
+\begin{table}[htbp]
+\centering
+\caption{Parameter estimates of Model \eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
+Both fixed effect parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.}
+\begin{tabular}{rrrcccrrcccrrcc}
+  \hline
+& \multicolumn{4}{c} {Experiment 5} & &\multicolumn{4}{c} {Experiment 6} && \multicolumn{4}{c} {Experiment 7}\\
+\cline{2-5} \cline{7-10} \cline{12-15} 
+ Effect& Est & SE & Zval & pvalue && Est & SE & Zval & pvalue && Est & SE & Zval & pvalue \\ 
+  \hline
+Fixed &  &  &  & && &  & & & & & &  & \\ 
+$\mu$ & 3.82 & 0.04 & 97.38 & $<$0.001 & & 3.90 & 0.03 & 118.19 & $<$0.001 && 3.73 & 0.05 & 69.04 & $<$0.001 \\ 
+  $\alpha_1$ & 0.33 & 0.04 & 9.35 & $<$0.001 && 0.34 & 0.03 & 11.4 & $<$0.001 && 0.28 & 0.05 & 5.63 & $<$0.001 \\ 
+  $\alpha$ & -0.04 & 0.00 & -9.3 & $<$0.001 && -0.04 & 0.00 & -10.19 & $<$0.001 && -0.03 & 0.01 & -4.22 & $<$0.001 \\ 
+Rand &  &  &  &  &  & & &&& & & &  & \\ 
+  $\sigma^2_l$ & 0.01 & 0.09 &  &  && 0.24 & 0.49 &  &&  & 0.13 & 0.37 &  &  \\ 
+  $\sigma^2_u$ & 0.26 & 0.51 &  &  & &0.00 & 0.05 & & &  & 0.00 & 0.05 &  &  \\ 
+   $\sigma^2_a$ & 0.00 & 0.03 &  &  & &0.04 & 0.20 & & &  & 0.06 & 0.24 &  &  \\ 
+  $\sigma^2$ & 0.21 & 0.46 &  &  & &0.25 & 0.50 &  &  && 0.21 & 0.45 &  &  \\ 
+   \hline
+\end{tabular}
+\label{tbl:model_result_time}
+\end{table}
 
@@
   5 & Hist vs density & 234 &  37 & 17.85 & 41.57 & 17.76 & 1.00 & 3.36 \\ 
-  6 & Violine vs boxplot & 417 &  17 & 17.95 & 105.87 & 25.39 & 1.00 & 3.34 \\ 
+  6 & Violin vs boxplot & 417 &  17 & 17.95 & 105.87 & 25.39 & 1.00 & 3.34 \\ 
   7 & Group separation & 106 &   6 & 16.13 & 5.15 & 4.86 & 1.00 & 3.72 \\ 
@@
 \section{Electoral Building Lineups and Results}
-Figures~\ref{fig:elect.1} through~\ref{fig:elect.5} show the five lineups shown to (different) Amazon Turk workers in an experiment.
+Figures~\ref{fig:elect-1} and \ref{fig:elect.2} through~\ref{fig:elect.5} are the five lineups shown to (different) Amazon Turk workers in an experiment.
 They all are created as described in the introduction of the paper.

### 2013-06-08 R/sociology_paper.R

@@ -215,7 +215,7 @@ qplot(variable_level, value, geom="boxplot", data=mdat[complete.cases(mdat),]) +
   facet_grid(variable~variable_name, scales="free") +
   stat_summary(fun.y=mean, geom="point") + xlab("Levels of demographic factors")+
   theme(axis.text.x=element_text(angle=90, hjust=1))
-ggsave("../images/demographic_effect.pdf", width=8, height=8)
+ggsave("../images/demographic_effect.pdf", width=6, height=6.5)
 
 # testing significance of demographic factor main effects
 
@@ -453,17 +453,36 @@ fit3 <- lmer(response ~ factor(attempt) + (attempt|id) + (1|pic_id), family="bin
 anova(fit3,fit2)
 
 
-
-# Fitting gamma mixed effect models with log(time_taken)
+# Fitting mixed effect models with log(time_taken)
 
 qplot(log(time_taken), data=dt, geom="histogram", binwidth=.1)
 
-# f0 <- lmer(time_taken ~ factor(attempt) + (attempt|id) + (1|pic_id), family=Gamma(link="inverse"), data=dt)
-# f0 <- lmer(time_taken ~ factor(attempt) + (attempt|id) + (1|pic_id), family=Gamma(link="log"), data=dt)
-f0 <- lmer(log(time_taken) ~I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), data=dt)
+get_estimates <- function(fit){
+  Vcov <- vcov(fit, useScale = FALSE)
+  betas <- fixef(fit)
+  se <- sqrt(diag(Vcov))
+  zval <- betas / se
+  pval <- round(2 * pnorm(abs(zval), lower.tail = FALSE),4)
+  pval[pval==0] <- "<0.001"
+  fe <- data.frame(Est=round(betas,3),SE= round(se,3), Zval=round(zval,2), pvalue=pval)
+  rem <- summary(fit)@REmat
+  re <- data.frame(Est=rem[,3], SE= rem[,4], Zval="",pvalue="")
+  re$Est <- round(as.numeric(as.character(re$Est)),2)
+  re$SE <- round(as.numeric(as.character(re$SE)),2)
+  rownames(re) <- paste(rem[,1],rem[,2])
+  return(rbind(fe,re))
+}
 
 
-
+dt <- subset(dtrend, experiment==5)
+f0 <- lmer(log(time_taken) ~I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), data=dt) 
[…]

### 2013-06-08 sociology_paper.tex
model fitting description added

@@
 
-\subsection{Learning Trend} Model \eqref{eqn:trend_time} is fitted to the data from experiment 5, 6, and 7.
+\subsection{Learning Trend} Model \eqref{eqn:trend_time} is fitted to the data from experiment 5, 6, and 7 separately.
+For this the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
+To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
 The results are shown in Table \ref{tbl:model_result_time}.
-The model is fitted using R function $lmer()$ from package $lme4$.
-To obtain the $p$-values of fixed effect parameters, normal approximation is used $Z$ scores are computed accordingly.
+Both fixed effect parameters of covariate Attempt ($\alpha_1$ and $\alpha$) are highly significant.
 

### 2013-06-08 R/sociology_paper.R
results of model

@@ -467,22 +467,30 @@ get_estimates <- function(fit){
   fe <- data.frame(Est=round(betas,3),SE= round(se,3), Zval=round(zval,2), pvalue=pval)
   rem <- summary(fit)@REmat
   re <- data.frame(Est=rem[,3], SE= rem[,4], Zval="",pvalue="")
-  re$Est <- round(as.numeric(as.character(re$Est)),2)
-  re$SE <- round(as.numeric(as.character(re$SE)),2)
+  re$Est <- round(as.numeric(as.character(re$Est)),3)
+  re$SE <- round(as.numeric(as.character(re$SE)),3)
   rownames(re) <- paste(rem[,1],rem[,2])
-  return(rbind(fe,re))
+  return(rbind(fe,re[order(rownames(re)),]))
 }
 
 
-dt <- subset(dtrend, experiment==5)
-f0 <- lmer(log(time_taken) ~I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), data=dt) 
-estimates <- get_estimates(f0)
-for (i in 6:7){
-  dt <- subset(dtrend, experiment==i)
-  f0 <- lmer(log(time_taken) ~I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), data=dt) 
-  estimates <- cbind(estimates,get_estimates(f0))
-}
-xtable(estimates)
+dt5 <- subset(dtrend, experiment==5)
+f5 <- lmer(log(time_taken) ~I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), data=dt5) 
+estimates5 <- get_estimates(f5)
+dt6 <- subset(dtrend, experiment==6)
+f6 <- lmer(log(time_taken) ~I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), data=dt6)
+estimates6 <- get_estimates(f6)
+dt7 <- subset(dtrend, experiment==7)
+f7 <- lmer(log(time_taken) ~I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), data=dt7)
+estimates7 <- get_estimates(f7)
+
+estimates <- data.frame(estimates5,g1=" ",estimates6,g2=" ", estimates7)
+rownames(estimates) <- c("$\\mu$","$\\alpha_1$","$\\alpha$", "$\\sigma^2_a$",
+                         "$\\sigma^2_u$","$\\sigma^2_l$","$\\sigma^2$")
+print(xtable(estimates),  sanitize.text.function = function(x){x})
+
+
+
 
[…]

### 2013-06-08 R/sociology_paper.R
time model result updated

@@ -463,7 +463,7 @@ get_estimates <- function(fit){
   se <- sqrt(diag(Vcov))
   zval <- betas / se
   pval <- round(2 * pnorm(abs(zval), lower.tail = FALSE),4)
-  pval[pval==0] <- "<0.001"
+  pval[pval==0] <- "$<$0.001"
   fe <- data.frame(Est=round(betas,3),SE= round(se,3), Zval=round(zval,2), pvalue=pval)
   rem <- summary(fit)@REmat
   re <- data.frame(Est=rem[,3], SE= rem[,4], Zval="",pvalue="")
@@ -487,8 +487,9 @@ estimates7 <- get_estimates(f7)
 estimates <- data.frame(estimates5,g1=" ",estimates6,g2=" ", estimates7)
 rownames(estimates) <- c("$\\mu$","$\\alpha_1$","$\\alpha$", "$\\sigma^2_a$",
                          "$\\sigma^2_u$","$\\sigma^2_l$","$\\sigma^2$")
-print(xtable(estimates),  sanitize.text.function = function(x){x})
-
+dgt <- c(rep(0,7), rep(3,14), rep(2,7), rep(0,7))
+dgts <- matrix(rep(dgt,3), ncol=15)
+print(xtable(estimates, digits=dgts),  sanitize.text.function = function(x){x})
 
 
 

### 2013-06-08 sociology_paper.tex
table alignment
results of model
time model result updated
model 1 results added

@@
 Both fixed effect parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.}
-\begin{tabular}{rrrcccrrcccrrcc}
+\scalebox{0.9}{
+\begin{tabular}{rrrrccrrrccrrrc}
   \hline
 & \multicolumn{4}{c} {Experiment 5} & &\multicolumn{4}{c} {Experiment 6} && \multicolumn{4}{c} {Experiment 7}\\
+
 \cline{2-5} \cline{7-10} \cline{12-15} 
- Effect& Est & SE & Zval & pvalue && Est & SE & Zval & pvalue && Est & SE & Zval & pvalue \\ 
+ Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
   \hline
 Fixed &  &  &  & && &  & & & & & &  & \\ 
-$\mu$ & 3.82 & 0.04 & 97.38 & $<$0.001 & & 3.90 & 0.03 & 118.19 & $<$0.001 && 3.73 & 0.05 & 69.04 & $<$0.001 \\ 
-  $\alpha_1$ & 0.33 & 0.04 & 9.35 & $<$0.001 && 0.34 & 0.03 & 11.4 & $<$0.001 && 0.28 & 0.05 & 5.63 & $<$0.001 \\ 
-  $\alpha$ & -0.04 & 0.00 & -9.3 & $<$0.001 && -0.04 & 0.00 & -10.19 & $<$0.001 && -0.03 & 0.01 & -4.22 & $<$0.001 \\ 
-Rand &  &  &  &  &  & & &&& & & &  & \\ 
-  $\sigma^2_l$ & 0.01 & 0.09 &  &  && 0.24 & 0.49 &  &&  & 0.13 & 0.37 &  &  \\ 
-  $\sigma^2_u$ & 0.26 & 0.51 &  &  & &0.00 & 0.05 & & &  & 0.00 & 0.05 &  &  \\ 
-   $\sigma^2_a$ & 0.00 & 0.03 &  &  & &0.04 & 0.20 & & &  & 0.06 & 0.24 &  &  \\ 
-  $\sigma^2$ & 0.21 & 0.46 &  &  & &0.25 & 0.50 &  &  && 0.21 & 0.45 &  &  \\ 
+$\mu$ & 3.817 & 0.039 & 97.38 & $<$0.001 &   & 3.901 & 0.033 & 118.19 & $<$0.001 &   & 3.731 & 0.054 & 69.04 & $<$0.001 \\ 
+  $\alpha_1$ & 0.326 & 0.035 & 9.35 & $<$0.001 &   & 0.335 & 0.029 & 11.4 & $<$0.001 &   & 0.280 & 0.050 & 5.63 & $<$0.001 \\ 
+  $\alpha$ & -0.038 & 0.004 & -9.3 & $<$0.001 &   & -0.039 & 0.004 & -10.19 & $<$0.001 &   & -0.029 & 0.007 & -4.22 & $<$0.001 \\ 
+\multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
+  $\sigma^2_a$ & 0.001 & 0.027 &  &  &   & 0.002 & 0.045 &  &  &   & 0.002 & 0.049 &  &  \\ 
+  $\sigma^2_u$ & 0.259 & 0.509 &  &  &   & 0.245 & 0.495 &  &  &   & 0.134 & 0.366 &  &  \\ 
+  $\sigma^2_l$ & 0.008 & 0.091 &  &  &   & 0.040 & 0.199 &  &  &   & 0.055 & 0.235 &  &  \\ 
+  $\sigma^2$ & 0.211 & 0.460 &  &  &   & 0.251 & 0.501 &  &  &   & 0.206 & 0.454 &  &  \\ 
    \hline
 \end{tabular}
+}
 \label{tbl:model_result_time}
+\end{table}
+
+\begin{table}[htbp]
+\centering
+\caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for log time taken to evaluate a lineup.
+None of the fixed factor effects of attempt (2-10) seem to be significantly different from each other at \%5 level in all three experiments 5, 6 and 7.}
+\scalebox{0.9}{
+\begin{tabular}{rrrrccrrrccrrrc}
+  \hline
+& \multicolumn{4}{c} {Experiment 5} & &\multicolumn{4}{c} {Experiment 6} && \multicolumn{4}{c} {Experiment 7}\\
+
+\cline{2-5} \cline{7-10} \cline{12-15} 
+ Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
+  \hline
+Fixed &  &  &  & && &  & & & & & &  & \\ 
+$\mu$ & -1.304 & 0.179 & -7.28 & $<$0.001 &   & -0.220 & 0.147 & -1.5 & 0.134 &   & -1.737 & 0.481 & -3.61 & $<$0.001 \\ 
+  $\alpha2$ & 0.270 & 0.219 & 1.24 & 0.217 &   & 0.262 & 0.158 & 1.66 & 0.098 &   & -0.456 & 0.385 & -1.18 & 0.237 \\ 
+  $\alpha3$ & -0.178 & 0.226 & -0.79 & 0.432 &   & 0.342 & 0.157 & 2.18 & 0.029 &   & -0.105 & 0.386 & -0.27 & 0.786 \\ 
+  $\alpha4$ & 0.083 & 0.224 & 0.37 & 0.712 &   & 0.358 & 0.159 & 2.26 & 0.024 &   & -0.378 & 0.381 & -0.99 & 0.322 \\ 
+  $\alpha5$ & 0.298 & 0.224 & 1.33 & 0.183 &   & 0.376 & 0.159 & 2.36 & 0.018 &   & -0.107 & 0.385 & -0.28 & 0.781 \\ 
+  $\alpha6$ & 0.042 & 0.231 & 0.18 & 0.857 &   & 0.246 & 0.158 & 1.56 & 0.12 &   & 0.026 & 0.407 & 0.06 & 0.949 \\ 
+  $\alpha7$ & 0.283 & 0.230 & 1.23 & 0.217 &   & 0.160 & 0.159 & 1.01 & 0.314 &   & 0.057 & 0.401 & 0.14 & 0.886 \\ 
+  $\alpha8$ & -0.045 & 0.233 & -0.19 & 0.847 &   & 0.341 & 0.160 & 2.13 & 0.033 &   & -0.003 & 0.394 & -0.01 & 0.994 \\ 
+  $\alpha9$ & -0.195 & 0.232 & -0.84 & 0.400 &   & 0.378 & 0.160 & 2.36 & 0.018 &   & 0.204 & 0.436 & 0.47 & 0.639 \\ 
+  $\alpha10$ & 0.513 & 0.228 & 2.25 & 0.024 &   & 0.192 & 0.163 & 1.18 & 0.238 &   & -0.213 & 0.432 & -0.49 & 0.622 \\ 
+\multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
+  $\sigma^2_a$ & $<$0.001 & 0.017 &  &  &   & 0.001 & 0.034 &  &  &   & 0.027 & 0.163 &  &  \\ 
+  $\sigma^2_u$ & 0.720 & 0.848 &  &  &   & 0.815 & 0.903 &  &  &   & $<$0.001 & $<$0.001 &  &  \\ 
+  $\sigma^2_l$ & 2.178 & 1.476 &  &  &   & 2.009 & 1.418 &  &  &   & 10.980 & 3.314 &  &  \\ 
+   \hline
+\end{tabular}
+}
+\label{tbl:model_result_response}
 \end{table}

### 2013-06-08 R/sociology_paper.R

@@ -325,6 +325,7 @@ gregexpr("\\W+", reason.correct)
 
 
 # =====================================================
+# Model fitting
 # Examining learning trend while giving feedback
 
 # function two obtain sequential attempt information from start_time
@@ -352,12 +353,17 @@ for (i in 1:7){
 
 
 # Attempt vs mean time taken for each lineup
-
-qplot(attempt, time_taken, data=subset(dtrend, time_taken< 200),colour=factor(experiment)) +
+qplot(attempt, log(time_taken), data=dtrend,colour=factor(experiment)) +
   stat_smooth(method="loess")
 
+# Log time taken appears to be normal
+qplot(log(time_taken), geom="density",
+      data=subset(dtrend, experiment %in% c(5,6,7)), color=factor(experiment)) +
+  scale_colour_hue(name="Experiment")
+
+# Modeling time taken
 # fitting linear random effect model to check 
-# if time_taken has any trend over sequential attempts
+# if log time_taken has any trend over sequential attempts
 
 dpt <- NULL
 for (i in 5:7){
@@ -397,16 +403,46 @@ ggplot() +
 
 ggsave("../images/learning_trend_time.pdf", width=10.5, height = 3.5)
 
+# Getting results of mixed effect model with log(time_taken)
 
-# Fitting generalized mixed effect model with time taken
-# Link function is inverse link (1/mu) since response is gamma
+get_estimates <- function(fit){
+  Vcov <- vcov(fit, useScale = FALSE)
+  betas <- fixef(fit)
+  se <- sqrt(diag(Vcov))
+  zval <- betas / se
+  pval <- round(2 * pnorm(abs(zval), lower.tail = FALSE),4)
+  pval[pval==0] <- "$<$0.001"
+  fe <- data.frame(Est=round(betas,3),SE= round(se,3), Zval=round(zval,2), pvalue=pval)
+  rem <- summary(fit)@REmat
+  re <- data.frame(Est=rem[,3], SE= rem[,4], Zval="",pvalue="")
+  re$Est <- round(as.numeric(as.character(re$Est)),3)
+  re$SE <- round(as.numeric(as.character(re$SE)),3)
+  rownames(re) <- paste(rem[,1],rem[,2])
[…]

### 2013-06-08 sociology_paper.tex
updated overview of the data

@@
 
-\subsection{Data collection methods}  The human observers who evaluate the experimental lineups are recruited through Amazon Mechanical Turk \cite{turk} or MTurk Web site.
+\subsection{Data collection methods}  Human observers were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \cite{turk} or MTurk Web site.
 It is an online work place where people from around the world can perform some tasks and get paid.
@@
 
-Because of its convenience it is getting popular for scientific research study.
+Because of its convenience it is getting popular for scientific research study as well.
 In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time and cost while provided more convenience.
@@
 We design and develop a web site which enables to display the lineups to the observers as per experimental design.
-The MTurk workers are redirect to the web site and the data are collected and automatically stored into a local server.
+The MTurk workers are redirect to the web site and the data are collected, stored automatically into a local database server.
 Demographic informations such as age group, gender and education levels were collected.
-The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback is received.
+The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
 The location of the observer is determined by the ip address of the observer.
@@
 
-\subsection{Overview of the Data} Figure \ref{fig:turker_location} shows the location from where we received data.
-Notice that we received data from around the world and we observed diversity of location in the collected data.
-We notice diversity in not only location of the observers but also their gender, age groups and education levels.
-Table \ref{tbl:demographics} shows that number of male and female participants are almost same.
+\subsection{Overview of the Data} Figure \ref{fig:turker_location} displays the location around the world from where a total of 2321 participants provided feedback data on the lineups in all the ten experimental studies.
+Even though most of the participants were from the United States and India, we have participants from 76 different countries.
+This provides a diverse pool of participants in terms of geographical locations.
+The diversity in not only in geographical locations of the participants but also in their gender, age groups and education levels as we see in Table \ref{tbl:demographics}.
+It is interesting that the number of male and female participants are almost same even though there are lot of people from developing countries.
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=4.5in]{turker_country.pdf} 
+   \includegraphics[width=4.5in]{turker_location.pdf}    
+   \caption{Location of the Amazon Mechanical Turk workers participating our study. Most of the people are coming from India and United States even though there are people from around the world.}
+   \label{fig:turker_location}
+\end{figure}
+
+Besides United States and India, countries such as Canada, Romania, United Kingdom and Macedonia have more than 10 participants each. The rest of the 70 countries have less than 10 participants.
 
@@
 
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=4.5in]{turker_country.pdf} 
-   \includegraphics[width=4.5in]{turker_location.pdf}    
-   \caption{Location of the Amazon Mechanical Turk workers participating our study. Most of the people are coming from India and United States even though there are people from around the world.}
-   \label{fig:turker_location}
-\end{figure}
-
-The large number of participants are from age group 18 to 35.
+The largest number of participants are from age group 18 to 25 which is the youngest age group in the study.
 Interestingly we see many participants from older age groups as well.
@@
 
-\subsection{Demographic Factors} Proportions of correct responses and the logarithm of average time taken for each lineup are computed and their distributions are shown in Figure \ref{fig:demographic_effect} for different demographic factor variables.
+\subsection{Demographic Factors} Proportions of correct responses and the logarithm of average time taken for each lineup are computed and their distributions are shown as boxplots in Figure \ref{fig:demographic_effect} for different demographic factor levels.
 
@@
 
+Observer demographics don't seem to have significant influence on performance in time taken to evaluate a lineup.
+Some variation is observed in terms of median proportion correct, but the mean proportion correct over different demographic factors levels seem to be similar.
+
 The simulation experiment reveals that the effect of location of actual data plot in the lineup is not significant.
@@
 
-More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of statistical graphics.
-The numerous pilot studies done with the more trained participants suggest that power of visual statical inference may be higher with skilled observer.
-Thus the variations observed with individual performances could be attributed to individual abilities.
+More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of knowledge about statistical graphics.
+The subjects participating this study may not necessarily knowledgable about statistical graphics.
+The numerous pilot studies done with the more trained participants suggest that power of visual statical inference may be higher with knowledgeable observer.
+Thus the variations observed with individual performances could be attributed to individual abilities in terms of their training on statistical graphics.
 

### 2013-06-08 sociology_paper.tex
discussion in demographic factors

@@
 Notice that fewer people participated from India beyond age 40 compared to united states.
-Even thoug total participations from india is larger compared to United States, they are mostly young people who are capable of using internet and computer.
-
-\subsection{Demographic Factors} Proportions of correct responses and the logarithm of average time taken for each lineup are computed and their distributions are shown as boxplots in Figure \ref{fig:demographic_effect} for different demographic factor levels.
+Even though total participations from india is larger compared to United States, they are mostly young people who are capable of using internet and computer.
+
+Most of the participants are graduate degree holder.
+They are mostly from India as any university degree is considered a graduate degree unlike north America where graduate degree means Masters level education.
+This is apparent in Figure \ref{fig:demographic_info}.
+Most of the unites states participants are with an Undergraduate degree or at least have some undergraduate course.
+
+\subsection{Demographic Factors} Proportions of correct responses and the logarithm of average time taken to evaluate  each lineup are computed and their distributions are shown as boxplots in Figure \ref{fig:demographic_effect} for different demographic factor levels.
+The youngest age group (18-25) took least average time to evaluate a lineup.
+People from India took more time than others.
+People who have only a high school degree took less time and this is related to what we have seen for young age group.
+The time taken by male and female participants are almost similar.
+None of the factors are significant in describing time taken to evaluate a lineup.
 
@@
 \end{figure}
+
+We observed some variations in median proportion of correct responses for different levels of the demographic factors as in Figure \ref{fig:demographic_effect}.
+But the variability of proportion correct responses are huge as per design of the experiments since we have some easy and some difficult lineups.
+There were lineups which were not expected to be evaluated correctly and some were very easy and expected to be evaluated correctly 100\% of the time.
 

### 2013-06-09 sociology_paper.tex
some allignment in the table
factor effects updated

@@
 
-There have been \dc{ten} experiments conducted using Amazon's Mechanical Turk~\cite{turk} that are being used to evaluate the the effect of the observer.
+There have been ten experiments conducted using Amazon's Mechanical Turk~\cite{turk} that are being used to evaluate the the effect of the observer.
 Table \ref{tbl:visual_stat} summarizes these.
@@
 
-The visual test statistic is chosen so that it displays a specific patter in case the null hypothesis not true.
+The visual test statistic is chosen so that it displays a specific pattern in case the null hypothesis is not true.
 Thus the most important factor that help an observer to correctly evaluate a lineup is the presence of any detectable signal in the data.
@@
 
+Table \ref{tbl:visual_stat} shows the visual test statistics used in this study.
+For both experiment 2 and 3 a scatterplot is used but a regression line fitted through the points is overlaid for experiment 2.
+It is easier to spot the slope of the line compared to just the scatterplot itself.
+But it appears that people are better in noticing the unusual pattern in the data which is some contamination purposefully added for experiment 3.
+It is observed that for experiment 3 the performance of the observers are much better compared to experiment 2 \citep{majumder:2013}.
+
 %Plots to do: for the same signal strength show the proportion correct for scatterplot and boxplot. eg., qplot(effect, ump-visual power, color= plot-type)
@@
 These questions are very crucial for the power of visual test.
-They help observer think in a controlled direction.
+They help observer think in a meaningful direction.
 Notice that there may be unnecessary patterns in the actual data plot which may not necessarily indicate the existence of the significant signal in the plot.
@@
 For example, a high school student may not respond same as a well educated person while evaluating a lineup.
-Also, to meet the Institutional Regulatory Board (IRB) requirement is is necessary to exclude any subject less than 18 years of old to participate human subject experiment.
+There may be variation in different age groups as well.
+Recruiting subjects with a well variety of age groups is be necessary to study this variability.
+
+To meet the Institutional Regulatory Board (IRB) requirement, it is necessary to exclude any subject less than 18 years of old to participate human subject experiment.
 Thus we intend to only focus on the performance of observer with age 18 years or older.
-But still there may be variation in different age groups.
-Recruiting subjects with a well variety of age groups may be necessary to have control over this variability.
+We don't expect many older people to participate our study and we are interested of age from 18 to 65 years.
 
@@
 
-\subsection{Learning Trend of the Observer} When an observer evaluates a lineup he or she may learn something from the experience.
+We don't expect much difference in performances between male and female observers.
+Gender may have some influence through the education level and geographical locations since some location may have small number of educated female population.
+We intend to include similar number of male and female subjects in our study so that this can be examined.
+
+\subsection{Learning Trend of the Observer} When an observer evaluates a lineup he or she may learn something from the experiences of the earlier observations.
 If the same observer is shown another lineup the learning from previous evaluation may help.
@@
 This learning trend may or may not be significant overall but it could influence the performance of the observer in some way.
-This paper investigates the learning trend with multiple simulation experiments.
+
+Learning may occur in two different ways.
+An observer can become more trained on the lineup structure and the pattern in the plots shown.
+This could be learning in proportion of correct evaluation.
+Or, he or she may become efficient in responding faster in the later evaluations.
+This could be observed as the time taken for evaluating a lineup.
+This paper investigates both of these learning trends with multiple simulation experiments.
+
+When an observer evaluate a lineup for the first time, he or she has to read through many instructions and become familiar with the environment of the experiment.
+We expect that this would require much more time compared to the rest of the evaluations.
+Time taken for the second or successive attempt would be much lower than the time taken for the first lineup.
+We attribute this to the adaptation to the experiment environment, not a learning of evaluating a lineup.
 
@@
 
-We design and develop a web site which enables to display the lineups to the observers as per experimental design.
-The MTurk workers are redirect to the web site and the data are collected, stored automatically into a local database server.
-Demographic informations such as age group, gender and education levels were collected.
+We designed and developed a web site which enables to display the lineups to the observers as per experimental design.
+The MTurk workers were redirected to the web site and the data were collected, stored automatically into a local database server.
+Demographic informations such as age group, gender and education levels were also collected.
 The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
@@
 $\mu$ & 3.817 & 0.039 & 97.38 & $<$0.001 &   & 3.901 & 0.033 & 118.19 & $<$0.001 &   & 3.731 & 0.054 & 69.04 & $<$0.001 \\ 
-  $\alpha_1$ & 0.326 & 0.035 & 9.35 & $<$0.001 &   & 0.335 & 0.029 & 11.4 & $<$0.001 &   & 0.280 & 0.050 & 5.63 & $<$0.001 \\ 
-  $\alpha$ & -0.038 & 0.004 & -9.3 & $<$0.001 &   & -0.039 & 0.004 & -10.19 & $<$0.001 &   & -0.029 & 0.007 & -4.22 & $<$0.001 \\ 
+  $\alpha_1$ & 0.326 & 0.035 & 9.35 & $<$0.001 &   & 0.335 & 0.029 & 11.40 & $<$0.001 &   & 0.280 & 0.050 & 5.63 & $<$0.001 \\ 
+  $\alpha$ & -0.038 & 0.004 & -9.30 & $<$0.001 &   & -0.039 & 0.004 & -10.19 & $<$0.001 &   & -0.029 & 0.007 & -4.22 & $<$0.001 \\ 
 \multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
@@
 \caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for log time taken to evaluate a lineup.
-None of the fixed factor effects of attempt (2-10) seem to be significantly different from each other at \%5 level in all three experiments 5, 6 and 7.}
+None of the fixed factor effects of attempt (2-10) seem to be significantly different from each other at \%5 level in all three experiments 5, 6 and 7.
+For experiment 7 subject specific variation is very small on the other hand lineup variance is much higher compared to the other two experiments.}
 \scalebox{0.9}{

### 2013-06-09 sociology_paper.tex
resuls section updated

@@
 
-To meet the Institutional Regulatory Board (IRB) requirement, it is necessary to exclude any subject less than 18 years of old to participate human subject experiment.
+To meet the Institutional Review Board (IRB) requirement, it is necessary to exclude any subject less than 18 years of old to participate human subject experiment.
 Thus we intend to only focus on the performance of observer with age 18 years or older.
@@
 This is apparent in Figure \ref{fig:demographic_info}.
-Most of the unites states participants are with an Undergraduate degree or at least have some undergraduate course.
+Most of the unites states participants are with an Undergraduate degree or at least have some undergraduate courses.
+
+Male and female participants are similarly distributed among all age groups except age 18-25 in India where less female participants were observed as we see in Figure \ref{fig:gender_country_bar}.
+Most of the participants from India are below age 40 while in United States the distribution of participants are similar beyond age 40.
+
+A total of XX lineups were evaluated in all these 10 experiments.
+Each person was supposed to evaluate 10 lineups except experiment 9 where three lineups were evaluated by each person.
+To clean the data we used the test plot as the basis and did not include the results of that test plot.
+In some cases the demographic information were not provided by the participants.
+Also, for some ip address, the actual geographical locations could not be retrieved.
+This resulted some missing demographic information.
 
@@
 There were lineups which were not expected to be evaluated correctly and some were very easy and expected to be evaluated correctly 100\% of the time.
+This is why we see the boxes ranges from 0 to 1 in most of the factor levels.
+
+The averages of proportion correct are for different factor levels are shown as dots in Figure \ref{fig:demographic_effect}.
+Unlike median, less variations are observed in the averages for different factor levels.
+As expected, there is no significant difference between Male and female performances.
+Among all the demographic factor levels only graduate degree, India and age group 31-35 appear to be significantly different from others.
+Most of the Indian participants reported to have a graduate degree.
+Thus the effect of graduate degree is actually the effect of India.
 
@@
 %------------------------------------------------------------------------------------
-This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the internal review board IRB 10-347.
+This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
 
@@
 For United States number of participants are similar beyond age 40 while few number of participants coming from India beyond that age.}
-   \label{fig:gender_country}
+   \label{fig:gender_country_bar}
 \end{figure}
@@
 }
-   \label{fig:gender_country}
+   \label{fig:gender_country_time}
 \end{figure}
@@
 }
-   \label{fig:gender_country}
+   \label{fig:gender_country_response}
 \end{figure}

### 2013-06-09 sociology_paper.tex
discussion on model 1 result

@@
 Male and female participants are similarly distributed among all age groups except age 18-25 in India where less female participants were observed as we see in Figure \ref{fig:gender_country_bar}.
+The distribution of education levels are also different across the countries for this age group.
 Most of the participants from India are below age 40 while in United States the distribution of participants are similar beyond age 40.
@@
 \subsection{Demographic Factors} Proportions of correct responses and the logarithm of average time taken to evaluate  each lineup are computed and their distributions are shown as boxplots in Figure \ref{fig:demographic_effect} for different demographic factor levels.
+Averages of these distributions are represented by dots inside the boxplots.
 The youngest age group (18-25) took least average time to evaluate a lineup.
@@
 
-The averages of proportion correct are for different factor levels are shown as dots in Figure \ref{fig:demographic_effect}.
 Unlike median, less variations are observed in the averages for different factor levels.
@@
 
-\subsection{Learning Trend} Model \eqref{eqn:trend_time} is fitted to the data from experiment 5, 6, and 7 separately.
+\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiment 5, 6, and 7 separately.
+For this the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
+To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
+
+Table \ref{tbl:model_result_response} presents the parameter estimates and $p$-values of fixed effect estimates of Model \eqref{eqn:trend_response}.
+The larger $p$-values suggest that none of the levels of attempt ($\alpha_2$ through $\alpha_{10}$) appears to be significant at \%1 significance level.
+Moreover, some of the estimates are positive and some are negative and they show up seemingly in random order suggesting later attempts not necessarily improved.
+This indicates that there may be no learning effect on the proportion of correct evaluations.
+
+\begin{table}[htbp]
+\centering
+\caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for log time taken to evaluate a lineup.
+None of the fixed factor effects of attempt ($\alpha_2$ through $\alpha_{10}$) seem to be significantly different from each other at \%1 level in all three experiments 5, 6 and 7.
+For experiment 7 subject specific variation is very small on the other hand lineup variance is much higher compared to the other two experiments.}
+\scalebox{0.9}{
+\begin{tabular}{rrrrccrrrccrrrc}
+  \hline
+& \multicolumn{4}{c} {Experiment 5} & &\multicolumn{4}{c} {Experiment 6} && \multicolumn{4}{c} {Experiment 7}\\
+
+\cline{2-5} \cline{7-10} \cline{12-15} 
+ Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
+  \hline
+Fixed &  &  &  & && &  & & & & & &  & \\ 
+$\mu$ & -1.304 & 0.179 & -7.28 & $<$0.001 &   & -0.220 & 0.147 & -1.5 & 0.134 &   & -1.737 & 0.481 & -3.61 & $<$0.001 \\ 
+  $\alpha_2$ & 0.270 & 0.219 & 1.24 & 0.217 &   & 0.262 & 0.158 & 1.66 & 0.098 &   & -0.456 & 0.385 & -1.18 & 0.237 \\ 
+  $\alpha_3$ & -0.178 & 0.226 & -0.79 & 0.432 &   & 0.342 & 0.157 & 2.18 & 0.029 &   & -0.105 & 0.386 & -0.27 & 0.786 \\ 
+  $\alpha_4$ & 0.083 & 0.224 & 0.37 & 0.712 &   & 0.358 & 0.159 & 2.26 & 0.024 &   & -0.378 & 0.381 & -0.99 & 0.322 \\ 
+  $\alpha_5$ & 0.298 & 0.224 & 1.33 & 0.183 &   & 0.376 & 0.159 & 2.36 & 0.018 &   & -0.107 & 0.385 & -0.28 & 0.781 \\ 
+  $\alpha_6$ & 0.042 & 0.231 & 0.18 & 0.857 &   & 0.246 & 0.158 & 1.56 & 0.12 &   & 0.026 & 0.407 & 0.06 & 0.949 \\ 
+  $\alpha_7$ & 0.283 & 0.230 & 1.23 & 0.217 &   & 0.160 & 0.159 & 1.01 & 0.314 &   & 0.057 & 0.401 & 0.14 & 0.886 \\ 
+  $\alpha_8$ & -0.045 & 0.233 & -0.19 & 0.847 &   & 0.341 & 0.160 & 2.13 & 0.033 &   & -0.003 & 0.394 & -0.01 & 0.994 \\ 
+  $\alpha_9$ & -0.195 & 0.232 & -0.84 & 0.400 &   & 0.378 & 0.160 & 2.36 & 0.018 &   & 0.204 & 0.436 & 0.47 & 0.639 \\ 
+  $\alpha_{10}$ & 0.513 & 0.228 & 2.25 & 0.024 &   & 0.192 & 0.163 & 1.18 & 0.238 &   & -0.213 & 0.432 & -0.49 & 0.622 \\ 
+\multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
+  $\sigma^2_a$ & $<$0.001 & 0.017 &  &  &   & 0.001 & 0.034 &  &  &   & 0.027 & 0.163 &  &  \\ 
+  $\sigma^2_u$ & 0.720 & 0.848 &  &  &   & 0.815 & 0.903 &  &  &   & $<$0.001 & $<$0.001 &  &  \\ 
+  $\sigma^2_l$ & 2.178 & 1.476 &  &  &   & 2.009 & 1.418 &  &  &   & 10.980 & 3.314 &  &  \\ 
+   \hline
+\end{tabular}
+}
+\label{tbl:model_result_response}
+\end{table}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=6.5in]{learning_trend.pdf} 
+   \caption{ Mean residuals of proportion correct obtained from the model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
+Linear regression lines fitted through the points are shown.
+None of the slopes in these three experiments are statistically significant.}
+   \label{fig:learning_trend_response}
+\end{figure}
+
+The estimated variance ($\sigma^2_a$) of the random effect attempt is much small compared to other random effects except for subject variability ($\sigma^2_u$) for experiment 7.
+This suggest that most of the variability for attempt is accounted for by the fixed effect factors of attempt.
+For experiment 7 the lineup variability ($\sigma^2_l$) was the largest as per experimental design.
+Some of the lineups were very easy and some were really difficult in experiment 7.
+The overall average probability for picking out the data plot from a lineup ($\mu$) appears to be significant for both experiments 5 and 7.
+But for experiment 6 it does not seem to be significant suggesting that experiment 6 lineups were difficult compared to the other experimental lineups.
+This shows the designs of experiments such that there exists the mixture of difficult and easy lineups within the experiments as well as between the experiments.
+For easy lineups there may be small chance to improve performances.
+But for harder lineups, the scope to improve performances is large.
+Since for none of these experiments improvement in performances is observed, it is important in the sense that performances did not improve over attempts in both difficult and easy lineup situations.
+
+%The results are shown in Tables \ref{tbl:model_result_response} \ref{tbl:model_result_time}.
+
+%Model \eqref{eqn:trend_time} is fitted to the data from experiment 5, 6, and 7 separately.
 For this the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
@@
 But to obtain this we have to adjust for lineup difficulty and individual performances.
+
+To visualize further how the performance in correct response improves over successive attempt, we fitted Model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
+The averages of these residuals for each of the attempts are plotted against attempt as shown in Figure \ref{fig:learning_trend_response}.
+Least square linear regression lines are drown through the points.
+Even though positive slopes over the attempts are observed, none of the trends is statistically significant.
 
@@
 
-\begin{table}[htbp]
-\centering
-\caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for log time taken to evaluate a lineup.
-None of the fixed factor effects of attempt (2-10) seem to be significantly different from each other at \%5 level in all three experiments 5, 6 and 7.
-For experiment 7 subject specific variation is very small on the other hand lineup variance is much higher compared to the other two experiments.}
-\scalebox{0.9}{
-\begin{tabular}{rrrrccrrrccrrrc}
-  \hline
-& \multicolumn{4}{c} {Experiment 5} & &\multicolumn{4}{c} {Experiment 6} && \multicolumn{4}{c} {Experiment 7}\\
-
-\cline{2-5} \cline{7-10} \cline{12-15} 
- Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
-  \hline
-Fixed &  &  &  & && &  & & & & & &  & \\ 
-$\mu$ & -1.304 & 0.179 & -7.28 & $<$0.001 &   & -0.220 & 0.147 & -1.5 & 0.134 &   & -1.737 & 0.481 & -3.61 & $<$0.001 \\ 
-  $\alpha2$ & 0.270 & 0.219 & 1.24 & 0.217 &   & 0.262 & 0.158 & 1.66 & 0.098 &   & -0.456 & 0.385 & -1.18 & 0.237 \\ 
-  $\alpha3$ & -0.178 & 0.226 & -0.79 & 0.432 &   & 0.342 & 0.157 & 2.18 & 0.029 &   & -0.105 & 0.386 & -0.27 & 0.786 \\ 
-  $\alpha4$ & 0.083 & 0.224 & 0.37 & 0.712 &   & 0.358 & 0.159 & 2.26 & 0.024 &   & -0.378 & 0.381 & -0.99 & 0.322 \\ 
-  $\alpha5$ & 0.298 & 0.224 & 1.33 & 0.183 &   & 0.376 & 0.159 & 2.36 & 0.018 &   & -0.107 & 0.385 & -0.28 & 0.781 \\ 
-  $\alpha6$ & 0.042 & 0.231 & 0.18 & 0.857 &   & 0.246 & 0.158 & 1.56 & 0.12 &   & 0.026 & 0.407 & 0.06 & 0.949 \\ 
-  $\alpha7$ & 0.283 & 0.230 & 1.23 & 0.217 &   & 0.160 & 0.159 & 1.01 & 0.314 &   & 0.057 & 0.401 & 0.14 & 0.886 \\ 
-  $\alpha8$ & -0.045 & 0.233 & -0.19 & 0.847 &   & 0.341 & 0.160 & 2.13 & 0.033 &   & -0.003 & 0.394 & -0.01 & 0.994 \\ 
-  $\alpha9$ & -0.195 & 0.232 & -0.84 & 0.400 &   & 0.378 & 0.160 & 2.36 & 0.018 &   & 0.204 & 0.436 & 0.47 & 0.639 \\ 
-  $\alpha10$ & 0.513 & 0.228 & 2.25 & 0.024 &   & 0.192 & 0.163 & 1.18 & 0.238 &   & -0.213 & 0.432 & -0.49 & 0.622 \\ 
-\multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
-  $\sigma^2_a$ & $<$0.001 & 0.017 &  &  &   & 0.001 & 0.034 &  &  &   & 0.027 & 0.163 &  &  \\ 
-  $\sigma^2_u$ & 0.720 & 0.848 &  &  &   & 0.815 & 0.903 &  &  &   & $<$0.001 & $<$0.001 &  &  \\ 
-  $\sigma^2_l$ & 2.178 & 1.476 &  &  &   & 2.009 & 1.418 &  &  &   & 10.980 & 3.314 &  &  \\ 
-   \hline
-\end{tabular}
-}
-\label{tbl:model_result_response}
-\end{table}
-
 To accomplish this we fit generalized mixed effect model \eqref{eqn:trend_response} with lineups and subjects as random effect and no fixed effect.
@@
 But none of apparent trend are statistically significant.
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=6.5in]{learning_trend.pdf} 
-   \caption{ Mean residuals of proportion correct obtained from the model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
-Linear regression lines fitted through the points are shown.
-None of the slopes in these three experiments are statistically significant.}
-   \label{fig:learning_trend}
-\end{figure}
 

### 2013-06-09 sociology_paper.tex
model 3 results discussion

@@
 
+Table \ref{tbl:model_result_time} presents the results of Model \eqref{eqn:trend_time}.
+The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments.
+The negative estimates suggest that later attempt took less time for an evaluation.
+This suggest that even though observers did not improve the performance over attempt, they became efficient in responding faster.
+The parameter $\alpha_1$ for first attempt is also highly significant.
+The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more times than other attempts.
+It is because for initial attempt the observer might have gone through instruction and became familiar with the experimental environment.
+
 \begin{table}[htbp]
@@
 
-To accomplish this we fit generalized mixed effect model \eqref{eqn:trend_response} with lineups and subjects as random effect and no fixed effect.
+%To accomplish this we fit generalized mixed effect model \eqref{eqn:trend_response} with lineups and subjects as random effect and no fixed effect.
 This gives the estimation of proportion correct responses  and we obtain residuals of the fitted model.

### 2013-06-09 sociology_paper.tex
discussion on location effect updated

@@
 Because each poll has a margin of error, this is used to simulate different scenarios that might have resulted on election day, if the polls were on target.
-A null data set is then generate as a set of draws from normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
+A null data set is then generated as a set of draws from normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
 These samples are plotted as electoral buildings, and the plot made with the results from the election is placed randomly among them in a lineup of size $5\times 4$.
-If the null hypothesis is true then the actual data plot should look just like any of the other plots, and not be identified by an observer.
+If the null hypothesis is true the actual data plot should look just like any of the other plots, and not be identified by an observer.
  \hh{Figure~\ref{fig:tower} shows a plot of the electoral building with added context information and labels.}
@@
 Some people start looking from the center of the lineup while others start from the top left corner.
-In the earlier phase of the exercise, the observer tend to scan the plots and then start comparing plots to make a final decision.
+In the earlier phase of the exercise, the observer tend to scan the plots and start comparing plots to make a final decision.
 Beside right-left or up-down directions observer show some diagonal movement too.
@@
 This should give the observer more time to compare it with the rest of the null plots.
-Those who start looking from the center may then scan towards right or left direction and thus don't have the chance to scan the null plots continuously as they could if they would start from a corner.
+Those who start looking from the center may scan towards right or left direction and thus don't have the chance to scan the null plots continuously as they could if they would start from a corner.
 Thus they should scan the center plot over and over again while scanning the left or right side plots.
@@
 \paragraph{Design for Location effect:} The experiment was set up based on the gene expression data with two groups.
-The groups are defined based on the presence or absence of iron sufficiency.
-The null hypothesis is that there is no difference between this two groups.
-Null plots were obtained by randomizing the group structure of the data.
-The actual plot was randomly placed in five different locations.
-For each location, five different sets of null plots were used to produce five different lineups.
-This produced 25 different lineups.
-Similar procedure was applied to obtain 25 lineups to test the interaction between Genotype and Empty Vector (EV).
+The groups are defined based on the Genotype and Empty Vector (EV).
+For each of the of the Genotype, gene expression data were collected in presence or absence of iron sufficiency.
+Two hypothesis were of primary interest.
+The Genotype effect and the Genotype and iron sufficiency interaction effect.
+
+Null plots were generated by randomizing the group structure of the data.
+The actual plot was randomly placed in five different locations in a lineup of size 20.
+The locations are 2,9,12,16,20 for Interaction effect and 1,8,12,17,20 for Genotype effect.
+For a lineup with specific data plot location, five different sets of null plots were used to produce 5 lineups for each location.
+In total we have 25 lineups for Interaction effect and 25 lineups for Genotype effect.
+
+%We set up a Turk experiment to examine whether there is any difference of performance based on the location of actual data plot in the lineup.
+For this five locations of a lineup were randomly chosen to put the same actual data plot.
+The locations are 2,9,12,16,20 for Interaction effect and 1,8,12,17,20 for Genotype effect.
+For a lineup with specific data plot location, five sets of null plots are used giving 5 lineups for each location.
+In total we have 25 lineups for Interaction effect and 25 lineups for Genotype effect.
+These 50 lineups are evaluated by 100 people recruited from Amazon Mechanical Turk.
+Each person evaluated 3 lineups one from Interaction, one from Genotype and the third one was a test lineup.
+The test lineup is used to process the payment and examine the quality of the data and the responses of the test plot is not added to our analysis.
 
@@
 The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more times than other attempts.
-It is because for initial attempt the observer might have gone through instruction and became familiar with the experimental environment.
+It is because for initial attempt the observer might have gone through instructions and became familiar with the experimental environment.
+Each page in the web site contains information about choice reason and the observers's confidence level.
+Also, the first page asks for observer Identification to be typed.
+The later pages of the web site was similar just the lineup was changed.
+Thus in the later attempts an observer does not need to spend any time for reading instructions.
+The model reflects that fact.
 
@@
 
+The lineup variance of experiment 7 is estimated as $\hat{\sigma}^2 = 10.98$ from Model \eqref{eqn:trend_response}.
+But from Model \eqref{eqn:trend_time} it is estimated to be much smaller in all these experiments.
+This suggest that the harder lineups not necessarily took more time than easier lineups.
+The observers spent enough times to evaluate a easy lineup and for difficult lineup they might just give up at some point and gave the feedback.
+
+To visualize how the time taken reduces over the successive attempts, we fitted Model \eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
+The averages of these residuals for each attempt are plotted in Figure \ref{fig:learning_trend_time} for all three experiments.
+Least square regression lines are fitted to the points excluding the first attempt.
+The downward trends are evident in the plots.
+All the slopes are highly significant.
+For attempt 1, we observed large positive residuals for each of the experiments.
+These results indicate that after working with the multiple tasks each observer became efficient which helped them finish the task faster in later attempts.
+
 %To accomplish this we fit generalized mixed effect model \eqref{eqn:trend_response} with lineups and subjects as random effect and no fixed effect.
@@
 
-A plot of residuals against the attempts $A_k$ should reveal the trend if there is any.
+%A plot of residuals against the attempts $A_k$ should reveal the trend if there is any.
 Figure \ref{fig:learning_trend} shows such plots for three experiments naming experiment 5, 6 and 7.
@@
 
-Learning trend is not only related to the proportion correct responses.
+%Learning trend is not only related to the proportion correct responses.
 It may be possible that people learn to answer fast.
@@
 
-We set up a Turk experiment to examine whether there is any difference of performance based on the location of actual data plot in the lineup.
-For this five locations of a lineup were randomly chosen to put the same actual data plot.
-The locations are 2,9,12,16,20 for Interaction effect and 1,8,12,17,20 for Genotype effect.
-For a lineup with specific data plot location, five sets of null plots are used giving 5 lineups for each location.
-In total we have 25 lineups for Interaction effect and 25 lineups for Genotype effect.
-These 50 lineups are then evaluated by 100 people recruited from Amazon Mechanical Turk.
-Each person evaluated 3 lineups one from Interaction, one from Genotype and the third one was a test lineup.
-The test lineup is used to process the payment and examine the quality of the data and the responses of the test plot is not added to our analysis.
-
-The proportion of correct responses by the turk observers are shown in Figure \ref{fig:location_effect}.
-Notice that the data plot is same for each location but we see some variability of performance based on different null plots.
-This may happen if some null plots appear to be more similar to the actual plot while the others are not making some lineups harder than others even though the actual data plot is same.
-This pattern is evident in the Figure \ref{fig:location_effect} as we see proportion correct for null plot 5 is consistently above the null plot 1 for each of the locations.
-But this variability could be controlled at some extent by collecting more data for each location.
+The proportion of correct responses for each data plot location is shown in Figure \ref{fig:location_effect}.
+We observe some variability of performance for different sets of null plots even though same data plot was used for all these null sets.
+This may happen when for some set of null plots, one null plot appears to be more similar to the actual plot.
+In another set of null plots this may not happen making some lineups easier than others even though the actual data plot is the same.
+This pattern is evident in the figure as we see proportion correct for null plot 5 is consistently above the null plot 1 for each of the locations.
+The overall average proportions correct responses for each location are shown using dashed lines.
+As we see in the plot  the overall proportion is not varying much for different locations.
+
+The sizes of the dots in  Figure \ref{fig:location_effect} represents the number of responses obtained for that location and null set.
+For some locations we have as many as 10 responses.
+For location 1, we did not have any response for null set 1 in one of the interaction lineups.
+The larger variability observed for Interaction lineups could be controlled at some extent by collecting more data for each location.
 Since the variability is larger when we have small number of responses as we see in the figure.
@@
 
-It is also interesting to check whether the proportion correct differs if the actual plot is on the outer boundary of the lineup or inside the outer boundary.
+We fitted Model \eqref{manova} to test if the mean performance vectors are similar for different locations.
+For this we use {\it anova()} function of {\it stats} package of \cite{R:2012}.
+The results are shown in Table \ref{tbl:manova}.
+The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05 and we failed to reject the hypothesis that there is no difference in location.
+
+We also examined whether the proportion of correct responses differs if the actual plot is on the outer boundary or in the inner locations.
+The locations 7,8,9,12,13,14 in the lineup are considered inner locations and the rest of the boundary locations are considered to be outer location.
 As we see in Figure \ref{fig:location_effect}, location 9, 12 are inside for Interaction effect and location 8, 12 are inside for Genotype.
 It does not appear to have any difference whether the actual plot is inside or outer border of the lineup.
-
-The overall proportion of correct response does not seem to vary much for different location indicating the non-existence of location effect.
-Model \eqref{manova} is used to test if the mean vectors are similar for different locations.
-To fit the model we use anova function of stats package of \cite{R:2012}.
-The results are shown in Table \ref{tbl:manova}.
-The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05 and we failed to reject the hypothesis that there is no difference in location.
+We also fitted Model \eqref{manova} with two locations, inner and outer, as covariate and observed no significant differences.
 
@@
 
-We also fit Model \eqref{manova} with only two locations based on outer or inner plot location.
-By outer location we mean the data plot location which is on the border area of the lineup.
-For example inner locations of a lineup would be 7,8,9,12,13,14.
-It turns out that inner and outer locations are also not significantly different.
-
+%
 %% latex table generated in R 2.15.0 by xtable 1.7-0 package
@@
 %\end{table}
-
+%
 % \footnote{Signif. codes: 0 Ô***Õ 0.001 Ô**Õ 0.01 Ô*Õ 0.05 Ô.Õ 0.1}
-
-%------------------------------------------------------------------------------------
-\paragraph{Acknowledgments}
-%------------------------------------------------------------------------------------
-This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
 
@@
 The future research may involve doing more experiments to check if fewer or more than 10 lineups have any significant impact on performance of the observer.
+The learning trend in terms of time taken shows downward trends.
+At some points it should level of.
+It would be useful to know for how many lineups time taken levels off, since that would indicate the limit beyond which no learning occurs.
 
@@
 
-\section{Some figures and tables (for review)}
+%------------------------------------------------------------------------------------
+\paragraph{Acknowledgments}
+%------------------------------------------------------------------------------------
+This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
+
+\section{Some figures and tables (May go to Appendix)}
 

### 2013-06-09 R/sociology_paper.R

@@ -112,7 +112,7 @@ for (i in 1:10){
   di$academic_study[di$academic_study==0] <- NA
   di$experiment = paste("experiment_",i, sep="")
   di$id = paste("exp",i,"_",di$id, sep="")
-  degree <- factor(c("High school or less", "Some under graduate courses",
+  degree <- factor(c("High school or less", "Some under grad courses",
                   "Under graduate degree","Some graduate courses",
                   "Graduate degree")[di$academic_study])
   di$degree <- factor(degree, levels=levels(degree)[order(c(5,1,4,2,3))], ordered=T)
@@ -132,7 +132,7 @@ demographics$country[(demographics$country_code != "IN") & (demographics$country
 demographics$country[demographics$country=="Namibia"] <- "Rest of the world"
 demographics$country[demographics$country==NA] <- "Rest of the world"
 levels(demographics$age_level)[7:9] <- 7
-levels(demographics$age_level)[1:7] <- c(levels(turker$age_level)[1:6], "above 50")
+levels(demographics$age_level)[1:7] <- c(levels(demographics$age_level)[1:6], "above 50")
 
 
 # getting unique participants for plotting purpose only
@@ -192,9 +192,6 @@ qplot(degree, per_correct, geom="bar", stat="identity", data=pdat[complete.cases
 ggsave("../images/age_gender_within_country_correct.pdf", width=10, height=8)
 
 # Demographic factor main effect for percent correct and average time aggregated by lineup
-gdat <- ddply(demographics, .(gender_level, pic_name), summarize,
-              prop_correct = mean(response),
-              avg_time = mean(time_taken))
 get_effect <- function(dat, var){
   res <- ddply(dat,c(var), summarize,
                log_avg_time = round(log(mean(time_taken)),2),
@@ -211,11 +208,12 @@ adat <- get_effect(demographics, c("age_level", "pic_name"))
 
 mdat <- melt(rbind(gdat,edat,cdat,adat), id=c("variable_level", "pic_name", "variable_name"))
 
-qplot(variable_level, value, geom="boxplot", data=mdat[complete.cases(mdat),]) +
+qplot(variable_level, value, geom="boxplot",data=mdat[complete.cases(mdat),]) +
   facet_grid(variable~variable_name, scales="free") +
   stat_summary(fun.y=mean, geom="point") + xlab("Levels of demographic factors")+
   theme(axis.text.x=element_text(angle=90, hjust=1))
[…]

### 2013-06-10 sociology_paper.tex
caption of table 3

@@
 (Although this example is more simplistic than most of the tests conducted to date, it will serve the purpose of illustrating the lineup protocol.) The data is looking at the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
-Each panel in Figure~\ref{fig:elect-1} shows an `electoral building' \citet{nytimes} where each state in the union is represented by a rectangle.
+Each panel in Figure~\ref{fig:elect-1} shows an `electoral building' \citep{nytimes} where each state in the union is represented by a rectangle.
 This difference is plotted horizontally, and the height of each box corresponds to the state's electoral votes.
@@
 A lineup can be evaluated by a single person or multiple observers.
-A \dc{multinomial} distribution is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
+A binomial distribution is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
 Observers should not be aware of the data that constitutes a lineup, and should not have seen the actual data plot before seeing the lineup.
@@
 
-There have been ten experiments conducted using Amazon's Mechanical Turk~\cite{turk} that are being used to evaluate the the effect of the observer.
+There have been ten experiments conducted using Amazon's Mechanical Turk~\citep{turk} that are being used to evaluate the the effect of the observer.
 Table \ref{tbl:visual_stat} summarizes these.
@@
 \centering
-\caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for log time taken to evaluate a lineup.
-None of the fixed factor effects of attempt ($\alpha_2$ through $\alpha_{10}$) seem to be significantly different from each other at \%1 level in all three experiments 5, 6 and 7.
-For experiment 7 subject specific variation is very small on the other hand lineup variance is much higher compared to the other two experiments.}
+\caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for \hh{probability of correct lineup evaluation.} %log time taken to evaluate a lineup. 
+None of the fixed factor effects of attempt ($\alpha_2$ through $\alpha_{10}$) seem to be significantly different from %each other 
+\hh{the first attempt $\alpha_1$}
+at \%1 level in all three experiments 5, 6 and 7. For experiment 7 subject specific variation is very small on the other hand lineup variance is much higher compared to the other two experiments.}
 \scalebox{0.9}{
@@
 Fixed &  &  &  & && &  & & & & & &  & \\ 
-$\mu$ & -1.304 & 0.179 & -7.28 & $<$0.001 &   & -0.220 & 0.147 & -1.5 & 0.134 &   & -1.737 & 0.481 & -3.61 & $<$0.001 \\ 
+$\mu$ & -1.304 & 0.179 & -7.28 & $<$0.001 &   & -0.220 & 0.147 & -1.50 & 0.134 &   & -1.737 & 0.481 & -3.61 & $<$0.001 \\ 
+  $\alpha_1$ & 0.000 & --- & --- & --- &   & 0.000 & --- & --- & --- &  & 0.000 & --- & --- & ---  \\ 
   $\alpha_2$ & 0.270 & 0.219 & 1.24 & 0.217 &   & 0.262 & 0.158 & 1.66 & 0.098 &   & -0.456 & 0.385 & -1.18 & 0.237 \\ 
@@
   $\alpha_5$ & 0.298 & 0.224 & 1.33 & 0.183 &   & 0.376 & 0.159 & 2.36 & 0.018 &   & -0.107 & 0.385 & -0.28 & 0.781 \\ 
-  $\alpha_6$ & 0.042 & 0.231 & 0.18 & 0.857 &   & 0.246 & 0.158 & 1.56 & 0.12 &   & 0.026 & 0.407 & 0.06 & 0.949 \\ 
+  $\alpha_6$ & 0.042 & 0.231 & 0.18 & 0.857 &   & 0.246 & 0.158 & 1.56 & 0.120 &   & 0.026 & 0.407 & 0.06 & 0.949 \\ 
   $\alpha_7$ & 0.283 & 0.230 & 1.23 & 0.217 &   & 0.160 & 0.159 & 1.01 & 0.314 &   & 0.057 & 0.401 & 0.14 & 0.886 \\ 

### 2013-06-10 sociology_paper.tex
some materials are moved to turk paper

@@
 
-\begin{table}[hbtp]
-\caption{Amazon mechanical turk experiments and their properties. Duration in hours per 100 tasks show the popularity of some tasks compared to others.}
-\centering
-\begin{tabular}{rlrrrrrrr}
-  \hline
-& Experiment& \multicolumn{2}{c}{ Total Task}& Average & \multicolumn{2}{c}{Duration (hour)} & Payment & Pay rate\\
-\cline{3-4} \cline{6-7}
-Serial & description & submitted & rejected & time(min) & Actual & 100 task& \$/task & \$/hour\\ 
-  \hline
-1 & Boxplot & 406 & 106 & 10.68 & 146.48 & 36.08 & 0.50 & 2.81 \\ 
-  2 & Scatterplot & 359 &   9 & 10.80 & 42.68 & 11.89 & 1.00 & 5.58 \\ 
-  3 & Contaminated plot & 219 &  19 & 13.53 & 126.17 & 57.61 & 1.00 & 2.22 \\ 
-  4 & Polar vs Cartesian & 110 &  10 & 20.65 & 11.65 & 10.59 & 1.00 & 2.91 \\ 
-  5 & Hist vs density & 234 &  37 & 17.85 & 41.57 & 17.76 & 1.00 & 3.36 \\ 
-  6 & Violin vs boxplot & 417 &  17 & 17.95 & 105.87 & 25.39 & 1.00 & 3.34 \\ 
-  7 & Group separation & 106 &   6 & 16.13 & 5.15 & 4.86 & 1.00 & 3.72 \\ 
-  8 & Sine Illusion & 101 &   1 & 16.52 & 78.38 & 77.60 & 1.00 & 3.63 \\ 
-  9 & Gene expression & 103 &   3 & 12.47 & 11.27 & 10.94 & 0.50 & 2.41 \\ 
-  10 & Test normality & 406 &   6 & 22.70 & 74.35 & 18.31 & 1.00 & 2.64 \\ 
-   \hline
-\end{tabular}
-\label{tbl:mturk}
-\end{table}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=3in]{rejected_task.pdf}
-      \includegraphics[width=3in]{task_duration.pdf} 
-   \caption{Percentage of rejected tasks and duration of each experiment in hour per 100 tasks for each of the 10 experiments.
+%\begin{table}[hbtp]
+%\caption{Amazon mechanical turk experiments and their properties. Duration in hours per 100 tasks show the popularity of some tasks compared to others.}
+%\centering
+%\begin{tabular}{rlrrrrrrr}
+%  \hline
+%& Experiment& \multicolumn{2}{c}{ Total Task}& Average & \multicolumn{2}{c}{Duration (hour)} & Payment & Pay rate\\
+%\cline{3-4} \cline{6-7}
+%Serial & description & submitted & rejected & time(min) & Actual & 100 task& \$/task & \$/hour\\ 
+%  \hline
+%1 & Boxplot & 406 & 106 & 10.68 & 146.48 & 36.08 & 0.50 & 2.81 \\ 
+%  2 & Scatterplot & 359 &   9 & 10.80 & 42.68 & 11.89 & 1.00 & 5.58 \\ 
+%  3 & Contaminated plot & 219 &  19 & 13.53 & 126.17 & 57.61 & 1.00 & 2.22 \\ 
+%  4 & Polar vs Cartesian & 110 &  10 & 20.65 & 11.65 & 10.59 & 1.00 & 2.91 \\ 
+%  5 & Hist vs density & 234 &  37 & 17.85 & 41.57 & 17.76 & 1.00 & 3.36 \\ 
+%  6 & Violin vs boxplot & 417 &  17 & 17.95 & 105.87 & 25.39 & 1.00 & 3.34 \\ 
+%  7 & Group separation & 106 &   6 & 16.13 & 5.15 & 4.86 & 1.00 & 3.72 \\ 
+%  8 & Sine Illusion & 101 &   1 & 16.52 & 78.38 & 77.60 & 1.00 & 3.63 \\ 
+%  9 & Gene expression & 103 &   3 & 12.47 & 11.27 & 10.94 & 0.50 & 2.41 \\ 
+%  10 & Test normality & 406 &   6 & 22.70 & 74.35 & 18.31 & 1.00 & 2.64 \\ 
+%   \hline
+%\end{tabular}
+%\label{tbl:mturk}
+%\end{table}
+%
+%
+%\begin{figure}[htbp] 
+%   \centering
+%   \includegraphics[width=3in]{rejected_task.pdf}
+%      \includegraphics[width=3in]{task_duration.pdf} 
+%   \caption{Percentage of rejected tasks and duration of each experiment in hour per 100 tasks for each of the 10 experiments.
 Most of the tasks got rejected for box plot experiment.
 Even though the sine illusion experiment took longest to finish the rejection rate is lowest for this experiment.}
-   \label{fig:task_duration}
-\end{figure}
+%   \label{fig:task_duration}
+%\end{figure}
 

### 2013-06-10 sociology_paper.tex
see some comments

@@
    \includegraphics[width=0.99\textwidth]{electoral-5-13.pdf} 
-   \caption{ \label{fig:elect-1} \dc{Which one of the plots is the most different from the others?
-Maybe this question will have people picking 4, because it is the only one where the red side is higher.}}
+   \caption{ \label{fig:elect-1} Which one of the plots is the most different from the others?}
 \end{figure}
@@
 (Although this example is more simplistic than most of the tests conducted to date, it will serve the purpose of illustrating the lineup protocol.) The data is looking at the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
-Each panel in Figure~\ref{fig:elect-1} shows an `electoral building' \citep{nytimes} where each state in the union is represented by a rectangle.
+Each panel in Figure~\ref{fig:elect-1} shows an `electoral building' where each state in the union is represented by a rectangle.
 This difference is plotted horizontally, and the height of each box corresponds to the state's electoral votes.
@@
 
-There have been ten experiments conducted using Amazon's Mechanical Turk~\citep{turk} that are being used to evaluate the the effect of the observer.
-Table \ref{tbl:visual_stat} summarizes these.
-All of these experiments are used to evaluate the effects of demographic factors.
-Experiments 5,6 and 7 are used to study learning trend of the observer.
-The design of experiment 9 enabled effect of position in the lineup, and the sample of nulls, to be evaluated.
-Section \ref{sec:factor_performance} describes the factors that may affect the performance of the observer.
-Section \ref{sec:exp_design} describes the methods used to assess the effects, and Section \ref{sec:result_socio} describes the results.
-
-%\subsection{Experiments pooled}
-
 \begin{table*}[hbtp] 
@@
 
-\section{Factors Affecting the performance of the observer} \label{sec:factor_performance}
+There have been ten experiments conducted using Amazon's Mechanical Turk~\citep{turk} that are being used to evaluate the effects of observers' demographic factors on the inference.
+Table \ref{tbl:visual_stat} summarizes these experiments.
+Each of these experiments collected demographic details of the subjects.
+Experiments 5, 6 and 7 are used to study the learning trend of the observer.
+The design of experiment 9 incorporated components that allows position of the actual data plot in the lineup, and the sample of nulls, to be evaluated.
+Section \ref{sec:factor_performance} discusses human factors that may affect the performance of the observer.
+Section \ref{sec:exp_design} describes the methods used to assess the effects, and Section \ref{sec:result_socio} describes the results.
+
+%\subsection{Experiments pooled}
+
+\section{Factors Affecting  Observer Performance} \label{sec:factor_performance}
 
 A brief description of the factors that may affect the power of visual inference.
+
+\dc{You need to have more than one sentence to open this section, it needs a paragraph overviewing the factors.}
 
@@
 
-\section{Simulation Experiment and Methods}\label{sec:exp_design}
+\section{Experimental Designs and Methods}\label{sec:exp_design}
 The performance of the observer while evaluating a lineup would depend on the factors described in Section \ref{sec:factor_performance}.
 This section presents the simulation experimental setup, data collection methods and data analysis plans with models to asses the influence of those factors on visual statistical inference.
+
+\dc{Learning trend and location effect need to be separate sections, cannot be simply bolded.}
 

### 2013-06-10 sociology_paper.tex
overall edits, giving reference to figures

@@
 The experiments suggest that individual skills vary substantially,  but demographics do not have a huge effect.
-There is evidence that a learning effect exists that observers get better and faster with repeated evaluations which suggests a potential for using the lineup protocol for training people to read statistical graphics.
-The placement in the lineup does not affect the inference.
+There is evidence that a learning effect exists that observers get faster with repeated evaluations which suggests a potential for using the lineup protocol for training people to read statistical graphics.
+The placement of data plot in the lineup does not affect the inference.
 
@@
 If the null hypothesis is true the actual data plot should look just like any of the other plots, and not be identified by an observer.
- \hh{Figure~\ref{fig:tower} shows a plot of the electoral building with added context information and labels.}
+Figure~\ref{fig:tower} shows a plot of the electoral building with added context information and labels.
  %A party needs 270 electoral votes or more to win the presidential election. 
@@
 
-A brief description of the factors that may affect the power of visual inference.
-
-\dc{You need to have more than one sentence to open this section, it needs a paragraph overviewing the factors.}
+Based on human evaluation on a lineup a decision is made in visual statistical inference.
+While the visual test statistic is defined to have a direct influence on the observer to pick a plot based on the signal in the data, there are other human factors that may affect the observer performance.
+It is important to study the effect of those factors and examine the extent of their influence.
+This section presents a brief description of the factors that may affect the power of visual inference.
+
+%\dc{You need to have more than one sentence to open this section, it needs a paragraph overviewing the factors.}
 
@@
 
-\dc{Learning trend and location effect need to be separate sections, cannot be simply bolded.}
+%\dc{Learning trend and location effect need to be separate sections, cannot be simply bolded.}
 
@@
 
-\paragraph{Learning Trend:} Learning trend of an observer can be observed in terms of performance over successive feedback when multiple lineups are shown for evaluation.
+\subsubsection{Learning Trend} Learning trend of an observer can be observed in terms of performance over successive feedback when multiple lineups are shown for evaluation.
 Experiments 5, 6 and 7 were used for this.
@@
 
-\paragraph{Design for Location effect:} The experiment was set up based on the gene expression data with two groups.
+\subsubsection{Design for Location effect} The experiment was set up based on the gene expression data with two groups.
 The groups are defined based on the Genotype and Empty Vector (EV).
@@
 
-Besides United States and India, countries such as Canada, Romania, United Kingdom and Macedonia have more than 10 participants each. The rest of the 70 countries have less than 10 participants.
+Besides United States and India, countries such as Canada, Romania, United Kingdom and Macedonia have more than 10 participants each.
+The rest of the 70 countries have less than 10 participants.
+Figure \ref{fig:turker_location_experiment} displays the location of turk workers by each experiment.
+Notice that For all the ten experiments the locations are similar with some small variations.
+That may be due to the time of the experiment when it is started.
+For some experiments India got more participants than United States and for some experiments the numbers just got reversed (\ref{fig:participation_time}).
+It is also observed in some experiments number of  participants are similar no matter whether the experiment is run on day time or nigh time.
 
@@
 
-A total of XX lineups were evaluated in all these 10 experiments.
+A total of 1911 lineups were evaluated in all these 10 experiments.
 Each person was supposed to evaluate 10 lineups except experiment 9 where three lineups were evaluated by each person.
@@
 \centering
-\caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for \hh{probability of correct lineup evaluation.} %log time taken to evaluate a lineup. 
-None of the fixed factor effects of attempt ($\alpha_2$ through $\alpha_{10}$) seem to be significantly different from %each other 
-\hh{the first attempt $\alpha_1$}
+\caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for probability of correct lineup evaluation.
+None of the fixed factor effects of attempt ($\alpha_2$ through $\alpha_{10}$) seem to be significantly different from the first attempt $\alpha_1$
 at \%1 level in all three experiments 5, 6 and 7. For experiment 7 subject specific variation is very small on the other hand lineup variance is much higher compared to the other two experiments.}
@@
 
-With controlled simulation experiments, it is observed that the performance of human observers does not increase through successive attempts while evaluating multiple lineups.
+Observer demographics don't seem to have significant influence on performance in time taken to evaluate a lineup.
+Some variation is observed in terms of median proportion correct, but the overall average proportion of correct responses over different demographic factors levels seem to be similar.
+
+With simulation experiments, it is observed that the performance of human observers does not increase through successive attempts while evaluating multiple lineups.
 But their skill in evaluating the lineup in shorter time gets improved over successive evaluations.
@@
 
-Observer demographics don't seem to have significant influence on performance in time taken to evaluate a lineup.
-Some variation is observed in terms of median proportion correct, but the mean proportion correct over different demographic factors levels seem to be similar.
-
 The simulation experiment reveals that the effect of location of actual data plot in the lineup is not significant.
@@
 
-More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of knowledge about statistical graphics.
-The subjects participating this study may not necessarily knowledgable about statistical graphics.
+The subjects participating this study may not necessarily know about statistical graphics.
 The numerous pilot studies done with the more trained participants suggest that power of visual statical inference may be higher with knowledgeable observer.
 Thus the variations observed with individual performances could be attributed to individual abilities in terms of their training on statistical graphics.
+More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of knowledge about statistical graphics.
 

### 2013-06-10 sociology_paper.tex
experiment setup updated

@@
 \subsection{Individual Performance of the Observer} Each person is different from others in some way.
-For example, in a controlled experimental study \citet{zhao:2012}  notice that some people spend a lot of time to decide no matter whether the lineup is difficult or easy while some simply glance at lineups to make a decision.
+For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.
 This influences the response of the observer.
@@
 The performance of the observer while evaluating a lineup would depend on the factors described in Section \ref{sec:factor_performance}.
+Some of these factors such as signal in the data and individual performances are studied in \cite{majumder:2013}.
+In this paper we intend to study effect of the factors such as human demographics, learning trend, actual plot locations and selection of null plots.
 This section presents the simulation experimental setup, data collection methods and data analysis plans with models to asses the influence of those factors on visual statistical inference.

### 2013-06-12 R/sociology_paper.R
fitting model with demographics

@@ -215,13 +215,16 @@ qplot(variable_level, value, geom="boxplot",data=mdat[complete.cases(mdat),]) +
 
 ggsave("../images/demographic_effect.pdf", width=6, height=6)
 
+# --------------------------------------------------------------
+# Model fitting with demographic factors
 # testing significance of demographic factor main effects
-
-ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+ (1|id)+(1|pic_name), data=demographics)
+# --------------------------------------------------------------
+ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name), 
+           data=subset(demographics, experiment=="experiment_6"))
 summary(ft)
 
-fp <- lmer(response~age_level+country+degree+gender_level+ (1|id)+(1|pic_name), 
-           family="binomial", data=demographics)
+fp <- lmer(response~age_level+country+degree+gender_level+(1|pic_name), 
+           family="binomial", subset(demographics, experiment=="experiment_5"))
 summary(fp)
 
 
@@ -323,8 +326,9 @@ gregexpr("\\W+", reason.correct)
 
 
 # =====================================================
-# Model fitting
+# Model fitting with trend
 # Examining learning trend while giving feedback
+# -----------------------------------------------------
 
 # function two obtain sequential attempt information from start_time
 get_trend <- function(dat){

### 2013-06-12 sociology_paper.tex
gene expression data reference updated

@@
 
-\subsubsection{Design for Location effect} The experiment was set up based on the gene expression data with two groups.
+\subsubsection{Design for Location effect} The experiment was set up based on the gene expression data \citep{atwood:2013} with two groups.
 The groups are defined based on the Genotype and Empty Vector (EV).
@@
 Two hypothesis were of primary interest.
-The Genotype effect and the Genotype and iron sufficiency interaction effect.
+The Genotype main effect and the Genotype and iron sufficiency interaction effect.
 
@@
 Each observer was shown three lineups.
-First lineups shown is a randomly selected Interaction lineup, one from randomly selected Genotype lineup.
+First lineup shown was a randomly selected Interaction lineup, the second one was from randomly selected Genotype lineup.
 Finally a test lineup was shown to screen out unusual evaluation.

### 2013-06-12 R/sociology_paper.R

@@ -221,6 +221,9 @@ ggsave("../images/demographic_effect.pdf", width=6, height=6)
 # --------------------------------------------------------------
 ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name), 
            data=subset(demographics, experiment=="experiment_6"))
+
+ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+experiment+(1|pic_name), 
+           data=subset(demographics,))
 summary(ft)
 
 fp <- lmer(response~age_level+country+degree+gender_level+(1|pic_name), 

### 2013-06-12 sociology_paper.tex
reference updated

@@
 
-\subsection{Signal in the data}
+\subsection{Signal in the Data}
 
@@
 
-\subsubsection{Design for Location effect} The experiment was set up based on the gene expression data \citep{atwood:2013} with two groups.
+\subsubsection{Design for Location Effect} The experiment was set up based on the gene expression data \citep{atwood:2013} with two groups.
 The groups are defined based on the Genotype and Empty Vector (EV).
@@
 
-\subsection{Data collection methods}  Human observers were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \cite{turk} or MTurk Web site.
+\subsection{Data Collection Methods}  Human observers were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \cite{turk} or MTurk Web site.
 It is an online work place where people from around the world can perform some tasks and get paid.
@@
 
-\section{Some figures and tables (May go to Appendix)}
-
-\subsection{How people pick the data plot} In the experiment setup we have a question for the turk observers to choose the reason for their selection of a particular data plot.
+\section{Some Figures and Tables (May go to Appendix)}
+
+\subsection{How People Pick the Data Plot} In the experiment setup we have a question for the turk observers to choose the reason for their selection of a particular data plot.
 It is shown by \cite{majumder:2013} that these choice reasons reveal the way people picks the data plot in the lineup.

### 2013-06-13 sociology_paper.tex
demographic model result added

@@
 This indicates that there may be no learning effect on the proportion of correct evaluations.
+
+\begin{table}[htbp]
+\centering
+\caption{Parameter estimates of Model \eqref{eqn:model_demo} fitted for avarage log time taken and probability of correct lineup evaluation.}
+\scalebox{0.9}{
+\begin{tabular}{rrrrrrrrrrr}
+  \hline
+& \multicolumn{4}{c} {Log(Time)} & &\multicolumn{4}{c} {Proportion correct} \\
+
+\cline{2-5} \cline{7-10} 
+ Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
+  \hline
+Fixed &  &  &  & && &  &  \\ 
+$\mu$ & 3.465 & 0.014 & 241.52 & $<$0.001 &   & -0.834 & 0.073 & -11.43 & $<$0.001 \\ 
+\multicolumn{2}{l}{Age}   &  &  &  &  & & && \\   
+  26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.209 \\ 
+  31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.038 \\ 
+  36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
+  41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.050 \\ 
+  46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.143 \\ 
+  above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.039 \\ 
+\multicolumn{2}{l}{Country}   &  &  &  &  & & && \\   
+  Rest world & -0.086 & 0.016 & -5.52 & $<$0.001 &   & 0.267 & 0.060 & 4.49 & $<$0.001 \\ 
+  USA & -0.230 & 0.012 & -19.39 & $<$0.001 &   & 0.186 & 0.046 & 4.05 & $<$0.001 \\ 
+\multicolumn{2}{l}{Education}   &  &  &  &  & & && \\       
+  UG course & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
+  UG degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.330 \\ 
+  Gr. course & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.157 \\ 
+  Gr. degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
+\multicolumn{2}{l}{Gender}   &  &  &  &  & & && \\     
+  Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.133 \\ 
+\multicolumn{2}{l}{Random}   &  &  &  &  & & && \\ 
+  Lineup & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
+  Error & 0.479 & 0.692 &  &  & & && &\\ 
+   \hline
+\end{tabular}
+}
+\label{tbl:model_result_demographics}
+\end{table}
+
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{rrrlllrrll}
+%  \hline
+% & Est & SE & Zval & pvalue & g1 & Est & SE & Zval & pvalue \\ 
+%  \hline
+%(Intercept) & 3.465 & 0.014 & 241.52 & $<$0.001 &   & -0.834 & 0.073 & -11.43 & $<$0.001 \\ 
+%  age_level26-30 & 0.058 & 0.013 & 4.5 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
+%  age_level31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0375 \\ 
+%  age_level36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
+%  age_level41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.0501 \\ 
+%  age_level46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.1425 \\ 
+%  age_levelabove 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.0394 \\ 
+%  countryRest of the world & -0.086 & 0.016 & -5.52 & $<$0.001 &   & 0.267 & 0.060 & 4.49 & $<$0.001 \\ 
+%  countryUnited States & -0.230 & 0.012 & -19.39 & $<$0.001 &   & 0.186 & 0.046 & 4.05 & 1e-04 \\ 
+%  degree.L & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
+%  degree.Q & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.3306 \\ 
+%  degree.C & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.1566 \\ 
+%  degree^4 & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
+%  gender_levelMale & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.5 & 0.1329 \\ 
+%  pic_name (Intercept) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
+%   \hline
+%\end{tabular}
+%\end{table}
 

### 2013-06-13 R/sociology_paper.R

@@ -215,6 +215,27 @@ qplot(variable_level, value, geom="boxplot",data=mdat[complete.cases(mdat),]) +
 
 ggsave("../images/demographic_effect.pdf", width=6, height=6)
 
+#----------------------------------------------------------------
+# Function to obtain estimates of fitted model for latex xtable
+# Getting results of mixed effect model with log(time_taken)
+#----------------------------------------------------------------
+get_estimates <- function(fit){
+  Vcov <- vcov(fit, useScale = FALSE)
+  betas <- fixef(fit)
+  se <- sqrt(diag(Vcov))
+  zval <- betas / se
+  pval <- round(2 * pnorm(abs(zval), lower.tail = FALSE),4)
+  pval[pval==0] <- "$<$0.001"
+  fe <- data.frame(Est=round(betas,3),SE= round(se,3), Zval=round(zval,2), pvalue=pval)
+  rem <- summary(fit)@REmat
+  re <- data.frame(Est=rem[,3], SE= rem[,4], Zval="",pvalue="")
+  re$Est <- round(as.numeric(as.character(re$Est)),3)
+  re$SE <- round(as.numeric(as.character(re$SE)),3)
+  rownames(re) <- paste(rem[,1],rem[,2])
+  return(rbind(fe,re[order(rownames(re)),]))
+}
+
+
 # --------------------------------------------------------------
 # Model fitting with demographic factors
 # testing significance of demographic factor main effects
@@ -222,13 +243,21 @@ ggsave("../images/demographic_effect.pdf", width=6, height=6)
 ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name), 
            data=subset(demographics, experiment=="experiment_6"))
 
-ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+experiment+(1|pic_name), 
-           data=subset(demographics,))
+ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name), 
+           data=demographics)
 summary(ft)
 
 fp <- lmer(response~age_level+country+degree+gender_level+(1|pic_name), 
            family="binomial", subset(demographics, experiment=="experiment_5"))
+fp <- lmer(response~age_level+country+degree+gender_level+(1|pic_name), 
[…]

### 2013-06-13 sociology_paper.tex
demographic factor results updated

--- 2013-06-12 sociology_paper.tex
@@
   \hline
- Factor & levels & subsjects & avg\_time & response \\ 
+ Factor & levels & subjects & avg\_time & response \\ 
   \hline
@@
 This indicates that there may be no learning effect on the proportion of correct evaluations.
+
+\begin{table}[htbp]
+\centering
+\caption{Parameter estimates of Model \eqref{eqn:model_demo} fitted for avarage log time taken and probability of correct lineup evaluation.}
+\scalebox{0.9}{
+\begin{tabular}{rlrrrrrrrrrr}
+  \hline
+Demographic&& \multicolumn{4}{c} {Model Log(Time)} & &\multicolumn{4}{c} {Model Proportion Correct} \\
+
+\cline{3-6} \cline{8-11} 
+Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
+  \hline
+\multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
+&$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
+\hline
+Age&18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
+ & 26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
+  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0376 \\ 
+  &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
+  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.0501 \\ 
+  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.1425 \\ 
+  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.0394 \\ 
+  \hline
+Country& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
+  &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
+  &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
+  \hline
+Education &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
+ &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
+  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.3307 \\ 
+  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.1566 \\ 
+  &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
+  \hline
+Gender & Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
+  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.5 & 0.1331 \\ 
+  \hline
+\multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
+& lineup & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
+ & Error & 0.479 & 0.692 &  &  & & && &\\ 
+   \hline
+\end{tabular}
+}
+\label{tbl:model_result_demographics}
+\end{table}
+
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{rrrlllrrll}
+%  \hline
+% & Est & SE & Zval & pvalue & g1 & Est & SE & Zval & pvalue \\ 
+%  \hline
+%(Intercept) & 3.465 & 0.014 & 241.52 & $<$0.001 &   & -0.834 & 0.073 & -11.43 & $<$0.001 \\ 
+%  age_level26-30 & 0.058 & 0.013 & 4.5 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
+%  age_level31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0375 \\ 
+%  age_level36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
+%  age_level41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.0501 \\ 
+%  age_level46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.1425 \\ 
+%  age_levelabove 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.0394 \\ 
+%  countryRest of the world & -0.086 & 0.016 & -5.52 & $<$0.001 &   & 0.267 & 0.060 & 4.49 & $<$0.001 \\ 
+%  countryUnited States & -0.230 & 0.012 & -19.39 & $<$0.001 &   & 0.186 & 0.046 & 4.05 & 1e-04 \\ 
+%  degree.L & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
+%  degree.Q & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.3306 \\ 
+%  degree.C & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.1566 \\ 
+%  degree^4 & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
+%  gender_levelMale & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.5 & 0.1329 \\ 
+%  pic_name (Intercept) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
+%   \hline
+%\end{tabular}
+%\end{table}
 

### 2013-06-13 R/sociology_paper.R

@@ -131,8 +131,13 @@ demographics$country <- as.character(demographics$country_name)
 demographics$country[(demographics$country_code != "IN") & (demographics$country_code != "US")] <- "Rest of the world"
 demographics$country[demographics$country=="Namibia"] <- "Rest of the world"
 demographics$country[demographics$country==NA] <- "Rest of the world"
+country <- factor(demographics$country)
+demographics$country <- factor(country,levels=levels(country)[c(3,1,2)], ordered=T)
+
 levels(demographics$age_level)[7:9] <- 7
 levels(demographics$age_level)[1:7] <- c(levels(demographics$age_level)[1:6], "above 50")
+
+
 
 
 # getting unique participants for plotting purpose only
@@ -255,9 +260,56 @@ fp <- lmer(response~age_level+country+degree+gender_level+(1|pic_name),
 summary(fp)
 
 est.factor <- cbind(get_estimates(ft)[-16,],g1=" ",get_estimates(fp))
+rownames(est.factor) <- c("$mu$",substr(rownames(est.factor)[2:7],10,nchar(rownames(est.factor)[2:7])),
+                          levels(demographics$country)[-1], levels(demographics$degree)[-1],
+                          "Male", "lineup")
 dgt <- c(rep(0,15), rep(3,30), rep(2,15),rep(0,15))
 dgts <- matrix(rep(dgt,2), ncol=10)
 print(xtable(est.factor, digits=dgts),  sanitize.text.function = function(x){x})
+
+fta <- lmer(log(time_taken)~country+degree+gender_level+(1|pic_name), 
+           data=demographics)
+ftc <- lmer(log(time_taken)~age_level+degree+gender_level+(1|pic_name), 
+           data=demographics)
+ftd <- lmer(log(time_taken)~age_level+country+gender_level+(1|pic_name), 
+           data=demographics)
+ftg <- lmer(log(time_taken)~age_level+country+degree+(1|pic_name), 
+           data=demographics)
+
+tag <- anova(ft,fta)
+tcn <- anova(ft,ftc)
+tde <- anova(ft,ftd)
+tge <- anova(ft,ftg)
+
+anova.time <- rbind(data.frame(round(tag[1,1:4],2),round(tag[2,5:7]/1,2)),
+                    data.frame(round(tcn[1,1:4],2),round(tcn[2,5:7]/1,2)),
+                    data.frame(round(tde[1,1:4],2),round(tde[2,5:7]/1,2)),
[…]

### 2013-06-13 sociology_paper.tex
anova analysis added

@@
 The dots inside the box represents mean.
-None of the demographic factors appear to be statistically significant except India effect for proportion correct.
-Effect of graduate degree in proportion correct is also significant but is actually the India effect since, Most of the graduate degree holders are coming from India.}
+Demographic factors appear to be statistically significant for explaining time taken.
+Effect of graduate degree and India in proportion correct is much different but is actually the India effect since, Most of the graduate degree holders are coming from India.}
    \label{fig:demographic_effect}
@@
 \label{tbl:model_result_demographics}
+\end{table}
+
+\begin{table}[htbp]
+\centering
+\caption{Anova of full model with all the demographic factors vs reduced model with removing respective factor.}
+\begin{tabular}{lrrrrrrrr}
+  \hline
+Model && Df & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
+  \hline
+Log Time &Full& 16 &  51904.00 &52034.00 & -25936.00 & &  &  \\ 
+&Age & 10 & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
+ & Country & 14 & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
+  &Education & 12 & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
+  &Gender & 15 & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
+\hline
+Proportion Correct &Full& 15 & 23822.00 &23943.00 &-11896.00  & &  &  \\ 
+  &Age & 9 & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
+  &Country & 13 & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
+  &Education & 11 & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
+  &Gender & 14 & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
+   \hline
+\end{tabular}
 \end{table}

### 2013-06-13 sociology_paper.tex
anova analysis table updated

@@
 Log Time &Full& 16 &  51904.00 &52034.00 & -25936.00 & &  &  \\ 
+& \multicolumn{2}{l}{Reduced} &  &  &  &  & \\
 &Age & 10 & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
@@
 Proportion Correct &Full& 15 & 23822.00 &23943.00 &-11896.00  & &  &  \\ 
+& \multicolumn{2}{l}{Reduced} &  &  &  &  & \\
   &Age & 9 & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 

### 2013-06-13 sociology_paper.tex
model for demographic factor

@@
 
+\subsection{Model to Estimate Demographic Factor Effect} Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$.
+Let $\pi_{ij}=  E(Y_{ij})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$.
+We model this in a generalized mixed effects model of the form
+\begin{equation} \label{eqn:demographic_response}
+g( \pi_{ij} )= \mu + X_{i}\beta + \ell_j,  
+\end{equation}
+
+where $\mu$ is an overall average probability for picking out the data plot from a lineup.
+$X_{i}=(X_{1i},X_{2i}, .., X_{pi})'$ is the vector of demographic factors of the observer $i$ and $\beta$ is the corresponding fixed effect coefficient vector.
+$p$ indicates the number of demographic factor variables.
+For $\ell_j$ we assume a normal distribution, $N(0, \sigma_\ell^2)$.
+$\ell_j$ is a random intercept predicting lineup difficulty level.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=g_1(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+
+The effect of demographic factors can be observed as time taken for each evaluation by an observer.
+Suppose $Z_{ij}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$.
+Let $\mu_{ij}=  E(Z_{ij})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$.
+We model this in a mixed effects model of the form
+%\hh{
+\begin{equation} \label{eqn:demographic_time}
+Z_{ij} = \mu + X_{i}\beta + \ell_j+ \epsilon_{ij},  
+\end{equation}
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
+$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
+
 \subsection{Model to Estimate Learning Trend}
@@
 \centering
-\caption{Parameter estimates of Model \eqref{eqn:model_demo} fitted for avarage log time taken and probability of correct lineup evaluation.}
+\caption{Parameter estimates of Models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for avarage log time taken and probability of correct lineup evaluations respectively.
+For time taken all the demographic factors appear to be significant.
+For probability of correct response age group 36-40, rest of the world and graduate degree appears to be significantly different.
+Gender does not show any difference in performance.}
 \scalebox{0.9}{
@@
   \hline
-Demographic&& \multicolumn{4}{c} {Model Log(Time)} & &\multicolumn{4}{c} {Model Proportion Correct} \\
+Demographic&& \multicolumn{4}{c} {Model \eqref{eqn:demographic_time} Log Time} & &\multicolumn{4}{c} {Model \eqref{eqn:demographic_response} Proportion Correct} \\
 

### 2013-06-14 sociology_paper.tex
subjectwise learning trend added

@@
 Results from multiple visual inference studies using Amazon's Mechanical Turk are examined to provide an assessment of these.
-The experiments suggest that individual skills vary substantially,  but demographics do not have a huge effect.
+The experiments suggest that individual skills vary substantially,  but demographics do not have a huge effect on performance.
 There is evidence that a learning effect exists that observers get faster with repeated evaluations which suggests a potential for using the lineup protocol for training people to read statistical graphics.
@@
 
-\subsection{Data Collection Methods}  Human observers were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \cite{turk} or MTurk Web site.
+\subsection{Data Collection Methods}  Human observers were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \citep{turk} or MTurk Web site.
 It is an online work place where people from around the world can perform some tasks and get paid.
-Usually tasks are very simple and no specialized training is required.
+Usually tasks are very simple and no specialized training is required to do them.
 Being a human is the main requirement.
 Tasks are designed for anyone to do but some tasks may require that workers satisfy some skill level depending on the recruiters need.
-The tasks are designed such that it does not take much time to complete.
+Each task is usually planned to complete in a short time.
 Humans are still better than computers in performing these types of tasks.
@@
 
-It is very fast, cheap and reliable to recruit people from MTurk.
+%It is very fast, cheap and reliable to recruit people from MTurk.
 Thats why it is getting very popular among the researchers who perform human subject experiment.
@@
 
-Because of its convenience it is getting popular for scientific research study as well.
+%Because of its convenience it is getting popular for scientific research study as well.
 In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time and cost while provided more convenience.
@@
 
-We designed and developed a web site which enables to display the lineups to the observers as per experimental design.
+We designed and developed a web application which enables to display the lineups to the observers as per experimental design.
 The MTurk workers were redirected to the web site and the data were collected, stored automatically into a local database server.
@@
 
-\subsection{Model to Estimate Demographic Factor Effect} Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$.
+\subsection{Model to Estimate Demographic Factor Effect} The feedback provided by each observer on a lineup is a binary response variable.
+Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$.
+Yij = 1 if the response is correct otherwise Yij = 0.
 Let $\pi_{ij}=  E(Y_{ij})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$.
@@
 \end{equation}
-
 where $\mu$ is an overall average probability for picking out the data plot from a lineup.
@@
 
-The feedback provided by each observer on a lineup is a binary response variable.
+Multiple lineups were sequentially shown to each observer for evaluation.
 Suppose $Y_{ijk}$ denotes the response from observer $i$ on a  lineup $j$ in his/her $k$th attempt.
 $Y_{ijk}=1$ if the response is correct otherwise  $Y_{ijk}=0$.
-
-%As per the design of the experiment for learning trend we have two different response variables.
-One is observer feedback on a lineup which is a binary variable.
-The other variable is the time taken for each evaluation which is a continuous variable.
-Suppose random variable $Y_{\ell, i}$ denotes the response from subject $ i=1,2, ..., I$ on lineup $\ell=1,2, ..., L$.
-We model the mean $\mu_{\ell, i} = E(Y_{\ell, i})$ with a generalized liner mixed effect model
-%
-%\begin{equation} \label{eqn:trend}
-%g( \mu_{\ell, i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
-%\end{equation}
-
-%\hh{
 Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in his/her $k$th attempt.
@@
 
+To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
+To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
+
 \subsection{Model to Estimate Location Effect}
 
-Since the actual data plot is same for each of the null sets of plot, the response data constitutes a multivariate response.
-Thus, to examine if the difference in proportion correct among the locations is statistically significant we fit a one way multi variate analysis of variance (MANOVA) model to the data.
+As per the design of experiment for location effect, the actual data plot is same for each of the null sets of plot.
+Thus the response data of this experiment constitute a multivariate response.
+To examine if the difference in proportion correct among the locations is statistically significant we fit a one way multi variate analysis of variance (MANOVA) model to the data.
 
@@
 \subsection{Overview of the Data} Figure \ref{fig:turker_location} displays the location around the world from where a total of 2321 participants provided feedback data on the lineups in all the ten experimental studies.
-Even though most of the participants were from the United States and India, we have participants from 76 different countries.
+Even though most of the participants were from the United States and India, we have participants from 76 other different countries.
 This provides a diverse pool of participants in terms of geographical locations.
 The diversity in not only in geographical locations of the participants but also in their gender, age groups and education levels as we see in Table \ref{tbl:demographics}.
-It is interesting that the number of male and female participants are almost same even though there are lot of people from developing countries.
+It is interesting that there were large number of female participants even though there were lot of people from developing countries.
 
@@
 Most of the participants from India are below age 40 while in United States the distribution of participants are similar beyond age 40.
+There were not many participants beyond age 50 and hence these age groups are merged to form one age group called above 50.
+In exploratory analysis and model fitting this new age group is used instead of actual age groups beyond age 50.
 
@@
 Each person was supposed to evaluate 10 lineups except experiment 9 where three lineups were evaluated by each person.
-To clean the data we used the test plot as the basis and did not include the results of that test plot.
+A test plot was shown to each observer and the feedback received from that plot is used to examine the data quality.
+But the data received for the test lineup is not included in the analysis.
 In some cases the demographic information were not provided by the participants.
@@
 The time taken by male and female participants are almost similar.
-None of the factors are significant in describing time taken to evaluate a lineup.
+
+%All of the factors are significant in describing time taken to evaluate a lineup.
 
@@
 
-We observed some variations in median proportion of correct responses for different levels of the demographic factors as in Figure \ref{fig:demographic_effect}.
-But the variability of proportion correct responses are huge as per design of the experiments since we have some easy and some difficult lineups.
-There were lineups which were not expected to be evaluated correctly and some were very easy and expected to be evaluated correctly 100\% of the time.
-This is why we see the boxes ranges from 0 to 1 in most of the factor levels.
-
-Unlike median, less variations are observed in the averages for different factor levels.
-As expected, there is no significant difference between Male and female performances.
-Among all the demographic factor levels only graduate degree, India and age group 31-35 appear to be significantly different from others.
-Most of the Indian participants reported to have a graduate degree.
-Thus the effect of graduate degree is actually the effect of India.
-
-\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiment 5, 6, and 7 separately.
-For this the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
-To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
-
-Table \ref{tbl:model_result_response} presents the parameter estimates and $p$-values of fixed effect estimates of Model \eqref{eqn:trend_response}.
-The larger $p$-values suggest that none of the levels of attempt ($\alpha_2$ through $\alpha_{10}$) appears to be significant at \%1 significance level.
-Moreover, some of the estimates are positive and some are negative and they show up seemingly in random order suggesting later attempts not necessarily improved.
-This indicates that there may be no learning effect on the proportion of correct evaluations.
+The distribution of time taken to evaluate a lineup is positively skewed.
+But in log scale it is symmetric for all the demographic factor variables as we see in Figure \ref{fig:demographic_effect}.
+Mean and medians are very similar as well.
+This allowed us to fit linear mixed model to the log time taken with normal error structure.
+
+We observed some differences in median proportion of correct responses for different levels of the demographic factors as in Figure \ref{fig:demographic_effect}.
+But the variability of proportion correct responses are huge as per design of the experiments since we have some super easy and some extremely difficult lineups.
+There were lineups which were not expected to be evaluated correctly since some of the null plots in those lineups were showing more signal than the actual plot itself.
+In those scenarios,  one may not reject the null hypothesis.
+We call these difficult lineup.
+Some of the lineups were so easy that data plot is detectable in a first glance.
+These were expected to be evaluated correctly 100\% of the time.
+This is why we see the boxes range from 0 to 1 in most of the factor levels.
+Unlike median, less differences are observed in the average proportion of correct responses for different factor levels.
+
+Model \eqref{eqn:demographic_response} is fitted to the data for different demographic factors such as age, country, education and gender.
+The results are shown in Table \ref{tbl:model_result_demographics}.
+For parameter estimation the first factor levels shown the Table are are used as the base line.
+The age group 36-40 is significantly different than the base line age group 18-25.
+The rest of the age levels appears to be similar in explaining the probability of correct response.
+For country, the rest of the world is different from USA but India appears to be similar to USA.
+We see from Table \ref{tbl:demographics} that most of the participants (about 90\%) are from India and USA.
+The rest of the 10\% data are from 76 different countries.
+This diversity in the rest of the world may make it different from India and USA.
 
@@
 
+The graduate degree holders are significantly different as well.
+The positive parameter estimate indicates that they perform better and the probability of correct response is higher than other education levels.
+As expected, there is no significant difference between Male and female performances.
+
+Notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in Model \eqref{eqn:demographic_response}.
+This indicates that the major and most important factor affecting the probability of correct response is the lineup difficulty.
+This is also evident from the large variability in the proportion correct in Figure~\ref{fig:demographic_effect}.
+Because of this huge impact of lineup specific variance, the actual effect of other factor variables on the probability of correct response is in fact very small even though they appear to be significant.
+
+Model \eqref{eqn:demographic_time} is fitted for the log time taken to evaluate a lineup and the results are also shown in  Table~\ref{tbl:model_result_demographics}.
+The effect of all the demographic factors on log time taken to evaluate a lineup appears to be significant.
+
 \begin{table}[htbp]
@@
 
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{rrrlllrrll}
-%  \hline
-% & Est & SE & Zval & pvalue & g1 & Est & SE & Zval & pvalue \\ 
-%  \hline
-%(Intercept) & 3.465 & 0.014 & 241.52 & $<$0.001 &   & -0.834 & 0.073 & -11.43 & $<$0.001 \\ 
-%  age_level26-30 & 0.058 & 0.013 & 4.5 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
-%  age_level31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0375 \\ 
-%  age_level36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
-%  age_level41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.0501 \\ 
-%  age_level46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.1425 \\ 
-%  age_levelabove 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.0394 \\ 
-%  countryRest of the world & -0.086 & 0.016 & -5.52 & $<$0.001 &   & 0.267 & 0.060 & 4.49 & $<$0.001 \\ 
-%  countryUnited States & -0.230 & 0.012 & -19.39 & $<$0.001 &   & 0.186 & 0.046 & 4.05 & 1e-04 \\ 
-%  degree.L & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
-%  degree.Q & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.3306 \\ 
-%  degree.C & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.1566 \\ 
-%  degree^4 & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
-%  gender_levelMale & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.5 & 0.1329 \\ 
-%  pic_name (Intercept) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
-%   \hline
-%\end{tabular}
-%\end{table}
+\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiment 5, 6, and 7 separately.
+Table \ref{tbl:model_result_response} presents the parameter estimates and $p$-values of fixed effect estimates of Model \eqref{eqn:trend_response}.
+The larger $p$-values suggest that none of the levels of attempt ($\alpha_2$ through $\alpha_{10}$) appears to be significant at \%1 significance level.
+Moreover, some of the estimates are positive and some are negative and they show up seemingly in random order suggesting later attempts not necessarily improved.
+This indicates that there may be no learning effect on the probability of correct evaluations.
 
@@
    \includegraphics[width=6.5in]{learning_trend.pdf} 
+    \includegraphics[width=6.5in]{learning_trend_subject.pdf} 
    \caption{ Mean residuals of proportion correct obtained from the model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
@@
    \includegraphics[width=6.5in]{learning_trend_time.pdf}    
-%   \includegraphics[width=6.5in]{learning_trend_time_subject.pdf}    
+ \includegraphics[width=6.5in]{learning_trend_time_subject.pdf}    
    \caption{ Mean residuals of log(time taken) obtained from model \eqref{eqn:trend_time} fitted without covariate attempt plotted against attempt.
@@
 
-Observer demographics don't seem to have significant influence on performance in time taken to evaluate a lineup.
-Some variation is observed in terms of median proportion correct, but the overall average proportion of correct responses over different demographic factors levels seem to be similar.
-
-With simulation experiments, it is observed that the performance of human observers does not increase through successive attempts while evaluating multiple lineups.
-But their skill in evaluating the lineup in shorter time gets improved over successive evaluations.
-When an observer evaluates multiple lineups, the earlier evaluations take significantly longer than the later evaluations.
+Observer demographics seem to have significant influence on performance in time taken to evaluate a lineup.
+Some variation is observed in terms of probability of correct evaluation.
+Age group 36-40, Countries other than India and United states, People who have a graduate degree appear to have significantly more probability to evaluate a lineup correctly.
+Gender does not have any effect on performance.
+
+Individual learning trend is observed in both time taken and observer performance.
+But the overall performance of human observers does not increase through successive attempts while evaluating multiple lineups.
+The skill in evaluating the lineup in shorter time gets improved over successive evaluations.
+The earlier evaluations take significantly longer than the later evaluations.
 This result is very important as it suggests that the power estimated for visual inference using human subject experiment is robust and may not change if those participants are allowed to give feedback again.
@@
 \section{Electoral Building Lineups and Results}
-Figures~\ref{fig:elect-1} and \ref{fig:elect.2} through~\ref{fig:elect.5} are the five lineups shown to (different) Amazon Turk workers in an experiment.
-They all are created as described in the introduction of the paper.
+%Figures~\ref{fig:elect-1} and \ref{fig:elect.2} through~\ref{fig:elect.5}
+
+Five lineups were shown to (different) Amazon Turk workers in an experiment.
+They all were created as described in the introduction of the paper.
 In order to not bias observers, no context information was given about how these plots were created or what data they were displaying.
@@
 
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.8\linewidth]{electoral-2-17.pdf} 
-   \caption{Lineup \#2 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.2}
-\end{figure}
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.8\linewidth]{electoral-3-18.pdf} 
-   \caption{Lineup \#3 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.3}
-\end{figure}
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.8\linewidth]{electoral-4-17.pdf} 
-   \caption{Lineup \#4 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.4}
-\end{figure}
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=0.8\linewidth]{electoral-1-1.pdf} 
-   \caption{Lineup \#5 - electoral building. Which building looks the most different from the other buildings? }
-   \label{fig:elect.5}
-\end{figure}
+%\begin{figure}[htbp] 
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{electoral-2-17.pdf} 
+%   \caption{Lineup \#2 - electoral building. Which building looks the most different from the other buildings? }
+%   \label{fig:elect.2}
+%\end{figure}
+%\begin{figure}[htbp] 
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{electoral-3-18.pdf} 
+%   \caption{Lineup \#3 - electoral building. Which building looks the most different from the other buildings? }
+%   \label{fig:elect.3}
+%\end{figure}
+%\begin{figure}[htbp] 
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{electoral-4-17.pdf} 
+%   \caption{Lineup \#4 - electoral building. Which building looks the most different from the other buildings? }
+%   \label{fig:elect.4}
+%\end{figure}
+%\begin{figure}[htbp] 
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{electoral-1-1.pdf} 
+%   \caption{Lineup \#5 - electoral building. Which building looks the most different from the other buildings? }
+%   \label{fig:elect.5}
+%\end{figure}
 

### 2013-06-14 R/sociology_paper.R

@@ -306,7 +306,7 @@ anova.prop <- rbind(data.frame(round(pag[1,1:4],2),round(pag[2,5:7]/1,2)),
 
 
 anova.reseult <- rbind(anova.time, anova.prop)
-
+xtable(anova.reseult)
 
 
 
@@ -467,14 +467,29 @@ qplot(attempt,resid, group=id, data= dpt, geom="line", alpha=I(0.1)) +
   facet_wrap(~experiment, scales="free_y") +
   scale_x_continuous(breaks = seq(2,10,by=2)) 
 
+
+
 dmt <- ddply(dpt, .(experiment,attempt), summarise,
              mean_resid = mean(resid))
 
 ggplot() + 
   geom_line(aes(attempt,resid, group=id), data= dpt, alpha=I(0.1)) +
   geom_smooth(aes(attempt,mean_resid), data= dmt, size=1.5, method="lm", se=F) +
-  facet_wrap(~experiment, scales="free_y") + ylab("Residual time taken") 
+  facet_wrap(~experiment, scales="free_y") + ylab("Residual time taken") +
   scale_x_continuous(breaks = seq(2,10,by=2)) 
+
+ggplot() +
+  geom_smooth(aes(attempt,mean_resid), data= dmt, size=1.5, method="lm", se=F) +
+  geom_smooth(aes(attempt,resid, group=id),method="lm", se=F, data=dpt, alpha=I(.01))+
+  facet_wrap(~experiment, scales="free_y")+
+  scale_x_continuous(breaks = seq(2,10,by=2)) 
+
+ggplot() + 
+  geom_point(aes(attempt,mean_resid), data= dmt) +
+  geom_smooth(aes(attempt,mean_resid), data= subset(dmt, attempt>1), method="lm", se=F) +
+  geom_smooth(aes(attempt,resid, group=id),method="lm", se=F, data=subset(dpt, attempt>1), alpha=I(.01))+
+  facet_wrap(~experiment, scales="free_y") + ylab("Residual log(time taken)") +
+  scale_x_continuous(breaks = seq(2,10,by=2))
 
 ggsave("../images/learning_trend_time_subject.pdf", width=10.5, height = 3.5)
 
@@ -539,6 +554,16 @@ qplot(attempt,mean_resid, data= ddt) + geom_point(size=2.5) +
   scale_x_continuous(breaks = seq(2,10,by=2)) 
 
 ggsave("../images/learning_trend.pdf", width=10.5, height = 3.5)
+
+
[…]

### 2013-06-14 R/sociology_paper.R
supplimetary typo fixed

@@ -557,10 +557,10 @@ ggsave("../images/learning_trend.pdf", width=10.5, height = 3.5)
 
 
 ggplot() + 
-  geom_point(aes(attempt,mean_resid), data= dmt) +
-  geom_smooth(aes(attempt,mean_resid), data= subset(dmt, attempt>1), method="lm", se=F) +
-  geom_smooth(aes(attempt,resid, group=id),method="lm", se=F, data=subset(dpt, attempt>1), alpha=I(.01))+
-  facet_wrap(~experiment, scales="free_y") + ylab("Residual log(time taken)") +
+  geom_point(aes(attempt,mean_resid), data= ddt) +
+  geom_smooth(aes(attempt,mean_resid), data= ddt, method="lm", se=F) +
+  geom_smooth(aes(attempt,resid, group=id),method="lm", se=F, data=trend.dat, alpha=I(.01))+
+  facet_wrap(~experiment, scales="free_y") + ylab("Mean residual proportion correct") +
   scale_x_continuous(breaks = seq(2,10,by=2))
 
 ggsave("../images/learning_trend_subject.pdf", width=10.5, height = 3.5)

### 2013-06-14 sociology_paper.tex
supplimetary typo fixed
cleaned abstrac

@@
 The experiments suggest that individual skills vary substantially,  but demographics do not have a huge effect on performance.
-There is evidence that a learning effect exists that observers get faster with repeated evaluations which suggests a potential for using the lineup protocol for training people to read statistical graphics.
+There is evidence that a learning effect exists but only in that observers get faster with repeated evaluations, but not more often correct.
 The placement of data plot in the lineup does not affect the inference.
@@
 There could be different other features people would use to evaluate a lineup depending on the situation.
-But these words we have a general idea how people may think to evaluate a lineup.
+But with these words we get a general idea on how people may think while evaluating a lineup.
 

### 2013-06-14 sociology_paper.tex
subject specific demographic variables

--- 2013-06-14 sociology_paper.tex
@@
 Linear regression lines fitted through the points are shown.
-None of the slopes in these three experiments are statistically significant.}
+None of the slopes in these three experiments are statistically significant.
+Subject specific variations in the slopes are observed where some of the observers did worse and some did better over the successive attempts.}
    \label{fig:learning_trend_response}
@@
 Linear regression lines are fitted through the points excluding the first attempt.
-For all the three experiments the downward slopes are statistically significant which indicates that MTurk workers take less time as they progress through their attempts.}
+For all the three experiments the downward slopes are statistically significant which indicates that MTurk workers take less time as they progress through their attempts.
+But subject specific variations in slopes are also observed where some of the subjects did worse over successive attempts while others did better.}
    \label{fig:learning_trend_time}
@@
 Observer demographics seem to have significant influence on performance in time taken to evaluate a lineup.
-Some variation is observed in terms of probability of correct evaluation.
+Some variations among the factors are observed in terms of probability of correct evaluation.
 Age group 36-40, Countries other than India and United states, People who have a graduate degree appear to have significantly more probability to evaluate a lineup correctly.
@@
 Individual learning trend is observed in both time taken and observer performance.
+Some individuals improved the performances while others showed decrease in their performances over attempts.
 But the overall performance of human observers does not increase through successive attempts while evaluating multiple lineups.

### 2013-06-14 R/sociology_paper.R

@@ -485,11 +485,12 @@ ggplot() +
   scale_x_continuous(breaks = seq(2,10,by=2)) 
 
 ggplot() + 
+  geom_smooth(aes(attempt,resid, group=id),method="lm", se=F, data=subset(dpt, attempt>1), colour=rgb(0,0,0, alpha=.05))+
   geom_point(aes(attempt,mean_resid), data= dmt) +
   geom_smooth(aes(attempt,mean_resid), data= subset(dmt, attempt>1), method="lm", se=F) +
-  geom_smooth(aes(attempt,resid, group=id),method="lm", se=F, data=subset(dpt, attempt>1), alpha=I(.01))+
   facet_wrap(~experiment, scales="free_y") + ylab("Residual log(time taken)") +
-  scale_x_continuous(breaks = seq(2,10,by=2))
+  scale_x_continuous(breaks = seq(2,10,by=2))  
+#  coord_cartesian(ylim=c(-0.5,0.5))
 
 ggsave("../images/learning_trend_time_subject.pdf", width=10.5, height = 3.5)
 
@@ -557,9 +558,9 @@ ggsave("../images/learning_trend.pdf", width=10.5, height = 3.5)
 
 
 ggplot() + 
-  geom_point(aes(attempt,mean_resid), data= ddt) +
-  geom_smooth(aes(attempt,mean_resid), data= ddt, method="lm", se=F) +
-  geom_smooth(aes(attempt,resid, group=id),method="lm", se=F, data=trend.dat, alpha=I(.01))+
+  geom_smooth(aes(attempt,resid, group=id),method="lm", se=F,  data=trend.dat, colour=rgb(0,0,0, alpha=0.05))+
+  geom_point(aes(attempt,mean_resid, group=1), data= ddt) +
+  geom_smooth(aes(attempt,mean_resid, group=1), data= ddt, method="lm", se=F) +
   facet_wrap(~experiment, scales="free_y") + ylab("Mean residual proportion correct") +
   scale_x_continuous(breaks = seq(2,10,by=2))
 

### 2013-06-15 sociology_paper.tex
practical impact of factors discussed

@@
 Factors that may also affect the power of visual inference are the observers' demographics, visual skills, and experience, the sample of null plots taken from the null distribution, the position of the data plot in the lineup, and the signal strength in the data.
-This paper examine these factors.
+This paper examines these factors.
 Results from multiple visual inference studies using Amazon's Mechanical Turk are examined to provide an assessment of these.
@@
 \begin{enumerate} \leftmargin 5cm  \itemsep 0in
-\item Age level
+\item Age group
 \item Gender
@@
 \begin{equation} \label{eqn:demographic_response}
-g( \pi_{ij} )= \mu + X_{i}\beta + \ell_j,  
+g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
 \end{equation}
 where $\mu$ is an overall average probability for picking out the data plot from a lineup.
-$X_{i}=(X_{1i},X_{2i}, .., X_{pi})'$ is the vector of demographic factors of the observer $i$ and $\beta$ is the corresponding fixed effect coefficient vector.
-$p$ indicates the number of demographic factor variables.
-For $\ell_j$ we assume a normal distribution, $N(0, \sigma_\ell^2)$.
-$\ell_j$ is a random intercept predicting lineup difficulty level.
+$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effect of age group $k(i)$, gender category $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$.
+$\ell_j$ is a random intercept predicting lineup difficulty level and we assume  $\ell_j \sim N(0, \sigma_\ell^2)$.
 $g(.)$ denotes the {\it logit} link function $g(\pi)=g_1(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+
+% and $k(i) \in \{18-25, 26-30, 31-35, 36-40, 41-45, 46-50, \text{above } 50 \}$.
+$\gamma$ is the effect of gender category $l(i) \in \{\text{Male, Female}\}$.
+$\tau$ is the effect of education levels $m(i) \in \{ \text{High school, Undergrad course, Undergrad degree, Graduate course, }\}$
 
@@
 \begin{equation} \label{eqn:demographic_time}
-Z_{ij} = \mu + X_{i}\beta + \ell_j+ \epsilon_{ij},  
+Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
 \end{equation}
 where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
+$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model \eqref{eqn:demographic_response}.
 $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
@@
 The distribution of time taken to evaluate a lineup is positively skewed.
-But in log scale it is symmetric for all the demographic factor variables as we see in Figure \ref{fig:demographic_effect}.
+But in log scale it appears to be symmetric for all the demographic factor variables as we see in Figure \ref{fig:demographic_effect}.
 Mean and medians are very similar as well.
@@
 
-Model \eqref{eqn:demographic_response} is fitted to the data for different demographic factors such as age, country, education and gender.
-The results are shown in Table \ref{tbl:model_result_demographics}.
-For parameter estimation the first factor levels shown the Table are are used as the base line.
-The age group 36-40 is significantly different than the base line age group 18-25.
-The rest of the age levels appears to be similar in explaining the probability of correct response.
-For country, the rest of the world is different from USA but India appears to be similar to USA.
-We see from Table \ref{tbl:demographics} that most of the participants (about 90\%) are from India and USA.
-The rest of the 10\% data are from 76 different countries.
-This diversity in the rest of the world may make it different from India and USA.
+Model \eqref{eqn:demographic_response} is fitted to the data with fixed effect factors such as age, country, education and gender.
+To estimate the overall factor main effect, a reduced model is fitted removing that factor from Model \eqref{eqn:demographic_response}.
+Analysis of variance (ANOVA)  results shown in Table~\ref{tbl:anova_factor} suggest that all the factor variables are significantly different in describing the the probability of correct response except gender.
+Gender does not seem to show any difference in performance.
+The similar ANOVA results are obtained by fitting Model \eqref{eqn:demographic_time} to the data of log time taken and are shown in Table~\ref{tbl:anova_factor}.
+All the factor variables are significantly different for log time taken including gender.
+
+\begin{table}[htbp]
+\centering
+\caption{Anova of full model with all the demographic factors vs reduced model with removing respective factor variable. Gender does not seem to have any effect on probability of correct response.}
+\begin{tabular}{lrrrrrrr}
+  \hline
+Model & & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
+  \hline
+Proportion Correct &Full & 23822.00 &23943.00 &-11896.00  & &  &  \\ 
+& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
+  &Age & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
+  &Country  & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
+  &Education  & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
+  &Gender  & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
+   \hline
+   Log Time &Full &  51904.00 &52034.00 & -25936.00 & &  &  \\ 
+& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
+&Age  & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
+ & Country  & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
+  &Education  & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
+  &Gender  & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
+\hline
+\end{tabular}
+\label{tbl:anova_factor}
+\end{table}
+
+To further illustrate how the individual factor levels differ from each other the detailed results of Models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
+For parameter estimation the first factor levels shown in the Table are are used as the base line.
+The effect of all the demographic factors on log time taken to evaluate a lineup appears to be significant.
+But for probability of correct responses we observe that not all the factor levels are significant.
 
@@
 
+Age group 36-40 is significantly different than the base line age group 18-25.
+The rest of the age levels appears to be similar in explaining the probability of correct response.
+For country, the rest of the world is different from USA but India appears to be similar to USA.
+We see from Table \ref{tbl:demographics} that most of the participants (about 90\%) are from India and USA.
+The rest of the 10\% data are from 76 different countries.
+This diversity in the rest of the world may make it different from India and USA.
+
 The graduate degree holders are significantly different as well.
@@
 Because of this huge impact of lineup specific variance, the actual effect of other factor variables on the probability of correct response is in fact very small even though they appear to be significant.
-
-Model \eqref{eqn:demographic_time} is fitted for the log time taken to evaluate a lineup and the results are also shown in  Table~\ref{tbl:model_result_demographics}.
+We illustrate this with the following example of the significant effect of graduate degree.
+
+Effect of graduate degree is estimated as 0.182 which appears to be highly significant. This causes an increase of probability of correct response for graduate degree over high school by 
+$$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.042.$$  If we move away one standard deviation of lineup variation, in other words if we add 2.293 to teach of the exponent term in above equation, this increase reduces to 0.024.
+For two standard deviations it further reduces to 0.003 which is practically very small.
+The similar calculation shows the same negligible impact of age level 36-40 and country on the probability of correct response.
+Thus even though some of the factor levels are significant, their practical impact is negligible because of high variability of lineup difficulties.
+
+%Model \eqref{eqn:demographic_time} is fitted for the log time taken to evaluate a lineup and the results are also shown in  Table~\ref{tbl:model_result_demographics}.
 The effect of all the demographic factors on log time taken to evaluate a lineup appears to be significant.
 
-\begin{table}[htbp]
-\centering
-\caption{Anova of full model with all the demographic factors vs reduced model with removing respective factor.}
-\begin{tabular}{lrrrrrrrr}
-  \hline
-Model && Df & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
-  \hline
-Log Time &Full& 16 &  51904.00 &52034.00 & -25936.00 & &  &  \\ 
-& \multicolumn{2}{l}{Reduced} &  &  &  &  & \\
-&Age & 10 & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
- & Country & 14 & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
-  &Education & 12 & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
-  &Gender & 15 & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
-\hline
-Proportion Correct &Full& 15 & 23822.00 &23943.00 &-11896.00  & &  &  \\ 
-& \multicolumn{2}{l}{Reduced} &  &  &  &  & \\
-  &Age & 9 & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
-  &Country & 13 & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
-  &Education & 11 & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
-  &Gender & 14 & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
-   \hline
-\end{tabular}
-\end{table}
+%
+%\begin{table}[htbp]
+%\centering
+%\caption{Anova of full model with all the demographic factors vs reduced model with removing respective factor.}
+%\begin{tabular}{lrrrrrrrr}
+%  \hline
+%Model && Df & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
+%  \hline
+%Log Time &Full& 16 &  51904.00 &52034.00 & -25936.00 & &  &  \\ 
+%& \multicolumn{2}{l}{Reduced} &  &  &  &  & \\
+%&Age & 10 & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
+% & Country & 14 & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
+%  &Education & 12 & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
+%  &Gender & 15 & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
+%\hline
+%Proportion Correct &Full& 15 & 23822.00 &23943.00 &-11896.00  & &  &  \\ 
+%& \multicolumn{2}{l}{Reduced} &  &  &  &  & \\
+%  &Age & 9 & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
+%  &Country & 13 & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
+%  &Education & 11 & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
+%  &Gender & 14 & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
+%   \hline
+%\end{tabular}
+%\end{table}
+%
 
@@
    \centering
-   \includegraphics[width=6.5in]{learning_trend.pdf} 
+%   \includegraphics[width=6.5in]{learning_trend.pdf} 
     \includegraphics[width=6.5in]{learning_trend_subject.pdf} 
@@
    \centering
-   \includegraphics[width=6.5in]{learning_trend_time.pdf}    
+%   \includegraphics[width=6.5in]{learning_trend_time.pdf}    
  \includegraphics[width=6.5in]{learning_trend_time_subject.pdf}    

### 2013-06-15 R/sociology_paper.R

@@ -603,22 +603,53 @@ print(xtable(results, digits=3),  sanitize.text.function = function(x){x})
 
 
 # ---------------------------------------------------------------
-# Examining different models with response (percent correct)
+# Examining different models 
+# with response (percent correct)
 # ---------------------------------------------------------------
 
 dt <- subset(dtrend, experiment==5)
-fit0 <- lmer(response ~ I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), family="binomial", data=dt)
-fit1 <- lmer(response ~ factor(attempt) + (attempt|id) + (1|pic_id), family="binomial", data=dt)
+fit0 <- lmer(response ~ factor(attempt) + (attempt|id) + (1|pic_id), family="binomial", data=dt)
+fit1 <- lmer(response ~ I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), family="binomial", data=dt)
 fit2 <- lmer(response ~ attempt + (attempt|id) + (1|pic_id), family="binomial", data=dt)
 fit3 <- lmer(response ~ attempt + (attempt|id) , family="binomial", data=dt)
 
+anova(fit0,fit1)
 anova(fit0,fit2)
 anova(fit1,fit2)
 
+# ---------------------------------------------------------------
+# Examining different models 
+# with response (log time taken)
+# ft1 apears to be the best model since no difference with ft0
+# ---------------------------------------------------------------
+
+dt <- subset(dtrend, experiment %in% c(5,6,7))
+
+ft0 <- lmer(log(time_taken) ~ factor(attempt) + (attempt|id) + (1|pic_id),  data=dt, method="ML")
+ft1 <- lmer(log(time_taken) ~ I(attempt==1)+ attempt + (attempt|id) + (1|pic_id),  data=dt, method="ML")
+ft2 <- lmer(log(time_taken) ~ attempt + (attempt|id) + (1|pic_id),  data=dt, method="ML")
+ft3 <- lmer(log(time_taken) ~ attempt + (attempt|id) , data=dt, method="ML")
+
+anova(ft0,ft1)
+anova(ft0,ft2)
+anova(ft1,ft2)
+
+# ==============================================================
+# examining the practical significance of demographics
+# ---------------------------------------------------------------
+
+xb0 <- -0.683
+xb1 <- xb0+0.182
[…]

### 2013-06-15 sociology_paper.tex
demographics

@@
 People who have only a high school degree took less time and this is related to what we have seen for young age group.
-The time taken by male and female participants are almost similar.
+The time taken by male and female participants are almost the same.
 
@@
    \caption{Boxplots of average time taken in log scale and proportion correct responses of all the lineups plotted for each demographic factor variables.
-The dots inside the box represents mean.
+The dots inside the boxes represent means.
 Demographic factors appear to be statistically significant for explaining time taken.
@@
 Mean and medians are very similar as well.
-This allowed us to fit linear mixed model to the log time taken with normal error structure.
+This allows us to fit linear mixed model to the log time taken with normal error structure.
 
 We observed some differences in median proportion of correct responses for different levels of the demographic factors as in Figure \ref{fig:demographic_effect}.
-But the variability of proportion correct responses are huge as per design of the experiments since we have some super easy and some extremely difficult lineups.
-There were lineups which were not expected to be evaluated correctly since some of the null plots in those lineups were showing more signal than the actual plot itself.
+But the variability of proportion correct responses are huge as per design of the experiments since we have some very easy and some extremely difficult lineups.
+Some of the lineups  were not expected to be evaluated correctly as some of the null plots in those lineups were showing more signal than the actual plot itself.
 In those scenarios,  one may not reject the null hypothesis.
 We call these difficult lineup.
-Some of the lineups were so easy that data plot is detectable in a first glance.
+Some of the lineups were so easy that data plot is detectable at a first glance.
 These were expected to be evaluated correctly 100\% of the time.
 This is why we see the boxes range from 0 to 1 in most of the factor levels.
-Unlike median, less differences are observed in the average proportion of correct responses for different factor levels.
+Unlike median, fewer differences are observed in the average proportion of correct responses for different factor levels.
 
@@
 Analysis of variance (ANOVA)  results shown in Table~\ref{tbl:anova_factor} suggest that all the factor variables are significantly different in describing the the probability of correct response except gender.
-Gender does not seem to show any difference in performance.
-The similar ANOVA results are obtained by fitting Model \eqref{eqn:demographic_time} to the data of log time taken and are shown in Table~\ref{tbl:anova_factor}.
+Gender does not show any significant difference in performance.
+Similarly, ANOVA results are obtained by fitting Model \eqref{eqn:demographic_time} to the data of log time taken and are shown in Table~\ref{tbl:anova_factor}.
 All the factor variables are significantly different for log time taken including gender.
@@
 
-To further illustrate how the individual factor levels differ from each other the detailed results of Models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
-For parameter estimation the first factor levels shown in the Table are are used as the base line.
-The effect of all the demographic factors on log time taken to evaluate a lineup appears to be significant.
-But for probability of correct responses we observe that not all the factor levels are significant.
+To further illustrate how the individual factor levels differ from each other  detailed results of Models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
+For parameter estimation the first factor levels shown in the Table are  used as the base line.
+Demographic factors have significant effects on log time taken to evaluate a lineup on all their levels.
+For the probability of correct responses not all the factor levels are significant.
 
@@
 \centering
-\caption{Parameter estimates of Models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for avarage log time taken and probability of correct lineup evaluations respectively.
-For time taken all the demographic factors appear to be significant.
-For probability of correct response age group 36-40, rest of the world and graduate degree appears to be significantly different.
-Gender does not show any difference in performance.}
+\caption{Parameter estimates of Models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and probability of correct lineup evaluations respectively.
+For time taken all the demographic factors are significant.
+For probability of correct response age group 36-40, rest of the world and graduate degree are significantly different.
+For gender we do not see any difference in performance.}
 \scalebox{0.9}{
@@
 
-Age group 36-40 is significantly different than the base line age group 18-25.
-The rest of the age levels appears to be similar in explaining the probability of correct response.
+Age group 36-40 is significantly different than the base line age group of 18-25 year olds.
+The other age levels appears to be similar in explaining the probability of correct response.
 For country, the rest of the world is different from USA but India appears to be similar to USA.
@@
 The positive parameter estimate indicates that they perform better and the probability of correct response is higher than other education levels.
-As expected, there is no significant difference between Male and female performances.
+As expected, there is no significant difference between male and female performances.
 
 Notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in Model \eqref{eqn:demographic_response}.
-This indicates that the major and most important factor affecting the probability of correct response is the lineup difficulty.
+This indicates that the major and most important factor affecting the probability of correct response is  lineup difficulty.
 This is also evident from the large variability in the proportion correct in Figure~\ref{fig:demographic_effect}.
-Because of this huge impact of lineup specific variance, the actual effect of other factor variables on the probability of correct response is in fact very small even though they appear to be significant.
+Because of this huge impact of lineup specific variance, the practical impact of other factor variables on the probability of correct response is in fact very small.
 We illustrate this with the following example of the significant effect of graduate degree.
 
-Effect of graduate degree is estimated as 0.182 which appears to be highly significant. This causes an increase of probability of correct response for graduate degree over high school by 
-$$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.042.$$  If we move away one standard deviation of lineup variation, in other words if we add 2.293 to teach of the exponent term in above equation, this increase reduces to 0.024.
+Effect of graduate degree is estimated as 0.182 which is highly significant.
+For an 18-25 year old female in the United States, the expected difference in the probability of correct response, if she has a graduate degree as compared to a high school degree, is, for a lineup with random effect of zero:
+$$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.042.$$  
+For a lineup with a predicted random effect of one standard deviation, i.e.
+we move away one standard deviation of lineup variation by adding 2.293 to each of the exponent terms in the equation, this difference reduces to 0.024.
 For two standard deviations it further reduces to 0.003 which is practically very small.
-The similar calculation shows the same negligible impact of age level 36-40 and country on the probability of correct response.
+A similar calculation shows the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of correct response.
 Thus even though some of the factor levels are significant, their practical impact is negligible because of high variability of lineup difficulties.

### 2013-06-15 sociology_paper.tex
model discussion

--- 2013-06-15 sociology_paper.tex
@@
 \subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiment 5, 6, and 7 separately.
+As an alternative to Model \eqref{eqn:trend_response} we examined a reduced model with attempt as a continuous covariate.
+But that reduced model appears to be significantly different with a $p$-value of XXX.
+Thus we consider the bigger Model \eqref{eqn:trend_response} with attempt as a factor variable with 10 levels.
+Attempt 1 is considered to be base level while fitting the model.
+
 Table \ref{tbl:model_result_response} presents the parameter estimates and $p$-values of fixed effect estimates of Model \eqref{eqn:trend_response}.
@@
 
-The estimated variance ($\sigma^2_a$) of the random effect attempt is much small compared to other random effects except for subject variability ($\sigma^2_u$) for experiment 7.
+The variance of random slope for attempt ($\sigma^2_a$) is estimated to be very small compared to other random effects except for subject variability ($\sigma^2_u$) for experiment 7.
 This suggest that most of the variability for attempt is accounted for by the fixed effect factors of attempt.
@@
 But for experiment 6 it does not seem to be significant suggesting that experiment 6 lineups were difficult compared to the other experimental lineups.
-This shows the designs of experiments such that there exists the mixture of difficult and easy lineups within the experiments as well as between the experiments.
+This shows a feature of the experimental designs where a mixture of difficult and easy lineups were included within the experiments as well as between the experiments.
 For easy lineups there may be small chance to improve performances.
@@
 Even though positive slopes over the attempts are observed, none of the trends is statistically significant.
-
+Subject specific least square lines shows large overall variability.
+Some subjects show improvement over time and some show the decrease in performance.
+The subject specific slopes are very different as well.
+
+Model \eqref{eqn:trend_time} was compared to couple of other alternatives.
+A bigger model with attempt as a factor was considered but they did not appear to be significantly different with $p$ value XXX.
+A reduced model with attempt as a continuos covariate was also considered but they were significantly different with $p$-value XXX.
 Table \ref{tbl:model_result_time} presents the results of Model \eqref{eqn:trend_time}.
@@
 
-\subsection{How People Pick the Data Plot} In the experiment setup we have a question for the turk observers to choose the reason for their selection of a particular data plot.
+\subsection{How People Pick the Data Plot} In each of the experiments we had a question for the turk observers to choose the reason for their selection of a particular data plot.
 It is shown by \cite{majumder:2013} that these choice reasons reveal the way people picks the data plot in the lineup.

### 2013-06-15 R/sociology_paper.R

@@ -607,7 +607,7 @@ print(xtable(results, digits=3),  sanitize.text.function = function(x){x})
 # with response (percent correct)
 # ---------------------------------------------------------------
 
-dt <- subset(dtrend, experiment==5)
+dt <- subset(dtrend, experiment %in% c(5,6,7))
 fit0 <- lmer(response ~ factor(attempt) + (attempt|id) + (1|pic_id), family="binomial", data=dt)
 fit1 <- lmer(response ~ I(attempt==1)+ attempt + (attempt|id) + (1|pic_id), family="binomial", data=dt)
 fit2 <- lmer(response ~ attempt + (attempt|id) + (1|pic_id), family="binomial", data=dt)

### 2013-06-15 sociology_paper.tex
last update before sending

@@
 For an 18-25 year old female in the United States, the expected difference in the probability of correct response, if she has a graduate degree as compared to a high school degree, is, for a lineup with random effect of zero:
-$$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.042.$$  
+$$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.041.$$ 
 For a lineup with a predicted random effect of one standard deviation, i.e.
@@
 As an alternative to Model \eqref{eqn:trend_response} we examined a reduced model with attempt as a continuous covariate.
-But that reduced model appears to be significantly different with a $p$-value of XXX.
-Thus we consider the bigger Model \eqref{eqn:trend_response} with attempt as a factor variable with 10 levels.
+They did not appear to be significantly different with a $p$-value of 0.856.
+But we consider the bigger Model \eqref{eqn:trend_response} with attempt as a factor variable with 10 levels.
 Attempt 1 is considered to be base level while fitting the model.
@@
 None of the slopes in these three experiments are statistically significant.
-Subject specific variations in the slopes are observed where some of the observers did worse and some did better over the successive attempts.}
+Subject specific least square lines show large subject specific variability.
+Some of the observers did worse and some did better over the successive attempts.}
    \label{fig:learning_trend_response}
@@
 The averages of these residuals for each of the attempts are plotted against attempt as shown in Figure \ref{fig:learning_trend_response}.
-Least square linear regression lines are drown through the points.
+Least square linear regression lines are fitted through the points.
 Even though positive slopes over the attempts are observed, none of the trends is statistically significant.
-Subject specific least square lines shows large overall variability.
+Subject specific least square lines show large overall variability.
 Some subjects show improvement over time and some show the decrease in performance.
@@
 Model \eqref{eqn:trend_time} was compared to couple of other alternatives.
-A bigger model with attempt as a factor was considered but they did not appear to be significantly different with $p$ value XXX.
-A reduced model with attempt as a continuos covariate was also considered but they were significantly different with $p$-value XXX.
+A bigger model with attempt as a factor was considered but they did not appear to be significantly different with $p$ value 0.236.
+A reduced model with attempt as a continuos covariate was also considered but they were significantly different with $p$-value $<0.001$.
 Table \ref{tbl:model_result_time} presents the results of Model \eqref{eqn:trend_time}.
@@
 The negative estimates suggest that later attempt took less time for an evaluation.
-This suggest that even though observers did not improve the performance over attempt, they became efficient in responding faster.
+It is interesting to note that even though observers did not improve the performance over attempt, they became efficient in responding faster.
 The parameter $\alpha_1$ for first attempt is also highly significant.
@@
 This suggest that the harder lineups not necessarily took more time than easier lineups.
-The observers spent enough times to evaluate a easy lineup and for difficult lineup they might just give up at some point and gave the feedback.
+The observers spent enough times to evaluate a easy lineup and for difficult lineup they might just give up at some point and provided the feedback.
 
@@
 These results indicate that after working with the multiple tasks each observer became efficient which helped them finish the task faster in later attempts.
+Subject specific least square lines show variations in slopes and overall time taken.
+Even though overall skill improved, there are some subjects who took more times in later attempts.
 
@@
 This pattern is evident in the figure as we see proportion correct for null plot 5 is consistently above the null plot 1 for each of the locations.
-The overall average proportions correct responses for each location are shown using dashed lines.
+The overall average proportion of correct responses for each location are shown using dashed lines.
 As we see in the plot  the overall proportion is not varying much for different locations.
@@
 For location 1, we did not have any response for null set 1 in one of the interaction lineups.
-The larger variability observed for Interaction lineups could be controlled at some extent by collecting more data for each location.
-Since the variability is larger when we have small number of responses as we see in the figure.
+We observed larger variability for interaction lineups as compared to Genotype lineups.
+It is because some of the interaction lineups were evaluated only once.
 For example, we have only one response for null plot 3 and 5 for Interaction effect which should lead to the extreme proportion of correct since one response would either be correct or wrong.
+The larger variability observed for Interaction lineups can be controlled at some extent by evaluating them multiple times.
 
@@
 The results are shown in Table \ref{tbl:manova}.
-The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05 and we failed to reject the hypothesis that there is no difference in location.
+The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05.
+This suggest that there may be no difference in location.
 
@@
 Some variations among the factors are observed in terms of probability of correct evaluation.
-Age group 36-40, Countries other than India and United states, People who have a graduate degree appear to have significantly more probability to evaluate a lineup correctly.
+Age group 36-40, Countries other than India and United states, People who have a graduate degree appear to be significantly different.
+But their practical impact on probability to correctly evaluate a lineup is very small and in some cases very negligible as the lineup difficulties increase or decreases.
 Gender does not have any effect on performance.
@@
 Some individuals improved the performances while others showed decrease in their performances over attempts.
-But the overall performance of human observers does not increase through successive attempts while evaluating multiple lineups.
+But the overall performance of the observers does not increase through successive attempts while evaluating multiple lineups.
+This result is very important as it suggests that the power estimated for visual inference using human subject experiment is robust and may not change if those participants are allowed to give feedback again.
 The skill in evaluating the lineup in shorter time gets improved over successive evaluations.
 The earlier evaluations take significantly longer than the later evaluations.
-This result is very important as it suggests that the power estimated for visual inference using human subject experiment is robust and may not change if those participants are allowed to give feedback again.
 This investigation suggest that the skilled person may only do it faster.
@@
 The numerous pilot studies done with the more trained participants suggest that power of visual statical inference may be higher with knowledgeable observer.
-Thus the variations observed with individual performances could be attributed to individual abilities in terms of their training on statistical graphics.
+The fact that a person with graduate degree performs better may be associated with this.
+But a graduate degree does not necessarily mean to have training on statistical graphics.
 More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of knowledge about statistical graphics.
 
-This investigation allowed each observer to evaluate 10 lineups assuming that it would not cause fatigue or disinterest of the observers towards the experiment based on the pilot studies.
+This investigation allowed each observer to evaluate 10 lineups assuming that it would not cause fatigue or disinterest toward the task of evaluating lineups.
+This assumption is made based on the pilot studies.
 The future research may involve doing more experiments to check if fewer or more than 10 lineups have any significant impact on performance of the observer.
@@
 At some points it should level of.
-It would be useful to know for how many lineups time taken levels off, since that would indicate the limit beyond which no learning occurs.
-
-Also, a lineup with fewer or larger than 20 plots may yield different results.
+It would be useful to know for how many lineups time taken levels off.
+
+A lineup with fewer or larger than 20 plots may yield different results.
 If the size of the lineup is much higher than 20, it is intuitive that there may be location effect of the data plot in the lineup.
 It is because, the observer may get tiered of scanning and make decision based on the partial scanning of the lineup.
-On the other hand fewer than 20 plots would give observer compare the actual plot with fewer null plots give more chance of picking actual plot as an error not as an actual plot.
-These are the some of the issues that needs to investigated.
+On the other hand fewer than 20 plots will allow observer to compare the actual plot with fewer null plots giving more chance of picking actual plot as an error not as an actual plot.
+These are the some of the issues that require further investigation.
 
@@
 
+\section{Some Plots of Exploratory Data Analysis}
+
 \begin{figure}[htbp] 

### 2013-06-15 R/sociology_paper.R

@@ -487,7 +487,7 @@ ggplot() +
 ggplot() + 
   geom_smooth(aes(attempt,resid, group=id),method="lm", se=F, data=subset(dpt, attempt>1), colour=rgb(0,0,0, alpha=.05))+
   geom_point(aes(attempt,mean_resid), data= dmt) +
-  geom_smooth(aes(attempt,mean_resid), data= subset(dmt, attempt>1), method="lm", se=F) +
+  geom_smooth(aes(attempt,mean_resid), data= subset(dmt, attempt>1), method="lm", se=F, size=I(1.2)) +
   facet_wrap(~experiment, scales="free_y") + ylab("Residual log(time taken)") +
   scale_x_continuous(breaks = seq(2,10,by=2))  
 #  coord_cartesian(ylim=c(-0.5,0.5))
@@ -560,7 +560,7 @@ ggsave("../images/learning_trend.pdf", width=10.5, height = 3.5)
 ggplot() + 
   geom_smooth(aes(attempt,resid, group=id),method="lm", se=F,  data=trend.dat, colour=rgb(0,0,0, alpha=0.05))+
   geom_point(aes(attempt,mean_resid, group=1), data= ddt) +
-  geom_smooth(aes(attempt,mean_resid, group=1), data= ddt, method="lm", se=F) +
+  geom_smooth(aes(attempt,mean_resid, group=1), data= ddt, method="lm", se=F, size=I(1.2)) +
   facet_wrap(~experiment, scales="free_y") + ylab("Mean residual proportion correct") +
   scale_x_continuous(breaks = seq(2,10,by=2))
 

### 2013-06-15 sociology_paper.tex
some typo

@@
 For some experiments India got more participants than United States and for some experiments the numbers just got reversed (\ref{fig:participation_time}).
-It is also observed in some experiments number of  participants are similar no matter whether the experiment is run on day time or nigh time.
+It is also observed in some experiments that number of  participants are similar no matter whether the experiment is run on day time or nigh time.
 
@@
 Notice that fewer people participated from India beyond age 40 compared to united states.
-Even though total participations from india is larger compared to United States, they are mostly young people who are capable of using internet and computer.
+Even though total participations from india is very close to United States, they are mostly young people who are capable of using internet and computer.
 

### 2013-06-15 sociology_paper.tex
typo fixed

@@
 The corresponding $p$-value is 0.00023, indicating sufficient evidence to reject the null hypothesis.
-, indicating that we have enough evidence to reject the null hypothesis.
+
 What does this mean, though?
@@
 Because each poll has a margin of error, this is used to simulate different scenarios that might have resulted on election day, if the polls were on target.
-A null data set is then generated as a set of draws from normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
+A null data set is generated as a set of draws from normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
 These samples are plotted as electoral buildings, and the plot made with the results from the election is placed randomly among them in a lineup of size $5\times 4$.

### 2013-06-15 sociology_paper.tex
reviewd upto location effect

@@
 \subsection{Choice of Visual Test Statistic}  The visual test statistic should be highly associated with the hypothesis under investigation.
-To achieve this purposes it is very important to decide which plot type and plot features should be adopted.
+To achieve this purpose it is very important to decide which plot type and plot features should be adopted.
 In a linear regression setting, the visual test statistics are presented in \citep{majumder:2013} for some common hypothesis testing.
@@
 Table \ref{tbl:visual_stat} shows the visual test statistics used in this study.
-For both experiment 2 and 3 a scatterplot is used but a regression line fitted through the points is overlaid for experiment 2.
+For both experiments 2 and 3 a scatterplot is used but a regression line fitted through the points is overlaid for experiment 2.
 It is easier to spot the slope of the line compared to just the scatterplot itself.
 But it appears that people are better in noticing the unusual pattern in the data which is some contamination purposefully added for experiment 3.
-It is observed that for experiment 3 the performance of the observers are much better compared to experiment 2 \citep{majumder:2013}.
+For experiment 3 the performance of the observers was much better compared to experiment 2 \citep{majumder:2013}.
 
@@
 \subsection{Question that Human Observer Answers} The researcher knows about the underlying hypothesis but the observer does not necessarily know the underlying details of the lineup.
-So, the researcher needs to ask a question observer is to answer while evaluating the lineup.
+So, the researcher needs to ask a question to the observer to answer while evaluating the lineup.
 This question should provide the observer a little clue so that the answer reflects the hypothesized patterns in the actual plot.
-For example Table \ref{tbl:visual_stat} shows the questions asked for the simulation experiments done by  \citet{majumder:2013}.
+For example Table \ref{tbl:visual_stat} shows the questions asked for the simulation experiments considered for this study.
 Notice that for case 1 if the observer can identify the actual plot in the lineup that should indicate that the plot chosen has the most vertical difference between groups A and B which is exactly what the researchers intend to examine.
@@
 Notice that there may be unnecessary patterns in the actual data plot which may not necessarily indicate the existence of the significant signal in the plot.
-These question help observer not to be misguided by those patterns.
-To review further on this \citet{majumder:2013} have also asked why the observers choose a specific plot.
+These questions help observer not to be misguided by those patterns.
+To follow up further on this \citet{majumder:2013} also collected reasons of choice made by the observers and noticed that Type-III error may occur where the hypothesis may be correctly rejected but for a wrong reason.
 
@@
 
-\subsection{Demographics of Observer}  Some of the demographics of the observer such as age, gender, education level and geographical location may have effect on how an observer would examine the lineup.
+\subsection{Demographics of the Observer}  Some of the demographics of the observer such as age, gender, education level and geographical location may have effect on how an observer would examine the lineup.
 This may produce some variability in the performance of the observer.
 For example, a high school student may not respond same as a well educated person while evaluating a lineup.
-There may be variation in different age groups as well.
-Recruiting subjects with a well variety of age groups is be necessary to study this variability.
+There may be variations in different age groups as well.
+Investigation on subjects with a well variety of age groups is necessary to study this variability.
 
@@
 The more evaluations an observer makes the more skillful the observer may become.
-This learning trend may or may not be significant overall but it could influence the performance of the observer in some way.
+This learning trend may or may not be significant overall but it may influence the performance of the observer in some way.
 
@@
 An observer can become more trained on the lineup structure and the pattern in the plots shown.
-This could be learning in proportion of correct evaluation.
+This can be learning in proportion of correct evaluation.
 Or, he or she may become efficient in responding faster in the later evaluations.
-This could be observed as the time taken for evaluating a lineup.
+This can be observed as the time taken for evaluating a lineup.
 This paper investigates both of these learning trends with multiple simulation experiments.
@@
 When an observer evaluate a lineup for the first time, he or she has to read through many instructions and become familiar with the environment of the experiment.
-We expect that this would require much more time compared to the rest of the evaluations.
-Time taken for the second or successive attempt would be much lower than the time taken for the first lineup.
+We expect that this will require much more time compared to the rest of the evaluations.
+Time taken for the second or successive attempt should be much lower than the time taken for the first lineup.
 We attribute this to the adaptation to the experiment environment, not a learning of evaluating a lineup.
@@
 While evaluating the lineup some people may start looking from some specific part of the lineup.
-With the help of eye-tracking equipment \citet{zhao:2012} track the observers' eyes to see how they go through the plots in a lineup to come to their answers.
+With the help of eye-tracking equipments \citet{zhao:2012} tracked the observers' eyes to see how they went through the plots in a lineup to come to their answers.
 The results suggest that people have particular methods of reading lineups.
-Some people read lineups from left to write direction while some read from upward to downward.
+Some people read lineups from left to right direction while some read from upward to downward.
 Some people start looking from the center of the lineup while others start from the top left corner.
 In the earlier phase of the exercise, the observer tend to scan the plots and start comparing plots to make a final decision.
-Beside right-left or up-down directions observer show some diagonal movement too.
+Beside right-left or up-down directions observers show some diagonal movement too.
 
@@
 
-It may be possible that some part of the lineup is little explored or not scanned by the observer if he or she feels that the actual plot is found before even seriously scanning the whole lineup.
+It may be possible that some part of the lineup is less explored or even never scanned by the observer if he or she feels that the actual plot is found before even seriously scanning the whole lineup.
 In those scenarios the location where the observer first start scanning is important.

### 2013-06-15 sociology_paper.tex
reviewed upto model specifications

@@
 This influences the response of the observer.
-Also subject specific variation in the power of visual test is observed in \citep{majumder:2013}.
+Also subject specific variation in the power of visual test is observed in \cite{majumder:2013}.
 
 \section{Experimental Designs and Methods}\label{sec:exp_design}
-The performance of the observer while evaluating a lineup would depend on the factors described in Section \ref{sec:factor_performance}.
-Some of these factors such as signal in the data and individual performances are studied in \cite{majumder:2013}.
+While evaluating a lineup, the performance of the observer may depend on the factors described in Section \ref{sec:factor_performance}.
+Some of these factors such as signal in the data and individual abilities were studied in \cite{majumder:2013}.
 In this paper we intend to study effect of the factors such as human demographics, learning trend, actual plot locations and selection of null plots.
-This section presents the simulation experimental setup, data collection methods and data analysis plans with models to asses the influence of those factors on visual statistical inference.
+This section presents the simulation experiment setup, data collection methods and data analysis plans with models to asses the influence of those factors on visual statistical inference.
 
@@
 
-\subsubsection{Learning Trend} Learning trend of an observer can be observed in terms of performance over successive feedback when multiple lineups are shown for evaluation.
+\subsubsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive feedbacks received when multiple lineups are shown for evaluation.
 Experiments 5, 6 and 7 were used for this.
@@
 The lineups are not necessarily with the same difficulty levels.
-But the order of lineups is randomized.
-The responses of the lineups are recorded by attempt 1 through 10.
-Attempt 1 means the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
+But the order of lineups were randomized.
+The responses of the lineups were recorded by attempt 1 through 10.
+Attempt 1 means that the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
 The goal is to estimate whether performance of the observer improves from attempt 1 to attempt 10.
 
-\subsubsection{Design for Location Effect} The experiment was set up based on the gene expression data \citep{atwood:2013} with two groups.
-The groups are defined based on the Genotype and Empty Vector (EV).
-For each of the of the Genotype, gene expression data were collected in presence or absence of iron sufficiency.
-Two hypothesis were of primary interest.
-The Genotype main effect and the Genotype and iron sufficiency interaction effect.
-
+\subsubsection{Design for Location Effect} To assess the location effect of a data plot in a lineup, we set up an experiment based on the gene expression data \citep{atwood:2013} with two groups.
+One is the Genotype and the other the Empty Vector (EV).
+For each of the groups, gene expression data were collected in presence or absence of iron sufficiency.
+Two factors were of primary interest.
+One is the Genotype main effect and the other is the interaction effect between Genotype and iron sufficiency.
+
+For Genotype main effect side by side dot plots of two groups are used as the visual test statistics.
+For interaction effect the same plot type is used except that the means of iron sufficiency or insufficiency were connected by lines colored by the groups.
+For no interaction, the visual test statistic shows two parallel lines representing two groups and for significant interaction the lines are not parallel.
 Null plots were generated by randomizing the group structure of the data.
@@
 
-Each observer was shown three lineups.
-First lineup shown was a randomly selected Interaction lineup, the second one was from randomly selected Genotype lineup.
+Each observer saw three lineups, first an Interaction lineup then a Genotype lineup.
 Finally a test lineup was shown to screen out unusual evaluation.
-The subjects were not aware which one is the test plot but it was informed in the starting of the task about having a test plot.
-The test plot was very easy and everyone should detect the data plot.
-The MTurk worker was paid based on whether he or she could correctly detect the data plot.
-But the data on test plot are excluded from the analysis.
-
-\subsection{Data Collection Methods}  Human observers were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \citep{turk} or MTurk Web site.
+The subjects did not know which one was the test lineup but they were informed before accepting the task that there would be one.
+The test lineup was very easy and everyone should detect the data plot.
+The MTurk workers were paid based on whether they could correctly evaluate the test lineup.
+But the data on the test plot are excluded from the analysis.
+
+\subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \citep{turk} or MTurk Web site.
 It is an online work place where people from around the world can perform some tasks and get paid.
@@
 Being a human is the main requirement.
-Tasks are designed for anyone to do but some tasks may require that workers satisfy some skill level depending on the recruiters need.
+Tasks are designed for anyone to do but some tasks may require some skills depending on the recruiters' need.
 Each task is usually planned to complete in a short time.
@@
 
-We designed and developed a web application which enables to display the lineups to the observers as per experimental design.
-The MTurk workers were redirected to the web site and the data were collected, stored automatically into a local database server.
+We designed and developed a web application which enables to display the lineups to the observers as per experimental need.
+The MTurk workers were redirected to the web site to evaluate lineups.
+The data were collected, stored automatically into a local database server.
 Demographic informations such as age group, gender and education levels were also collected.
@@
 \centering
+\scalebox{.9}{
 \begin{tabular}{rlrrr}
@@
    \hline
-\end{tabular}\label{tbl:demographics}
+\end{tabular}
+}
+\label{tbl:demographics}
 \end{table}

### 2013-06-16 sociology_paper.tex
reviewed upto demographic factors

@@
 The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
+It is measured in seconds.
 The location of the observer is determined by the ip address of the observer.
@@
 Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$.
-Yij = 1 if the response is correct otherwise Yij = 0.
+$Y_{ij} =1$ if the response is correct otherwise $Y_{ij} =0$.
 Let $\pi_{ij}=  E(Y_{ij})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$.
@@
 where $\mu$ is an overall average probability for picking out the data plot from a lineup.
-$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effect of age group $k(i)$, gender category $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$.
+$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender category $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$.
 $\ell_j$ is a random intercept predicting lineup difficulty level and we assume  $\ell_j \sim N(0, \sigma_\ell^2)$.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=g_1(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
 
@@
 $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$. $u_i$ is a random intercept, describing a basic subject-specific ability. We assume $u_i \sim N(0, \sigma_u^2)$. 
-$a_i$ is a random slope capturing an individuals' specific learning effect over the course of $K$ attempts. We assume $a_i \sim N(0, \sigma_a^2)$. 
+$a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts. We assume $a_i \sim N(0, \sigma_a^2)$. 
 For $\ell_j$ we again assume a normal distribution, $N(0, \sigma_\ell^2)$.
 $\ell_j$ is a random intercept predicting lineup difficulty level.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=g_1(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ .
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ .
 
@@
 
-Suppose $Y=(Y_1,Y_2, ...
-, Y_p)$ is a vector of random variable with dimension $p=5$ representing the response for five null sets.
-let $Y_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., 5$.
+Suppose $\mathbf{Y}=(Y_1,Y_2, ...
+, Y_p)^\top$ be a vector of random variable with dimension $p$, the total number of null sets.
+let $\mathbf{Y}_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., I$.
 We fit the following MANOVA model
 \begin{equation}\label{manova}
-Y_{ij} = \mu_{i} + \epsilon_{ij}
+\mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
 \end{equation}
-where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$.
+where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$.
 
@@
 
-\subsection{Overview of the Data} Figure \ref{fig:turker_location} displays the location around the world from where a total of 2321 participants provided feedback data on the lineups in all the ten experimental studies.
-Even though most of the participants were from the United States and India, we have participants from 76 other different countries.
-This provides a diverse pool of participants in terms of geographical locations.
+\subsection{Overview of the Data} A total of 2321 participants provided feedback data on the lineups in ten different experimental studies.
+Figure \ref{fig:turker_location} displays the locations of participants around the world.
+Most of the participants were from the United States and India.
+There were 76 other different countries from where we received feedbacks.
+This provides a diverse pool of participants.
 The diversity in not only in geographical locations of the participants but also in their gender, age groups and education levels as we see in Table \ref{tbl:demographics}.
@@
 The rest of the 70 countries have less than 10 participants.
-Figure \ref{fig:turker_location_experiment} displays the location of turk workers by each experiment.
-Notice that For all the ten experiments the locations are similar with some small variations.
-That may be due to the time of the experiment when it is started.
-For some experiments India got more participants than United States and for some experiments the numbers just got reversed (\ref{fig:participation_time}).
+The distribution of participants remains almost similar in all ten experiments with some small variations (Figure~\ref{fig:turker_location_experiment}).
+That may be due to the time of the experiment when it first started.
+For some experiments India got more participants than United States and for some experiments the numbers just got reversed (Figure~\ref{fig:participation_time}).
 It is also observed in some experiments that number of  participants are similar no matter whether the experiment is run on day time or nigh time.
@@
 \centering
-\scalebox{.9}{
-\begin{tabular}{rlrrr}
-  \hline
- Factor & levels & subjects & avg\_time & response \\ 
-  \hline
-Gender & Male & 1348 & 48.51 & 13493 \\ 
-   & Female & 991 & 43.75 & 10564 \\ 
-\hline
-  Education & High school or less & 193 & 37.21 & 2241 \\ 
-   & Some under graduate courses & 418 & 42.84 & 4070 \\ 
-   & Under graduate degree & 584 & 44.29 & 5775 \\ 
-   & Some graduate courses & 245 & 43.43 & 2460 \\ 
-   & Graduate degree & 902 & 52.18 & 9511 \\ 
-\hline
-  Age & 18-25 & 740 & 42.97 & 7311 \\ 
-   & 26-30 & 547 & 46.27 & 5585 \\ 
-   & 31-35 & 376 & 44.27 & 3923 \\ 
-   & 36-40 & 257 & 55.03 & 2714 \\ 
-   & 41-45 & 141 & 43.90 & 1519 \\ 
-   & 46-50 &  95 & 49.29 & 1003 \\ 
-   & 51-55 &  83 & 48.67 & 867 \\ 
-   & 56-60 &  64 & 59.73 & 678 \\ 
-   & above 60 &  38 & 48.67 & 457 \\ 
-\hline   
-  Country  & United States & 1087 & 39.64 & 10769 \\  
-  & India & 980 & 52.63 & 10227 \\ 
-  & Rest of the world & 254 & 46.86 & 2819 \\
+\scalebox{.85}{
+\begin{tabular}{rlrrcr}
+  \hline
+  && \multicolumn{2}{c}{Participants}&Average&\\
+  \cline{3-4}
+ Factor & levels & Total &\%Factor & time & response \\ 
+  \hline
+Gender & Male & 1348&57.63 & 48.51 & 13493 \\ 
+   & Female & 991&42.37 & 43.75 & 10564 \\ 
+\hline
+  Education & High school or less & 193 &8.24& 37.21 & 2241 \\ 
+   & Some under graduate courses & 418 & 17.85&42.84 & 4070 \\ 
+   & Under graduate degree & 584 &24.93& 44.29 & 5775 \\ 
+   & Some graduate courses & 245 &10.46& 43.43 & 2460 \\ 
+   & Graduate degree & 902 &38.51& 52.18 & 9511 \\ 
+\hline
+  Age & 18-25 & 740 &31.61& 42.97 & 7311 \\ 
+   & 26-30 & 547 &23.36& 46.27 & 5585 \\ 
+   & 31-35 & 376 &16.06& 44.27 & 3923 \\ 
+   & 36-40 & 257 &10.98& 55.03 & 2714 \\ 
+   & 41-45 & 141 &6.02& 43.90 & 1519 \\ 
+   & 46-50 &  95 &4.05& 49.29 & 1003 \\ 
+   & 51-55 &  83 &3.54& 48.67 & 867 \\ 
+   & 56-60 &  64 & 2.73& 59.73 & 678 \\ 
+   & above 60 & 38 &1.62& 48.67 & 457 \\ 
+\hline   
+  Country  & United States & 1087 &46.83& 39.64 & 10769 \\  
+  & India & 980 &42.22& 52.63 & 10227 \\ 
+  & Rest of the world & 254&10.94 & 46.86 & 2819 \\
 
@@
 
+%\begin{table}[hbtp]
+%\caption{Demographic information of the subjects participated the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
+%\centering
+%\scalebox{.9}{
+%\begin{tabular}{rlrrr}
+%  \hline
+% Factor & levels & subjects & avg\_time & response \\ 
+%  \hline
+%Gender & Male & 1348 & 48.51 & 13493 \\ 
+%   & Female & 991 & 43.75 & 10564 \\ 
+%\hline
+%  Education & High school or less & 193 & 37.21 & 2241 \\ 
+%   & Some under graduate courses & 418 & 42.84 & 4070 \\ 
+%   & Under graduate degree & 584 & 44.29 & 5775 \\ 
+%   & Some graduate courses & 245 & 43.43 & 2460 \\ 
+%   & Graduate degree & 902 & 52.18 & 9511 \\ 
+%\hline
+%  Age & 18-25 & 740 & 42.97 & 7311 \\ 
+%   & 26-30 & 547 & 46.27 & 5585 \\ 
+%   & 31-35 & 376 & 44.27 & 3923 \\ 
+%   & 36-40 & 257 & 55.03 & 2714 \\ 
+%   & 41-45 & 141 & 43.90 & 1519 \\ 
+%   & 46-50 &  95 & 49.29 & 1003 \\ 
+%   & 51-55 &  83 & 48.67 & 867 \\ 
+%   & 56-60 &  64 & 59.73 & 678 \\ 
+%   & above 60 &  38 & 48.67 & 457 \\ 
+%\hline   
+%  Country  & United States & 1087 & 39.64 & 10769 \\  
+%  & India & 980 & 52.63 & 10227 \\ 
+%  & Rest of the world & 254 & 46.86 & 2819 \\ 
+%  
+%   \hline
+%\end{tabular}
+%}
+%\label{tbl:demographics}
+%\end{table}
+
 The largest number of participants are from age group 18 to 25 which is the youngest age group in the study.
-Interestingly we see many participants from older age groups as well.
-Specially for united states almost all the age groups show uniform participations after age 30 as in Figure \ref{fig:demographic_info}.
+The majority of the people have ages between 18 to 35.
+Interestingly there are many participants from older age groups as well.
+Specially for united states almost all the age groups show uniform participations beyond age 30 (Figure~\ref{fig:demographic_info}).
 Notice that fewer people participated from India beyond age 40 compared to united states.
-Even though total participations from india is very close to United States, they are mostly young people who are capable of using internet and computer.
-
-Most of the participants are graduate degree holder.
+Total number of participations from india and United States are almost same with United States having 107 more participants.
+Unlike USA Indian participants are mostly young people.
+
+In terms of participant's academic background, the largest group is graduate degree.
+A total of 902 participants have a graduate degree which is about 38.51\%.
 They are mostly from India as any university degree is considered a graduate degree unlike north America where graduate degree means Masters level education.
-This is apparent in Figure \ref{fig:demographic_info}.
-Most of the unites states participants are with an Undergraduate degree or at least have some undergraduate courses.
-
-Male and female participants are similarly distributed among all age groups except age 18-25 in India where less female participants were observed as we see in Figure \ref{fig:gender_country_bar}.
+Most of the USA participants are with an Undergraduate degree or at least have some undergraduate courses (Figure~\ref{fig:demographic_info}).
+
+The distribution of male and female participants are similar among all age groups except age 18-25 in India where fewer female participants were observed (Figure \ref{fig:gender_country_bar}).
 The distribution of education levels are also different across the countries for this age group.
@@
 There were not many participants beyond age 50 and hence these age groups are merged to form one age group called above 50.
-In exploratory analysis and model fitting this new age group is used instead of actual age groups beyond age 50.
-
-A total of 1911 lineups were evaluated in all these 10 experiments.
-Each person was supposed to evaluate 10 lineups except experiment 9 where three lineups were evaluated by each person.
-A test plot was shown to each observer and the feedback received from that plot is used to examine the data quality.
+The exploratory analysis and the models fitted in the following sections use this new age group instead of actual age groups beyond age 50.
+
+A total of 1911 lineups were evaluated in the ten experiments.
+Each person evaluated at least 10 lineups except for experiment 9 where three lineups were evaluated by each person.
+A test plot was shown to each observer and the feedback received from that plot is used to examine the data quality or process payment.
 But the data received for the test lineup is not included in the analysis.
@@
 \caption{Anova of full model with all the demographic factors vs reduced model with removing respective factor variable. Gender does not seem to have any effect on probability of correct response.}
-\begin{tabular}{lrrrrrrr}
+\scalebox{.9}{
+\begin{tabular}{rlrrrrrr}
   \hline
@@
 \end{tabular}
+}
 \label{tbl:anova_factor}

### 2013-06-16 R/sociology_paper.R

@@ -779,6 +779,16 @@ int.dat$locs <- "out"
 int.dat$locs[int.dat$plot_location %in% c(9,12)] <- "in"
 summary(manova(cbind(null_1,null_2,null_3,null_4,null_5)~factor(locs), data=int.dat))
 
+# ===========================================================
+# Modeling location effect
+# Fiting generalized mixed model for location effect
+# Locations are not significant
+# -----------------------------------------------------------
+
+colnames(df)
+fi <- lmer(response~factor(plot_location)+nulls+(1|pic_name), family="binomial",data=subset(df,plot_type=="Interaction"))
+fg <- lmer(response~factor(plot_location)+nulls+(1|pic_name), family="binomial",data=subset(df,plot_type=="Genotype"))
+
 # test for equivalency
 library(equivalence)
 

### 2013-06-16 sociology_paper.tex
figure showing diminishing impact of factors

@@
 For two standard deviations it further reduces to 0.003 which is practically very small.
+Figure~\ref{fig:practical_impact_demographics} illustrates this example showing how the change in probability diminishes for larger variability of lineup difficulty.
 A similar calculation shows the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of correct response.
 Thus even though some of the factor levels are significant, their practical impact is negligible because of high variability of lineup difficulties.
+
+\begin{figure}[htbp] 
+   \centering
+%   \includegraphics[width=6.5in]{learning_trend.pdf} 
+    \includegraphics[width=3in]{practical_impact_demographics.pdf} 
+   \caption{ Change in probability of correct response due to graduate degree as compared to high school degree diminishes rapidly as the variability in lineup difficulty $\sigma_{\ell}$ increases.}
+   \label{fig:practical_impact_demographics}
+\end{figure}
 

### 2013-06-16 R/sociology_paper.R

@@ -308,7 +308,59 @@ anova.prop <- rbind(data.frame(round(pag[1,1:4],2),round(pag[2,5:7]/1,2)),
 anova.reseult <- rbind(anova.time, anova.prop)
 xtable(anova.reseult)
 
+# ==============================================================
+# Exploring the practical significance of demographics
+# ---------------------------------------------------------------
 
+et <- 6
+xb <- seq(-et,et, by=.1)
+px <- function(xb){exp(xb)/(1+exp(xb))}
+pdat <- data.frame(xb,px=px(xb))
+
+s1 <- 2.293
+s2 <- s1*2
+xx <- c(0,0.182)
+
+x <- c(xx,-et,-et,xx+s1,-et,-et,xx+s2,-et,-et)
+xend <- c(xx,xx,xx+s1,xx+s1,xx+s2,xx+s2 )
+y <- c(0,0,px(xx),0,0,px(xx+s1),0,0,px(xx+s2))
+yend <- px(xend)
+ldat <- data.frame(x,xend,y,yend)
+
+ggplot()+geom_line(aes(xb,px), data=pdat)+
+  geom_segment(aes(x=x,xend=xend,y=y,yend=yend), data=ldat, linetype="dashed") + 
+  theme_bw() + xlab(expression(eta)) + ylab("Prability of correct evaluation")
+
+
+get_change <- function(s){
+  xb0 <- -0.683 + s
+  xb1 <- xb0 + 0.182 
+  px(xb1) - px(xb0)
+}
+
+sgl <- 2.293
+get_change(c(0,sgl,sgl*2))
+get_change(2.293)
+
+xx <- (1:2)*sgl
+x <- c(xx,0,0)
+xend <- c(xx,xx)
+y <- c(0,0, get_change(xx))
+yen <- get_change(xend)
+ldat <- data.frame(x=c(xx,0,0), xend=c(xx,xx),y=c(0,0,get_change(xx)), yend=get_change(c(xx,xx)))
+
+ss <- seq(0,6.5, by=.1)
+qplot(ss,get_change(ss), geom="line") + 
+  geom_segment(aes(x=x,xend=xend,y=y,yend=yend), data=ldat, linetype="dashed") +
+  ylab("Change in probability of correct response \n due to graduate degree") +
+  xlab("Lineup variability") +
+  scale_y_continuous(breaks =c(0,.003237246,.01,.02,.02376092,.03,.04), expand = c(0.02,0),
+                     labels =c(0,0.003,.01,.02,.024,.03,.04)) +
[…]

### 2013-06-16 sociology_paper.tex
up to learning trend

@@
 
+% This table contains the demographic information without factor wise percentages.
 %\begin{table}[hbtp]
@@
 
-\subsection{Demographic Factors} Proportions of correct responses and the logarithm of average time taken to evaluate  each lineup are computed and their distributions are shown as boxplots in Figure \ref{fig:demographic_effect} for different demographic factor levels.
+\subsection{Demographic Factors} Proportions of correct responses and natural logarithms of average time taken to evaluate  each lineup are computed  for different demographic factor levels.
+Their distributions are shown using boxplots in Figure \ref{fig:demographic_effect}.
 Averages of these distributions are represented by dots inside the boxplots.
@@
 People who have only a high school degree took less time and this is related to what we have seen for young age group.
-The time taken by male and female participants are almost the same.
+Averages of log time taken by male and female participants look similar but there are some differences as we see in Table~\ref{tbl:demographics}.
 
@@
    \includegraphics[width=4in]{demographic_effect.pdf} 
-   \caption{Boxplots of average time taken in log scale and proportion correct responses of all the lineups plotted for each demographic factor variables.
+   \caption{Boxplots of average log time taken and proportion correct responses of all the lineups plotted for each demographic factor levels.
 The dots inside the boxes represent means.
-Demographic factors appear to be statistically significant for explaining time taken.
-Effect of graduate degree and India in proportion correct is much different but is actually the India effect since, Most of the graduate degree holders are coming from India.}
+Some differences in means of various demographic factors are observed.
+Variability in proportion correct indicates large variability in lineup difficulties.}
    \label{fig:demographic_effect}
@@
 
-We observed some differences in median proportion of correct responses for different levels of the demographic factors as in Figure \ref{fig:demographic_effect}.
-But the variability of proportion correct responses are huge as per design of the experiments since we have some very easy and some extremely difficult lineups.
-Some of the lineups  were not expected to be evaluated correctly as some of the null plots in those lineups were showing more signal than the actual plot itself.
+We observed some differences in median proportion of correct responses for different demographic factor levels.
+But the variability of proportion correct responses are huge as per design of the experiments since there were some very easy as well as extremely difficult lineups.
+Some of the lineups  were not expected to be evaluated correctly as the null plots in those lineups were showing more signal than the actual data plot itself.
 In those scenarios,  one may not reject the null hypothesis.
-We call these difficult lineup.
+We call these difficult lineups.
 Some of the lineups were so easy that data plot is detectable at a first glance.
@@
 Gender does not show any significant difference in performance.
-Similarly, ANOVA results are obtained by fitting Model \eqref{eqn:demographic_time} to the data of log time taken and are shown in Table~\ref{tbl:anova_factor}.
+Similarly, ANOVA results are obtained by fitting Model \eqref{eqn:demographic_time} to the data of log time taken and are shown in Table~\ref{tbl:anova_factor} as well.
 All the factor variables are significantly different for log time taken including gender.
@@
 
-To further illustrate how the individual factor levels differ from each other  detailed results of Models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
+To further illustrate how the individual factor levels differ from each other  the detailed results of Models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
 For parameter estimation the first factor levels shown in the Table are  used as the base line.
-Demographic factors have significant effects on log time taken to evaluate a lineup on all their levels.
-For the probability of correct responses not all the factor levels are significant.
+Demographic factor levels have significant effects on log time taken to evaluate a lineup.
+But not all the levels are significant for the probability of correct responses.
 
@@
 For probability of correct response age group 36-40, rest of the world and graduate degree are significantly different.
-For gender we do not see any difference in performance.}
+For gender no difference in performance is observed.
+Lineup variability is estimated to be very large for Model \eqref{eqn:demographic_response}.}
 \scalebox{0.9}{
@@
 Because of this huge impact of lineup specific variance, the practical impact of other factor variables on the probability of correct response is in fact very small.
-We illustrate this with the following example of the significant effect of graduate degree.
+We illustrate this with the following example of graduate degree.
 

### 2013-06-16 R/sociology_paper.R

@@ -112,10 +112,10 @@ for (i in 1:10){
   di$academic_study[di$academic_study==0] <- NA
   di$experiment = paste("experiment_",i, sep="")
   di$id = paste("exp",i,"_",di$id, sep="")
-  degree <- factor(c("High school or less", "Some under grad courses",
-                  "Under graduate degree","Some graduate courses",
+  degree <- factor(c("High school or less", "Under grad courses",
+                  "Under grad degree","Graduate courses",
                   "Graduate degree")[di$academic_study])
-  di$degree <- factor(degree, levels=levels(degree)[order(c(5,1,4,2,3))], ordered=T)
+  di$degree <- factor(degree, levels=levels(degree)[order(c(4,5,1,2,3))], ordered=T)
   di$age_level <- factor(c("below 18","18-25","26-30","31-35","36-40",
                   "41-45","46-50","51-55","56-60","above 60")[di$age])
   di$gender_level <- factor(c("Male","Female")[di$gender])
@@ -218,7 +218,7 @@ qplot(variable_level, value, geom="boxplot",data=mdat[complete.cases(mdat),]) +
   stat_summary(fun.y=mean, geom="point") + xlab("Levels of demographic factors")+
   theme(axis.text.x=element_text(angle=90, hjust=1))
 
-ggsave("../images/demographic_effect.pdf", width=6, height=6)
+ggsave("../images/demographic_effect.pdf", width=6.5, height=6)
 
 #----------------------------------------------------------------
 # Function to obtain estimates of fitted model for latex xtable
@@ -246,14 +246,9 @@ get_estimates <- function(fit){
 # testing significance of demographic factor main effects
 # --------------------------------------------------------------
 ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name), 
-           data=subset(demographics, experiment=="experiment_6"))
-
-ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name), 
            data=demographics)
 summary(ft)
 
[…]

### 2013-06-17 sociology_paper.tex
upto location effect

@@
 These concepts have been developed further by \citet{majumder:2013} who refined the terminology and validated the lineup protocol with a head to head comparison with conventional inference.
-One of the major contributions of  \citet{majumder:2013}  is to define the power of the visual test and how to estimate the power for a particular lineup.
+One of the major contributions of  \citet{majumder:2013}  is to define the power of the visual test and provide methods to obtain the power for a particular lineup.
 It was observed that the power can be as good or better than that of a conventional test in some scenarios.
@@
 It is easier to spot the slope of the line compared to just the scatterplot itself.
-But it appears that people are better in noticing the unusual pattern in the data which is some contamination purposefully added for experiment 3.
+But people are better in noticing the unusual pattern in the data which is some contamination purposefully added for experiment 3.
 For experiment 3 the performance of the observers was much better compared to experiment 2 \citep{majumder:2013}.
@@
 Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $kth$ attempt by a single observer as 
-\begin{equation} \label{eqn:trend_power}
-\hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
+\begin{equation} \label{eqn:trend_time_est}
+\hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j.
 \end{equation}
@@
 \centering
-\caption{Anova of full model with all the demographic factors vs reduced model with removing respective factor variable. Gender does not seem to have any effect on probability of correct response.}
+\caption{Anova of full model with all the demographic factors vs reduced model with removing respective factor variable. Gender does not have any effect on probability of correct response.}
 \scalebox{.9}{
@@
 Age group 36-40 is significantly different than the base line age group of 18-25 year olds.
-The other age levels appears to be similar in explaining the probability of correct response.
+The other age levels are similar in explaining the probability of correct response.
 For country, the rest of the world is different from USA but India appears to be similar to USA.
@@
 They did not appear to be significantly different with a $p$-value of 0.856.
-But we consider the bigger Model \eqref{eqn:trend_response} with attempt as a factor variable with 10 levels.
+But we consider the bigger Model \eqref{eqn:trend_response} with attempt as a factor variable with 10 levels to examine the effect of each level.
 Attempt 1 is considered to be base level while fitting the model.
@@
 Table \ref{tbl:model_result_response} presents the parameter estimates and $p$-values of fixed effect estimates of Model \eqref{eqn:trend_response}.
-The larger $p$-values suggest that none of the levels of attempt ($\alpha_2$ through $\alpha_{10}$) appears to be significant at \%1 significance level.
-Moreover, some of the estimates are positive and some are negative and they show up seemingly in random order suggesting later attempts not necessarily improved.
+The larger $p$-values suggest that none of the levels of attempt ($\alpha_2$ through $\alpha_{10}$) are significant at \%1 significance level.
+Moreover, some of the estimates are positive and some are negative and they show up seemingly in random order suggesting later attempts not necessarily have improved.
 This indicates that there may be no learning effect on the probability of correct evaluations.
@@
 \caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for probability of correct lineup evaluation.
-None of the fixed factor effects of attempt ($\alpha_2$ through $\alpha_{10}$) seem to be significantly different from the first attempt $\alpha_1$
+None of the fixed factor effects of attempt ($\alpha_2$ through $\alpha_{10}$) are significantly different from the first attempt $\alpha_1$
 at \%1 level in all three experiments 5, 6 and 7. For experiment 7 subject specific variation is very small on the other hand lineup variance is much higher compared to the other two experiments.}
@@
   $\sigma^2_u$ & 0.720 & 0.848 &  &  &   & 0.815 & 0.903 &  &  &   & $<$0.001 & $<$0.001 &  &  \\ 
-  $\sigma^2_l$ & 2.178 & 1.476 &  &  &   & 2.009 & 1.418 &  &  &   & 10.980 & 3.314 &  &  \\ 
+  $\sigma^2_{\ell}$ & 2.178 & 1.476 &  &  &   & 2.009 & 1.418 &  &  &   & 10.980 & 3.314 &  &  \\ 
    \hline
@@
 
-\begin{figure}[htbp] 
-   \centering
-%   \includegraphics[width=6.5in]{learning_trend.pdf} 
-    \includegraphics[width=6.5in]{learning_trend_subject.pdf} 
-   \caption{ Mean residuals of proportion correct obtained from the model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
-Linear regression lines fitted through the points are shown.
-None of the slopes in these three experiments are statistically significant.
-Subject specific least square lines show large subject specific variability.
-Some of the observers did worse and some did better over the successive attempts.}
-   \label{fig:learning_trend_response}
-\end{figure}
-
 The variance of random slope for attempt ($\sigma^2_a$) is estimated to be very small compared to other random effects except for subject variability ($\sigma^2_u$) for experiment 7.
-This suggest that most of the variability for attempt is accounted for by the fixed effect factors of attempt.
-For experiment 7 the lineup variability ($\sigma^2_l$) was the largest as per experimental design.
+This suggests that most of the variability for attempt is accounted for by the fixed effect factor attempt.
+For all three experiments the lineup variabilities ($\sigma^2_l$) were estimated to be very large making the practical impact of other factor levels even more negligible.
 Some of the lineups were very easy and some were really difficult in experiment 7.
-The overall average probability for picking out the data plot from a lineup ($\mu$) appears to be significant for both experiments 5 and 7.
-But for experiment 6 it does not seem to be significant suggesting that experiment 6 lineups were difficult compared to the other experimental lineups.
+The overall average probability for picking out the data plot from a lineup ($\mu$) is significant for both experiments 5 and 7.
+But for experiment 6 it is not significant suggesting that experiment 6 lineups were difficult compared to the other experimental lineups.
 This shows a feature of the experimental designs where a mixture of difficult and easy lineups were included within the experiments as well as between the experiments.
@@
 
-To visualize further how the performance in correct response improves over successive attempt, we fitted Model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
-The averages of these residuals for each of the attempts are plotted against attempt as shown in Figure \ref{fig:learning_trend_response}.
-Least square linear regression lines are fitted through the points.
-Even though positive slopes over the attempts are observed, none of the trends is statistically significant.
-Subject specific least square lines show large overall variability.
+To visualize how the performance in correct response improves over successive attempts, we fitted Model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
+Least square regression lines were fitted through the subject specific residuals as shown in Figure~\ref{fig:learning_trend_response}.
+Two important features were observed; one is subject specific variability and the other is random slope with attempts which indicates subjects specific learning trend.
 Some subjects show improvement over time and some show the decrease in performance.
-The subject specific slopes are very different as well.
-
-Model \eqref{eqn:trend_time} was compared to couple of other alternatives.
-A bigger model with attempt as a factor was considered but they did not appear to be significantly different with $p$ value 0.236.
-A reduced model with attempt as a continuos covariate was also considered but they were significantly different with $p$-value $<0.001$.
+
+\begin{figure}[htbp] 
+   \centering
+%   \includegraphics[width=6.5in]{learning_trend.pdf} 
+    \includegraphics[width=6.5in]{learning_trend_subject.pdf} 
+   \caption{ Least square lines fitted through the subject specific residual proportion correct obtained from Model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
+Subject specific positive and negative slopes are observed.
+Mean residuals are shown as dots and least square regression lines fitted through the points show no overall learning trend in each of the three experiments.}
+   \label{fig:learning_trend_response}
+\end{figure}
+
+The averages of these residuals for each of the attempts are shown as dots in Figure~\ref{fig:learning_trend_response}.
+Least square linear regression lines are fitted through the points for each of the experiments.
+Positive slopes over the attempts are observed but none of them is statistically significant.
+
 Table \ref{tbl:model_result_time} presents the results of Model \eqref{eqn:trend_time}.
 The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments.
-The negative estimates suggest that later attempt took less time for an evaluation.
-It is interesting to note that even though observers did not improve the performance over attempt, they became efficient in responding faster.
+The negative estimates suggest that on an average later attempt took less time for an evaluation.
+Even though observers did not improve the performance over attempt, they became efficient in responding faster.
 The parameter $\alpha_1$ for first attempt is also highly significant.
@@
   $\sigma^2_u$ & 0.259 & 0.509 &  &  &   & 0.245 & 0.495 &  &  &   & 0.134 & 0.366 &  &  \\ 
-  $\sigma^2_l$ & 0.008 & 0.091 &  &  &   & 0.040 & 0.199 &  &  &   & 0.055 & 0.235 &  &  \\ 
+  $\sigma^2_{\ell}$ & 0.008 & 0.091 &  &  &   & 0.040 & 0.199 &  &  &   & 0.055 & 0.235 &  &  \\ 
   $\sigma^2$ & 0.211 & 0.460 &  &  &   & 0.251 & 0.501 &  &  &   & 0.206 & 0.454 &  &  \\ 
@@
 
-The lineup variance of experiment 7 is estimated as $\hat{\sigma}^2 = 10.98$ from Model \eqref{eqn:trend_response}.
+The lineup variance of experiment 7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from Model \eqref{eqn:trend_response}.
 But from Model \eqref{eqn:trend_time} it is estimated to be much smaller in all these experiments.
-This suggest that the harder lineups not necessarily took more time than easier lineups.
+This suggests that the harder lineups not necessarily took more time than easier lineups.
 The observers spent enough times to evaluate a easy lineup and for difficult lineup they might just give up at some point and provided the feedback.
 
+Model \eqref{eqn:trend_time} was compared to couple of other alternatives.
+A bigger model with attempt as a factor was considered but it was not significantly different with $p$ value 0.236.
+A reduced model without the first attempt was also considered but it was significantly different with $p$-value $<0.001$.
+
 To visualize how the time taken reduces over the successive attempts, we fitted Model \eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
-The averages of these residuals for each attempt are plotted in Figure \ref{fig:learning_trend_time} for all three experiments.
-Least square regression lines are fitted to the points excluding the first attempt.
+Least square regression lines are fitted through the subject specific residuals.
+Subject specific slopes are much different in each of the three experiments as we see in Figure \ref{fig:learning_trend_time}.
+Some subjects improved over attempts by taking less time in the later attempts while others got worse.
+The averages of these residuals for each attempt are plotted as dots.
+Least square regression lines are fitted to these points excluding the first attempt since for first attempt we fitted an indicator covariate.
 The downward trends are evident in the plots.
 All the slopes are highly significant.
-For attempt 1, we observed large positive residuals for each of the experiments.
-These results indicate that after working with the multiple tasks each observer became efficient which helped them finish the task faster in later attempts.
+As expected we observed large positive residuals for each of the experiments for first attempt.
+
+%These results indicate that after working with the multiple tasks each observer became efficient which helped them finish the task faster in later attempts.
 Subject specific least square lines show variations in slopes and overall time taken.
@@
  \includegraphics[width=6.5in]{learning_trend_time_subject.pdf}    
-   \caption{ Mean residuals of log(time taken) obtained from model \eqref{eqn:trend_time} fitted without covariate attempt plotted against attempt.
-Linear regression lines are fitted through the points excluding the first attempt.
-For all the three experiments the downward slopes are statistically significant which indicates that MTurk workers take less time as they progress through their attempts.
-But subject specific variations in slopes are also observed where some of the subjects did worse over successive attempts while others did better.}
+   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting Model \eqref{eqn:trend_time} without covariate attempt.
+Differences in subject specific slopes are observed.
+Some of the subjects did worse over successive attempts while others did better.
+Averages of these residuals are plotted as dots and least square regression lines are fitted to obtain overall trends.
+For all the three experiments the overall downward slopes are statistically significant which indicates that MTurk workers take less time as they progress through their attempts.}
    \label{fig:learning_trend_time}
@@
 
-Observer demographics seem to have significant influence on performance in time taken to evaluate a lineup.
+Observer demographics have significant influence on performance in time taken to evaluate a lineup.
 Some variations among the factors are observed in terms of probability of correct evaluation.
-Age group 36-40, Countries other than India and United states, People who have a graduate degree appear to be significantly different.
+Age group 36-40, Countries other than India and United states, People who have a graduate degree are significantly different.
 But their practical impact on probability to correctly evaluate a lineup is very small and in some cases very negligible as the lineup difficulties increase or decreases.
@@
 If the size of the lineup is much higher than 20, it is intuitive that there may be location effect of the data plot in the lineup.
-It is because, the observer may get tiered of scanning and make decision based on the partial scanning of the lineup.
+It is because, the observer may get tired of scanning and make decision based on the partial scanning of the lineup.
 On the other hand fewer than 20 plots will allow observer to compare the actual plot with fewer null plots giving more chance of picking actual plot as an error not as an actual plot.
@@
 %\end{figure}
-
+\clearpage
 \section{Some Plots of Exploratory Data Analysis}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=5.8in]{turker_location_experiment.pdf} 
+   \caption{World maps showing where the participants are coming from for all the 10 experiments.}
+   \label{fig:turker_location_experiment}
+\end{figure}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=6.5in]{participation_time.pdf} 
+   \caption{Number of participants by time of the day feedbacks received (central time).
+Experiment 1 shows MTurk workers participated the experiments around the clock.
+Other experiments did not take a whole day to finish.
+For experiment 3 most of the participants are from India because of timing.
+No matter when the experiment is started, subjects from India shows participations.
+For United States, subjects participated if the experiment is not in the mid night, except for experiment 6.}
+   \label{fig:participation_time}
+\end{figure}
 
@@
    \label{fig:gender_country_response}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=6in]{turker_location_experiment.pdf} 
-   \caption{World maps showing where the participants are coming from for all the 10 experiments.}
-   \label{fig:turker_location_experiment}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=6.5in]{participation_time.pdf} 
-   \caption{Time of the day when the participants work (central time).
-Experiment 1 shows MTurk workers participated the experiments around the clock.
-Other experiments did not take a whole day to finish.
-For experiment 3 most of the participants are from India because of timing.
-No matter when the experiment is started, subjects from India shows participations.
-For United States, subjects participated if the experiment is not in the mid night, except for experiment 6.}
-   \label{fig:participation_time}
 \end{figure}

### 2013-06-17 R/sociology_paper.R

@@ -303,6 +303,13 @@ anova.prop <- rbind(data.frame(round(pag[1,1:4],2),round(pag[2,5:7]/1,2)),
 anova.reseult <- rbind(anova.time, anova.prop)
 xtable(anova.reseult)
 
+# Model with demographic factor interaction
+# India:Undergrad course is hghly significant
+fit <- lmer(response~age_level+country+degree+gender_level+country:degree+(1|pic_name), 
+               family="binomial", demographics)
+
+
+fit
 # ==============================================================
 # Exploring the practical significance of demographics
 # ---------------------------------------------------------------

### 2013-06-17 sociology_paper.tex
first revision completed

@@
 
-\subsubsection{Design for Location Effect} To assess the location effect of a data plot in a lineup, we set up an experiment based on the gene expression data \citep{atwood:2013} with two groups.
+\subsubsection{Design for Location Effect} \label{sec:location_design} Experiment 9 shown in Table~\ref{tbl:visual_stat} is designed to to assess the location effect of a data plot in a lineup.
+It is set up based on the gene expression data \citep{atwood:2013} with two groups.
 One is the Genotype and the other the Empty Vector (EV).
@@
 For interaction effect the same plot type is used except that the means of iron sufficiency or insufficiency were connected by lines colored by the groups.
+One of the interaction test statistics is shown in Table~\ref{tbl:visual_stat} for experiment 9.
 For no interaction, the visual test statistic shows two parallel lines representing two groups and for significant interaction the lines are not parallel.
@@
 
-The proportion of correct responses for each data plot location is shown in Figure \ref{fig:location_effect}.
-We observe some variability of performance for different sets of null plots even though same data plot was used for all these null sets.
-This may happen when for some set of null plots, one null plot appears to be more similar to the actual plot.
+A total of 111 subjects were recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
+Each subject evaluated two lineups;  one for Interaction effect and the other for Genotype.
+In total there were 222 feedbacks or responses on 50 lineups.
+The data on the test lineup were excluded from the analysis.
+
+The proportion of correct responses for each data plot location is shown in Figure \ref{fig:location_effect} colored by null sets.
+We observe some variability of performance for different null sets even though same data plot was used for all these null sets.
+This may happen when for some set of null plots, one null plot appears to be very similar to the actual plot.
 In another set of null plots this may not happen making some lineups easier than others even though the actual data plot is the same.
 This pattern is evident in the figure as we see proportion correct for null plot 5 is consistently above the null plot 1 for each of the locations.
-The overall average proportion of correct responses for each location are shown using dashed lines.
-As we see in the plot  the overall proportion is not varying much for different locations.
-
-The sizes of the dots in  Figure \ref{fig:location_effect} represents the number of responses obtained for that location and null set.
-For some locations we have as many as 10 responses.
-For location 1, we did not have any response for null set 1 in one of the interaction lineups.
-We observed larger variability for interaction lineups as compared to Genotype lineups.
-It is because some of the interaction lineups were evaluated only once.
-For example, we have only one response for null plot 3 and 5 for Interaction effect which should lead to the extreme proportion of correct since one response would either be correct or wrong.
-The larger variability observed for Interaction lineups can be controlled at some extent by evaluating them multiple times.
+A test for differences in average performances of null sets shows significant difference only for null set 3 of interaction lineup.
+The rest of the null sets don't show any differences at 1\% significance level.
 
@@
 
+The overall average proportion of correct responses for each location are shown using dashed lines in Figure \ref{fig:location_effect}.
+There is variability in performances for each null set depending on the locations.
+But the overall proportion is not varying much for different locations.
+
+The sizes of the dots in  Figure \ref{fig:location_effect} represents the number of responses obtained for that location and null set.
+For some locations we have as many as 10 responses.
+For location 1, we did not have any response for null set 1 in one of the interaction lineups.
+We observe larger variability for interaction lineups as compared to Genotype lineups.
+It is because some of the interaction lineups were evaluated only once as we see for null plot 3 and 5 for Interaction effect.
+This leads to the extreme proportion of correct since one response will either be correct or wrong.
+The larger variability observed for Interaction lineups can be controlled at some extent by evaluating them multiple times.
+
 We fitted Model \eqref{manova} to test if the mean performance vectors are similar for different locations.
@@
 This suggest that there may be no difference in location.
-
-We also examined whether the proportion of correct responses differs if the actual plot is on the outer boundary or in the inner locations.
-The locations 7,8,9,12,13,14 in the lineup are considered inner locations and the rest of the boundary locations are considered to be outer location.
-As we see in Figure \ref{fig:location_effect}, location 9, 12 are inside for Interaction effect and location 8, 12 are inside for Genotype.
-It does not appear to have any difference whether the actual plot is inside or outer border of the lineup.
-We also fitted Model \eqref{manova} with two locations, inner and outer, as covariate and observed no significant differences.
 
@@
  \cline{5-7}
- Effect & DF & Pillai & Approx. F & Numerator& Denumerator &Residual & p value\\
+ Effect & DF & Pillai & Approx. F & Numerator& Denumerator &Residual & $p$-value\\
   \hline
@@
 \end{table}
+
+We also examined whether the proportion of correct responses differs if the actual plot is on the outer boundary or in the inner locations.
+The locations 7,8,9,12,13,14 in the lineup are considered inner locations and the rest of the boundary locations are considered to be outer location.
+As we see in Figure \ref{fig:location_effect}, location 9, 12 are inside for Interaction effect and location 8, 12 are inside for Genotype.
+It does not show any differences whether the actual plot is inside or outer border of the lineup.
+We also fitted Model \eqref{manova} with two locations, inner and outer, as covariate and observed no significant differences.
 
@@
 
-Observer demographics have significant influence on performance in time taken to evaluate a lineup.
+It is found that observer demographics have significant influence on performance in time taken to evaluate a lineup.
 Some variations among the factors are observed in terms of probability of correct evaluation.
@@
 
-The simulation experiment reveals that the effect of location of actual data plot in the lineup is not significant.
+The simulation experiment reveals that there is no significant effect of location of actual data plot in the lineup.
 This is important as the visual statistical inference procedure suggests that the data plot be placed at random anywhere in the lineup.
-Even though there are variations on the performance depending on actual plot locations, it is not statistically significant to be of any concern.
 This paper suggests that any random place in a lineup is as good as other places in the lineup.
+Even though there are variations on the performance depending on different null sets, their impact on probability to correctly evaluate a lineup is very negligible.
 
 The subjects participating this study may not necessarily know about statistical graphics.
-The numerous pilot studies done with the more trained participants suggest that power of visual statical inference may be higher with knowledgeable observer.
+The numerous pilot studies done with more trained participants suggest that power of visual statical inference may be higher for observers knowledgeable about statistical graphics.
 The fact that a person with graduate degree performs better may be associated with this.
 But a graduate degree does not necessarily mean to have training on statistical graphics.
-More experiments may be needed to make a concrete decision about the difference in power with skilled observer and a non-skilled observer in terms of knowledge about statistical graphics.
+More experiments may be needed to learn about the differences in power with trained and a non-trained observer in terms of knowledge about statistical graphics.
 
@@
 The learning trend in terms of time taken shows downward trends.
-At some points it should level of.
-It would be useful to know for how many lineups time taken levels off.
+At some point it should level of.
+It will be useful to know for how many lineups time taken levels off to decide how many lineups is appropriate to show.
 
@@
 If the size of the lineup is much higher than 20, it is intuitive that there may be location effect of the data plot in the lineup.
-It is because, the observer may get tired of scanning and make decision based on the partial scanning of the lineup.
+It is because, the observer may get tired of scanning and may make decision based on the partial scanning of the lineup.
 On the other hand fewer than 20 plots will allow observer to compare the actual plot with fewer null plots giving more chance of picking actual plot as an error not as an actual plot.
@@
 
-\subsection{How People Pick the Data Plot} In each of the experiments we had a question for the turk observers to choose the reason for their selection of a particular data plot.
-It is shown by \cite{majumder:2013} that these choice reasons reveal the way people picks the data plot in the lineup.
-To investigate this further we added free text input option for their choice reason instead of some fixed reasons to select from.
-In this experiment people could write whatever they think their reasons for choice are.
+\subsection{How People Pick the Data Plot} In each of the experiments observers were asked to choose the reason for their selection of a particular data plot.
+\cite{majumder:2013} showed that these choice reasons may provide the clue on how people picks the data plot in the lineup.
+To investigate this further we added free text input option in experiment 9 for their choice reason instead of some fixed reasons to select from.
+This allowed observers to write whatever they think their reasons for choice are.
 Figure \ref{fig:wordle} shows the words used to explain their reasons for choice.
-The most common words used to explain their choice are points and green which indicates two important feature of a plot.
-One is the indicator of plot aesthetics and the other is the color.
+The most common words used to explain their choice are points and green which indicates the use of two important features of the grammar of graphics \citep{wilkinson:1999, hadley:2009}.
+One is the indicator of geometric shape and the other is the aesthetics.
 Spread, steepest, line and apart are some other important words used frequently.
@@
 Figure \ref {fig:wordle} also shows some insight about peoples way of reading a plot.
-We notice that variability in the data, color and aesthetics used to generate plots and existence of any systematic pattern in the plot are some of the important features revealed from the figure.
+We notice that variability in the data, geometric shapes and aesthetics used to generate plots and existence of any systematic pattern in the plot are some of the important features revealed from the figure.
 These features are commonly used by human brain to examine and compare plots.
 Notice that these features may be specific to this particular experiment.
-There could be different other features people would use to evaluate a lineup depending on the situation.
+There can be different other features people may use to evaluate a lineup depending on the situation.
 But with these words we get a general idea on how people may think while evaluating a lineup.
 
-Turk workers are not necessarily trained on statistics or aware of specific terms used in statistics or statistical plots.
-Thus it is also interesting to note that how they explain things that have specific definitions and meaning.
-Notice in Figure \ref {fig:wordle} that some keywords like spread, apart should be analogous to larger variability while together,  close may be for indicating smaller variability.
+Turk workers are not necessarily trained on statistics or aware of specific terms used in statistics or statistical plots let alone having knowledge about grammar of graphics.
+It is interesting that people explain things that have specific definitions and meaning in the literature.
+For example, some keywords like spread, apart should be analogous to larger variability while together,  close may be for indicating smaller variability.
+Thus the perception from a statistical graphics is intuitive and human intelligence learn this even without having specific training.
+This is why visual inference can be used as a tool for teaching inference in the basic statistics classes.
 

### 2013-06-17 R/sociology_paper.R

@@ -817,11 +817,24 @@ summary(manova(cbind(null_1,null_2,null_3,null_4,null_5)~factor(locs), data=int.
 # Modeling location effect
 # Fiting generalized mixed model for location effect
 # Locations are not significant
+# Model fi1 shows the effect of null plots variability is negligible 
+# as location effects are much bigger compared to null variability
 # -----------------------------------------------------------
 
 colnames(df)
-fi <- lmer(response~factor(plot_location)+nulls+(1|pic_name), family="binomial",data=subset(df,plot_type=="Interaction"))
+fi0 <- lmer(response~factor(plot_location)+nulls+(1|pic_name), family="binomial",data=subset(df,plot_type=="Interaction"))
+fi1 <- lmer(response~factor(plot_location)+(1|nulls), family="binomial",data=subset(df,plot_type=="Interaction"))
+fi2 <- glm(response~factor(plot_location), family="binomial",data=subset(df,plot_type=="Interaction"))
+
+# this model shows null set 3 for interaction is significant
+fi3 <- lmer(response~factor(nulls)+(1|plot_location), family="binomial",data=subset(df,plot_type=="Interaction"))
+
 fg <- lmer(response~factor(plot_location)+nulls+(1|pic_name), family="binomial",data=subset(df,plot_type=="Genotype"))
+
+# the following plot shows null set 3 is different in interaction lineup
+dd <- ddply(df, .(plot_type,nulls), summarize, prop_correct=mean(response))
+qplot(nulls, prop_correct, geom="bar", stat="identity", facets=~plot_type, data=dd)
+
 
 # test for equivalency
 library(equivalence)

### 2013-06-20 sociology_paper.tex
impact of graduate degree plot and discussion

@@
 
-{\bf Keywords: \sf statistical graphics, non-parametric test, cognitive phycology, data visualization, exploratory data analysis, data mining, visual analytics} 
+{\bf Keywords: \sf statistical graphics, non-parametric test, cognitive phycology, data visualization, exploratory data analysis, data mining, visual analytics.} 
 \end {abstract}
@@
    \includegraphics[width=0.99\textwidth]{electoral-5-13.pdf} 
+%      \includegraphics[width=0.99\textwidth]{electoral-1-1.pdf} 
    \caption{ \label{fig:elect-1} Which one of the plots is the most different from the others?}
@@
 \hline
-Age&18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
- & 26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
+Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
+ &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
   &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0376 \\ 
@@
   \hline
-Country& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
+Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
   &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
@@
   \hline
-Education &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
+Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
  &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
@@
   \hline
-Gender & Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
-  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.5 & 0.1331 \\ 
+Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
+  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.1331 \\ 
   \hline
 \multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
-& lineup & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
- & Error & 0.479 & 0.692 &  &  & & && &\\ 
+& lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
+ & Error($\sigma$) & 0.479 & 0.692 &  &  & & && &\\ 
    \hline
@@
 The positive parameter estimate indicates that they perform better and the probability of correct response is higher than other education levels.
-As expected, there is no significant difference between male and female performances.
+There is no significant difference between male and female performances.
 
@@
 For two standard deviations it further reduces to 0.003 which is practically very small.
-Figure~\ref{fig:practical_impact_demographics} illustrates this example showing how the change in probability diminishes for larger variability of lineup difficulty.
+Figure~\ref{fig:practical_impact_graduate} illustrates this example showing the proportions of correct responses due to graduate degree as compared to high school degree.
+The highest practical difference in proportions of correct response is estimated as 0.045 which is very negligible.
+The difference further diminishes if we move one or two standard deviation of lineup variability.
 A similar calculation shows the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of correct response.
-Thus even though some of the factor levels are significant, their practical impact is negligible because of high variability of lineup difficulties.
+Thus even though some of the factor levels are statistically significant, their practical impact is negligible.
 
@@
 %   \includegraphics[width=6.5in]{learning_trend.pdf} 
-    \includegraphics[width=3in]{practical_impact_demographics.pdf} 
-   \caption{ Change in probability of correct response due to graduate degree as compared to high school degree diminishes rapidly as the variability in lineup difficulty $\sigma_{\ell}$ increases.}
-   \label{fig:practical_impact_demographics}
+    \includegraphics[width=4in]{practical_impact_graduate.pdf} 
+   \caption{Proportion of correct responses due to graduate degree as compared to high school degree for an 18-25 year old female in the United States.
+Even though graduate degree is statistically significant, the largest practical difference in proportion correct is  0.045 which is very negligible.
+The difference even diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) of lineup variability.}
+   \label{fig:practical_impact_graduate}
 \end{figure}
+
+%\begin{figure}[htbp] 
+%   \centering
+%%   \includegraphics[width=6.5in]{learning_trend.pdf} 
+%    \includegraphics[width=3in]{practical_impact_demographics.pdf} 
+%   \caption{ Change in probability of correct response due to graduate degree as compared to high school degree diminishes rapidly as the variability in lineup difficulty $\sigma_{\ell}$ increases.}
+%   \label{fig:practical_impact_demographics}
+%\end{figure}
 
@@
 
-It is found that observer demographics have significant influence on performance in time taken to evaluate a lineup.
+Human demographics have significant influence on performance in time taken to evaluate a lineup.
 Some variations among the factors are observed in terms of probability of correct evaluation.
 Age group 36-40, Countries other than India and United states, People who have a graduate degree are significantly different.
-But their practical impact on probability to correctly evaluate a lineup is very small and in some cases very negligible as the lineup difficulties increase or decreases.
-Gender does not have any effect on performance.
+But their practical impact on probability to correctly evaluate a lineup is very negligible and in some cases it diminishes  as the lineup difficulties increase or decrease.
+Gender does not have any significant effect on performance.
+Thus there may be differences in time taken to evaluate a lineup for different human demographics but the practical impact of demographics on the performance is very negligible.
+This result is very important for the power of visual test as it demonstrates the robustness of the test for different human factors.
 
@@
 But the overall performance of the observers does not increase through successive attempts while evaluating multiple lineups.
-This result is very important as it suggests that the power estimated for visual inference using human subject experiment is robust and may not change if those participants are allowed to give feedback again.
+This result suggests that the power estimated for visual inference using human subject experiment is robust and may not change if those participants are allowed to give feedback again.
 The skill in evaluating the lineup in shorter time gets improved over successive evaluations.
-The earlier evaluations take significantly longer than the later evaluations.
-This investigation suggest that the skilled person may only do it faster.
+The earlier evaluations take significantly longer time than the later evaluations.
+It suggests that the skilled person may only do it faster.
 
@@
 The subjects participating this study may not necessarily know about statistical graphics.
-The numerous pilot studies done with more trained participants suggest that power of visual statical inference may be higher for observers knowledgeable about statistical graphics.
+The numerous pilot studies done with more trained participants suggest that power of visual statical inference may be higher for observers who have advanced knowledge about statistical graphics.
 The fact that a person with graduate degree performs better may be associated with this.
@@
 %\end{figure}
-\clearpage
+
+%\clearpage
+\section{Electoral Building Lineups and Results}
+%Figures~\ref{fig:elect-1} and \ref{fig:elect.2} through~\ref{fig:elect.5}
+
+Five lineups were shown to (different) Amazon Turk workers in an experiment.
+They all were created as described in the introduction of the paper.
+In order to not bias observers, no context information was given about how these plots were created or what data they were displaying.
+This also made it necessary to slightly more stylize the display.
+
+% latex table generated in R 3.0.1 by xtable 1.7-1 package
+% Wed Jun  5 17:39:09 2013
+\begin{table}[hbtp]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ lineup & \#1 & \#2 & \#3 & \#4 & \#5 \\ 
+  \hline
+  \# correct/ \#evaluation &  12/72 & 11/66 & 5/74 & 14/72 & 19/57 \\ 
+$p$-value & 0.00023 & 0.00041 & 0.31 & 1.2e-05 &  1.9e-11 \\ 
+data panel & $3\cdot4+1$ & $2^4+1$ & $4^2+2 $ & $12+\sqrt{25}$ & $2^3-7$ \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+% latex table generated in R 3.0.1 by xtable 1.7-1 package
+% Wed Jun  5 18:00:55 2013
+\begin{table}[hbtp]
+\centering
+\caption{\label{tbl:response}Overview of all choices by observers for each of the lineups.
+The correct choice is bolded.
+In most lineups there are null plots that were picked more often by observers, but the actual result is among the plots being picked most often, indicating that there is some indication that the election result is not completely consistent with the polls.}
+\begin{tabular}{rrrrrrrrrrrrrrrrrrrrr}
+  \hline
+& \multicolumn{10}{l}{panel chosen}\\
+Lineup & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\ 
+  \hline
+\#  1 & 2 & 2 & 0 & 10 & 2 & 2 & 6 &  23 & 1 & 1 & 0 & 1 & \bf 12 & 3 & 3 & 1 & 0 & 1 & 1 & 1 \\ 
+\#  2 & 0 & 16 & 1 & 1 & 5 & 1 & 0 & 8 & 0 & 2 & 0 & 0 & 0 & 4 & 2 & 1 & \bf 11 & 1 & 0 & 13 \\ 
+\#  3 & 7 & 26 & 0 & 2 & 0 & 5 & 3 & 0 & 2 & 1 & 0 & 4 & 0 & 0 & 2 & 0 & 9 & \bf 5 & 0 & 6 \\ 
+\#  4 & 0 & 0 & 0 & 2 & 0 & 0 & 0 & 3 & 1 & 10 & 2 & 18 & 1 & 0 & 4 & 2 &\bf 14 & 0 & 13 & 0 \\ 
+\# 5 &\bf 19 & 1 & 4 & 1 & 0 & 1 & 0 & 12 & 0 & 0 & 0 & 4 & 1 & 0 & 0 & 12 & 1 & 1 & 0 & 0 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+%\clearpage
 \section{Some Plots of Exploratory Data Analysis}
@@
 \end{figure}
-
-\clearpage
-\section{Electoral Building Lineups and Results}
-%Figures~\ref{fig:elect-1} and \ref{fig:elect.2} through~\ref{fig:elect.5}
-
-Five lineups were shown to (different) Amazon Turk workers in an experiment.
-They all were created as described in the introduction of the paper.
-In order to not bias observers, no context information was given about how these plots were created or what data they were displaying.
-This also made it necessary to slightly more stylize the display.
-
-% latex table generated in R 3.0.1 by xtable 1.7-1 package
-% Wed Jun  5 17:39:09 2013
-\begin{table}[hbtp]
-\centering
-\begin{tabular}{rrrrrr}
-  \hline
- lineup & \#1 & \#2 & \#3 & \#4 & \#5 \\ 
-  \hline
-  \# correct/ \#evaluation &  12/72 & 11/66 & 5/74 & 14/72 & 19/57 \\ 
-$p$-value & 0.00023 & 0.00041 & 0.31 & 1.2e-05 &  1.9e-11 \\ 
-data panel & $3\cdot4+1$ & $2^4+1$ & $4^2+2 $ & $12+\sqrt{25}$ & $2^3-7$ \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-% latex table generated in R 3.0.1 by xtable 1.7-1 package
-% Wed Jun  5 18:00:55 2013
-\begin{table}[hbtp]
-\centering
-\caption{\label{tbl:response}Overview of all choices by observers for each of the lineups.
-The correct choice is bolded.
-In most lineups there are null plots that were picked more often by observers, but the actual result is among the plots being picked most often, indicating that there is some indication that the election result is not completely consistent with the polls.}
-\begin{tabular}{rrrrrrrrrrrrrrrrrrrrr}
-  \hline
-& \multicolumn{10}{l}{panel chosen}\\
-Lineup & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\ 
-  \hline
-\#  1 & 2 & 2 & 0 & 10 & 2 & 2 & 6 &  23 & 1 & 1 & 0 & 1 & \bf 12 & 3 & 3 & 1 & 0 & 1 & 1 & 1 \\ 
-\#  2 & 0 & 16 & 1 & 1 & 5 & 1 & 0 & 8 & 0 & 2 & 0 & 0 & 0 & 4 & 2 & 1 & \bf 11 & 1 & 0 & 13 \\ 
-\#  3 & 7 & 26 & 0 & 2 & 0 & 5 & 3 & 0 & 2 & 1 & 0 & 4 & 0 & 0 & 2 & 0 & 9 & \bf 5 & 0 & 6 \\ 
-\#  4 & 0 & 0 & 0 & 2 & 0 & 0 & 0 & 3 & 1 & 10 & 2 & 18 & 1 & 0 & 4 & 2 &\bf 14 & 0 & 13 & 0 \\ 
-\# 5 &\bf 19 & 1 & 4 & 1 & 0 & 1 & 0 & 12 & 0 & 0 & 0 & 4 & 1 & 0 & 0 & 12 & 1 & 1 & 0 & 0 \\ 
-   \hline
-\end{tabular}
-\end{table}
 

### 2013-06-20 R/sociology_paper.R

@@ -362,6 +362,26 @@ qplot(ss,get_change(ss), geom="line") +
 ggsave("../images/practical_impact_demographics.pdf", width=5.5, height=5)
 
 
+difficulty <- seq(-6,6, by=.1)
+hs_xb <- -0.683 + difficulty
+gd_xb <- hs_xb + 0.182
+ddat <- data.frame(difficulty=difficulty, prop_hs = px(hs_xb),prop_gd=px(gd_xb))
+
+# Maximum difference in prop correct is 0.045
+qplot(difficulty, prop_gd-prop_hs, data=ddat, geom="line")
+with(ddat, max(prop_gd-prop_hs))
+
+mddat <- melt(ddat,id=c("difficulty"))
+qplot(difficulty, value, data=mddat, color=variable, geom="line")+
+  xlab("Lineup difficulty") + ylab("Proportion of correct response") +
+  scale_x_continuous(breaks=c(-6,-2*2.293,-2.293,0,2.293,2*2.293,6), expand = c(0.02,0),
+                     labels=c("Difficult",expression(-2*sigma[l],-sigma[l],0,sigma[l], 
+                                                     2*sigma[l],"Easy"))) +
+  scale_color_discrete(name="Education",
+                      labels=c("High school", "Grad. degree"))
+
+ggsave("../images/practical_impact_graduate.pdf", width=6.5, height=4.5)
+
 
 # Time of the day when Mturk worker works
 turker$hours <- as.factor(hour(as.POSIXlt(turker$start_time)))

### 2013-06-20 sociology_paper.tex
some rearrangements of tables and figures

@@
 Lineup variability is estimated to be very large for Model \eqref{eqn:demographic_response}.}
-\scalebox{0.9}{
+\scalebox{0.86}{
 \begin{tabular}{rlrrrrrrrrrr}
@@
 %   \includegraphics[width=6.5in]{learning_trend.pdf} 
-    \includegraphics[width=4in]{practical_impact_graduate.pdf} 
+    \includegraphics[width=3.5in]{practical_impact_graduate.pdf} 
    \caption{Proportion of correct responses due to graduate degree as compared to high school degree for an 18-25 year old female in the United States.
-Even though graduate degree is statistically significant, the largest practical difference in proportion correct is  0.045 which is very negligible.
-The difference even diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) of lineup variability.}
+Even though graduate degree is statistically significant, the largest difference in proportion correct is  0.045 which is very negligible.
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) of lineup variability.}
    \label{fig:practical_impact_graduate}
@@
 at \%1 level in all three experiments 5, 6 and 7. For experiment 7 subject specific variation is very small on the other hand lineup variance is much higher compared to the other two experiments.}
-\scalebox{0.9}{
+\scalebox{0.81}{
 \begin{tabular}{rrrrccrrrccrrrc}
@@
 %   \includegraphics[width=6.5in]{learning_trend.pdf} 
-    \includegraphics[width=6.5in]{learning_trend_subject.pdf} 
+    \includegraphics[width=6.3in]{learning_trend_subject.pdf} 
    \caption{ Least square lines fitted through the subject specific residual proportion correct obtained from Model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
@@
 Both fixed effect parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.}
-\scalebox{0.9}{
+\scalebox{0.81}{
 \begin{tabular}{rrrrccrrrccrrrc}
@@
 %   \includegraphics[width=6.5in]{learning_trend_time.pdf}    
- \includegraphics[width=6.5in]{learning_trend_time_subject.pdf}    
+ \includegraphics[width=6.3in]{learning_trend_time_subject.pdf}    
    \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting Model \eqref{eqn:trend_time} without covariate attempt.
@@
 %\clearpage
-\section{Electoral Building Lineups and Results}
-%Figures~\ref{fig:elect-1} and \ref{fig:elect.2} through~\ref{fig:elect.5}
-
-Five lineups were shown to (different) Amazon Turk workers in an experiment.
-They all were created as described in the introduction of the paper.
-In order to not bias observers, no context information was given about how these plots were created or what data they were displaying.
-This also made it necessary to slightly more stylize the display.
-
-% latex table generated in R 3.0.1 by xtable 1.7-1 package
-% Wed Jun  5 17:39:09 2013
-\begin{table}[hbtp]
-\centering
-\begin{tabular}{rrrrrr}
-  \hline
- lineup & \#1 & \#2 & \#3 & \#4 & \#5 \\ 
-  \hline
-  \# correct/ \#evaluation &  12/72 & 11/66 & 5/74 & 14/72 & 19/57 \\ 
-$p$-value & 0.00023 & 0.00041 & 0.31 & 1.2e-05 &  1.9e-11 \\ 
-data panel & $3\cdot4+1$ & $2^4+1$ & $4^2+2 $ & $12+\sqrt{25}$ & $2^3-7$ \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-% latex table generated in R 3.0.1 by xtable 1.7-1 package
-% Wed Jun  5 18:00:55 2013
-\begin{table}[hbtp]
-\centering
-\caption{\label{tbl:response}Overview of all choices by observers for each of the lineups.
-The correct choice is bolded.
-In most lineups there are null plots that were picked more often by observers, but the actual result is among the plots being picked most often, indicating that there is some indication that the election result is not completely consistent with the polls.}
-\begin{tabular}{rrrrrrrrrrrrrrrrrrrrr}
-  \hline
-& \multicolumn{10}{l}{panel chosen}\\
-Lineup & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\ 
-  \hline
-\#  1 & 2 & 2 & 0 & 10 & 2 & 2 & 6 &  23 & 1 & 1 & 0 & 1 & \bf 12 & 3 & 3 & 1 & 0 & 1 & 1 & 1 \\ 
-\#  2 & 0 & 16 & 1 & 1 & 5 & 1 & 0 & 8 & 0 & 2 & 0 & 0 & 0 & 4 & 2 & 1 & \bf 11 & 1 & 0 & 13 \\ 
-\#  3 & 7 & 26 & 0 & 2 & 0 & 5 & 3 & 0 & 2 & 1 & 0 & 4 & 0 & 0 & 2 & 0 & 9 & \bf 5 & 0 & 6 \\ 
-\#  4 & 0 & 0 & 0 & 2 & 0 & 0 & 0 & 3 & 1 & 10 & 2 & 18 & 1 & 0 & 4 & 2 &\bf 14 & 0 & 13 & 0 \\ 
-\# 5 &\bf 19 & 1 & 4 & 1 & 0 & 1 & 0 & 12 & 0 & 0 & 0 & 4 & 1 & 0 & 0 & 12 & 1 & 1 & 0 & 0 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-%\clearpage
 \section{Some Plots of Exploratory Data Analysis}
+
+This section includes some plots showing results from exploratory data analysis. The results are discussed in Section~\ref{sec:result_socio}.
 
@@
 
+\clearpage
+\section{Electoral Building Lineups and Results}
+%Figures~\ref{fig:elect-1} and \ref{fig:elect.2} through~\ref{fig:elect.5}
+
+Five lineups were shown to (different) Amazon Turk workers in an experiment.
+They all were created as described in the introduction of the paper.
+In order to not bias observers, no context information was given about how these plots were created or what data they were displaying.
+This also made it necessary to slightly more stylize the display.
+
+% latex table generated in R 3.0.1 by xtable 1.7-1 package
+% Wed Jun  5 17:39:09 2013
+\begin{table}[hbtp]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ lineup & \#1 & \#2 & \#3 & \#4 & \#5 \\ 
+  \hline
+  \# correct/ \#evaluation &  12/72 & 11/66 & 5/74 & 14/72 & 19/57 \\ 
+$p$-value & 0.00023 & 0.00041 & 0.31 & 1.2e-05 &  1.9e-11 \\ 
+data panel & $3\cdot4+1$ & $2^4+1$ & $4^2+2 $ & $12+\sqrt{25}$ & $2^3-7$ \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+% latex table generated in R 3.0.1 by xtable 1.7-1 package
+% Wed Jun  5 18:00:55 2013
+\begin{table}[hbtp]
+\centering
+\caption{\label{tbl:response}Overview of all choices by observers for each of the lineups.
+The correct choice is bolded.
+In most lineups there are null plots that were picked more often by observers, but the actual result is among the plots being picked most often, indicating that there is some indication that the election result is not completely consistent with the polls.}
+\scalebox{.9}{
+\begin{tabular}{rrrrrrrrrrrrrrrrrrrrr}
+  \hline
+& \multicolumn{10}{l}{panel chosen}\\
+Lineup & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\ 
+  \hline
+\#  1 & 2 & 2 & 0 & 10 & 2 & 2 & 6 &  23 & 1 & 1 & 0 & 1 & \bf 12 & 3 & 3 & 1 & 0 & 1 & 1 & 1 \\ 
+\#  2 & 0 & 16 & 1 & 1 & 5 & 1 & 0 & 8 & 0 & 2 & 0 & 0 & 0 & 4 & 2 & 1 & \bf 11 & 1 & 0 & 13 \\ 
+\#  3 & 7 & 26 & 0 & 2 & 0 & 5 & 3 & 0 & 2 & 1 & 0 & 4 & 0 & 0 & 2 & 0 & 9 & \bf 5 & 0 & 6 \\ 
+\#  4 & 0 & 0 & 0 & 2 & 0 & 0 & 0 & 3 & 1 & 10 & 2 & 18 & 1 & 0 & 4 & 2 &\bf 14 & 0 & 13 & 0 \\ 
+\# 5 &\bf 19 & 1 & 4 & 1 & 0 & 1 & 0 & 12 & 0 & 0 & 0 & 4 & 1 & 0 & 0 & 12 & 1 & 1 & 0 & 0 \\ 
+   \hline
+\end{tabular}
+}
+\end{table}
+
 \bibliographystyle{asa}

### 2013-06-21 sociology_paper.tex
cleaned up paragraph on demographics

@@
 
-Effect of graduate degree is estimated as 0.182 which is highly significant.
-For an 18-25 year old female in the United States, the expected difference in the probability of correct response, if she has a graduate degree as compared to a high school degree, is, for a lineup with random effect of zero:
+While some of the demographic factors are strongly statistically significant, the main source of variation in proportion correct is the lineup difficulty.
+For example, let's examine the effect of graduate degree.
+To see just how large the effect is, we examine the change in proportion correct for a (hypothetical) 18-25 year old female in the United States, with a graduate degree as compared to a high school degree, for an average difficulty lineup (random effect = 0).
+Plugging in the relevant quantities to the fitted model gives a difference equal to:
 $$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.041.$$ 
-For a lineup with a predicted random effect of one standard deviation, i.e.
-we move away one standard deviation of lineup variation by adding 2.293 to each of the exponent terms in the equation, this difference reduces to 0.024.
-For two standard deviations it further reduces to 0.003 which is practically very small.
-Figure~\ref{fig:practical_impact_graduate} illustrates this example showing the proportions of correct responses due to graduate degree as compared to high school degree.
-The highest practical difference in proportions of correct response is estimated as 0.045 which is very negligible.
-The difference further diminishes if we move one or two standard deviation of lineup variability.
-A similar calculation shows the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of correct response.
-Thus even though some of the factor levels are statistically significant, their practical impact is negligible.
+The person with a high school education on average picks the data plot in 33.5\% of lineups having average difficulty, as compare to 37.6\% if they have a graduate degree.
+This difference is reduced to 2\% for a lineup with one standard deviation order of magnitude difference in difficulty.
+For two standard deviations it further reduces to 0.3\%.
+Thus although there is statstically significant difference in proportion correct for some demographic factors, these are not practically significant differences.
+Figure~\ref{fig:practical_impact_graduate} illustrates this example showing fitted models for a US 18-25 female with either a high school education or a graduate degree.
+Similar calculations show the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of correct response.
+Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.
 

### 2013-06-21 sociology_paper.tex
table 1 scatter plot changed

@@
 between group A and B? \\
-2 &  Scatter plot & \begin{minipage}[t]{3cm}  \begin{center} \scalebox{0.45}{\includegraphics{stat_beta_k.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+2 &  Scatter plot & \begin{minipage}[t]{3cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
 \\
 
-  3 & Contaminated plot &\begin{minipage}[t]{3cm} \begin{center} \scalebox{0.45}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+  3 & Contaminated plot &\begin{minipage}[t]{3cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
 \\  \\
@@
 Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
- &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
-  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0376 \\ 
+ &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.206 \\ 
+  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.038 \\ 
   &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
-  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.0501 \\ 
-  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.1425 \\ 
-  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.0394 \\ 
+  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.050 \\ 
+  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.143 \\ 
+  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.039 \\ 
   \hline
@@
  &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
-  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.3307 \\ 
-  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.1566 \\ 
+  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.331 \\ 
+  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.157 \\ 
   &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
@@
 Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
-  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.1331 \\ 
+  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.133 \\ 
   \hline
@@
 \end{table}
+
+% This table had 4 digits $p$-value for last column
+%\begin{table}[htbp]
+%\centering
+%\caption{Parameter estimates of Models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and probability of correct lineup evaluations respectively.
+For time taken all the demographic factors are significant.
+For probability of correct response age group 36-40, rest of the world and graduate degree are significantly different.
+For gender no difference in performance is observed.
+Lineup variability is estimated to be very large for Model \eqref{eqn:demographic_response}.}
+%\scalebox{0.86}{
+%\begin{tabular}{rlrrrrrrrrrr}
+%  \hline
+%Demographic&& \multicolumn{4}{c} {Model \eqref{eqn:demographic_time} Log Time} & &\multicolumn{4}{c} {Model \eqref{eqn:demographic_response} Proportion Correct} \\
+%
+%\cline{3-6} \cline{8-11} 
+%Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
+%  \hline
+%\multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
+%&$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
+%\hline
+%Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
+% &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
+%  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0376 \\ 
+%  &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
+%  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.0501 \\ 
+%  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.1425 \\ 
+%  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.0394 \\ 
+%  \hline
+%Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
+%  &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
+%  &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
+%  \hline
+%Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
+% &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
+%  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.3307 \\ 
+%  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.1566 \\ 
+%  &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
+%  \hline
+%Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
+%  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.1331 \\ 
+%  \hline
+%\multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
+%& lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
+% & Error($\sigma$) & 0.479 & 0.692 &  &  & & && &\\ 
+%   \hline
+%\end{tabular}
+%}
+%\label{tbl:model_result_demographics}
+%\end{table}
 

### 2013-07-09 sociology_paper.tex
Figure and test statistics in table updated

@@
 \caption{Visual test statistics used in 10 different simulation experiments. The observers are asked different questions to answer while evaluating a lineup} 
-\begin{tabular}{m{.5cm}m{3cm}m{3cm}m{6.5cm}} 
+\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{6.5cm}} 
 \hline\hline 
@@
 \hline 
-1  & Box plot & \begin{minipage}[t]{3cm} \begin{center}	\scalebox{0.4}{\includegraphics{stat_category.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
+1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
 between group A and B? \\
-2 &  Scatter plot & \begin{minipage}[t]{3cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
 \\
 
-  3 & Contaminated plot &\begin{minipage}[t]{3cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+  3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
 \\  \\
-  4 & Polar vs Cartesian & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different? \\ 
-  5 & Hist vs density & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?\\
-  6 & Violin vs boxplot & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?\\
-  7 & Group separation & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?\\
-  8 & Sine Illusion & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?\\
-  9 & Gene expression &\begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?\\
-  10 & Test normality & \begin{minipage}[t]{3cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?\\
+  4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different? \\ 
+  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?\\
+  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?\\
+  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?\\
+  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?\\
+  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?\\
+  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?\\
 

### 2013-07-09 R/sociology_paper.R

@@ -188,12 +188,15 @@ pdat <- ddply(demographics, .(country,gender_level, age_level,degree), summarize
               per_correct = mean(response),
               avg_time = mean(time_taken))
 qplot(degree, avg_time, geom="bar", stat="identity", data=pdat[complete.cases(pdat),]) +
-  facet_grid(age_level~country+gender_level) + coord_flip() + xlab("Academic Study") 
+  facet_grid(age_level~country+gender_level) + coord_flip() + 
+  xlab("Academic Study") + ylab("Average time taken")
 
 ggsave("../images/age_gender_within_country_time.pdf", width=10, height=8)
 
-qplot(degree, per_correct, geom="bar", stat="identity", data=pdat[complete.cases(pdat),]) +
-  facet_grid(age_level~country+gender_level) + coord_flip() + xlab("Academic Study") 
+qplot(degree, per_correct*100, geom="bar", stat="identity", data=pdat[complete.cases(pdat),]) +
+  facet_grid(age_level~country+gender_level) + coord_flip() + 
+  xlab("Academic Study") + ylab("Percent correct") +
+  scale_y_continuous(breaks=c(25,50,75,100)) 
 
 ggsave("../images/age_gender_within_country_correct.pdf", width=10, height=8)
 

### 2013-07-09 sociology_paper.tex
Table 1 new column added

@@
 \caption{Visual test statistics used in 10 different simulation experiments. The observers are asked different questions to answer while evaluating a lineup} 
-\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{6.5cm}} 
+\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{4.5cm}m{3.5cm}} 
 \hline\hline 
-Serial & Experiment &  Test Statistic  & Lineup question \\ [0.5ex] % inserts table %heading 
+Serial & Experiment &  Test Statistic  & Lineup question & Provided data to estimate\\ [0.5ex] % inserts table %heading 
 \hline 
 1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
-between group A and B? \\
+between group A and B? & Demographic factor\\
 2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
-\\
-
+& Demographic factor\\
   3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
-\\  \\
-  4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different? \\ 
-  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?\\
-  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?\\
-  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?\\
-  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?\\
-  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?\\
-  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?\\
+& Demographic factor\\
+ 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& Demographic factor\\
+  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?& Learning trend and Demographic factor\\
+  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?
+& Learning trend and Demographic factor\\
+  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?
+& Learning trend and  Demographic factor\\
+  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?
+& Demographic factor\\
+  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?
+& Location effect and Demographic factor\\
+  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?
+& Demographic factor\\
 

### 2013-07-09 sociology_paper.tex
some section headings are modified

@@
 \centering 
-\caption{Visual test statistics used in 10 different simulation experiments. The observers are asked different questions to answer while evaluating a lineup} 
+\caption{Visual test statistics used in 10 different simulation experiments.
+The observers are asked different questions to answer while evaluating a lineup.
+All the experiment data are used to estimate the effect of demographic factors on visual inference while some are used to estimate the learning trend and location effect of actual data in the lineup.}
 \begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{4.5cm}m{3.5cm}} 
@@
 
-\section{Factors Affecting  Observer Performance} \label{sec:factor_performance}
+\section{Factors Affecting Visual Inference} \label{sec:factor_performance}
 
@@
 
-\subsection{Selection of Null Plots} A lineup becomes difficult to evaluate if one of the null plots appears to be very similar to the actual plot.
+\subsection{Sample of Null Plots} A lineup becomes difficult to evaluate if one of the null plots appears to be very similar to the actual plot.
 When null hypothesis is true it is a common scenario.
@@
 But we have only specific number of null plots in a lineup.
-So, a different set of null plots may yield some variation in the difficulty of the lineup with the same actual plot.
+So, a different sample of null plots may yield some variation in the difficulty of the lineup with the same actual plot.
 This may affect the performance of the observer while evaluating a lineup.
 
-\subsection{Individual Performance of the Observer} Each person is different from others in some way.
+\subsection{Individual Skill or Ability of the Observer} Each person is different from others in some way.
 For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.
@@
 
-\subsection{Location Effect}
+\subsection{Lineup Design}
+
+\subsubsection{Sample of lineup}
+\subsubsection{Sample of Null Plots}
+
+\subsubsection{Location Effect}
 

### 2013-07-18 sociology_paper.tex
reorganised the sections

@@
    \begin{minipage}[c]{0.45\textwidth}
-   Freedom's Lighthouse Averages \\
+   Poll aggregator 1 \\
       \includegraphics[width=\textwidth]{freedom-sim.pdf} 
-   RealClearPolitics Averages\\
+   Poll aggregator 2\\
       \includegraphics[width=\textwidth]{rcp-sim.pdf} 
@@
 For the histogram on the top, the $p$-value of observing results as extreme as  the  2012 U.S.
-election results based on the bootstrap is 0.0533 (with Bootstrap standard error of 0.002), making the election results almost significantly different from the polls.
-There is no indication of any inconsistency between polls and election results based on the bootstrap simulation below.
-The lineups are based on the top source.
-}
+election results based on the bootstrap is 0.0533 (with Bootstrap standard error of 0.002), making the election results almost significantly different from the polls (data source for poll aggregator 1 is Freedom`s Lighthouse Averages).
+There is no indication of any inconsistency between polls and election results based on the bootstrap simulation below (data source for poll aggregator 2 is RealClearPolitics Averages).
+The lineups are based on the top source.}
 \end{figure}
@@
 The observers are asked different questions to answer while evaluating a lineup.
-All the experiment data are used to estimate the effect of demographic factors on visual inference while some are used to estimate the learning trend and location effect of actual data in the lineup.}
-\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{4.5cm}m{3.5cm}} 
+All the experiment data are used to estimate the effect of demographic factors (DF) on visual inference while some are used to estimate the learning trend (LT) and location effect (LE) of actual data in the lineup as shown in the right most column.}
+\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
 \hline\hline 
-Serial & Experiment &  Test Statistic  & Lineup question & Provided data to estimate\\ [0.5ex] % inserts table %heading 
+Serial & Experiment &  Test Statistic  & Lineup question & Designed to study\\ [0.5ex] % inserts table %heading 
 \hline 
 1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
-between group A and B? & Demographic factor\\
+between group A and B? & DF\\
 2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
-& Demographic factor\\
+& DF\\
   3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
-& Demographic factor\\
- 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& Demographic factor\\
-  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?& Learning trend and Demographic factor\\
+& DF\\
+ 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
+  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF and LT \\
   6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?
-& Learning trend and Demographic factor\\
+&  DF and LT \\
   7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?
-& Learning trend and  Demographic factor\\
+& DF and LT \\
   8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?
-& Demographic factor\\
+& DF\\
   9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?
-& Location effect and Demographic factor\\
+& DF and LE\\
   10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?
-& Demographic factor\\
+& DF\\
 
@@
 
-\section{Factors Affecting Visual Inference} \label{sec:factor_performance}
+\section{Factors Potentially Affecting Visual Inference} \label{sec:factor_performance}
 
@@
 
-\section{Experimental Designs and Methods}\label{sec:exp_design}
+\section{Experimental Methods}\label{sec:exp_design}
 While evaluating a lineup, the performance of the observer may depend on the factors described in Section \ref{sec:factor_performance}.
@@
 
-\subsection{Experiment Setup}  In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table \ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
+\subsection{Demographic Factors}  In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table \ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
 
@@
 
-%Multiple simulation experiments are designed to investigate the learning trend of the observer and the location effect of the actual data in the lineup. These are explained in tis section.
-
-\subsubsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive feedbacks received when multiple lineups are shown for evaluation.
-Experiments 5, 6 and 7 were used for this.
-Each subject was shown a total of 10 lineups randomly selected from a pool of lineups.
-The lineups are not necessarily with the same difficulty levels.
-But the order of lineups were randomized.
-The responses of the lineups were recorded by attempt 1 through 10.
-Attempt 1 means that the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
-The goal is to estimate whether performance of the observer improves from attempt 1 to attempt 10.
-
-\subsubsection{Design for Location Effect} \label{sec:location_design} Experiment 9 shown in Table~\ref{tbl:visual_stat} is designed to to assess the location effect of a data plot in a lineup.
-It is set up based on the gene expression data \citep{atwood:2013} with two groups.
-One is the Genotype and the other the Empty Vector (EV).
-For each of the groups, gene expression data were collected in presence or absence of iron sufficiency.
-Two factors were of primary interest.
-One is the Genotype main effect and the other is the interaction effect between Genotype and iron sufficiency.
-
-For Genotype main effect side by side dot plots of two groups are used as the visual test statistics.
-For interaction effect the same plot type is used except that the means of iron sufficiency or insufficiency were connected by lines colored by the groups.
-One of the interaction test statistics is shown in Table~\ref{tbl:visual_stat} for experiment 9.
-For no interaction, the visual test statistic shows two parallel lines representing two groups and for significant interaction the lines are not parallel.
-Null plots were generated by randomizing the group structure of the data.
-The actual plot was randomly placed in five different locations in a lineup of size 20.
-The locations are 2,9,12,16,20 for Interaction effect and 1,8,12,17,20 for Genotype effect.
-For a lineup with specific data plot location, five different sets of null plots were used to produce 5 lineups for each location.
-In total we have 25 lineups for Interaction effect and 25 lineups for Genotype effect.
-
-%We set up a Turk experiment to examine whether there is any difference of performance based on the location of actual data plot in the lineup.
-For this five locations of a lineup were randomly chosen to put the same actual data plot.
-The locations are 2,9,12,16,20 for Interaction effect and 1,8,12,17,20 for Genotype effect.
-For a lineup with specific data plot location, five sets of null plots are used giving 5 lineups for each location.
-In total we have 25 lineups for Interaction effect and 25 lineups for Genotype effect.
-These 50 lineups are evaluated by 100 people recruited from Amazon Mechanical Turk.
-Each person evaluated 3 lineups one from Interaction, one from Genotype and the third one was a test lineup.
-The test lineup is used to process the payment and examine the quality of the data and the responses of the test plot is not added to our analysis.
-
-%One such lineup is shown in Figure \ref{fig:lineup_geno}.
-
-Each observer saw three lineups, first an Interaction lineup then a Genotype lineup.
-Finally a test lineup was shown to screen out unusual evaluation.
-The subjects did not know which one was the test lineup but they were informed before accepting the task that there would be one.
-The test lineup was very easy and everyone should detect the data plot.
-The MTurk workers were paid based on whether they could correctly evaluate the test lineup.
-But the data on the test plot are excluded from the analysis.
-
-\subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \citep{turk} or MTurk Web site.
-It is an online work place where people from around the world can perform some tasks and get paid.
-Usually tasks are very simple and no specialized training is required to do them.
-Being a human is the main requirement.
-Tasks are designed for anyone to do but some tasks may require some skills depending on the recruiters' need.
-Each task is usually planned to complete in a short time.
-Humans are still better than computers in performing these types of tasks.
-The the amount of money paid for each task is very small as well.
-
-%It is very fast, cheap and reliable to recruit people from MTurk.
-Thats why it is getting very popular among the researchers who perform human subject experiment.
-The another benefit is that a very diverse pool of subjects can be recruited which is otherwise very hard to obtain for a study.
-The researchers can easily filter the workers based on their experimental design, such as recruiting people only from a specific geographical location or a group of people who satisfy certain criteria etc.
-The recruiter can decide who they pay or not.
-Workers have to satisfy the task requirement to ensure payment.
-But at the end it is the recruiter who has the final say.
-Usually recruiters pay promptly after the task has been done properly and thats why MTurk is very popular among the online job seeker.
-At any point of time thousands of tasks are available for the thousands of workers around the world.
-
-%Because of its convenience it is getting popular for scientific research study as well.
-In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time and cost while provided more convenience.
-\cite{majumder:2013} recruited people from MTurk for their simulation study in estimating the power of visual statistical inference.
-They have done numerous pilot studies in lab before doing actual MTurk study and found similar results.
-\cite{mason:2012} explains various features of MTurk and describes how it can be used as part of human behavioral study.
-
-We designed and developed a web application which enables to display the lineups to the observers as per experimental need.
-The MTurk workers were redirected to the web site to evaluate lineups.
-The data were collected, stored automatically into a local database server.
-Demographic informations such as age group, gender and education levels were also collected.
-The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
-It is measured in seconds.
-The location of the observer is determined by the ip address of the observer.
-
-\subsection{Model to Estimate Demographic Factor Effect} The feedback provided by each observer on a lineup is a binary response variable.
+The feedback provided by each observer on a lineup is a binary response variable.
 Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$.
@@
 
-% and $k(i) \in \{18-25, 26-30, 31-35, 36-40, 41-45, 46-50, \text{above } 50 \}$.
-$\gamma$ is the effect of gender category $l(i) \in \{\text{Male, Female}\}$.
-$\tau$ is the effect of education levels $m(i) \in \{ \text{High school, Undergrad course, Undergrad degree, Graduate course, }\}$
-
 The effect of demographic factors can be observed as time taken for each evaluation by an observer.
@@
 
-\subsection{Model to Estimate Learning Trend}
+\subsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive feedbacks received when multiple lineups are shown for evaluation.
+Experiments 5, 6 and 7 were used for this.
+Each subject was shown a total of 10 lineups randomly selected from a pool of lineups.
+The lineups are not necessarily with the same difficulty levels.
+But the order of lineups were randomized.
+The responses of the lineups were recorded by attempt 1 through 10.
+Attempt 1 means that the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
+The goal is to estimate whether performance of the observer improves from attempt 1 to attempt 10.
+
+%\subsection{Model to Estimate Learning Trend}
 
 Multiple lineups were sequentially shown to each observer for evaluation.
-Suppose $Y_{ijk}$ denotes the response from observer $i$ on a  lineup $j$ in his/her $k$th attempt.
+Suppose $Y_{ijk}$ denotes the response from observer $i$ on a  lineup $j$ in his/her $k^{th}$ attempt.
 $Y_{ijk}=1$ if the response is correct otherwise  $Y_{ijk}=0$.
-Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in his/her $k$th attempt.
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in his/her $k^{th}$ attempt.
 We model this in a generalized mixed effects model of the form
@@
 \end{equation}
-where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k$th attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$.
+where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k^{th}$ attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$.
 $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$. $u_i$ is a random intercept, describing a basic subject-specific ability. We assume $u_i \sim N(0, \sigma_u^2)$. 
@@
 
-%$W$ is a design matrix of covariates corresponding to specifics of lineup $\ell$ and subject $i$, and $\delta$ is the vector of corresponding parameters.
-Covariates could include  demographic information of individuals, such as age, gender, education level etc.,  as well as the attempt made by the observer.
-$Z_{\ell i}$,  $1 \le i \le K$, $1 \le \ell \le L$,  is a design matrix corresponding to random effects specific to individual $i$ and lineup $\ell$; and
-%$\tau$  is a vector of independent normally distributed random variables $\tau_{\ell i}$ with  variance matrix $\sigma_\tau I_{KL \times KL}$.
-$\tau$ will usually include a component incorporating an individual's ability or skill to evaluate lineups.
-
-The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $kth$ attempt by a single observer as
+The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k^{th}$ attempt by a single observer as
 \begin{equation} \label{eqn:trend_power}
@@
 The learning of each observer over a course of $K$ evaluations may be observed as the improvement of the time taken to evaluate a lineup in the later attempts.
-Suppose $Z_{ijk}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
-Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k$th attempt.
+Suppose $Z_{ijk}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k^{th}$ attempt.
+Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k^{th}$ attempt.
 We model this in a mixed effects model of the form
@@
 $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-%}
-
-%\begin{equation} \label{eqn:trend_time}
-%Z_{ijk} = \mu_1 + \alpha_k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
-%\end{equation}
-%where the link function $g_1(\mu)=1/\mu;  \mu >0$ and $\mu_1$ represents the inverse of overall average time taken to evaluate a lineup.
 
@@
 
-%For $Y_{l,i}$ being binary response we use {\it logit} link function $g(\mu)=g_1(\mu)=\log(\mu) - \log(1-\mu); 0 \le \mu \le 1$ and $\hat p_{\ell i}$ gives the estimate of proportion correct by subject $i$ with lineup $\ell$.
-The response time taken is a positive random variable with a right skewed distribution.
-For this the usual convention is to model a gamma distribution with time.
-Thus we used inverse link function $g(\mu)=g_2(\mu)=1/\mu;  \mu >0$ for gamma response $Y_{l,i}$ and $\hat p_{\ell i}$ gives the estimate of mean time taken by subject $i$ for lineup $\ell$.
-
 To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
@@
 
-\subsection{Model to Estimate Location Effect}
+\subsection{Location Effect} \label{sec:location_design} Experiment 9 shown in Table~\ref{tbl:visual_stat} is designed to to assess the location effect of a data plot in a lineup.
+It is set up based on the gene expression data \citep{atwood:2013} with two groups.
+One is the Genotype and the other the Empty Vector (EV).
+For each of the groups, gene expression data were collected in presence or absence of iron sufficiency.
+Two factors were of primary interest.
+One is the Genotype main effect and the other is the interaction effect between Genotype and iron sufficiency.
+
+For Genotype main effect side by side dot plots of two groups are used as the visual test statistics.
+For interaction effect the same plot type is used except that the means of iron sufficiency or insufficiency were connected by lines colored by the groups.
+One of the interaction test statistics is shown in Table~\ref{tbl:visual_stat} for experiment 9.
+For no interaction, the visual test statistic shows two parallel lines representing two groups and for significant interaction the lines are not parallel.
+Null plots were generated by randomizing the group structure of the data.
+The actual plot was randomly placed in five different locations in a lineup of size 20.
+The locations are 2,9,12,16,20 for Interaction effect and 1,8,12,17,20 for Genotype effect.
+For a lineup with specific data plot location, five different sets of null plots were used to produce 5 lineups for each location.
+In total we have 25 lineups for Interaction effect and 25 lineups for Genotype effect.
+
+Each observer saw three lineups, first an Interaction lineup then a Genotype lineup.
+Finally a test lineup was shown to screen out unusual evaluation.
+The subjects did not know which one was the test lineup but they were informed before accepting the task that there would be one.
+The test lineup was very easy and everyone should detect the data plot.
+The MTurk workers were paid based on whether they could correctly evaluate the test lineup.
+But the data on the test plot are excluded from the analysis.
+
+%\subsection{Model to Estimate Location Effect}
 
@@
 where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$.
+
+\subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \citep{turk} or MTurk Web site.
+It is an online work place where people from around the world can perform some tasks and get paid.
+Usually tasks are very simple and no specialized training is required to do them.
+Being a human is the main requirement.
+Tasks are designed for anyone to do but some tasks may require some skills depending on the recruiters' need.
+Each task is usually planned to complete in a short time.
+Humans are still better than computers in performing these types of tasks.
+The the amount of money paid for each task is very small as well.
+
+%It is very fast, cheap and reliable to recruit people from MTurk.
+Thats why it is getting very popular among the researchers who perform human subject experiment.
+The another benefit is that a very diverse pool of subjects can be recruited which is otherwise very hard to obtain for a study.
+The researchers can easily filter the workers based on their experimental design, such as recruiting people only from a specific geographical location or a group of people who satisfy certain criteria etc.
+The recruiter can decide who they pay or not.
+Workers have to satisfy the task requirement to ensure payment.
+But at the end it is the recruiter who has the final say.
+Usually recruiters pay promptly after the task has been done properly and thats why MTurk is very popular among the online job seeker.
+At any point of time thousands of tasks are available for the thousands of workers around the world.
+
+%Because of its convenience it is getting popular for scientific research study as well.
+In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time and cost while provided more convenience.
+\cite{majumder:2013} recruited people from MTurk for their simulation study in estimating the power of visual statistical inference.
+They have done numerous pilot studies in lab before doing actual MTurk study and found similar results.
+\cite{mason:2012} explains various features of MTurk and describes how it can be used as part of human behavioral study.
+
+We designed and developed a web application which enables to display the lineups to the observers as per experimental need.
+The MTurk workers were redirected to the web site to evaluate lineups.
+The data were collected, stored automatically into a local database server.
+Demographic informations such as age group, gender and education levels were also collected.
+The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
+It is measured in seconds.
+The location of the observer is determined by the ip address of the observer.
+
+%\subsection{Model to Estimate Demographic Factor Effect} The feedback provided by each observer on a lineup is a binary response variable.
+Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$.
+$Y_{ij} =1$ if the response is correct otherwise $Y_{ij} =0$.
+Let $\pi_{ij}=  E(Y_{ij})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$.
+We model this in a generalized mixed effects model of the form
+%\begin{equation} \label{eqn:demographic_response}
+%g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
+%\end{equation}
+%where $\mu$ is an overall average probability for picking out the data plot from a lineup.
+$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender category $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$.
+$\ell_j$ is a random intercept predicting lineup difficulty level and we assume  $\ell_j \sim N(0, \sigma_\ell^2)$.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+%
+%
+%The effect of demographic factors can be observed as time taken for each evaluation by an observer.
+Suppose $Z_{ij}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$.
+Let $\mu_{ij}=  E(Z_{ij})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$.
+We model this in a mixed effects model of the form
+%%\hh{
+%\begin{equation} \label{eqn:demographic_time}
+%Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
+%\end{equation}
+%where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
+$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model \eqref{eqn:demographic_response}.
+$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
+
+%\subsection{Model to Estimate Learning Trend}
+%
+%Multiple lineups were sequentially shown to each observer for evaluation.
+Suppose $Y_{ijk}$ denotes the response from observer $i$ on a  lineup $j$ in his/her $k^{th}$ attempt.
+$Y_{ijk}=1$ if the response is correct otherwise  $Y_{ijk}=0$.
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in his/her $k^{th}$ attempt.
+We model this in a generalized mixed effects model of the form
+%\begin{equation} \label{eqn:trend_response}
+%g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
+%\end{equation}
+%where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k^{th}$ attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$.
+%$u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$. $u_i$ is a random intercept, describing a basic subject-specific ability. We assume $u_i \sim N(0, \sigma_u^2)$. 
+%$a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts. We assume $a_i \sim N(0, \sigma_a^2)$. 
+%For $\ell_j$ we again assume a normal distribution, $N(0, \sigma_\ell^2)$.
+$\ell_j$ is a random intercept predicting lineup difficulty level.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ .
+%
+%The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k^{th}$ attempt by a single observer as
+%\begin{equation} \label{eqn:trend_power}
+%\hat p_{ijk} =  g^{-1}(\hat{\mu} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
+%\end{equation}
+%
+%The learning of each observer over a course of $K$ evaluations may be observed as the improvement of the time taken to evaluate a lineup in the later attempts.
+Suppose $Z_{ijk}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k^{th}$ attempt.
+Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k^{th}$ attempt.
+We model this in a mixed effects model of the form
+%%\hh{
+%\begin{equation} \label{eqn:trend_time}
+%Z_{ijk} = \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
+%\end{equation}
+%where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
+$\alpha$ is the average change in log time taken for each additional attempt.
+$\alpha_1$ is an offset in log time taken for the first attempt.
+All other effects are random effects: as before, $u_i$ is a subject-specific intercept representing  individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$.
+$a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$.
+We assume $a_i \sim N(0, \sigma_a^2)$.
+$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
+%
+%
+%Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $kth$ attempt by a single observer as 
+%\begin{equation} \label{eqn:trend_time_est}
+%\hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j.
+%\end{equation}
+%
+%To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
+To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
+
+%\subsection{Model to Estimate Location Effect}
+%
+%
+%As per the design of experiment for location effect, the actual data plot is same for each of the null sets of plot.
+Thus the response data of this experiment constitute a multivariate response.
+To examine if the difference in proportion correct among the locations is statistically significant we fit a one way multi variate analysis of variance (MANOVA) model to the data.
+%
+%Suppose $\mathbf{Y}=(Y_1,Y_2, ...
+, Y_p)^\top$ be a vector of random variable with dimension $p$, the total number of null sets.
+let $\mathbf{Y}_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., I$.
+We fit the following MANOVA model
+%\begin{equation}\label{manova}
+%\mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
+%\end{equation}
+%where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$.
 

### 2013-07-18 sociology_paper.tex
tweaking little things

@@
 \centering 
-\caption{Visual test statistics used in 10 different simulation experiments.
-The observers are asked different questions to answer while evaluating a lineup.
-All the experiment data are used to estimate the effect of demographic factors (DF) on visual inference while some are used to estimate the learning trend (LT) and location effect (LE) of actual data in the lineup as shown in the right most column.}
+\caption{Overview of 10 different Turk experiments, from where data was taken to study human factor effects.
+All of the experimental data was used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.}
 \begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
 \hline\hline 
-Serial & Experiment &  Test Statistic  & Lineup question & Designed to study\\ [0.5ex] % inserts table %heading 
+ & Experiment &  Test Statistic  & Lineup question & Used in study of\\ [0.5ex] % inserts table %heading 
 \hline 
@@
 
-Based on human evaluation on a lineup a decision is made in visual statistical inference.
+Because visual statistical inference relies on human evaluation, it is important to understand the factors that might affect results.
+This section provides a brief description of the factors that may influence results.
+
+%ased on human evaluation on a lineup a decision is made in visual statistical inference.
 While the visual test statistic is defined to have a direct influence on the observer to pick a plot based on the signal in the data, there are other human factors that may affect the observer performance.

### 2013-07-18 sociology_paper.tex
working on section 2

@@
 
-Because visual statistical inference relies on human evaluation, it is important to understand the factors that might affect results.
-This section provides a brief description of the factors that may influence results.
+Because visual statistical inference relies on plotting choices, lineup design and human evaluation, it is important to understand how these factors that might affect results.
+This section provides a brief description of the factors that are expected to have some impact on the performance of visual statistical inference.
 
@@
 %\dc{You need to have more than one sentence to open this section, it needs a paragraph overviewing the factors.}
+
+\subsection{Choice of Visual Test Statistic}
+
+There are a variety of ways that data can be plotted and enhanced.
+The choice of plot that is used as the visual test statistic should be closely associated with the problem being investigated.
+For example, side-by-side boxplots may be more appropriate than side-by-side dotplots, when one variable is categorical.
+Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying some types of problems.
+Some choices of visual test statistics provide better results than others when studying the same problem, which is discussed in \cite{heike:2012}.
+This makes sense, because some plot types will make it easier for see some structure.
+It also presents as a flip-side to visual inference, that the same methods could be used to determine the optimal plot choice.
+
+Table \ref{tbl:visual_stat} shows the visual test statistics used in the Turk studies completed so far.
+Side-by-side boxplots were used in experiment 1 to study the importance of a categorical variable in a regression model.
+For both experiments 2 and 3 a scatterplot is used but a regression line fitted through the points is overlaid for experiment 2.
+It is easier to spot the slope of the line compared to just the scatterplot itself.
+But people are better in noticing the unusual pattern in the data which is some contamination purposefully added for experiment 3.
+Experiment 4 compared polar coordinates to euclidean coordinates on the reading of structure in a barchart.
+In addition, this experiment also examined the style of plot for a bivariate data problem: a scatterplot, side-by-side dot plots, overlaid densities or side-by-side boxplots.
+....
+expand description of experiments here...
 
@@
 It force the observer to chose a wrong plot when the null hypothesis should not be rejected.
-
-\subsection{Choice of Visual Test Statistic}  The visual test statistic should be highly associated with the hypothesis under investigation.
-To achieve this purpose it is very important to decide which plot type and plot features should be adopted.
-In a linear regression setting, the visual test statistics are presented in \citep{majumder:2013} for some common hypothesis testing.
-It is also observed that a scatterplot may do a better job than a box plot when using as a visual test statistic for regression parameters.
-Some of the effective features of visual test statistics are discussed in \citep{heike:2012} including plot type, color and shape of the plots.
-\cite{niladri:2012} presents some \textit{distance measure} to determine how a plot may be different from each other.
-
-Table \ref{tbl:visual_stat} shows the visual test statistics used in this study.
-For both experiments 2 and 3 a scatterplot is used but a regression line fitted through the points is overlaid for experiment 2.
-It is easier to spot the slope of the line compared to just the scatterplot itself.
-But people are better in noticing the unusual pattern in the data which is some contamination purposefully added for experiment 3.
-For experiment 3 the performance of the observers was much better compared to experiment 2 \citep{majumder:2013}.
 

### 2013-07-19 R/sociology_paper.R
new figure on education effect added

@@ -387,6 +387,25 @@ qplot(difficulty, value, data=mddat, color=variable, geom="line")+
 ggsave("../images/practical_impact_graduate.pdf", width=6.5, height=4.5)
 
 
+diff <- c(-2*2.293, -0.683, 2*2.293)
+hs_xb <- -0.683 + diff
+uc_xb <- hs_xb -0.083
+ud_xb <- hs_xb -0.044
+gc_xb <- hs_xb + 0.070
+gd_xb <- hs_xb + 0.182
+ddat.all <- data.frame(difficulty=diff, high.school = px(hs_xb),u.grad.course = px(uc_xb),
+                   u.grad.degree=px(ud_xb), grad.course=px(gc_xb), grad.degree=px(gd_xb))
+mddat.all <- melt(ddat.all,id=c("difficulty"))
+
+qplot(variable,value, color=factor(difficulty), data=mddat.all, size=I(3.5)) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  ylab("Proportion of correct responses") + xlab("Education")+
+  scale_color_discrete(name="Lineup difficulty", 
+                       labels=c("Difficult", "Medium","Easy"))
+
+ggsave("../images/practical_impact_degree.pdf", width=5.5, height=4.5)
+
+
 # Time of the day when Mturk worker works
 turker$hours <- as.factor(hour(as.POSIXlt(turker$start_time)))
 qplot(hours, data=subset(turker, complete.cases(turker))) +

### 2013-07-19 sociology_paper.tex
new figure on education effect added
still working on section 2

@@
 
-There are a variety of ways that data can be plotted and enhanced.
-The choice of plot that is used as the visual test statistic should be closely associated with the problem being investigated.
-For example, side-by-side boxplots may be more appropriate than side-by-side dotplots, when one variable is categorical.
-Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying some types of problems.
+There are a variety of ways that data can be plotted and for plots to be enhanced by axes, legends or representations of models.
+The choice of plot chosen as the visual test statistic, primarily should be appropriate for the problem being investigated.
+For example, for studying the association between two quantitative variables a scatterplot is usually optimal.
+Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.
 Some choices of visual test statistics provide better results than others when studying the same problem, which is discussed in \cite{heike:2012}.
-This makes sense, because some plot types will make it easier for see some structure.
-It also presents as a flip-side to visual inference, that the same methods could be used to determine the optimal plot choice.
-
-Table \ref{tbl:visual_stat} shows the visual test statistics used in the Turk studies completed so far.
+This makes sense, because some plot types will make it easier for certain structures to be seen.
+
+Table \ref{tbl:visual_stat} shows the large variety of visual test statistics used in the Turk studies completed so far.
+These experiments were conducted to examine many different visual tasks.
 Side-by-side boxplots were used in experiment 1 to study the importance of a categorical variable in a regression model.
-For both experiments 2 and 3 a scatterplot is used but a regression line fitted through the points is overlaid for experiment 2.
-It is easier to spot the slope of the line compared to just the scatterplot itself.
-But people are better in noticing the unusual pattern in the data which is some contamination purposefully added for experiment 3.
+For both experiments 2 and 3, a scatterplot is the visual test statistic, but a regression line fitted through the points is overlaid for experiment 2.
+The experiments studied slightly different problems, for which the difference in plot type was important.
 Experiment 4 compared polar coordinates to euclidean coordinates on the reading of structure in a barchart.
-In addition, this experiment also examined the style of plot for a bivariate data problem: a scatterplot, side-by-side dot plots, overlaid densities or side-by-side boxplots.
-....
-expand description of experiments here...
+Experiment 5 examined the style of plot for a bivariate data problem, by comparing results when using a scatterplot, side-by-side dot plots, overlaid densities or side-by-side boxplots.
+Experiment 6 studied the difference between variations of boxplots.
+Experiment 7 studied the effects of high dimension low sample size on the separation of groups in low-dimensional projections.
+The sine illusion, where people read the strength of differences between curves as different when there is a strong sinusoidal structure, is examined in experiment 8.
+Experiment 9 was conducted to test if there was any structure in an RNA-seq data set measured on soybeans, and used interaction plots classically constructed to examine interaction effects in $2\times 2$ factorial designs.
+Experiment 10 used different sorts of residual plots for hierarchical linear models as the visual test statistics.
+
+In experiments 4, 5 and 6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices.
+Results are published in \cite{heike:2012} and \cite{XXX}(The violin plot study needs to be referenced).
 
@@
 
-The visual test statistic is chosen so that it displays a specific pattern in case the null hypothesis is not true.
-Thus the most important factor that help an observer to correctly evaluate a lineup is the presence of any detectable signal in the data.
-On the other hand, if the null hypothesis is not true, visual test statistic should not display any distinguishable pattern.
+The visual test statistic is chosen for a particular task.
+If it is well-designed then departures from the null hypothesis will be visible in the plot, especially if the signal is strong.
+An important factor that helps an observer to identify the actual data plot in a lineup is the strength of the signal in the data.
+Simulated data was generated for all of the experiments, except 9, enabling the study of signal strength and actual data plot detection.
+Primarily the results were as expected, that as the signal strength increased observers more frequently picked the actual data plot as different from the null plots.
+
+%On the other hand, if the null hypothesis is not true, visual test statistic should not display any distinguishable pattern.
 In fact, some of the null plots may appear to be the most different plot in a lineup influencing the observer to chose a plot different than actual data plot.
@@
 
-\subsection{Question that Human Observer Answers} The researcher knows about the underlying hypothesis but the observer does not necessarily know the underlying details of the lineup.
+\subsection{Question Design} The researcher knows about the underlying hypothesis but the observer does not necessarily know the underlying details of the lineup.
 So, the researcher needs to ask a question to the observer to answer while evaluating the lineup.
@@
 
-\subsection{Demographics of the Observer}  Some of the demographics of the observer such as age, gender, education level and geographical location may have effect on how an observer would examine the lineup.
+\subsection{Demographics}  Some of the demographics of the observer such as age, gender, education level and geographical location may have effect on how an observer would examine the lineup.
 This may produce some variability in the performance of the observer.
@@
 
-\subsection{Learning Trend of the Observer} When an observer evaluates a lineup he or she may learn something from the experiences of the earlier observations.
+\subsection{Learning Trend} When an observer evaluates a lineup he or she may learn something from the experiences of the earlier observations.
 If the same observer is shown another lineup the learning from previous evaluation may help.
@@
 
-\subsection{Location of Actual Plot in the Lineup} The actual data plot is placed in a random spot in a lineup.
+\subsection{Location of Actual Data Plot in the Lineup} The actual data plot is placed in a random spot in a lineup.
 While evaluating the lineup some people may start looking from some specific part of the lineup.
@@
 
-\subsection{Individual Skill or Ability of the Observer} Each person is different from others in some way.
+\subsection{Individual Skill or Ability} Each person is different from others in some way.
 For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.
@@
 %   \includegraphics[width=6.5in]{learning_trend.pdf} 
-    \includegraphics[width=3.5in]{practical_impact_graduate.pdf} 
+    \includegraphics[width=3.0in]{practical_impact_graduate.pdf} 
+    \includegraphics[width=3.0in]{practical_impact_degree.pdf} 
    \caption{Proportion of correct responses due to graduate degree as compared to high school degree for an 18-25 year old female in the United States.

### 2013-07-22 sociology_paper.tex
section 2.3

@@
 
-The visual test statistic is chosen for a particular task.
+A visual test statistic is chosen for a particular task.
 If it is well-designed then departures from the null hypothesis will be visible in the plot, especially if the signal is strong.
@@
 
-\subsection{Question Design} The researcher knows about the underlying hypothesis but the observer does not necessarily know the underlying details of the lineup.
+\subsection{Question Design} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
+Visual inference supports discovery of srtucture, enabling the the unexpected to surprise us.
+(\citet{buja:2009} has a very complete explanation.) This is tightly connected to the task that the human observer is asked to perform.
+For most purposes the oberver is asked a very general question, such as to pick the plot that is most different from the others, and explain the reasons why they see it as different.
+By being general, all possible discoverable structures are included in the significance calculations.
+
+In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted.
+This was important in order to make a comparison of the results.
+For latter studies (4, 7, 10), question posed to the observer was broader.
+Table \ref{tbl:visual_stat} lists the questions used for each experiment.
+
+% The researcher knows about the underlying hypothesis but the observer does not necessarily know the underlying details of the lineup.
 So, the researcher needs to ask a question to the observer to answer while evaluating the lineup.
@@
 
-These questions are very crucial for the power of visual test.
+%These questions are very crucial for the power of visual test.
 They help observer think in a meaningful direction.

### 2013-07-22 sociology_paper.tex
sections 2.5/2.6

@@
 Visual inference supports discovery of srtucture, enabling the the unexpected to surprise us.
-(\citet{buja:2009} has a very complete explanation.) This is tightly connected to the task that the human observer is asked to perform.
+This is tightly connected to the task that the human observer is asked to perform.
 For most purposes the oberver is asked a very general question, such as to pick the plot that is most different from the others, and explain the reasons why they see it as different.
 By being general, all possible discoverable structures are included in the significance calculations.
+(\citet{buja:2009} has a very complete explanation.)
 
@@
 
-\subsection{Demographics}  Some of the demographics of the observer such as age, gender, education level and geographical location may have effect on how an observer would examine the lineup.
+\subsection{Demographics}  The experiments collected data on age, gender, education level and geographic location.
+Each observer self-reported their gender, age in roughly five year intervals, education level as high school or less, some college courses, an undergraduate degree, some graduate courses or a graduate degree.
+The IP address of the computer afforded the geographical location of the subject.
+The purpose of collecting this information with each experiment is to examine the effect that they have on the results of visual inference -- ideally very little.
+
+%Some of the demographics of the observer such as age, gender, education level and geographical location may have effect on how an observer would examine the lineup.
 This may produce some variability in the performance of the observer.
@@
 
-To meet the Institutional Review Board (IRB) requirement, it is necessary to exclude any subject less than 18 years of old to participate human subject experiment.
+%To meet the Institutional Review Board (IRB) requirement, it is necessary to exclude any subject less than 18 years of old to participate human subject experiment.
 Thus we intend to only focus on the performance of observer with age 18 years or older.
@@
 
-Education level has some relation with age.
+%Education level has some relation with age.
 Younger observer may not have higher degree.
@@
 
-We don't expect much difference in performances between male and female observers.
+%We don't expect much difference in performances between male and female observers.
 Gender may have some influence through the education level and geographical locations since some location may have small number of educated female population.
@@
 
-\subsection{Learning Trend} When an observer evaluates a lineup he or she may learn something from the experiences of the earlier observations.
+\subsection{Learning Trend} One might expect that as an observer evaluates more lineups they may become more skillful in their evaluations.
+Each subject in the experiments 5, 6 and 7 evaluated a block of of ten lineups of the same type of data plot.
+The ten lineups were randomly chosen from the lineups  produced for each study.
+Before evaluating the first lineup, the observer needs to read instructions and become accustomed to the type of plot used in the lineup.
+In subsequent lineups being evaluated the type of plot is the same, level of difficulty will be different.
+It is possible that the observer becomes more skillful at recognizing the most different plot, either by correctly finding the actual data plot or more quickly reporting their choice.
+These two ways of measuring learning trend are evaluated on data collected from experiments 5, 6 and 7.
+
+%When an observer evaluates a lineup he or she may learn something from the experiences of the earlier observations.
 If the same observer is shown another lineup the learning from previous evaluation may help.
@@
 
-Learning may occur in two different ways.
+%Learning may occur in two different ways.
 An observer can become more trained on the lineup structure and the pattern in the plots shown.
@@
 
-When an observer evaluate a lineup for the first time, he or she has to read through many instructions and become familiar with the environment of the experiment.
+%When an observer evaluate a lineup for the first time, he or she has to read through many instructions and become familiar with the environment of the experiment.
 We expect that this will require much more time compared to the rest of the evaluations.
@@
 
-\subsection{Location of Actual Data Plot in the Lineup} The actual data plot is placed in a random spot in a lineup.
-While evaluating the lineup some people may start looking from some specific part of the lineup.
-With the help of eye-tracking equipments \citet{zhao:2012} tracked the observers' eyes to see how they went through the plots in a lineup to come to their answers.
-The results suggest that people have particular methods of reading lineups.
-Some people read lineups from left to right direction while some read from upward to downward.
-Some people start looking from the center of the lineup while others start from the top left corner.
+\subsection{Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 points is used.
+A random number generator is used to determine the position where the actual data plot is placed in each lineup.
+With the help of eye-tracking equipments \citet{zhao:2012} tracked the observers' eyes to see how they searched the plots in a lineup.
+The results suggest that many people read lineups from left to right direction, some read from upward to downward.
+There is some concern that if the actual data plot is located in the top left of the lineup, observers may find it more easily than if it were placed in the bottom right.
+Ideally this does not happen.
+Experiment 9, conducted to test for structure in a real data set, enables this to be studied.
+Five different locations in the lineup were used to assess how fast and accurately observers identified the actual data plot.
+
+%While evaluating the lineup some people may start looking from some specific part of the lineup.
 In the earlier phase of the exercise, the observer tend to scan the plots and start comparing plots to make a final decision.
@@
 
-Given a specific pattern of eye movement of the observer in examining the lineup, the location of actual plot in the lineup may have some effect.
+%Given a specific pattern of eye movement of the observer in examining the lineup, the location of actual plot in the lineup may have some effect.
 Those who start exploring from the left top corner may get first glance of the actual plot if it is on that location.
@@
 
-It may be possible that some part of the lineup is less explored or even never scanned by the observer if he or she feels that the actual plot is found before even seriously scanning the whole lineup.
+%It may be possible that some part of the lineup is less explored or even never scanned by the observer if he or she feels that the actual plot is found before even seriously scanning the whole lineup.
 In those scenarios the location where the observer first start scanning is important.

### 2013-07-22 sociology_paper.tex
finished with section 2

@@
 
-\subsection{Sample of Null Plots} A lineup becomes difficult to evaluate if one of the null plots appears to be very similar to the actual plot.
+\subsection{Sample of Null Plots} In classical inference, the test statistic is compared with the full null distribution, to determine if it is extreme or not.
+In visual inference the actual plot is compared with plots of a finite number of samples from the null distribution.
+In the lineups used in the Turk experiments, there are 19 null plots upon which the comparison with the actual data plot is made.
+These null samples are random draws, and there is a chance that one or more null plots in the lineup may be similar or even more extreme than the actual data plot.
+The sample of null plots can affect the observer's decision.
+This is tested with the data from experiment 9, where the experiment was set up with lineups made from different null plots.
+
+%each lineup, there is a sample of draws from a null distribution.
+A lineup becomes difficult to evaluate if one of the null plots appears to be very similar to the actual plot.
 When null hypothesis is true it is a common scenario.
@@
 
-\subsection{Individual Skill or Ability} Each person is different from others in some way.
-For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.
+\subsection{Individual Skill or Ability} Each person may have different visual skill sets, and aptitude for reading statistical plots.
+This can be examined by having the same lineups evaluated by multiple subjects, so that it can be seen if some subjects consistently detect the actual data plot more often than others.
+ %For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.
 This influences the response of the observer.

### 2013-07-23 sociology_paper.tex
working on section 3

@@
 
-\subsection{Choice of Visual Test Statistic}
-
+\begin{itemize} \itemsep 0in
+%\subsection{Choice of Visual Test Statistic}  
+\item{\bf Choice of Visual Test Statistic:}  
 There are a variety of ways that data can be plotted and for plots to be enhanced by axes, legends or representations of models.
@@
 
-\subsection{Signal in the Data}
-
+%\subsection{Signal in the Data}
+\item{\bf Signal in the Data:}
 A visual test statistic is chosen for a particular task.
@@
 
-\subsection{Question Design} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
+%\subsection{Question Design} 
+\item{\bf Question Design:} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
 Visual inference supports discovery of srtucture, enabling the the unexpected to surprise us.
@@
 
-\subsection{Demographics}  The experiments collected data on age, gender, education level and geographic location.
+%\subsection{Demographics} 
+\item{\bf Demographics:}  The experiments collected data on age, gender, education level and geographic location.
 Each observer self-reported their gender, age in roughly five year intervals, education level as high school or less, some college courses, an undergraduate degree, some graduate courses or a graduate degree.
@@
 
-\subsection{Learning Trend} One might expect that as an observer evaluates more lineups they may become more skillful in their evaluations.
+%\subsection{Learning Trend} 
+\item{\bf Learning Trend:} One might expect that as an observer evaluates more lineups they may become more skillful in their evaluations.
 Each subject in the experiments 5, 6 and 7 evaluated a block of of ten lineups of the same type of data plot.
@@
 
-\subsection{Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 points is used.
+%\subsection{Location of Actual Data Plot in the Lineup}
+\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 points is used.
 A random number generator is used to determine the position where the actual data plot is placed in each lineup.
@@
 
-\subsection{Sample of Null Plots} In classical inference, the test statistic is compared with the full null distribution, to determine if it is extreme or not.
+%\subsection{Sample of Null Plots} 
+\item{\bf Sample of Null Plots:} In classical inference, the test statistic is compared with the full null distribution, to determine if it is extreme or not.
 In visual inference the actual plot is compared with plots of a finite number of samples from the null distribution.
@@
 
-\subsection{Individual Skill or Ability} Each person may have different visual skill sets, and aptitude for reading statistical plots.
+%\subsection{Individual Skill or Ability} 
+\item{\bf Individual Skill or Ability:} Each person may have different visual skill sets, and aptitude for reading statistical plots.
 This can be examined by having the same lineups evaluated by multiple subjects, so that it can be seen if some subjects consistently detect the actual data plot more often than others.
@@
 Also subject specific variation in the power of visual test is observed in \cite{majumder:2013}.
+\end{itemize}
 
 \section{Experimental Methods}\label{sec:exp_design}
-While evaluating a lineup, the performance of the observer may depend on the factors described in Section \ref{sec:factor_performance}.
-Some of these factors such as signal in the data and individual abilities were studied in \cite{majumder:2013}.
+
+%While evaluating a lineup, the performance of the observer may depend on the factors described in Section \ref{sec:factor_performance}. 
+Two of the factors, signal in the data and individual abilities, were studied in \cite{majumder:2013}.
+The choice of visual test statistic was examined in \cite{heike:2012}.
+In each of these analyses demographic factors were given a cursory glance, to ensure that they did not have large effects on the results.
+The design of experiments 5, 6 and 7 enables the examination of learning trend, which is studied in this paper.
+Experiment 9 was a real test case for visual statistical inference, and in order to understand the significance of the structure in the genomic data, multiple lineups were made in which location of the actual data plot, and the sample of nulls, were randomized.
+This enables the assessment of the effect of these factors on the results.
+This section describes the experimental methods used to examine the effects of demography, placement of the actual data plot, sample of null plots and the existence of a learning trend.
+
+% but for the most part data collected to examine demographic effects .
 In this paper we intend to study effect of the factors such as human demographics, learning trend, actual plot locations and selection of null plots.
@@
 
-\subsection{Demographic Factors}  In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table \ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
+\subsection{Demographic Factors}  %In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table \ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
+
+All the ten experiments shown in Table \ref{tbl:visual_stat}, had demographic information collected from subjects, as follows:
 
 \begin{enumerate} \leftmargin 5cm  \itemsep 0in
-\item Age group
-\item Gender
-\item Academic education level
-\item Geographical location
+\item Age group, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, 51-55, 56-60, above 60.
+\item Gender, male of female.
+\item Education level, with levels being high school or less, some under grad course, under graduate degree, some graduate courses, and graduate degree.
+\item Geographical location, collected from the IP address of the participants' computer, as latitude, longitude, city and country. 
 \end{enumerate}
 
-Instead of collecting exact age, 9 levels of age were collected.
+%Instead of collecting exact age, 9 levels of age were collected.
 They are 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, 51-55, 56-60, above 60.
@@
 
-The feedback provided by each observer on a lineup is a binary response variable.
-Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$.
-$Y_{ij} =1$ if the response is correct otherwise $Y_{ij} =0$.
-Let $\pi_{ij}=  E(Y_{ij})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$.
-We model this in a generalized mixed effects model of the form
+Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the response is correct otherwise $Y_{ij} =0$.
+The factors are examined in association with the observer's response using a mixed effects linear model:
 \begin{equation} \label{eqn:demographic_response}
@@
 \end{equation}
-where $\mu$ is an overall average probability for picking out the data plot from a lineup.
-$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender category $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$.
-$\ell_j$ is a random intercept predicting lineup difficulty level and we assume  $\ell_j \sim N(0, \sigma_\ell^2)$.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-
-The effect of demographic factors can be observed as time taken for each evaluation by an observer.
-Suppose $Z_{ij}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$.
+where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ correctly picks the data panel from lineup $j$, %We model this in a generalized mixed effects model of the form
+%\begin{equation} \label{eqn:demographic_response}
+%g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
+%\end{equation}
+ $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$.
+The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume  $\ell_j \sim N(0, \sigma_\ell^2)$, and $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+
+In addition, a model is constructed with time taken as the response variable.
+Let $Z_{ij}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$.
 Let $\mu_{ij}=  E(Z_{ij})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$.
@@
 \end{equation}
-where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
-$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model \eqref{eqn:demographic_response}.
-$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model \eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
 

### 2013-07-23 sociology_paper.tex
section 3.3

@@
 \end{equation}
-where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model \eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model \eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
+
+(*** No results are given for this model, are they????)
 
@@
 Each subject was shown a total of 10 lineups randomly selected from a pool of lineups.
-The lineups are not necessarily with the same difficulty levels.
-But the order of lineups were randomized.
+The lineups are not necessarily with the same difficulty levels, but the order of lineups was randomized.
 The responses of the lineups were recorded by attempt 1 through 10.
 Attempt 1 means that the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
-The goal is to estimate whether performance of the observer improves from attempt 1 to attempt 10.
+The goal is to estimate whether performance of the observer improves, or changes, from attempt 1 to attempt 10.
+
+It should be noted that we are examining the observer's performance, when we model response as correct or incorrect, but this is not the goal of visual inference.
+Visual inference is constructed to measure the significance of structure discovered in data.
+It is expected that some observers will be more skilled at reading data plots, and hence, more readily detect the plot that is different.
+It is also expected that as observer's gain experience in evaluating lineups that they become more proficient in reading data plots, particularly if feedback is given on whether the actual data plot was chosen or not.
+Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups, so ``correct'' is more of a convenient rather than ideal word to use for this analysis.
+So in this context, correct, or not, is used as a response to examine individual differences.
 
@@
 
-Multiple lineups were sequentially shown to each observer for evaluation.
-Suppose $Y_{ijk}$ denotes the response from observer $i$ on a  lineup $j$ in his/her $k^{th}$ attempt.
-$Y_{ijk}=1$ if the response is correct otherwise  $Y_{ijk}=0$.
-Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in his/her $k^{th}$ attempt.
-We model this in a generalized mixed effects model of the form
+%Multiple lineups were sequentially shown to each observer for evaluation. 
+Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k^{th}$ evaluation attempt, where $Y_{ijk}=1$ if the response is correct otherwise  $Y_{ijk}=0$.
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in their $k^{th}$ attempt.
+Learning trend is assessed using a generalized mixed effects model of the form
 \begin{equation} \label{eqn:trend_response}
@@
 \end{equation}
-where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k^{th}$ attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$.
-$u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$. $u_i$ is a random intercept, describing a basic subject-specific ability. We assume $u_i \sim N(0, \sigma_u^2)$. 
-$a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts. We assume $a_i \sim N(0, \sigma_a^2)$. 
-For $\ell_j$ we again assume a normal distribution, $N(0, \sigma_\ell^2)$.
-$\ell_j$ is a random intercept predicting lineup difficulty level.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ .
-
-The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k^{th}$ attempt by a single observer as
+where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k^{th}$ attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
+The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
+The term $a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts, where $a_i \sim N(0, \sigma_a^2)$. 
+For $\ell_j$ a normal distribution, $N(0, \sigma_\ell^2)$, is assumed, and $\ell_j$ is a random intercept predicting lineup difficulty level.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+The inverse link function, $g^{-1}(.)$, from equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k^{th}$ attempt by a single observer as
 \begin{equation} \label{eqn:trend_power}
@@
 
-The learning of each observer over a course of $K$ evaluations may be observed as the improvement of the time taken to evaluate a lineup in the later attempts.
-Suppose $Z_{ijk}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k^{th}$ attempt.
+When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k^{th}$ attempt.
 Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k^{th}$ attempt.
@@
 \end{equation}
-where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
-$\alpha$ is the average change in log time taken for each additional attempt.
-$\alpha_1$ is an offset in log time taken for the first attempt.
-All other effects are random effects: as before, $u_i$ is a subject-specific intercept representing  individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$.
-$a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$.
-We assume $a_i \sim N(0, \sigma_a^2)$.
-$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$ is the average change in log time taken for each additional attempt,  $\alpha_1$ is an offset in log time taken for the first attempt.
+All other effects are random effects: $u_i$ is a subject-specific intercept representing individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$, $a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$, with $a_i \sim N(0, \sigma_a^2)$, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
 Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $kth$ attempt by a single observer as 
@@
 
+*** Doesn't this need to go in a place so that it is obvious that all models described are fit with these routines, at the end of the section perhaps??
 To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
 To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
+
+*** These models treat attempt as a linear effect, so that only linear increase or decrease is modelled *** something like this.
+
+*** Also, you need to say something about examining the residuals, which is what is done in several of the parts of the results section.
 

### 2013-07-23 sociology_paper.tex
some typos

@@
 \newcommand{\dc}[1]{{\color{green} #1}}
+\newcommand{\mm}[1]{{\color{blue} #1}}
 
@@
 \item Age group, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, 51-55, 56-60, above 60.
-\item Gender, male of female.
+\item Gender, male or female.
 \item Education level, with levels being high school or less, some under grad course, under graduate degree, some graduate courses, and graduate degree.
@@
 
-(*** No results are given for this model, are they????)
+(\dc{*** No results are given for this model, are they????}  \mm{results are given in table \ref{tbl:model_result_demographics}})
 
@@
 
-*** Doesn't this need to go in a place so that it is obvious that all models described are fit with these routines, at the end of the section perhaps??
-To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
+\dc{*** Doesn't this need to go in a place so that it is obvious that all models described are fit with these routines, at the end of the section perhaps??} To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
 To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
-
+\dc{
 *** These models treat attempt as a linear effect, so that only linear increase or decrease is modelled *** something like this.
 
-*** Also, you need to say something about examining the residuals, which is what is done in several of the parts of the results section.
+*** Also, you need to say something about examining the residuals, which is what is done in several of the parts of the results section.}
 

### 2013-07-23 sociology_paper.tex
section 3.3

--- 2013-07-23 sociology_paper.tex
@@
 
-\subsection{Location Effect} \label{sec:location_design} Experiment 9 shown in Table~\ref{tbl:visual_stat} is designed to to assess the location effect of a data plot in a lineup.
-It is set up based on the gene expression data \citep{atwood:2013} with two groups.
+\subsection{Location Effect} \label{sec:location_design} Experiment 9 was designed to collect data to assess the location effect of a actual data plot in a lineup.
+The data used, documented in \citep{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
+Two factors were of primary interest: genotype and the interaction between genotype (call them RPA and EV) and iron sufficiency.
+Three replicates of all but one treatment were included, with the remaining treatment only having two replicates due one replicate being removed to a data quality concern.
+There are on the order of *** genes examined for significant expression.
+In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
+This overall significance is studied by visual inference.
+Full results are documented in \cite{tengfei:2013}.
+
+For the study, the routine processing of the data was conducted to find (1) the gene that exhibited the most significance difference in expression between genotypes, (2) the gene that exhibited the most different expression on RPA, but insignificant expression on EV.
+Two different types of plots would be made to examine this.
+For (1) side-by-side dotplots of the expression levels for the one gene by genotype were created.
+For (2) an interaction plot, with iron sufficiency represented on the horizontal axis, expression vertically, and genotype by color.
+Colored lines connect the means for the two genotypes over iron treatment.
+The null plots were created by randomizing the experimental design, and re-doing everything else as done with the actual data.
+Any genes found under these circumstances had expression differences that can only be considered to by consistent with random differences.
+
+Multiple lineups were produced.
+The actual plot was randomly placed in five different locations in a lineup of size 20.
+The locations were  1, 8, 12, 17, 20 for genotype effect (1) and 2, 9, 12, 16, 20 for interaction effect (2).
+Five different sets of null plots were used to produce 5 lineups for each location position, creating a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
+Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to filter observers.
+
+%It is set up based on the gene expression data \citep{atwood:2013} with two groups.
 One is the Genotype and the other the Empty Vector (EV).
@@
 
-For Genotype main effect side by side dot plots of two groups are used as the visual test statistics.
+%For Genotype main effect side by side dot plots of two groups are used as the visual test statistics.
 For interaction effect the same plot type is used except that the means of iron sufficiency or insufficiency were connected by lines colored by the groups.
@@
 
-Each observer saw three lineups, first an Interaction lineup then a Genotype lineup.
+%Each observer saw three lineups, first an Interaction lineup then a Genotype lineup.
 Finally a test lineup was shown to screen out unusual evaluation.
@@
 
-As per the design of experiment for location effect, the actual data plot is same for each of the null sets of plot.
-Thus the response data of this experiment constitute a multivariate response.
-To examine if the difference in proportion correct among the locations is statistically significant we fit a one way multi variate analysis of variance (MANOVA) model to the data.
-
-Suppose $\mathbf{Y}=(Y_1,Y_2, ...
-, Y_p)^\top$ be a vector of random variable with dimension $p$, the total number of null sets.
-let $\mathbf{Y}_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., I$.
-We fit the following MANOVA model
+%As per the design of experiment for location effect, the actual data plot is same for each of the null sets of plot. Thus the response data of this experiment constitute a multivariate response. 
+To examine if the difference in proportion correct among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
+Let $\mathbf{Y}=(Y_1,Y_2, ...
+, Y_p)^\top$ be a vector of random variables with dimension $p (=5)$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., I(=5)$.
+Because the same data plot is shown in each lineup, it is assumed that there could be some association between the responses for each null set, which suggests the MANOVA model rather than a univariate ANOVA.
+The MANOVA model
 \begin{equation}\label{manova}
@@
 \end{equation}
-where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$.
+where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$, tests for significant difference between the means.
+
+*** Why MANOVA rather than ANOVA? Explain this better here.
 

### 2013-07-23 R/sociology_paper.R
Figures updated

@@ -376,12 +376,12 @@ qplot(difficulty, prop_gd-prop_hs, data=ddat, geom="line")
 with(ddat, max(prop_gd-prop_hs))
 
 mddat <- melt(ddat,id=c("difficulty"))
-qplot(difficulty, value, data=mddat, color=variable, geom="line")+
+qplot(difficulty, value, data=mddat, linetype=variable, geom="line")+
   xlab("Lineup difficulty") + ylab("Proportion of correct response") +
   scale_x_continuous(breaks=c(-6,-2*2.293,-2.293,0,2.293,2*2.293,6), expand = c(0.02,0),
                      labels=c("Difficult",expression(-2*sigma[l],-sigma[l],0,sigma[l], 
                                                      2*sigma[l],"Easy"))) +
-  scale_color_discrete(name="Education",
+  scale_linetype_discrete(name="Education",
                       labels=c("High school", "Grad. degree"))
 
 ggsave("../images/practical_impact_graduate.pdf", width=6.5, height=4.5)
@@ -397,11 +397,18 @@ ddat.all <- data.frame(difficulty=diff, high.school = px(hs_xb),u.grad.course = 
                    u.grad.degree=px(ud_xb), grad.course=px(gc_xb), grad.degree=px(gd_xb))
 mddat.all <- melt(ddat.all,id=c("difficulty"))
 
-qplot(variable,value, color=factor(difficulty), data=mddat.all, size=I(3.5)) +
+qplot(variable,value, shape=factor(difficulty), geom=c("point"),data=mddat.all, size=I(3.5)) +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   ylab("Proportion of correct responses") + xlab("Education")+
-  scale_color_discrete(name="Lineup difficulty", 
+  scale_shape_discrete(name="Lineup difficulty", 
                        labels=c("Difficult", "Medium","Easy"))
+
+qplot(variable,value, shape=factor(difficulty), geom=c("point"),data=mddat.all, size=I(3.5)) +
+  ylab("Proportion of correct responses") + xlab("Education")+
+  scale_x_discrete(labels=c("hs","uc","ud","gc","gd"))+
+  scale_shape_discrete(name="Lineup difficulty", 
+                       labels=c("Difficult", "Medium","Easy"))
+
 
 ggsave("../images/practical_impact_degree.pdf", width=5.5, height=4.5)
 

### 2013-07-23 sociology_paper.tex
Figures updated
figure updated

@@
 For two standard deviations it further reduces to 0.3\%.
-Thus although there is statstically significant difference in proportion correct for some demographic factors, these are not practically significant differences.
+Thus although there is statistically significant difference in proportion correct for some demographic factors, these are not practically significant differences.
 Figure~\ref{fig:practical_impact_graduate} illustrates this example showing fitted models for a US 18-25 female with either a high school education or a graduate degree.
@@
 %   \includegraphics[width=6.5in]{learning_trend.pdf} 
-    \includegraphics[width=3.0in]{practical_impact_graduate.pdf} 
-    \includegraphics[width=3.0in]{practical_impact_degree.pdf} 
+    \includegraphics[width=3.2in]{practical_impact_graduate.pdf} 
+    \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
    \caption{Proportion of correct responses due to graduate degree as compared to high school degree for an 18-25 year old female in the United States.

### 2013-07-23 R/sociology_paper.R

@@ -403,14 +403,15 @@ qplot(variable,value, shape=factor(difficulty), geom=c("point"),data=mddat.all, 
   scale_shape_discrete(name="Lineup difficulty", 
                        labels=c("Difficult", "Medium","Easy"))
 
-qplot(variable,value, shape=factor(difficulty), geom=c("point"),data=mddat.all, size=I(3.5)) +
+qplot(variable,value, shape=factor(difficulty),data=mddat.all, size=I(3.5)) +
+  geom_line(aes(variable,value, group=factor(difficulty)))+
   ylab("Proportion of correct responses") + xlab("Education")+
-  scale_x_discrete(labels=c("hs","uc","ud","gc","gd"))+
+  scale_x_discrete(labels=c("h.school","u.gd.course","u.grad.deg","grad.course","grad.deg"))+
   scale_shape_discrete(name="Lineup difficulty", 
                        labels=c("Difficult", "Medium","Easy"))
 
 
-ggsave("../images/practical_impact_degree.pdf", width=5.5, height=4.5)
+ggsave("../images/practical_impact_degree.pdf", width=6.5, height=4.5)
 
 
 # Time of the day when Mturk worker works

### 2013-07-24 R/sociology_paper.R
Manova test = Wilks used

@@ -382,7 +382,7 @@ qplot(difficulty, value, data=mddat, linetype=variable, geom="line")+
                      labels=c("Difficult",expression(-2*sigma[l],-sigma[l],0,sigma[l], 
                                                      2*sigma[l],"Easy"))) +
   scale_linetype_discrete(name="Education",
-                      labels=c("High school", "Grad. degree"))
+                      labels=c("High school", "Grad. degree")) 
 
 ggsave("../images/practical_impact_graduate.pdf", width=6.5, height=4.5)
 
@@ -408,7 +408,7 @@ qplot(variable,value, shape=factor(difficulty),data=mddat.all, size=I(3.5)) +
   ylab("Proportion of correct responses") + xlab("Education")+
   scale_x_discrete(labels=c("h.school","u.gd.course","u.grad.deg","grad.course","grad.deg"))+
   scale_shape_discrete(name="Lineup difficulty", 
-                       labels=c("Difficult", "Medium","Easy"))
+                       labels=c("Difficult", "Medium","Easy")) 
 
 
 ggsave("../images/practical_impact_degree.pdf", width=6.5, height=4.5)
@@ -854,10 +854,10 @@ model.dat <- dcast(df,plot_type+plot_location+replicates~nulls, value.var="respo
 model <- as.formula(cbind(null_1,null_2,null_3,null_4,null_5)~factor(plot_location))
 int.dat <- subset(model.dat, plot_type=="Interaction")
 fit1 <- manova(model, data=int.dat )
-summary(fit1)
+summary(fit1, test="W")
 
 fit2 <- manova(model, data=subset(model.dat, plot_type=="Genotype") )
-summary(fit2)
+summary(fit2, test="W")
 
 
 int.dat$locs <- "out"

### 2013-07-24 sociology_paper.tex
Manova test = Wilks used
reference updated

@@
 Three replicates of all but one treatment were included, with the remaining treatment only having two replicates due one replicate being removed to a data quality concern.
-There are on the order of *** genes examined for significant expression.
+There are on the order of \dc{***} genes examined for significant expression.
 In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
 This overall significance is studied by visual inference.
-Full results are documented in \cite{tengfei:2013}.
+Full results are documented in \citep{atwood:2013}.
 
@@
 
-*** Why MANOVA rather than ANOVA? Explain this better here.
+\dc{*** Why MANOVA rather than ANOVA? Explain this better here.}
 
@@
 
-\subsubsection{Sample of lineup}
-\subsubsection{Sample of Null Plots}
-
-\subsubsection{Location Effect}
-
 A total of 111 subjects were recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
@@
 \begin{table}[hbtp]
-\caption{The results obtained by fitting MANOVA Model \eqref{manova}.}
+\caption{The summary of the results obtained by fitting MANOVA Model \eqref{manova} using Wilks test.}
 \begin{center}
@@
  \cline{5-7}
- Effect & DF & Pillai & Approx. F & Numerator& Denumerator &Residual & $p$-value\\
-  \hline
-  Interaction& 3&1.4783 &0.7772&15&12 & 6 & 0.6821 \\ 
-  Genotype &4&1.7796 &1.1221&20&28 & 8 &  0.3824 \\ 
+ Effect & DF & Wilks & Approx. F & Numerator& Denumerator &Residual & $p$-value\\
+  \hline
+  Interaction& 3&0.07986 & 0.592 &15&5.92 & 6 & 0.8082 \\ 
+  Genotype & 4&0.01689 & 1.722 & 20&14.22 & 8 & 0.1488 \\ 
    \hline

### 2013-07-24 sociology_paper.tex
typo fixed

@@
 Model \eqref{eqn:demographic_response} is fitted to the data with fixed effect factors such as age, country, education and gender.
-To estimate the overall factor main effect, a reduced model is fitted removing that factor from Model \eqref{eqn:demographic_response}.
+To estimate the overall factor main effect, a reduced model is fitted removing that factor from model \eqref{eqn:demographic_response}.
 Analysis of variance (ANOVA)  results shown in Table~\ref{tbl:anova_factor} suggest that all the factor variables are significantly different in describing the the probability of correct response except gender.
 Gender does not show any significant difference in performance.
-Similarly, ANOVA results are obtained by fitting Model \eqref{eqn:demographic_time} to the data of log time taken and are shown in Table~\ref{tbl:anova_factor} as well.
+Similarly, ANOVA results are obtained by fitting model \eqref{eqn:demographic_time} to the data of log time taken and are shown in Table~\ref{tbl:anova_factor} as well.
 All the factor variables are significantly different for log time taken including gender.
@@
 
-To further illustrate how the individual factor levels differ from each other  the detailed results of Models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
+To further illustrate how the individual factor levels differ from each other  the detailed results of models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
 For parameter estimation the first factor levels shown in the Table are  used as the base line.
@@
 \centering
-\caption{Parameter estimates of Models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and probability of correct lineup evaluations respectively.
+\caption{Parameter estimates of models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and probability of correct lineup evaluations respectively.
 For time taken all the demographic factors are significant.
@@
 For gender no difference in performance is observed.
-Lineup variability is estimated to be very large for Model \eqref{eqn:demographic_response}.}
+Lineup variability is estimated to be very large for model \eqref{eqn:demographic_response}.}
 \scalebox{0.86}{
@@
   \hline
-Demographic&& \multicolumn{4}{c} {Model \eqref{eqn:demographic_time} Log Time} & &\multicolumn{4}{c} {Model \eqref{eqn:demographic_response} Proportion Correct} \\
+Demographic&& \multicolumn{4}{c} {Log Time (model \ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Proportion Correct (model \ref{eqn:demographic_response})} \\
 
@@
 
-Notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in Model \eqref{eqn:demographic_response}.
+Notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
 This indicates that the major and most important factor affecting the probability of correct response is  lineup difficulty.
@@
 
-\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiment 5, 6, and 7 separately.
-As an alternative to Model \eqref{eqn:trend_response} we examined a reduced model with attempt as a continuous covariate.
+\subsection{Learning Trend} models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiment 5, 6, and 7 separately.
+As an alternative to model \eqref{eqn:trend_response} we examined a reduced model with attempt as a continuous covariate.
 They did not appear to be significantly different with a $p$-value of 0.856.
-But we consider the bigger Model \eqref{eqn:trend_response} with attempt as a factor variable with 10 levels to examine the effect of each level.
+But we consider the bigger model \eqref{eqn:trend_response} with attempt as a factor variable with 10 levels to examine the effect of each level.
 Attempt 1 is considered to be base level while fitting the model.
 
-Table \ref{tbl:model_result_response} presents the parameter estimates and $p$-values of fixed effect estimates of Model \eqref{eqn:trend_response}.
+Table \ref{tbl:model_result_response} presents the parameter estimates and $p$-values of fixed effect estimates of model \eqref{eqn:trend_response}.
 The larger $p$-values suggest that none of the levels of attempt ($\alpha_2$ through $\alpha_{10}$) are significant at \%1 significance level.
@@
 \centering
-\caption{Parameter estimates of Model \eqref{eqn:trend_response} fitted for probability of correct lineup evaluation.
+\caption{Parameter estimates of model \eqref{eqn:trend_response} fitted for probability of correct lineup evaluation.
 None of the fixed factor effects of attempt ($\alpha_2$ through $\alpha_{10}$) are significantly different from the first attempt $\alpha_1$
@@
 
-To visualize how the performance in correct response improves over successive attempts, we fitted Model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
+To visualize how the performance in correct response improves over successive attempts, we fitted model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
 Least square regression lines were fitted through the subject specific residuals as shown in Figure~\ref{fig:learning_trend_response}.
@@
     \includegraphics[width=6.3in]{learning_trend_subject.pdf} 
-   \caption{ Least square lines fitted through the subject specific residual proportion correct obtained from Model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
+   \caption{ Least square lines fitted through the subject specific residual proportion correct obtained from model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
 Subject specific positive and negative slopes are observed.
@@
 
-Table \ref{tbl:model_result_time} presents the results of Model \eqref{eqn:trend_time}.
+Table \ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time}.
 The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments.
@@
 \centering
-\caption{Parameter estimates of Model \eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
+\caption{Parameter estimates of model \eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
 Both fixed effect parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.}
@@
 
-The lineup variance of experiment 7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from Model \eqref{eqn:trend_response}.
-But from Model \eqref{eqn:trend_time} it is estimated to be much smaller in all these experiments.
+The lineup variance of experiment 7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model \eqref{eqn:trend_response}.
+But from model \eqref{eqn:trend_time} it is estimated to be much smaller in all these experiments.
 This suggests that the harder lineups not necessarily took more time than easier lineups.
@@
 
-To visualize how the time taken reduces over the successive attempts, we fitted Model \eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
+To visualize how the time taken reduces over the successive attempts, we fitted model \eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
 Least square regression lines are fitted through the subject specific residuals.
@@
  \includegraphics[width=6.3in]{learning_trend_time_subject.pdf}    
-   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting Model \eqref{eqn:trend_time} without covariate attempt.
+   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting model \eqref{eqn:trend_time} without covariate attempt.
 Differences in subject specific slopes are observed.
@@
 
-We fitted Model \eqref{manova} to test if the mean performance vectors are similar for different locations.
+We fitted model \eqref{manova} to test if the mean performance vectors are similar for different locations.
 For this we use {\it anova()} function of {\it stats} package of \cite{R:2012}.
@@
 \begin{table}[hbtp]
-\caption{The summary of the results obtained by fitting MANOVA Model \eqref{manova} using Wilks test.}
+\caption{The summary of the results obtained by fitting MANOVA model \eqref{manova} using Wilks test.}
 \begin{center}
@@
 It does not show any differences whether the actual plot is inside or outer border of the lineup.
-We also fitted Model \eqref{manova} with two locations, inner and outer, as covariate and observed no significant differences.
+We also fitted model \eqref{manova} with two locations, inner and outer, as covariate and observed no significant differences.
 

### 2013-08-23 sociology_paper.tex
changed p value and discussion of distribution for visual inference

@@
 
-{\bf Keywords: \sf statistical graphics, non-parametric test, cognitive phycology, data visualization, exploratory data analysis, data mining, visual analytics.} 
+{\bf Keywords: \sf statistical graphics, non-parametric test, cognitive \hh{psychology}, data visualization, exploratory data analysis, data mining, visual analytics.} 
 \end {abstract}
@@
 
-Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are rendered from data generated under a null model.
+Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are showing from data generated under a null model.
 Which one of the 20 plots is the most different from the others?
-When asked this question, 12 of 72 observers picked the same plot (with number equal to the result of $3^2+4$), with reasons given being 'asymmetry' (36\%), 'trend' (26\%) or `outliers' (16\%).
-The corresponding $p$-value is 0.00023, indicating sufficient evidence to reject the null hypothesis.
-
-What does this mean, though?
-For that, we need to know the context of the data and we need to have more information about the generation of the null plots.
-For that, we need to know the context of the data and we need to have more information about the generation of the null plots.
-This example investigates the results from the 2012 US presidential election in comparison to the poll results just prior to the election.
+When asked this question, 12 \hh{out} of 72 observers picked the  plot \hh{of the observed data} (located in spot number $3^2+4$ in the lineup).
+The corresponding $p$-value is \hh{0.0077}, indicating sufficient evidence to reject the null hypothesis.
+% , with reasons given being 'asymmetry' (36\%), 'trend' (26\%) or `outliers' (16\%).
+
+\hh{How do we interpret this finding, though?} For that, we need to know the context of the data and we need to have more information about the \hh{how null plots were generated}. 
+This particular example investigates the results from the 2012 US presidential election in comparison to the poll results just prior to the election. 
 (Although this example is more simplistic than most of the tests conducted to date, it will serve the purpose of illustrating the lineup protocol.) The data is looking at the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
@@
 A lineup can be evaluated by a single person or multiple observers.
-A binomial distribution is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
+A distribution \hh{similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario,} is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
 Observers should not be aware of the data that constitutes a lineup, and should not have seen the actual data plot before seeing the lineup.
@@
 \item{\bf Question Design:} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
-Visual inference supports discovery of srtucture, enabling the the unexpected to surprise us.
+Visual inference supports discovery of structure, enabling the the unexpected to surprise us.
 This is tightly connected to the task that the human observer is asked to perform.
-For most purposes the oberver is asked a very general question, such as to pick the plot that is most different from the others, and explain the reasons why they see it as different.
+For most purposes the observer is asked a very general question, such as to pick the plot that is most different from the others, and explain the reasons why they see it as different.
 By being general, all possible discoverable structures are included in the significance calculations.

### 2013-09-04 sociology_paper.tex
reference for expectation error added (sensory evaluation techniques)

@@
 A distribution \hh{similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario,} is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
-Observers should not be aware of the data that constitutes a lineup, and should not have seen the actual data plot before seeing the lineup.
+In order to avoid expectation errors \citep{meilgaard}, observers should not be aware of the data that constitutes a lineup, and should not have seen the actual data plot before seeing the lineup.
 This is the reason that in the election example, above, the scenario was explained after the lineup question, in the text.

### 2014-01-07 sociology_paper.tex
revising language

@@
 
-There have been ten experiments conducted using Amazon's Mechanical Turk~\citep{turk} that are being used to evaluate the effects of observers' demographic factors on the inference.
+There have been ten experiments conducted using Amazon's Mechanical Turk~\citep{turk} (MTurk) that are being used to evaluate the effects of observers' demographic factors on the inference.
 Table \ref{tbl:visual_stat} summarizes these experiments.
@@
 where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model \eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-
-(\dc{*** No results are given for this model, are they????}  \mm{results are given in table \ref{tbl:model_result_demographics}})
 
@@
 
-\dc{*** Doesn't this need to go in a place so that it is obvious that all models described are fit with these routines, at the end of the section perhaps??} To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
+To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
 To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
-\dc{
-*** These models treat attempt as a linear effect, so that only linear increase or decrease is modelled *** something like this.
-
-*** Also, you need to say something about examining the residuals, which is what is done in several of the parts of the results section.}
-
-\subsection{Location Effect} \label{sec:location_design} Experiment 9 was designed to collect data to assess the location effect of a actual data plot in a lineup.
-The data used, documented in \citep{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
-Two factors were of primary interest: genotype and the interaction between genotype (call them RPA and EV) and iron sufficiency.
-Three replicates of all but one treatment were included, with the remaining treatment only having two replicates due one replicate being removed to a data quality concern.
-There are on the order of \dc{***} genes examined for significant expression.
+
+\subsection{Location Effect} \label{sec:location_design} Experiment 9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could be assessed.
+The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
+Two factors were of primary interest: genotype (call them RPA and EV) and the interaction between genotype and iron condition (sufficient or insufficient).
+The $2\times2$ factorial design provides four treatment conditions.
+For each gene there were 11 data points, because each treatment contained two or three replicates.
+There are on the order of \dc{*** how many} genes examined for significant expression.
 In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
 This overall significance is studied by visual inference.
-Full results are documented in \citep{atwood:2013}.
 
@@
 Any genes found under these circumstances had expression differences that can only be considered to by consistent with random differences.
-
-Multiple lineups were produced.
+Full results are documented in \cite{tengfei:2013}.
+
+Multiple lineups of the data were produced.
 The actual plot was randomly placed in five different locations in a lineup of size 20.
@@
 Five different sets of null plots were used to produce 5 lineups for each location position, creating a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
-Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to filter observers.
+Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to help clean the data.
 
@@
 
-\dc{*** Why MANOVA rather than ANOVA? Explain this better here.}
-
-\subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through Amazon Mechanical Turk \citep{turk} or MTurk Web site.
-It is an online work place where people from around the world can perform some tasks and get paid.
+\subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through MTurk \citep{turk}.
+It is an online work place where people from around the world can perform tasks, that computers would have difficulty doing, and get paid.
 Usually tasks are very simple and no specialized training is required to do them.
-Being a human is the main requirement.
 Tasks are designed for anyone to do but some tasks may require some skills depending on the recruiters' need.
-Each task is usually planned to complete in a short time.
-Humans are still better than computers in performing these types of tasks.
-The the amount of money paid for each task is very small as well.
+Each task is usually planned to completed in a short time.
+The amount of money paid for each task is very small, on the order of minimum wage in the USA.
 
@@
 
-We designed and developed a web application which enables to display the lineups to the observers as per experimental need.
-The MTurk workers were redirected to the web site to evaluate lineups.
-The data were collected, stored automatically into a local database server.
-Demographic informations such as age group, gender and education levels were also collected.
+We designed and developed a web application which enables the display of lineups to the observers as per experimental need.
+The MTurk workers were redirected to this web application to complete their assigned tasks.
+The responses were collected, stored automatically into a local database server, along with demographic information, age group, gender and education level.
 The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
@@
 Most of the participants were from the United States and India.
-There were 76 other different countries from where we received feedbacks.
+There were 76 other different countries represented.
 This provides a diverse pool of participants.
-The diversity in not only in geographical locations of the participants but also in their gender, age groups and education levels as we see in Table \ref{tbl:demographics}.
-It is interesting that there were large number of female participants even though there were lot of people from developing countries.
+The diversity in not only in geographic but also in their gender, age group and education level as can be seen in Table \ref{tbl:demographics}.
+It is interesting that there were large number of female participants even though there were lot of participants worked from developing countries.
 
@@
 The rest of the 70 countries have less than 10 participants.
-The distribution of participants remains almost similar in all ten experiments with some small variations (Figure~\ref{fig:turker_location_experiment}).
-That may be due to the time of the experiment when it first started.
-For some experiments India got more participants than United States and for some experiments the numbers just got reversed (Figure~\ref{fig:participation_time}).
+The distribution of participants remains similar in all ten experiments.
+That may be due to the time of day when the experiment was made available.
+%For some experiments more people participated from India than the United States and for others numbers were reversed (Figure~\ref{fig:participation_time}).
 It is also observed in some experiments that number of  participants are similar no matter whether the experiment is run on day time or nigh time.
+(*** Took out reference to figure 11 in appendix.)
 
@@
   \hline
-  && \multicolumn{2}{c}{Participants}&Average&\\
+  && \multicolumn{2}{c}{Participants}&Average& Number of\\
   \cline{3-4}
- Factor & levels & Total &\%Factor & time & response \\ 
+ Factor & Levels & Total &\% & Time & Responses \\ 
   \hline
@@
 
-The largest number of participants are from age group 18 to 25 which is the youngest age group in the study.
-The majority of the people have ages between 18 to 35.
-Interestingly there are many participants from older age groups as well.
-Specially for united states almost all the age groups show uniform participations beyond age 30 (Figure~\ref{fig:demographic_info}).
-Notice that fewer people participated from India beyond age 40 compared to united states.
-Total number of participations from india and United States are almost same with United States having 107 more participants.
-Unlike USA Indian participants are mostly young people.
-
-In terms of participant's academic background, the largest group is graduate degree.
-A total of 902 participants have a graduate degree which is about 38.51\%.
-They are mostly from India as any university degree is considered a graduate degree unlike north America where graduate degree means Masters level education.
-Most of the USA participants are with an Undergraduate degree or at least have some undergraduate courses (Figure~\ref{fig:demographic_info}).
-
-The distribution of male and female participants are similar among all age groups except age 18-25 in India where fewer female participants were observed (Figure \ref{fig:gender_country_bar}).
-The distribution of education levels are also different across the countries for this age group.
-Most of the participants from India are below age 40 while in United States the distribution of participants are similar beyond age 40.
+The largest number of participants are from age group 18 to 25, with the majority between 18 to 35.
+Interestingly, there are many participants in the older age groups as well.
+The United States shows a more uniform participation beyond age 30 but Indian participants are primarily younger.
+
+In terms of academic background, the largest participant group is graduate degree, with a total of 902 participants, about 38.51\%.
+Examining the location in conjunction with the academic background reveals that there are more Indian participants that list themselves as having a graduate degree.
+In retrospect this was realized to be a language confusion, as any university degree is considered to be a graduate degree in India.
+Unlike the definition in North America where graduate degree means Masters level education.
+Most of the USA participants list themselves as having an undergraduate degree or at least have some undergraduate courses.
+
+The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where proportion of female participants is lower.
+The distribution of education levels also differ slightly across the countries for this age group.
+%Most of the participants from India are below age 40 while in United States the distribution of participants are similar beyond age 40.
 There were not many participants beyond age 50 and hence these age groups are merged to form one age group called above 50.
@@
 Each person evaluated at least 10 lineups except for experiment 9 where three lineups were evaluated by each person.
-A test plot was shown to each observer and the feedback received from that plot is used to examine the data quality or process payment.
-But the data received for the test lineup is not included in the analysis.
+A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
+Data from participants who got it wrong were removed, and then results for just this lineup were removed for all participants who responded correctly.
 In some cases the demographic information were not provided by the participants.
@@
 This resulted some missing demographic information.
+
+\dc{*** Di is up to here on paper}
 

### 2014-01-08 sociology_paper.tex
working on model results summary

@@
 This is the definition of the power of the test.
-When multiple observers evaluate a lineup, the proportion of correct response can be used to estimate the power.
+When multiple observers evaluate a lineup, the detection rate can be used to estimate the power.
 The ability of individual observers can vary, and examining this is the purpose of this paper.
+
+\dc{*** Am trying to use detection rate rather than correct response.
+This isharder for this paper because we are evaluating subject's abilities to some extent, but we should try to use this in our vis inference papers.}
 
@@
 In subsequent lineups being evaluated the type of plot is the same, level of difficulty will be different.
-It is possible that the observer becomes more skillful at recognizing the most different plot, either by correctly finding the actual data plot or more quickly reporting their choice.
+It is possible that the observer becomes more skillful at recognizing the most different plot, either by more often detecting the actual data plot or more quickly reporting their choice.
 These two ways of measuring learning trend are evaluated on data collected from experiments 5, 6 and 7.
@@
 
-Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the response is correct otherwise $Y_{ij} =0$.
+Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
 The factors are examined in association with the observer's response using a mixed effects linear model:
@@
 \end{equation}
-where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ correctly picks the data panel from lineup $j$, %We model this in a generalized mixed effects model of the form
+where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ picks the actual data plot from lineup $j$, %We model this in a generalized mixed effects model of the form
 %\begin{equation} \label{eqn:demographic_response}
@@
 
-It should be noted that we are examining the observer's performance, when we model response as correct or incorrect, but this is not the goal of visual inference.
+It should be noted that we are examining the observer's performance, when we model response as detected or not, but this is not the goal of visual inference.
 Visual inference is constructed to measure the significance of structure discovered in data.
@@
 It is also expected that as observer's gain experience in evaluating lineups that they become more proficient in reading data plots, particularly if feedback is given on whether the actual data plot was chosen or not.
-Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups, so ``correct'' is more of a convenient rather than ideal word to use for this analysis.
-So in this context, correct, or not, is used as a response to examine individual differences.
+Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups.%, so ``correct'' is more of a convenient rather than ideal word to use for this analysis.
+So in this context, detected, or not, is used as a response to examine individual differences.
 
@@
 %Multiple lineups were sequentially shown to each observer for evaluation. 
-Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k^{th}$ evaluation attempt, where $Y_{ijk}=1$ if the response is correct otherwise  $Y_{ijk}=0$.
-Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in their $k^{th}$ attempt.
+Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k^{th}$ evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the actual data plot from lineup $j$ in their $k^{th}$ attempt.
 Learning trend is assessed using a generalized mixed effects model of the form
@@
 
-\subsection{Location Effect} \label{sec:location_design} Experiment 9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could be assessed.
+\subsection{Location Effect} \label{sec:location_design} Experiment 9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
 The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
@@
 %As per the design of experiment for location effect, the actual data plot is same for each of the null sets of plot. Thus the response data of this experiment constitute a multivariate response. 
-To examine if the difference in proportion correct among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
+To examine if the difference in detection rate among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
 Let $\mathbf{Y}=(Y_1,Y_2, ...
@@
 
-\dc{*** Di is up to here on paper}
-
-\subsection{Demographic Factors} Proportions of correct responses and natural logarithms of average time taken to evaluate  each lineup are computed  for different demographic factor levels.
-Their distributions are shown using boxplots in Figure \ref{fig:demographic_effect}.
-Averages of these distributions are represented by dots inside the boxplots.
+\subsection{Demographic Factors}
+
+\dc{*** It seems repetitive to start with a description of what we learn from the plot, so start with the model summary and use the plot to support the model findings.}
+
+To explore the significance of the demographic factors model \eqref{eqn:demographic_response} is fit to the data, with age, country, education and gender as fixed effects.
+To estimate the significance of the factors, reduced models are fit with that factor removed from the model.
+Table ~\ref{tbl:anova_factor} summarizes the results.
+All of the demographic factors are significantly different in describing the the detection rate, except for gender.
+With respect to time taken, all of the demographic factors are significant.
+
+\begin{table}[htbp]
+\centering
+\caption{Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
+Gender does not have a significant effect on detection rate, but does on time to respond.
+All factors significantly affect time to respond.}
+\scalebox{.9}{
+\begin{tabular}{rlrrrrrr}
+  \hline
+Model & & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
+  \hline
+Proportion Correct &Full & 23822.00 &23943.00 &-11896.00  & &  &  \\ 
+& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
+  &Age & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
+  &Country  & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
+  &Education  & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
+  &Gender  & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
+   \hline
+   Log Time &Full &  51904.00 &52034.00 & -25936.00 & &  &  \\ 
+& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
+&Age  & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
+ & Country  & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
+  &Education  & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
+  &Gender  & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
+\hline
+\end{tabular}
+}
+\label{tbl:anova_factor}
+\end{table}
+
+Parameter estimates from the model fits are shown in Table \ref{tbl:model_result_demographics}.
+The first level of each factor serves as the baseline for the model.
+The age group 36-40 significantly has a higher detection rate than the 18-25, and to a lesser extent this is also true for age groups 31-35 and above 50.
+Participants from India have a similar detection rate to those from the USA, but ones from the rest of the world had a significantly improved detection rate.
+Subjects who listed themselves as having a graduate degree had a significantly higher detection rate.
+For time to respond, all age groups were significantly slower than the 18-25 year olds.
+Participants from India were slower, and those from the rest of the world were faster than Americans.
+A higher degree corresponded to longer time, and males were slower than females!
+
+\begin{table}[htbp]
+\centering
+\caption{Parameter estimates of models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively.
+For time taken all the demographic factors are significant.
+For detection rate, age group 36-40, rest of the world and graduate degree are significantly different.
+For gender no difference in performance is observed.
+Lineup variability is estimated to be very large for model \eqref{eqn:demographic_response}.}
+\scalebox{0.86}{
+\begin{tabular}{rlrrrrrrrrrr}
+  \hline
+Demographic&& \multicolumn{4}{c} {Log Time (model \ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model \ref{eqn:demographic_response})} \\
+
+\cline{3-6} \cline{8-11} 
+Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
+  \hline
+\multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
+&$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
+\hline
+Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
+ &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.206 \\ 
+  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.038 \\ 
+  &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
+  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.050 \\ 
+  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.143 \\ 
+  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.039 \\ 
+  \hline
+Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
+  &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
+  &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
+  \hline
+Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
+ &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
+  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.331 \\ 
+  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.157 \\ 
+  &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
+  \hline
+Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
+  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.133 \\ 
+  \hline
+\multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
+& lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
+ & Error($\sigma$) & 0.479 & 0.692 &  &  & & && &\\ 
+   \hline
+\end{tabular}
+}
+\label{tbl:model_result_demographics}
+\end{table}
+
+Figure \ref{fig:demographic_effect} supports the model results.
+The average time taken to respond (natural log) and detection rate are computed for the different demographic factor levels, and displayed using boxplots.
+Data averages are represented by dots inside the boxplots.
+%There are some small differences across demographic groups.
 The youngest age group (18-25) took least average time to evaluate a lineup.
@@
 Some differences in means of various demographic factors are observed.
-Variability in proportion correct indicates large variability in lineup difficulties.}
+Variability in proportion correct indicates large variability in lineup difficulties.
+\dc{*** Vertical axis needs to be labelled ``detection rate'' for bottom row, and ``value'' needs to be removed.}}
    \label{fig:demographic_effect}
@@
 
-The distribution of time taken to evaluate a lineup is positively skewed.
+%The distribution of time taken to evaluate a lineup is positively skewed.
 But in log scale it appears to be symmetric for all the demographic factor variables as we see in Figure \ref{fig:demographic_effect}.
@@
 
-We observed some differences in median proportion of correct responses for different demographic factor levels.
-But the variability of proportion correct responses are huge as per design of the experiments since there were some very easy as well as extremely difficult lineups.
+Some differences in median detection for different demographic factor levels can be seen,  but the variability in responses is huge.
+This variability is primarily in variability in the lineups -- some very easy and some extremely difficult lineups.
 Some of the lineups  were not expected to be evaluated correctly as the null plots in those lineups were showing more signal than the actual data plot itself.
@@
 
-Model \eqref{eqn:demographic_response} is fitted to the data with fixed effect factors such as age, country, education and gender.
-To estimate the overall factor main effect, a reduced model is fitted removing that factor from model \eqref{eqn:demographic_response}.
-Analysis of variance (ANOVA)  results shown in Table~\ref{tbl:anova_factor} suggest that all the factor variables are significantly different in describing the the probability of correct response except gender.
-Gender does not show any significant difference in performance.
-Similarly, ANOVA results are obtained by fitting model \eqref{eqn:demographic_time} to the data of log time taken and are shown in Table~\ref{tbl:anova_factor} as well.
-All the factor variables are significantly different for log time taken including gender.
-
-\begin{table}[htbp]
-\centering
-\caption{Anova of full model with all the demographic factors vs reduced model with removing respective factor variable. Gender does not have any effect on probability of correct response.}
-\scalebox{.9}{
-\begin{tabular}{rlrrrrrr}
-  \hline
-Model & & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
-  \hline
-Proportion Correct &Full & 23822.00 &23943.00 &-11896.00  & &  &  \\ 
-& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
-  &Age & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
-  &Country  & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
-  &Education  & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
-  &Gender  & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
-   \hline
-   Log Time &Full &  51904.00 &52034.00 & -25936.00 & &  &  \\ 
-& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
-&Age  & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
- & Country  & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
-  &Education  & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
-  &Gender  & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
-\hline
-\end{tabular}
-}
-\label{tbl:anova_factor}
-\end{table}
-
 To further illustrate how the individual factor levels differ from each other  the detailed results of models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
@@
 But not all the levels are significant for the probability of correct responses.
-
-\begin{table}[htbp]
-\centering
-\caption{Parameter estimates of models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and probability of correct lineup evaluations respectively.
-For time taken all the demographic factors are significant.
-For probability of correct response age group 36-40, rest of the world and graduate degree are significantly different.
-For gender no difference in performance is observed.
-Lineup variability is estimated to be very large for model \eqref{eqn:demographic_response}.}
-\scalebox{0.86}{
-\begin{tabular}{rlrrrrrrrrrr}
-  \hline
-Demographic&& \multicolumn{4}{c} {Log Time (model \ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Proportion Correct (model \ref{eqn:demographic_response})} \\
-
-\cline{3-6} \cline{8-11} 
-Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
-  \hline
-\multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
-&$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
-\hline
-Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
- &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.206 \\ 
-  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.038 \\ 
-  &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
-  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.050 \\ 
-  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.143 \\ 
-  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.039 \\ 
-  \hline
-Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
-  &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
-  &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
-  \hline
-Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
- &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
-  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.331 \\ 
-  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.157 \\ 
-  &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
-  \hline
-Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
-  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.133 \\ 
-  \hline
-\multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
-& lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
- & Error($\sigma$) & 0.479 & 0.692 &  &  & & && &\\ 
-   \hline
-\end{tabular}
-}
-\label{tbl:model_result_demographics}
-\end{table}
 
@@
 
-Age group 36-40 is significantly different than the base line age group of 18-25 year olds.
+%Age group 36-40 is significantly different than the base line age group of 18-25 year olds.
 The other age levels are similar in explaining the probability of correct response.
@@
 
-The graduate degree holders are significantly different as well.
+%The graduate degree holders are significantly different as well.
 The positive parameter estimate indicates that they perform better and the probability of correct response is higher than other education levels.

### 2014-01-08 sociology_paper.tex
up to lineup design

@@
 & lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
- & Error($\sigma$) & 0.479 & 0.692 &  &  & & && &\\ 
+ & Error($\sigma$) & 0.479 & &  &  &  & 0.692&& &\\ 
    \hline
@@
 Figure \ref{fig:demographic_effect} supports the model results.
-The average time taken to respond (natural log) and detection rate are computed for the different demographic factor levels, and displayed using boxplots.
-Data averages are represented by dots inside the boxplots.
+The average time taken to respond (natural log) and detection rate are computed by lineup \dc{*** check this, were these calculated by lineup??} for the different demographic factor levels, and displayed using boxplots.
+Since the models use the mean values, rather than median, these are represented by a dot inside the boxplots.
 %There are some small differences across demographic groups.
@@
 Averages of log time taken by male and female participants look similar but there are some differences as we see in Table~\ref{tbl:demographics}.
+The means and medians for response time (top row) are essentially equal.
+There are some differences between the two in the detection rate plots (bottom row).
+The differences in these values for different factors, that was tested by the models, can be seen in the plots.
+However, particularly for detection rate, the variation is huge.
+A large component of the variation is difficulty of the lineup.
+In some lineups the actual data plot was distinctively different from the null plots, and we would expect that the detection rate would be close to 100\%.
+In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond.
+In Table \ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
+This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
+So although the demographic factors emerge from the model to be statistically significant, the practical significance is minimal.
+We illustrate this with the following example of graduate degree.
+
+%Some differences in median detection rate for different demographic factor levels can be seen,  but the variability in responses is huge.
+This variability is primarily in variability in the lineups -- some very easy and some extremely difficult lineups.
+Some of the lineups  were not expected to be evaluated correctly as the null plots in those lineups were showing more signal than the actual data plot itself.
+In those scenarios,  one may not reject the null hypothesis.
+We call these difficult lineups.
+Some of the lineups were so easy that data plot is detectable at a first glance.
+These were expected to be evaluated correctly 100\% of the time.
+This is why we see the boxes range from 0 to 1 in most of the factor levels.
+Unlike median, fewer differences are observed in the average proportion of correct responses for different factor levels.
 
@@
 
-Some differences in median detection for different demographic factor levels can be seen,  but the variability in responses is huge.
-This variability is primarily in variability in the lineups -- some very easy and some extremely difficult lineups.
-Some of the lineups  were not expected to be evaluated correctly as the null plots in those lineups were showing more signal than the actual data plot itself.
-In those scenarios,  one may not reject the null hypothesis.
-We call these difficult lineups.
-Some of the lineups were so easy that data plot is detectable at a first glance.
-These were expected to be evaluated correctly 100\% of the time.
-This is why we see the boxes range from 0 to 1 in most of the factor levels.
-Unlike median, fewer differences are observed in the average proportion of correct responses for different factor levels.
-
-To further illustrate how the individual factor levels differ from each other  the detailed results of models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
+%To further illustrate how the individual factor levels differ from each other  the detailed results of models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
 For parameter estimation the first factor levels shown in the Table are  used as the base line.
@@
 
-Notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
-This indicates that the major and most important factor affecting the probability of correct response is  lineup difficulty.
-This is also evident from the large variability in the proportion correct in Figure~\ref{fig:demographic_effect}.
-Because of this huge impact of lineup specific variance, the practical impact of other factor variables on the probability of correct response is in fact very small.
+%In Table \ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
+This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
+Because of the huge impact of lineup variance, the practical impact of other factor variables on the probability of correct response is in fact very small.
 We illustrate this with the following example of graduate degree.
 
-While some of the demographic factors are strongly statistically significant, the main source of variation in proportion correct is the lineup difficulty.
+While some of the demographic factors are strongly statistically significant, the main source of variation in detection rate is the lineup difficulty.
 For example, let's examine the effect of graduate degree.
-To see just how large the effect is, we examine the change in proportion correct for a (hypothetical) 18-25 year old female in the United States, with a graduate degree as compared to a high school degree, for an average difficulty lineup (random effect = 0).
+To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with a graduate degree as compared to a high school degree, for an average difficulty lineup (random effect = 0).
 Plugging in the relevant quantities to the fitted model gives a difference equal to:
@@
 
-\subsection{Learning Trend} models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiment 5, 6, and 7 separately.
-As an alternative to model \eqref{eqn:trend_response} we examined a reduced model with attempt as a continuous covariate.
-They did not appear to be significantly different with a $p$-value of 0.856.
-But we consider the bigger model \eqref{eqn:trend_response} with attempt as a factor variable with 10 levels to examine the effect of each level.
-Attempt 1 is considered to be base level while fitting the model.
-
-Table \ref{tbl:model_result_response} presents the parameter estimates and $p$-values of fixed effect estimates of model \eqref{eqn:trend_response}.
-The larger $p$-values suggest that none of the levels of attempt ($\alpha_2$ through $\alpha_{10}$) are significant at \%1 significance level.
+\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments 5, 6, and 7 separately.
+In this model attempt is fitted as a factor variable to allow for all possible non-linear learning trends.
+It should be noted that we also examined an alternative to model \eqref{eqn:trend_response} where attempt was linearly fitted as a continuous covariate, but this also was not significant.
+
+Table \ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model \eqref{eqn:trend_response} examining detection rate.
+There is some evidence of a learning curve only for experiment 6.
+There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.
+
+%The large $p$-values suggest that none of the levels of attempt ($\alpha_2$ through $\alpha_{10}$) are significant at \%1 significance level.
 Moreover, some of the estimates are positive and some are negative and they show up seemingly in random order suggesting later attempts not necessarily have improved.
@@
 \centering
-\caption{Parameter estimates of model \eqref{eqn:trend_response} fitted for probability of correct lineup evaluation.
-None of the fixed factor effects of attempt ($\alpha_2$ through $\alpha_{10}$) are significantly different from the first attempt $\alpha_1$
-at \%1 level in all three experiments 5, 6 and 7. For experiment 7 subject specific variation is very small on the other hand lineup variance is much higher compared to the other two experiments.}
+\caption{Parameter estimates of models \eqref{eqn:trend_response} fitted to data from three different experiments using detection rate as the response to assess learning trend.
+Attempt number is fitted as a factor to enable modeling any sort of non-linear learning trend.
+Only experiment 6 shows evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
 \scalebox{0.81}{
@@
 
-The variance of random slope for attempt ($\sigma^2_a$) is estimated to be very small compared to other random effects except for subject variability ($\sigma^2_u$) for experiment 7.
+%The variance of random slope for attempt ($\sigma^2_a$) is estimated to be very small compared to other random effects except for subject variability ($\sigma^2_u$) for experiment 7.
 This suggests that most of the variability for attempt is accounted for by the fixed effect factor attempt.
@@
 
-To visualize how the performance in correct response improves over successive attempts, we fitted model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
+To visualize how the detection rate effect over successive attempts, we fitted model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
 Least square regression lines were fitted through the subject specific residuals as shown in Figure~\ref{fig:learning_trend_response}.
+The averages of these residuals for each of the attempts are shown as dots.
 Two important features were observed; one is subject specific variability and the other is random slope with attempts which indicates subjects specific learning trend.
 Some subjects show improvement over time and some show the decrease in performance.
+Although, the model fit experiment 6 suggested a statistically significant learning curve the plots indicates that is minimal.
 
@@
 
-The averages of these residuals for each of the attempts are shown as dots in Figure~\ref{fig:learning_trend_response}.
+%The averages of these residuals for each of the attempts are shown as dots in Figure~\ref{fig:learning_trend_response}.
 Least square linear regression lines are fitted through the points for each of the experiments.
@@
 
-Table \ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time}.
+Table \ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time} where response time is examined with respect to attempt.
+Attempt is modelled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated.
+Interestingly, time to respond significantly decreases as number of attempts increases, for all three experiments.
 The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments.
@@
 
-The lineup variance of experiment 7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model \eqref{eqn:trend_response}.
+%The lineup variance of experiment 7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model \eqref{eqn:trend_response}.
 But from model \eqref{eqn:trend_time} it is estimated to be much smaller in all these experiments.
@@
 
-Model \eqref{eqn:trend_time} was compared to couple of other alternatives.
+%Model \eqref{eqn:trend_time} was compared to couple of other alternatives.
 A bigger model with attempt as a factor was considered but it was not significantly different with $p$ value 0.236.

### 2014-01-08 sociology_paper.tex
up to conclusions

@@
 Each subject evaluated two lineups;  one for Interaction effect and the other for Genotype.
-In total there were 222 feedbacks or responses on 50 lineups.
+In total there were 222 responses on 50 lineups.
 The data on the test lineup were excluded from the analysis.
 
-The proportion of correct responses for each data plot location is shown in Figure \ref{fig:location_effect} colored by null sets.
-We observe some variability of performance for different null sets even though same data plot was used for all these null sets.
-This may happen when for some set of null plots, one null plot appears to be very similar to the actual plot.
+The detection rate for each data plot location is shown in Figure \ref{fig:location_effect} colored by null sets.
+The overall average detection rate for each location are shown using dashed lines.
+Size of the point represents the number of responses.
+For some locations we have as many as 10 responses.
+For location 1, we did not have any response for null set 1 in one of the interaction lineups.
+
+We observe some variability of performance for different null sets even though the same actual data plot was used for all these null sets.
+The lineup protocol is using a finite and small sample (19 in these experiments) of all possible null sets, so it is posisble to see some variability in performance depending on null set.
+If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate.
+From the plot we can see that null set 3 appeared more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot.
+This null set effect was tested, and results in significant difference only for null set 3 of interaction lineup.
+The rest of the null sets don't show any differences at 1\% significance level.
+
+%This may happen for some sets of null plots if one or more null plots are very similar to the actual plot.
 In another set of null plots this may not happen making some lineups easier than others even though the actual data plot is the same.
 This pattern is evident in the figure as we see proportion correct for null plot 5 is consistently above the null plot 1 for each of the locations.
-A test for differences in average performances of null sets shows significant difference only for null set 3 of interaction lineup.
-The rest of the null sets don't show any differences at 1\% significance level.
 
@@
 
-The overall average proportion of correct responses for each location are shown using dashed lines in Figure \ref{fig:location_effect}.
-There is variability in performances for each null set depending on the locations.
-But the overall proportion is not varying much for different locations.
-
-The sizes of the dots in  Figure \ref{fig:location_effect} represents the number of responses obtained for that location and null set.
+%The sizes of the dots in  Figure \ref{fig:location_effect} represents the number of responses obtained for that location and null set.
 For some locations we have as many as 10 responses.
@@
 
-We fitted model \eqref{manova} to test if the mean performance vectors are similar for different locations.
-For this we use {\it anova()} function of {\it stats} package of \cite{R:2012}.
+Model \eqref{manova} was fit to the data to test if the detection rates were significantly different for the different locations.
+For this we use {\it manova()} function of {\it stats} package of \cite{R:2012}.
 The results are shown in Table \ref{tbl:manova}.
 The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05.
-This suggest that there may be no difference in location.
+This suggest that location of the actual data plot in the lineup does not significantly affect performance.
 

### 2014-01-08 sociology_paper.tex
up to conclusions

@@
 
-Human demographics have significant influence on performance in time taken to evaluate a lineup.
+Human demographics have the potential to influence performance when the lineup protocol is used for statistical inference.
+In our study of the demographic effects on performance across a set of different experiments we found some statistically significant effects.
+Age group 36-40, Countries other than India and United states, People who have a graduate degree have a significantly higher detection rate.
+Gender does not have any significant effect.
+However, the effects are minimal, on the order of a few percentage points different from the average.
+This result is very important for the power of visual test as it demonstrates the robustness of the test for different human factors.
+
+%in time taken to evaluate a lineup.
 Some variations among the factors are observed in terms of probability of correct evaluation.
@@
 
-Individual learning trend is observed in both time taken and observer performance.
+Individual learning trend is observed in time taken but not so much in observer performance.
 Some individuals improved the performances while others showed decrease in their performances over attempts.
-But the overall performance of the observers does not increase through successive attempts while evaluating multiple lineups.
+This could be interpreted as good, in the sense that it means that observers could realisitically be recruited from source like MTurk, and that substantial training is not necessary in order to obtain useful evaulations of lineups in practice.
+However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea.
+
+%But the overall performance of the observers does not increase through successive attempts while evaluating multiple lineups.
 This result suggests that the power estimated for visual inference using human subject experiment is robust and may not change if those participants are allowed to give feedback again.
@@
 
-The simulation experiment reveals that there is no significant effect of location of actual data plot in the lineup.
-This is important as the visual statistical inference procedure suggests that the data plot be placed at random anywhere in the lineup.
+The simulation experiment reveals that there is no significant effect of location of the actual data plot in the lineup.
+This is important as the visual statistical inference procedure prescribes that the data plot be placed at random in the lineup.
 This paper suggests that any random place in a lineup is as good as other places in the lineup.
@@
 
-The subjects participating this study may not necessarily know about statistical graphics.
+%The subjects participating this study may not necessarily know about statistical graphics.
 The numerous pilot studies done with more trained participants suggest that power of visual statical inference may be higher for observers who have advanced knowledge about statistical graphics.
@@
 
-This investigation allowed each observer to evaluate 10 lineups assuming that it would not cause fatigue or disinterest toward the task of evaluating lineups.
+%This investigation allowed each observer to evaluate 10 lineups assuming that it would not cause fatigue or disinterest toward the task of evaluating lineups.
 This assumption is made based on the pilot studies.
@@
 
-A lineup with fewer or larger than 20 plots may yield different results.
+%A lineup with fewer or larger than 20 plots may yield different results.
 If the size of the lineup is much higher than 20, it is intuitive that there may be location effect of the data plot in the lineup.
@@
 
+There are many more ways that the lineup protocol might be tested to learn what we might expect about its performance for tackling real data mining problems.
+Possible changes in the lineup protocol in the pipeline are allowing observers to select more than one plot, and to vary the size of the lineup from 20 to a smaller number, and have observers all see different lineups with different null sets.
+This paper assessed the effect of human factors on the experiments conducted to date.
+As changes to the lineup protocol are suggested by other experiments the human factor effects me need to be examined again.
+
 %------------------------------------------------------------------------------------
@@
 This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
+
+\end{document}
+
+\dc{*** Suggest all of this is supplementary}
 

### 2014-01-09 sociology_paper.tex
done!!
author info updated

@@
 \author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is a PhD student (e-mail: mahbub72@gmail.com), Heike Hofmann is an Associate  Professor and Dianne Cook is a Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
+\thanks{Mahbubul Majumder is an Assistant Professor in the Department of Mathematics, University of Nebraska at Omaha (e-mail: mmajumder@unomaha.edu), Heike Hofmann and Dianne Cook are Professors in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
 This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}

### 2014-01-09 R/sociology_paper.R

@@ -91,6 +91,11 @@ dat9$response_all <- apply(subset(dat9, select=c(plot_location,response_no)), 1,
 # responses on the plot used for defense talk
 subset(dat9, pic_name=="plot_turk9_interaction_1_3.svg")
 
+# responses on the plot used for defense nsf grant
+subset(dat9, pic_name=="plot_turk9_geno_1_2.svg")
+subset(dat9, pic_name=="plot_turk9_interaction_2_1.svg")
+
+
 # ==============================================================
 # Summary statistics of demographic data
 # --------------------------------------------------------------

### 2014-01-09 sociology_paper.tex
figure 4 demographic_effect.pdf updated

@@
    \includegraphics[width=4in]{demographic_effect.pdf} 
-   \caption{Boxplots of average log time taken and proportion correct responses of all the lineups plotted for each demographic factor levels.
+   \caption{Boxplots of average log time taken and proportion correct responses (detection rate) of all the lineups plotted for each demographic factor levels.
 The dots inside the boxes represent means.
 Some differences in means of various demographic factors are observed.
-Variability in proportion correct indicates large variability in lineup difficulties.
-\dc{*** Vertical axis needs to be labelled ``detection rate'' for bottom row, and ``value'' needs to be removed.}}
+Variability in detection rate indicates large variability in lineup difficulties.
+}
    \label{fig:demographic_effect}
 \end{figure}
+
+%\dc{*** Vertical axis needs to be labelled ``detection rate'' for bottom row, and ``value'' needs to be removed.}
 

### 2014-01-09 R/sociology_paper.R

@@ -209,7 +209,7 @@ ggsave("../images/age_gender_within_country_correct.pdf", width=10, height=8)
 get_effect <- function(dat, var){
   res <- ddply(dat,c(var), summarize,
                log_avg_time = round(log(mean(time_taken)),2),
-               prop_correct = mean(response)
+               detection_rate = mean(response)
   )
   res$variable_name <- var[1]
   colnames(res) <- c("variable_level", colnames(res)[-1])
@@ -224,7 +224,7 @@ mdat <- melt(rbind(gdat,edat,cdat,adat), id=c("variable_level", "pic_name", "var
 
 qplot(variable_level, value, geom="boxplot",data=mdat[complete.cases(mdat),]) +
   facet_grid(variable~variable_name, scales="free") +
-  stat_summary(fun.y=mean, geom="point") + xlab("Levels of demographic factors")+
+  stat_summary(fun.y=mean, geom="point") + xlab("Levels of demographic factors")+ylab("")+
   theme(axis.text.x=element_text(angle=90, hjust=1))
 
 ggsave("../images/demographic_effect.pdf", width=6.5, height=6)

### 2014-01-09 sociology_paper.tex
references updated

@@
 \author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is an Assistant Professor in the Department of Mathematics, University of Nebraska at Omaha (e-mail: mmajumder@unomaha.edu), Heike Hofmann and Dianne Cook are Professors in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
+\thanks{Mahbubul Majumder is an Assistant Professor in the Department of Mathematics, University of Nebraska at Omaha, NE 68182 (e-mail: mmajumder@unomaha.edu), Heike Hofmann and Dianne Cook are Professors in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
 This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
@@
 This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
+
+\bibliographystyle{asa}
+\bibliography{references}
 

### 2014-01-11 sociology_paper.tex
number of genes analyzed
comments about detection rate

@@
 \newcommand{\dc}[1]{{\color{green} #1}}
-\newcommand{\mm}[1]{{\color{blue} #1}}
+\newcommand{\mm}[1]{{\color{cyan} #1}}
 
@@
 \dc{*** Am trying to use detection rate rather than correct response.
-This isharder for this paper because we are evaluating subject's abilities to some extent, but we should try to use this in our vis inference papers.}
+This is harder for this paper because we are evaluating subject's abilities to some extent, but we should try to use this in our vis inference papers.} \mm{Di, do you suggest that I change ``correct response''  to ``detection rate'' in all the texts here?}
 
@@
 For each gene there were 11 data points, because each treatment contained two or three replicates.
-There are on the order of \dc{*** how many} genes examined for significant expression.
+\mm{Di, There are more than 40,000 genes examined for significant expression in \citep{tengfei:2013}.
+Is that what you wanted to put here?}.
 In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.

### 2014-01-12 sociology_paper.tex
computation of detection rate

@@
 Figure \ref{fig:demographic_effect} supports the model results.
-The average time taken to respond (natural log) and detection rate are computed by lineup \dc{*** check this, were these calculated by lineup??} for the different demographic factor levels, and displayed using boxplots.
+The average time taken to respond (natural log) and detection rate are computed by lineup for the different demographic factor levels, and displayed using boxplots.
 Since the models use the mean values, rather than median, these are represented by a dot inside the boxplots.

### 2014-08-02 sociology_paper.tex
some comments added

@@
 \usepackage{color}
+\definecolor{mahbub}{rgb}{1.0, 0.6, 0.4}
+%\usepackage[usenames,dvipsnames]{color}
 \newcommand{\hh}[1]{{\color{magenta} #1}}
 \newcommand{\dc}[1]{{\color{green} #1}}
-\newcommand{\mm}[1]{{\color{cyan} #1}}
+\newcommand{\mm}[1]{{\color{mahbub} #1}}
 
@@
 where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model \eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
+\mm{ Since each person evaluate a lineup only once, should it be $\epsilon_{ij} \sim N(0, \sigma^2)$?}
 
@@
 & lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
- & Error($\sigma$) & 0.479 & &  &  &  & 0.692&& &\\ 
+ & Error($\sigma$) & 0.479 & &  &  &  & \mm{? 0.692}&& &\\ 
    \hline
@@
 Even though graduate degree is statistically significant, the largest difference in proportion correct is  0.045 which is very negligible.
-The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) of lineup variability.}
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) \mm{should it be 5.259 instead ?} of lineup variability.}
    \label{fig:practical_impact_graduate}

### 2014-08-12 sociology_paper.tex
comments added

@@
 
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{dfn}{Definition}[section]
-\newtheorem{cor}{Corollary}[thm]
-\newtheorem{con}{Conjecture}[thm]
-\newtheorem{lemma}[thm]{Lemma}
+%\newtheorem{thm}{Theorem}[section]
+%\newtheorem{dfn}{Definition}[section]
+%\newtheorem{cor}{Corollary}[thm]
+%\newtheorem{con}{Conjecture}[thm]
+%\newtheorem{lemma}[thm]{Lemma}
 
@@
 
-\tableofcontents
+% tableofcontents
 
@@
 
-\dc{*** Am trying to use detection rate rather than correct response.
-This is harder for this paper because we are evaluating subject's abilities to some extent, but we should try to use this in our vis inference papers.} \mm{Di, do you suggest that I change ``correct response''  to ``detection rate'' in all the texts here?}
+%\dc{*** Am trying to use detection rate rather than correct response.
+This is harder for this paper because we are evaluating subject's abilities to some extent, but we should try to use this in our vis inference papers.} % \mm{Di, do you suggest that I change ``correct response''  to ``detection rate'' in all the texts here?}
 
@@
 
-\dc{*** It seems repetitive to start with a description of what we learn from the plot, so start with the model summary and use the plot to support the model findings.}
+%\dc{*** It seems repetitive to start with a description of what we learn from the plot, so start with the model summary and use the plot to support the model findings.}
 
@@
 Even though graduate degree is statistically significant, the largest difference in proportion correct is  0.045 which is very negligible.
-The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) \mm{should it be 5.259 instead ?} of lineup variability.}
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) \mm{should it be 5.259 instead ?}
+   of lineup variability.}
    \label{fig:practical_impact_graduate}

### 2014-10-20 sociology_paper.tex
small edits throughout

@@
 Experiments 5, 6 and 7 are used to study the learning trend of the observer.
-The design of experiment 9 incorporated components that allows position of the actual data plot in the lineup, and the sample of nulls, to be evaluated.
+The design of experiment~9 incorporated components that allows position of the actual data plot in the lineup, and the sample of nulls, to be evaluated.
 Section \ref{sec:factor_performance} discusses human factors that may affect the performance of the observer.
@@
 These experiments were conducted to examine many different visual tasks.
-Side-by-side boxplots were used in experiment 1 to study the importance of a categorical variable in a regression model.
-For both experiments 2 and 3, a scatterplot is the visual test statistic, but a regression line fitted through the points is overlaid for experiment 2.
+Side-by-side boxplots were used in experiment~1 to study the importance of a categorical variable in a regression model.
+For both experiments~2 and~3, a scatterplot is the visual test statistic, but a regression line fitted through the points is overlaid for experiment~2.
 The experiments studied slightly different problems, for which the difference in plot type was important.
-Experiment 4 compared polar coordinates to euclidean coordinates on the reading of structure in a barchart.
-Experiment 5 examined the style of plot for a bivariate data problem, by comparing results when using a scatterplot, side-by-side dot plots, overlaid densities or side-by-side boxplots.
-Experiment 6 studied the difference between variations of boxplots.
-Experiment 7 studied the effects of high dimension low sample size on the separation of groups in low-dimensional projections.
-The sine illusion, where people read the strength of differences between curves as different when there is a strong sinusoidal structure, is examined in experiment 8.
-Experiment 9 was conducted to test if there was any structure in an RNA-seq data set measured on soybeans, and used interaction plots classically constructed to examine interaction effects in $2\times 2$ factorial designs.
-Experiment 10 used different sorts of residual plots for hierarchical linear models as the visual test statistics.
-
-In experiments 4, 5 and 6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices.
-Results are published in \cite{heike:2012} and \cite{XXX}(The violin plot study needs to be referenced).
+Experiment~4 compared polar coordinates to euclidean coordinates on the reading of structure in a barchart.
+Experiment~5 examined the style of plot for a bivariate data problem, by comparing results when using a scatterplot, side-by-side dot plots, overlaid densities or side-by-side boxplots.
+Experiment~6 studied the difference between variations of boxplots.
+Experiment~7 studied the effects of high dimension low sample size on the separation of groups in low-dimensional projections.
+The sine illusion, where people read the strength of differences between curves as different when there is a strong sinusoidal structure, is examined in experiment~8.
+Experiment~9 was conducted to test if there was any structure in an RNA-seq data set measured on soybeans, and used interaction plots classically constructed to examine interaction effects in $2\times 2$ factorial designs.
+Experiment~10 used different sorts of residual plots for hierarchical linear models as the visual test statistics.
+
+In experiments~4,~5 and~6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices.
+Results are published in \cite{heike:2012} and \cite{boxplots:arxiv}.
 
@@
 The IP address of the computer afforded the geographical location of the subject.
-The purpose of collecting this information with each experiment is to examine the effect that they have on the results of visual inference -- ideally very little.
+The purpose of collecting this information with each experiment~is to examine the effect that they have on the results of visual inference -- ideally very little.
 
@@
 Ideally this does not happen.
-Experiment 9, conducted to test for structure in a real data set, enables this to be studied.
+Experiment~9, conducted to test for structure in a real data set, enables this to be studied.
 Five different locations in the lineup were used to assess how fast and accurately observers identified the actual data plot.
@@
 The sample of null plots can affect the observer's decision.
-This is tested with the data from experiment 9, where the experiment was set up with lineups made from different null plots.
+This is tested with the data from experiment~9, where the experiment was set up with lineups made from different null plots.
 
@@
 The design of experiments 5, 6 and 7 enables the examination of learning trend, which is studied in this paper.
-Experiment 9 was a real test case for visual statistical inference, and in order to understand the significance of the structure in the genomic data, multiple lineups were made in which location of the actual data plot, and the sample of nulls, were randomized.
+Experiment~9 was a real test case for visual statistical inference, and in order to understand the significance of the structure in the genomic data, multiple lineups were made in which location of the actual data plot, and the sample of nulls, were randomized.
 This enables the assessment of the effect of these factors on the results.
@@
 Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
-The factors are examined in association with the observer's response using a mixed effects linear model:
+The factors are examined in association with the observer's response using a logistic regression with random effects terms:
 \begin{equation} \label{eqn:demographic_response}
@@
 
-In addition, a model is constructed with time taken as the response variable.
-Let $Z_{ij}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$.
-Let $\mu_{ij}=  E(Z_{ij})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$.
-We model this in a mixed effects model of the form
+In addition, we model the  time  an observer takes to identify a panel from a lineup.
+Let $Z_{ij}$ denote the logarithm of time taken for  observer $i$ to evaluate  lineup $j$.
+Let $\mu_{ij}=  E(Z_{ij})$ be the average of the (log) time taken by  observer $i$ to pick a panel from lineup $j$.
+We model this in a mixed effects model of the same structure as model~(\ref{eqn:demographic_response}) given as:
 %\hh{
@@
 \end{equation}
-where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model \eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-\mm{ Since each person evaluate a lineup only once, should it be $\epsilon_{ij} \sim N(0, \sigma^2)$?}
-
-\subsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive feedbacks received when multiple lineups are shown for evaluation.
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model \eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.
+
+\subsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive responses when multiple lineups are shown for evaluation.
 Experiments 5, 6 and 7 were used for this.
 Each subject was shown a total of 10 lineups randomly selected from a pool of lineups.
-The lineups are not necessarily with the same difficulty levels, but the order of lineups was randomized.
+The lineups are not necessarily of the same difficulty level, but the order of lineups was randomized.
 The responses of the lineups were recorded by attempt 1 through 10.
@@
 %Multiple lineups were sequentially shown to each observer for evaluation. 
-Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k^{th}$ evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
-Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the actual data plot from lineup $j$ in their $k^{th}$ attempt.
+Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k$th evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the actual data plot from lineup $j$ in their $k$th attempt.
 Learning trend is assessed using a generalized mixed effects model of the form
@@
 \end{equation}
-where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k^{th}$ attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
+where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k$th attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
 The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
@@
 $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-The inverse link function, $g^{-1}(.)$, from equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k^{th}$ attempt by a single observer as
+The inverse link function, $g^{-1}(.)$, from equation~\ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in the $k$th attempt by a single observer as
 \begin{equation} \label{eqn:trend_power}
@@
 
-When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k^{th}$ attempt.
-Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k^{th}$ attempt.
+When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
+Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k$th attempt.
 We model this in a mixed effects model of the form
@@
 All other effects are random effects: $u_i$ is a subject-specific intercept representing individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$, $a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$, with $a_i \sim N(0, \sigma_a^2)$, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $kth$ attempt by a single observer as 
+Equation~\ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
 \begin{equation} \label{eqn:trend_time_est}
@@
 
-To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
+To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2011}.
 To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
 
-\subsection{Location Effect} \label{sec:location_design} Experiment 9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
+\subsection{Location Effect} \label{sec:location_design} Experiment~9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
 The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
@@
 Let $\mathbf{Y}=(Y_1,Y_2, ...
-, Y_p)^\top$ be a vector of random variables with dimension $p (=5)$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., I(=5)$.
+, Y_p)^\top$ be a vector of random variables with dimension $p (=5)$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represents $j$th vector response for $i$th location with $i=1,2, ..., I(=5)$.
 Because the same data plot is shown in each lineup, it is assumed that there could be some association between the responses for each null set, which suggests the MANOVA model rather than a univariate ANOVA.
@@
 %Multiple lineups were sequentially shown to each observer for evaluation.
-Suppose $Y_{ijk}$ denotes the response from observer $i$ on a  lineup $j$ in his/her $k^{th}$ attempt.
+Suppose $Y_{ijk}$ denotes the response from observer $i$ on a  lineup $j$ in his/her $k$th attempt.
 $Y_{ijk}=1$ if the response is correct otherwise  $Y_{ijk}=0$.
-Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in his/her $k^{th}$ attempt.
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in his/her $k$th attempt.
 We model this in a generalized mixed effects model of the form
@@
 %\end{equation}
-%where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k^{th}$ attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$.
+%where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k$th attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$.
 %$u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$. $u_i$ is a random intercept, describing a basic subject-specific ability. We assume $u_i \sim N(0, \sigma_u^2)$. 
@@
 %
-%The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k^{th}$ attempt by a single observer as
+%The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k$th attempt by a single observer as
 %\begin{equation} \label{eqn:trend_power}
@@
 %The learning of each observer over a course of $K$ evaluations may be observed as the improvement of the time taken to evaluate a lineup in the later attempts.
-Suppose $Z_{ijk}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k^{th}$ attempt.
-Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k^{th}$ attempt.
+Suppose $Z_{ijk}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
+Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k$th attempt.
 We model this in a mixed effects model of the form
@@
 %
-%Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $kth$ attempt by a single observer as 
+%Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
 %\begin{equation} \label{eqn:trend_time_est}
@@
 , Y_p)^\top$ be a vector of random variable with dimension $p$, the total number of null sets.
-let $\mathbf{Y}_{ij}$ represents $jth$ vector response for $ith$ location with $i=1,2, ..., I$.
+let $\mathbf{Y}_{ij}$ represents $j$th vector response for $i$th location with $i=1,2, ..., I$.
 We fit the following MANOVA model
@@
 A total of 1911 lineups were evaluated in the ten experiments.
-Each person evaluated at least 10 lineups except for experiment 9 where three lineups were evaluated by each person.
+Each person evaluated at least 10 lineups except for experiment~9 where three lineups were evaluated by each person.
 A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
@@
 Table \ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model \eqref{eqn:trend_response} examining detection rate.
-There is some evidence of a learning curve only for experiment 6.
+There is some evidence of a learning curve only for experiment~6.
 There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.
@@
 Attempt number is fitted as a factor to enable modeling any sort of non-linear learning trend.
-Only experiment 6 shows evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
+Only experiment~6 shows evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
 \scalebox{0.81}{
@@
   \hline
-& \multicolumn{4}{c} {Experiment 5} & &\multicolumn{4}{c} {Experiment 6} && \multicolumn{4}{c} {Experiment 7}\\
+& \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
 
@@
 
-%The variance of random slope for attempt ($\sigma^2_a$) is estimated to be very small compared to other random effects except for subject variability ($\sigma^2_u$) for experiment 7.
+%The variance of random slope for attempt ($\sigma^2_a$) is estimated to be very small compared to other random effects except for subject variability ($\sigma^2_u$) for experiment~7.
 This suggests that most of the variability for attempt is accounted for by the fixed effect factor attempt.
 For all three experiments the lineup variabilities ($\sigma^2_l$) were estimated to be very large making the practical impact of other factor levels even more negligible.
-Some of the lineups were very easy and some were really difficult in experiment 7.
+Some of the lineups were very easy and some were really difficult in experiment~7.
 The overall average probability for picking out the data plot from a lineup ($\mu$) is significant for both experiments 5 and 7.
-But for experiment 6 it is not significant suggesting that experiment 6 lineups were difficult compared to the other experimental lineups.
+But for experiment~6 it is not significant suggesting that experiment~6 lineups were difficult compared to the other experimental lineups.
 This shows a feature of the experimental designs where a mixture of difficult and easy lineups were included within the experiments as well as between the experiments.
@@
 
-%Model \eqref{eqn:trend_time} is fitted to the data from experiment 5, 6, and 7 separately.
+%Model \eqref{eqn:trend_time} is fitted to the data from experiment~5, 6, and 7 separately.
 For this the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
@@
 %Each subject participating the experiment has seen multiple number of lineups for evaluations.
-Suppose $K$ be the number of lineups evaluated by a subject and $A_k$ be the $kth$ attempt, $k=1,2, ..., K$.
+Suppose $K$ be the number of lineups evaluated by a subject and $A_k$ be the $k$th attempt, $k=1,2, ..., K$.
 Subjects may have a chance for self learning and do better on lineups shown later in the sequence.
@@
 Some subjects show improvement over time and some show the decrease in performance.
-Although, the model fit experiment 6 suggested a statistically significant learning curve the plots indicates that is minimal.
+Although, the model fit experiment~6 suggested a statistically significant learning curve the plots indicates that is minimal.
 
@@
 
-%The lineup variance of experiment 7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model \eqref{eqn:trend_response}.
+%The lineup variance of experiment~7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model \eqref{eqn:trend_response}.
 But from model \eqref{eqn:trend_time} it is estimated to be much smaller in all these experiments.
@@
 %A plot of residuals against the attempts $A_k$ should reveal the trend if there is any.
-Figure \ref{fig:learning_trend} shows such plots for three experiments naming experiment 5, 6 and 7.
+Figure \ref{fig:learning_trend} shows such plots for three experiments naming experiment~5, 6 and 7.
 For all the experiments we see some increasing trend of the residuals.
@@
 %\caption{Fixed effect parameter estimates of generalized mixed model.
-Note that attempt is not significant for experiment 2.
+Note that attempt is not significant for experiment~2.
 The continuous covariate, lineup difficulty, was measured by the $p$-value of the actual plot in the lineup.}
@@
 %  \hline
-%\multicolumn{2}{l}{\bf{Experiment 1} } &&&& &\\
+%\multicolumn{2}{l}{\bf{Experiment~1} } &&&& &\\
 %&(Intercept) & -0.30 & 0.10 & -2.89 & 0.00 & ***\\ 
@@
 % & lineup difficulty & -11.32 & 1.01 & -11.19 & 0.00 & ***\\ 
-%\multicolumn{2}{l}{\bf{Experiment 2} } &&&&& \\
+%\multicolumn{2}{l}{\bf{Experiment~2} } &&&&& \\
 % & (Intercept) & 2.36 & 0.15 & 16.09 & 0.00 & ***\\ 
@@
 % & lineup difficulty & -55.03 & 2.20 & -25.00 & 0.00 & ***\\ 
-%\multicolumn{2}{l}{\bf{Experiment 3} } &&&& \\
+%\multicolumn{2}{l}{\bf{Experiment~3} } &&&& \\
 % & (Intercept) & 0.25 & 0.16 & 1.58 & 0.11 & .\\ 
@@
 %But the overall performance of the observers does not increase through successive attempts while evaluating multiple lineups.
-This result suggests that the power estimated for visual inference using human subject experiment is robust and may not change if those participants are allowed to give feedback again.
+This result suggests that the power estimated for visual inference using human subject experiment~is robust and may not change if those participants are allowed to give feedback again.
 The skill in evaluating the lineup in shorter time gets improved over successive evaluations.
@@
 \cite{majumder:2013} showed that these choice reasons may provide the clue on how people picks the data plot in the lineup.
-To investigate this further we added free text input option in experiment 9 for their choice reason instead of some fixed reasons to select from.
+To investigate this further we added free text input option in experiment~9 for their choice reason instead of some fixed reasons to select from.
 This allowed observers to write whatever they think their reasons for choice are.
@@
    \caption{Number of participants by time of the day feedbacks received (central time).
-Experiment 1 shows MTurk workers participated the experiments around the clock.
+Experiment~1 shows MTurk workers participated the experiments around the clock.
 Other experiments did not take a whole day to finish.
-For experiment 3 most of the participants are from India because of timing.
+For experiment~3 most of the participants are from India because of timing.
 No matter when the experiment is started, subjects from India shows participations.
-For United States, subjects participated if the experiment is not in the mid night, except for experiment 6.}
+For United States, subjects participated if the experiment is not in the mid night, except for experiment~6.}
    \label{fig:participation_time}

### 2014-10-20 sociology_paper.tex
small edits throught the first section

@@
 
-The lineup protocol introduced in \citet{buja:2009} can be used to test the significance of findings during the exploratory data analysis.
-The methodology is a part of what is called visual statistical inference.
-These concepts have been developed further by \citet{majumder:2013} who refined the terminology and validated the lineup protocol with a head to head comparison with conventional inference.
+The lineup protocol introduced in \citet{buja:2009} can be used to test the significance of \hh{graphical} findings in the exploratory data analysis process.
+This methodology is  part of what is called visual statistical inference, which has been developed further  and validated by \citet{majumder:2013} in simulation studies of head to head comparisons with conventional inference.
 One of the major contributions of  \citet{majumder:2013}  is to define the power of the visual test and provide methods to obtain the power for a particular lineup.
@@
 
-Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are showing from data generated under a null model.
+Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are showing  data generated under a null model.
 Which one of the 20 plots is the most different from the others?
-When asked this question, 12 \hh{out} of 72 observers picked the  plot \hh{of the observed data} (located in spot number $3^2+4$ in the lineup).
+When asked this question, 12 \hh{out} of 72 observers picked the  plot \hh{of the observed data} (located in panel number $3^2+4$ in the lineup).
 The corresponding $p$-value is \hh{0.0077}, indicating sufficient evidence to reject the null hypothesis.
@@
 
-\hh{How do we interpret this finding, though?} For that, we need to know the context of the data and we need to have more information about the \hh{how null plots were generated}. 
-This particular example investigates the results from the 2012 US presidential election in comparison to the poll results just prior to the election. 
-(Although this example is more simplistic than most of the tests conducted to date, it will serve the purpose of illustrating the lineup protocol.) The data is looking at the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
-Each panel in Figure~\ref{fig:elect-1} shows an `electoral building' where each state in the union is represented by a rectangle.
+\hh{How do we interpret this finding, though?} For that, we need to know the context of the data and we need to have more information about \hh{how the null plots were generated}. 
+This particular example investigates the results from the 2012 US presidential election in comparison to  poll results just prior to the election. 
+Although this example is more simplistic than most of the tests conducted to date, it will serve the purpose of illustrating the lineup protocol.
+The data consists of the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
+Each panel in Figure~\ref{fig:elect-1} shows an `electoral building' \citep{mosley2010} where each state in the union is represented by a rectangle.
 This difference is plotted horizontally, and the height of each box corresponds to the state's electoral votes.
@@
 The polling results provides the null model from which data is simulated.
-Because each poll has a margin of error, this is used to simulate different scenarios that might have resulted on election day, if the polls were on target.
-A null data set is generated as a set of draws from normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
+\hh{A normal model with mean and standard deviation based on a poll and its margin of error is used to} simulate different scenarios that might have resulted on election day, if the polls were on target.
+A null data set is generated as a set of draws from this model.
+% normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
 These samples are plotted as electoral buildings, and the plot made with the results from the election is placed randomly among them in a lineup of size $5\times 4$.
-If the null hypothesis is true the actual data plot should look just like any of the other plots, and not be identified by an observer.
-Figure~\ref{fig:tower} shows a plot of the electoral building with added context information and labels.
+If the null hypothesis is true, the actual data plot should look  like any of the other plots, and not be identifiable by an observer.
+Figure~\ref{fig:tower} shows the plot of the electoral building with added context information and labels.
  %A party needs 270 electoral votes or more to win the presidential election. 

### 2014-10-21 sociology_paper.tex
edits throughout section 1

@@
 % normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
-These samples are plotted as electoral buildings, and the plot made with the results from the election is placed randomly among them in a lineup of size $5\times 4$.
-If the null hypothesis is true, the actual data plot should look  like any of the other plots, and not be identifiable by an observer.
+These samples are plotted as electoral buildings, and the plot based on the actual election results is placed randomly among the null plots in a lineup of size $5\times 4$.
+If the null hypothesis is true, the  plot of the actual data should look  like any of the other plots, and not be identifiable by an observer.
 Figure~\ref{fig:tower} shows the plot of the electoral building with added context information and labels.
@@
 In order to avoid expectation errors \citep{meilgaard}, observers should not be aware of the data that constitutes a lineup, and should not have seen the actual data plot before seeing the lineup.
-This is the reason that in the election example, above, the scenario was explained after the lineup question, in the text.
-
-The question that is asked of the observer should be as general as possible, effectively asking the observer to pick the plot that is different, and allowing them to provide their reasons for seeing their pick as different.
-In some of the studies, the ones described in  \cite{majumder:2013} very specific questions were asked, because the experiments were being conducted to compare results from the lineup protocol with those of conventional tests.
-In those experiments, structure in the data was strictly controlled in the simulation process, which allowed for specific questions to be asked.
+This is the reason that in \hh{our first example we explained the context only after presenting the lineup.}%the election example, above, the scenario was explained after the lineup question, in the text.
+
+\hh{When presenting a lineup to an observer we should also phrase the accompanying question in as general a manner as possible, effectively}
+%The question that is asked of the observer should be as general as possible, 
+effectively asking the observer to pick the plot that is \hh{most} different, and allowing \hh{observers} to provide their reasons for their choice. 
+\hh{If lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments  in  \cite{majumder:2013}, these questions have to be phrased much more specifically, in order to make the situations more comparable.
+}
+%In some of the studies, the ones described in  \cite{majumder:2013} very specific questions were asked, because the experiments were being conducted to compare results from the lineup protocol with those of conventional tests.
+In those experiments, the structure in the data is strictly controlled in the simulation process, which allows for specific questions to be asked. 
 In the election example, observers were asked ``which plot is the most different?''.
-The type of plot, showing two (modified) stacked bar charts in different colors should suggest to the observer that the interesting feature is most difference between the two heights.
-Most observers got this, but it is possible that some observers might pick plot 4, where the red tower is slightly above the blue as the most different because it is the only plot with this feature.
-So a better question may have been ``Which plot shows the biggest height difference between the two towers?'' except that this tailors the inference to a specific feature which does not match the null hypothesis of interest.
+The type of plot, showing two (modified) stacked bar charts in different colors should suggest to the observer that the interesting feature is the difference between the two heights.
+\hh{Most observers based their decision on this -- besides  `asymmetry' between the towers (a reason suggested by us), most free-form responses for the reason of choice describe a comparison between the red and blue towers.
+In the responses for this lineup we see something that is very typical of responses in lineups: it is only a few panels from the lineup that observers pick, while most panels do not get picked or get picked only a few times.
+Out of a total of 72~responses, 23 observers picked panel~\#8, 12 observers picked the (data) panel \#13, 10 observers picked panel~\#4, 6 observers picked panel \#7, and all of the remaining panels were picked at most three times.
+From the panels picked, we see that the top three choices  are all extreme with respect to heights of blue versus red towers:  both panels \#8 and \#13 show a large difference between the height of the towers, while panel \#4 is the only panel, in which the red tower is higher than the blue one.
+}
+% results from the turk study:
+% panel:   1 10 11 12 13 14 15 16 18 19  2 20  3   4  5  6  7  8  9 
+% #picks:  2  1    0  1 12    3  3    1  1   1  2   1  0 10  2  2  6 23  1
+
+%Most observers got this, but it is possible that some observers might pick plot~4, where the red tower is slightly above the blue as the most different because it is the only plot with this feature \hh{(10 out of 72 observers actually did this)}.
+  %a better question may
+
+ \hh{If we consider this difference to be the main objective of the visualization, a more focused question would}
+ have been ``Which plot shows the biggest height difference between the two towers?'' %except that this tailors the inference to a specific feature which does not match the null hypothesis of interest.
+
 %=======

### 2014-10-21 sociology_paper.tex
edits end of section 1

@@
 
-%=======
-%A lineup can be evaluated by a single person or multiple observers.
-Based on the feedback of $K$ observers a $p$-value is computed with the help of a binomial random variable $X$ for the number of observers picking the data plot.
-Under the null hypothesis is $X$ is  distributed according to $B_{K, 1/20}$ \citep{majumder:2013}.
-The $p$-value for a lineup is then given as $P(X \ge x)$.
-%%This finally a decision is made about the hypothesis being tested.
-%>>>>>>> example
-
-%\begin{figure}[htbp] 
-%   \centering
-%   \includegraphics[width=6.5in]{plot_turk9_geno_4_3.png} 
-%   \caption{A lineup of 20 plots. One of the plots is observed data and rest of the 19 plots are obtained by randomly permuting the group structure. Can you identify the observed data?}
-%   \label{fig:lineup_geno}
-%\end{figure}
-
-%\subsection{Observer task}
-
-For any evaluation the observer may or may not identify the actual data plot.
+\hh{XXX why do we need the last paragraph? It distracts from the main flow somewhat}
+
+% HH already included:
+% For any evaluation the observer may or may not identify the actual data plot.
 Under the null hypothesis, the actual data plot should look similar to the null plots making it harder to detect.
@@
 But since there are limited number of plots in a lineup, which is 20 in the election example, there is a 1/20 chance that the observer would pick the actual plot.
-This proportion is associated with the Type~I error of the test.
+
+\hh{The lineup protocols allows us to calculate all relevant properties that we are familiar with from conventional statistical tests.
+In particular, the power of a lineup is calculated as the {\it detection rate} at which observers identify the actual data plot.
+Visual power depends unlike power in conventional tests not only on the strength of the signal, but also on individuals' ability.
+}
+%This proportion is associated with the Type~I error of the test.
 On the other hand, if null hypothesis is not true, the observed plot should look different from the null plots, making it easier to be detected.
@@
 When multiple observers evaluate a lineup, the detection rate can be used to estimate the power.
-The ability of individual observers can vary, and examining this is the purpose of this paper.
+The ability of individual observers varies, and \hh{the effects that might influence ability are the focus} of this paper.
 
@@
 
-There have been ten experiments conducted using Amazon's Mechanical Turk~\citep{turk} (MTurk) that are being used to evaluate the effects of observers' demographic factors on the inference.
-Table \ref{tbl:visual_stat} summarizes these experiments.
-Each of these experiments collected demographic details of the subjects.
-Experiments 5, 6 and 7 are used to study the learning trend of the observer.
-The design of experiment~9 incorporated components that allows position of the actual data plot in the lineup, and the sample of nulls, to be evaluated.
+\hh{We have conducted a series of experiments (see Table \ref{tbl:visual_stat} for an overview)}
+%There have been ten experiments conducted 
+using Amazon's Mechanical Turk~\citep{turk} (MTurk) that \hh{provide a basis for evaluating} 
+%are being used to evaluate 
+the effects of covariates on individual's abilities to evaluate lineups. 
+\hh{In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.  }
+ %Each of these experiments collected demographic details of the subjects. 
+ Experiments~5, 6 and 7 can be used \hh{to investigate short term learning trends}.
+Experiment~9 was also designed to check whether the  positioning of the  data plot in the lineup had an effect on detection rate.
+
  Section \ref{sec:factor_performance} discusses human factors that may affect the performance of the observer.

### 2014-10-23 sociology_paper.tex
intro and table 1

@@
 \centering 
-\caption{Overview of 10 different Turk experiments, from where data was taken to study human factor effects.
+\caption{Overview of 10 different Turk experiments, \hh{in which} data was \hh{collected} to study human factor effects.
 All of the experimental data was used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.}
@@
  4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
-  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF and LT \\
+  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
   6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?
-&  DF and LT \\
+&  DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
   7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?
-& DF and LT \\
+& DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
   8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?
@@
   9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?
-& DF and LE\\
+& DF  \hfill \phantom{LT} \hfill {LE} \hfill \phantom{.}\\
   10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?
@@
 %are being used to evaluate 
-the effects of covariates on individual's abilities to evaluate lineups. 
+the effects of different covariates on an individual's ability to evaluate lineups. 
 \hh{In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.  }

### 2014-10-24 sociology_paper.tex
factors - halfway

@@
 
-Because visual statistical inference relies on plotting choices, lineup design and human evaluation, it is important to understand how these factors that might affect results.
-This section provides a brief description of the factors that are expected to have some impact on the performance of visual statistical inference.
+\hh{Visual statistical inference relies not only on the strength of the signal in the data, but also on how this information is presented in the plot, on the lineup design and on the abilities of human observers.} It is important to understand how these factors might affect results.
+\hh{Here, we provide a brief discussion }%This section provides a brief description
+of the factors that are expected to have some impact on the performance of visual statistical inference.
 
@@
 
-\begin{itemize} \itemsep 0in
+\begin{itemize} \itemsep 0in \parsep-10pt %\parskip0pt 
 %\subsection{Choice of Visual Test Statistic}  
 \item{\bf Choice of Visual Test Statistic:}  
-There are a variety of ways that data can be plotted and for plots to be enhanced by axes, legends or representations of models.
-The choice of plot chosen as the visual test statistic, primarily should be appropriate for the problem being investigated.
-For example, for studying the association between two quantitative variables a scatterplot is usually optimal.
-Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.
-Some choices of visual test statistics provide better results than others when studying the same problem, which is discussed in \cite{heike:2012}.
-This makes sense, because some plot types will make it easier for certain structures to be seen.
+%There are a variety of ways that
+Data can be plotted in a variety of ways and we can enhance plots by axes, legends or representations of models.
+The choice of plot primarily should be appropriate for the problem being investigated, as the data plot is the visual test statistic.
+When for example studying the association between two quantitative variables a scatterplot is usually optimal.
+\hh{XXX do we really want to say optimal?
+that sounds like we have an underlying criterion to optimize.
+XXX} Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.
+Some choices of visual test statistics provide better results than others when studying the same problem \citep{heike:2012}, because some plot types make it easier for certain structures to be seen.
 
@@
 An important factor that helps an observer to identify the actual data plot in a lineup is the strength of the signal in the data.
-Simulated data was generated for all of the experiments, except 9, enabling the study of signal strength and actual data plot detection.
-Primarily the results were as expected, that as the signal strength increased observers more frequently picked the actual data plot as different from the null plots.
+Simulated data was generated for all of the experiments, except exps.~4 and~9, enabling the study of signal strength and actual data plot detection.
+\hh{In experiment~4 we took a different approach by varying the sampling size, which also allowed us to investigate the effect a design had on detection rate while varying signal/noise ratio.} Primarily the results were as expected, that as the signal strength increased observers more frequently picked the actual data plot as different from the null plots.
 
@@
 \item{\bf Question Design:} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
-Visual inference supports discovery of structure, enabling the the unexpected to surprise us.
+Visual inference supports discovery of structure, enabling the unexpected to surprise us.
 This is tightly connected to the task that the human observer is asked to perform.
 For most purposes the observer is asked a very general question, such as to pick the plot that is most different from the others, and explain the reasons why they see it as different.
-By being general, all possible discoverable structures are included in the significance calculations.
-(\citet{buja:2009} has a very complete explanation.)
+By being general, all possible discoverable structures are included in the significance calculations \citet[see also ][]{buja:2009}.
 
 In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted.
-This was important in order to make a comparison of the results.
-For latter studies (4, 7, 10), question posed to the observer was broader.
-Table \ref{tbl:visual_stat} lists the questions used for each experiment.
+This was important \hh{to allow a direct comparison of the results to results from conventional testing.}%in order to make a comparison of the results.
+For latter studies (4, 7, 10), questions posed to the observer were more general. Table \ref{tbl:visual_stat} lists the questions used for each experiment.
 
@@
 \item{\bf Demographics:}  The experiments collected data on age, gender, education level and geographic location.
-Each observer self-reported their gender, age in roughly five year intervals, education level as high school or less, some college courses, an undergraduate degree, some graduate courses or a graduate degree.
+Each observer self-reported gender, age in roughly five year intervals, education level as high school or less, some college courses, an undergraduate degree, some graduate courses or a graduate degree.
 The IP address of the computer afforded the geographical location of the subject.
-The purpose of collecting this information with each experiment~is to examine the effect that they have on the results of visual inference -- ideally very little.
+The purpose of collecting this information with each experiment is to examine the effect that they have on the results of visual inference -- ideally very little.
 

### 2014-10-26 sociology_paper.tex
edits factors

@@
 In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted.
-This was important \hh{to allow a direct comparison of the results to results from conventional testing.}%in order to make a comparison of the results.
+This was important \hh{to allow a direct comparison of the results to results from conventional testing.
+}%in order to make a comparison of the results.
 For latter studies (4, 7, 10), questions posed to the observer were more general. Table \ref{tbl:visual_stat} lists the questions used for each experiment.
@@
 \item{\bf Learning Trend:} One might expect that as an observer evaluates more lineups they may become more skillful in their evaluations.
-Each subject in the experiments 5, 6 and 7 evaluated a block of of ten lineups of the same type of data plot.
+Each participant in  experiments~5, 6 and~7 was asked to evaluate a block of  ten lineups of the same type of data plot.
 The ten lineups were randomly chosen from the lineups  produced for each study.
 Before evaluating the first lineup, the observer needs to read instructions and become accustomed to the type of plot used in the lineup.
-In subsequent lineups being evaluated the type of plot is the same, level of difficulty will be different.
+In subsequent lineups  the type of plot is the same.
 It is possible that the observer becomes more skillful at recognizing the most different plot, either by more often detecting the actual data plot or more quickly reporting their choice.
-These two ways of measuring learning trend are evaluated on data collected from experiments 5, 6 and 7.
+These two ways of measuring learning trend are evaluated on data collected from experiments~5, 6 and~7.
 
@@
 A random number generator is used to determine the position where the actual data plot is placed in each lineup.
-With the help of eye-tracking equipments \citet{zhao:2012} tracked the observers' eyes to see how they searched the plots in a lineup.
-The results suggest that many people read lineups from left to right direction, some read from upward to downward.
-There is some concern that if the actual data plot is located in the top left of the lineup, observers may find it more easily than if it were placed in the bottom right.
-Ideally this does not happen.
-Experiment~9, conducted to test for structure in a real data set, enables this to be studied.
+Eye tracking experiments \citep{zhao:2012} suggest that some observers traverse lineups in horizontal direction, while others have a vertical up-and-down approach to their search.
+Almost universal, observers start at the top left of a lineup.
+This leads to some concern that observers might be able to more easily identify the actual data plot, if it is placed at the top left of a lineup than when it the actual data is placed at the bottom right.
+%With the help of eye-tracking equipment \citet{zhao:2012} tracked the observers' eyes to see their approach in searching the plots in a lineup.
+The results suggest that many people read lineups from left to right, some read from upward to downward.
+%There is some concern that if the actual data plot is located in the top left of the lineup, observers may be able to identify it more easily than when the actual data plot is at the bottom right. 
+Ideally, this does not happen.
+Experiment~9  is designed to allow us to investigate the effects of location on detection.
 Five different locations in the lineup were used to assess how fast and accurately observers identified the actual data plot.
@@
 %\subsection{Sample of Null Plots} 
-\item{\bf Sample of Null Plots:} In classical inference, the test statistic is compared with the full null distribution, to determine if it is extreme or not.
-In visual inference the actual plot is compared with plots of a finite number of samples from the null distribution.
-In the lineups used in the Turk experiments, there are 19 null plots upon which the comparison with the actual data plot is made.
+\item{\bf Sample of Null Plots:} In classical inference, the test statistic is compared to the full null distribution, to decide if it is extreme or not.
+In visual inference the actual plot is compared to plots of a finite number of samples from the null distribution.
+In the lineups used in the Turk experiments, the actual data plot is compared on 19 null plots.
 These null samples are random draws, and there is a chance that one or more null plots in the lineup may be similar or even more extreme than the actual data plot.

### 2015-02-08 sociology_paper.tex
read-through up to page 8

@@
 
-\hh{When presenting a lineup to an observer we should also phrase the accompanying question in as general a manner as possible, effectively}
+\hh{When presenting a lineup to an observer we should also phrase the accompanying question in as general a manner as possible, }
 %The question that is asked of the observer should be as general as possible, 
 effectively asking the observer to pick the plot that is \hh{most} different, and allowing \hh{observers} to provide their reasons for their choice. 
-\hh{If lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments  in  \cite{majumder:2013}, these questions have to be phrased much more specifically, in order to make the situations more comparable.
+\hh{However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments  in  \cite{majumder:2013}, these questions have to be phrased much more specifically, in order to make the situations more comparable.
 }
@@
 
- \hh{If we consider this difference to be the main objective of the visualization, a more focused question would}
+ \hh{If we consider this difference to be the main objective of the visualization, a more focused question, that aligns more with a hypothesis test on the difference  in electoral votes  between parties, would}
  have been ``Which plot shows the biggest height difference between the two towers?'' %except that this tailors the inference to a specific feature which does not match the null hypothesis of interest.
-
-\hh{XXX why do we need the last paragraph? It distracts from the main flow somewhat}
 
@@
 For most purposes the observer is asked a very general question, such as to pick the plot that is most different from the others, and explain the reasons why they see it as different.
-By being general, all possible discoverable structures are included in the significance calculations \citet[see also ][]{buja:2009}.
+By being general, all possible discoverable structures are included in the significance calculations \citep[see also ][]{buja:2009}.
 
@@
 %\subsection{Demographics} 
-\item{\bf Demographics:}  The experiments collected data on age, gender, education level and geographic location.
+\item{\bf Demographics:}  \hh{During the experiments } data on age, gender, education level and geographic location \hh{was collected}.
 Each observer self-reported gender, age in roughly five year intervals, education level as high school or less, some college courses, an undergraduate degree, some graduate courses or a graduate degree.
@@
 %\subsection{Learning Trend} 
-\item{\bf Learning Trend:} One might expect that as an observer evaluates more lineups they may become more skillful in their evaluations.
+\item{\bf Learning Trend:} One might expect that as an observer evaluates more lineups they  become more skillful in their evaluations.
 Each participant in  experiments~5, 6 and~7 was asked to evaluate a block of  ten lineups of the same type of data plot.
@@
 %\subsection{Location of Actual Data Plot in the Lineup}
-\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 points is used.
+\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 \hh{panels} is used.
 A random number generator is used to determine the position where the actual data plot is placed in each lineup.
 Eye tracking experiments \citep{zhao:2012} suggest that some observers traverse lineups in horizontal direction, while others have a vertical up-and-down approach to their search.
-Almost universal, observers start at the top left of a lineup.
+Almost universally, observers start at the top left of a lineup.
 This leads to some concern that observers might be able to more easily identify the actual data plot, if it is placed at the top left of a lineup than when it the actual data is placed at the bottom right.
@@
 %\subsection{Individual Skill or Ability} 
-\item{\bf Individual Skill or Ability:} Each person may have different visual skill sets, and aptitude for reading statistical plots.
-This can be examined by having the same lineups evaluated by multiple subjects, so that it can be seen if some subjects consistently detect the actual data plot more often than others.
+\item{\bf Individual Skill or Ability:} Each person may have different  aptitude for reading statistical plots, and their visual skill sets might be differently developed.
+We \hh{We can evaluate the effect of individuals' skills and abilities}  by letting multiple subjects evaluate the same lineups, which allows us to estimate, if some subjects consistently detect the actual data plot more often than others.
  %For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.
@@
 
-All the ten experiments shown in Table \ref{tbl:visual_stat}, had demographic information collected from subjects, as follows:
+For all  ten experiments shown in Table \ref{tbl:visual_stat}, the following demographic information was collected from subjects:
 
 \begin{enumerate} \leftmargin 5cm  \itemsep 0in
-\item Age group, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, 51-55, 56-60, above 60.
-\item Gender, male or female.
-\item Education level, with levels being high school or less, some under grad course, under graduate degree, some graduate courses, and graduate degree.
-\item Geographical location, collected from the IP address of the participants' computer, as latitude, longitude, city and country. 
+\item {\it Age group}, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, 51-55, 56-60, above 60.
+\item {\it Gender}, male or female.
+\item {\it Education level}, with levels being high school or less, some undergraduate courses, undergraduate degree, some graduate courses, and graduate degree.
+\item {\it Geographical location}, collected from the IP address of the participants' computer, as latitude, longitude, city and country. 
 \end{enumerate}
@@
 Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
-The factors are examined in association with the observer's response using a logistic regression with random effects terms:
+The factors are examined in association with the observer's response using a logistic regression with random effect terms:
 \begin{equation} \label{eqn:demographic_response}
@@
 %\end{equation}
- $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$.
-The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume  $\ell_j \sim N(0, \sigma_\ell^2)$, and $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-
-In addition, we model the  time  an observer takes to identify a panel from a lineup.
+ $\mu$ is an overall population average, %probability for picking out the data plot from a lineup HH: mu is not a probability, it is on the scale of the link function
+ $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively for observer $i$.
+The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume independence and normality of the errors, i.e.\ $\ell_j \sim N(0, \sigma_\ell^2)$.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+
+Similarly, we model the  time  an observer takes to identify a panel from a lineup.
 Let $Z_{ij}$ denote the logarithm of time taken for  observer $i$ to evaluate  lineup $j$.

### 2015-02-08 sociology_paper.tex
proofread up to section 4.2

@@
 \subsection{Demographic Factors}  %In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table \ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
-
+\noindent
 For all  ten experiments shown in Table \ref{tbl:visual_stat}, the following demographic information was collected from subjects:
@@
 \end{enumerate}
-
+%
 %Instead of collecting exact age, 9 levels of age were collected.
@@
 This provides latitude and longitude of the ip address as well as the city and country information.
-
+%
 Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
@@
 \end{equation}
-where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k$th attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
+where $\mu$ is an overall population average, %probability for picking out the data plot from a lineup, 
+$\alpha_k$ is the effect of the $k$th attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
 The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
@@
 \subsection{Overview of the Data} A total of 2321 participants provided feedback data on the lineups in ten different experimental studies.
-Figure \ref{fig:turker_location} displays the locations of participants around the world.
+Figure~\ref{fig:turker_location} displays the locations of participants around the world.
 Most of the participants were from the United States and India.
@@
 This provides a diverse pool of participants.
-The diversity in not only in geographic but also in their gender, age group and education level as can be seen in Table \ref{tbl:demographics}.
-It is interesting that there were large number of female participants even though there were lot of participants worked from developing countries.
+The diversity in not only  geographic but also in gender, age group and education level as can be seen in Table \ref{tbl:demographics}.
+% It is interesting that there were large number of female participants even though there were a lot of participants  from developing countries.
+\hh{The large number of female participants from all countries was a pleasant surprise to us.}
 
@@
    \includegraphics[width=4.5in]{turker_location.pdf}    
-   \caption{Location of the Amazon Mechanical Turk workers participating our study. Most of the people are coming from India and United States even though there are people from around the world.}
+   \caption{Location of  Amazon Mechanical Turk workers participating in the experiments. Most subjects came from India and the United States, but subjects  from countries around the world took part.}
    \label{fig:turker_location}
@@
 Besides United States and India, countries such as Canada, Romania, United Kingdom and Macedonia have more than 10 participants each.
-The rest of the 70 countries have less than 10 participants.
-The distribution of participants remains similar in all ten experiments.
-That may be due to the time of day when the experiment was made available.
+The rest of the 70 countries have fewer than 10 participants.
+The distribution of participants is similar in all ten experiments.
+%This may be due to the time of day when the experiment was made available.
 %For some experiments more people participated from India than the United States and for others numbers were reversed (Figure~\ref{fig:participation_time}).
 It is also observed in some experiments that number of  participants are similar no matter whether the experiment is run on day time or nigh time.
-(*** Took out reference to figure 11 in appendix.)
+(*** Took out reference to Figure~11 in appendix.)
 
@@
 
-The largest number of participants are from age group 18 to 25, with the majority between 18 to 35.
-Interestingly, there are many participants in the older age groups as well.
-The United States shows a more uniform participation beyond age 30 but Indian participants are primarily younger.
-
-In terms of academic background, the largest participant group is graduate degree, with a total of 902 participants, about 38.51\%.
-Examining the location in conjunction with the academic background reveals that there are more Indian participants that list themselves as having a graduate degree.
-In retrospect this was realized to be a language confusion, as any university degree is considered to be a graduate degree in India.
-Unlike the definition in North America where graduate degree means Masters level education.
-Most of the USA participants list themselves as having an undergraduate degree or at least have some undergraduate courses.
-
-The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where proportion of female participants is lower.
-The distribution of education levels also differ slightly across the countries for this age group.
+The largest number of participants \hh{falls within the} age group of 18 to 25, with the majority between 18 to 35.
+\hh{ Many older participants took part in the experiments.} %Interestingly, there are many participants in the older age groups as well.
+The United States shows a more uniform participation beyond age 30, Indian participants are primarily younger.
+
+In terms of academic background, \hh{a majority of participants has a graduate degree} (a total of 902 participants, about 38.51\%).
+Examining the location in conjunction with the academic background reveals that there are more Indian participants that \hh{report to have } a graduate degree.
+In retrospect \hh{we realized this to be a culturally based misunderstanding}, as any university degree in India is considered to be a graduate degree, while the North American definition of graduate degree refers to a Masters level education.
+Most of the USA participants \hh{report to have} an undergraduate degree or at least have some undergraduate courses.
+
+The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where \hh{the} proportion of female participants is lower.
+The distribution of education levels also differs slightly across countries for this age group.
 %Most of the participants from India are below age 40 while in United States the distribution of participants are similar beyond age 40.
@@
 
-A total of 1911 lineups were evaluated in the ten experiments.
-Each person evaluated at least 10 lineups except for experiment~9 where three lineups were evaluated by each person.
-A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
-Data from participants who got it wrong were removed, and then results for just this lineup were removed for all participants who responded correctly.
+A total of 1911 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where three lineups were evaluated by each person. 
+\hh{In order to ensure data quality, a test plot was shown to each observer, and the performance on this test plot was used for screening purposes.} %A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
+\hh{Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.}
+Data from participants who \hh{gave the wrong answer on the test plot were considered as gaming the system, and their data} were removed.
+\hh{Results from test plots were not considered in any of the analyses}.
  In some cases the demographic information were not provided by the participants.
 Also, for some ip address, the actual geographical locations could not be retrieved.
-This resulted some missing demographic information.
+This resulted in some missing demographic information.
 
@@
 
-Figure \ref{fig:demographic_effect} supports the model results.
+Figure~\ref{fig:demographic_effect} supports the model results.
 The average time taken to respond (natural log) and detection rate are computed by lineup for the different demographic factor levels, and displayed using boxplots.
@@
 %The distribution of time taken to evaluate a lineup is positively skewed.
-But in log scale it appears to be symmetric for all the demographic factor variables as we see in Figure \ref{fig:demographic_effect}.
+But in log scale it appears to be symmetric for all the demographic factor variables as we see in Figure~\ref{fig:demographic_effect}.
 Mean and medians are very similar as well.
@@
 Least square regression lines are fitted through the subject specific residuals.
-Subject specific slopes are much different in each of the three experiments as we see in Figure \ref{fig:learning_trend_time}.
+Subject specific slopes are much different in each of the three experiments as we see in Figure~\ref{fig:learning_trend_time}.
 Some subjects improved over attempts by taking less time in the later attempts while others got worse.
@@
 %A plot of residuals against the attempts $A_k$ should reveal the trend if there is any.
-Figure \ref{fig:learning_trend} shows such plots for three experiments naming experiment~5, 6 and 7.
+Figure~\ref{fig:learning_trend} shows such plots for three experiments naming experiment~5, 6 and 7.
 For all the experiments we see some increasing trend of the residuals.
@@
 We fit a mixed effect linear model with time taken considering lineup and subjects as random effect.
-The resulting residuals are plotted in Figure \ref{fig:learning_trend_time}.
+The resulting residuals are plotted in Figure~\ref{fig:learning_trend_time}.
 It appears that in the later attempts observers take less time to evaluate a lineup.
 Even though it does not tell much about improving their performance over attempts, but it does indicate the improvement of their evaluation skill.
-Adding this information with what we observe in Figure \ref{fig:learning_trend}, we can say that the observers' improved skills could not contribute to their performance.
+Adding this information with what we observe in Figure~\ref{fig:learning_trend}, we can say that the observers' improved skills could not contribute to their performance.
 It only help them finish the task faster in later attempts.
@@
 
-The detection rate for each data plot location is shown in Figure \ref{fig:location_effect} colored by null sets.
+The detection rate for each data plot location is shown in Figure~\ref{fig:location_effect} colored by null sets.
 The overall average detection rate for each location are shown using dashed lines.
@@
 In another set of null plots this may not happen making some lineups easier than others even though the actual data plot is the same.
-This pattern is evident in the figure as we see proportion correct for null plot 5 is consistently above the null plot 1 for each of the locations.
+This pattern is evident in the Figure~as we see proportion correct for null plot 5 is consistently above the null plot 1 for each of the locations.
 
@@
 
-%The sizes of the dots in  Figure \ref{fig:location_effect} represents the number of responses obtained for that location and null set.
+%The sizes of the dots in  Figure~\ref{fig:location_effect} represents the number of responses obtained for that location and null set.
 For some locations we have as many as 10 responses.
@@
 The locations 7,8,9,12,13,14 in the lineup are considered inner locations and the rest of the boundary locations are considered to be outer location.
-As we see in Figure \ref{fig:location_effect}, location 9, 12 are inside for Interaction effect and location 8, 12 are inside for Genotype.
+As we see in Figure~\ref{fig:location_effect}, location 9, 12 are inside for Interaction effect and location 8, 12 are inside for Genotype.
 It does not show any differences whether the actual plot is inside or outer border of the lineup.
@@
 This allowed observers to write whatever they think their reasons for choice are.
-Figure \ref{fig:wordle} shows the words used to explain their reasons for choice.
+Figure~\ref{fig:wordle} shows the words used to explain their reasons for choice.
 The most common words used to explain their choice are points and green which indicates the use of two important features of the grammar of graphics \citep{wilkinson:1999, hadley:2009}.
@@
 
-Figure \ref {fig:wordle} also shows some insight about peoples way of reading a plot.
+Figure~\ref {fig:wordle} also shows some insight about peoples way of reading a plot.
 We notice that variability in the data, geometric shapes and aesthetics used to generate plots and existence of any systematic pattern in the plot are some of the important features revealed from the figure.

### 2015-02-08 sociology_paper.tex
table 4 variance component question

@@
 To explore the significance of the demographic factors model \eqref{eqn:demographic_response} is fit to the data, with age, country, education and gender as fixed effects.
-To estimate the significance of the factors, reduced models are fit with that factor removed from the model.
+To estimate the significance of each of the factors, reduced models are fit with that factor removed from the model.
 Table ~\ref{tbl:anova_factor} summarizes the results.
-All of the demographic factors are significantly different in describing the the detection rate, except for gender.
+Except for gender, all of the demographic factors show significant differences in  detection rates.
 With respect to time taken, all of the demographic factors are significant.
@@
 The first level of each factor serves as the baseline for the model.
-The age group 36-40 significantly has a higher detection rate than the 18-25, and to a lesser extent this is also true for age groups 31-35 and above 50.
-Participants from India have a similar detection rate to those from the USA, but ones from the rest of the world had a significantly improved detection rate.
-Subjects who listed themselves as having a graduate degree had a significantly higher detection rate.
-For time to respond, all age groups were significantly slower than the 18-25 year olds.
-Participants from India were slower, and those from the rest of the world were faster than Americans.
-A higher degree corresponded to longer time, and males were slower than females!
+The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for age groups 31-35 and above 50.
+Participants from India have a similar detection rate to those from the USA, but \hh{participants} from the rest of the world show a significantly \hh{higher} detection rate.
+Subjects  \hh{reporting to have}  a graduate degree had a significantly higher detection rate.
+For time to respond, all age groups were significantly slower than  18-25 year olds.
+Participants from India were slower, while those from the rest of the world were faster than US Americans.
+A higher degree corresponded to longer time, and males were slower than females.
 
@@
 For gender no difference in performance is observed.
-Lineup variability is estimated to be very large for model \eqref{eqn:demographic_response}.}
+Lineup variability is estimated to be very large for model \eqref{eqn:demographic_response}.
+
+\hh{XXX -- the way that the random effects are listed is a bit misleading -- the first value is the {\bf variance} component, not the estimate of the standard deviation, while  the second value is its square root value, not the standard error of the estimate.
+I don't believe that lmer provides a standard error of the variance component estimates.}}
 \scalebox{0.86}{
@@
 Even though graduate degree is statistically significant, the largest difference in proportion correct is  0.045 which is very negligible.
-The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) \mm{should it be 5.259 instead ?}
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) \mm{should it be 5.259 instead ?} \hh{$\sigma$ is standard deviation, 5.259 is variance, I believe.}
    of lineup variability.}

### 2015-02-08 sociology_paper.tex
proofread, done!

@@
 Lineup variability is estimated to be very large for model \eqref{eqn:demographic_response}.
-
+%
 \hh{XXX -- the way that the random effects are listed is a bit misleading -- the first value is the {\bf variance} component, not the estimate of the standard deviation, while  the second value is its square root value, not the standard error of the estimate.
@@
 Even though graduate degree is statistically significant, the largest difference in proportion correct is  0.045 which is very negligible.
-The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$) \mm{should it be 5.259 instead ?} \hh{$\sigma$ is standard deviation, 5.259 is variance, I believe.}
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$)
    of lineup variability.}
@@
 \subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments 5, 6, and 7 separately.
-In this model attempt is fitted as a factor variable to allow for all possible non-linear learning trends.
+In this model, attempt is fitted as a factor variable to allow for all possible non-linear learning trends.
 It should be noted that we also examined an alternative to model \eqref{eqn:trend_response} where attempt was linearly fitted as a continuous covariate, but this also was not significant.
@@
 Table \ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model \eqref{eqn:trend_response} examining detection rate.
-There is some evidence of a learning curve only for experiment~6.
+Only for experiment~6 there is some evidence of a learning curve.
 There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.
@@
 
-To visualize how the detection rate effect over successive attempts, we fitted model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
+To visualize how  detection rates change over successive attempts, we fitted model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
 Least square regression lines were fitted through the subject specific residuals as shown in Figure~\ref{fig:learning_trend_response}.
@@
 Some subjects show improvement over time and some show the decrease in performance.
-Although, the model fit experiment~6 suggested a statistically significant learning curve the plots indicates that is minimal.
+Although, the model fit experiment~6 suggested a statistically significant learning curve the plots indicates that it is minimal.
 
@@
 Table \ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time} where response time is examined with respect to attempt.
-Attempt is modelled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated.
+Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated.
 Interestingly, time to respond significantly decreases as number of attempts increases, for all three experiments.
 The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments.
-The negative estimates suggest that on an average later attempt took less time for an evaluation.
-Even though observers did not improve the performance over attempt, they became efficient in responding faster.
+The negative estimates suggest that on average later attempts took less time.
+Even though observers did not improve in their performance in successive attempts, they became more efficient in their response.
 The parameter $\alpha_1$ for first attempt is also highly significant.
-The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more times than other attempts.
-It is because for initial attempt the observer might have gone through instructions and became familiar with the experimental environment.
-Each page in the web site contains information about choice reason and the observers's confidence level.
-Also, the first page asks for observer Identification to be typed.
-The later pages of the web site was similar just the lineup was changed.
+The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts.
+\hh{A possible reason is that an observer might have gone through instructions and familiarized themselves  with the experimental environment.} Each of the experiments \hh{asks observers to give a reason for their choice (out of a list of pre-defined answers}.
+\hh{For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).}  Later pages of the web site were identical  except for the lineup.
 Thus in the later attempts an observer does not need to spend any time for reading instructions.
@@
 \caption{Parameter estimates of model \eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
-Both fixed effect parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.}
+Both fixed effect parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.
+\hh{XXX same problem with the random effects as in tables 4 and 5.}}
 \scalebox{0.81}{
@@
 We observe some variability of performance for different null sets even though the same actual data plot was used for all these null sets.
-The lineup protocol is using a finite and small sample (19 in these experiments) of all possible null sets, so it is posisble to see some variability in performance depending on null set.
+The lineup protocol is using a finite and small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set.
 If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate.
 From the plot we can see that null set 3 appeared more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot.
-This null set effect was tested, and results in significant difference only for null set 3 of interaction lineup.
+This null set effect was tested, and significant difference \hh{were found} for null set 3 of the interaction lineup.
 The rest of the null sets don't show any differences at 1\% significance level.
@@
 In our study of the demographic effects on performance across a set of different experiments we found some statistically significant effects.
-Age group 36-40, Countries other than India and United states, People who have a graduate degree have a significantly higher detection rate.
+Age group 36-40, Countries other than India and United states, people who have a graduate degree have a significantly higher detection rate.
 Gender does not have any significant effect.
@@
 Some individuals improved the performances while others showed decrease in their performances over attempts.
-This could be interpreted as good, in the sense that it means that observers could realisitically be recruited from source like MTurk, and that substantial training is not necessary in order to obtain useful evaulations of lineups in practice.
+This could be interpreted as good, in the sense that it means that observers could realistically be recruited from sources like MTurk, and that substantial training is not necessary in order to obtain useful evaluations of lineups in practice.
 However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea.

### 2015-05-17 sociology_paper.tex
problem with standard errors of random effects fixed

@@
 The choice of plot primarily should be appropriate for the problem being investigated, as the data plot is the visual test statistic.
-When for example studying the association between two quantitative variables a scatterplot is usually optimal.
-\hh{XXX do we really want to say optimal?
-that sounds like we have an underlying criterion to optimize.
-XXX} Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.
+When for example studying the association between two quantitative variables a scatterplot is usually \hh{considered best practice}.
+Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.
 Some choices of visual test statistics provide better results than others when studying the same problem \citep{heike:2012}, because some plot types make it easier for certain structures to be seen.
@@
 %\subsection{Demographics} 
-\item{\bf Demographics:}  \hh{During the experiments } data on age, gender, education level and geographic location \hh{was collected}.
+\item{\bf Demographics:}  \hh{During each of the experiments } data on age, gender, education level and geographic location \hh{was collected}.
 Each observer self-reported gender, age in roughly five year intervals, education level as high school or less, some college courses, an undergraduate degree, some graduate courses or a graduate degree.
@@
 %
-\hh{XXX -- the way that the random effects are listed is a bit misleading -- the first value is the {\bf variance} component, not the estimate of the standard deviation, while  the second value is its square root value, not the standard error of the estimate.
-I don't believe that lmer provides a standard error of the variance component estimates.}}
+}
 \scalebox{0.86}{
@@
 \multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
-& lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
+& lineup($\sigma_{\ell}$) & 0.287 &  &  &  &   & 2.293 &  &  &  \\ 
  & Error($\sigma$) & 0.479 & &  &  &  & \mm{? 0.692}&& &\\ 
@@
 \multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
-  $\sigma^2_a$ & $<$0.001 & 0.017 &  &  &   & 0.001 & 0.034 &  &  &   & 0.027 & 0.163 &  &  \\ 
-  $\sigma^2_u$ & 0.720 & 0.848 &  &  &   & 0.815 & 0.903 &  &  &   & $<$0.001 & $<$0.001 &  &  \\ 
-  $\sigma^2_{\ell}$ & 2.178 & 1.476 &  &  &   & 2.009 & 1.418 &  &  &   & 10.980 & 3.314 &  &  \\ 
+  $\sigma^2_a$ &  0.017 &&  &  &   &  0.034 &&  &  &   &  0.163 &&  &  \\ 
+  $\sigma^2_u$  & 0.848 &&  &  &    & 0.903 &&  &  &   &  $<$0.001 &&  &  \\ 
+  $\sigma^2_{\ell}$  & 1.476 &&  &  &    & 1.418 &&  &  &   &  3.314 &&  &  \\ 
    \hline
@@
 Both fixed effect parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.
-\hh{XXX same problem with the random effects as in tables 4 and 5.}}
+}
 \scalebox{0.81}{
@@
 \multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
-  $\sigma^2_a$ & 0.001 & 0.027 &  &  &   & 0.002 & 0.045 &  &  &   & 0.002 & 0.049 &  &  \\ 
-  $\sigma^2_u$ & 0.259 & 0.509 &  &  &   & 0.245 & 0.495 &  &  &   & 0.134 & 0.366 &  &  \\ 
-  $\sigma^2_{\ell}$ & 0.008 & 0.091 &  &  &   & 0.040 & 0.199 &  &  &   & 0.055 & 0.235 &  &  \\ 
-  $\sigma^2$ & 0.211 & 0.460 &  &  &   & 0.251 & 0.501 &  &  &   & 0.206 & 0.454 &  &  \\ 
+  $\sigma^2_a$ &   0.027 &  &  &   &  & 0.045 &  &  &   &  & 0.049 &  &  \\ 
+  $\sigma^2_u$ &   0.509 &  &  &   &  & 0.495 &  &  &   &  & 0.366 &  &  \\ 
+  $\sigma^2_{\ell}$ &   0.091 &  &  &   &  & 0.199 &  &  &   &  & 0.235 &  &  \\ 
+  $\sigma^2$ &   0.460 &  &  &   &  & 0.501 &  &  &   &  & 0.454 &  & &  \\ 
    \hline

### 2015-05-17 sociology_paper.tex
typos

@@
 \item{\bf Individual Skill or Ability:} Each person may have different  aptitude for reading statistical plots, and their visual skill sets might be differently developed.
-We \hh{We can evaluate the effect of individuals' skills and abilities}  by letting multiple subjects evaluate the same lineups, which allows us to estimate, if some subjects consistently detect the actual data plot more often than others.
+\hh{We can evaluate the effect of individuals' skills and abilities}  by letting multiple subjects evaluate the same lineups, which allows us to estimate, if some subjects consistently detect the actual data plot more often than others.
  %For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.

### 2015-05-17 sociology_paper.tex
footnote for calculus

@@
 \begin {abstract}  
-Visual statistical inference is a way to determine significance of patterns found while exploring data.
-It is dependent on the evaluation of a lineup, of a data plot among a sample of null plots, by human observers.
+Visual statistical inference is a way to determine significance of patterns found while  exploring data with graphics.
+Visual inference relies on human observers inspecting a lineup: one  data plot randomly placed among a set of decoy plots.
 Each individual is different in their cognitive psychology and judiciousness, which can affect the visual inference.
-The usual way to estimate the effectiveness of a statistical test is its power.
-The estimate of power of a lineup can be controlled by combining evaluations from multiple observers.
-Factors that may also affect the power of visual inference are the observers' demographics, visual skills, and experience, the sample of null plots taken from the null distribution, the position of the data plot in the lineup, and the signal strength in the data.
+The usual way of estimating the effectiveness of a statistical test is its power.
+The estimate of power of a lineup is controlled by combining evaluations from multiple observers.
+Factors that may  affect the power of visual inference are the observers' demographics, visual skills, and experience, the sample of null plots taken from the null distribution, the position of the data plot in the lineup, and the signal strength in the data.
 This paper examines these factors.
@@
 The experiments suggest that individual skills vary substantially,  but demographics do not have a huge effect on performance.
-There is evidence that a learning effect exists but only in that observers get faster with repeated evaluations, but not more often correct.
+There is evidence that a learning effect exists but only in the sense that observers become faster with repeated evaluations, while accuracy stays the same.
 The placement of data plot in the lineup does not affect the inference.
@@
 
-The lineup protocol introduced in \citet{buja:2009} can be used to test the significance of \hh{graphical} findings in the exploratory data analysis process.
+The lineup protocol introduced in \citet{buja:2009} can be used to quantify the significance of \hh{graphical} findings in the exploratory data analysis process.
 This methodology is  part of what is called visual statistical inference, which has been developed further  and validated by \citet{majumder:2013} in simulation studies of head to head comparisons with conventional inference.
@@
 
-Figure~\ref{fig:tower} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are showing  data generated under a null model.
+Figure~\ref{fig:elect-1} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are showing  data generated under a null model.
 Which one of the 20 plots is the most different from the others?
-When asked this question, 12 \hh{out} of 72 observers picked the  plot \hh{of the observed data} (located in panel number $3^2+4$ in the lineup).
+When asked this question, 12 \hh{out} of 72 observers picked the  plot \hh{of the observed data}, located in panel number $3^2+4$ in the lineup \footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
 The corresponding $p$-value is \hh{0.0077}, indicating sufficient evidence to reject the null hypothesis.

### 2015-05-18 sociology_paper.tex
mini changes
typo

@@
 
-A total of 111 subjects were recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
+A total of 111 subjects was recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
 Each subject evaluated two lineups;  one for Interaction effect and the other for Genotype.
-In total there were 222 responses on 50 lineups.
+In total there were 222 responses for 50 lineups.
 The data on the test lineup were excluded from the analysis.
@@
 The detection rate for each data plot location is shown in Figure~\ref{fig:location_effect} colored by null sets.
-The overall average detection rate for each location are shown using dashed lines.
-Size of the point represents the number of responses.
+The overall average detection rate for each location is shown using dashed lines.
+Size of the points represent the number of responses.
 For some locations we have as many as 10 responses.
-For location 1, we did not have any response for null set 1 in one of the interaction lineups.
-
-We observe some variability of performance for different null sets even though the same actual data plot was used for all these null sets.
+For location 1, we did not have any responses for null set 1 in one of the interaction lineups.
+
+We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets.
 The lineup protocol is using a finite and small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set.
 If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate.
-From the plot we can see that null set 3 appeared more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot.
-This null set effect was tested, and significant difference \hh{were found} for null set 3 of the interaction lineup.
-The rest of the null sets don't show any differences at 1\% significance level.
+From the plot we can see that null set 3 appeared to be more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot.
+This null set effect was tested, and a significant difference \hh{was found} for null set 3 of the interaction lineup.
+The rest of the null sets do not show any differences at  a 1\% significance level.
 
@@
 Model \eqref{manova} was fit to the data to test if the detection rates were significantly different for the different locations.
-For this we use {\it manova()} function of {\it stats} package of \cite{R:2012}.
+For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R:2012}.
 The results are shown in Table \ref{tbl:manova}.
-The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05.
-This suggest that location of the actual data plot in the lineup does not significantly affect performance.
+The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05 suggesting that location of the actual data plot in the lineup does not significantly affect performance.
 
@@
  \cline{5-7}
- Effect & DF & Wilks & Approx. F & Numerator& Denumerator &Residual & $p$-value\\
-  \hline
-  Interaction& 3&0.07986 & 0.592 &15&5.92 & 6 & 0.8082 \\ 
+ Effect & DF & Wilks & Approx. F & Numerator& Denominator &Residual & $p$-value\\
+  \hline
+  Interaction& 3&0.07986 & 0.592 &15& \phantom{0}5.92 & 6 & 0.8082 \\ 
   Genotype & 4&0.01689 & 1.722 & 20&14.22 & 8 & 0.1488 \\ 

### 2015-05-18 sociology_paper.tex
location effect

@@
 \begin{document}
-
-% tableofcontents
+\tableofcontents
 
@@
 The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
-Two factors were of primary interest: genotype (call them RPA and EV) and the interaction between genotype and iron condition (sufficient or insufficient).
+Two factors were of primary interest a main effect for genotype and an interaction effect between genotype and treatment condition.
+
+%Two factors were of primary interest: genotype (call them RPA and EV) and the interaction between genotype and iron condition (sufficient or insufficient).
 The $2\times2$ factorial design provides four treatment conditions.
@@
 In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
-This overall significance is studied by visual inference.
-
-For the study, the routine processing of the data was conducted to find (1) the gene that exhibited the most significance difference in expression between genotypes, (2) the gene that exhibited the most different expression on RPA, but insignificant expression on EV.
+This overall significance is studied using visual inference  in \citet{tengfei:2013}.
+
+%For the study, the routine processing of the data was conducted to find (1) the gene that exhibited the most significance difference in expression between genotypes, (2) the gene that exhibited the most different expression on RPA, but insignificant expression on EV.
 Two different types of plots would be made to examine this.
@@
 
-Multiple lineups of the data were produced.
-The actual plot was randomly placed in five different locations in a lineup of size 20.
-The locations were  1, 8, 12, 17, 20 for genotype effect (1) and 2, 9, 12, 16, 20 for interaction effect (2).
-Five different sets of null plots were used to produce 5 lineups for each location position, creating a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
+In order to study the effect that  location has on evaluating lineups, we used multiple lineups for each of the data plots.
+For each data plot, five sets of null plots were generated and the actual plot was randomly placed in one of five different locations in a lineup of size 20.
+For the genotype effect, the locations were  1, 8, 12, 17, 20 and locations 2, 9, 12, 16, 20 were used for the interaction effect.
+%Five different sets of null plots were used to produce 5 lineups for each location position, creating
+Overall this created a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
 Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to help clean the data.
@@
 
-\subsection{Lineup Design}
+\subsection{Location Effect}
 
@@
 
-The detection rate for each data plot location is shown in Figure~\ref{fig:location_effect} colored by null sets.
+We are investigating location effect in two ways: (1) location effect of each individual plot location, and (2) location effect of inside/outside panel positions.
+
+Figure~\ref{fig:location_effect} shows an overview of detection rates for both of the effects studies in experiment~9.
+Each dot corresponds to one lineup.
+The size of each dot gives the number of evaluations.
+Lineups with  the same set of null plots are connected by lines.
+%The detection rate for each data plot location is shown in colored by null sets. 
 The overall average detection rate for each location is shown using dashed lines.
@@
 
-Model \eqref{manova} was fit to the data to test if the detection rates were significantly different for the different locations.
+Model \eqref{manova} was fit to the data to test if  detection rates were significantly different for the different locations.
 For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R:2012}.
@@
 
-We also examined whether the proportion of correct responses differs if the actual plot is on the outer boundary or in the inner locations.
+For the second approach in investigating the location effect, we distinguish between inner panels 7,8,9,12,13,14 and outer panels (any panel with an outside edge).  
+%We also examined whether the proportion of correct responses differs if the actual plot is on the outer boundary or in the inner locations.
 The locations 7,8,9,12,13,14 in the lineup are considered inner locations and the rest of the boundary locations are considered to be outer location.
-As we see in Figure~\ref{fig:location_effect}, location 9, 12 are inside for Interaction effect and location 8, 12 are inside for Genotype.
-It does not show any differences whether the actual plot is inside or outer border of the lineup.
-We also fitted model \eqref{manova} with two locations, inner and outer, as covariate and observed no significant differences.
+From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype. 
+Model \eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.  
+%It does not show any differences whether the actual plot is inside or outer border of the lineup.
 

### 2015-05-18 sociology_paper.tex
references in intro

@@
 %      \includegraphics[width=0.99\textwidth]{electoral-1-1.pdf} 
-   \caption{ \label{fig:elect-1} Which one of the plots is the most different from the others?}
+   \caption{ \label{fig:elect} Which one of the plots is the most different from the others?}
 \end{figure}
 
-Figure~\ref{fig:elect-1} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are showing  data generated under a null model.
+Figure~\ref{fig:elect} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are showing  data generated under a null model.
 Which one of the 20 plots is the most different from the others?
@@
 The data consists of the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
-Each panel in Figure~\ref{fig:elect-1} shows an `electoral building' \citep{mosley2010} where each state in the union is represented by a rectangle.
+Each panel in Figure~\ref{fig:elect} shows an `electoral building' \citep{mosley2010} where each state in the union is represented by a rectangle.
 This difference is plotted horizontally, and the height of each box corresponds to the state's electoral votes.

### 2015-05-19 sociology_paper.tex
checked through changes to intro

@@
 \begin{document}
-\tableofcontents
+%\tableofcontents
 
@@
 
-{\bf Keywords: \sf statistical graphics, non-parametric test, cognitive \hh{psychology}, data visualization, exploratory data analysis, data mining, visual analytics.} 
+{\bf Keywords: \sf statistical graphics, non-parametric test, cognitive psychology, data visualization, exploratory data analysis, data mining, visual analytics.} 
 \end {abstract}
@@
 
-The lineup protocol introduced in \citet{buja:2009} can be used to quantify the significance of \hh{graphical} findings in the exploratory data analysis process.
+The lineup protocol introduced in \citet{buja:2009} can be used to quantify the significance of graphical findings in the exploratory data analysis process.
 This methodology is  part of what is called visual statistical inference, which has been developed further  and validated by \citet{majumder:2013} in simulation studies of head to head comparisons with conventional inference.
@@
 %      \includegraphics[width=0.99\textwidth]{electoral-1-1.pdf} 
-   \caption{ \label{fig:elect} Which one of the plots is the most different from the others?}
+   \caption{ \label{fig:elect} Which one of the plots, labelled 1 through 20 from top left to bottom right, is the most different from the others?}
 \end{figure}
@@
 Which one of the 20 plots is the most different from the others?
-When asked this question, 12 \hh{out} of 72 observers picked the  plot \hh{of the observed data}, located in panel number $3^2+4$ in the lineup \footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
-The corresponding $p$-value is \hh{0.0077}, indicating sufficient evidence to reject the null hypothesis.
+When asked this question, 12 out of 72 observers picked the  plot of the observed data, located in panel number $3^2+4$ in the lineup \footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
+The corresponding $p$-value is 0.0077, indicating sufficient evidence to reject the null hypothesis.
 % , with reasons given being 'asymmetry' (36\%), 'trend' (26\%) or `outliers' (16\%).
 
-\hh{How do we interpret this finding, though?} For that, we need to know the context of the data and we need to have more information about \hh{how the null plots were generated}. 
+How do we interpret this finding, though? For that, we need to know the context of the data and we need to have more information about how the null plots were generated. 
 This particular example investigates the results from the 2012 US presidential election in comparison to  poll results just prior to the election. 
@@
 The null hypothesis is ``that the election results were consistent with the polls''.
-The polling results provides the null model from which data is simulated.
-\hh{A normal model with mean and standard deviation based on a poll and its margin of error is used to} simulate different scenarios that might have resulted on election day, if the polls were on target.
+The polling results provide the null model from which data is simulated.
+A normal model with mean and standard deviation based on a poll and its margin of error is used to simulate different scenarios that might have resulted on election day, if the polls were on target.
 A null data set is generated as a set of draws from this model.
@@
 If the null hypothesis is true, the  plot of the actual data should look  like any of the other plots, and not be identifiable by an observer.
-Figure~\ref{fig:tower} shows the plot of the electoral building with added context information and labels.
+Figure~\ref{fig:tower} shows the plot of the electoral building with the additional context information and labels.
  %A party needs 270 electoral votes or more to win the presidential election. 
@@
 Presidential Election (left).
-On the right two histograms of 10,000 simulations each based on polling averages from two different sources.
-For the histogram on the top, the $p$-value of observing results as extreme as  the  2012 U.S.
-election results based on the bootstrap is 0.0533 (with Bootstrap standard error of 0.002), making the election results almost significantly different from the polls (data source for poll aggregator 1 is Freedom`s Lighthouse Averages).
-There is no indication of any inconsistency between polls and election results based on the bootstrap simulation below (data source for poll aggregator 2 is RealClearPolitics Averages).
-The lineups are based on the top source.}
+On the right two histograms showing simulated poll results for the Democratic presidential candidate using  polling averages collected at two different sources, Freedom's Lighthouse Averages (top), RealClearPolitics (bottom).
+If the polls matched the electoral result the simulated results would be centered at the electoral result.
+This happens for the RealClearPolitics polls (bottom) but not for Freedom's Lighthouse.
+For the latter only 5.33\% of simulated results were as high as the actual election result, which converts to a $p$-value of 0.0533, almost a statistically significant difference.
+It hints at the polls from this source under-estimating the vote for Obama.
+This source was used to generate the null plots in the lineup in Figure \ref{fig:elect}.
+The $p$-value from the lineup protocol, was 0.0077, which more strongly concludes that these polls mismatched the actual result.}
 \end{figure}
@@
 A lineup can be evaluated by a single person or multiple observers.
-A distribution \hh{similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario,} is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
-In order to avoid expectation errors \citep{meilgaard}, observers should not be aware of the data that constitutes a lineup, and should not have seen the actual data plot before seeing the lineup.
-This is the reason that in \hh{our first example we explained the context only after presenting the lineup.}%the election example, above, the scenario was explained after the lineup question, in the text.
-
-\hh{When presenting a lineup to an observer we should also phrase the accompanying question in as general a manner as possible, }
+A distribution similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario, is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
+To avoid expectation errors \citep{meilgaard}, particularly in emotionally charged areas such as election results, observers should not be aware of the data that constitutes a lineup, and should not have seen the actual data plot before seeing the lineup.
+This is the reason that we explained the context of the election problem only after you had evaluated the lineup.%the election example, above, the scenario was explained after the lineup question, in the text.
+
+When presenting a lineup to an observer  the accompanying question should be phrased in as general a manner as possible, 
 %The question that is asked of the observer should be as general as possible, 
-effectively asking the observer to pick the plot that is \hh{most} different, and allowing \hh{observers} to provide their reasons for their choice. 
-\hh{However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments  in  \cite{majumder:2013}, these questions have to be phrased much more specifically, in order to make the situations more comparable.
-}
+effectively asking the observer to pick the plot that is most different, and allowing observers to provide their reasons for their choice.
+This effectively allows all possible deviations from the null hypothesis to be detected.
+However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments  in  \cite{majumder:2013}, these questions have to be phrased much more specifically, in order to make the situations more comparable.
 %In some of the studies, the ones described in  \cite{majumder:2013} very specific questions were asked, because the experiments were being conducted to compare results from the lineup protocol with those of conventional tests.
@@
 The type of plot, showing two (modified) stacked bar charts in different colors should suggest to the observer that the interesting feature is the difference between the two heights.
-\hh{Most observers based their decision on this -- besides  `asymmetry' between the towers (a reason suggested by us), most free-form responses for the reason of choice describe a comparison between the red and blue towers.
+Most observers based their decision on this -- besides  `asymmetry' between the towers (a reason suggested by us), most free-form responses for the reason of choice describe a comparison between the red and blue towers.
 In the responses for this lineup we see something that is very typical of responses in lineups: it is only a few panels from the lineup that observers pick, while most panels do not get picked or get picked only a few times.
@@
 From the panels picked, we see that the top three choices  are all extreme with respect to heights of blue versus red towers:  both panels \#8 and \#13 show a large difference between the height of the towers, while panel \#4 is the only panel, in which the red tower is higher than the blue one.
-}
 % results from the turk study:
@@
 
- \hh{If we consider this difference to be the main objective of the visualization, a more focused question, that aligns more with a hypothesis test on the difference  in electoral votes  between parties, would}
- have been ``Which plot shows the biggest height difference between the two towers?'' %except that this tailors the inference to a specific feature which does not match the null hypothesis of interest.
+% \hh{If we consider this difference to be the main objective of the visualization, a more focused question, that aligns more with a hypothesis test on the difference  in electoral votes  between parties, would} have been ``Which plot shows the biggest height difference between the two towers?'' %except that this tailors the inference to a specific feature which does not match the null hypothesis of interest.
 
@@
 
-\hh{The lineup protocols allows us to calculate all relevant properties that we are familiar with from conventional statistical tests.
+The lineup protocols allows us to calculate all relevant properties that we are familiar with from conventional statistical tests.
 In particular, the power of a lineup is calculated as the {\it detection rate} at which observers identify the actual data plot.
 Visual power depends unlike power in conventional tests not only on the strength of the signal, but also on individuals' ability.
-}
 %This proportion is associated with the Type~I error of the test.
@@
 When multiple observers evaluate a lineup, the detection rate can be used to estimate the power.
-The ability of individual observers varies, and \hh{the effects that might influence ability are the focus} of this paper.
+The ability of individual observers varies, and the effects that might influence this ability are the focus of this paper.
 
@@
 
-\hh{We have conducted a series of experiments (see Table \ref{tbl:visual_stat} for an overview)}
+We have conducted a series of experiments (see Table \ref{tbl:visual_stat} for an overview)
 %There have been ten experiments conducted 
-using Amazon's Mechanical Turk~\citep{turk} (MTurk) that \hh{provide a basis for evaluating} 
+using Amazon's Mechanical Turk~\citep{turk} (MTurk) that provide a basis for evaluating
 %are being used to evaluate 
 the effects of different covariates on an individual's ability to evaluate lineups. 
-\hh{In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.  }
+In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.  
  %Each of these experiments collected demographic details of the subjects. 
- Experiments~5, 6 and 7 can be used \hh{to investigate short term learning trends}.
+ Experiments~5, 6 and 7 can be used to investigate short term learning trends.
 Experiment~9 was also designed to check whether the  positioning of the  data plot in the lineup had an effect on detection rate.

### 2015-05-19 sociology_paper.tex
checked through changes completely

@@
 \centering 
-\caption{Overview of 10 different Turk experiments, \hh{in which} data was \hh{collected} to study human factor effects.
+\caption{Overview of 10 different Turk experiments, from which data was collected to study human factor effects.
 All of the experimental data was used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.}
@@
 %There have been ten experiments conducted 
-using Amazon's Mechanical Turk~\citep{turk} (MTurk) that provide a basis for evaluating
+using Amazon's Mechanical Turk~\citep{turk} (MTurk) for a variety to purposes: validating the protocol against existing tests, comparing plot designs, and evaluating structure in data analysis problems.
+In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.
+Mining this data will provide a basis for evaluating
 %are being used to evaluate 
 the effects of different covariates on an individual's ability to evaluate lineups.
-In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.  
- %Each of these experiments collected demographic details of the subjects. 
 Experiments~5, 6 and 7 can be used to investigate short term learning trends.
 Experiment~9 was also designed to check whether the  positioning of the  data plot in the lineup had an effect on detection rate.
-
- Section \ref{sec:factor_performance} discusses human factors that may affect the performance of the observer.
-Section \ref{sec:exp_design} describes the methods used to assess the effects, and Section \ref{sec:result_socio} describes the results.
+ %Each of these experiments collected demographic details of the subjects. 
+Section \ref{sec:factor_performance} discusses human factors that we expect may affect the performance of the observer.
+Section \ref{sec:exp_design} describes the methods used to assess the effects, and Section \ref{sec:result_socio} describes the results on what we have learned about human factors affecting lineup evaluation.
 
@@
 
-\hh{Visual statistical inference relies not only on the strength of the signal in the data, but also on how this information is presented in the plot, on the lineup design and on the abilities of human observers.} It is important to understand how these factors might affect results.
-\hh{Here, we provide a brief discussion }%This section provides a brief description
+Visual statistical inference relies not only on the strength of the signal in the data, but also on how this information is presented in the plot, on the lineup design and on the abilities of human observers.
+It is important to understand how these factors might affect results.
+Here, we provide a brief discussion %This section provides a brief description
 of the factors that are expected to have some impact on the performance of visual statistical inference.
@@
 The choice of plot primarily should be appropriate for the problem being investigated, as the data plot is the visual test statistic.
-When for example studying the association between two quantitative variables a scatterplot is usually \hh{considered best practice}.
+When for example studying the association between two quantitative variables a scatterplot is usually considered best practice.
 Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.
@@
 Simulated data was generated for all of the experiments, except exps.~4 and~9, enabling the study of signal strength and actual data plot detection.
-\hh{In experiment~4 we took a different approach by varying the sampling size, which also allowed us to investigate the effect a design had on detection rate while varying signal/noise ratio.} Primarily the results were as expected, that as the signal strength increased observers more frequently picked the actual data plot as different from the null plots.
+In experiment~4 we took a different approach by varying the sampling size, which also allowed us to investigate the effect a design had on detection rate while varying signal/noise ratio.
+Primarily the results were as expected, that as the signal strength increased observers more frequently picked the actual data plot as different from the null plots.
 
@@
 In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted.
-This was important \hh{to allow a direct comparison of the results to results from conventional testing.
-}%in order to make a comparison of the results.
+This was important to allow a direct comparison of the results to those from conventional testing.
+%in order to make a comparison of the results.
 For latter studies (4, 7, 10), questions posed to the observer were more general. Table \ref{tbl:visual_stat} lists the questions used for each experiment.
@@
 %\subsection{Demographics} 
-\item{\bf Demographics:}  \hh{During each of the experiments } data on age, gender, education level and geographic location \hh{was collected}.
+\item{\bf Demographics:}  During each of the experiments, data on age, gender, education level and geographic location was collected.
 Each observer self-reported gender, age in roughly five year intervals, education level as high school or less, some college courses, an undergraduate degree, some graduate courses or a graduate degree.
@@
 %\subsection{Location of Actual Data Plot in the Lineup}
-\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 \hh{panels} is used.
+\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 panels is used.
 A random number generator is used to determine the position where the actual data plot is placed in each lineup.
@@
 \item{\bf Individual Skill or Ability:} Each person may have different  aptitude for reading statistical plots, and their visual skill sets might be differently developed.
-\hh{We can evaluate the effect of individuals' skills and abilities}  by letting multiple subjects evaluate the same lineups, which allows us to estimate, if some subjects consistently detect the actual data plot more often than others.
+We can examine the effect of individuals' skills and abilities because multiple subjects evaluated the same lineups, which allows us to estimate, if some subjects consistently detect the actual data plot more often than others.
  %For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.
@@
 % It is interesting that there were large number of female participants even though there were a lot of participants  from developing countries.
-\hh{The large number of female participants from all countries was a pleasant surprise to us.}
+The large number of female participants from all countries was a pleasant surprise to us.
 
@@
 
-The largest number of participants \hh{falls within the} age group of 18 to 25, with the majority between 18 to 35.
-\hh{ Many older participants took part in the experiments.} %Interestingly, there are many participants in the older age groups as well.
+The largest number of participants falls within the age group of 18 to 25, with the majority between 18 to 35.
+Many older participants also took part in the experiments.%Interestingly, there are many participants in the older age groups as well.
 The United States shows a more uniform participation beyond age 30, Indian participants are primarily younger.
 
-In terms of academic background, \hh{a majority of participants has a graduate degree} (a total of 902 participants, about 38.51\%).
-Examining the location in conjunction with the academic background reveals that there are more Indian participants that \hh{report to have } a graduate degree.
-In retrospect \hh{we realized this to be a culturally based misunderstanding}, as any university degree in India is considered to be a graduate degree, while the North American definition of graduate degree refers to a Masters level education.
-Most of the USA participants \hh{report to have} an undergraduate degree or at least have some undergraduate courses.
-
-The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where \hh{the} proportion of female participants is lower.
+In terms of academic background, the majority of participants had a graduate degree (a total of 902 participants, about 38.51\%).
+Examining the participants' location in conjunction with the academic background reveals that there are more Indian participants that report to have a graduate degree.
+In retrospect we realized this to be a culturally based misunderstanding, as any university degree in India is considered to be a graduate degree, while the North American the definition of graduate degree refers to a Masters level, or above, education.
+Most of the USA participants report to have an undergraduate degree or at least have some undergraduate courses.
+
+The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where the proportion of female participants is lower.
 The distribution of education levels also differs slightly across countries for this age group.
@@
 A total of 1911 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where three lineups were evaluated by each person. 
-\hh{In order to ensure data quality, a test plot was shown to each observer, and the performance on this test plot was used for screening purposes.} %A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
-\hh{Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.}
-Data from participants who \hh{gave the wrong answer on the test plot were considered as gaming the system, and their data} were removed.
-\hh{Results from test plots were not considered in any of the analyses}.
+In each of the experiments, to ensure quality evaluations, a test plot was shown to each observer, and the performance on this test plot was used for screening purposes.%A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
+Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.
+Data from participants who gave the wrong answer on the test plot were considered as gaming the system, and their data were removed.
+Results from test plots were not considered in any of the analyses.
  In some cases the demographic information were not provided by the participants.
@@
 The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for age groups 31-35 and above 50.
-Participants from India have a similar detection rate to those from the USA, but \hh{participants} from the rest of the world show a significantly \hh{higher} detection rate.
-Subjects  \hh{reporting to have}  a graduate degree had a significantly higher detection rate.
+Participants from India have a similar detection rate to those from the USA, but participants from the rest of the world show a significantly higher detection rate.
+Subjects  reporting to have  a graduate degree had a significantly higher detection rate.
 For time to respond, all age groups were significantly slower than  18-25 year olds.
@@
 The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts.
-\hh{A possible reason is that an observer might have gone through instructions and familiarized themselves  with the experimental environment.} Each of the experiments \hh{asks observers to give a reason for their choice (out of a list of pre-defined answers}.
-\hh{For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).}  Later pages of the web site were identical  except for the lineup.
+Most observers, before evaluating the first lineup, carefully read through instructions and familiarize themselves  with the types of plots used in the experiment, which would inflate the time taken on the first lineup.
+Each of the experiments asks observers to give a reason for their choice (out of a list of pre-defined answers.
+For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).
+Later pages of the web site were identical  except for the lineup.
 Thus in the later attempts an observer does not need to spend any time for reading instructions.
@@
 From the plot we can see that null set 3 appeared to be more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot.
-This null set effect was tested, and a significant difference \hh{was found} for null set 3 of the interaction lineup.
+This null set effect was tested, and a significant difference was found for null set 3 of the interaction lineup.
 The rest of the null sets do not show any differences at  a 1\% significance level.
@@
 
-For the second approach in investigating the location effect, we distinguish between inner panels 7,8,9,12,13,14 and outer panels (any panel with an outside edge).  
+For the second approach in investigating the location effect, we distinguish between inner panels 7, 8, 9, 12, 13, 14 and outer panels (any panel with an outside edge).  
 %We also examined whether the proportion of correct responses differs if the actual plot is on the outer boundary or in the inner locations.

### 2015-08-28 sociology_paper.tex
update references

@@
 %\documentclass{article}
-\documentclass[10pt]{article}
+\documentclass[10pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 \usepackage{times}
@@
 \pdfminorversion=4 % as instructed by JASA file upload
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is an Assistant Professor in the Department of Mathematics, University of Nebraska at Omaha, NE 68182 (e-mail: mmajumder@unomaha.edu), Heike Hofmann and Dianne Cook are Professors in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210.
+\thanks{Mahbubul Majumder is an Assistant Professor in the Department of Mathematics, University of Nebraska at Omaha, NE 68182 (e-mail: mmajumder@unomaha.edu), Heike Hofmann is Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210, and  Dianne Cook is Professor in the Department of Econometrics and Business Statistics, Monash University, Australia.
 This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
@@
 %
-% \footnote{Signif. codes: 0 Ô***Õ 0.001 Ô**Õ 0.01 Ô*Õ 0.05 Ô.Õ 0.1}
+% \footnote{Signif. codes: 0 ?***? 0.001 ?**? 0.01 ?*? 0.05 ?.? 0.1}
 

### 2015-08-28 sociology_paper.Rnw
updating references
typographic changes

@@
 \usepackage[colorlinks=true, citecolor=blue, linkcolor=blue]{hyperref}
-\usepackage{amsmath, amssymb, fullpage, amsthm, array,,graphicx,asa, url}
+\usepackage{amsmath, amssymb, fullpage, amsthm, array, graphicx, asa, url}
 %\usepackage[dvips]{graphics}
@@
 There is evidence that a learning effect exists but only in the sense that observers become faster with repeated evaluations, while accuracy stays the same.
-The placement of data plot in the lineup does not affect the inference.
+The placement of data plot in the lineup does not affect inferential results.
 
@@
 It hints at the polls from this source under-estimating the vote for Obama.
-This source was used to generate the null plots in the lineup in Figure \ref{fig:elect}.
+This source was used to generate the null plots in the lineup in Figure~\ref{fig:elect}.
 The $p$-value from the lineup protocol, was 0.0077, which more strongly concludes that these polls mismatched the actual result.}
@@
 
-%Some of the factors that may affect the performance of the observer are discussed in Section \ref{sec:factor_performance}.
+%Some of the factors that may affect the performance of the observer are discussed in Section~\ref{sec:factor_performance}.
 In this paper we focus on estimating the learning effect of the observer in terms of both proportion correct and time taken to evaluate a lineup through the attempts made during the evaluation of multiple plots.
@@
 
-We have conducted a series of experiments (see Table \ref{tbl:visual_stat} for an overview)
+We have conducted a series of experiments (see Table~\ref{tbl:visual_stat} for an overview)
 %There have been ten experiments conducted 
@@
 the effects of different covariates on an individual's ability to evaluate lineups.
-Experiments~5, 6 and 7 can be used to investigate short term learning trends.
-Experiment~9 was also designed to check whether the  positioning of the  data plot in the lineup had an effect on detection rate.
+The design of some of the experimets allow a further investigation into aspects beyond participants' demographics: experiments~5, 6 and 7 can be used to investigate short term learning trends, while experiment~9 allows insight on whether the  positioning of the  data plot in the lineup has an effect on detection rate.
  %Each of these experiments collected demographic details of the subjects. 
-Section \ref{sec:factor_performance} discusses human factors that we expect may affect the performance of the observer.
-Section \ref{sec:exp_design} describes the methods used to assess the effects, and Section \ref{sec:result_socio} describes the results on what we have learned about human factors affecting lineup evaluation.
+Section~\ref{sec:factor_performance} discusses human factors that we expect to be influential on the performance of the observer.
+Section~\ref{sec:exp_design} describes the methods used to assess the effects, and Section~\ref{sec:result_socio} describes the results on what we have learned about human factors affecting lineup evaluation.
 
@@
 
-Table \ref{tbl:visual_stat} shows the large variety of visual test statistics used in the Turk studies completed so far.
+Table~\ref{tbl:visual_stat} shows the large variety of visual test statistics used in the Turk studies completed so far.
 These experiments were conducted to examine many different visual tasks.
@@
 %in order to make a comparison of the results.
-For latter studies (4, 7, 10), questions posed to the observer were more general. Table \ref{tbl:visual_stat} lists the questions used for each experiment.
+For latter studies (4, 7, 10), questions posed to the observer were more general. Table~\ref{tbl:visual_stat} lists the questions used for each experiment.
 
@@
 This question should provide the observer a little clue so that the answer reflects the hypothesized patterns in the actual plot.
-For example Table \ref{tbl:visual_stat} shows the questions asked for the simulation experiments considered for this study.
+For example Table~\ref{tbl:visual_stat} shows the questions asked for the simulation experiments considered for this study.
 Notice that for case 1 if the observer can identify the actual plot in the lineup that should indicate that the plot chosen has the most vertical difference between groups A and B which is exactly what the researchers intend to examine.
@@
 
-%While evaluating a lineup, the performance of the observer may depend on the factors described in Section \ref{sec:factor_performance}. 
+%While evaluating a lineup, the performance of the observer may depend on the factors described in Section~\ref{sec:factor_performance}. 
 Two of the factors, signal in the data and individual abilities, were studied in \cite{majumder:2013}.
@@
 
-\subsection{Demographic Factors}  %In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table \ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
+\subsection{Demographic Factors}  %In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table~\ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
 \noindent
-For all  ten experiments shown in Table \ref{tbl:visual_stat}, the following demographic information was collected from subjects:
+For all  ten experiments shown in Table~\ref{tbl:visual_stat}, the following demographic information was collected from subjects:
 
@@
 
-To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2011}.
-To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
+To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2015, lme4:paper}.
+To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used to compute $Z$ scores  as the ratio of estimates to  estimated standard errors.
 
@@
 %
-%The inverse link function, $g^{-1}(.)$, from  Equation \ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k$th attempt by a single observer as
+%The inverse link function, $g^{-1}(.)$, from  Equation~\ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k$th attempt by a single observer as
 %\begin{equation} \label{eqn:trend_power}
@@
 %
-%Equation \ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
+%Equation~\ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
 %\begin{equation} \label{eqn:trend_time_est}
@@
 This provides a diverse pool of participants.
-The diversity in not only  geographic but also in gender, age group and education level as can be seen in Table \ref{tbl:demographics}.
+The diversity in not only  geographic but also in gender, age group and education level as can be seen in Table~\ref{tbl:demographics}.
 % It is interesting that there were large number of female participants even though there were a lot of participants  from developing countries.
@@
 
-Parameter estimates from the model fits are shown in Table \ref{tbl:model_result_demographics}.
+Parameter estimates from the model fits are shown in Table~\ref{tbl:model_result_demographics}.
 The first level of each factor serves as the baseline for the model.
@@
   \hline
-Demographic&& \multicolumn{4}{c} {Log Time (model \ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model \ref{eqn:demographic_response})} \\
+Demographic&& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})} \\
 
@@
 In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond.
-In Table \ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
+In Table~\ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
 This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
@@
 
-%To further illustrate how the individual factor levels differ from each other  the detailed results of models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table \ref{tbl:model_result_demographics}.
+%To further illustrate how the individual factor levels differ from each other  the detailed results of models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table~\ref{tbl:model_result_demographics}.
 For parameter estimation the first factor levels shown in the Table are  used as the base line.
@@
 For country, the rest of the world is different from USA but India appears to be similar to USA.
-We see from Table \ref{tbl:demographics} that most of the participants (about 90\%) are from India and USA.
+We see from Table~\ref{tbl:demographics} that most of the participants (about 90\%) are from India and USA.
 The rest of the 10\% data are from 76 different countries.
@@
 
-%In Table \ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
+%In Table~\ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
 This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
@@
 
-Table \ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model \eqref{eqn:trend_response} examining detection rate.
+Table~\ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model \eqref{eqn:trend_response} examining detection rate.
 Only for experiment~6 there is some evidence of a learning curve.
@@
 
-%The results are shown in Tables \ref{tbl:model_result_response} \ref{tbl:model_result_time}.
+%The results are shown in Tables~\ref{tbl:model_result_response}~\ref{tbl:model_result_time}.
 
@@
 To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
-The results are shown in Table \ref{tbl:model_result_time}.
+The results are shown in Table~\ref{tbl:model_result_time}.
 Both fixed effect parameters of covariate Attempt ($\alpha_1$ and $\alpha$) are highly significant.
@@
 
-Table \ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time} where response time is examined with respect to attempt.
+Table~\ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time} where response time is examined with respect to attempt.
 Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated.
@@
 For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R:2012}.
-The results are shown in Table \ref{tbl:manova}.
+The results are shown in Table~\ref{tbl:manova}.
 The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05 suggesting that location of the actual data plot in the lineup does not significantly affect performance.
@@
 \section{Electoral Building Lineups and Results}
-%Figures~\ref{fig:elect-1} and \ref{fig:elect.2} through~\ref{fig:elect.5}
+%Figures~\ref{fig:elect-1} and~\ref{fig:elect.2} through~\ref{fig:elect.5}
 

### 2015-08-28 sociology_paper.Rnw
small edits

@@
 To create a lineup, this plot, called the actual data plot, is placed in a layout of null plots.
-The null plots  are generated from the model specified by a null hypothesis, essentially describing what the plot might look like if the data had no structure.
-An observer is asked to evaluate the lineup.
-If the  actual data plot is detected by the observer, the null hypothesis is rejected.
+The null plots  are generated from the model specified by a null hypothesis.
+Commonly, this null hypothesis assumes independence/no structure in the data.
+An observer is then asked to evaluate the lineup.
+If the  actual data plot is detected by the observer, this counts as evidence against the null hypothesis.
+Based on the number of evaluations by independent observers and the number of times that the data plot is identified, we are then able to evaluate the significance of the graphical relationship in the same way and the same rigor as a conventional hypothesis test.
+% is rejected.
 This means that the structure in the actual data plot has significant structure, a pattern that is not simply due to randomness.
@@
 Which one of the 20 plots is the most different from the others?
-When asked this question, 12 out of 72 observers picked the  plot of the observed data, located in panel number $3^2+4$ in the lineup \footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
+When asked this question, 12 out of 72 observers picked the  plot of the observed data, located in panel number $4^2-3$ in the lineup \footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
 The corresponding $p$-value is 0.0077, indicating sufficient evidence to reject the null hypothesis.
@@
 Each panel in Figure~\ref{fig:elect} shows an `electoral building' \citep{mosley2010} where each state in the union is represented by a rectangle.
-This difference is plotted horizontally, and the height of each box corresponds to the state's electoral votes.
+The difference in poll results is plotted horizontally, while the height of each box corresponds to the number of that state's electoral votes.
 Color indicates party affiliation.
-The null hypothesis is ``that the election results were consistent with the polls''.
-The polling results provide the null model from which data is simulated.
+The null hypothesis is that ``the election results were consistent with the polls''.
+Therefore the polling results provide the null model from which data is simulated.
 A normal model with mean and standard deviation based on a poll and its margin of error is used to simulate different scenarios that might have resulted on election day, if the polls were on target.
-A null data set is generated as a set of draws from this model.
+Each null data set is generated as a set of draws from this model.
 % normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
@@
 A distribution similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario, is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
-To avoid expectation errors \citep{meilgaard}, particularly in emotionally charged areas such as election results, observers should not be aware of the data that constitutes a lineup, and should not have seen the actual data plot before seeing the lineup.
-This is the reason that we explained the context of the election problem only after you had evaluated the lineup.%the election example, above, the scenario was explained after the lineup question, in the text.
+To avoid expectation errors \citep{meilgaard}, particularly in emotionally charged areas such as election results, observers should not be aware of the data that constitutes a lineup, and should therefore not have seen the actual data plot before inspecting the lineup.
+This is the reason that we explained the context of the election problem only after you had a chance to evaluate the lineup.%the election example, above, the scenario was explained after the lineup question, in the text.
 
@@
 %The question that is asked of the observer should be as general as possible, 
-effectively asking the observer to pick the plot that is most different, and allowing observers to provide their reasons for their choice.
-This effectively allows all possible deviations from the null hypothesis to be detected.
-However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments  in  \cite{majumder:2013}, these questions have to be phrased much more specifically, in order to make the situations more comparable.
+effectively asking the observer to pick the plot that is most different, and allowing observers to provide their own reasons for their choice.
+This ensures all possible deviations from the null hypothesis to be detected.
+However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments  in  \cite{majumder:2013} or \cite{tengfei:2013}, these questions have to be phrased much more specifically, in order to avoid errors of type III \citep{mosteller:48} consisting of correctly rejecting the null hypothesis for a wrong reason.
 %In some of the studies, the ones described in  \cite{majumder:2013} very specific questions were asked, because the experiments were being conducted to compare results from the lineup protocol with those of conventional tests.

### 2015-08-28 sociology_paper.Rnw
intro

@@
 This methodology is  part of what is called visual statistical inference, which has been developed further  and validated by \citet{majumder:2013} in simulation studies of head to head comparisons with conventional inference.
-One of the major contributions of  \citet{majumder:2013}  is to define the power of the visual test and provide methods to obtain the power for a particular lineup.
-It was observed that the power can be as good or better than that of a conventional test in some scenarios.
-
-In visual inference, the test statistic is a plot of the observed data.
-To create a lineup, this plot, called the actual data plot, is placed in a layout of null plots.
-The null plots  are generated from the model specified by a null hypothesis.
-Commonly, this null hypothesis assumes independence/no structure in the data.
+One of their major contributions   is to define the power of  visual tests and to provide a method for obtaining the power for a particular lineup.
+In some scenarios, the power of a visual test  was seen to outperform that of a conventional test.
+
+In visual inference, the test statistic is a plot of the observed data, called the {\it (actual) data plot}.
+To create a lineup the data plot is placed randomly in a grid of null plots.
+Null plots  are rendered from data generated as  specified by a null hypothesis.
+Often, this null hypothesis assumes independence or no structure in the data.
 An observer is then asked to evaluate the lineup.
 If the  actual data plot is detected by the observer, this counts as evidence against the null hypothesis.
-Based on the number of evaluations by independent observers and the number of times that the data plot is identified, we are then able to evaluate the significance of the graphical relationship in the same way and the same rigor as a conventional hypothesis test.
+Based on the number of evaluations by independent observers and the number of times that the data plot is identified, we are  able to evaluate the significance of the graphical relationship in the same way and the same rigor as a conventional hypothesis test.
 % is rejected.

### 2015-08-28 sociology_paper.Rnw
read through Data Collection

@@
 
-Figure~\ref{fig:elect} displays a lineup of 20 plots where one of the plots is observed data, while the remaining 19 plots are showing  data generated under a null model.
+Figure~\ref{fig:elect} displays an example lineup, one of the plots is based on the observed data, while the remaining 19 plots show  data generated from a null model.
 Which one of the 20 plots is the most different from the others?
-When asked this question, 12 out of 72 observers picked the  plot of the observed data, located in panel number $4^2-3$ in the lineup \footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
+When asked this question, 12 out of 72 observers picked the  data plot located in panel number $4^2-3$ in the lineup\footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
 The corresponding $p$-value is 0.0077, indicating sufficient evidence to reject the null hypothesis.
@@
 This particular example investigates the results from the 2012 US presidential election in comparison to  poll results just prior to the election. 
-Although this example is more simplistic than most of the tests conducted to date, it will serve the purpose of illustrating the lineup protocol.
+Although this example is more simplistic than most of the tests conducted to date, it serves the purpose of illustrating the lineup protocol.
 The data consists of the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
@@
 Color indicates party affiliation.
-The null hypothesis is that ``the election results were consistent with the polls''.
+The null hypothesis is that ``the election results are consistent with the polls''.
 Therefore the polling results provide the null model from which data is simulated.
@@
 These samples are plotted as electoral buildings, and the plot based on the actual election results is placed randomly among the null plots in a lineup of size $5\times 4$.
-If the null hypothesis is true, the  plot of the actual data should look  like any of the other plots, and not be identifiable by an observer.
+If the null hypothesis is true, the  data plot should look  like any of the other plots, and not be identifiable by an observer.
 Figure~\ref{fig:tower} shows the plot of the electoral building with the additional context information and labels.
@@
 Presidential Election (left).
-On the right two histograms showing simulated poll results for the Democratic presidential candidate using  polling averages collected at two different sources, Freedom's Lighthouse Averages (top), RealClearPolitics (bottom).
+On the right two histograms showing simulated poll results for the Democratic presidential candidate using  polling averages collected from two different sources, Freedom's Lighthouse Averages (top), RealClearPolitics (bottom).
 If the polls matched the electoral result the simulated results would be centered at the electoral result.
 This happens for the RealClearPolitics polls (bottom) but not for Freedom's Lighthouse.
-For the latter only 5.33\% of simulated results were as high as the actual election result, which converts to a $p$-value of 0.0533, almost a statistically significant difference.
+For the latter, only 5.33\% of simulated results were as high as the actual election result, which converts to a $p$-value of 0.0533, indicating an almost statistically significant difference.
 It hints at the polls from this source under-estimating the vote for Obama.
@@
 To avoid expectation errors \citep{meilgaard}, particularly in emotionally charged areas such as election results, observers should not be aware of the data that constitutes a lineup, and should therefore not have seen the actual data plot before inspecting the lineup.
-This is the reason that we explained the context of the election problem only after you had a chance to evaluate the lineup.%the election example, above, the scenario was explained after the lineup question, in the text.
+This is the reason that we revealed the context of the election problem only after the interested reader had a chance to evaluate the lineup.%the election example, above, the scenario was explained after the lineup question, in the text.
 
@@
 This ensures all possible deviations from the null hypothesis to be detected.
-However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments  in  \cite{majumder:2013} or \cite{tengfei:2013}, these questions have to be phrased much more specifically, in order to avoid errors of type III \citep{mosteller:48} consisting of correctly rejecting the null hypothesis for a wrong reason.
+However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments  in  \cite{majumder:2013} or \cite{tengfei:2013}, these questions have to be phrased much more specifically, in order to avoid  type III errors \citep{mosteller:48} consisting of correctly rejecting the null hypothesis for a wrong reason.
 %In some of the studies, the ones described in  \cite{majumder:2013} very specific questions were asked, because the experiments were being conducted to compare results from the lineup protocol with those of conventional tests.
@@
 Section~\ref{sec:factor_performance} discusses human factors that we expect to be influential on the performance of the observer.
-Section~\ref{sec:exp_design} describes the methods used to assess the effects, and Section~\ref{sec:result_socio} describes the results on what we have learned about human factors affecting lineup evaluation.
+Section~\ref{sec:exp_design} describes the methods used to assess the effects, and Section~\ref{sec:result_socio} summarizes the results of what we have learned about human factors affecting lineup evaluation.
 
@@
 A visual test statistic is chosen for a particular task.
-If it is well-designed then departures from the null hypothesis will be visible in the plot, especially if the signal is strong.
+If it is well-designed  departures from the null hypothesis will be visible in the plot, especially if the signal is strong.
 An important factor that helps an observer to identify the actual data plot in a lineup is the strength of the signal in the data.
@@
  $\mu$ is an overall population average, %probability for picking out the data plot from a lineup HH: mu is not a probability, it is on the scale of the link function
- $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively for observer $i$.
+ $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively, for observer $i$.
 The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume independence and normality of the errors, i.e.\ $\ell_j \sim N(0, \sigma_\ell^2)$.
@@
 It is expected that some observers will be more skilled at reading data plots, and hence, more readily detect the plot that is different.
-It is also expected that as observer's gain experience in evaluating lineups that they become more proficient in reading data plots, particularly if feedback is given on whether the actual data plot was chosen or not.
-Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups.%, so ``correct'' is more of a convenient rather than ideal word to use for this analysis.
+It is also expected that as observers gain experience in evaluating lineups that they become more proficient in reading data plots, particularly if feedback is given on whether the actual data plot was chosen or not.
+Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups.
+%, so ``correct'' is more of a convenient rather than ideal word to use for this analysis.
 So in this context, detected, or not, is used as a response to examine individual differences.
@@
 where $\mu$ is an overall population average, %probability for picking out the data plot from a lineup, 
-$\alpha_k$ is the effect of the $k$th attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
+$\alpha_k$ is the effect of the $k$th attempt on the probability, using the first attempt as reference, $\alpha_1 = 0$, and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
 The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
@@
 When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
-Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k$th attempt.
-We model this in a mixed effects model of the form
+Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of the (log) of time taken by  observer $i$ to choosing a  panel from lineup $j$ in his/her $k$th attempt.
+We evaluate this in a mixed effects model of the form
 %\hh{
@@
 To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2015, lme4:paper}.
-To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used to compute $Z$ scores  as the ratio of estimates to  estimated standard errors.
+We employ a normal approximation to obtain $p$-values corresponding to fixed effect parameters estimates.
+% normal approximation is used to compute $Z$ scores  as the ratio of estimates to  estimated standard errors.
 

### 2015-08-28 sociology_paper.Rnw
first stab at proof reading the results

@@
 Let $\mathbf{Y}=(Y_1,Y_2, ...
-, Y_p)^\top$ be a vector of random variables with dimension $p (=5)$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represents $j$th vector response for $i$th location with $i=1,2, ..., I(=5)$.
+, Y_p)^\top$ be a vector of random variables with dimension $p = 5$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represent the $j$th vector response for  location $i$ with $i=1,2, ..., I$ for $U=5$.
 Because the same data plot is shown in each lineup, it is assumed that there could be some association between the responses for each null set, which suggests the MANOVA model rather than a univariate ANOVA.
@@
 \subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through MTurk \citep{turk}.
-It is an online work place where people from around the world can perform tasks, that computers would have difficulty doing, and get paid.
+It is an online work place where people from around the world can sign up for so-called `HIT's, human intelligence tasks, generally short tasks that are humans are typically better at solving than computers.
 Usually tasks are very simple and no specialized training is required to do them.
 Tasks are designed for anyone to do but some tasks may require some skills depending on the recruiters' need.
-Each task is usually planned to completed in a short time.
-The amount of money paid for each task is very small, on the order of minimum wage in the USA.
+%Each task is usually planned to completed in a short time.
+ For completing a HIT workers are paid a small amount of money, on the order of minimum wage in the USA.
 
@@
 The another benefit is that a very diverse pool of subjects can be recruited which is otherwise very hard to obtain for a study.
-The researchers can easily filter the workers based on their experimental design, such as recruiting people only from a specific geographical location or a group of people who satisfy certain criteria etc.
+%The researchers can easily filter the workers based on their experimental design, such as recruiting people only from a specific geographical location or a group of people who satisfy certain criteria etc.
 The recruiter can decide who they pay or not.
@@
 
-We designed and developed a web application which enables the display of lineups to the observers as per experimental need.
-The MTurk workers were redirected to this web application to complete their assigned tasks.
-The responses were collected, stored automatically into a local database server, along with demographic information, age group, gender and education level.
+We designed and developed a web application which enables the display of lineups to  observers as per experimental need.
+The MTurk workers were re-directed to this web application to complete their assigned tasks.
+Responses were collected, stored automatically into a local database server, along with demographic information, age group, gender and education level.
 The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
-It is measured in seconds.
-The location of the observer is determined by the ip address of the observer.
+%It is measured in seconds.
+Location of an observer is determined based on the ip address of the observer.
 
@@
 
-Besides United States and India, countries such as Canada, Romania, United Kingdom and Macedonia have more than 10 participants each.
-The rest of the 70 countries have fewer than 10 participants.
+Besides the United States and India, countries such as Canada, Romania, the United Kingdom and Macedonia have more than 10 participants each.
+The remaining 70 countries have fewer than 10 participants each.
 The distribution of participants is similar in all ten experiments.
@@
 \begin{table}[hbtp]
-\caption{Demographic information of the subjects participated the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
+\caption{Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
 \centering
@@
 
-The largest number of participants falls within the age group of 18 to 25, with the majority between 18 to 35.
-Many older participants also took part in the experiments.%Interestingly, there are many participants in the older age groups as well.
-The United States shows a more uniform participation beyond age 30, Indian participants are primarily younger.
+The largest number of participants falls within the age group of 18 to 25, with the majority being between 18 to 35.
+Many older participants also took part in the experiments.
+%Interestingly, there are many participants in the older age groups as well.
+The United States shows a more uniform participation beyond age 30,  participants from India are primarily younger.
 
 In terms of academic background, the majority of participants had a graduate degree (a total of 902 participants, about 38.51\%).
-Examining the participants' location in conjunction with the academic background reveals that there are more Indian participants that report to have a graduate degree.
-In retrospect we realized this to be a culturally based misunderstanding, as any university degree in India is considered to be a graduate degree, while the North American the definition of graduate degree refers to a Masters level, or above, education.
-Most of the USA participants report to have an undergraduate degree or at least have some undergraduate courses.
+Examining the participants' location in conjunction with the academic background reveals that there is an above average  number of participants from India reporting to have a graduate degree.
+In retrospect this turned out to be a culturally based misunderstanding, as any university degree in India is considered to be a graduate degree, while the North American definition of a graduate degree refers to a Masters level education or above.
+Most of the  participants from the U.S.
+report to have an undergraduate degree or at least have some undergraduate courses.
 
@@
 
-A total of 1911 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where three lineups were evaluated by each person. 
-In each of the experiments, to ensure quality evaluations, a test plot was shown to each observer, and the performance on this test plot was used for screening purposes.%A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
+A total of 1911 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where each participant evaulated three lineups.
+As a measure to ensure a high quality in the evaluations,
+ a test plot was shown to each observer, and the performance on this test plot was used for screening purposes.
+%A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
 Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.
-Data from participants who gave the wrong answer on the test plot were considered as gaming the system, and their data were removed.
-Results from test plots were not considered in any of the analyses.
- In some cases the demographic information were not provided by the participants.
+Data from participants who gave the wrong answer on the test plot % were considered as gaming the system, and their data were removed.
+led to the removal of all of their answers from the analysis. 
+Results from test plots were also not considered in any of the analyses.
+ In some cases  participants chose to not provide their demographic information.
 Also, for some ip address, the actual geographical locations could not be retrieved.
@@
   \hline
-Model & & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
-  \hline
-Proportion Correct &Full & 23822.00 &23943.00 &-11896.00  & &  &  \\ 
-& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
-  &Age & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
-  &Country  & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
-  &Education  & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
-  &Gender  & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
-   \hline
-   Log Time &Full &  51904.00 &52034.00 & -25936.00 & &  &  \\ 
-& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
-&Age  & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
- & Country  & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
-  &Education  & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
-  &Gender  & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
+\multicolumn{2}{c}{Model}  & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
+  \hline
+Proportion Correct &Full & 23822.00 &23943.00 &-11896.00  & &  &  \\ [3pt]
+%& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
+{Reduced Models}  &- Age & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
+  &- Country  & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
+  &- Education  & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
+  &- Gender  & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
+   \hline
+   Log Time Taken &Full &  51904.00 &52034.00 & -25936.00 & &  &  \\ [3pt]
+%& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
+Reduced Models &- Age  & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
+ &- Country  & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
+  &- Education  & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
+  &- Gender  & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
 \hline
@@
 Participants from India were slower, while those from the rest of the world were faster than US Americans.
-A higher degree corresponded to longer time, and males were slower than females.
+A higher degree corresponded to longer time, and males were on average slower than females.
 
@@
 For gender no difference in performance is observed.
-Lineup variability is estimated to be very large for model \eqref{eqn:demographic_response}.
+Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.
 %
@@
 The average time taken to respond (natural log) and detection rate are computed by lineup for the different demographic factor levels, and displayed using boxplots.
-Since the models use the mean values, rather than median, these are represented by a dot inside the boxplots.
+Mean values are shown as dots inside the boxplot, as this statistic rather than the median is used in the models.
 %There are some small differences across demographic groups.
@@
 $$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.041.$$ 
-The person with a high school education on average picks the data plot in 33.5\% of lineups having average difficulty, as compare to 37.6\% if they have a graduate degree.
+The person with a high school education on average picks the data plot in 33.5\% of lineups having average difficulty, as compared to 37.6\% if they have a graduate degree.
 This difference is reduced to 2\% for a lineup with one standard deviation order of magnitude difference in difficulty.
 For two standard deviations it further reduces to 0.3\%.
-Thus although there is statistically significant difference in proportion correct for some demographic factors, these are not practically significant differences.
+Thus, although there is a statistically significant difference in the proportion at which participants identify the  data plot for some demographic factors, these are not  significant in any practical sense.
 Figure~\ref{fig:practical_impact_graduate} illustrates this example showing fitted models for a US 18-25 female with either a high school education or a graduate degree.
-Similar calculations show the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of correct response.
+Similar calculations show the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of a correct response.
 Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.
@@
     \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
-   \caption{Proportion of correct responses due to graduate degree as compared to high school degree for an 18-25 year old female in the United States.
-Even though graduate degree is statistically significant, the largest difference in proportion correct is  0.045 which is very negligible.
+   \caption{Proportion of correct responses of an 18-25 year old female in the United States with a graduate degree compared to a high school degree.
+Even though having a graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.045 which is  negligible from a practical point of view.
 The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$)
@@
 
-\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments 5, 6, and 7 separately.
+\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments~5,~6, and~7 separately.
 In this model, attempt is fitted as a factor variable to allow for all possible non-linear learning trends.

### 2015-08-28 sociology_paper.Rnw
up to Location Effect

@@
 The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts.
-Most observers, before evaluating the first lineup, carefully read through instructions and familiarize themselves  with the types of plots used in the experiment, which would inflate the time taken on the first lineup.
+One explanation might be that participants take the chance to  carefully read through instructions and familiarize themselves  with the types of plots used in the experiment before evaluating their first lineup, which would inflate the time taken on the first lineup.
 Each of the experiments asks observers to give a reason for their choice (out of a list of pre-defined answers.
@@
 Later pages of the web site were identical  except for the lineup.
-Thus in the later attempts an observer does not need to spend any time for reading instructions.
+%Thus in the later attempts an observer does not need to spend any time for reading instructions.
 The model reflects that fact.
@@
 \caption{Parameter estimates of model \eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
-Both fixed effect parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.
+Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.
 }
@@
 Averages of these residuals are plotted as dots and least square regression lines are fitted to obtain overall trends.
-For all the three experiments the overall downward slopes are statistically significant which indicates that MTurk workers take less time as they progress through their attempts.}
+For all the three experiments the overall downward slopes are statistically significant which indicates that MTurk workers become more efficient as they progress through their attempts.}
    \label{fig:learning_trend_time}

### 2015-08-28 sociology_paper.Rnw
first whole read-through

@@
 
-Figure~\ref{fig:location_effect} shows an overview of detection rates for both of the effects studies in experiment~9.
+Figure~\ref{fig:location_effect} shows an overview of detection rates for both of the effects studied in experiment~9.
 Each dot corresponds to one lineup.
@@
 We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets.
-The lineup protocol is using a finite and small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set.
+The lineup protocol is using a finite, small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set.
 If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate.
@@
 Each colored line represents a null set and the size of the dots represents number of responses.
-The overall average proportions are shown by dashed line.
+The overall average proportions are shown by dashed lines.
 The actual data plot locations are shaded grey on the top panels to demonstrate their relative positions on a lineup.}
@@
 
-Model \eqref{manova} was fit to the data to test if  detection rates were significantly different for the different locations.
+Model~\eqref{manova} was fit to the data to test, if  detection rates were significantly different for the different locations.
 For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R:2012}.
 The results are shown in Table~\ref{tbl:manova}.
-The $p$-values for both Interaction and Genotype effect are much bigger than the conventional threshold of 0.05 suggesting that location of the actual data plot in the lineup does not significantly affect performance.
+The $p$-values for both the Interaction and the Genotype effect are much bigger than the conventional threshold of 0.05 suggesting that location of the actual data plot in the lineup does not significantly affect performance.
 
@@
 However, the effects are minimal, on the order of a few percentage points different from the average.
-This result is very important for the power of visual test as it demonstrates the robustness of the test for different human factors.
+These results are very important for the power of visual tests as they demonstrate the robustness of the test against different human factors.
 
@@
 Individual learning trend is observed in time taken but not so much in observer performance.
-Some individuals improved the performances while others showed decrease in their performances over attempts.
+Some individuals improved on their performances while others showed a decrease on successive attempts.
 This could be interpreted as good, in the sense that it means that observers could realistically be recruited from sources like MTurk, and that substantial training is not necessary in order to obtain useful evaluations of lineups in practice.
-However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea.
+However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea, at least for short term learning.
 
@@
 
-The simulation experiment reveals that there is no significant effect of location of the actual data plot in the lineup.
+The simulation experiment reveals that there is no significant effect of location on the actual data plot in the lineup.
 This is important as the visual statistical inference procedure prescribes that the data plot be placed at random in the lineup.
@@
 This paper assessed the effect of human factors on the experiments conducted to date.
-As changes to the lineup protocol are suggested by other experiments the human factor effects me need to be examined again.
+As changes to the lineup protocol are suggested by other experiments the human factor effects may need to be examined again.
 

### 2015-08-28 sociology_paper.Rnw
references
references added

@@
 \end{figure}
+The lineup of Figure~\ref{fig:elect}, as well as the other lineups in the manuscript are created using the nullabor package by \citep{nullabor}.
 
@@
 A lineup can be evaluated by a single person or multiple observers.
-A distribution similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario, is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis.
+A distribution similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario, is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis \citep{vinference}.
 To avoid expectation errors \citep{meilgaard}, particularly in emotionally charged areas such as election results, observers should not be aware of the data that constitutes a lineup, and should therefore not have seen the actual data plot before inspecting the lineup.
@@
 
-In experiments~4,~5 and~6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices.
-Results are published in \cite{heike:2012} and \cite{boxplots:arxiv}.
+In experiments~4,~5 and~6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices, the use of lineup tests for comparisons and two examples are discussed in \citet{heike:2012}
 
@@
 Model~\eqref{manova} was fit to the data to test, if  detection rates were significantly different for the different locations.
-For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R:2012}.
+For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R}.
 The results are shown in Table~\ref{tbl:manova}.

### 2015-08-28 R/sociology_paper.R
figure 4

@@ -221,9 +221,12 @@ cdat <- get_effect(demographics, c("country", "pic_name"))
 adat <- get_effect(demographics, c("age_level", "pic_name"))
 
 mdat <- melt(rbind(gdat,edat,cdat,adat), id=c("variable_level", "pic_name", "variable_name"))
+levels(mdat$variable) <- c("(Log) Time Taken in seconds", "Detection Rate")
+mdat$variable_name <- factor(mdat$variable_name)
+levels(mdat$variable_name) <- c("Age Categories", "Country", "Education", "Gender")
 
 qplot(variable_level, value, geom="boxplot",data=mdat[complete.cases(mdat),]) +
-  facet_grid(variable~variable_name, scales="free") +
+  facet_grid(variable~variable_name, scales="free", space="free_x") +
   stat_summary(fun.y=mean, geom="point") + xlab("Levels of demographic factors")+ylab("")+
   theme(axis.text.x=element_text(angle=90, hjust=1))
 

### 2015-08-28 R/sociology_paper.R
adjusted get_estimates for new version of lme4

@@ -244,8 +244,9 @@ get_estimates <- function(fit){
   pval <- round(2 * pnorm(abs(zval), lower.tail = FALSE),4)
   pval[pval==0] <- "$<$0.001"
   fe <- data.frame(Est=round(betas,3),SE= round(se,3), Zval=round(zval,2), pvalue=pval)
-  rem <- summary(fit)@REmat
-  re <- data.frame(Est=rem[,3], SE= rem[,4], Zval="",pvalue="")
+#  browser()
+  rem <- as.data.frame(VarCorr(fit))
+  re <- data.frame(Est=rem$vcov, SE= rem$sdcor, Zval="",pvalue="")
   re$Est <- round(as.numeric(as.character(re$Est)),3)
   re$SE <- round(as.numeric(as.character(re$SE)),3)
   rownames(re) <- paste(rem[,1],rem[,2])
@@ -261,8 +262,8 @@ ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name),
            data=demographics)
 summary(ft)
 
-fp <- lmer(response~age_level+country+degree+gender_level+(1|pic_name), 
-           family="binomial", demographics)
+fp <- glmer(response~age_level+country+degree+gender_level+(1|pic_name), 
+           family="binomial", demographics, control=glmerControl(optimizer="bobyqa"))
 
 summary(fp)
 

### 2015-08-28 R/sociology_paper.R
figure 5

@@ -318,8 +318,8 @@ xtable(anova.reseult)
 
 # Model with demographic factor interaction
 # India:Undergrad course is hghly significant
-fit <- lmer(response~age_level+country+degree+gender_level+country:degree+(1|pic_name), 
-               family="binomial", demographics)
+fit <- glmer(response~age_level+country+degree+gender_level+country:degree+(1|pic_name), 
+               family="binomial", demographics, control=glmerControl(optimizer="bobyqa"))
 
 
 fit
@@ -386,7 +386,7 @@ with(ddat, max(prop_gd-prop_hs))
 
 mddat <- melt(ddat,id=c("difficulty"))
 qplot(difficulty, value, data=mddat, linetype=variable, geom="line")+
-  xlab("Lineup difficulty") + ylab("Proportion of correct response") +
+  xlab("Lineup difficulty") + ylab("Proportion of data identification") +
   scale_x_continuous(breaks=c(-6,-2*2.293,-2.293,0,2.293,2*2.293,6), expand = c(0.02,0),
                      labels=c("Difficult",expression(-2*sigma[l],-sigma[l],0,sigma[l], 
                                                      2*sigma[l],"Easy"))) +
@@ -414,7 +414,7 @@ qplot(variable,value, shape=factor(difficulty), geom=c("point"),data=mddat.all, 
 
 qplot(variable,value, shape=factor(difficulty),data=mddat.all, size=I(3.5)) +
   geom_line(aes(variable,value, group=factor(difficulty)))+
-  ylab("Proportion of correct responses") + xlab("Education")+
+  ylab("Proportion of data identification") + xlab("Education")+
   scale_x_discrete(labels=c("h.school","u.gd.course","u.grad.deg","grad.course","grad.deg"))+
   scale_shape_discrete(name="Lineup difficulty", 
                        labels=c("Difficult", "Medium","Easy")) 

### 2015-08-28 sociology_paper.Rnw
map

@@
     \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
-   \caption{Proportion of correct responses of an 18-25 year old female in the United States with a graduate degree compared to a high school degree.
+   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with a graduate degree compared to a high school degree.
 Even though having a graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.045 which is  negligible from a practical point of view.

### 2015-08-28 R/sociology_paper.R

@@ -481,6 +481,18 @@ ggsave("../images/rejected_task.pdf", width=6, height=4)
 
 # Drawing the map of turk participants
 
+map_theme <- list(theme(panel.grid.minor = element_blank(),
+                        panel.grid.major = element_blank(),
+                        panel.background = element_blank(),
+                        panel.border = element_blank(),
+                        axis.line = element_blank(),
+                        axis.text.x = element_blank(),
+                        axis.text.y = element_blank(),
+                        axis.ticks = element_blank(),
+                        axis.title.x = element_blank(),
+                        axis.title.y = element_blank(),
+                        plot.margin=unit(c(0,0,0,0), unit="cm")))
+
 library(maps)
 map.dat <- as.data.frame(map("world",ylim=c(-45,70), plot=FALSE)[c("x","y")])
 map.dat <- map_data("world")
@@ -488,8 +500,8 @@ map.dat <- map_data("world")
 ggplot() +
   geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) +
   geom_point(aes(longitude,latitude, colour=factor("A")), data=ip.details, alpha=.6) +
-  theme_bw()+
-  theme(legend.position="none", axis.text = element_blank(), axis.title=element_blank()) 
+  theme_bw()+ map_theme
+
 ggsave("../images/turker_location.pdf", width=8, height=4)
 
 

### 2015-08-29 R/sociology_paper.R
figure 7

@@ -262,8 +262,8 @@ ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name),
            data=demographics)
 summary(ft)
 
-fp <- glmer(response~age_level+country+degree+gender_level+(1|pic_name), 
-           family="binomial", demographics, control=glmerControl(optimizer="bobyqa"))
+fp <- glmer(response~age_level+factor(country, ordered=FALSE)+factor(degree, ordered=FALSE)+gender_level+(1|pic_name), 
+           family="binomial", data=demographics, control=glmerControl(optimizer="bobyqa"))
 
 summary(fp)
 
@@ -484,14 +484,13 @@ ggsave("../images/rejected_task.pdf", width=6, height=4)
 map_theme <- list(theme(panel.grid.minor = element_blank(),
                         panel.grid.major = element_blank(),
                         panel.background = element_blank(),
-                        panel.border = element_blank(),
                         axis.line = element_blank(),
                         axis.text.x = element_blank(),
                         axis.text.y = element_blank(),
                         axis.ticks = element_blank(),
                         axis.title.x = element_blank(),
                         axis.title.y = element_blank(),
-                        plot.margin=unit(c(0,0,0,0), unit="cm")))
+                        plot.margin=unit(c(0,0,-1,-1), unit="line")))
 
 library(maps)
 map.dat <- as.data.frame(map("world",ylim=c(-45,70), plot=FALSE)[c("x","y")])
@@ -500,7 +499,8 @@ map.dat <- map_data("world")
 ggplot() +
   geom_polygon(aes(long,lat, group=group), fill="grey65", data=map.dat) +
   geom_point(aes(longitude,latitude, colour=factor("A")), data=ip.details, alpha=.6) +
-  theme_bw()+ map_theme
+  theme_bw()+ map_theme +
+  theme(legend.position="none")
 
 ggsave("../images/turker_location.pdf", width=8, height=4)
 
@@ -516,7 +516,7 @@ ggsave("../images/turker_location_experiment.pdf", width=8, height=10)
 # country-wise participants counts
 qplot(factor(country, levels=names(table(country))[order(table(country))]), 
       geom="bar", fill="A", data=subset(turker,complete.cases(turker)))+ 
[…]

### 2015-08-29 R/sociology_paper.R
figure 6

@@ -627,7 +627,7 @@ qplot(attempt,mean_resid, data= dmt) + geom_point(size=2.5) +
 ggplot() + 
   geom_point(aes(attempt,mean_resid), data= dmt) +
   geom_smooth(aes(attempt,mean_resid), data= subset(dmt, attempt>1), method="lm", se=F) +
-  facet_wrap(~experiment, scales="free_y") + ylab("Residual log(time taken)") +
+  facet_grid(.~Experiment,  labeller="label_both") + ylab("Residual log(time taken)") +
   scale_x_continuous(breaks = seq(2,10,by=2))
 
 ggsave("../images/learning_trend_time.pdf", width=10.5, height = 3.5)
@@ -662,7 +662,7 @@ for (i in 5:7){
   d <- subset(dtrend, experiment==i)
   dd <- subset(d, id %in% id[attempt > 9])
   model <- as.formula(response ~ (1|pic_name) + (1|id))
-  fit <- lmer(model,family="binomial",data=dd)
+  fit <- glmer(model,family="binomial",data=dd, control=glmerControl(optimizer="bobyqa"))
   dd$resid <- (dd$response - fitted(fit))
   trend.dat <- rbind(trend.dat ,dd)
 }
@@ -674,9 +674,10 @@ ggplot(trend.dat, aes(attempt,resid, colour=factor(experiment)))+
 ddt <- ddply(trend.dat,.(experiment, attempt), summarise,
              mean_resid = mean(resid))
 
+ddt$Experiment <- ddt$experiment
 qplot(attempt,mean_resid, data= ddt) + geom_point(size=2.5) +
   geom_smooth(method="lm", se=F) + ylab("Mean residual proportion correct") +
-  facet_wrap(~experiment, scales="free_y") +
+  facet_grid(.~Experiment, scales="free_y", labeller="label_both") +
   scale_x_continuous(breaks = seq(2,10,by=2)) 
 
 ggsave("../images/learning_trend.pdf", width=10.5, height = 3.5)
@@ -686,8 +687,8 @@ ggplot() +
   geom_smooth(aes(attempt,resid, group=id),method="lm", se=F,  data=trend.dat, colour=rgb(0,0,0, alpha=0.05))+
   geom_point(aes(attempt,mean_resid, group=1), data= ddt) +
   geom_smooth(aes(attempt,mean_resid, group=1), data= ddt, method="lm", se=F, size=I(1.2)) +
[…]

### 2015-08-29 R/sociology_paper.R
figure 6

@@ -682,7 +682,7 @@ qplot(attempt,mean_resid, data= ddt) + geom_point(size=2.5) +
 
 ggsave("../images/learning_trend.pdf", width=10.5, height = 3.5)
 
-
+trend.dat$Experiment <- trend.dat$experiment
 ggplot() + 
   geom_smooth(aes(attempt,resid, group=id),method="lm", se=F,  data=trend.dat, colour=rgb(0,0,0, alpha=0.05))+
   geom_point(aes(attempt,mean_resid, group=1), data= ddt) +

### 2015-08-29 sociology_paper.Rnw
reference added

@@
 
-In experiments~4,~5 and~6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices, the use of lineup tests for comparisons and two examples are discussed in \citet{heike:2012}
+In experiments~4,~5 and~6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices, the use of lineup tests for comparisons and  examples are discussed in \citet{heike:2012, qqplots:2015}
 

### 2015-08-29 R/sociology_paper.R

@@ -688,7 +688,7 @@ ggplot() +
   geom_point(aes(attempt,mean_resid, group=1), data= ddt) +
   geom_smooth(aes(attempt,mean_resid, group=1), data= ddt, method="lm", se=F, size=I(1.2)) +
   facet_grid(.~Experiment, scales="free_y", labeller="label_both") + ylab("Mean residual proportion correct") +
-  scale_x_continuous(breaks = 1:10)
+  scale_x_continuous(breaks = 1:10) + xlab("Attempt")
 
 ggsave("../images/learning_trend_subject.pdf", width=10.5, height = 3.5)
 

### 2015-08-29 R/sociology_paper.R
updated ip_details.csv

@@ -294,13 +294,13 @@ anova.time <- rbind(data.frame(round(tag[1,1:4],2),round(tag[2,5:7]/1,2)),
                     data.frame(round(tde[1,1:4],2),round(tde[2,5:7]/1,2)),
                     data.frame(round(tge[1,1:4],2),round(tge[2,5:7]/1,2)))
 
-fpa <- lmer(response~country+degree+gender_level+(1|pic_name), 
+fpa <- glmer(response~country+degree+gender_level+(1|pic_name), 
            family="binomial", demographics)
-fpc <- lmer(response~age_level+degree+gender_level+(1|pic_name), 
+fpc <- glmer(response~age_level+degree+gender_level+(1|pic_name), 
            family="binomial", demographics)
-fpd <- lmer(response~age_level+country+gender_level+(1|pic_name), 
+fpd <- glmer(response~age_level+country+gender_level+(1|pic_name), 
            family="binomial", demographics)
-fpg <- lmer(response~age_level+country+degree+(1|pic_name), 
+fpg <- glmer(response~age_level+country+degree+(1|pic_name), 
            family="binomial", demographics)
 
 pag <- anova(fp,fpa)

### 2015-08-29 R/sociology_paper.R
figure 3

@@ -262,10 +262,13 @@ ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name),
            data=demographics)
 summary(ft)
 
-fp <- glmer(response~age_level+factor(country, ordered=FALSE)+factor(degree, ordered=FALSE)+gender_level+(1|pic_name), 
+fp <- glmer(response~age_level+country+degree+gender_level+(1|pic_name), 
            family="binomial", data=demographics, control=glmerControl(optimizer="bobyqa"))
 
 summary(fp)
+resid <- residuals(fp)
+# residual error
+sqrt(sum(resid^2)/(length(resid)-length(fixef(fp))))
 
 est.factor <- cbind(get_estimates(ft)[-16,],g1=" ",get_estimates(fp))
 rownames(est.factor) <- c("$mu$",substr(rownames(est.factor)[2:7],10,nchar(rownames(est.factor)[2:7])),
@@ -295,13 +298,13 @@ anova.time <- rbind(data.frame(round(tag[1,1:4],2),round(tag[2,5:7]/1,2)),
                     data.frame(round(tge[1,1:4],2),round(tge[2,5:7]/1,2)))
 
 fpa <- glmer(response~country+degree+gender_level+(1|pic_name), 
-           family="binomial", demographics)
+           family="binomial", demographics, control=glmerControl(optimizer="bobyqa"))
 fpc <- glmer(response~age_level+degree+gender_level+(1|pic_name), 
-           family="binomial", demographics)
+           family="binomial", demographics, control=glmerControl(optimizer="bobyqa"))
 fpd <- glmer(response~age_level+country+gender_level+(1|pic_name), 
-           family="binomial", demographics)
+           family="binomial", demographics, control=glmerControl(optimizer="bobyqa"))
 fpg <- glmer(response~age_level+country+degree+(1|pic_name), 
-           family="binomial", demographics)
+           family="binomial", demographics, control=glmerControl(optimizer="bobyqa"))
 
 pag <- anova(fp,fpa)
 pcn <- anova(fp,fpc)
@@ -675,8 +678,9 @@ ddt <- ddply(trend.dat,.(experiment, attempt), summarise,
              mean_resid = mean(resid))
 
 ddt$Experiment <- ddt$experiment
-qplot(attempt,mean_resid, data= ddt) + geom_point(size=2.5) +
[…]

### 2015-08-29 R/sociology_paper.R
figures 6,7

@@ -689,8 +689,8 @@ ggsave("../images/learning_trend.pdf", width=10.5, height = 3.5)
 trend.dat$Experiment <- trend.dat$experiment
 ggplot() + 
   geom_smooth(aes(attempt,resid, group=id),method="lm", se=F,  data=trend.dat, colour=rgb(0,0,0, alpha=0.05))+
-  geom_point(aes(attempt,mean_resid, group=1), data= ddt) +
   geom_smooth(aes(attempt,mean_resid, group=1), data= ddt, method="lm", se=F, size=I(1.2)) +
+  geom_point(aes(attempt,mean_resid, group=1), data= ddt) +
   facet_grid(.~Experiment, scales="free_y", labeller="label_both") + ylab("Mean residual proportion correct") +
   scale_x_continuous(breaks = 1:10) + xlab("Attempt")
 
@@ -707,19 +707,19 @@ fit3 <- lm(resid ~ attempt, data=subset(trend.dat, experiment==7))
 # Getting results of model with proportion correct
 model <- as.formula(response ~ factor(attempt) + (1|pic_name) + (attempt|id))
 dt5 <- subset(dtrend, experiment==5)
-fp5 <- lmer(model,family="binomial",data=dt5)
+fp5 <- glmer(model,family="binomial",data=dt5)
 res5 <- get_estimates(fp5)
 res5$pvalue <- as.character(res5$pvalue)
 res5$pvalue[2:10] <- round(as.numeric(res5$pvalue[2:10]),3)
 
 dt6 <- subset(dtrend, experiment==6)
-fp6 <- lmer(model,family="binomial",data=dt6)
+fp6 <- glmer(model,family="binomial",data=dt6)
 res6 <- get_estimates(fp6)
 res6$pvalue <- as.character(res6$pvalue)
 res6$pvalue[1:10] <- round(as.numeric(res6$pvalue[1:10]),3)
 
 dt7 <- subset(dtrend, experiment==7)
-fp7 <- lmer(model,family="binomial",data=dt7)
+fp7 <- glmer(model,family="binomial",data=dt7)
 res7 <- get_estimates(fp7)
 res7$pvalue <- as.character(res7$pvalue)
 res7$pvalue[1:10] <- round(as.numeric(res7$pvalue[1:10]),3)

### 2015-08-30 sociology_paper.Rnw
updated ip_details.csv
figure 3
re-running models with unordered factors - will change things downstream, but not by much

@@
 To estimate the significance of each of the factors, reduced models are fit with that factor removed from the model.
-Table ~\ref{tbl:anova_factor} summarizes the results.
+Table~\ref{tbl:anova_factor} summarizes the results.
 Except for gender, all of the demographic factors show significant differences in  detection rates.
@@
 
+% latex table generated in R 3.2.2 by xtable 1.7-4 package
+% Sat Aug 29 14:17:22 2015
+\begin{table}[hbtp]
+\caption{Parameter estimates of models~\eqref{eqn:demographic_time} and~\eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively.
+For time taken all the demographic factors are significant.
+For detection rate, age group 36-40, rest of the world and graduate degree are significantly different.
+For gender no difference in performance is observed.
+Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
+\centering
+\scalebox{.85}{
+\begin{tabular}{rrr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
+ && \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
+
+\cline{3-6} \cline{8-11} 
+\multicolumn{2}{c}{{Demographic Factor} \hfill Level} & Est & 2.5\% & 97.5\% &   && Est & 2.5\% & 97.5\% &   \\ 
+\cline{3-6} \cline{8-11}
+\bf Fixed Effects & Average $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.61 & -0.79 & -0.43 & *** \\ [3pt]
+\bf Age Category $(\alpha)$ & 18-25 & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
+&  26-30 & 0.06 & 0.03 & 0.08 & *** &   & 0.06 & -0.03 & 0.16 &  \\ 
+&  31-35 & 0.07 & 0.04 & 0.10 & *** &   & 0.11 & 0.01 & 0.22 & * \\ 
+&  36-40 & 0.23 & 0.20 & 0.26 & *** &   & 0.31 & 0.19 & 0.43 & *** \\ 
+&  41-45 & 0.18 & 0.14 & 0.22 & *** &   & 0.16 & 0.00 & 0.31 & * \\ 
+&  46-50 & 0.27 & 0.22 & 0.32 & *** &   & 0.14 & -0.05 & 0.33 &  \\ 
+&  above 50 & 0.35 & 0.32 & 0.39 & *** &   & 0.15 & 0.01 & 0.28 & * \\ [3pt]
+\bf Country $(\kappa)$& U.S. & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
+&  India & 0.23 & 0.21 & 0.25 & *** &   & -0.18 & -0.27 & -0.10 & *** \\ 
+&  Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.08 & -0.04 & 0.20 &  \\ [3pt]
+\bf Education $(\tau)$& High School & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
+&  Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & -0.03 & -0.18 & 0.11 &  \\ 
+&  Under grad degree & 0.13 & 0.10 & 0.17 & *** &   & 0.12 & -0.01 & 0.26 & . \\ 
+&  Graduate courses & 0.01 & -0.04 & 0.05 &  &   & -0.17 & -0.33 & -0.01 & * \\ 
+&  Graduate degree & 0.13 & 0.09 & 0.16 & *** &   & -0.06 & -0.19 & 0.07 &  \\ [3pt]
+\bf Gender $(\gamma)$& Female & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
+&  Male & 0.08 & 0.06 & 0.10 & *** &   & 0.06 & -0.02 & 0.13 &  \\
+\cline{3-6} \cline{8-11}
+\bf Random Effects &  lineup $(\sigma_\ell)$& \multicolumn{5}{l}{0.29} & \multicolumn{4}{l}{\phantom{-}2.29} \\ 
+                   &  Error $(\sigma)$& \multicolumn{5}{l}{0.69} & \multicolumn{4}{l}{\phantom{-}0.89}  \\ 
+\cline{3-6} \cline{8-11}
+\end{tabular}}
+\end{table}
+
 \begin{table}[htbp]
@@
 & lineup($\sigma_{\ell}$) & 0.287 &  &  &  &   & 2.293 &  &  &  \\ 
- & Error($\sigma$) & 0.479 & &  &  &  & \mm{? 0.692}&& &\\ 
+ & Error($\sigma$) & 0.692 & &  &  &  & 0.893 && &\\ 
    \hline

### 2015-08-30 sociology_paper.Rnw
question: no sigma in model 1
figure 5

@@
 \end{equation}
-where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model \eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model~\eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.
 
@@
 %where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
-$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model \eqref{eqn:demographic_response}.
+$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model~\eqref{eqn:demographic_response}.
 $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
@@
 
-To explore the significance of the demographic factors model \eqref{eqn:demographic_response} is fit to the data, with age, country, education and gender as fixed effects.
+To explore the significance of the demographic factors model~\eqref{eqn:demographic_response} is fit to the data, with age, country, education and gender as fixed effects.
 To estimate the significance of each of the factors, reduced models are fit with that factor removed from the model.
@@
 \begin{table}[hbtp]
-\caption{Parameter estimates of models~\eqref{eqn:demographic_time} and~\eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively.
+\caption{\label{tbl:model_result_demographics}Parameter estimates of models~\eqref{eqn:demographic_time} and~\eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively.
 For time taken all the demographic factors are significant.
@@
 
-\begin{table}[htbp]
-\centering
-\caption{Parameter estimates of models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively.
+% \begin{table}[htbp]
+% \centering
+% \caption{Parameter estimates of models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively.
 For time taken all the demographic factors are significant.
@@
 Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.
-%
-}
-\scalebox{0.86}{
-\begin{tabular}{rlrrrrrrrrrr}
-  \hline
-Demographic&& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})} \\
-
-\cline{3-6} \cline{8-11} 
-Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
-  \hline
-\multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
-&$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
-\hline
-Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
- &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.206 \\ 
-  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.038 \\ 
-  &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
-  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.050 \\ 
-  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.143 \\ 
-  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.039 \\ 
-  \hline
-Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
-  &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
-  &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
-  \hline
-Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
- &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
-  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.331 \\ 
-  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.157 \\ 
-  &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
-  \hline
-Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
-  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.133 \\ 
-  \hline
-\multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
-& lineup($\sigma_{\ell}$) & 0.287 &  &  &  &   & 2.293 &  &  &  \\ 
- & Error($\sigma$) & 0.692 & &  &  &  & 0.893 && &\\ 
-   \hline
-\end{tabular}
-}
-\label{tbl:model_result_demographics}
-\end{table}
+% %
+% }
+% \scalebox{0.86}{
+% \begin{tabular}{rlrrrrrrrrrr}
+%   \hline
+% Demographic&& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})} \\
+% 
+% \cline{3-6} \cline{8-11} 
+% Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
+%   \hline
+% \multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
+% &$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
+% \hline
+% Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
+%  &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.206 \\ 
+%   &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.038 \\ 
+%   &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
+%   &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.050 \\ 
+%   &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.143 \\ 
+%   &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.039 \\ 
+%   \hline
+% Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
+%   &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
+%   &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
+%   \hline
+% Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
+%  &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
+%   &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.331 \\ 
+%   &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.157 \\ 
+%   &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
+%   \hline
+% Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
+%   &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.133 \\ 
+%   \hline
+% \multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
+% & lineup($\sigma_{\ell}$) & 0.287 &  &  &  &   & 2.293 &  &  &  \\ 
+%  & Error($\sigma$) & 0.692 & &  &  &  & 0.893 && &\\ 
+%    \hline
+% \end{tabular}
+% }
+% \label{tbl:model_result_demographics}
+% \end{table}
 
@@
 In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond.
-In Table~\ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
+In Table~\ref{tbl:model_result_demographics} notice that lineup specific error is estimated as 2.29 which is much higher than any of the other parameter estimates in the model.
 This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
@@
 For gender no difference in performance is observed.
-Lineup variability is estimated to be very large for Model \eqref{eqn:demographic_response}.}
+Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
 %\scalebox{0.86}{
@@
 %  \hline
-%Demographic&& \multicolumn{4}{c} {Model \eqref{eqn:demographic_time} Log Time} & &\multicolumn{4}{c} {Model \eqref{eqn:demographic_response} Proportion Correct} \\
+%Demographic&& \multicolumn{4}{c} {Model~\eqref{eqn:demographic_time} Log Time} & &\multicolumn{4}{c} {Model~\eqref{eqn:demographic_response} Proportion Correct} \\
 %
@@
 
-%In Table~\ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model \eqref{eqn:demographic_response}.
+%In Table~\ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model~\eqref{eqn:demographic_response}.
 This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
@@
 For example, let's examine the effect of graduate degree.
-To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with a graduate degree as compared to a high school degree, for an average difficulty lineup (random effect = 0).
+To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with some graduate course work as compared to an undergraduate degree, for an average difficulty lineup (random effect of zero).
 Plugging in the relevant quantities to the fitted model gives a difference equal to:
-$$ \frac {\exp(-0.683+0.182)}{1+\exp(-0.683+0.182)}-  \frac {\exp(-0.683)}{1+\exp(-0.683)} = 0.377- 0.336 =0.041.$$ 
-The person with a high school education on average picks the data plot in 33.5\% of lineups having average difficulty, as compared to 37.6\% if they have a graduate degree.
+$$ \frac {\exp(-0.61-0.17)}{1+\exp(-0.61-0.17)}-  \frac {\exp(-0.61+0.12)}{1+\exp(-0.61+0.12)} = 0.314- 0.379 =-0.065.$$ 
+The person with some graduate course work picks the data plot  on average in 33.5\% of lineups of average difficulty, as compared to 37.9\% if they have an undergraduate degree.
 This difference is reduced to 2\% for a lineup with one standard deviation order of magnitude difference in difficulty.
@@
 Figure~\ref{fig:practical_impact_graduate} illustrates this example showing fitted models for a US 18-25 female with either a high school education or a graduate degree.
-Similar calculations show the same negligible impact of age level 36-40 (0.0533 at most)  and country (0.0424 at most) on the probability of a correct response.
+Similar calculations show the same negligible impact of age level 36-40 (\Sexpr{round(exp(-0.61+0.31)/(1+ exp(-0.61+0.31)) - exp(-0.61)/(1 + exp(-0.61)),3)} at most)  and country (\Sexpr{round(exp(-0.61+0.08)/(1+ exp(-0.61+0.08)) - exp(-0.61-0.18)/(1 + exp(-0.61-0.18)), 3)} at most) on the probability of a correct response.
 Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.
@@
 Even though having a graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.045 which is  negligible from a practical point of view.
-The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.293$)
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.29$)
    of lineup variability.}
@@
 
-%Model \eqref{eqn:demographic_time} is fitted for the log time taken to evaluate a lineup and the results are also shown in  Table~\ref{tbl:model_result_demographics}.
+%Model~\eqref{eqn:demographic_time} is fitted for the log time taken to evaluate a lineup and the results are also shown in  Table~\ref{tbl:model_result_demographics}.
 The effect of all the demographic factors on log time taken to evaluate a lineup appears to be significant.
@@
 In this model, attempt is fitted as a factor variable to allow for all possible non-linear learning trends.
-It should be noted that we also examined an alternative to model \eqref{eqn:trend_response} where attempt was linearly fitted as a continuous covariate, but this also was not significant.
-
-Table~\ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model \eqref{eqn:trend_response} examining detection rate.
+It should be noted that we also examined an alternative to model~\eqref{eqn:trend_response} where attempt was linearly fitted as a continuous covariate, but this also was not significant.
+
+Table~\ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model~\eqref{eqn:trend_response} examining detection rate.
 Only for experiment~6 there is some evidence of a learning curve.
@@
 
-%Model \eqref{eqn:trend_time} is fitted to the data from experiment~5, 6, and 7 separately.
+%Model~\eqref{eqn:trend_time} is fitted to the data from experiment~5, 6, and 7 separately.
 For this the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
@@
 
-To visualize how  detection rates change over successive attempts, we fitted model \eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
+To visualize how  detection rates change over successive attempts, we fitted model~\eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
 Least square regression lines were fitted through the subject specific residuals as shown in Figure~\ref{fig:learning_trend_response}.
@@
     \includegraphics[width=6.3in]{learning_trend_subject.pdf} 
-   \caption{ Least square lines fitted through the subject specific residual proportion correct obtained from model \eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
+   \caption{ Least square lines fitted through the subject specific residual proportion correct obtained from model~\eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
 Subject specific positive and negative slopes are observed.
@@
 
-Table~\ref{tbl:model_result_time} presents the results of model \eqref{eqn:trend_time} where response time is examined with respect to attempt.
+Table~\ref{tbl:model_result_time} presents the results of model~\eqref{eqn:trend_time} where response time is examined with respect to attempt.
 Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated.
@@
 \centering
-\caption{Parameter estimates of model \eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
+\caption{Parameter estimates of model~\eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
 Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.
@@
 
-%The lineup variance of experiment~7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model \eqref{eqn:trend_response}.
-But from model \eqref{eqn:trend_time} it is estimated to be much smaller in all these experiments.
+%The lineup variance of experiment~7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model~\eqref{eqn:trend_response}.
+But from model~\eqref{eqn:trend_time} it is estimated to be much smaller in all these experiments.
 This suggests that the harder lineups not necessarily took more time than easier lineups.
@@
 
-%Model \eqref{eqn:trend_time} was compared to couple of other alternatives.
+%Model~\eqref{eqn:trend_time} was compared to couple of other alternatives.
 A bigger model with attempt as a factor was considered but it was not significantly different with $p$ value 0.236.
@@
 
-To visualize how the time taken reduces over the successive attempts, we fitted model \eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
+To visualize how the time taken reduces over the successive attempts, we fitted model~\eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
 Least square regression lines are fitted through the subject specific residuals.
@@
 
-%To accomplish this we fit generalized mixed effect model \eqref{eqn:trend_response} with lineups and subjects as random effect and no fixed effect.
+%To accomplish this we fit generalized mixed effect model~\eqref{eqn:trend_response} with lineups and subjects as random effect and no fixed effect.
 This gives the estimation of proportion correct responses  and we obtain residuals of the fitted model.
@@
  \includegraphics[width=6.3in]{learning_trend_time_subject.pdf}    
-   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting model \eqref{eqn:trend_time} without covariate attempt.
+   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting model~\eqref{eqn:trend_time} without covariate attempt.
 Differences in subject specific slopes are observed.
@@
 \begin{table}[hbtp]
-\caption{The summary of the results obtained by fitting MANOVA model \eqref{manova} using Wilks test.}
+\caption{The summary of the results obtained by fitting MANOVA model~\eqref{manova} using Wilks test.}
 \begin{center}
@@
 From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype. 
-Model \eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.  
+Model~\eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.  
 %It does not show any differences whether the actual plot is inside or outer border of the lineup.

### 2015-08-30 R/sociology_paper.R

@@ -410,9 +410,11 @@ ggsave("../images/practical_impact_demographics.pdf", width=5.5, height=5)
 
 
 difficulty <- seq(-6,6, by=.1)
-hs_xb <- -0.683 + difficulty
-gd_xb <- hs_xb + 0.182
-ddat <- data.frame(difficulty=difficulty, prop_hs = px(hs_xb),prop_gd=px(gd_xb))
+hs_xb <- -0.61335621 + difficulty
+gd_xb <- hs_xb + -0.17436466
+ugd_xb <- hs_xb + 0.12496329 
+
+ddat <- data.frame(difficulty=difficulty, prop_hs = px(ugd_xb),prop_gd=px(gd_xb))
 
 # Maximum difference in prop correct is 0.045
 qplot(difficulty, prop_gd-prop_hs, data=ddat, geom="line")
@@ -425,7 +427,7 @@ qplot(difficulty, value, data=mddat, linetype=variable, geom="line")+
                      labels=c("Difficult",expression(-2*sigma[l],-sigma[l],0,sigma[l], 
                                                      2*sigma[l],"Easy"))) +
   scale_linetype_discrete(name="Education",
-                      labels=c("High school", "Grad. degree")) 
+                      labels=c("U.Grad degree", "Grad. courses")) 
 
 ggsave("../images/practical_impact_graduate.pdf", width=6.5, height=4.5)
 

### 2015-08-30 R/sociology_paper.R
fixed get_estimates for larger sizes

@@ -277,11 +277,16 @@ get_estimates <- function(fit){
   pval <- round(2 * pnorm(abs(zval), lower.tail = FALSE),4)
   pval <- stars(pval)
   fe <- data.frame(Est=round(betas,3),LB= round(cis[,1],3), UB= round(cis[,2],3), pvalue=pval)
-  #  browser()
+#    browser()
   rem <- as.data.frame(VarCorr(fit))
   re <- data.frame(Est=rem$sdcor, LB=NA, UB=NA, pvalue=NA)
   re$Est <- round(as.numeric(as.character(re$Est)),3)
-  rownames(re) <- rem[,1]
+  nvars <- ncol(rem) - 2
+  nms <- rem[,1]
+  if (nvars > 2) {
+    for (i in 2:nvars) nms <- paste(nms, rem[,i], sep="|")
+  }
+  rownames(re) <- nms
   return(rbind(fe,re[order(rownames(re)),]))
 }
 

### 2015-08-30 R/sociology_paper.R
models for prop correct for individual experiments

@@ -745,19 +745,19 @@ fit3 <- lm(resid ~ attempt, data=subset(trend.dat, experiment==7))
 # Getting results of model with proportion correct
 model <- as.formula(response ~ factor(attempt) + (1|pic_name) + (attempt|id))
 dt5 <- subset(dtrend, experiment==5)
-fp5 <- glmer(model,family="binomial",data=dt5)
+fp5 <- glmer(model,family="binomial",data=dt5, control=glmerControl(optimizer="bobyqa"))
 res5 <- get_estimates(fp5)
 res5$pvalue <- as.character(res5$pvalue)
 res5$pvalue[2:10] <- round(as.numeric(res5$pvalue[2:10]),3)
 
 dt6 <- subset(dtrend, experiment==6)
-fp6 <- glmer(model,family="binomial",data=dt6)
+fp6 <- glmer(model,family="binomial",data=dt6, control=glmerControl(optimizer="bobyqa"))
 res6 <- get_estimates(fp6)
 res6$pvalue <- as.character(res6$pvalue)
 res6$pvalue[1:10] <- round(as.numeric(res6$pvalue[1:10]),3)
 
 dt7 <- subset(dtrend, experiment==7)
-fp7 <- glmer(model,family="binomial",data=dt7)
+fp7 <- glmer(model,family="binomial",data=dt7, control=glmerControl(optimizer="bobyqa"))
 res7 <- get_estimates(fp7)
 res7$pvalue <- as.character(res7$pvalue)
 res7$pvalue[1:10] <- round(as.numeric(res7$pvalue[1:10]),3)

### 2015-08-30 R/sociology_paper.R
getting fp7 to converge

@@ -747,20 +747,16 @@ model <- as.formula(response ~ factor(attempt) + (1|pic_name) + (attempt|id))
 dt5 <- subset(dtrend, experiment==5)
 fp5 <- glmer(model,family="binomial",data=dt5, control=glmerControl(optimizer="bobyqa"))
 res5 <- get_estimates(fp5)
-res5$pvalue <- as.character(res5$pvalue)
-res5$pvalue[2:10] <- round(as.numeric(res5$pvalue[2:10]),3)
 
 dt6 <- subset(dtrend, experiment==6)
 fp6 <- glmer(model,family="binomial",data=dt6, control=glmerControl(optimizer="bobyqa"))
 res6 <- get_estimates(fp6)
-res6$pvalue <- as.character(res6$pvalue)
-res6$pvalue[1:10] <- round(as.numeric(res6$pvalue[1:10]),3)
 
 dt7 <- subset(dtrend, experiment==7)
-fp7 <- glmer(model,family="binomial",data=dt7, control=glmerControl(optimizer="bobyqa"))
+library(optimx)
+fp7 <- glmer(model,family="binomial",data=dt7, control=glmerControl(optimizer="optimx",
+                                                                    optCtrl=list(method="L-BFGS-B")))
 res7 <- get_estimates(fp7)
-res7$pvalue <- as.character(res7$pvalue)
-res7$pvalue[1:10] <- round(as.numeric(res7$pvalue[1:10]),3)
 
 results <- data.frame(res5,g1=" ",res6,g2=" ", res7)
 rownames(results) <- c("$\\mu$", paste("$\\alpha", 2:10, "$",sep=""),

### 2015-08-30 sociology_paper.Rnw
getting fp7 to converge
table 5 to confidence intervals

@@
 \subsection{Demographic Factors}  %In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table~\ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
+\hh{XXX compare demographics to US Census data - we see relatively young, relatively highly educated participants compared to general population.
+Being amazon turkers implicitly assures a relatively high level of technological fluency.
+}
 \noindent
@@
                    &  Error $(\sigma)$& \multicolumn{5}{l}{0.69} & \multicolumn{4}{l}{\phantom{-}0.89}  \\ 
-\cline{3-6} \cline{8-11}
+\cline{3-6} \cline{8-11}\\
+\hline
+\multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
 \end{tabular}}
@@
 
-\begin{table}[htbp]
-\centering
-\caption{Parameter estimates of models \eqref{eqn:trend_response} fitted to data from three different experiments using detection rate as the response to assess learning trend.
+% \begin{table}[htbp]
+% \centering
+% \caption{Parameter estimates of models \eqref{eqn:trend_response} fitted to data from three different experiments using detection rate as the response to assess learning trend.
 Attempt number is fitted as a factor to enable modeling any sort of non-linear learning trend.
 Only experiment~6 shows evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
-\scalebox{0.81}{
-\begin{tabular}{rrrrccrrrccrrrc}
-  \hline
+% \scalebox{0.81}{
+% \begin{tabular}{rrrrccrrrccrrrc}
+%   \hline
+% & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
+% 
+% \cline{2-5} \cline{7-10} \cline{12-15} 
+%  Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
+%   \hline
+% Fixed &  &  &  & && &  & & & & & &  & \\ 
+% $\mu$ & -1.304 & 0.179 & -7.28 & $<$0.001 &   & -0.220 & 0.147 & -1.50 & 0.134 &   & -1.737 & 0.481 & -3.61 & $<$0.001 \\ 
+%   $\alpha_1$ & 0.000 & --- & --- & --- &   & 0.000 & --- & --- & --- &  & 0.000 & --- & --- & ---  \\ 
+%   $\alpha_2$ & 0.270 & 0.219 & 1.24 & 0.217 &   & 0.262 & 0.158 & 1.66 & 0.098 &   & -0.456 & 0.385 & -1.18 & 0.237 \\ 
+%   $\alpha_3$ & -0.178 & 0.226 & -0.79 & 0.432 &   & 0.342 & 0.157 & 2.18 & 0.029 &   & -0.105 & 0.386 & -0.27 & 0.786 \\ 
+%   $\alpha_4$ & 0.083 & 0.224 & 0.37 & 0.712 &   & 0.358 & 0.159 & 2.26 & 0.024 &   & -0.378 & 0.381 & -0.99 & 0.322 \\ 
+%   $\alpha_5$ & 0.298 & 0.224 & 1.33 & 0.183 &   & 0.376 & 0.159 & 2.36 & 0.018 &   & -0.107 & 0.385 & -0.28 & 0.781 \\ 
+%   $\alpha_6$ & 0.042 & 0.231 & 0.18 & 0.857 &   & 0.246 & 0.158 & 1.56 & 0.120 &   & 0.026 & 0.407 & 0.06 & 0.949 \\ 
+%   $\alpha_7$ & 0.283 & 0.230 & 1.23 & 0.217 &   & 0.160 & 0.159 & 1.01 & 0.314 &   & 0.057 & 0.401 & 0.14 & 0.886 \\ 
+%   $\alpha_8$ & -0.045 & 0.233 & -0.19 & 0.847 &   & 0.341 & 0.160 & 2.13 & 0.033 &   & -0.003 & 0.394 & -0.01 & 0.994 \\ 
+%   $\alpha_9$ & -0.195 & 0.232 & -0.84 & 0.400 &   & 0.378 & 0.160 & 2.36 & 0.018 &   & 0.204 & 0.436 & 0.47 & 0.639 \\ 
+%   $\alpha_{10}$ & 0.513 & 0.228 & 2.25 & 0.024 &   & 0.192 & 0.163 & 1.18 & 0.238 &   & -0.213 & 0.432 & -0.49 & 0.622 \\ 
+% \multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
+%   $\sigma^2_a$ &  0.017 &&  &  &   &  0.034 &&  &  &   &  0.163 &&  &  \\ 
+%   $\sigma^2_u$  & 0.848 &&  &  &    & 0.903 &&  &  &   &  $<$0.001 &&  &  \\ 
+%   $\sigma^2_{\ell}$  & 1.476 &&  &  &    & 1.418 &&  &  &   &  3.314 &&  &  \\ 
+%    \hline
+% \end{tabular}
+% }
+% \label{tbl:model_result_response}
+% \end{table}
+
+% latex table generated in R 3.2.2 by xtable 1.7-4 package
+% Sun Aug 30 11:15:41 2015
+\begin{table}[hbtp]
+\centering
+\caption{\label{tbl:model_result_response}Parameter estimates of models \eqref{eqn:trend_response} fitted to data from three different experiments using detection rate as the response to assess learning trend.
+Attempt number is fitted as a factor to enable modeling any sort of non-linear learning trend.
+Only experiment~6 shows some evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
+\scalebox{.81}{
+\begin{tabular}{rr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
+\cline{2-5} \cline{7-10} \cline{12-15} 
 & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
-
 \cline{2-5} \cline{7-10} \cline{12-15} 
- Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
-  \hline
-Fixed &  &  &  & && &  & & & & & &  & \\ 
-$\mu$ & -1.304 & 0.179 & -7.28 & $<$0.001 &   & -0.220 & 0.147 & -1.50 & 0.134 &   & -1.737 & 0.481 & -3.61 & $<$0.001 \\ 
-  $\alpha_1$ & 0.000 & --- & --- & --- &   & 0.000 & --- & --- & --- &  & 0.000 & --- & --- & ---  \\ 
-  $\alpha_2$ & 0.270 & 0.219 & 1.24 & 0.217 &   & 0.262 & 0.158 & 1.66 & 0.098 &   & -0.456 & 0.385 & -1.18 & 0.237 \\ 
-  $\alpha_3$ & -0.178 & 0.226 & -0.79 & 0.432 &   & 0.342 & 0.157 & 2.18 & 0.029 &   & -0.105 & 0.386 & -0.27 & 0.786 \\ 
-  $\alpha_4$ & 0.083 & 0.224 & 0.37 & 0.712 &   & 0.358 & 0.159 & 2.26 & 0.024 &   & -0.378 & 0.381 & -0.99 & 0.322 \\ 
-  $\alpha_5$ & 0.298 & 0.224 & 1.33 & 0.183 &   & 0.376 & 0.159 & 2.36 & 0.018 &   & -0.107 & 0.385 & -0.28 & 0.781 \\ 
-  $\alpha_6$ & 0.042 & 0.231 & 0.18 & 0.857 &   & 0.246 & 0.158 & 1.56 & 0.120 &   & 0.026 & 0.407 & 0.06 & 0.949 \\ 
-  $\alpha_7$ & 0.283 & 0.230 & 1.23 & 0.217 &   & 0.160 & 0.159 & 1.01 & 0.314 &   & 0.057 & 0.401 & 0.14 & 0.886 \\ 
-  $\alpha_8$ & -0.045 & 0.233 & -0.19 & 0.847 &   & 0.341 & 0.160 & 2.13 & 0.033 &   & -0.003 & 0.394 & -0.01 & 0.994 \\ 
-  $\alpha_9$ & -0.195 & 0.232 & -0.84 & 0.400 &   & 0.378 & 0.160 & 2.36 & 0.018 &   & 0.204 & 0.436 & 0.47 & 0.639 \\ 
-  $\alpha_{10}$ & 0.513 & 0.228 & 2.25 & 0.024 &   & 0.192 & 0.163 & 1.18 & 0.238 &   & -0.213 & 0.432 & -0.49 & 0.622 \\ 
-\multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
-  $\sigma^2_a$ &  0.017 &&  &  &   &  0.034 &&  &  &   &  0.163 &&  &  \\ 
-  $\sigma^2_u$  & 0.848 &&  &  &    & 0.903 &&  &  &   &  $<$0.001 &&  &  \\ 
-  $\sigma^2_{\ell}$  & 1.476 &&  &  &    & 1.418 &&  &  &   &  3.314 &&  &  \\ 
-   \hline
-\end{tabular}
-}
-\label{tbl:model_result_response}
+Effect & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  \\ 
+\cline{2-5} \cline{7-10} \cline{12-15} 
+Fixed \\
+$\mu$ & -1.30 & -1.69 & -0.92 & *** &   & -0.22 & -0.51 & 0.07 &  &   & -1.71 & -2.77 & -0.65 & ** \\ [3pt]
+  $\alpha_2$ & 0.27 & -0.19 & 0.73 &  &   & 0.26 & -0.06 & 0.58 &  &   & -0.49 & -1.31 & 0.32 &  \\ 
+  $\alpha_3$ & -0.18 & -0.65 & 0.29 &  &   & 0.34 & 0.02 & 0.66 & * &   & -0.19 & -1.01 & 0.63 &  \\ 
+  $\alpha_4$ & 0.08 & -0.39 & 0.55 &  &   & 0.36 & 0.04 & 0.68 & * &   & -0.47 & -1.28 & 0.35 &  \\ 
+  $\alpha_5$ & 0.30 & -0.17 & 0.77 &  &   & 0.37 & 0.05 & 0.70 & * &   & -0.16 & -0.96 & 0.65 &  \\ 
+  $\alpha_6$ & 0.04 & -0.44 & 0.53 &  &   & 0.24 & -0.07 & 0.56 &  &   & -0.03 & -0.87 & 0.82 &  \\ 
+  $\alpha_7$ & 0.28 & -0.20 & 0.77 &  &   & 0.16 & -0.16 & 0.48 &  &   & 0.02 & -0.81 & 0.84 &  \\ 
+  $\alpha_8$ & -0.04 & -0.54 & 0.45 &  &   & 0.34 & 0.02 & 0.66 & * &   & -0.08 & -0.89 & 0.73 &  \\ 
+  $\alpha_9$ & -0.20 & -0.69 & 0.30 &  &   & 0.38 & 0.05 & 0.70 & * &   & 0.15 & -0.71 & 1.02 &  \\ 
+  $\alpha_{10}$ & 0.51 & 0.03 & 1.00 & * &   & 0.19 & -0.14 & 0.52 &  &   & -0.28 & -1.14 & 0.58 &  \\ 
+  Random \\
+  $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.85}  & \multicolumn{5}{l}{\phantom{-}0.90} & 0.76   \\ 
+  $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.02}  & \multicolumn{5}{l}{\phantom{-}0.03} & 0.06   \\ 
+  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}1.48}  & \multicolumn{5}{l}{\phantom{-}1.42} & 3.37   \\ 
+\cline{2-5} \cline{7-10} \cline{12-15} 
+\end{tabular}}
 \end{table}

### 2015-08-30 R/sociology_paper.R

@@ -759,10 +759,12 @@ fp7 <- glmer(model,family="binomial",data=dt7, control=glmerControl(optimizer="o
 res7 <- get_estimates(fp7)
 
 results <- data.frame(res5,g1=" ",res6,g2=" ", res7)
+# remove covariance term
+results <- results[-11, ]
 rownames(results) <- c("$\\mu$", paste("$\\alpha", 2:10, "$",sep=""),
-                         "$\\sigma^2_a$","$\\sigma^2_u$","$\\sigma^2_l$")
+                         "$\\sigma^2_u$","$\\sigma^2_a$","$\\sigma^2_l$")
 
-print(xtable(results, digits=3),  sanitize.text.function = function(x){x})
+print(xtable(results, digits=2),  sanitize.text.function = function(x){x})
 
 
 

### 2015-08-30 R/sociology_paper.R
first stab at table 7

@@ -686,9 +686,10 @@ f7 <- lmer(model, data=dt7)
 estimates7 <- get_estimates(f7)
 
 estimates <- data.frame(estimates5,g1=" ",estimates6,g2=" ", estimates7)
-rownames(estimates) <- c("$\\mu$","$\\alpha_1$","$\\alpha$", "$\\sigma^2_a$",
-                         "$\\sigma^2_u$","$\\sigma^2_l$","$\\sigma^2$")
-dgt <- c(rep(0,7), rep(3,14), rep(2,7), rep(0,7))
+estimates <- estimates[-4,]
+rownames(estimates) <- c("$\\mu$","$\\alpha_1$","$\\alpha$", 
+                         "$\\sigma^2_u$","$\\sigma^2_a$", "$\\sigma^2_l$","$\\sigma^2$")
+dgt <- c(rep(0,7), rep(2,14), rep(2,7), rep(0,7))
 dgts <- matrix(rep(dgt,3), ncol=15)
 print(xtable(estimates, digits=dgts),  sanitize.text.function = function(x){x})
 

### 2015-08-31 sociology_paper.Rnw
first stab at table 7
stab at table 3

@@
 \label{tbl:anova_factor}
+\end{table}
+
+% latex table generated in R 3.2.2 by xtable 1.7-4 package
+% Sun Aug 30 22:43:05 2015
+\begin{table}[ht]
+\caption{\label{tbl:anova_factor}}
+\centering
+\begin{tabular}{rrrrrlrrrr}
+  \hline
+ & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\   \hline
+Full & 51872.44 & 0.00 & 6.00 & 0.00 &  & 23791.91 & 0.00 & 6.00 & 0.00 \\ 
+  Age & 51329.26 & 543.18 & 6.00 & 0.00 &  & 23766.41 & 25.50 & 6.00 & 0.00 \\ 
+  Country & 51496.37 & 376.07 & 2.00 & 0.00 &  & 23765.88 & 26.03 & 2.00 & 0.00 \\ 
+  Degree & 51756.21 & 116.23 & 4.00 & 0.00 &  & 23768.15 & 23.76 & 4.00 & 0.00 \\ 
+  Gender & 51804.32 & 68.12 & 1.00 & 0.00 &  & 23789.74 & 2.17 & 1.00 & 0.14 \\ 
+   \hline
+\end{tabular}
 \end{table}
@@
 
-\begin{table}[htbp]
-\centering
-\caption{Parameter estimates of model~\eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
+% \begin{table}[htbp]
+% \centering
+% \caption{Parameter estimates of model~\eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
 Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.
 }
-\scalebox{0.81}{
-\begin{tabular}{rrrrccrrrccrrrc}
-  \hline
+% \scalebox{0.81}{
+% \begin{tabular}{rrrrccrrrccrrrc}
+%   \hline
+% & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
+% 
+% \cline{2-5} \cline{7-10} \cline{12-15} 
+%  Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
+%   \hline
+% Fixed &  &  &  & && &  & & & & & &  & \\ 
+% $\mu$ & 3.817 & 0.039 & 97.38 & $<$0.001 &   & 3.901 & 0.033 & 118.19 & $<$0.001 &   & 3.731 & 0.054 & 69.04 & $<$0.001 \\ 
+%   $\alpha_1$ & 0.326 & 0.035 & 9.35 & $<$0.001 &   & 0.335 & 0.029 & 11.40 & $<$0.001 &   & 0.280 & 0.050 & 5.63 & $<$0.001 \\ 
+%   $\alpha$ & -0.038 & 0.004 & -9.30 & $<$0.001 &   & -0.039 & 0.004 & -10.19 & $<$0.001 &   & -0.029 & 0.007 & -4.22 & $<$0.001 \\ 
+% \multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
+%   $\sigma^2_a$ &   0.027 &  &  &   &  & 0.045 &  &  &   &  & 0.049 &  &  \\ 
+%   $\sigma^2_u$ &   0.509 &  &  &   &  & 0.495 &  &  &   &  & 0.366 &  &  \\ 
+%   $\sigma^2_{\ell}$ &   0.091 &  &  &   &  & 0.199 &  &  &   &  & 0.235 &  &  \\ 
+%   $\sigma^2$ &   0.460 &  &  &   &  & 0.501 &  &  &   &  & 0.454 &  & &  \\ 
+%    \hline
+% \end{tabular}
+% }
+% \label{tbl:model_result_time}
+% \end{table}
+
+% latex table generated in R 3.2.2 by xtable 1.7-4 package
+% Sun Aug 30 11:42:08 2015
+\begin{table}[bhtp]
+\caption{\label{tbl:model_result_time}Parameter estimates of model~\eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
+Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments~5,~6 and~7.}
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
+\cline{2-5} \cline{7-10} \cline{12-15} 
 & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
@@
 \cline{2-5} \cline{7-10} \cline{12-15} 
- Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
-  \hline
-Fixed &  &  &  & && &  & & & & & &  & \\ 
-$\mu$ & 3.817 & 0.039 & 97.38 & $<$0.001 &   & 3.901 & 0.033 & 118.19 & $<$0.001 &   & 3.731 & 0.054 & 69.04 & $<$0.001 \\ 
-  $\alpha_1$ & 0.326 & 0.035 & 9.35 & $<$0.001 &   & 0.335 & 0.029 & 11.40 & $<$0.001 &   & 0.280 & 0.050 & 5.63 & $<$0.001 \\ 
-  $\alpha$ & -0.038 & 0.004 & -9.30 & $<$0.001 &   & -0.039 & 0.004 & -10.19 & $<$0.001 &   & -0.029 & 0.007 & -4.22 & $<$0.001 \\ 
-\multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
-  $\sigma^2_a$ &   0.027 &  &  &   &  & 0.045 &  &  &   &  & 0.049 &  &  \\ 
-  $\sigma^2_u$ &   0.509 &  &  &   &  & 0.495 &  &  &   &  & 0.366 &  &  \\ 
-  $\sigma^2_{\ell}$ &   0.091 &  &  &   &  & 0.199 &  &  &   &  & 0.235 &  &  \\ 
-  $\sigma^2$ &   0.460 &  &  &   &  & 0.501 &  &  &   &  & 0.454 &  & &  \\ 
-   \hline
-\end{tabular}
-}
-\label{tbl:model_result_time}
+\multicolumn{2}{c}{Effect \hfill Est} & 2.5\% & 97.5\% &  &  & Est & 2.5\% & 97.5\% & &  & Est & 2.5\% & 97.5\% & \\ 
+  \cline{2-5} \cline{7-10} \cline{12-15}
+\multicolumn{2}{l}{Fixed}\\
+$\mu$ & 3.82 & 3.74 & 3.89 & *** &   & 3.90 & 3.84 & 3.97 & *** &   & 3.73 & 3.62 & 3.84 & *** \\ 
+  $\alpha_1$ & 0.33 & 0.26 & 0.39 & *** &   & 0.34 & 0.28 & 0.39 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
+  $\alpha$ & -0.04 & -0.05 & -0.03 & *** &   & -0.04 & -0.05 & -0.03 & *** &   & -0.03 & -0.04 & -0.02 & *** \\ 
+\multicolumn{2}{l}{Random}\\
+  $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.51}   & \multicolumn{5}{l}{\phantom{-}0.49}  & 0.37   \\ 
+  $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.03}  & \multicolumn{5}{l}{\phantom{-}0.04}   & 0.05   \\ 
+  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}0.09}   & \multicolumn{5}{l}{\phantom{-}0.20}   & 0.23   \\ 
+  $\sigma^2$ & \multicolumn{5}{l}{\phantom{-}0.46}   & \multicolumn{5}{l}{\phantom{-}0.50}  & 0.45  \\ 
+   \cline{2-5} \cline{7-10} \cline{12-15}
+\end{tabular}}
 \end{table}
-
 %The lineup variance of experiment~7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model~\eqref{eqn:trend_response}.

### 2015-08-31 R/sociology_paper.R

@@ -136,7 +136,7 @@ demographics <- merge(demographics,ip.details,all.x=TRUE, by="ip_address")
 demographics$country <- as.character(demographics$country_name)
 demographics$country[(demographics$country_code != "IN") & (demographics$country_code != "US")] <- "Rest of the world"
 demographics$country[demographics$country=="Namibia"] <- "Rest of the world"
-demographics$country[demographics$country==NA] <- "Rest of the world"
+demographics$country[is.na(demographics$country)] <- "Rest of the world"
 country <- factor(demographics$country)
 demographics$country <- factor(country,levels=levels(country)[c(3,1,2)])
 
@@ -314,45 +314,52 @@ dgt <- c(rep(0,15), rep(2,45), rep(0,15))
 dgts <- matrix(rep(dgt,2), ncol=10)
 print(xtable(est.factor, digits=dgts),  sanitize.text.function = function(x){x})
 
-fta <- lmer(log(time_taken)~country+degree+gender_level+(1|pic_name), 
-           data=demographics)
-ftc <- lmer(log(time_taken)~age_level+degree+gender_level+(1|pic_name), 
-           data=demographics)
-ftd <- lmer(log(time_taken)~age_level+country+gender_level+(1|pic_name), 
-           data=demographics)
-ftg <- lmer(log(time_taken)~age_level+country+degree+(1|pic_name), 
-           data=demographics)
+ftdata <- ft@frame
+names(ftdata)[1] <- "time_taken"
+ftdata$time_taken <- exp(ftdata$time_taken)
+
+ft <- update(ft, data=ftdata)
+fta <- update(ft, .~.-age_level, data=ftdata)
+ftc <- update(ft, .~.-country, data=ftdata)
+ftd <- update(ft, .~.-degree, data=ftdata)
+ftg <- update(ft, .~.-gender_level, data=ftdata)
 
 tag <- anova(ft,fta)
 tcn <- anova(ft,ftc)
 tde <- anova(ft,ftd)
 tge <- anova(ft,ftg)
 
-anova.time <- rbind(data.frame(round(tag[1,1:4],2),round(tag[2,5:7]/1,2)),
-                    data.frame(round(tcn[1,1:4],2),round(tcn[2,5:7]/1,2)),
-                    data.frame(round(tde[1,1:4],2),round(tde[2,5:7]/1,2)),
[…]

### 2015-08-31 sociology_paper.Rnw
table 6
table 3

@@
 
-\begin{table}[htbp]
-\centering
-\caption{Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
+% \begin{table}[htbp]
+% \centering
+% \caption{Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
 Gender does not have a significant effect on detection rate, but does on time to respond.
 All factors significantly affect time to respond.}
-\scalebox{.9}{
-\begin{tabular}{rlrrrrrr}
-  \hline
-\multicolumn{2}{c}{Model}  & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
-  \hline
-Proportion Correct &Full & 23822.00 &23943.00 &-11896.00  & &  &  \\ [3pt]
-%& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
-{Reduced Models}  &- Age & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
-  &- Country  & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
-  &- Education  & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
-  &- Gender  & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
-   \hline
-   Log Time Taken &Full &  51904.00 &52034.00 & -25936.00 & &  &  \\ [3pt]
-%& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
-Reduced Models &- Age  & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
- &- Country  & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
-  &- Education  & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
-  &- Gender  & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
-\hline
-\end{tabular}
-}
-\label{tbl:anova_factor}
-\end{table}
+% \scalebox{.9}{
+% \begin{tabular}{rlrrrrrr}
+%   \hline
+% \multicolumn{2}{c}{Model}  & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
+%   \hline
+% Proportion Correct &Full & 23822.00 &23943.00 &-11896.00  & &  &  \\ [3pt]
+% %& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
+% {Reduced Models}  &- Age & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
+%   &- Country  & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
+%   &- Education  & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
+%   &- Gender  & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
+%    \hline
+%    Log Time Taken &Full &  51904.00 &52034.00 & -25936.00 & &  &  \\ [3pt]
+% %& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
+% Reduced Models &- Age  & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
+%  &- Country  & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
+%   &- Education  & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
+%   &- Gender  & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
+% \hline
+% \end{tabular}
+% }
+% \label{tbl:anova_factor}
+% \end{table}
 
@@
 \begin{table}[ht]
-\caption{\label{tbl:anova_factor}}
-\centering
-\begin{tabular}{rrrrrlrrrr}
-  \hline
- & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\   \hline
-Full & 51872.44 & 0.00 & 6.00 & 0.00 &  & 23791.91 & 0.00 & 6.00 & 0.00 \\ 
-  Age & 51329.26 & 543.18 & 6.00 & 0.00 &  & 23766.41 & 25.50 & 6.00 & 0.00 \\ 
-  Country & 51496.37 & 376.07 & 2.00 & 0.00 &  & 23765.88 & 26.03 & 2.00 & 0.00 \\ 
-  Degree & 51756.21 & 116.23 & 4.00 & 0.00 &  & 23768.15 & 23.76 & 4.00 & 0.00 \\ 
-  Gender & 51804.32 & 68.12 & 1.00 & 0.00 &  & 23789.74 & 2.17 & 1.00 & 0.14 \\ 
-   \hline
-\end{tabular}
+\caption{\label{tbl:anova_factor}Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
+Gender does not have a significant effect on detection rate, but does on time to respond.
+All factors significantly affect time to respond.}
+\centering
+\scalebox{0.9}{
+\begin{tabular}{lrrrrlrrrr}
+& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
+\cline{2-5} \cline{7-10} 
+ & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
+Full  & 51872.4 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 23792.0 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
+ - Age & 51329.3 & 543.2 & 6 & $<$0.001 & \hspace{0.5cm} & 23766.4 & 26.0 & 6 & $<$0.001 \\ 
+ - Country & 51496.4 & 376.1 & 2 & $<$0.001 &  & 23765.9 & 26.0 & 2 & $<$0.001 \\ 
+ - Degree & 51756.2 & 116.2 & 4 & $<$0.001 &  & 23768.2 & 23.8 & 4 & $<$0.001 \\ 
+ - Gender & 51804.3 & 68.1 & 1 & $<$0.001 &  & 23789.7 & 2.2 & 1 & 0.140 \\ 
+\cline{2-5} \cline{7-10}
+\end{tabular}}
 \end{table}
@@
   $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.02}  & \multicolumn{5}{l}{\phantom{-}0.03} & 0.06   \\ 
-  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}1.48}  & \multicolumn{5}{l}{\phantom{-}1.42} & 3.37   \\ 
-\cline{2-5} \cline{7-10} \cline{12-15} 
+  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}1.48}  & \multicolumn{5}{l}{\phantom{-}1.42} & 3.37   \\ \cline{2-5} \cline{7-10} \cline{12-15} 
+\\
+\hline
+\multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
 \end{tabular}}
@@
 \centering
-\scalebox{0.8}{
+\scalebox{0.9}{
 \begin{tabular}{lr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}

### 2015-08-31 sociology_paper.Rnw
done for now, I think
added two references and comparison to surveys

@@
 \subsection{Demographic Factors}  %In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table~\ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
-\hh{XXX compare demographics to US Census data - we see relatively young, relatively highly educated participants compared to general population.
-Being amazon turkers implicitly assures a relatively high level of technological fluency.
-}
 \noindent
@@
 
+Compared to data from the 2010 US Census \cite{census}, participants were both relatively younger, and more highly educated than the general US population. 
+Being turkers implicitly assures that the participants in these experiments have a relatively high level of technological fluency.
+In fact, the demographics of participants align somewhat better with demographics of internet users published by
+\citet{pew}.
+
 The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where the proportion of female participants is lower.
@@
 \centering
-\scalebox{.85}{
+\scalebox{.9}{
 \begin{tabular}{rrr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
@@
 Only experiment~6 shows some evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
-\scalebox{.81}{
+\scalebox{.9}{
 \begin{tabular}{rr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}

### 2015-09-02 R/add_interaction.R
make_interactive

@@ -1,0 +1,18 @@
+make_interactive <- function(filename, script, toggle="toggle", background="#e5e5e5") {
+# use #ffffff for white background
+  require(gridSVG)
+  grobs <- grid.ls()
+  
+  idx <- grep("panel-", grobs$name)
+  for (i in idx) { 
+    grid.garnish(grobs$name[i],
+                 onmouseover=paste("frame('",grobs$name[i+2], ".1')", sep=""),
+                 onmouseout=paste("deframe('",grobs$name[i+2], ".1')", sep=""), 
+                 onmousedown=paste(sprintf("%shigh(evt, '", toggle, background),grobs$name[i+2], ".1')", sep=""))
+  }
+
+  # use script on server to get locally executable javascript code
+  # or use inline option
+  grid.script(filename=script)
+  gridToSVG(filename)
+}

### 2016-11-30 sociology_paper.Rnw
Roy citation update
typo fixed

@@
 Visual statistical inference is a way to determine significance of patterns found while  exploring data with graphics.
-Visual inference relies on human observers inspecting a lineup: one  data plot randomly placed among a set of decoy plots.
+It relies on human observers inspecting a lineup: one  data plot randomly placed among a set of decoy plots.
 Each individual is different in their cognitive psychology and judiciousness, which can affect the visual inference.
@@
 In our study of the demographic effects on performance across a set of different experiments we found some statistically significant effects.
-Age group 36-40, Countries other than India and United states, people who have a graduate degree have a significantly higher detection rate.
+Age group 36-40, countries other than India and United states, people who have a graduate degree have a significantly higher detection rate.
 Gender does not have any significant effect.

### 2017-10-12 R/sociology_paper.R
Cleaning the codes

@@ -293,6 +293,7 @@ get_estimates <- function(fit){
 # --------------------------------------------------------------
 # Model fitting with demographic factors
 # testing significance of demographic factor main effects
+# Finally it will generate results table 4 in the paper
 # --------------------------------------------------------------
 ft <- lmer(log(time_taken)~age_level+country+degree+gender_level+(1|pic_name), 
            data=demographics)
@@ -526,9 +527,10 @@ ggplot(turk.summary, aes(x = factor(plot, levels=plot[order(percent_rejected)]),
 ggsave("../images/rejected_task.pdf", width=6, height=4)
 
 
-
+# ----------------------------------------------------------------
 # Drawing the map of turk participants
-
+# This will create the map of figure 3 in the paper
+# ----------------------------------------------------------------
 map_theme <- list(theme(panel.grid.minor = element_blank(),
                         panel.grid.major = element_blank(),
                         panel.background = element_blank(),
@@ -561,7 +563,10 @@ p<- ggplot() +
   theme(legend.position="none", axis.text = element_blank(), axis.title=element_blank()) 
 ggsave("../images/turker_location_experiment.pdf", width=8, height=10)
 
+# ----------------------------------------------------------------
 # country-wise participants counts
+# This will produce figure 3 bar chart
+# ----------------------------------------------------------------
 qplot(factor(country, levels=names(table(country))[order(table(country))]), 
       geom="bar", fill="A", data=subset(turker,complete.cases(turker)))+ 
   coord_flip() +ylab("Number of participants") + xlab("Country") +
@@ -657,6 +662,10 @@ ggplot() +
 
 dpt$Experiment <- dpt$experiment
 dmt$Experiment <- dmt$experiment
+
+# ----------------------------------------------------------------
+# Generating figure 7 of the paper
+# ----------------------------------------------------------------
 ggplot() + 
   geom_smooth(aes(attempt,resid, group=id),method="lm", se=F, data=subset(dpt, attempt>1), colour=rgb(0,0,0, alpha=.05))+
   geom_smooth(aes(attempt,mean_resid), data= subset(dmt, attempt>1), method="lm", se=F, size=I(1.2)) +
@@ -680,6 +689,10 @@ ggplot() +
 
 ggsave("../images/learning_trend_time.pdf", width=10.5, height = 3.5)
 
+# ----------------------------------------------------------------
[…]

### 2017-10-12 R/sociology_paper.R
adding description

@@ -887,7 +887,7 @@ ggplot(subset(dd, plot_type=="Interaction")) +
   facet_wrap(~plot_type)
 
 
-# proportion of correct evaluation by location and null
+# proportion of correct evaluation (Detection Rate) by location and null
 
 nulls <- strsplit(as.character(dat9$param_value),split="_")
 dat9$nulls <- paste("null_",sapply(nulls, function(x) return(x[2])), sep="")
@@ -912,6 +912,10 @@ p_prop <- ggplot(p.dat, aes(plot_loc, prop_correct,  group=nulls))+
 
 ggsave(p_prop,file= "../images/proportion_nulls.pdf", width=9, height=4)
 
+# ----------------------------------------------------------------
+# Generating figure 8 of the paper
+# ----------------------------------------------------------------
+
 l.dat <- data.frame(plot_location=1:20, plot_int=0, plot_gen=0)
 plot_int <- unique(subset(dat9, plot_type=="Interaction")$plot_location)
 plot_gen <- unique(subset(dat9, plot_type=="Genotype")$plot_location)
@@ -933,7 +937,10 @@ print(p_int,vp = vp1)
 print(p_gen,vp = vp2)
 dev.off()
 
+# ----------------------------------------------------------------
 # Fitting MANOVA model
+# This produces results table 7
+# ----------------------------------------------------------------
 df <- ddply(subset(dat9, plot_type != "Filter"), 
             .(plot_type, plot_location, nulls, pic_name),summarise,
             prop_correct = mean(response))

### 2017-10-12 sociology_paper.Rnw
typo fixed

@@
 \caption{Overview of 10 different Turk experiments, from which data was collected to study human factor effects.
-All of the experimental data was used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.}
+All of the experimental data were used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.}
 \begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 

### 2017-10-17 sociology_paper.Rnw
updated age group

@@
 \begin{enumerate} \leftmargin 5cm  \itemsep 0in
-\item {\it Age group}, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, 51-55, 56-60, above 60.
+\item {\it Age group}, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, above 50.
 \item {\it Gender}, male or female.
@@
   \hline
-Gender & Male & 1348&57.63 & 48.51 & 13493 \\ 
-   & Female & 991&42.37 & 43.75 & 10564 \\ 
+Gender & Male & 1351 & 57.74 & 48.75 & 13542 \\ 
+   & Female & 989 & 42.26 & 43.69 & 10538 \\ 
 \hline
-  Education & High school or less & 193 &8.24& 37.21 & 2241 \\ 
-   & Some under graduate courses & 418 & 17.85&42.84 & 4070 \\ 
-   & Under graduate degree & 584 &24.93& 44.29 & 5775 \\ 
-   & Some graduate courses & 245 &10.46& 43.43 & 2460 \\ 
-   & Graduate degree & 902 &38.51& 52.18 & 9511 \\ 
+  Education & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
+   & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
+   & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
+   & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
+   & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
 \hline
-  Age & 18-25 & 740 &31.61& 42.97 & 7311 \\ 
-   & 26-30 & 547 &23.36& 46.27 & 5585 \\ 
-   & 31-35 & 376 &16.06& 44.27 & 3923 \\ 
-   & 36-40 & 257 &10.98& 55.03 & 2714 \\ 
-   & 41-45 & 141 &6.02& 43.90 & 1519 \\ 
-   & 46-50 &  95 &4.05& 49.29 & 1003 \\ 
-   & 51-55 &  83 &3.54& 48.67 & 867 \\ 
-   & 56-60 &  64 & 2.73& 59.73 & 678 \\ 
-   & above 60 & 38 &1.62& 48.67 & 457 \\ 
+  Age & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
+   & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
+   & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
+   & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
+   & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
+   & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
+   & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
 \hline   
-  Country  & United States & 1087 &46.83& 39.64 & 10769 \\  
-  & India & 980 &42.22& 52.63 & 10227 \\ 
-  & Rest of the world & 254&10.94 & 46.86 & 2819 \\
+  Country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
+   & India & 980 & 41.88 & 52.63 & 10238 \\ 
+   & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\
 

### 2017-10-17 R/sociology_paper.R

@@ -192,13 +192,15 @@ ggsave("../images/age_gender_within_country_bar.pdf", width=10, height=8)
 pdat <- ddply(demographics, .(country,gender_level, age_level,degree), summarize,
               per_correct = mean(response),
               avg_time = mean(time_taken))
-qplot(degree, avg_time, geom="bar", stat="identity", data=pdat[complete.cases(pdat),]) +
+ggplot(data=pdat[complete.cases(pdat),], aes(degree, avg_time)) +
+  geom_bar(stat="identity") +
   facet_grid(age_level~country+gender_level) + coord_flip() + 
   xlab("Academic Study") + ylab("Average time taken")
 
 ggsave("../images/age_gender_within_country_time.pdf", width=10, height=8)
 
-qplot(degree, per_correct*100, geom="bar", stat="identity", data=pdat[complete.cases(pdat),]) +
+ggplot(data=pdat[complete.cases(pdat),], aes(degree, per_correct*100)) +
+  geom_bar(stat="identity") +
   facet_grid(age_level~country+gender_level) + coord_flip() + 
   xlab("Academic Study") + ylab("Percent correct") +
   scale_y_continuous(breaks=c(25,50,75,100)) 
@@ -225,12 +227,17 @@ levels(mdat$variable) <- c("(Log) Time Taken in seconds", "Detection Rate")
 mdat$variable_name <- factor(mdat$variable_name)
 levels(mdat$variable_name) <- c("Age Categories", "Country", "Education", "Gender")
 
-qplot(variable_level, value, geom="boxplot",data=mdat[complete.cases(mdat),]) +
+ggplot(data=mdat[complete.cases(mdat),], aes(variable_level, value)) +
+  geom_boxplot() +
   facet_grid(variable~variable_name, scales="free", space="free_x") +
   stat_summary(fun.y=mean, geom="point") + xlab("Levels of demographic factors")+ylab("")+
   theme(axis.text.x=element_text(angle=90, hjust=1))
 
+# ------------------------------------------------------------------
+# saving figure 4
 ggsave("../images/demographic_effect.pdf", width=6.5, height=6)
+# ------------------------------------------------------------------
+
 
 #----------------------------------------------------------------
 # Function to obtain estimates of fitted model for latex xtable
[…]

### 2017-10-17 sociology_paper.Rnw
updated section 4.1

@@
 
-\subsection{Overview of the Data} A total of 2321 participants provided feedback data on the lineups in ten different experimental studies.
+\subsection{Overview of the Data} A total of 2340 participants provided feedback data on the lineups in ten different experimental studies.
 Figure~\ref{fig:turker_location} displays the locations of participants around the world.
 Most of the participants were from the United States and India.
-There were 76 other different countries represented.
+There were 70 other different countries represented.
 This provides a diverse pool of participants.
@@
 Besides the United States and India, countries such as Canada, Romania, the United Kingdom and Macedonia have more than 10 participants each.
-The remaining 70 countries have fewer than 10 participants each.
+The remaining 64 countries have fewer than 10 participants each.
 The distribution of participants is similar in all ten experiments.
@@
 
-In terms of academic background, the majority of participants had a graduate degree (a total of 902 participants, about 38.51\%).
+In terms of academic background, the majority of participants had a graduate degree (a total of 899 participants, about 38.42\%).
 Examining the participants' location in conjunction with the academic background reveals that there is an above average  number of participants from India reporting to have a graduate degree.
@@
 
-A total of 1911 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where each participant evaulated three lineups.
+A total of 1912 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where each participant evaulated three lineups.
 As a measure to ensure a high quality in the evaluations,
@@
 Also, for some ip address, the actual geographical locations could not be retrieved.
-This resulted in some missing demographic information.
+This resulted in some missing geographic information.
 

### 2017-10-17 R/sociology_paper.R

@@ -7,6 +7,7 @@ library(lme4)
 library(xtable)
 library(grid)
 library(lubridate)
+library(dplyr)
 
 # ===================================================
 # Loading the data and some common functions
@@ -161,6 +162,18 @@ turker$exp <- factor(turker$experiment, levels=names(table(turker$experiment))[o
 #   coord_flip() + xlab("Academic Study") + labs(title="Gender") +
 #   scale_x_discrete( aes(breaks=academic_study,labels=as.charater(study)))
 
+# ------------------------------------------------------------------
+# generating statistics for section 4.1 "Overview of the Data"
+# showing number of countries the subjects are coming from
+length(unique(demographics$country_code))
+# showing countries with more than 10 participants
+countryCount <- demographics %>% 
+  group_by(country_name) %>%
+  summarise(counts = length(unique(id))) %>%
+  arrange(desc(counts))
+# showing number of linesups (pic_name) evaluated
+length(unique(demographics$pic_name))
+# ------------------------------------------------------------------
 
 # Plotting age, gender and country together
 p1 <- qplot(degree, facets=age_level~gender_level, data=turker) +
@@ -234,7 +247,8 @@ ggplot(data=mdat[complete.cases(mdat),], aes(variable_level, value)) +
   theme(axis.text.x=element_text(angle=90, hjust=1))
 
 # ------------------------------------------------------------------
-# saving figure 4
+# saving figure 4 showing effect of demographic factors and their
+# effects on log(time taken) and detection rate
 ggsave("../images/demographic_effect.pdf", width=6.5, height=6)
 # ------------------------------------------------------------------
 

### 2017-10-19 R/sociology_paper.R
updated results based on data

@@ -13,19 +13,19 @@ library(dplyr)
 # Loading the data and some common functions
 # ---------------------------------------------------
 
-raw.dat1 <- read.csv("../data/raw_data_turk1.csv")
-raw.dat2 <- read.csv("../data/raw_data_turk2.csv")
-raw.dat3 <- read.csv("../data/raw_data_turk3.csv")
-raw.dat4 <- read.csv("../data/raw_data_turk4.csv")
-raw.dat5 <- read.csv("../data/raw_data_turk5.csv")
-raw.dat6 <- read.csv("../data/raw_data_turk6.csv")
-raw.dat7 <- read.csv("../data/raw_data_turk7.csv")
-raw.dat8 <- read.csv("../data/raw_data_turk8.csv")
-raw.dat9 <- read.csv("../data/raw_data_turk9.csv")
-raw.dat10 <- read.csv("../data/raw_data_turk10.csv")
+raw.dat1 <- read.csv("../data/raw_data_turk1.csv", stringsAsFactors = FALSE)
+raw.dat2 <- read.csv("../data/raw_data_turk2.csv", stringsAsFactors = FALSE)
+raw.dat3 <- read.csv("../data/raw_data_turk3.csv", stringsAsFactors = FALSE)
+raw.dat4 <- read.csv("../data/raw_data_turk4.csv", stringsAsFactors = FALSE)
+raw.dat5 <- read.csv("../data/raw_data_turk5.csv", stringsAsFactors = FALSE)
+raw.dat6 <- read.csv("../data/raw_data_turk6.csv", stringsAsFactors = FALSE)
+raw.dat7 <- read.csv("../data/raw_data_turk7.csv", stringsAsFactors = FALSE)
+raw.dat8 <- read.csv("../data/raw_data_turk8.csv", stringsAsFactors = FALSE)
+raw.dat9 <- read.csv("../data/raw_data_turk9.csv", stringsAsFactors = FALSE)
+raw.dat10 <- read.csv("../data/raw_data_turk10.csv", stringsAsFactors = FALSE)
 
-ip.details <- read.csv("../data/ip_details.csv")
-turk.summary <- read.csv("../data/turk_summary.csv")
+ip.details <- read.csv("../data/ip_details.csv", stringsAsFactors = FALSE)
+turk.summary <- read.csv("../data/turk_summary.csv", stringsAsFactors = FALSE)
 
 source("calculate_ump_power.R") # functions to compute power
 
[…]

### 2017-10-19 submitted-v1/cover-letter.tex
submission version 1

\documentclass{letter}
\usepackage{graphics}
\usepackage{fullpage}
\topmargin=-0.2in
%\headheight=0in
\pdfminorversion=4 % as instructed by JASA file upload
\setlength\oddsidemargin{-0.2in}
\textheight=8.5in
\textwidth=7in
\def\apl#1#2{
\begin{letter}{#1}
\signature{Mahbubul Majumder, PhD.\\ Assistant Professor \\ Department of Mathematics \\University of Nebraska at Omaha \\ Omaha, NE 68118 \\ Phone: (402) 554 2734 \\ {\tt mmajumder@unomaha.edu} \\ with Dr. Heike Hoffman, Iowa State university and Dianne Cook, Monash University, Australia}
\opening{{Dear #2,}}

Please find enclosed a manuscript  ``Human Factors Influencing Visual Statistical Inference'' which we are
submitting for consideration for publication in the
Applications and Case Studies section of the Journal of American Statistical
Association.

This paper is the complement of the two foundational papers describing research which bridges
graphics, exploratory data analysis/data mining, and statistical
inference. The first paper by Buja, et al (2009) published in RSS Phil. Trans. A,
sets up the conceptual foundation of visual inference. The second paper by Majumder, et al (2013) published in JASA Theory and applications section, clarifies and defines the approach and validates it with a user study. It was one of the most read JASA articles and selected to be a feature article. This paper studies the human factors that may influence the visual inference conducting a series of 10 experiments with human subjects. We believe this paper is of interest to the JASA readership.

Thank you for your consideration of our work.

\closing{Sincerely yours,
}
\end{letter}}

\begin{document}

{{\pdfimage width 7in {letterhead.png}}}
\apl{Professor Montserrat Fuentes\\Applications and Case Studies Editor\\Journal of American Statistical Association}{Professor Fuentes}

\end{document}

### 2017-10-19 submitted-v1/R/sociology-paper-v1.R

@@ -1,0 +1,575 @@
+# The complete R code for Sociology Paper
+
+library(ggplot2)
+library(plyr)
+library(reshape2)
+library(lme4)
+library(xtable)
+library(grid)
+library(lubridate)
+library(dplyr)
+library(maps)
+
+# ==================================================================
+# Loading the data 
+# ------------------------------------------------------------------
+
+demographics <- readRDS("../data/sociology-data.rds")
+dat9 <- readRDS("../data/exp9.RDS")
+
+
+# ==================================================================
+# generating statistics for section 4.1 "Overview of the Data"
+# ------------------------------------------------------------------
+
+# showing number of countries the subjects are coming from
+length(unique(demographics$country_code))
+# showing countries with more than 10 participants
+countryCount <- demographics %>% 
+  group_by(country_name) %>%
+  summarise(counts = length(unique(id))) %>%
+  arrange(desc(counts))
+# showing number of linesups evaluated
+length(unique(demographics$lineup_id))
+
+
+
+# ==================================================================
+# figure 3 map of turk participants and country-wise bar chart
+# ----------------------------------------------------------------
+# getting unique participants information
+# Since each participant has multiple responses in demographics data
+turker <- demographics[!duplicated(demographics$id),]
+turker$country <- factor(turker$country, 
+                         levels=names(table(turker$country))[order(table(turker$country))])
+
+map_theme <- list(theme(panel.grid.minor = element_blank(),
+                        panel.grid.major = element_blank(),
+                        panel.background = element_blank(),
+                        axis.line = element_blank(),
+                        axis.text.x = element_blank(),
+                        axis.text.y = element_blank(),
+                        axis.ticks = element_blank(),
+                        axis.title.x = element_blank(),
[…]

### 2017-10-19 submitted-v1/sociology_paper.Rnw
updated paper

--- 2017-10-17 sociology_paper.Rnw
@@
   \hline
-Gender & Male & 1351 & 57.74 & 48.75 & 13542 \\ 
-   & Female & 989 & 42.26 & 43.69 & 10538 \\ 
+Gender & Female & 989 & 42.26 & 43.69 & 10538 \\ 
+   & Male & 1351 & 57.74 & 48.75 & 13542 \\  
 \hline
@@
 \end{table}
+
+% 
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 16:56:32 2017
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{llrrrr}
+%   \hline
+% var & lbls & subsjects & percents & avg\_time & response \\ 
+%   \hline
+% gender\_level & Female & 989 & 42.26 & 43.69 & 10538 \\ 
+%   gender\_level & Male & 1351 & 57.74 & 48.75 & 13542 \\ 
+%   degree & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
+%   degree & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
+%   degree & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
+%   degree & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
+%   degree & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
+%   age\_level & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
+%   age\_level & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
+%   age\_level & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
+%   age\_level & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
+%   age\_level & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
+%   age\_level & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
+%   age\_level & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
+%   country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
+%   country & India & 980 & 41.88 & 52.63 & 10238 \\ 
+%   country & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\ 
+%    \hline
+% \end{tabular}
+% \end{table}
 
@@
 
-% latex table generated in R 3.2.2 by xtable 1.7-4 package
-% Sun Aug 30 22:43:05 2015
+% latex table generated in R 3.4.1 by xtable 1.8-2 package
+% Wed Oct 18 17:43:02 2017
 \begin{table}[ht]
@@
  & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
-Full  & 51872.4 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 23792.0 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
- - Age & 51329.3 & 543.2 & 6 & $<$0.001 & \hspace{0.5cm} & 23766.4 & 26.0 & 6 & $<$0.001 \\ 
- - Country & 51496.4 & 376.1 & 2 & $<$0.001 &  & 23765.9 & 26.0 & 2 & $<$0.001 \\ 
- - Degree & 51756.2 & 116.2 & 4 & $<$0.001 &  & 23768.2 & 23.8 & 4 & $<$0.001 \\ 
- - Gender & 51804.3 & 68.1 & 1 & $<$0.001 &  & 23789.7 & 2.2 & 1 & 0.140 \\ 
+Full  & 52609.1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 24140.5 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
+ - Age & 52081.3 & 527.8 & 6 & $<$0.001 & \hspace{0.5cm} & 24113.2 & 27.3 & 6 & $<$0.001 \\ 
+ - Country & 52245.5 & 363.6 & 2 & $<$0.001 &  & 24114.8 & 25.7 & 2 & $<$0.001 \\ 
+ - Degree & 52498.7 & 110.4 & 4 & $<$0.001 &  & 24116.0 & 24.5 & 4 & $<$0.001 \\ 
+ - Gender & 52540.3 & 68.8 & 1 & $<$0.001 &  & 24138.9 & 1.6 & 1 & 0.200 \\ 
 \cline{2-5} \cline{7-10}
@@
 \end{table}
+
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 17:43:02 2017
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{rrrrrlrrrr}
+%   \hline
+%  & deviance & Chisq & Chi.Df & Pr..Chisq. & g1 & deviance & Chisq & Chi.Df & Pr..Chisq. \\ 
+%   \hline
+% Full & 52609.06 & 0.00 & 6.00 & 0.00 &  & 24140.49 & 0.00 & 6.00 & 0.00 \\ 
+%   Age & 52081.29 & 527.77 & 6.00 & 0.00 &  & 24113.24 & 27.25 & 6.00 & 0.00 \\ 
+%   Country & 52245.46 & 363.60 & 2.00 & 0.00 &  & 24114.79 & 25.70 & 2.00 & 0.00 \\ 
+%   Degree & 52498.65 & 110.41 & 4.00 & 0.00 &  & 24116.02 & 24.47 & 4.00 & 0.00 \\ 
+%   Gender & 52540.30 & 68.76 & 1.00 & 0.00 &  & 24138.85 & 1.64 & 1.00 & 0.20 \\ 
+%    \hline
+% \end{tabular}
+% \end{table}
 
@@
 
-% latex table generated in R 3.2.2 by xtable 1.7-4 package
-% Sat Aug 29 14:17:22 2015
+% latex table generated in R 3.4.1 by xtable 1.8-2 package
+% Wed Oct 18 17:22:24 2017
 \begin{table}[hbtp]
@@
 \scalebox{.9}{
-\begin{tabular}{rrr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
+\begin{tabular}{rrr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
  && \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
@@
 \cline{3-6} \cline{8-11}
-\bf Fixed Effects & Average $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.61 & -0.79 & -0.43 & *** \\ [3pt]
+\bf Fixed Effects & Average $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.63 & -0.81 & -0.45 & *** \\ [3pt]
 \bf Age Category $(\alpha)$ & 18-25 & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
-&  26-30 & 0.06 & 0.03 & 0.08 & *** &   & 0.06 & -0.03 & 0.16 &  \\ 
-&  31-35 & 0.07 & 0.04 & 0.10 & *** &   & 0.11 & 0.01 & 0.22 & * \\ 
-&  36-40 & 0.23 & 0.20 & 0.26 & *** &   & 0.31 & 0.19 & 0.43 & *** \\ 
-&  41-45 & 0.18 & 0.14 & 0.22 & *** &   & 0.16 & 0.00 & 0.31 & * \\ 
-&  46-50 & 0.27 & 0.22 & 0.32 & *** &   & 0.14 & -0.05 & 0.33 &  \\ 
-&  above 50 & 0.35 & 0.32 & 0.39 & *** &   & 0.15 & 0.01 & 0.28 & * \\ [3pt]
+&  26-30 & 0.05 & 0.03 & 0.07 & *** &   & 0.03 & -0.06 & 0.13 &  \\ 
+&  31-35 & 0.06 & 0.03 & 0.09 & *** &   & 0.12 & 0.02 & 0.23 & * \\ 
+&  36-40 & 0.22 & 0.19 & 0.25 & *** &   & 0.29 & 0.17 & 0.41 & *** \\ 
+&  41-45 & 0.17 & 0.13 & 0.21 & *** &   & 0.19 & 0.03 & 0.34 & * \\ 
+&  46-50 & 0.28 & 0.23 & 0.32 & *** &   & 0.20 & 0.01 & 0.39 & * \\ 
+&  above 50 & 0.34 & 0.31 & 0.38 & *** &   & 0.16 & 0.02 & 0.29 & * \\ [3pt]
 \bf Country $(\kappa)$& U.S. & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
-&  India & 0.23 & 0.21 & 0.25 & *** &   & -0.18 & -0.27 & -0.10 & *** \\ 
-&  Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.08 & -0.04 & 0.20 &  \\ [3pt]
+&  India & 0.23 & 0.20 & 0.25 & *** &   & -0.19 & -0.28 & -0.10 & *** \\ 
+&  Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.06 & -0.05 & 0.18 &  \\ [3pt]
 \bf Education $(\tau)$& High School & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
-&  Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & -0.03 & -0.18 & 0.11 &  \\ 
-&  Under grad degree & 0.13 & 0.10 & 0.17 & *** &   & 0.12 & -0.01 & 0.26 & . \\ 
-&  Graduate courses & 0.01 & -0.04 & 0.05 &  &   & -0.17 & -0.33 & -0.01 & * \\ 
-&  Graduate degree & 0.13 & 0.09 & 0.16 & *** &   & -0.06 & -0.19 & 0.07 &  \\ [3pt]
+&  Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & 0.00 & -0.14 & 0.14 &  \\ 
+&  Under grad degree & 0.14 & 0.10 & 0.17 & *** &   & 0.16 & 0.03 & 0.29 & * \\ 
+&  Graduate courses & 0.02 & -0.02 & 0.06 &  &   & -0.13 & -0.29 & 0.03 &  \\ 
+&  Graduate degree & 0.13 & 0.10 & 0.16 & *** &   & -0.04 & -0.17 & 0.09 &  \\ [3pt]
 \bf Gender $(\gamma)$& Female & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
-&  Male & 0.08 & 0.06 & 0.10 & *** &   & 0.06 & -0.02 & 0.13 &  \\
+&  Male & 0.08 & 0.06 & 0.10 & *** &   & 0.05 & -0.02 & 0.12 &  \\
 \cline{3-6} \cline{8-11}
-\bf Random Effects &  lineup $(\sigma_\ell)$& \multicolumn{5}{l}{0.29} & \multicolumn{4}{l}{\phantom{-}2.29} \\ 
+\bf Random Effects &  lineup $(\sigma_\ell)$& \multicolumn{5}{l}{0.29} & \multicolumn{4}{l}{\phantom{-}2.27} \\ 
                    &  Error $(\sigma)$& \multicolumn{5}{l}{0.69} & \multicolumn{4}{l}{\phantom{-}0.89}  \\ 
@@
 \end{table}
+
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 17:22:24 2017
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{rrrrllrrrl}
+%   \hline
+%  & Est & LB & UB & pvalue & g1 & Est & LB & UB & pvalue \\ 
+%   \hline
+% $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.63 & -0.81 & -0.45 & *** \\ 
+%   26-30 & 0.05 & 0.03 & 0.07 & *** &   & 0.03 & -0.06 & 0.13 &  \\ 
+%   31-35 & 0.06 & 0.03 & 0.09 & *** &   & 0.12 & 0.02 & 0.23 & * \\ 
+%   36-40 & 0.22 & 0.19 & 0.25 & *** &   & 0.29 & 0.17 & 0.41 & *** \\ 
+%   41-45 & 0.17 & 0.13 & 0.21 & *** &   & 0.19 & 0.03 & 0.34 & * \\ 
+%   46-50 & 0.28 & 0.23 & 0.32 & *** &   & 0.20 & 0.01 & 0.39 & * \\ 
+%   above 50 & 0.34 & 0.31 & 0.38 & *** &   & 0.16 & 0.02 & 0.29 & * \\ 
+%   India & 0.23 & 0.20 & 0.25 & *** &   & -0.19 & -0.28 & -0.10 & *** \\ 
+%   Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.06 & -0.05 & 0.18 &  \\ 
+%   Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & 0.00 & -0.14 & 0.14 &  \\ 
+%   Under grad degree & 0.14 & 0.10 & 0.17 & *** &   & 0.16 & 0.03 & 0.29 & * \\ 
+%   Graduate courses & 0.02 & -0.02 & 0.06 &  &   & -0.13 & -0.29 & 0.03 &  \\ 
+%   Graduate degree & 0.13 & 0.10 & 0.16 & *** &   & -0.04 & -0.17 & 0.09 &  \\ 
+%   Male & 0.08 & 0.06 & 0.10 & *** &   & 0.05 & -0.02 & 0.12 &  \\ 
+%   lineup & 0.29 &  &  &  &   & 2.27 &  &  &  \\ 
+%    \hline
+% \end{tabular}
+% \end{table}
 
@@
 In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond.
-In Table~\ref{tbl:model_result_demographics} notice that lineup specific error is estimated as 2.29 which is much higher than any of the other parameter estimates in the model.
+In Table~\ref{tbl:model_result_demographics} notice that lineup specific error is estimated as 2.27 which is much higher than any of the other parameter estimates in the model.
 This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
@@
 Even though having a graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.045 which is  negligible from a practical point of view.
-The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.29$)
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.27$)
    of lineup variability.}
@@
     \includegraphics[width=6.5in]{proportion_nulls_guide.pdf} 
-   \caption{Location of data plot in the lineup and proportion correct for both Interaction and Genotype effect.
+   \caption{Location of data plot in the lineup and proportion correct (detection rate) for both Interaction and Genotype effect.
 Each colored line represents a null set and the size of the dots represents number of responses.

### 2017-10-19 sociology_paper.Rnw
results updated

@@
 
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 17:43:02 2017
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{rrrrrlrrrr}
+%   \hline
+%  & deviance & Chisq & Chi.Df & Pr..Chisq. & g1 & deviance & Chisq & Chi.Df & Pr..Chisq. \\ 
+%   \hline
+% Full & 52609.06 & 0.00 & 6.00 & 0.00 &  & 24140.49 & 0.00 & 6.00 & 0.00 \\ 
+%   Age & 52081.29 & 527.77 & 6.00 & 0.00 &  & 24113.24 & 27.25 & 6.00 & 0.00 \\ 
+%   Country & 52245.46 & 363.60 & 2.00 & 0.00 &  & 24114.79 & 25.70 & 2.00 & 0.00 \\ 
+%   Degree & 52498.65 & 110.41 & 4.00 & 0.00 &  & 24116.02 & 24.47 & 4.00 & 0.00 \\ 
+%   Gender & 52540.30 & 68.76 & 1.00 & 0.00 &  & 24138.85 & 1.64 & 1.00 & 0.20 \\ 
+%    \hline
+% \end{tabular}
+% \end{table}
+
 Parameter estimates from the model fits are shown in Table~\ref{tbl:model_result_demographics}.
 The first level of each factor serves as the baseline for the model.
-The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for age groups 31-35 and above 50.
-Participants from India have a similar detection rate to those from the USA, but participants from the rest of the world show a significantly higher detection rate.
-Subjects  reporting to have  a graduate degree had a significantly higher detection rate.
-For time to respond, all age groups were significantly slower than  18-25 year olds.
-Participants from India were slower, while those from the rest of the world were faster than US Americans.
-A higher degree corresponded to longer time, and males were on average slower than females.
-
-% latex table generated in R 3.2.2 by xtable 1.7-4 package
-% Sat Aug 29 14:17:22 2015
+The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for other age groups except for 26-30.
+Participants from the rest of the world have a similar detection rate to those from the USA, but participants from India show a significantly lower detection rate.
+Subjects reporting to have  an undergraduate degree had a significantly higher detection rate.
+For time to respond, all age groups were significantly faster than  18-25 year olds.
+Participants from India as well as the rest of the world were faster than US Americans.
+A higher degree corresponded to longer time, and males were on average faster than females.
+
+% latex table generated in R 3.4.1 by xtable 1.8-2 package
+% Wed Oct 18 17:22:24 2017
 \begin{table}[hbtp]
@@
 While some of the demographic factors are strongly statistically significant, the main source of variation in detection rate is the lineup difficulty.
-For example, let's examine the effect of graduate degree.
+For example, let's examine the effect of an undergraduate degree.
 To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with some graduate course work as compared to an undergraduate degree, for an average difficulty lineup (random effect of zero).
 Plugging in the relevant quantities to the fitted model gives a difference equal to:
-$$ \frac {\exp(-0.61-0.17)}{1+\exp(-0.61-0.17)}-  \frac {\exp(-0.61+0.12)}{1+\exp(-0.61+0.12)} = 0.314- 0.379 =-0.065.$$ 
-The person with some graduate course work picks the data plot  on average in 33.5\% of lineups of average difficulty, as compared to 37.9\% if they have an undergraduate degree.
-This difference is reduced to 2\% for a lineup with one standard deviation order of magnitude difference in difficulty.
-For two standard deviations it further reduces to 0.3\%.
+$$ \frac {\exp(-0.63-0.13)}{1+\exp(-0.63-0.13)}-  \frac {\exp(-0.63+0.16)}{1+\exp(-0.63+0.16)} = 0.319- 0.385 =-0.066.$$ 
+The person with some graduate course work picks the data plot  on average in 31.9\% of lineups of average difficulty, as compared to 38.5\% if they have an undergraduate degree.
+This difference is reduced to 4\% for a lineup with one standard deviation order of magnitude difference in difficulty.
+For two standard deviations it further reduces to 0.6\%.
 Thus, although there is a statistically significant difference in the proportion at which participants identify the  data plot for some demographic factors, these are not  significant in any practical sense.
-Figure~\ref{fig:practical_impact_graduate} illustrates this example showing fitted models for a US 18-25 female with either a high school education or a graduate degree.
-Similar calculations show the same negligible impact of age level 36-40 (\Sexpr{round(exp(-0.61+0.31)/(1+ exp(-0.61+0.31)) - exp(-0.61)/(1 + exp(-0.61)),3)} at most)  and country (\Sexpr{round(exp(-0.61+0.08)/(1+ exp(-0.61+0.08)) - exp(-0.61-0.18)/(1 + exp(-0.61-0.18)), 3)} at most) on the probability of a correct response.
+Figure~\ref{practical_impact_und_und_graduate.pdf} illustrates this example showing fitted models for a US 18-25 female with either a high school education or an undergraduate degree.
+Similar calculations show the same negligible impact of age level 36-40 (\Sexpr{round(exp(-0.63+0.29)/(1+ exp(-0.63+0.29)) - exp(-0.63)/(1 + exp(-0.63)),3)} at most)  and country India (\Sexpr{round(exp(-0.63-0.19)/(1+ exp(-0.63-0.19)) - exp(-0.63)/(1 + exp(-0.63)), 3)} at most) on the probability of a correct response.
 Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.
@@
 %   \includegraphics[width=6.5in]{learning_trend.pdf} 
-    \includegraphics[width=3.2in]{practical_impact_graduate.pdf} 
+    \includegraphics[width=3.2in]{practical_impact_und_graduate.pdf} 
     \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
-   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with a graduate degree compared to a high school degree.
-Even though having a graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.045 which is  negligible from a practical point of view.
-The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.29$)
+   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with an under graduate degree compared to a high school degree.
+Even though having an under graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.04 which is  negligible from a practical point of view.
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} = 2.27$)
    of lineup variability.}
-   \label{fig:practical_impact_graduate}
+   \label{practical_impact_und_und_graduate.pdf}
 \end{figure}
@@
 
-% latex table generated in R 3.2.2 by xtable 1.7-4 package
-% Sun Aug 30 11:15:41 2015
+% latex table generated in R 3.4.1 by xtable 1.8-2 package
+% Wed Oct 18 23:54:25 2017
 \begin{table}[hbtp]
@@
 \scalebox{.9}{
-\begin{tabular}{rr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
+\begin{tabular}{rr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
 \cline{2-5} \cline{7-10} \cline{12-15} 
@@
 Fixed \\
-$\mu$ & -1.30 & -1.69 & -0.92 & *** &   & -0.22 & -0.51 & 0.07 &  &   & -1.71 & -2.77 & -0.65 & ** \\ [3pt]
-  $\alpha_2$ & 0.27 & -0.19 & 0.73 &  &   & 0.26 & -0.06 & 0.58 &  &   & -0.49 & -1.31 & 0.32 &  \\ 
-  $\alpha_3$ & -0.18 & -0.65 & 0.29 &  &   & 0.34 & 0.02 & 0.66 & * &   & -0.19 & -1.01 & 0.63 &  \\ 
-  $\alpha_4$ & 0.08 & -0.39 & 0.55 &  &   & 0.36 & 0.04 & 0.68 & * &   & -0.47 & -1.28 & 0.35 &  \\ 
-  $\alpha_5$ & 0.30 & -0.17 & 0.77 &  &   & 0.37 & 0.05 & 0.70 & * &   & -0.16 & -0.96 & 0.65 &  \\ 
-  $\alpha_6$ & 0.04 & -0.44 & 0.53 &  &   & 0.24 & -0.07 & 0.56 &  &   & -0.03 & -0.87 & 0.82 &  \\ 
-  $\alpha_7$ & 0.28 & -0.20 & 0.77 &  &   & 0.16 & -0.16 & 0.48 &  &   & 0.02 & -0.81 & 0.84 &  \\ 
-  $\alpha_8$ & -0.04 & -0.54 & 0.45 &  &   & 0.34 & 0.02 & 0.66 & * &   & -0.08 & -0.89 & 0.73 &  \\ 
-  $\alpha_9$ & -0.20 & -0.69 & 0.30 &  &   & 0.38 & 0.05 & 0.70 & * &   & 0.15 & -0.71 & 1.02 &  \\ 
-  $\alpha_{10}$ & 0.51 & 0.03 & 1.00 & * &   & 0.19 & -0.14 & 0.52 &  &   & -0.28 & -1.14 & 0.58 &  \\ 
+$\mu$ & -1.32 & -1.73 & -0.92 & *** &   & -0.16 & -0.46 & 0.13 &  &   & -1.73 & -2.81 & -0.66 & ** \\ [3pt]
+  $\alpha2$ & 0.30 & -0.17 & 0.78 &  &   & 0.26 & -0.07 & 0.58 &  &   & -0.40 & -1.23 & 0.44 &  \\ 
+  $\alpha3$ & -0.20 & -0.68 & 0.29 &  &   & 0.31 & -0.02 & 0.63 & . &   & -0.15 & -0.99 & 0.68 &  \\ 
+  $\alpha4$ & 0.13 & -0.35 & 0.61 &  &   & 0.34 & 0.01 & 0.66 & * &   & -0.41 & -1.24 & 0.42 &  \\ 
+  $\alpha5$ & 0.35 & -0.14 & 0.83 &  &   & 0.34 & 0.02 & 0.67 & * &   & -0.10 & -0.92 & 0.73 &  \\ 
+  $\alpha6$ & 0.10 & -0.40 & 0.60 &  &   & 0.20 & -0.12 & 0.53 &  &   & 0.03 & -0.84 & 0.89 &  \\ 
+  $\alpha7$ & 0.33 & -0.17 & 0.82 &  &   & 0.11 & -0.22 & 0.43 &  &   & -0.00 & -0.84 & 0.84 &  \\ 
+  $\alpha8$ & -0.01 & -0.51 & 0.50 &  &   & 0.31 & -0.01 & 0.64 & . &   & -0.06 & -0.88 & 0.76 &  \\ 
+  $\alpha9$ & -0.20 & -0.70 & 0.30 &  &   & 0.34 & 0.01 & 0.67 & * &   & 0.21 & -0.67 & 1.08 &  \\ 
+  $\alpha10$ & 0.51 & 0.02 & 1.01 & * &   & 0.19 & -0.14 & 0.53 &  &   & -0.20 & -1.07 & 0.67 &  \\ 
   Random \\
-  $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.85}  & \multicolumn{5}{l}{\phantom{-}0.90} & 0.76   \\ 
-  $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.02}  & \multicolumn{5}{l}{\phantom{-}0.03} & 0.06   \\ 
-  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}1.48}  & \multicolumn{5}{l}{\phantom{-}1.42} & 3.37   \\ \cline{2-5} \cline{7-10} \cline{12-15} 
+  $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.91}  & \multicolumn{5}{l}{\phantom{-}0.89} & 0.82   \\ 
+  $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.04}  & \multicolumn{5}{l}{\phantom{-}0.04} & 0.05   \\ 
+  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}1.47}  & \multicolumn{5}{l}{\phantom{-}1.43} & 3.38   \\ \cline{2-5} \cline{7-10} \cline{12-15} 
 \\
@@
 \end{table}
+
+% 
+% 
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 23:54:25 2017
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{rrrrllrrrllrrrl}
+%   \hline
+%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
+%   \hline
+% $\mu$ & -1.32 & -1.73 & -0.92 & *** &   & -0.16 & -0.46 & 0.13 &  &   & -1.73 & -2.81 & -0.66 & ** \\ 
+%   $\alpha2$ & 0.30 & -0.17 & 0.78 &  &   & 0.26 & -0.07 & 0.58 &  &   & -0.40 & -1.23 & 0.44 &  \\ 
+%   $\alpha3$ & -0.20 & -0.68 & 0.29 &  &   & 0.31 & -0.02 & 0.63 & . &   & -0.15 & -0.99 & 0.68 &  \\ 
+%   $\alpha4$ & 0.13 & -0.35 & 0.61 &  &   & 0.34 & 0.01 & 0.66 & * &   & -0.41 & -1.24 & 0.42 &  \\ 
+%   $\alpha5$ & 0.35 & -0.14 & 0.83 &  &   & 0.34 & 0.02 & 0.67 & * &   & -0.10 & -0.92 & 0.73 &  \\ 
+%   $\alpha6$ & 0.10 & -0.40 & 0.60 &  &   & 0.20 & -0.12 & 0.53 &  &   & 0.03 & -0.84 & 0.89 &  \\ 
+%   $\alpha7$ & 0.33 & -0.17 & 0.82 &  &   & 0.11 & -0.22 & 0.43 &  &   & -0.00 & -0.84 & 0.84 &  \\ 
+%   $\alpha8$ & -0.01 & -0.51 & 0.50 &  &   & 0.31 & -0.01 & 0.64 & . &   & -0.06 & -0.88 & 0.76 &  \\ 
+%   $\alpha9$ & -0.20 & -0.70 & 0.30 &  &   & 0.34 & 0.01 & 0.67 & * &   & 0.21 & -0.67 & 1.08 &  \\ 
+%   $\alpha10$ & 0.51 & 0.02 & 1.01 & * &   & 0.19 & -0.14 & 0.53 &  &   & -0.20 & -1.07 & 0.67 &  \\ 
+%   $\sigma^2_u$ & 0.91 &  &  &  &   & 0.89 &  &  &  &   & 0.82 &  &  &  \\ 
+%   $\sigma^2_a$ & 0.04 &  &  &  &   & 0.04 &  &  &  &   & 0.05 &  &  &  \\ 
+%   $\sigma^2_l$ & 1.47 &  &  &  &   & 1.43 &  &  &  &   & 3.38 &  &  &  \\ 
+%    \hline
+% \end{tabular}
+% \end{table}
+% 
+%
 
@@
 
-% latex table generated in R 3.2.2 by xtable 1.7-4 package
-% Sun Aug 30 11:42:08 2015
+% latex table generated in R 3.4.1 by xtable 1.8-2 package
+% Wed Oct 18 23:37:55 2017
 \begin{table}[bhtp]
@@
 \scalebox{0.9}{
-\begin{tabular}{lr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
+\begin{tabular}{lr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
 \cline{2-5} \cline{7-10} \cline{12-15} 
@@
 \multicolumn{2}{l}{Fixed}\\
-$\mu$ & 3.82 & 3.74 & 3.89 & *** &   & 3.90 & 3.84 & 3.97 & *** &   & 3.73 & 3.62 & 3.84 & *** \\ 
-  $\alpha_1$ & 0.33 & 0.26 & 0.39 & *** &   & 0.34 & 0.28 & 0.39 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
+$\mu$ & 3.80 & 3.72 & 3.88 & *** &   & 3.89 & 3.83 & 3.96 & *** &   & 3.75 & 3.64 & 3.85 & *** \\ 
+  $\alpha_1$ & 0.32 & 0.25 & 0.39 & *** &   & 0.34 & 0.28 & 0.40 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
   $\alpha$ & -0.04 & -0.05 & -0.03 & *** &   & -0.04 & -0.05 & -0.03 & *** &   & -0.03 & -0.04 & -0.02 & *** \\ 
 \multicolumn{2}{l}{Random}\\
-  $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.51}   & \multicolumn{5}{l}{\phantom{-}0.49}  & 0.37   \\ 
-  $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.03}  & \multicolumn{5}{l}{\phantom{-}0.04}   & 0.05   \\ 
-  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}0.09}   & \multicolumn{5}{l}{\phantom{-}0.20}   & 0.23   \\ 
+  $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.52}   & \multicolumn{5}{l}{\phantom{-}0.49}  & 0.37   \\ 
+  $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.03}  & \multicolumn{5}{l}{\phantom{-}0.05}   & 0.05   \\ 
+  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}0.09}   & \multicolumn{5}{l}{\phantom{-}0.20}   & 0.24   \\
   $\sigma^2$ & \multicolumn{5}{l}{\phantom{-}0.46}   & \multicolumn{5}{l}{\phantom{-}0.50}  & 0.45  \\ 
@@
 \end{table}
+
+% 
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 23:37:55 2017
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{rrrrllrrrllrrrl}
+%   \hline
+%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
+%   \hline
+% $\mu$ & 3.80 & 3.72 & 3.88 & *** &   & 3.89 & 3.83 & 3.96 & *** &   & 3.75 & 3.64 & 3.85 & *** \\ 
+%   $\alpha_1$ & 0.32 & 0.25 & 0.39 & *** &   & 0.34 & 0.28 & 0.40 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
+%   $\alpha$ & -0.04 & -0.05 & -0.03 & *** &   & -0.04 & -0.05 & -0.03 & *** &   & -0.03 & -0.04 & -0.02 & *** \\ 
+%   $\sigma^2_u$ & 0.52 &  &  &  &   & 0.49 &  &  &  &   & 0.37 &  &  &  \\ 
+%   $\sigma^2_a$ & 0.03 &  &  &  &   & 0.05 &  &  &  &   & 0.05 &  &  &  \\ 
+%   $\sigma^2_l$ & 0.09 &  &  &  &   & 0.20 &  &  &  &   & 0.24 &  &  &  \\ 
+%   $\sigma^2$ & 0.46 &  &  &  &   & 0.50 &  &  &  &   & 0.45 &  &  &  \\ 
+%    \hline
+% \end{tabular}
+% \end{table}
+
 %The lineup variance of experiment~7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model~\eqref{eqn:trend_response}.

### 2017-10-19 R/sociology_paper.R

@@ -159,8 +159,10 @@ sociology_data <- demographics %>%
 saveRDS(sociology_data, file = "../submitted-v1/data/sociology-data.RDS")
 
 exp9 <- dat9 %>% 
-  mutate(lineup_name = pic_name) %>%
-  select(plot_type, plot_location, nulls, lineup_name, response)
+  mutate(lineup_name = pic_name,
+         id = paste0("exp9_",id),
+         lineup_type = plot_type) %>%
+  select(id, lineup_type, lineup_name, plot_location, nulls, response)
 
 saveRDS(exp9, file = "../submitted-v1/data/exp9.RDS")
 

### 2017-10-19 submitted-v1/R/sociology-paper-v1.R

@@ -252,84 +252,46 @@ ggsave("../images/demographic_effect.pdf", width=6.5, height=6)
 # ==================================================================
 # figure 5 practical significance of demographics
 # ------------------------------------------------------------------
-
-et <- 6
-xb <- seq(-et,et, by=.1)
+# function to compute estimated proportion
 px <- function(xb){exp(xb)/(1+exp(xb))}
-pdat <- data.frame(xb,px=px(xb))
 
-s1 <- 2.293
-s2 <- s1*2
-xx <- c(0,0.182)
-
-ldat <- data.frame(x =c(xx,-et,-et,xx+s1,-et,-et,xx+s2,-et,-et),
-                   xend = c(xx,xx,xx+s1,xx+s1,xx+s2,xx+s2 ),
-                   y = c(0,0,px(xx),0,0,px(xx+s1),0,0,px(xx+s2)),
-                   yend= px(c(xx,xx,xx+s1,xx+s1,xx+s2,xx+s2 )))
-
-ggplot()+geom_line(aes(xb,px), data=pdat)+
-  geom_segment(aes(x=x,xend=xend,y=y,yend=yend), data=ldat, linetype="dashed") + 
-  theme_bw() + xlab(expression(eta)) + ylab("Prability of detection") +
-  theme(panel.grid = element_blank())+
-  scale_y_continuous(expand = c(0.01,0)) +
-  scale_x_continuous(expand = c(0.01,0))
-
-
-get_change <- function(s){
-  xb0 <- -0.683 + s
-  xb1 <- xb0 + 0.182 
-  px(xb1) - px(xb0)
-}
-
-get_change(c(0,s1,s2))
-get_change(2.293)
-
-xx <- (1:2)*s1
-ldat <- data.frame(x=c(xx,0,0), xend=c(xx,xx),y=c(0,0,get_change(xx)), yend=get_change(c(xx,xx)))
-
-ss <- seq(0,6.5, by=.1)
-qplot(ss,get_change(ss), geom="line") + 
-  geom_segment(aes(x=x,xend=xend,y=y,yend=yend), data=ldat, linetype="dashed") +
-  ylab("Change in probability of detection \n due to graduate degree") +
-  xlab("Lineup variability") +
-  scale_y_continuous(breaks =c(0,.003237246,.01,.02,.02376092,.03,.04), expand = c(0.02,0),
-                     labels =c(0,0.003,.01,.02,.024,.03,.04)) +
[…]

### 2017-10-19 R/sociology_paper.R
changed data name

@@ -156,7 +156,7 @@ levels(demographics$age_level)[1:7] <- c(levels(demographics$age_level)[1:6], "a
 sociology_data <- demographics %>%
   select(id, response, start_time, time_taken, lineup_id, gender_level, age_level, 
          degree, country_code, country_name, country, longitude, latitude, experiment)
-saveRDS(sociology_data, file = "../submitted-v1/data/sociology-data.RDS")
+saveRDS(sociology_data, file = "../submitted-v1/data/experiment-data-all.RDS")
 
 exp9 <- dat9 %>% 
   mutate(lineup_name = pic_name,
@@ -164,7 +164,7 @@ exp9 <- dat9 %>%
          lineup_type = plot_type) %>%
   select(id, lineup_type, lineup_name, plot_location, nulls, response)
 
-saveRDS(exp9, file = "../submitted-v1/data/exp9.RDS")
+saveRDS(exp9, file = "../submitted-v1/data/exp9-extra.RDS")
 
 # getting unique participants information
 # Since each participant has multiple responses in demographics data

### 2017-10-19 submitted-v1/R/sociology-paper-v1.R

@@ -14,8 +14,8 @@ library(maps)
 # Loading the data 
 # ------------------------------------------------------------------
 
-demographics <- readRDS("../data/sociology-data.rds")
-dat9 <- readRDS("../data/exp9.RDS")
+demographics <- readRDS("../data/experiment-data-all.rds")
+dat9 <- readRDS("../data/exp9-extra.RDS")
 
 
 # ==================================================================

### 2017-10-19 submitted-v1/R/sociology-paper-v1.R
added data

@@ -14,7 +14,7 @@ library(maps)
 # Loading the data 
 # ------------------------------------------------------------------
 
-demographics <- readRDS("../data/experiment-data-all.rds")
+expDat <- readRDS("../data/experiment-data-all.rds")
 dat9 <- readRDS("../data/exp9-extra.RDS")
 
 
@@ -23,14 +23,14 @@ dat9 <- readRDS("../data/exp9-extra.RDS")
 # ------------------------------------------------------------------
 
 # showing number of countries the subjects are coming from
-length(unique(demographics$country_code))
+length(unique(expDat$country_code))
 # showing countries with more than 10 participants
-countryCount <- demographics %>% 
+countryCount <- expDat %>% 
   group_by(country_name) %>%
   summarise(counts = length(unique(id))) %>%
   arrange(desc(counts))
 # showing number of linesups evaluated
-length(unique(demographics$lineup_id))
+length(unique(expDat$lineup_id))
 
 
 
@@ -39,7 +39,7 @@ length(unique(demographics$lineup_id))
 # ----------------------------------------------------------------
 # getting unique participants information
 # Since each participant has multiple responses in demographics data
-turker <- demographics[!duplicated(demographics$id),]
+turker <- expDat[!duplicated(expDat$id),]
 turker$country <- factor(turker$country, 
                          levels=names(table(turker$country))[order(table(turker$country))])
 
@@ -94,10 +94,10 @@ get_summary <- function(dat, var){
   return(data.frame(var=var,lbls=res[,1], res[,-1]))
 }
 
-sg <- get_summary(demographics, "gender_level")
-se <- get_summary(demographics, "degree")
-sa <- get_summary(demographics, "age_level")
-sl <- get_summary(demographics, "country")
+sg <- get_summary(expDat, "gender_level")
+se <- get_summary(expDat, "degree")
+sa <- get_summary(expDat, "age_level")
+sl <- get_summary(expDat, "country")
 
 sdat <- rbind(sg,se,sa,sl)
 print(xtable(sdat[,-7]), include.rownames=FALSE)
@@ -108,11 +108,11 @@ print(xtable(sdat[,-7]), include.rownames=FALSE)
 # ----------------------------------------------------------------
 
[…]

### 2017-10-21 R/sociology_paper.R
uploaded data as .RData

@@ -96,7 +96,7 @@ dat9$nulls <- paste("null_",sapply(nulls, function(x) return(x[2])), sep="")
 # responses on the plot used for defense talk
 subset(dat9, pic_name=="plot_turk9_interaction_1_3.svg")
 
-# responses on the plot used for defense nsf grant
+# responses on the plot used for nsf grant
 subset(dat9, pic_name=="plot_turk9_geno_1_2.svg")
 subset(dat9, pic_name=="plot_turk9_interaction_2_1.svg")
 
@@ -153,18 +153,20 @@ levels(demographics$age_level)[1:7] <- c(levels(demographics$age_level)[1:6], "a
 # saving data for submission
 # ------------------------------------------------------------------
 # write.csv(demographics, "../submitted-v1/data/sociology-data.csv", row.names = FALSE)
-sociology_data <- demographics %>%
+allDat <- demographics %>%
   select(id, response, start_time, time_taken, lineup_id, gender_level, age_level, 
          degree, country_code, country_name, country, longitude, latitude, experiment)
-saveRDS(sociology_data, file = "../submitted-v1/data/experiment-data-all.RDS")
+# saveRDS(allDat, file = "../submitted-v1/data/experiment-data-all.RDS")
 
-exp9 <- dat9 %>% 
+exp9Dat <- dat9 %>% 
   mutate(lineup_name = pic_name,
          id = paste0("exp9_",id),
          lineup_type = plot_type) %>%
   select(id, lineup_type, lineup_name, plot_location, nulls, response)
+# saveRDS(exp9Dat, file = "../submitted-v1/data/exp9-extra.RDS")
 
-saveRDS(exp9, file = "../submitted-v1/data/exp9-extra.RDS")
+save(allDat, exp9Dat, file= "../submitted-v1/data/experiment-data-all.RData")
+
 
 # getting unique participants information
 # Since each participant has multiple responses in demographics data

### 2017-10-21 submitted-v1/R/sociology-paper-v1.R

@@ -1,4 +1,9 @@
-# The complete R code for Sociology Paper
+# The complete R code for Sociology Paper submitted to JASA
+# Last updated by Mahbubul Majumder Oct 20, 2017
+
+# ==================================================================
+# Loading necessary libraries 
+# ------------------------------------------------------------------
 
 library(ggplot2)
 library(plyr)
@@ -13,9 +18,9 @@ library(maps)
 # ==================================================================
 # Loading the data 
 # ------------------------------------------------------------------
-
-expDat <- readRDS("../data/experiment-data-all.rds")
-dat9 <- readRDS("../data/exp9-extra.RDS")
+load("../data/experiment-data-all.RData")
+expDat <- allDat
+dat9 <- exp9Dat
 
 
 # ==================================================================
@@ -33,12 +38,11 @@ countryCount <- expDat %>%
 length(unique(expDat$lineup_id))
 
 
-
 # ==================================================================
 # figure 3 map of turk participants and country-wise bar chart
 # ----------------------------------------------------------------
 # getting unique participants information
-# Since each participant has multiple responses in demographics data
+# Since each participant has multiple responses in the actual data
 turker <- expDat[!duplicated(expDat$id),]
 turker$country <- factor(turker$country, 
                          levels=names(table(turker$country))[order(table(turker$country))])
@@ -116,7 +120,7 @@ fp <- glmer(response~age_level+country+degree+gender_level+(1|lineup_id),
 
 summary(fp)
 resid <- residuals(fp)
-# residual error
+# residual error estimate (sigma)
 sqrt(sum(resid^2)/(length(resid)-length(fixef(fp))))
 
 # function to get estimates with confidence interval
@@ -163,6 +167,7 @@ rownames(est.factor) <- c("$\\mu$",substr(rownames(est.factor)[2:7],10,nchar(row
 dgt <- c(rep(0,15), rep(2,45), rep(0,15))
 dgts <- matrix(rep(dgt,2), ncol=10)
 print(xtable(est.factor, digits=dgts),  sanitize.text.function = function(x){x})
+
 
 # ==================================================================
 # table 3 ANOVA Model results of demographic factor main effects
[…]

### 2017-10-21 submitted-v1/sociology_paper.Rnw
updated author form

@@
-%\documentclass{article}
 \documentclass[10pt]{article}
@@
 the effects of different covariates on an individual's ability to evaluate lineups.
-The design of some of the experimets allow a further investigation into aspects beyond participants' demographics: experiments~5, 6 and 7 can be used to investigate short term learning trends, while experiment~9 allows insight on whether the  positioning of the  data plot in the lineup has an effect on detection rate.
+The design of some of the experiments allow a further investigation into aspects beyond participants' demographics: experiments~5, 6 and 7 can be used to investigate short term learning trends, while experiment~9 allows insight on whether the  positioning of the  data plot in the lineup has an effect on detection rate.
  %Each of these experiments collected demographic details of the subjects. 
@@
 
-A total of 1912 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where each participant evaulated three lineups.
+A total of 1912 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where each participant evaluated three lineups.
 As a measure to ensure a high quality in the evaluations,
@@
 The first level of each factor serves as the baseline for the model.
-The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for age groups 31-35 and above 50.
-Participants from India have a similar detection rate to those from the USA, but participants from the rest of the world show a significantly higher detection rate.
-Subjects  reporting to have  a graduate degree had a significantly higher detection rate.
-For time to respond, all age groups were significantly slower than  18-25 year olds.
-Participants from India were slower, while those from the rest of the world were faster than US Americans.
-A higher degree corresponded to longer time, and males were on average slower than females.
+The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for other age groups except for 26-30.
+Participants from the rest of the world have a similar detection rate to those from the USA, but participants from India show a significantly lower detection rate.
+Subjects reporting to have  an undergraduate degree had a significantly higher detection rate.
+For time to respond, all age groups were significantly faster than  18-25 year olds.
+Participants from India as well as the rest of the world were faster than US Americans.
+A higher degree corresponded to longer time, and males were on average faster than females.
 
@@
 %   \includegraphics[width=6.5in]{learning_trend.pdf} 
-    \includegraphics[width=3.2in]{practical_impact_graduate.pdf} 
+    \includegraphics[width=3.2in]{practical_impact_und_graduate.pdf} 
     \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
-   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with a graduate degree compared to a high school degree.
-Even though having a graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.045 which is  negligible from a practical point of view.
-The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.27$)
+   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with an under graduate degree compared to a high school degree.
+Even though having an under graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.04 which is  negligible from a practical point of view.
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} = 2.27$)
    of lineup variability.}
-   \label{fig:practical_impact_graduate}
+   \label{practical_impact_und_und_graduate.pdf}
 \end{figure}
@@
 In our study of the demographic effects on performance across a set of different experiments we found some statistically significant effects.
-Age group 36-40, countries other than India and United states, people who have a graduate degree have a significantly higher detection rate.
+Age group 36-40, people who have a graduate degree have a significantly higher detection rate.
+India shows significantly lower detection rate compared to United States.
 Gender does not have any significant effect.
@@
 \end{document}
-
-\dc{*** Suggest all of this is supplementary}
-
-\section{Some Figures and Tables (May go to Appendix)}
-
-\subsection{How People Pick the Data Plot} In each of the experiments observers were asked to choose the reason for their selection of a particular data plot.
-\cite{majumder:2013} showed that these choice reasons may provide the clue on how people picks the data plot in the lineup.
-To investigate this further we added free text input option in experiment~9 for their choice reason instead of some fixed reasons to select from.
-This allowed observers to write whatever they think their reasons for choice are.
-Figure~\ref{fig:wordle} shows the words used to explain their reasons for choice.
-The most common words used to explain their choice are points and green which indicates the use of two important features of the grammar of graphics \citep{wilkinson:1999, hadley:2009}.
-One is the indicator of geometric shape and the other is the aesthetics.
-Spread, steepest, line and apart are some other important words used frequently.
-Spread and apart are indicative of variability in the data.
-Steepest and line indicates some sort of systematic pattern in the data.
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=3.5in]{reasoning_words.pdf} 
-   \caption{Words used to explain the reasons for selection of data plot in a lineup show what features of a lineup may help a non-statistician to evaluate it.
-Larger font indicate more people choosing that word.
-Different color is used just to separate the words.}
-   \label{fig:wordle}
-\end{figure}
-
-Figure~\ref {fig:wordle} also shows some insight about peoples way of reading a plot.
-We notice that variability in the data, geometric shapes and aesthetics used to generate plots and existence of any systematic pattern in the plot are some of the important features revealed from the figure.
-These features are commonly used by human brain to examine and compare plots.
-Notice that these features may be specific to this particular experiment.
-There can be different other features people may use to evaluate a lineup depending on the situation.
-But with these words we get a general idea on how people may think while evaluating a lineup.
-
-Turk workers are not necessarily trained on statistics or aware of specific terms used in statistics or statistical plots let alone having knowledge about grammar of graphics.
-It is interesting that people explain things that have specific definitions and meaning in the literature.
-For example, some keywords like spread, apart should be analogous to larger variability while together,  close may be for indicating smaller variability.
-Thus the perception from a statistical graphics is intuitive and human intelligence learn this even without having specific training.
-This is why visual inference can be used as a tool for teaching inference in the basic statistics classes.
-
-%\begin{table}[hbtp]
-%\caption{Amazon mechanical turk experiments and their properties. Duration in hours per 100 tasks show the popularity of some tasks compared to others.}
-%\centering
-%\begin{tabular}{rlrrrrrrr}
-%  \hline
-%& Experiment& \multicolumn{2}{c}{ Total Task}& Average & \multicolumn{2}{c}{Duration (hour)} & Payment & Pay rate\\
-%\cline{3-4} \cline{6-7}
-%Serial & description & submitted & rejected & time(min) & Actual & 100 task& \$/task & \$/hour\\ 
-%  \hline
-%1 & Boxplot & 406 & 106 & 10.68 & 146.48 & 36.08 & 0.50 & 2.81 \\ 
-%  2 & Scatterplot & 359 &   9 & 10.80 & 42.68 & 11.89 & 1.00 & 5.58 \\ 
-%  3 & Contaminated plot & 219 &  19 & 13.53 & 126.17 & 57.61 & 1.00 & 2.22 \\ 
-%  4 & Polar vs Cartesian & 110 &  10 & 20.65 & 11.65 & 10.59 & 1.00 & 2.91 \\ 
-%  5 & Hist vs density & 234 &  37 & 17.85 & 41.57 & 17.76 & 1.00 & 3.36 \\ 
-%  6 & Violin vs boxplot & 417 &  17 & 17.95 & 105.87 & 25.39 & 1.00 & 3.34 \\ 
-%  7 & Group separation & 106 &   6 & 16.13 & 5.15 & 4.86 & 1.00 & 3.72 \\ 
-%  8 & Sine Illusion & 101 &   1 & 16.52 & 78.38 & 77.60 & 1.00 & 3.63 \\ 
-%  9 & Gene expression & 103 &   3 & 12.47 & 11.27 & 10.94 & 0.50 & 2.41 \\ 
-%  10 & Test normality & 406 &   6 & 22.70 & 74.35 & 18.31 & 1.00 & 2.64 \\ 
-%   \hline
-%\end{tabular}
-%\label{tbl:mturk}
-%\end{table}
-%
-%
-%\begin{figure}[htbp] 
-%   \centering
-%   \includegraphics[width=3in]{rejected_task.pdf}
-%      \includegraphics[width=3in]{task_duration.pdf} 
-%   \caption{Percentage of rejected tasks and duration of each experiment in hour per 100 tasks for each of the 10 experiments.
-Most of the tasks got rejected for box plot experiment.
-Even though the sine illusion experiment took longest to finish the rejection rate is lowest for this experiment.}
-%   \label{fig:task_duration}
-%\end{figure}
-
-%\clearpage
-\section{Some Plots of Exploratory Data Analysis}
-
-This section includes some plots showing results from exploratory data analysis. The results are discussed in Section~\ref{sec:result_socio}.
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=5.8in]{turker_location_experiment.pdf} 
-   \caption{World maps showing where the participants are coming from for all the 10 experiments.}
-   \label{fig:turker_location_experiment}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=6.5in]{participation_time.pdf} 
-   \caption{Number of participants by time of the day feedbacks received (central time).
-Experiment~1 shows MTurk workers participated the experiments around the clock.
-Other experiments did not take a whole day to finish.
-For experiment~3 most of the participants are from India because of timing.
-No matter when the experiment is started, subjects from India shows participations.
-For United States, subjects participated if the experiment is not in the mid night, except for experiment~6.}
-   \label{fig:participation_time}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=6in]{demographic_info.pdf} 
-   \caption{Countrywise distribution of age and academic levels of the MTurk workers participating the experiments shows the diversity of the subjects in all the demographic aspect.
-Almost equal number of male and female subjects participated the online experiments.}
-   \label{fig:demographic_info}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=6.5in]{age_gender_within_country_bar.pdf} 
-   \caption{Countrywise distribution of age and academic levels of the MTurk workers participating the experiments shows the diversity of the subjects in all the demographic aspect.
-Male and female participants differ in India specially for agelevel 18-25.
-For United States number of participants are similar beyond age 40 while few number of participants coming from India beyond that age.}
-   \label{fig:gender_country_bar}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=6.5in]{age_gender_within_country_time.pdf} 
-   \caption{Countrywise average time taken for different age and academic levels of the MTurk workers participating the experiments shows that the demographic factors may not have effect on time taken.
-}
-   \label{fig:gender_country_time}
-\end{figure}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=6.5in]{age_gender_within_country_correct.pdf} 
-   \caption{Countrywise percentage of correct responses for different age and academic levels of the MTurk workers participating the experiments shows that the demographic factors may not have effect on the percentage of correct responses.
-}
-   \label{fig:gender_country_response}
-\end{figure}
-
-%\begin{figure}[htbp] 
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{electoral-2-17.pdf} 
-%   \caption{Lineup \#2 - electoral building. Which building looks the most different from the other buildings? }
-%   \label{fig:elect.2}
-%\end{figure}
-%\begin{figure}[htbp] 
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{electoral-3-18.pdf} 
-%   \caption{Lineup \#3 - electoral building. Which building looks the most different from the other buildings? }
-%   \label{fig:elect.3}
-%\end{figure}
-%\begin{figure}[htbp] 
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{electoral-4-17.pdf} 
-%   \caption{Lineup \#4 - electoral building. Which building looks the most different from the other buildings? }
-%   \label{fig:elect.4}
-%\end{figure}
-%\begin{figure}[htbp] 
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{electoral-1-1.pdf} 
-%   \caption{Lineup \#5 - electoral building. Which building looks the most different from the other buildings? }
-%   \label{fig:elect.5}
-%\end{figure}
-
-\clearpage
-\section{Electoral Building Lineups and Results}
-%Figures~\ref{fig:elect-1} and~\ref{fig:elect.2} through~\ref{fig:elect.5}
-
-Five lineups were shown to (different) Amazon Turk workers in an experiment.
-They all were created as described in the introduction of the paper.
-In order to not bias observers, no context information was given about how these plots were created or what data they were displaying.
-This also made it necessary to slightly more stylize the display.
-
-% latex table generated in R 3.0.1 by xtable 1.7-1 package
-% Wed Jun  5 17:39:09 2013
-\begin{table}[hbtp]
-\centering
-\begin{tabular}{rrrrrr}
-  \hline
- lineup & \#1 & \#2 & \#3 & \#4 & \#5 \\ 
-  \hline
-  \# correct/ \#evaluation &  12/72 & 11/66 & 5/74 & 14/72 & 19/57 \\ 
-$p$-value & 0.00023 & 0.00041 & 0.31 & 1.2e-05 &  1.9e-11 \\ 
-data panel & $3\cdot4+1$ & $2^4+1$ & $4^2+2 $ & $12+\sqrt{25}$ & $2^3-7$ \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-% latex table generated in R 3.0.1 by xtable 1.7-1 package
-% Wed Jun  5 18:00:55 2013
-\begin{table}[hbtp]
-\centering
-\caption{\label{tbl:response}Overview of all choices by observers for each of the lineups.
-The correct choice is bolded.
-In most lineups there are null plots that were picked more often by observers, but the actual result is among the plots being picked most often, indicating that there is some indication that the election result is not completely consistent with the polls.}
-\scalebox{.9}{
-\begin{tabular}{rrrrrrrrrrrrrrrrrrrrr}
-  \hline
-& \multicolumn{10}{l}{panel chosen}\\
-Lineup & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 & 19 & 20 \\ 
-  \hline
-\#  1 & 2 & 2 & 0 & 10 & 2 & 2 & 6 &  23 & 1 & 1 & 0 & 1 & \bf 12 & 3 & 3 & 1 & 0 & 1 & 1 & 1 \\ 
-\#  2 & 0 & 16 & 1 & 1 & 5 & 1 & 0 & 8 & 0 & 2 & 0 & 0 & 0 & 4 & 2 & 1 & \bf 11 & 1 & 0 & 13 \\ 
-\#  3 & 7 & 26 & 0 & 2 & 0 & 5 & 3 & 0 & 2 & 1 & 0 & 4 & 0 & 0 & 2 & 0 & 9 & \bf 5 & 0 & 6 \\ 
-\#  4 & 0 & 0 & 0 & 2 & 0 & 0 & 0 & 3 & 1 & 10 & 2 & 18 & 1 & 0 & 4 & 2 &\bf 14 & 0 & 13 & 0 \\ 
-\# 5 &\bf 19 & 1 & 4 & 1 & 0 & 1 & 0 & 12 & 0 & 0 & 0 & 4 & 1 & 0 & 0 & 12 & 1 & 1 & 0 & 0 \\ 
-   \hline
-\end{tabular}
-}
-\end{table}
-
-\bibliographystyle{asa}
-\bibliography{references}
-
-\end{document}

### 2017-10-21 submitted-v1/R/sociology-paper-v1.R

@@ -22,6 +22,9 @@ load("../data/experiment-data-all.RData")
 expDat <- allDat
 dat9 <- exp9Dat
 
+# getting the data table information
+# data.frame(sapply(expDat, class))
+# data.frame(sapply(dat9, class))
 
 # ==================================================================
 # generating statistics for section 4.1 "Overview of the Data"

### 2017-10-21 submitted-v1/sociology_paper.Rnw
cleaned the source

@@
-\documentclass[10pt]{article}
+\documentclass[12pt]{article}
+
 \usepackage{times}
-%\usepackage{natbib}
-%\usepackage{multicol}
 \RequirePackage{natbib}
@@
 \usepackage{amsmath, amssymb, fullpage, amsthm, array, graphicx, asa, url}
-%\usepackage[dvips]{graphics}
-
-%\usepackage{hyperref} % for hyper reference
 
@@
 
-% \usepackage{pifont} % this package is used to print check mark \checkmark
-% \linespread{1.6} % factor 1.6 = double space
-
-%\usepackage{setspace}
-%\doublespacing
+% JASA guideline to have double spaced fonts
+\usepackage{setspace}
+\doublespacing
 
@@
 \definecolor{mahbub}{rgb}{1.0, 0.6, 0.4}
-%\usepackage[usenames,dvipsnames]{color}
+
 \newcommand{\hh}[1]{{\color{magenta} #1}}
@@
 
-%\newtheorem{thm}{Theorem}[section]
-%\newtheorem{dfn}{Definition}[section]
-%\newtheorem{cor}{Corollary}[thm]
-%\newtheorem{con}{Conjecture}[thm]
-%\newtheorem{lemma}[thm]{Lemma}
-
-%\topmargin -0.10in   % when making pdf
-%\textheight 9.15in  % when making pdf
-
 \pdfminorversion=4 % as instructed by JASA file upload
+% NOTE: To produce blinded version, replace "1" with "0" below. 
+% Blinded version will also remove acknowledgement at the bottom.
+\newcommand{\blindDoc}{1}
 
 \begin{document}
-%\tableofcontents
-
-% Article top matter
-\title{Human Factors Influencing Visual Statistical Inference }
+
+% ===================================================================================
+%Article top matter 
+%------------------------------------------------------------------------------------
+
+\if1\blindDoc
+{
+  \title{Human Factors Influencing Visual Statistical Inference}
   \author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
@@
 \date{\vspace{-.5in}}
-%\date{\today}  %\today is replaced with the current date
   \maketitle
-
+} \fi
+
+\if0\blindDoc
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+    {\bf \LARGE Human Factors Influencing Visual Statistical Inference}
+\end{center}
+  \medskip
+} \fi
+
+% ===================================================================================
 \begin {abstract} 
+%------------------------------------------------------------------------------------
+
 Visual statistical inference is a way to determine significance of patterns found while  exploring data with graphics.
@@
 
-%\begin{multicols}{2}
-%\twocolumn
-
+% ===================================================================================
 \section{Introduction}  
+%------------------------------------------------------------------------------------
 
@@
 Based on the number of evaluations by independent observers and the number of times that the data plot is identified, we are  able to evaluate the significance of the graphical relationship in the same way and the same rigor as a conventional hypothesis test.
-% is rejected.
-This means that the structure in the actual data plot has significant structure, a pattern that is not simply due to randomness.
-Combining the choices of multiple observers provides more stability in the estimation of significance.
-
-%\subsection{Visual inference example}
 
@@
    \includegraphics[width=0.99\textwidth]{electoral-5-13.pdf} 
-%      \includegraphics[width=0.99\textwidth]{electoral-1-1.pdf} 
    \caption{ \label{fig:elect} Which one of the plots, labelled 1 through 20 from top left to bottom right, is the most different from the others?}
@@
 The corresponding $p$-value is 0.0077, indicating sufficient evidence to reject the null hypothesis.
-% , with reasons given being 'asymmetry' (36\%), 'trend' (26\%) or `outliers' (16\%).
 
@@
 Figure~\ref{fig:tower} shows the plot of the electoral building with the additional context information and labels.
- %A party needs 270 electoral votes or more to win the presidential election. 
-%\begin{figure}[htbp] 
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{electoral-5-13.pdf} 
-%   \caption{Which plot looks the most different from the other plots?  Lineup \#1 - electoral building; another set of four lineups can be found in figs.~{fig:elect.2} through~{fig:elect.5}}
-%   \label{fig:elect.1}
-%\end{figure}
 
@@
 
-%The rest of the 19 plots are generated from the assumption that there is no difference between two groups of data.
-For this, the group structure of the data is broken by random permutation of the groups.
-
-%If the null hypothesis is true, the observed data should look similar to the rest of the plots in the lineup making it difficult to detect the observed plot.
-On the other hand, if the null hypothesis is not true, the observed data should be different from the rest of the plots in the lineup and observers should be able to pick the data plot out from the rest.
-
 A lineup can be evaluated by a single person or multiple observers.
@@
 To avoid expectation errors \citep{meilgaard}, particularly in emotionally charged areas such as election results, observers should not be aware of the data that constitutes a lineup, and should therefore not have seen the actual data plot before inspecting the lineup.
-This is the reason that we revealed the context of the election problem only after the interested reader had a chance to evaluate the lineup.%the election example, above, the scenario was explained after the lineup question, in the text.
-
-When presenting a lineup to an observer  the accompanying question should be phrased in as general a manner as possible, 
-%The question that is asked of the observer should be as general as possible, 
-effectively asking the observer to pick the plot that is most different, and allowing observers to provide their own reasons for their choice.
+This is the reason that we revealed the context of the election problem only after the interested reader had a chance to evaluate the lineup.
+
+When presenting a lineup to an observer the accompanying question should be phrased in as general a manner as possible, effectively asking the observer to pick the plot that is most different, and allowing observers to provide their own reasons for their choice.
 This ensures all possible deviations from the null hypothesis to be detected.
 However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments in \cite{majumder:2013} or \cite{tengfei:2013}, these questions have to be phrased much more specifically, in order to avoid  type III errors \citep{mosteller:48} consisting of correctly rejecting the null hypothesis for a wrong reason.
-%In some of the studies, the ones described in  \cite{majumder:2013} very specific questions were asked, because the experiments were being conducted to compare results from the lineup protocol with those of conventional tests.
 In those experiments, the structure in the data is strictly controlled in the simulation process, which allows for specific questions to be asked.
+
 In the election example, observers were asked ``which plot is the most different?''.
@@
 From the panels picked, we see that the top three choices  are all extreme with respect to heights of blue versus red towers:  both panels \#8 and \#13 show a large difference between the height of the towers, while panel \#4 is the only panel, in which the red tower is higher than the blue one.
-% results from the turk study:
-% panel:   1 10 11 12 13 14 15 16 18 19  2 20  3   4  5  6  7  8  9 
-% #picks:  2  1    0  1 12    3  3    1  1   1  2   1  0 10  2  2  6 23  1
-
-%Most observers got this, but it is possible that some observers might pick plot~4, where the red tower is slightly above the blue as the most different because it is the only plot with this feature \hh{(10 out of 72 observers actually did this)}.
-  %a better question may
-
-% \hh{If we consider this difference to be the main objective of the visualization, a more focused question, that aligns more with a hypothesis test on the difference  in electoral votes  between parties, would} have been ``Which plot shows the biggest height difference between the two towers?'' %except that this tailors the inference to a specific feature which does not match the null hypothesis of interest.
-
-% HH already included:
-% For any evaluation the observer may or may not identify the actual data plot.
-Under the null hypothesis, the actual data plot should look similar to the null plots making it harder to detect.
-It is not expected that an observer would be able to detect the data plot in this scenario.
-But since there are limited number of plots in a lineup, which is 20 in the election example, there is a 1/20 chance that the observer would pick the actual plot.
 
@@
 Visual power depends unlike power in conventional tests not only on the strength of the signal, but also on individuals' ability.
-%This proportion is associated with the Type~I error of the test.
-On the other hand, if null hypothesis is not true, the observed plot should look different from the null plots, making it easier to be detected.
-This is the definition of the power of the test.
-When multiple observers evaluate a lineup, the detection rate can be used to estimate the power.
 The ability of individual observers varies, and the effects that might influence this ability are the focus of this paper.
-
-%\dc{*** Am trying to use detection rate rather than correct response.
-This is harder for this paper because we are evaluating subject's abilities to some extent, but we should try to use this in our vis inference papers.} % \mm{Di, do you suggest that I change ``correct response''  to ``detection rate'' in all the texts here?}
-
-%For a lineup with fixed difficulty, the proportion correct responses may vary for different observers.
-If an observer evaluates multiple lineups, the proportion of correct evaluations by the observer could be used as a measure of performance by an observer.
-The another way of measuring the performance of the observer could be the time taken by an observer to evaluate a lineup.
-Since performance would also depend on the lineup difficulty it is essential to adjust for lineup difficulty before estimating the performance of the observer.
-
-%Some of the factors that may affect the performance of the observer are discussed in Section~\ref{sec:factor_performance}.
-In this paper we focus on estimating the learning effect of the observer in terms of both proportion correct and time taken to evaluate a lineup through the attempts made during the evaluation of multiple plots.
-We also estimate the location effect of the observed data plot in a lineup.
-The other factors are controlled in simulation experiments so that learning effect and location effect can be estimated.
 
@@
 
-%\subsection{Experiments pooled}
-
+% ===================================================================================
 \section{Factors Potentially Affecting Visual Inference} \label{sec:factor_performance} 
+%------------------------------------------------------------------------------------
 
@@
 of the factors that are expected to have some impact on the performance of visual statistical inference.
-
-%ased on human evaluation on a lineup a decision is made in visual statistical inference.
-While the visual test statistic is defined to have a direct influence on the observer to pick a plot based on the signal in the data, there are other human factors that may affect the observer performance.
-It is important to study the effect of those factors and examine the extent of their influence.
-This section presents a brief description of the factors that may affect the power of visual inference.
-
-%\dc{You need to have more than one sentence to open this section, it needs a paragraph overviewing the factors.}
 
@@
 
-%On the other hand, if the null hypothesis is not true, visual test statistic should not display any distinguishable pattern.
-In fact, some of the null plots may appear to be the most different plot in a lineup influencing the observer to chose a plot different than actual data plot.
-This is an elegant feature of the lineup.
-It force the observer to chose a wrong plot when the null hypothesis should not be rejected.
-
-%Plots to do: for the same signal strength show the proportion correct for scatterplot and boxplot. eg., qplot(effect, ump-visual power, color= plot-type)
-
-%\subsection{Question Design} 
 \item{\bf Question Design:} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
@@
 
-% The researcher knows about the underlying hypothesis but the observer does not necessarily know the underlying details of the lineup.
-So, the researcher needs to ask a question to the observer to answer while evaluating the lineup.
-This question should provide the observer a little clue so that the answer reflects the hypothesized patterns in the actual plot.
-For example Table~\ref{tbl:visual_stat} shows the questions asked for the simulation experiments considered for this study.
-Notice that for case 1 if the observer can identify the actual plot in the lineup that should indicate that the plot chosen has the most vertical difference between groups A and B which is exactly what the researchers intend to examine.
-Similarly for case 2, a correct evaluation would indicate that the slope is different than the slope that may show up just from randomness.
-
-%These questions are very crucial for the power of visual test.
-They help observer think in a meaningful direction.
-Notice that there may be unnecessary patterns in the actual data plot which may not necessarily indicate the existence of the significant signal in the plot.
-These questions help observer not to be misguided by those patterns.
-To follow up further on this \citet{majumder:2013} also collected reasons of choice made by the observers and noticed that Type-III error may occur where the hypothesis may be correctly rejected but for a wrong reason.
-
-%\subsection{Observer Personality}
-
-%\subsection{Visual Perception of Human Eye} Different people observe the lineup in different ways.
-With the help of eye-tracking equipment \citet{zhao:2012} track the observers' eyes to see how they go through the plots in a lineup to come to their answers.
-The result suggest that people have particular methods of reading lineups.
-Some people read lineups from left to write direction while some read from upward to downward.
-Some people start looking from the center of the lineup while others start from the top left corner.
-In the earlier phase of the exercise, the observer tend to scan the plots and then start comparing plots to make a final decision.
-Beside right-left or up-down directions observer show some diagonal movement too.
-%
-%Given this pattern of human eye tracks, it may effect the performance of an observer depending on where the actual plot is placed in the lineup.
-If the actual plot is on the top left corner and the observer starts from that point it may be easier to detect the actual plot earlier in the exercise and the observer could get plenty of time to make comparison.
-Thus the position of the actual plot in a lineup may have some impact.
-
 %\subsection{Demographics} 
@@
 
-%Some of the demographics of the observer such as age, gender, education level and geographical location may have effect on how an observer would examine the lineup.
-This may produce some variability in the performance of the observer.
-For example, a high school student may not respond same as a well educated person while evaluating a lineup.
-There may be variations in different age groups as well.
-Investigation on subjects with a well variety of age groups is necessary to study this variability.
-
-%To meet the Institutional Review Board (IRB) requirement, it is necessary to exclude any subject less than 18 years of old to participate human subject experiment.
-Thus we intend to only focus on the performance of observer with age 18 years or older.
-We don't expect many older people to participate our study and we are interested of age from 18 to 65 years.
-
-%Education level has some relation with age.
-Younger observer may not have higher degree.
-Thus some of the variability in performance for different education level may be confounded with age.
-On the other hand, it could be possible to have undergrad degree with higher age.
-Specially for any geographical location where not many certain gender group have higher education, this may occur.
-Thus the effect of all these demographical factors may be confounded at certain level.
-
-%We don't expect much difference in performances between male and female observers.
-Gender may have some influence through the education level and geographical locations since some location may have small number of educated female population.
-We intend to include similar number of male and female subjects in our study so that this can be examined.
-
-%\subsection{Learning Trend} 
 \item{\bf Learning Trend:} One might expect that as an observer evaluates more lineups they  become more skillful in their evaluations.
@@
 These two ways of measuring learning trend are evaluated on data collected from experiments~5, 6 and~7.
-
-%When an observer evaluates a lineup he or she may learn something from the experiences of the earlier observations.
-If the same observer is shown another lineup the learning from previous evaluation may help.
-The more evaluations an observer makes the more skillful the observer may become.
-This learning trend may or may not be significant overall but it may influence the performance of the observer in some way.
-
-%Learning may occur in two different ways.
-An observer can become more trained on the lineup structure and the pattern in the plots shown.
-This can be learning in proportion of correct evaluation.
-Or, he or she may become efficient in responding faster in the later evaluations.
-This can be observed as the time taken for evaluating a lineup.
-This paper investigates both of these learning trends with multiple simulation experiments.
-
-%When an observer evaluate a lineup for the first time, he or she has to read through many instructions and become familiar with the environment of the experiment.
-We expect that this will require much more time compared to the rest of the evaluations.
-Time taken for the second or successive attempt should be much lower than the time taken for the first lineup.
-We attribute this to the adaptation to the experiment environment, not a learning of evaluating a lineup.
 
@@
 
-%While evaluating the lineup some people may start looking from some specific part of the lineup.
-In the earlier phase of the exercise, the observer tend to scan the plots and start comparing plots to make a final decision.
-Beside right-left or up-down directions observers show some diagonal movement too.
-
-%Given a specific pattern of eye movement of the observer in examining the lineup, the location of actual plot in the lineup may have some effect.
-Those who start exploring from the left top corner may get first glance of the actual plot if it is on that location.
-This should give the observer more time to compare it with the rest of the null plots.
-Those who start looking from the center may scan towards right or left direction and thus don't have the chance to scan the null plots continuously as they could if they would start from a corner.
-Thus they should scan the center plot over and over again while scanning the left or right side plots.
-If the data plot is in the center location it may have multiple chance to be examined and hence be identified.
-
-%It may be possible that some part of the lineup is less explored or even never scanned by the observer if he or she feels that the actual plot is found before even seriously scanning the whole lineup.
-In those scenarios the location where the observer first start scanning is important.
-Placing the actual data plot in that location may yield different results.
-
-%\subsection{Sample of Null Plots} 
 \item{\bf Sample of Null Plots:} In classical inference, the test statistic is compared to the full null distribution, to decide if it is extreme or not.
@@
 This is tested with the data from experiment~9, where the experiment was set up with lineups made from different null plots.
-
-%each lineup, there is a sample of draws from a null distribution.
-A lineup becomes difficult to evaluate if one of the null plots appears to be very similar to the actual plot.
-When null hypothesis is true it is a common scenario.
-With alternative hypothesis being true,  it may happen if  we compare actual plot with many null plots.
-But we have only specific number of null plots in a lineup.
-So, a different sample of null plots may yield some variation in the difficulty of the lineup with the same actual plot.
-This may affect the performance of the observer while evaluating a lineup.
 
@@
 
+% ===================================================================================
 \section{Experimental Methods}\label{sec:exp_design}
-
-%While evaluating a lineup, the performance of the observer may depend on the factors described in Section~\ref{sec:factor_performance}. 
+%------------------------------------------------------------------------------------
+
 Two of the factors, signal in the data and individual abilities, were studied in \cite{majumder:2013}.
@@
 
-% but for the most part data collected to examine demographic effects .
-In this paper we intend to study effect of the factors such as human demographics, learning trend, actual plot locations and selection of null plots.
-This section presents the simulation experiment setup, data collection methods and data analysis plans with models to asses the influence of those factors on visual statistical inference.
-
-%\dc{Learning trend and location effect need to be separate sections, cannot be simply bolded.}
-
-\subsection{Demographic Factors}  %In addition to obtain evaluation responses of the lineups, all the 10 experiments shown in Table~\ref{tbl:visual_stat} were designed to collect the following demographic information of the subjects participating the experiments:
+\subsection{Demographic Factors}
 \noindent
@@
 \end{enumerate}
-%
-%Instead of collecting exact age, 9 levels of age were collected.
-They are 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, 51-55, 56-60, above 60.
-There were five levels of academic information.
-They are (1) High school or less , (2) Some under grad course, (3) Under graduate degree, (4) Some graduate courses,  (5) Graduate degree.
-Geographical locations were collected using the true public ip addresses of the participants' computer.
-This provides latitude and longitude of the ip address as well as the city and country information.
-%
+
 Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
 The factors are examined in association with the observer's response using a logistic regression with random effect terms:
+
 \begin{equation} \label{eqn:demographic_response}
@@
 \end{equation}
-where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ picks the actual data plot from lineup $j$, %We model this in a generalized mixed effects model of the form
-%\begin{equation} \label{eqn:demographic_response}
-%g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
-%\end{equation}
- $\mu$ is an overall population average, %probability for picking out the data plot from a lineup HH: mu is not a probability, it is on the scale of the link function
- $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively, for observer $i$.
+
+where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ picks the actual data plot from lineup $j$, $\mu$ is an overall population average, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively, for observer $i$.
 The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume independence and normality of the errors, i.e.\ $\ell_j \sim N(0, \sigma_\ell^2)$.
@@
 Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups.
-%, so ``correct'' is more of a convenient rather than ideal word to use for this analysis.
 So in this context, detected, or not, is used as a response to examine individual differences.
 
-%\subsection{Model to Estimate Learning Trend}
-
-%Multiple lineups were sequentially shown to each observer for evaluation. 
 Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k$th evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
@@
 \end{equation}
-where $\mu$ is an overall population average, %probability for picking out the data plot from a lineup, 
-$\alpha_k$ is the effect of the $k$th attempt on the probability, using the first attempt as reference, $\alpha_1 = 0$, and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
+where $\mu$ is an overall population average, $\alpha_k$ is the effect of the $k$th attempt on the probability, using the first attempt as reference, $\alpha_1 = 0$, and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
 The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
@@
 We evaluate this in a mixed effects model of the form
-%\hh{
+
 \begin{equation} \label{eqn:trend_time}
@@
 We employ a normal approximation to obtain $p$-values corresponding to fixed effect parameters estimates.
-% normal approximation is used to compute $Z$ scores  as the ratio of estimates to  estimated standard errors.
 
@@
 
-%Two factors were of primary interest: genotype (call them RPA and EV) and the interaction between genotype and iron condition (sufficient or insufficient).
-The $2\times2$ factorial design provides four treatment conditions.
-For each gene there were 11 data points, because each treatment contained two or three replicates.
-\mm{Di, There are more than 40,000 genes examined for significant expression in \citep{tengfei:2013}.
-Is that what you wanted to put here?}.
 In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
 This overall significance is studied using visual inference  in \citet{tengfei:2013}.
-
-%For the study, the routine processing of the data was conducted to find (1) the gene that exhibited the most significance difference in expression between genotypes, (2) the gene that exhibited the most different expression on RPA, but insignificant expression on EV.
-Two different types of plots would be made to examine this.
-For (1) side-by-side dotplots of the expression levels for the one gene by genotype were created.
-For (2) an interaction plot, with iron sufficiency represented on the horizontal axis, expression vertically, and genotype by color.
-Colored lines connect the means for the two genotypes over iron treatment.
-The null plots were created by randomizing the experimental design, and re-doing everything else as done with the actual data.
-Any genes found under these circumstances had expression differences that can only be considered to by consistent with random differences.
-Full results are documented in \cite{tengfei:2013}.
 
@@
 
-%It is set up based on the gene expression data \citep{atwood:2013} with two groups.
-One is the Genotype and the other the Empty Vector (EV).
-For each of the groups, gene expression data were collected in presence or absence of iron sufficiency.
-Two factors were of primary interest.
-One is the Genotype main effect and the other is the interaction effect between Genotype and iron sufficiency.
-
-%For Genotype main effect side by side dot plots of two groups are used as the visual test statistics.
-For interaction effect the same plot type is used except that the means of iron sufficiency or insufficiency were connected by lines colored by the groups.
-One of the interaction test statistics is shown in Table~\ref{tbl:visual_stat} for experiment 9.
-For no interaction, the visual test statistic shows two parallel lines representing two groups and for significant interaction the lines are not parallel.
-Null plots were generated by randomizing the group structure of the data.
-The actual plot was randomly placed in five different locations in a lineup of size 20.
-The locations are 2,9,12,16,20 for Interaction effect and 1,8,12,17,20 for Genotype effect.
-For a lineup with specific data plot location, five different sets of null plots were used to produce 5 lineups for each location.
-In total we have 25 lineups for Interaction effect and 25 lineups for Genotype effect.
-
-%Each observer saw three lineups, first an Interaction lineup then a Genotype lineup.
-Finally a test lineup was shown to screen out unusual evaluation.
-The subjects did not know which one was the test lineup but they were informed before accepting the task that there would be one.
-The test lineup was very easy and everyone should detect the data plot.
-The MTurk workers were paid based on whether they could correctly evaluate the test lineup.
-But the data on the test plot are excluded from the analysis.
-
-%\subsection{Model to Estimate Location Effect}
-
-%As per the design of experiment for location effect, the actual data plot is same for each of the null sets of plot. Thus the response data of this experiment constitute a multivariate response. 
 To examine if the difference in detection rate among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
@@
 
-%It is very fast, cheap and reliable to recruit people from MTurk.
-Thats why it is getting very popular among the researchers who perform human subject experiment.
-The another benefit is that a very diverse pool of subjects can be recruited which is otherwise very hard to obtain for a study.
-%The researchers can easily filter the workers based on their experimental design, such as recruiting people only from a specific geographical location or a group of people who satisfy certain criteria etc.
-The recruiter can decide who they pay or not.
-Workers have to satisfy the task requirement to ensure payment.
-But at the end it is the recruiter who has the final say.
-Usually recruiters pay promptly after the task has been done properly and thats why MTurk is very popular among the online job seeker.
-At any point of time thousands of tasks are available for the thousands of workers around the world.
-
-%Because of its convenience it is getting popular for scientific research study as well.
-In comparison with a lab study \cite{suri:2010} perform the same study using MTurk and demonstrate that their study results are as good as the lab study results even though MTurk study required less time and cost while provided more convenience.
-\cite{majumder:2013} recruited people from MTurk for their simulation study in estimating the power of visual statistical inference.
-They have done numerous pilot studies in lab before doing actual MTurk study and found similar results.
-\cite{mason:2012} explains various features of MTurk and describes how it can be used as part of human behavioral study.
-
 We designed and developed a web application which enables the display of lineups to  observers as per experimental need.
@@
 The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
-%It is measured in seconds.
 Location of an observer is determined based on the ip address of the observer.
 
-%\subsection{Model to Estimate Demographic Factor Effect} The feedback provided by each observer on a lineup is a binary response variable.
-Suppose $Y_{ij}$ denotes the response from observer $i$ on a  lineup $j$.
-$Y_{ij} =1$ if the response is correct otherwise $Y_{ij} =0$.
-Let $\pi_{ij}=  E(Y_{ij})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$.
-We model this in a generalized mixed effects model of the form
-%\begin{equation} \label{eqn:demographic_response}
-%g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
-%\end{equation}
-%where $\mu$ is an overall average probability for picking out the data plot from a lineup.
-$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender category $l(i)$, education level $m(i)$ and country name $s(i)$ respectively for observer $i$.
-$\ell_j$ is a random intercept predicting lineup difficulty level and we assume  $\ell_j \sim N(0, \sigma_\ell^2)$.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-%
-%
-%The effect of demographic factors can be observed as time taken for each evaluation by an observer.
-Suppose $Z_{ij}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$.
-Let $\mu_{ij}=  E(Z_{ij})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$.
-We model this in a mixed effects model of the form
-%%\hh{
-%\begin{equation} \label{eqn:demographic_time}
-%Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
-%\end{equation}
-%where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
-$\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in Model~\eqref{eqn:demographic_response}.
-$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-
-%\subsection{Model to Estimate Learning Trend}
-%
-%Multiple lineups were sequentially shown to each observer for evaluation.
-Suppose $Y_{ijk}$ denotes the response from observer $i$ on a  lineup $j$ in his/her $k$th attempt.
-$Y_{ijk}=1$ if the response is correct otherwise  $Y_{ijk}=0$.
-Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ correctly picks the data panel from lineup $j$ in his/her $k$th attempt.
-We model this in a generalized mixed effects model of the form
-%\begin{equation} \label{eqn:trend_response}
-%g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
-%\end{equation}
-%where $\mu$ is an overall average probability for picking out the data plot from a lineup, $\alpha_k$ is the effect of the $k$th attempt on the probability, with $\alpha_1 = 0$ and $k = 1, ..., K$.
-%$u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$. $u_i$ is a random intercept, describing a basic subject-specific ability. We assume $u_i \sim N(0, \sigma_u^2)$. 
-%$a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts. We assume $a_i \sim N(0, \sigma_a^2)$. 
-%For $\ell_j$ we again assume a normal distribution, $N(0, \sigma_\ell^2)$.
-$\ell_j$ is a random intercept predicting lineup difficulty level.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$ .
-%
-%The inverse link function, $g^{-1}(.)$, from  Equation~\ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in $k$th attempt by a single observer as
-%\begin{equation} \label{eqn:trend_power}
-%\hat p_{ijk} =  g^{-1}(\hat{\mu} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
-%\end{equation}
-%
-%The learning of each observer over a course of $K$ evaluations may be observed as the improvement of the time taken to evaluate a lineup in the later attempts.
-Suppose $Z_{ijk}$ denotes the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
-Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of log(time taken) by  observer $i$ to pick the data panel from lineup $j$ in his/her $k$th attempt.
-We model this in a mixed effects model of the form
-%%\hh{
-%\begin{equation} \label{eqn:trend_time}
-%Z_{ijk} = \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
-%\end{equation}
-%where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup.
-$\alpha$ is the average change in log time taken for each additional attempt.
-$\alpha_1$ is an offset in log time taken for the first attempt.
-All other effects are random effects: as before, $u_i$ is a subject-specific intercept representing  individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$.
-$a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$.
-We assume $a_i \sim N(0, \sigma_a^2)$.
-$\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup; $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-%
-%
-%Equation~\ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
-%\begin{equation} \label{eqn:trend_time_est}
-%\hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j.
-%\end{equation}
-%
-%To fit all these mixed effect models the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
-To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
-
-%\subsection{Model to Estimate Location Effect}
-%
-%
-%As per the design of experiment for location effect, the actual data plot is same for each of the null sets of plot.
-Thus the response data of this experiment constitute a multivariate response.
-To examine if the difference in proportion correct among the locations is statistically significant we fit a one way multi variate analysis of variance (MANOVA) model to the data.
-%
-%Suppose $\mathbf{Y}=(Y_1,Y_2, ...
-, Y_p)^\top$ be a vector of random variable with dimension $p$, the total number of null sets.
-let $\mathbf{Y}_{ij}$ represents $j$th vector response for $i$th location with $i=1,2, ..., I$.
-We fit the following MANOVA model
-%\begin{equation}\label{manova}
-%\mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
-%\end{equation}
-%where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$.
-
+% ===================================================================================
 \section{Results}\label{sec:result_socio}
+%------------------------------------------------------------------------------------
 
@@
 The diversity in not only  geographic but also in gender, age group and education level as can be seen in Table~\ref{tbl:demographics}.
-% It is interesting that there were large number of female participants even though there were a lot of participants  from developing countries.
 The large number of female participants from all countries was a pleasant surprise to us.
@@
 The distribution of participants is similar in all ten experiments.
-%This may be due to the time of day when the experiment was made available.
-%For some experiments more people participated from India than the United States and for others numbers were reversed (Figure~\ref{fig:participation_time}).
-It is also observed in some experiments that number of  participants are similar no matter whether the experiment is run on day time or nigh time.
-(*** Took out reference to Figure~11 in appendix.)
 
@@
 
-% This table contains the demographic information without factor wise percentages.
-%\begin{table}[hbtp]
-%\caption{Demographic information of the subjects participated the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
-%\centering
-%\scalebox{.9}{
-%\begin{tabular}{rlrrr}
-%  \hline
-% Factor & levels & subjects & avg\_time & response \\ 
-%  \hline
-%Gender & Male & 1348 & 48.51 & 13493 \\ 
-%   & Female & 991 & 43.75 & 10564 \\ 
-%\hline
-%  Education & High school or less & 193 & 37.21 & 2241 \\ 
-%   & Some under graduate courses & 418 & 42.84 & 4070 \\ 
-%   & Under graduate degree & 584 & 44.29 & 5775 \\ 
-%   & Some graduate courses & 245 & 43.43 & 2460 \\ 
-%   & Graduate degree & 902 & 52.18 & 9511 \\ 
-%\hline
-%  Age & 18-25 & 740 & 42.97 & 7311 \\ 
-%   & 26-30 & 547 & 46.27 & 5585 \\ 
-%   & 31-35 & 376 & 44.27 & 3923 \\ 
-%   & 36-40 & 257 & 55.03 & 2714 \\ 
-%   & 41-45 & 141 & 43.90 & 1519 \\ 
-%   & 46-50 &  95 & 49.29 & 1003 \\ 
-%   & 51-55 &  83 & 48.67 & 867 \\ 
-%   & 56-60 &  64 & 59.73 & 678 \\ 
-%   & above 60 &  38 & 48.67 & 457 \\ 
-%\hline   
-%  Country  & United States & 1087 & 39.64 & 10769 \\  
-%  & India & 980 & 52.63 & 10227 \\ 
-%  & Rest of the world & 254 & 46.86 & 2819 \\ 
-%  
-%   \hline
-%\end{tabular}
-%}
-%\label{tbl:demographics}
-%\end{table}
-
 The largest number of participants falls within the age group of 18 to 25, with the majority being between 18 to 35.
@@
 The distribution of education levels also differs slightly across countries for this age group.
-%Most of the participants from India are below age 40 while in United States the distribution of participants are similar beyond age 40.
-There were not many participants beyond age 50 and hence these age groups are merged to form one age group called above 50.
-The exploratory analysis and the models fitted in the following sections use this new age group instead of actual age groups beyond age 50.
 
@@
 
-%\dc{*** It seems repetitive to start with a description of what we learn from the plot, so start with the model summary and use the plot to support the model findings.}
-
 To explore the significance of the demographic factors model~\eqref{eqn:demographic_response} is fit to the data, with age, country, education and gender as fixed effects.
@@
 With respect to time taken, all of the demographic factors are significant.
-
-% \begin{table}[htbp]
-% \centering
-% \caption{Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
-Gender does not have a significant effect on detection rate, but does on time to respond.
-All factors significantly affect time to respond.}
-% \scalebox{.9}{
-% \begin{tabular}{rlrrrrrr}
-%   \hline
-% \multicolumn{2}{c}{Model}  & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
-%   \hline
-% Proportion Correct &Full & 23822.00 &23943.00 &-11896.00  & &  &  \\ [3pt]
-% %& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
-% {Reduced Models}  &- Age & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
-%   &- Country  & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
-%   &- Education  & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
-%   &- Gender  & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
-%    \hline
-%    Log Time Taken &Full &  51904.00 &52034.00 & -25936.00 & &  &  \\ [3pt]
-% %& \multicolumn{2}{l}{Reduced}  &  &  &  & \\
-% Reduced Models &- Age  & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
-%  &- Country  & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
-%   &- Education  & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
-%   &- Gender  & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
-% \hline
-% \end{tabular}
-% }
-% \label{tbl:anova_factor}
-% \end{table}
 
@@
 
-% \begin{table}[htbp]
-% \centering
-% \caption{Parameter estimates of models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively.
-For time taken all the demographic factors are significant.
-For detection rate, age group 36-40, rest of the world and graduate degree are significantly different.
-For gender no difference in performance is observed.
-Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.
-% %
-% }
-% \scalebox{0.86}{
-% \begin{tabular}{rlrrrrrrrrrr}
-%   \hline
-% Demographic&& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})} \\
-% 
-% \cline{3-6} \cline{8-11} 
-% Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
-%   \hline
-% \multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
-% &$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
-% \hline
-% Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
-%  &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.206 \\ 
-%   &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.038 \\ 
-%   &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
-%   &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.050 \\ 
-%   &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.143 \\ 
-%   &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.039 \\ 
-%   \hline
-% Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
-%   &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
-%   &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
-%   \hline
-% Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
-%  &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
-%   &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.331 \\ 
-%   &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.157 \\ 
-%   &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
-%   \hline
-% Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
-%   &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.133 \\ 
-%   \hline
-% \multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
-% & lineup($\sigma_{\ell}$) & 0.287 &  &  &  &   & 2.293 &  &  &  \\ 
-%  & Error($\sigma$) & 0.692 & &  &  &  & 0.893 && &\\ 
-%    \hline
-% \end{tabular}
-% }
-% \label{tbl:model_result_demographics}
-% \end{table}
-
 Figure~\ref{fig:demographic_effect} supports the model results.
@@
 Mean values are shown as dots inside the boxplot, as this statistic rather than the median is used in the models.
-%There are some small differences across demographic groups.
-The youngest age group (18-25) took least average time to evaluate a lineup.
-People from India took more time than others.
-People who have only a high school degree took less time and this is related to what we have seen for young age group.
-Averages of log time taken by male and female participants look similar but there are some differences as we see in Table~\ref{tbl:demographics}.
 The means and medians for response time (top row) are essentially equal.
@@
 
-%Some differences in median detection rate for different demographic factor levels can be seen,  but the variability in responses is huge.
-This variability is primarily in variability in the lineups -- some very easy and some extremely difficult lineups.
-Some of the lineups  were not expected to be evaluated correctly as the null plots in those lineups were showing more signal than the actual data plot itself.
-In those scenarios,  one may not reject the null hypothesis.
-We call these difficult lineups.
-Some of the lineups were so easy that data plot is detectable at a first glance.
-These were expected to be evaluated correctly 100\% of the time.
-This is why we see the boxes range from 0 to 1 in most of the factor levels.
-Unlike median, fewer differences are observed in the average proportion of correct responses for different factor levels.
-
-%All of the factors are significant in describing time taken to evaluate a lineup.
-
 \begin{figure}[htbp] 
@@
 \end{figure}
-
-%\dc{*** Vertical axis needs to be labelled ``detection rate'' for bottom row, and ``value'' needs to be removed.}
-
-%The distribution of time taken to evaluate a lineup is positively skewed.
-But in log scale it appears to be symmetric for all the demographic factor variables as we see in Figure~\ref{fig:demographic_effect}.
-Mean and medians are very similar as well.
-This allows us to fit linear mixed model to the log time taken with normal error structure.
-
-%To further illustrate how the individual factor levels differ from each other  the detailed results of models \eqref{eqn:demographic_response} and \eqref{eqn:demographic_time} are shown in Table~\ref{tbl:model_result_demographics}.
-For parameter estimation the first factor levels shown in the Table are  used as the base line.
-Demographic factor levels have significant effects on log time taken to evaluate a lineup.
-But not all the levels are significant for the probability of correct responses.
-
-% This table had 4 digits $p$-value for last column
-%\begin{table}[htbp]
-%\centering
-%\caption{Parameter estimates of Models \eqref{eqn:demographic_time} and \eqref{eqn:demographic_response} fitted for average log time taken and probability of correct lineup evaluations respectively.
-For time taken all the demographic factors are significant.
-For probability of correct response age group 36-40, rest of the world and graduate degree are significantly different.
-For gender no difference in performance is observed.
-Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
-%\scalebox{0.86}{
-%\begin{tabular}{rlrrrrrrrrrr}
-%  \hline
-%Demographic&& \multicolumn{4}{c} {Model~\eqref{eqn:demographic_time} Log Time} & &\multicolumn{4}{c} {Model~\eqref{eqn:demographic_response} Proportion Correct} \\
-%
-%\cline{3-6} \cline{8-11} 
-%Factor & Level& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value  \\ 
-%  \hline
-%\multicolumn{2}{l}{Fixed Effect} &  &  &  &  &  & & && \\ 
-%&$\mu$ & 3.360 & 0.013 & 249.21 & $<$0.001 &   & -0.683 & 0.071 & -9.64 & $<$0.001 \\ 
-%\hline
-%Age ($\alpha$)& 18-25&0.000   &---  &---  &---  &  &--- &--- &---& ---\\   
-% &26-30 & 0.058 & 0.013 & 4.50 & $<$0.001 &   & 0.062 & 0.049 & 1.27 & 0.2058 \\ 
-%  &31-35 & 0.068 & 0.014 & 4.72 & $<$0.001 &   & 0.115 & 0.055 & 2.08 & 0.0376 \\ 
-%  &36-40 & 0.231 & 0.016 & 14.05 & $<$0.001 &   & 0.310 & 0.063 & 4.93 & $<$0.001 \\ 
-%  &41-45 & 0.176 & 0.021 & 8.56 & $<$0.001 &   & 0.158 & 0.081 & 1.96 & 0.0501 \\ 
-%  &46-50 & 0.272 & 0.024 & 11.29 & $<$0.001 &   & 0.141 & 0.096 & 1.47 & 0.1425 \\ 
-%  &above 50 & 0.352 & 0.018 & 19.19 & $<$0.001 &   & 0.147 & 0.071 & 2.06 & 0.0394 \\ 
-%  \hline
-%Country($\kappa$)& United States&0.000   &---  &---  &---  &  &--- &--- &---& ---\\  
-%  &India & 0.101 & 0.011 & 9.11 & $<$0.001 &   & 0.058 & 0.043 & 1.33 & 0.183 \\ 
-%  &Rest of world & -0.129 & 0.009 & -13.82 & $<$0.001 &   & 0.185 & 0.035 & 5.22 & $<$0.001 \\ 
-%  \hline
-%Education($\tau$) &High school or less&0.000   &---  &---  &---  &  &--- &--- &---& ---\\     
-% &Under grad courses & 0.042 & 0.013 & 3.25 & 0.0011 &   & -0.083 & 0.050 & -1.65 & 0.098 \\ 
-%  &Under grad degree & -0.037 & 0.012 & -3.21 & 0.0013 &   & -0.044 & 0.045 & -0.97 & 0.3307 \\ 
-%  &Graduate courses & 0.117 & 0.013 & 9.12 & $<$0.001 &   & 0.070 & 0.050 & 1.42 & 0.1566 \\ 
-%  &Graduate degree & 0.046 & 0.011 & 4.12 & $<$0.001 &   & 0.182 & 0.043 & 4.22 & $<$0.001 \\ 
-%  \hline
-%Gender ($\gamma$)& Female&0.000   &---  &---  &---  &  &--- &--- &---& ---\\      
-%  &Male & 0.078 & 0.009 & 8.26 & $<$0.001 &   & 0.055 & 0.036 & 1.50 & 0.1331 \\ 
-%  \hline
-%\multicolumn{2}{l}{Random Effect} &  &  &  &  &  & & && \\ 
-%& lineup($\sigma_{\ell}$) & 0.082 & 0.287 &  &  &   & 5.259 & 2.293 &  &  \\ 
-% & Error($\sigma$) & 0.479 & 0.692 &  &  & & && &\\ 
-%   \hline
-%\end{tabular}
-%}
-%\label{tbl:model_result_demographics}
-%\end{table}
-
-%Age group 36-40 is significantly different than the base line age group of 18-25 year olds.
-The other age levels are similar in explaining the probability of correct response.
-For country, the rest of the world is different from USA but India appears to be similar to USA.
-We see from Table~\ref{tbl:demographics} that most of the participants (about 90\%) are from India and USA.
-The rest of the 10\% data are from 76 different countries.
-This diversity in the rest of the world may make it different from India and USA.
-
-%The graduate degree holders are significantly different as well.
-The positive parameter estimate indicates that they perform better and the probability of correct response is higher than other education levels.
-There is no significant difference between male and female performances.
-
-%In Table~\ref{tbl:model_result_demographics} notice that lineup specific variance estimate is 5.259 with a standard error of 2.293 which is much higher than the other significant parameter estimates in model~\eqref{eqn:demographic_response}.
-This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
-Because of the huge impact of lineup variance, the practical impact of other factor variables on the probability of correct response is in fact very small.
-We illustrate this with the following example of graduate degree.
 
@@
    \centering
-%   \includegraphics[width=6.5in]{learning_trend.pdf} 
     \includegraphics[width=3.2in]{practical_impact_und_graduate.pdf} 
@@
 
-%\begin{figure}[htbp] 
-%   \centering
-%%   \includegraphics[width=6.5in]{learning_trend.pdf} 
-%    \includegraphics[width=3in]{practical_impact_demographics.pdf} 
-%   \caption{ Change in probability of correct response due to graduate degree as compared to high school degree diminishes rapidly as the variability in lineup difficulty $\sigma_{\ell}$ increases.}
-%   \label{fig:practical_impact_demographics}
-%\end{figure}
-
-%Model~\eqref{eqn:demographic_time} is fitted for the log time taken to evaluate a lineup and the results are also shown in  Table~\ref{tbl:model_result_demographics}.
-The effect of all the demographic factors on log time taken to evaluate a lineup appears to be significant.
-
-%
-%\begin{table}[htbp]
-%\centering
-%\caption{Anova of full model with all the demographic factors vs reduced model with removing respective factor.}
-%\begin{tabular}{lrrrrrrrr}
-%  \hline
-%Model && Df & AIC & BIC & logLik & Chisq & Chi.Df & $p$-value \\ 
-%  \hline
-%Log Time &Full& 16 &  51904.00 &52034.00 & -25936.00 & &  &  \\ 
-%& \multicolumn{2}{l}{Reduced} &  &  &  &  & \\
-%&Age & 10 & 52435.63 & 52516.41 & -26207.81 & 543.18 & 6 & $<$0.001 \\ 
-% & Country & 14 & 52849.92 & 52963.15 & -26410.96 & 949.47 & 2 & $<$0.001 \\ 
-%  &Education & 12 & 52012.68 & 52109.61 & -25994.34 & 116.23 & 4 & $<$0.001 \\ 
-%  &Gender & 15 & 51970.57 & 52091.74 & -25970.28 & 68.12 & 1 & $<$0.001 \\ 
-%\hline
-%Proportion Correct &Full& 15 & 23822.00 &23943.00 &-11896.00  & &  &  \\ 
-%& \multicolumn{2}{l}{Reduced} &  &  &  &  & \\
-%  &Age & 9 & 23835.22 & 23907.93 & -11908.61 & 25.50 & 6 & $<$0.001 \\ 
-%  &Country & 13 & 24099.57 & 24204.72 & -12036.78 & 281.85 & 2 & $<$0.001 \\ 
-%  &Education & 11 & 23837.48 & 23926.34 & -11907.74 & 23.77 & 4 & $<$0.001 \\ 
-%  &Gender & 14 & 23821.88 & 23934.98 & -11896.94 & 2.17 & 1 & 0.140 \\ 
-%   \hline
-%\end{tabular}
-%\end{table}
-%
-
 \subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments~5,~6, and~7 separately.
@@
 There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.
-
-%The large $p$-values suggest that none of the levels of attempt ($\alpha_2$ through $\alpha_{10}$) are significant at \%1 significance level.
-Moreover, some of the estimates are positive and some are negative and they show up seemingly in random order suggesting later attempts not necessarily have improved.
-This indicates that there may be no learning effect on the probability of correct evaluations.
-
-% \begin{table}[htbp]
-% \centering
-% \caption{Parameter estimates of models \eqref{eqn:trend_response} fitted to data from three different experiments using detection rate as the response to assess learning trend.
-Attempt number is fitted as a factor to enable modeling any sort of non-linear learning trend.
-Only experiment~6 shows evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
-% \scalebox{0.81}{
-% \begin{tabular}{rrrrccrrrccrrrc}
-%   \hline
-% & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
-% 
-% \cline{2-5} \cline{7-10} \cline{12-15} 
-%  Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
-%   \hline
-% Fixed &  &  &  & && &  & & & & & &  & \\ 
-% $\mu$ & -1.304 & 0.179 & -7.28 & $<$0.001 &   & -0.220 & 0.147 & -1.50 & 0.134 &   & -1.737 & 0.481 & -3.61 & $<$0.001 \\ 
-%   $\alpha_1$ & 0.000 & --- & --- & --- &   & 0.000 & --- & --- & --- &  & 0.000 & --- & --- & ---  \\ 
-%   $\alpha_2$ & 0.270 & 0.219 & 1.24 & 0.217 &   & 0.262 & 0.158 & 1.66 & 0.098 &   & -0.456 & 0.385 & -1.18 & 0.237 \\ 
-%   $\alpha_3$ & -0.178 & 0.226 & -0.79 & 0.432 &   & 0.342 & 0.157 & 2.18 & 0.029 &   & -0.105 & 0.386 & -0.27 & 0.786 \\ 
-%   $\alpha_4$ & 0.083 & 0.224 & 0.37 & 0.712 &   & 0.358 & 0.159 & 2.26 & 0.024 &   & -0.378 & 0.381 & -0.99 & 0.322 \\ 
-%   $\alpha_5$ & 0.298 & 0.224 & 1.33 & 0.183 &   & 0.376 & 0.159 & 2.36 & 0.018 &   & -0.107 & 0.385 & -0.28 & 0.781 \\ 
-%   $\alpha_6$ & 0.042 & 0.231 & 0.18 & 0.857 &   & 0.246 & 0.158 & 1.56 & 0.120 &   & 0.026 & 0.407 & 0.06 & 0.949 \\ 
-%   $\alpha_7$ & 0.283 & 0.230 & 1.23 & 0.217 &   & 0.160 & 0.159 & 1.01 & 0.314 &   & 0.057 & 0.401 & 0.14 & 0.886 \\ 
-%   $\alpha_8$ & -0.045 & 0.233 & -0.19 & 0.847 &   & 0.341 & 0.160 & 2.13 & 0.033 &   & -0.003 & 0.394 & -0.01 & 0.994 \\ 
-%   $\alpha_9$ & -0.195 & 0.232 & -0.84 & 0.400 &   & 0.378 & 0.160 & 2.36 & 0.018 &   & 0.204 & 0.436 & 0.47 & 0.639 \\ 
-%   $\alpha_{10}$ & 0.513 & 0.228 & 2.25 & 0.024 &   & 0.192 & 0.163 & 1.18 & 0.238 &   & -0.213 & 0.432 & -0.49 & 0.622 \\ 
-% \multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
-%   $\sigma^2_a$ &  0.017 &&  &  &   &  0.034 &&  &  &   &  0.163 &&  &  \\ 
-%   $\sigma^2_u$  & 0.848 &&  &  &    & 0.903 &&  &  &   &  $<$0.001 &&  &  \\ 
-%   $\sigma^2_{\ell}$  & 1.476 &&  &  &    & 1.418 &&  &  &   &  3.314 &&  &  \\ 
-%    \hline
-% \end{tabular}
-% }
-% \label{tbl:model_result_response}
-% \end{table}
 
@@
 
-%The variance of random slope for attempt ($\sigma^2_a$) is estimated to be very small compared to other random effects except for subject variability ($\sigma^2_u$) for experiment~7.
-This suggests that most of the variability for attempt is accounted for by the fixed effect factor attempt.
-For all three experiments the lineup variabilities ($\sigma^2_l$) were estimated to be very large making the practical impact of other factor levels even more negligible.
-Some of the lineups were very easy and some were really difficult in experiment~7.
-The overall average probability for picking out the data plot from a lineup ($\mu$) is significant for both experiments 5 and 7.
-But for experiment~6 it is not significant suggesting that experiment~6 lineups were difficult compared to the other experimental lineups.
-This shows a feature of the experimental designs where a mixture of difficult and easy lineups were included within the experiments as well as between the experiments.
-For easy lineups there may be small chance to improve performances.
-But for harder lineups, the scope to improve performances is large.
-Since for none of these experiments improvement in performances is observed, it is important in the sense that performances did not improve over attempts in both difficult and easy lineup situations.
-
-%The results are shown in Tables~\ref{tbl:model_result_response}~\ref{tbl:model_result_time}.
-
-%Model~\eqref{eqn:trend_time} is fitted to the data from experiment~5, 6, and 7 separately.
-For this the function $lmer()$ is used from R package $lme4$ by \cite{lme4:2011}.
-To obtain the $p$-values of fixed effect parameters estimates, normal approximation is used for $Z$ scores computed as the ratio of estimates to the estimated standard errors.
-The results are shown in Table~\ref{tbl:model_result_time}.
-Both fixed effect parameters of covariate Attempt ($\alpha_1$ and $\alpha$) are highly significant.
-
-%Each subject participating the experiment has seen multiple number of lineups for evaluations.
-Suppose $K$ be the number of lineups evaluated by a subject and $A_k$ be the $k$th attempt, $k=1,2, ..., K$.
-Subjects may have a chance for self learning and do better on lineups shown later in the sequence.
-They may learn from their previous mistake or they may learn about the features of the plots being evaluated as they progress.
-Thus if there is any learning trend that should be apparent in proportion correct over $A_k$.
-But to obtain this we have to adjust for lineup difficulty and individual performances.
-
 To visualize how  detection rates change over successive attempts, we fitted model~\eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
@@
    \centering
-%   \includegraphics[width=6.5in]{learning_trend.pdf} 
     \includegraphics[width=6.3in]{learning_trend_subject.pdf} 
@@
 \end{figure}
-
-%The averages of these residuals for each of the attempts are shown as dots in Figure~\ref{fig:learning_trend_response}.
-Least square linear regression lines are fitted through the points for each of the experiments.
-Positive slopes over the attempts are observed but none of them is statistically significant.
 
@@
 Later pages of the web site were identical  except for the lineup.
-%Thus in the later attempts an observer does not need to spend any time for reading instructions.
-The model reflects that fact.
-
-% \begin{table}[htbp]
-% \centering
-% \caption{Parameter estimates of model~\eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
-Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments 5, 6 and 7.
-}
-% \scalebox{0.81}{
-% \begin{tabular}{rrrrccrrrccrrrc}
-%   \hline
-% & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
-% 
-% \cline{2-5} \cline{7-10} \cline{12-15} 
-%  Effect& Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value && Est & SE & Zval & $p$-value \\ 
-%   \hline
-% Fixed &  &  &  & && &  & & & & & &  & \\ 
-% $\mu$ & 3.817 & 0.039 & 97.38 & $<$0.001 &   & 3.901 & 0.033 & 118.19 & $<$0.001 &   & 3.731 & 0.054 & 69.04 & $<$0.001 \\ 
-%   $\alpha_1$ & 0.326 & 0.035 & 9.35 & $<$0.001 &   & 0.335 & 0.029 & 11.40 & $<$0.001 &   & 0.280 & 0.050 & 5.63 & $<$0.001 \\ 
-%   $\alpha$ & -0.038 & 0.004 & -9.30 & $<$0.001 &   & -0.039 & 0.004 & -10.19 & $<$0.001 &   & -0.029 & 0.007 & -4.22 & $<$0.001 \\ 
-% \multicolumn{2}{l}{Random}   &  &  &  &  & & &&& & & &  & \\ 
-%   $\sigma^2_a$ &   0.027 &  &  &   &  & 0.045 &  &  &   &  & 0.049 &  &  \\ 
-%   $\sigma^2_u$ &   0.509 &  &  &   &  & 0.495 &  &  &   &  & 0.366 &  &  \\ 
-%   $\sigma^2_{\ell}$ &   0.091 &  &  &   &  & 0.199 &  &  &   &  & 0.235 &  &  \\ 
-%   $\sigma^2$ &   0.460 &  &  &   &  & 0.501 &  &  &   &  & 0.454 &  & &  \\ 
-%    \hline
-% \end{tabular}
-% }
-% \label{tbl:model_result_time}
-% \end{table}
 
@@
 
-%The lineup variance of experiment~7 is estimated as $\hat{\sigma}^2_{\ell} = 10.98$ from model~\eqref{eqn:trend_response}.
-But from model~\eqref{eqn:trend_time} it is estimated to be much smaller in all these experiments.
-This suggests that the harder lineups not necessarily took more time than easier lineups.
-The observers spent enough times to evaluate a easy lineup and for difficult lineup they might just give up at some point and provided the feedback.
-
-%Model~\eqref{eqn:trend_time} was compared to couple of other alternatives.
-A bigger model with attempt as a factor was considered but it was not significantly different with $p$ value 0.236.
-A reduced model without the first attempt was also considered but it was significantly different with $p$-value $<0.001$.
-
 To visualize how the time taken reduces over the successive attempts, we fitted model~\eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
@@
 
-%These results indicate that after working with the multiple tasks each observer became efficient which helped them finish the task faster in later attempts.
-Subject specific least square lines show variations in slopes and overall time taken.
-Even though overall skill improved, there are some subjects who took more times in later attempts.
-
-%To accomplish this we fit generalized mixed effect model~\eqref{eqn:trend_response} with lineups and subjects as random effect and no fixed effect.
-This gives the estimation of proportion correct responses  and we obtain residuals of the fitted model.
-The residuals give variations in proportion correct adjusted for lineup difficulties and individual skills.
-If there is any learning trend over different attempt $A_k$, the residuals should have that information since $A_k$ was not included as a covariate in the model.
-%At this point we focus on absolute residuals over different attempts to study the deviations due to attempts.
-An smaller absolute residual at attempt $A_k$ should indicate learning from all other attempts less than $k$.
-
-%A plot of residuals against the attempts $A_k$ should reveal the trend if there is any.
-Figure~\ref{fig:learning_trend} shows such plots for three experiments naming experiment~5, 6 and 7.
-For all the experiments we see some increasing trend of the residuals.
-This could be attributed to the exhaustion of subjects' attempt after some earlier evaluations are made.
-But none of apparent trend are statistically significant.
-
-%Learning trend is not only related to the proportion correct responses.
-It may be possible that people learn to answer fast.
-We examine this by the time taken for each evaluation.
-We fit a mixed effect linear model with time taken considering lineup and subjects as random effect.
-The resulting residuals are plotted in Figure~\ref{fig:learning_trend_time}.
-It appears that in the later attempts observers take less time to evaluate a lineup.
-Even though it does not tell much about improving their performance over attempts, but it does indicate the improvement of their evaluation skill.
-Adding this information with what we observe in Figure~\ref{fig:learning_trend}, we can say that the observers' improved skills could not contribute to their performance.
-It only help them finish the task faster in later attempts.
-
 \begin{figure}[htbp] 
@@
 
-%This may happen for some sets of null plots if one or more null plots are very similar to the actual plot.
-In another set of null plots this may not happen making some lineups easier than others even though the actual data plot is the same.
-This pattern is evident in the Figure~as we see proportion correct for null plot 5 is consistently above the null plot 1 for each of the locations.
-
 \begin{figure}[htbp] 
@@
 \end{figure}
-
-%The sizes of the dots in  Figure~\ref{fig:location_effect} represents the number of responses obtained for that location and null set.
-For some locations we have as many as 10 responses.
-For location 1, we did not have any response for null set 1 in one of the interaction lineups.
-We observe larger variability for interaction lineups as compared to Genotype lineups.
-It is because some of the interaction lineups were evaluated only once as we see for null plot 3 and 5 for Interaction effect.
-This leads to the extreme proportion of correct since one response will either be correct or wrong.
-The larger variability observed for Interaction lineups can be controlled at some extent by evaluating them multiple times.
 
@@
 For the second approach in investigating the location effect, we distinguish between inner panels 7, 8, 9, 12, 13, 14 and outer panels (any panel with an outside edge).
-%We also examined whether the proportion of correct responses differs if the actual plot is on the outer boundary or in the inner locations.
-The locations 7,8,9,12,13,14 in the lineup are considered inner locations and the rest of the boundary locations are considered to be outer location.
 From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype.
 Model~\eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.
-%It does not show any differences whether the actual plot is inside or outer border of the lineup.
-
-%
-%% latex table generated in R 2.15.0 by xtable 1.7-0 package
-%% Fri Sep  7 10:28:47 2012
-%\begin{table}[hbtp]
-%\caption{Fixed effect parameter estimates of generalized mixed model.
-Note that attempt is not significant for experiment~2.
-The continuous covariate, lineup difficulty, was measured by the $p$-value of the actual plot in the lineup.}
-%\begin{center}
-%\begin{tabular}{llrrrrl}
-%  \hline
-%& Parameters & Estimate & Std..Error & z.value & P-value  & \\ 
-%  \hline
-%\multicolumn{2}{l}{\bf{Experiment~1} } &&&& &\\
-%&(Intercept) & -0.30 & 0.10 & -2.89 & 0.00 & ***\\ 
-% & attempt & 0.08 & 0.02 & 4.21 & 0.00 & ***\\ 
-% & lineup difficulty & -11.32 & 1.01 & -11.19 & 0.00 & ***\\ 
-%\multicolumn{2}{l}{\bf{Experiment~2} } &&&&& \\
-% & (Intercept) & 2.36 & 0.15 & 16.09 & 0.00 & ***\\ 
-% & attempt & 0.01 & 0.03 & 0.35 & 0.72 \\ 
-% & lineup difficulty & -55.03 & 2.20 & -25.00 & 0.00 & ***\\ 
-%\multicolumn{2}{l}{\bf{Experiment~3} } &&&& \\
-% & (Intercept) & 0.25 & 0.16 & 1.58 & 0.11 & .\\ 
-% & attempt & 0.10 & 0.03 & 3.34 & 0.00 & ***\\ 
-% & lineup difficulty & 3.00 & 0.56 & 5.36 & 0.00 & ***\\ 
-%   \hline
-%\end{tabular}
-%\end{center}
-%\label{tbl:model_par}
-%\end{table}
-%
-% \footnote{Signif. codes: 0 ?***? 0.001 ?**? 0.01 ?*? 0.05 ?.? 0.1}
-
+
+% ===================================================================================
 \section{Conclusion}
+%------------------------------------------------------------------------------------
 
@@
 
-%in time taken to evaluate a lineup.
-Some variations among the factors are observed in terms of probability of correct evaluation.
-Age group 36-40, Countries other than India and United states, People who have a graduate degree are significantly different.
-But their practical impact on probability to correctly evaluate a lineup is very negligible and in some cases it diminishes  as the lineup difficulties increase or decrease.
-Gender does not have any significant effect on performance.
-Thus there may be differences in time taken to evaluate a lineup for different human demographics but the practical impact of demographics on the performance is very negligible.
-This result is very important for the power of visual test as it demonstrates the robustness of the test for different human factors.
-
 Individual learning trend is observed in time taken but not so much in observer performance.
@@
 
-%But the overall performance of the observers does not increase through successive attempts while evaluating multiple lineups.
-This result suggests that the power estimated for visual inference using human subject experiment~is robust and may not change if those participants are allowed to give feedback again.
-The skill in evaluating the lineup in shorter time gets improved over successive evaluations.
-The earlier evaluations take significantly longer time than the later evaluations.
-It suggests that the skilled person may only do it faster.
-
 The simulation experiment reveals that there is no significant effect of location on the actual data plot in the lineup.
@@
 
-%The subjects participating this study may not necessarily know about statistical graphics.
-The numerous pilot studies done with more trained participants suggest that power of visual statical inference may be higher for observers who have advanced knowledge about statistical graphics.
-The fact that a person with graduate degree performs better may be associated with this.
-But a graduate degree does not necessarily mean to have training on statistical graphics.
-More experiments may be needed to learn about the differences in power with trained and a non-trained observer in terms of knowledge about statistical graphics.
-
-%This investigation allowed each observer to evaluate 10 lineups assuming that it would not cause fatigue or disinterest toward the task of evaluating lineups.
-This assumption is made based on the pilot studies.
-The future research may involve doing more experiments to check if fewer or more than 10 lineups have any significant impact on performance of the observer.
-The learning trend in terms of time taken shows downward trends.
-At some point it should level of.
-It will be useful to know for how many lineups time taken levels off to decide how many lineups is appropriate to show.
-
-%A lineup with fewer or larger than 20 plots may yield different results.
-If the size of the lineup is much higher than 20, it is intuitive that there may be location effect of the data plot in the lineup.
-It is because, the observer may get tired of scanning and may make decision based on the partial scanning of the lineup.
-On the other hand fewer than 20 plots will allow observer to compare the actual plot with fewer null plots giving more chance of picking actual plot as an error not as an actual plot.
-These are the some of the issues that require further investigation.
-
 There are many more ways that the lineup protocol might be tested to learn what we might expect about its performance for tackling real data mining problems.
@@
 
+\if1\blindDoc
+{
 %------------------------------------------------------------------------------------
@@
 This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
+} \fi
 

### 2024-09-16 wiley_paper/wiley_paper.Rnw
Add files via upload

--- 2017-10-21 submitted-v1/sociology_paper.Rnw
@@
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
 
@@
 \newcommand{\mm}[1]{{\color{mahbub} #1}}
+\renewcommand*\familydefault{\sfdefault} % To make font Halvatica, Wiley requirement
 
@@
 {
-  \title{Human Factors Influencing Visual Statistical Inference}
+  \title{Effect of Sociological Factors on Visual Statistical Inference}
   \author{{Mahbubul Majumder, Heike Hofmann, Dianne Cook}
-\thanks{Mahbubul Majumder is an Assistant Professor in the Department of Mathematics, University of Nebraska at Omaha, NE 68182 (e-mail: mmajumder@unomaha.edu), Heike Hofmann is Professor in the Department of Statistics and Statistical Laboratory, Iowa State University, Ames, IA 50011-1210, and  Dianne Cook is Professor in the Department of Econometrics and Business Statistics, Monash University, Australia.
+\thanks{Mahbubul Majumder is an Associate Professor in the Department of Mathematical and Statistical Sciences, University of Nebraska Omaha, NE 68182 (e-mail: mmajumder@unomaha.edu), Heike Hofmann is Professor in the Department of Statistics, University of Nebraska Lincoln, NE, and  Dianne Cook is Professor in the Department of Econometrics and Business Statistics, Monash University, Australia.
 This research is supported in part by the National Science Foundation Grant \# DMS 1007697.}}
@@
 
-Visual statistical inference is a way to determine significance of patterns found while  exploring data with graphics.
-It relies on human observers inspecting a lineup: one  data plot randomly placed among a set of decoy plots.
-Each individual is different in their cognitive psychology and judiciousness, which can affect the visual inference.
-The usual way of estimating the effectiveness of a statistical test is its power.
-The estimate of power of a lineup is controlled by combining evaluations from multiple observers.
-Factors that may  affect the power of visual inference are the observers' demographics, visual skills, and experience, the sample of null plots taken from the null distribution, the position of the data plot in the lineup, and the signal strength in the data.
-This paper examines these factors.
-Results from multiple visual inference studies using Amazon's Mechanical Turk are examined to provide an assessment of these.
-The experiments suggest that individual skills vary substantially,  but demographics do not have a huge effect on performance.
-There is evidence that a learning effect exists but only in the sense that observers become faster with repeated evaluations, while accuracy stays the same.
-The placement of data plot in the lineup does not affect inferential results.
+Visual statistical inference determines the significance of patterns found in data exploration through graphics.
+It involves human observers inspecting a lineup of plots, with one real data plot randomly placed among decoys.
+Each observer's cognitive skills and judiciousness can influence results.
+The effectiveness of this method, measured by power, depends on combining evaluations from multiple observers.
+Factors influencing power include observer demographics, visual skills, experience, the sample of null plots, plot position, and signal strength.
+This paper examines these factors through studies using Amazon's Mechanical Turk, finding individual skills vary but demographics have little impact.
+Learning increases speed but not accuracy, and plot position doesn’t affect results.
 
@@
 
-The lineup protocol introduced in \citet{buja:2009} can be used to quantify the significance of graphical findings in the exploratory data analysis process.
+%The lineup protocol introduced in \citet{buja:2009} can be used to quantify the significance of graphical findings in the exploratory data analysis process.
 This methodology is  part of what is called visual statistical inference, which has been developed further  and validated by \citet{majumder:2013} in simulation studies of head to head comparisons with conventional inference.
@@
 
-In visual inference, the test statistic is a plot of the observed data, called the {\it (actual) data plot}.
+%In visual inference, the test statistic is a plot of the observed data, called the {\it (actual) data plot}.
 To create a lineup the data plot is placed randomly in a grid of null plots.
@@
 
-\begin{figure}[hbtp] 
-   \centering
-   \includegraphics[width=0.99\textwidth]{electoral-5-13.pdf} 
-   \caption{ \label{fig:elect} Which one of the plots, labelled 1 through 20 from top left to bottom right, is the most different from the others?}
-\end{figure}
-
-Figure~\ref{fig:elect} displays an example lineup, one of the plots is based on the observed data, while the remaining 19 plots show  data generated from a null model.
+The lineup protocol introduced by \citet{buja:2009} quantifies the significance of graphical findings in exploratory data analysis, forming part of visual statistical inference.
+\citet{majumder:2013} extended and validated this approach through simulation studies, showing that visual inference can sometimes outperform conventional methods.
+They also defined the power of visual tests and proposed a way to calculate it for a given lineup.
+In visual inference, the test statistic is a plot of observed data, randomly placed among null plots generated under a null hypothesis, often assuming no structure.
+Observers evaluate the lineup, and if they consistently identify the data plot, this provides evidence against the null hypothesis, akin to a conventional hypothesis test.
+
+% \begin{figure}[hbtp] 
+%    \centering
+%    \includegraphics[width=0.99\textwidth]{electoral-5-13.pdf} 
+%    \caption{ \label{fig:elect} Which one of the plots, labelled 1 through 20 from top left to bottom right, is the most different from the others?}
+% \end{figure}
+
+%Figure~\ref{fig:elect} displays an example lineup, one of the plots is based on the observed data, while the remaining 19 plots show  data generated from a null model.
 Which one of the 20 plots is the most different from the others?
-When asked this question, 12 out of 72 observers picked the  data plot located in panel number $4^2-3$ in the lineup\footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
+%When asked this question, 12 out of 72 observers picked the  data plot located in panel number $4^2-3$ in the lineup\footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
 The corresponding $p$-value is 0.0077, indicating sufficient evidence to reject the null hypothesis.
 
-How do we interpret this finding, though? For that, we need to know the context of the data and we need to have more information about how the null plots were generated. 
-This particular example investigates the results from the 2012 US presidential election in comparison to  poll results just prior to the election. 
-Although this example is more simplistic than most of the tests conducted to date, it serves the purpose of illustrating the lineup protocol.
+%How do we interpret this finding, though? For that, we need to know the context of the data and we need to have more information about how the null plots were generated. 
+%This particular example investigates the results from the 2012 US presidential election in comparison to  poll results just prior to the election. 
+%Although this example is more simplistic than most of the tests conducted to date, it serves the purpose of illustrating the lineup protocol.
 The data consists of the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
@@
 % normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
-These samples are plotted as electoral buildings, and the plot based on the actual election results is placed randomly among the null plots in a lineup of size $5\times 4$.
+%These samples are plotted as electoral buildings, and the plot based on the actual election results is placed randomly among the null plots in a lineup of size $5\times 4$.
 If the null hypothesis is true, the  data plot should look  like any of the other plots, and not be identifiable by an observer.
@@
 
-\begin{figure}[hbtp] 
-   \centering
-   \begin{minipage}[c]{0.49\textwidth}
-   \includegraphics[width=\textwidth]{tower.pdf}
-   \end{minipage}
-    \hfill
-   \begin{minipage}[c]{0.45\textwidth}
-   Poll aggregator 1 \\
-      \includegraphics[width=\textwidth]{freedom-sim.pdf} 
-   Poll aggregator 2\\
-      \includegraphics[width=\textwidth]{rcp-sim.pdf} 
-   \end{minipage}
-   \caption{ \label{fig:tower} Electoral building plot of the results of the 2012 U.S.
+% \begin{figure}[hbtp] 
+%    \centering
+%    \begin{minipage}[c]{0.49\textwidth}
+%    \includegraphics[width=\textwidth]{tower.pdf}
+%    \end{minipage}
+%     \hfill
+%    \begin{minipage}[c]{0.45\textwidth}
+%    Poll aggregator 1 \\
+%       \includegraphics[width=\textwidth]{freedom-sim.pdf} 
+%    Poll aggregator 2\\
+%       \includegraphics[width=\textwidth]{rcp-sim.pdf} 
+%    \end{minipage}
+%    \caption{ \label{fig:tower} Electoral building plot of the results of the 2012 U.S.
 Presidential Election (left).
@@
 The $p$-value from the lineup protocol, was 0.0077, which more strongly concludes that these polls mismatched the actual result.}
-\end{figure}
-The lineup of Figure~\ref{fig:elect}, as well as the other lineups in the manuscript are created using the nullabor package by \citep{nullabor}.
-
-A lineup can be evaluated by a single person or multiple observers.
+%\end{figure}
+%The lineup of Figure~\ref{fig:elect}, as well as the other lineups in the manuscript are created using the nullabor package by \citep{nullabor}.
+
+%A lineup can be evaluated by a single person or multiple observers.
 A distribution similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario, is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis \citep{vinference}.
@@
 
-When presenting a lineup to an observer the accompanying question should be phrased in as general a manner as possible, effectively asking the observer to pick the plot that is most different, and allowing observers to provide their own reasons for their choice.
+%When presenting a lineup to an observer the accompanying question should be phrased in as general a manner as possible, effectively asking the observer to pick the plot that is most different, and allowing observers to provide their own reasons for their choice.
 This ensures all possible deviations from the null hypothesis to be detected.
@@
 
-In the election example, observers were asked ``which plot is the most different?''.
+A lineup can be evaluated by one or more observers, and a distribution similar to the binomial, adjusted for dependencies from the lineup scenario, is used to calculate the $p$-value based on how often the actual data plot is identified \citep{vinference}.
+To avoid expectation bias \citep{meilgaard}, especially in sensitive areas like election results, observers must not know the data constituting the lineup or have seen the actual data plot beforehand.
+The context of the election problem was revealed only after lineup evaluation to prevent bias.
+The accompanying question in a lineup should be broad, asking observers to pick the most distinct plot, allowing detection of any deviations from the null hypothesis.
+However, in head-to-head tests with conventional methods \citep{majumder:2013, tengfei:2013}, specific questions are needed to avoid type III errors \citep{mosteller:48} consisting of correctly rejecting the null hypothesis for a wrong reason.
+
+%In the election example, observers were asked ``which plot is the most different?''.
 The type of plot, showing two (modified) stacked bar charts in different colors should suggest to the observer that the interesting feature is the difference between the two heights.
@@
 
-\begin{table*}[hbtp] 
-\centering 
-\caption{Overview of 10 different Turk experiments, from which data was collected to study human factor effects.
-All of the experimental data were used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.}
-\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
-\hline\hline 
- & Experiment &  Test Statistic  & Lineup question & Used in study of\\ [0.5ex] % inserts table %heading 
-\hline 
-1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
-between group A and B? & DF\\
-2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
-& DF\\
-  3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
-& DF\\
- 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
-  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
-  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?
-&  DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
-  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?
-& DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
-  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?
-& DF\\
-  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?
-& DF  \hfill \phantom{LT} \hfill {LE} \hfill \phantom{.}\\
-  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?
-& DF\\
-
-\hline 
-\end{tabular} 
-\label{tbl:visual_stat} 
-\end{table*}
-
-We have conducted a series of experiments (see Table~\ref{tbl:visual_stat} for an overview)
-%There have been ten experiments conducted 
-using Amazon's Mechanical Turk~\citep{turk} (MTurk) for a variety to purposes: validating the protocol against existing tests, comparing plot designs, and evaluating structure in data analysis problems.
+%We have conducted a series of experiments (see Table~\ref{tbl:visual_stat} for an overview) using Amazon's Mechanical Turk~\citep{turk} (MTurk) for a variety to purposes: validating the protocol against existing tests, comparing plot designs, and evaluating structure in data analysis problems.
 In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.
-Mining this data will provide a basis for evaluating
-%are being used to evaluate 
-the effects of different covariates on an individual's ability to evaluate lineups.
+Mining this data will provide a basis for evaluating the effects of different covariates on an individual's ability to evaluate lineups.
 The design of some of the experiments allow a further investigation into aspects beyond participants' demographics: experiments~5, 6 and 7 can be used to investigate short term learning trends, while experiment~9 allows insight on whether the  positioning of the  data plot in the lineup has an effect on detection rate.
- %Each of these experiments collected demographic details of the subjects. 
-Section~\ref{sec:factor_performance} discusses human factors that we expect to be influential on the performance of the observer.
+
+We conducted several experiments (see Table~\ref{tbl:visual_stat}) using Amazon's Mechanical Turk~\citep{turk} (MTurk) for validating the lineup protocol, comparing plot designs, and evaluating data analysis structures.
+Demographic data such as gender, age, and educational background were collected to assess the influence of these covariates on lineup evaluations.
+Additionally, experiments investigated short-term learning trends as well as explored the effect of data plot positioning on detection rates.
+
+%Section~\ref{sec:factor_performance} discusses human factors that we expect to be influential on the performance of the observer.
 Section~\ref{sec:exp_design} describes the methods used to assess the effects, and Section~\ref{sec:result_socio} summarizes the results of what we have learned about human factors affecting lineup evaluation.
+
+Section~\ref{sec:factor_performance} discusses expected human factors influencing observer performance and Section~\ref{sec:result_socio} presents findings on factors affecting lineup evaluations.
+The methods used to assess the effects are oulined in Appendix~\ref{sec:exp_design}.
 
@@
 \begin{itemize} \itemsep 0in \parsep-10pt %\parskip0pt
-%\subsection{Choice of Visual Test Statistic}  
-\item{\bf Choice of Visual Test Statistic:}  
-%There are a variety of ways that
-Data can be plotted in a variety of ways and we can enhance plots by axes, legends or representations of models.
+
+%\item{\bf Choice of Visual Test Statistic:}  
+%Data can be plotted in a variety of ways and we can enhance plots by axes, legends or representations of models.
 The choice of plot primarily should be appropriate for the problem being investigated, as the data plot is the visual test statistic.
-When for example studying the association between two quantitative variables a scatterplot is usually considered best practice.
+%When for example studying the association between two quantitative variables a scatterplot is usually considered best practice.
 Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.
@@
 
-Table~\ref{tbl:visual_stat} shows the large variety of visual test statistics used in the Turk studies completed so far.
+%Table~\ref{tbl:visual_stat} shows the large variety of visual test statistics used in the Turk studies completed so far.
 These experiments were conducted to examine many different visual tasks.
@@
 
-In experiments~4,~5 and~6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices, the use of lineup tests for comparisons and  examples are discussed in \citet{heike:2012, qqplots:2015}
-
-%\subsection{Signal in the Data}
-\item{\bf Signal in the Data:}
-A visual test statistic is chosen for a particular task.
+%In experiments~4,~5 and~6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices, the use of lineup tests for comparisons and  examples are discussed in \citet{heike:2012, qqplots:2015}
+
+%\item{\bf Signal in the Data:}
+%A visual test statistic is chosen for a particular task.
 If it is well-designed  departures from the null hypothesis will be visible in the plot, especially if the signal is strong.
@@
 
-\item{\bf Question Design:} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
+%\item{\bf Question Design:} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
 Visual inference supports discovery of structure, enabling the unexpected to surprise us.
@@
 
-In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted.
+%In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted.
 This was important to allow a direct comparison of the results to those from conventional testing.
 %in order to make a comparison of the results.
-For latter studies (4, 7, 10), questions posed to the observer were more general. Table~\ref{tbl:visual_stat} lists the questions used for each experiment.
-
-%\subsection{Demographics} 
+%For latter studies (4, 7, 10), questions posed to the observer were more general. Table~\ref{tbl:visual_stat} lists the questions used for each experiment.
+
 \item{\bf Demographics:}  During each of the experiments, data on age, gender, education level and geographic location was collected.
@@
 
-% ===================================================================================
-\section{Experimental Methods}\label{sec:exp_design}
-%------------------------------------------------------------------------------------
-
-Two of the factors, signal in the data and individual abilities, were studied in \cite{majumder:2013}.
+% 
+% % ===================================================================================
+% \section{Experimental Methods}\label{sec:exp_design}
+% %------------------------------------------------------------------------------------
+% 
+% Two of the factors, signal in the data and individual abilities, were studied in \cite{majumder:2013}.
 The choice of visual test statistic was examined in \cite{heike:2012}.
@@
 This section describes the experimental methods used to examine the effects of demography, placement of the actual data plot, sample of null plots and the existence of a learning trend.
-
-\subsection{Demographic Factors}
-\noindent
-For all  ten experiments shown in Table~\ref{tbl:visual_stat}, the following demographic information was collected from subjects:
-
-\begin{enumerate} \leftmargin 5cm  \itemsep 0in
-\item {\it Age group}, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, above 50.
-\item {\it Gender}, male or female.
-\item {\it Education level}, with levels being high school or less, some undergraduate courses, undergraduate degree, some graduate courses, and graduate degree.
-\item {\it Geographical location}, collected from the IP address of the participants' computer, as latitude, longitude, city and country. 
-\end{enumerate}
-
-Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
+% 
+% \subsection{Demographic Factors}
+% \noindent
+% For all  ten experiments shown in Table~\ref{tbl:visual_stat}, the following demographic information was collected from subjects:
+% 
+% \begin{enumerate} \leftmargin 5cm  \itemsep 0in
+% \item {\it Age group}, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, above 50.
+% \item {\it Gender}, male or female.
+% \item {\it Education level}, with levels being high school or less, some undergraduate courses, undergraduate degree, some graduate courses, and graduate degree.
+% \item {\it Geographical location}, collected from the IP address of the participants' computer, as latitude, longitude, city and country. 
+% \end{enumerate}
+% 
+% 
+% Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
 The factors are examined in association with the observer's response using a logistic regression with random effect terms:
-
-\begin{equation} \label{eqn:demographic_response}
-g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
-\end{equation}
-
-where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ picks the actual data plot from lineup $j$, $\mu$ is an overall population average, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively, for observer $i$.
+% 
+% \begin{equation} \label{eqn:demographic_response}
+% g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
+% \end{equation}
+% 
+% where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ picks the actual data plot from lineup $j$, $\mu$ is an overall population average, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively, for observer $i$.
 The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume independence and normality of the errors, i.e.\ $\ell_j \sim N(0, \sigma_\ell^2)$.
 $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-
-Similarly, we model the  time  an observer takes to identify a panel from a lineup.
+% 
+% 
+% Similarly, we model the  time  an observer takes to identify a panel from a lineup.
 Let $Z_{ij}$ denote the logarithm of time taken for  observer $i$ to evaluate  lineup $j$.
@@
 We model this in a mixed effects model of the same structure as model~(\ref{eqn:demographic_response}) given as:
-%\hh{
-\begin{equation} \label{eqn:demographic_time}
-Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
-\end{equation}
-where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model~\eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.
-
-\subsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive responses when multiple lineups are shown for evaluation.
+% %\hh{
+% \begin{equation} \label{eqn:demographic_time}
+% Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
+% \end{equation}
+% where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model~\eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.
+% 
+% \subsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive responses when multiple lineups are shown for evaluation.
 Experiments 5, 6 and 7 were used for this.
@@
 The goal is to estimate whether performance of the observer improves, or changes, from attempt 1 to attempt 10.
-
-It should be noted that we are examining the observer's performance, when we model response as detected or not, but this is not the goal of visual inference.
+% 
+% It should be noted that we are examining the observer's performance, when we model response as detected or not, but this is not the goal of visual inference.
 Visual inference is constructed to measure the significance of structure discovered in data.
@@
 So in this context, detected, or not, is used as a response to examine individual differences.
-
-Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k$th evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
+% 
+% Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k$th evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
 Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the actual data plot from lineup $j$ in their $k$th attempt.
 Learning trend is assessed using a generalized mixed effects model of the form
-\begin{equation} \label{eqn:trend_response}
-g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
-\end{equation}
-where $\mu$ is an overall population average, $\alpha_k$ is the effect of the $k$th attempt on the probability, using the first attempt as reference, $\alpha_1 = 0$, and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
+% \begin{equation} \label{eqn:trend_response}
+% g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
+% \end{equation}
+% where $\mu$ is an overall population average, $\alpha_k$ is the effect of the $k$th attempt on the probability, using the first attempt as reference, $\alpha_1 = 0$, and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
 The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
-The term $a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts, where $a_i \sim N(0, \sigma_a^2)$. 
-For $\ell_j$ a normal distribution, $N(0, \sigma_\ell^2)$, is assumed, and $\ell_j$ is a random intercept predicting lineup difficulty level.
+% The term $a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts, where $a_i \sim N(0, \sigma_a^2)$. 
+% For $\ell_j$ a normal distribution, $N(0, \sigma_\ell^2)$, is assumed, and $\ell_j$ is a random intercept predicting lineup difficulty level.
 $g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
 The inverse link function, $g^{-1}(.)$, from equation~\ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in the $k$th attempt by a single observer as
-\begin{equation} \label{eqn:trend_power}
-\hat p_{ijk} =  g^{-1}(\hat{\mu} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
-\end{equation}
-
-When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
+% \begin{equation} \label{eqn:trend_power}
+% \hat p_{ijk} =  g^{-1}(\hat{\mu} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
+% \end{equation}
+% 
+% When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
 Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of the (log) of time taken by  observer $i$ to choosing a  panel from lineup $j$ in his/her $k$th attempt.
 We evaluate this in a mixed effects model of the form
-
-\begin{equation} \label{eqn:trend_time}
-Z_{ijk} = \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
-\end{equation}
-where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$ is the average change in log time taken for each additional attempt,  $\alpha_1$ is an offset in log time taken for the first attempt.
+% 
+% \begin{equation} \label{eqn:trend_time}
+% Z_{ijk} = \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
+% \end{equation}
+% where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$ is the average change in log time taken for each additional attempt,  $\alpha_1$ is an offset in log time taken for the first attempt.
 All other effects are random effects: $u_i$ is a subject-specific intercept representing individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$, $a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$, with $a_i \sim N(0, \sigma_a^2)$, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-Equation~\ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
-\begin{equation} \label{eqn:trend_time_est}
-\hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j.
-\end{equation}
-
-To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2015, lme4:paper}.
+% Equation~\ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
+% \begin{equation} \label{eqn:trend_time_est}
+% \hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j.
+% \end{equation}
+% 
+% To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2015, lme4:paper}.
 We employ a normal approximation to obtain $p$-values corresponding to fixed effect parameters estimates.
-
-\subsection{Location Effect} \label{sec:location_design} Experiment~9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
+% 
+% \subsection{Location Effect} \label{sec:location_design} Experiment~9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
 The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
 Two factors were of primary interest a main effect for genotype and an interaction effect between genotype and treatment condition.
-
-In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
+% 
+% In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
 This overall significance is studied using visual inference  in \citet{tengfei:2013}.
-
-In order to study the effect that  location has on evaluating lineups, we used multiple lineups for each of the data plots.
+% 
+% In order to study the effect that  location has on evaluating lineups, we used multiple lineups for each of the data plots.
 For each data plot, five sets of null plots were generated and the actual plot was randomly placed in one of five different locations in a lineup of size 20.
-For the genotype effect, the locations were  1, 8, 12, 17, 20 and locations 2, 9, 12, 16, 20 were used for the interaction effect.
+% For the genotype effect, the locations were  1, 8, 12, 17, 20 and locations 2, 9, 12, 16, 20 were used for the interaction effect.
 %Five different sets of null plots were used to produce 5 lineups for each location position, creating
-Overall this created a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
+% Overall this created a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
 Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to help clean the data.
-
-To examine if the difference in detection rate among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
-Let $\mathbf{Y}=(Y_1,Y_2, ...
+% 
+% To examine if the difference in detection rate among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
+% Let $\mathbf{Y}=(Y_1,Y_2, ...
 , Y_p)^\top$ be a vector of random variables with dimension $p = 5$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represent the $j$th vector response for  location $i$ with $i=1,2, ..., I$ for $U=5$.
@@
 The MANOVA model
-\begin{equation}\label{manova}
-\mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
-\end{equation}
-where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$, tests for significant difference between the means.
-
-\subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through MTurk \citep{turk}.
+% \begin{equation}\label{manova}
+% \mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
+% \end{equation}
+% where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$, tests for significant difference between the means. 
+% 
+% \subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through MTurk \citep{turk}.
 It is an online work place where people from around the world can sign up for so-called `HIT's, human intelligence tasks, generally short tasks that are humans are typically better at solving than computers.
@@
 %Each task is usually planned to completed in a short time.
- For completing a HIT workers are paid a small amount of money, on the order of minimum wage in the USA.
-
-We designed and developed a web application which enables the display of lineups to  observers as per experimental need.
+%  For completing a HIT workers are paid a small amount of money, on the order of minimum wage in the USA. 
+% 
+% 
+% We designed and developed a web application which enables the display of lineups to  observers as per experimental need.
 The MTurk workers were re-directed to this web application to complete their assigned tasks.
@@
 Location of an observer is determined based on the ip address of the observer.
+%
 
@@
 
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=4.5in]{turker_country.pdf} 
-   \includegraphics[width=4.5in]{turker_location.pdf}    
-   \caption{Location of Amazon Mechanical Turk workers participating in the experiments. Most subjects came from India and the United States, but subjects from countries around the world took part.}
-   \label{fig:turker_location}
-\end{figure}
+% \begin{figure}[htbp] 
+%    \centering
+%    \includegraphics[width=4.5in]{turker_country.pdf} 
+%    \includegraphics[width=4.5in]{turker_location.pdf}    
+%    \caption{Location of Amazon Mechanical Turk workers participating in the experiments. Most subjects came from India and the United States, but subjects from countries around the world took part.}
+%    \label{fig:turker_location}
+% \end{figure}
 
@@
 
-\begin{table}[hbtp]
-\caption{Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
-\centering
-\scalebox{.85}{
-\begin{tabular}{rlrrcr}
-  \hline
-  && \multicolumn{2}{c}{Participants}&Average& Number of\\
-  \cline{3-4}
- Factor & Levels & Total &\% & Time & Responses \\ 
-  \hline
-Gender & Female & 989 & 42.26 & 43.69 & 10538 \\ 
-   & Male & 1351 & 57.74 & 48.75 & 13542 \\  
-\hline
-  Education & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
-   & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
-   & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
-   & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
-   & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
-\hline
-  Age & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
-   & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
-   & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
-   & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
-   & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
-   & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
-   & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
-\hline   
-  Country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
-   & India & 980 & 41.88 & 52.63 & 10238 \\ 
-   & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\
-
-   \hline
-\end{tabular}
-}
-\label{tbl:demographics}
-\end{table}
-
-% 
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 16:56:32 2017
-% \begin{table}[ht]
+% \begin{table}[hbtp]
+% \caption{Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
 % \centering
-% \begin{tabular}{llrrrr}
-%   \hline
-% var & lbls & subsjects & percents & avg\_time & response \\ 
-%   \hline
-% gender\_level & Female & 989 & 42.26 & 43.69 & 10538 \\ 
-%   gender\_level & Male & 1351 & 57.74 & 48.75 & 13542 \\ 
-%   degree & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
-%   degree & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
-%   degree & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
-%   degree & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
-%   degree & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
-%   age\_level & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
-%   age\_level & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
-%   age\_level & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
-%   age\_level & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
-%   age\_level & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
-%   age\_level & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
-%   age\_level & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
-%   country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
-%   country & India & 980 & 41.88 & 52.63 & 10238 \\ 
-%   country & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\ 
+% \scalebox{.85}{
+% \begin{tabular}{rlrrcr}
+%   \hline
+%   && \multicolumn{2}{c}{Participants}&Average& Number of\\
+%   \cline{3-4}
+%  Factor & Levels & Total &\% & Time & Responses \\ 
+%   \hline
+% Gender & Female & 989 & 42.26 & 43.69 & 10538 \\ 
+%    & Male & 1351 & 57.74 & 48.75 & 13542 \\  
+% \hline
+%   Education & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
+%    & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
+%    & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
+%    & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
+%    & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
+% \hline
+%   Age & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
+%    & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
+%    & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
+%    & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
+%    & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
+%    & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
+%    & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
+% \hline   
+%   Country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
+%    & India & 980 & 41.88 & 52.63 & 10238 \\ 
+%    & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\ 
+%   
 %    \hline
 % \end{tabular}
+% }
+% \label{tbl:demographics}
 % \end{table}
+%
 
@@
 
-% latex table generated in R 3.4.1 by xtable 1.8-2 package
-% Wed Oct 18 17:43:02 2017
-\begin{table}[ht]
-\caption{\label{tbl:anova_factor}Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
+% 
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 17:43:02 2017
+% \begin{table}[ht]
+% \caption{\label{tbl:anova_factor}Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
 Gender does not have a significant effect on detection rate, but does on time to respond.
 All factors significantly affect time to respond.}
-\centering
-\scalebox{0.9}{
-\begin{tabular}{lrrrrlrrrr}
-& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
-\cline{2-5} \cline{7-10} 
- & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
-Full  & 52609.1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 24140.5 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
- - Age & 52081.3 & 527.8 & 6 & $<$0.001 & \hspace{0.5cm} & 24113.2 & 27.3 & 6 & $<$0.001 \\ 
- - Country & 52245.5 & 363.6 & 2 & $<$0.001 &  & 24114.8 & 25.7 & 2 & $<$0.001 \\ 
- - Degree & 52498.7 & 110.4 & 4 & $<$0.001 &  & 24116.0 & 24.5 & 4 & $<$0.001 \\ 
- - Gender & 52540.3 & 68.8 & 1 & $<$0.001 &  & 24138.9 & 1.6 & 1 & 0.200 \\ 
-\cline{2-5} \cline{7-10}
-\end{tabular}}
-\end{table}
+% \centering
+% \scalebox{0.9}{
+% \begin{tabular}{lrrrrlrrrr}
+% & \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
+% \cline{2-5} \cline{7-10} 
+%  & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
+% Full  & 52609.1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 24140.5 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
+%  - Age & 52081.3 & 527.8 & 6 & $<$0.001 & \hspace{0.5cm} & 24113.2 & 27.3 & 6 & $<$0.001 \\ 
+%  - Country & 52245.5 & 363.6 & 2 & $<$0.001 &  & 24114.8 & 25.7 & 2 & $<$0.001 \\ 
+%  - Degree & 52498.7 & 110.4 & 4 & $<$0.001 &  & 24116.0 & 24.5 & 4 & $<$0.001 \\ 
+%  - Gender & 52540.3 & 68.8 & 1 & $<$0.001 &  & 24138.9 & 1.6 & 1 & 0.200 \\ 
+% \cline{2-5} \cline{7-10}
+% \end{tabular}}
+% \end{table}
 
@@
 
-\subsection{Location Effect}
-
-A total of 111 subjects was recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
+% \subsection{Location Effect}
+% 
+% A total of 111 subjects was recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
 Each subject evaluated two lineups;  one for Interaction effect and the other for Genotype.
@@
 The data on the test lineup were excluded from the analysis.
-
-We are investigating location effect in two ways: (1) location effect of each individual plot location, and (2) location effect of inside/outside panel positions.
-
-Figure~\ref{fig:location_effect} shows an overview of detection rates for both of the effects studied in experiment~9.
+% 
+% We are investigating location effect in two ways: (1) location effect of each individual plot location, and (2) location effect of inside/outside panel positions.
+% 
+% Figure~\ref{fig:location_effect} shows an overview of detection rates for both of the effects studied in experiment~9.
 Each dot corresponds to one lineup.
@@
 Lineups with  the same set of null plots are connected by lines.
-%The detection rate for each data plot location is shown in colored by null sets. 
-The overall average detection rate for each location is shown using dashed lines.
+% %The detection rate for each data plot location is shown in colored by null sets. 
+% The overall average detection rate for each location is shown using dashed lines.
 Size of the points represent the number of responses.
@@
 For location 1, we did not have any responses for null set 1 in one of the interaction lineups.
-
-We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets.
+% 
+% We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets.
 The lineup protocol is using a finite, small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set.
@@
 The rest of the null sets do not show any differences at  a 1\% significance level.
-
-\begin{figure}[htbp] 
-   \centering
-    \includegraphics[width=6.5in]{proportion_nulls_guide.pdf} 
-   \caption{Location of data plot in the lineup and proportion correct (detection rate) for both Interaction and Genotype effect.
+% 
+% 
+% \begin{figure}[htbp] 
+%    \centering
+%     \includegraphics[width=6.5in]{proportion_nulls_guide.pdf} 
+%    \caption{Location of data plot in the lineup and proportion correct (detection rate) for both Interaction and Genotype effect.
 Each colored line represents a null set and the size of the dots represents number of responses.
@@
 The actual data plot locations are shaded grey on the top panels to demonstrate their relative positions on a lineup.}
-   \label{fig:location_effect}
-\end{figure}
-
-Model~\eqref{manova} was fit to the data to test, if  detection rates were significantly different for the different locations.
+%    \label{fig:location_effect}
+% \end{figure}
+% 
+% 
+% Model~\eqref{manova} was fit to the data to test, if  detection rates were significantly different for the different locations.
 For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R}.
@@
 The $p$-values for both the Interaction and the Genotype effect are much bigger than the conventional threshold of 0.05 suggesting that location of the actual data plot in the lineup does not significantly affect performance.
-
-\begin{table}[hbtp]
-\caption{The summary of the results obtained by fitting MANOVA model~\eqref{manova} using Wilks test.}
-\begin{center}
-\begin{tabular}{ccccccccc}
-  \hline \hline
- Location & && & \multicolumn{3}{c} {Degrees of Freedom}  & F test \\
- \cline{5-7}
- Effect & DF & Wilks & Approx. F & Numerator& Denominator &Residual & $p$-value\\
-  \hline
-  Interaction& 3&0.07986 & 0.592 &15& \phantom{0}5.92 & 6 & 0.8082 \\ 
-  Genotype & 4&0.01689 & 1.722 & 20&14.22 & 8 & 0.1488 \\ 
-   \hline
-\end{tabular}
-\end{center}
-\label{tbl:manova}
-\end{table}
-
-For the second approach in investigating the location effect, we distinguish between inner panels 7, 8, 9, 12, 13, 14 and outer panels (any panel with an outside edge).
+% 
+% \begin{table}[hbtp]
+% \caption{The summary of the results obtained by fitting MANOVA model~\eqref{manova} using Wilks test.}
+% \begin{center}
+% \begin{tabular}{ccccccccc}
+%   \hline \hline
+%  Location & && & \multicolumn{3}{c} {Degrees of Freedom}  & F test \\
+%  \cline{5-7}
+%  Effect & DF & Wilks & Approx. F & Numerator& Denominator &Residual & $p$-value\\
+%   \hline
+%   Interaction& 3&0.07986 & 0.592 &15& \phantom{0}5.92 & 6 & 0.8082 \\ 
+%   Genotype & 4&0.01689 & 1.722 & 20&14.22 & 8 & 0.1488 \\ 
+%    \hline
+% \end{tabular}
+% \end{center}
+% \label{tbl:manova}
+% \end{table}
+% 
+% For the second approach in investigating the location effect, we distinguish between inner panels 7, 8, 9, 12, 13, 14 and outer panels (any panel with an outside edge).
 From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype.
@@
 
+% ---------------------------------------- Appendix ----------------------------------
+
+\appendix
+
+% ===================================================================================
+\section{Experimental Methods}\label{sec:exp_design}
+%------------------------------------------------------------------------------------
+
+Two of the factors, signal in the data and individual abilities, were studied in \cite{majumder:2013}.
+The choice of visual test statistic was examined in \cite{heike:2012}.
+In each of these analyses demographic factors were given a cursory glance, to ensure that they did not have large effects on the results.
+The design of experiments 5, 6 and 7 enables the examination of learning trend, which is studied in this paper.
+Experiment~9 was a real test case for visual statistical inference, and in order to understand the significance of the structure in the genomic data, multiple lineups were made in which location of the actual data plot, and the sample of nulls, were randomized.
+This enables the assessment of the effect of these factors on the results.
+This section describes the experimental methods used to examine the effects of demography, placement of the actual data plot, sample of null plots and the existence of a learning trend.
+
+\subsection{Demographic Factors}
+\noindent
+For all  ten experiments shown in Table~\ref{tbl:visual_stat}, the following demographic information was collected from subjects:
+
+\begin{enumerate} \leftmargin 5cm  \itemsep 0in
+\item {\it Age group}, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, above 50.
+\item {\it Gender}, male or female.
+\item {\it Education level}, with levels being high school or less, some undergraduate courses, undergraduate degree, some graduate courses, and graduate degree.
+\item {\it Geographical location}, collected from the IP address of the participants' computer, as latitude, longitude, city and country. 
+\end{enumerate}
+
+Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
+The factors are examined in association with the observer's response using a logistic regression with random effect terms:
+
+\begin{equation} \label{eqn:demographic_response}
+g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
+\end{equation}
+
+where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ picks the actual data plot from lineup $j$, $\mu$ is an overall population average, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively, for observer $i$.
+The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume independence and normality of the errors, i.e.\ $\ell_j \sim N(0, \sigma_\ell^2)$.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+
+Similarly, we model the  time  an observer takes to identify a panel from a lineup.
+Let $Z_{ij}$ denote the logarithm of time taken for  observer $i$ to evaluate  lineup $j$.
+Let $\mu_{ij}=  E(Z_{ij})$ be the average of the (log) time taken by  observer $i$ to pick a panel from lineup $j$.
+We model this in a mixed effects model of the same structure as model~(\ref{eqn:demographic_response}) given as:
+%\hh{
+\begin{equation} \label{eqn:demographic_time}
+Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
+\end{equation}
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model~\eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.
+
+\subsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive responses when multiple lineups are shown for evaluation.
+Experiments 5, 6 and 7 were used for this.
+Each subject was shown a total of 10 lineups randomly selected from a pool of lineups.
+The lineups are not necessarily of the same difficulty level, but the order of lineups was randomized.
+The responses of the lineups were recorded by attempt 1 through 10.
+Attempt 1 means that the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
+The goal is to estimate whether performance of the observer improves, or changes, from attempt 1 to attempt 10.
+
+It should be noted that we are examining the observer's performance, when we model response as detected or not, but this is not the goal of visual inference.
+Visual inference is constructed to measure the significance of structure discovered in data.
+It is expected that some observers will be more skilled at reading data plots, and hence, more readily detect the plot that is different.
+It is also expected that as observers gain experience in evaluating lineups that they become more proficient in reading data plots, particularly if feedback is given on whether the actual data plot was chosen or not.
+Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups.
+So in this context, detected, or not, is used as a response to examine individual differences.
+
+Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k$th evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the actual data plot from lineup $j$ in their $k$th attempt.
+Learning trend is assessed using a generalized mixed effects model of the form
+\begin{equation} \label{eqn:trend_response}
+g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
+\end{equation}
+where $\mu$ is an overall population average, $\alpha_k$ is the effect of the $k$th attempt on the probability, using the first attempt as reference, $\alpha_1 = 0$, and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
+The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
+The term $a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts, where $a_i \sim N(0, \sigma_a^2)$. 
+For $\ell_j$ a normal distribution, $N(0, \sigma_\ell^2)$, is assumed, and $\ell_j$ is a random intercept predicting lineup difficulty level.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+The inverse link function, $g^{-1}(.)$, from equation~\ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in the $k$th attempt by a single observer as
+\begin{equation} \label{eqn:trend_power}
+\hat p_{ijk} =  g^{-1}(\hat{\mu} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
+\end{equation}
+
+When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
+Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of the (log) of time taken by  observer $i$ to choosing a  panel from lineup $j$ in his/her $k$th attempt.
+We evaluate this in a mixed effects model of the form
+
+\begin{equation} \label{eqn:trend_time}
+Z_{ijk} = \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
+\end{equation}
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$ is the average change in log time taken for each additional attempt,  $\alpha_1$ is an offset in log time taken for the first attempt.
+All other effects are random effects: $u_i$ is a subject-specific intercept representing individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$, $a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$, with $a_i \sim N(0, \sigma_a^2)$, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
+Equation~\ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
+\begin{equation} \label{eqn:trend_time_est}
+\hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j.
+\end{equation}
+
+To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2015, lme4:paper}.
+We employ a normal approximation to obtain $p$-values corresponding to fixed effect parameters estimates.
+
+\subsection{Location Effect} \label{sec:location_design} Experiment~9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
+The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
+Two factors were of primary interest a main effect for genotype and an interaction effect between genotype and treatment condition.
+
+In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
+This overall significance is studied using visual inference  in \citet{tengfei:2013}.
+
+In order to study the effect that  location has on evaluating lineups, we used multiple lineups for each of the data plots.
+For each data plot, five sets of null plots were generated and the actual plot was randomly placed in one of five different locations in a lineup of size 20.
+For the genotype effect, the locations were  1, 8, 12, 17, 20 and locations 2, 9, 12, 16, 20 were used for the interaction effect.
+%Five different sets of null plots were used to produce 5 lineups for each location position, creating
+Overall this created a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
+Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to help clean the data.
+
+To examine if the difference in detection rate among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
+Let $\mathbf{Y}=(Y_1,Y_2, ...
+, Y_p)^\top$ be a vector of random variables with dimension $p = 5$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represent the $j$th vector response for  location $i$ with $i=1,2, ..., I$ for $U=5$.
+Because the same data plot is shown in each lineup, it is assumed that there could be some association between the responses for each null set, which suggests the MANOVA model rather than a univariate ANOVA.
+The MANOVA model
+\begin{equation}\label{manova}
+\mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
+\end{equation}
+where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$, tests for significant difference between the means.
+
+\subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through MTurk \citep{turk}.
+It is an online work place where people from around the world can sign up for so-called `HIT's, human intelligence tasks, generally short tasks that are humans are typically better at solving than computers.
+Usually tasks are very simple and no specialized training is required to do them.
+Tasks are designed for anyone to do but some tasks may require some skills depending on the recruiters' need.
+%Each task is usually planned to completed in a short time.
+ For completing a HIT workers are paid a small amount of money, on the order of minimum wage in the USA.
+
+We designed and developed a web application which enables the display of lineups to  observers as per experimental need.
+The MTurk workers were re-directed to this web application to complete their assigned tasks.
+Responses were collected, stored automatically into a local database server, along with demographic information, age group, gender and education level.
+The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
+Location of an observer is determined based on the ip address of the observer.
+
+\section{\\Some tables and figures}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=4.5in]{turker_country.pdf} 
+   \includegraphics[width=4.5in]{turker_location.pdf}    
+   \caption{Location of Amazon Mechanical Turk workers participating in the experiments. Most subjects came from India and the United States, but subjects from countries around the world took part.}
+   \label{fig:turker_location}
+\end{figure}
+
+\begin{table}[hbtp]
+\caption{Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
+\centering
+\scalebox{.85}{
+\begin{tabular}{rlrrcr}
+  \hline
+  && \multicolumn{2}{c}{Participants}&Average& Number of\\
+  \cline{3-4}
+ Factor & Levels & Total &\% & Time & Responses \\ 
+  \hline
+Gender & Female & 989 & 42.26 & 43.69 & 10538 \\ 
+   & Male & 1351 & 57.74 & 48.75 & 13542 \\  
+\hline
+  Education & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
+   & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
+   & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
+   & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
+   & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
+\hline
+  Age & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
+   & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
+   & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
+   & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
+   & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
+   & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
+   & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
+\hline   
+  Country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
+   & India & 980 & 41.88 & 52.63 & 10238 \\ 
+   & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\
+
+   \hline
+\end{tabular}
+}
+\label{tbl:demographics}
+\end{table}
+
+% latex table generated in R 3.4.1 by xtable 1.8-2 package
+% Wed Oct 18 17:43:02 2017
+\begin{table}[ht]
+\caption{\label{tbl:anova_factor}Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
+Gender does not have a significant effect on detection rate, but does on time to respond.
+All factors significantly affect time to respond.}
+\centering
+\scalebox{0.9}{
+\begin{tabular}{lrrrrlrrrr}
+& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
+\cline{2-5} \cline{7-10} 
+ & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
+Full  & 52609.1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 24140.5 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
+ - Age & 52081.3 & 527.8 & 6 & $<$0.001 & \hspace{0.5cm} & 24113.2 & 27.3 & 6 & $<$0.001 \\ 
+ - Country & 52245.5 & 363.6 & 2 & $<$0.001 &  & 24114.8 & 25.7 & 2 & $<$0.001 \\ 
+ - Degree & 52498.7 & 110.4 & 4 & $<$0.001 &  & 24116.0 & 24.5 & 4 & $<$0.001 \\ 
+ - Gender & 52540.3 & 68.8 & 1 & $<$0.001 &  & 24138.9 & 1.6 & 1 & 0.200 \\ 
+\cline{2-5} \cline{7-10}
+\end{tabular}}
+\end{table}
+
+\begin{table*}[hbtp] 
+\centering 
+\caption{Overview of 10 different Turk experiments, from which data was collected to study human factor effects.
+All of the experimental data were used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.}
+\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
+\hline\hline 
+ & Experiment &  Test Statistic  & Lineup question & Used in study of\\ [0.5ex] % inserts table %heading 
+\hline 
+1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
+between group A and B? & DF\\
+2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+& DF\\
+  3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+& DF\\
+ 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
+  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
+  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?
+&  DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
+  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?
+& DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
+  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?
+& DF\\
+  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?
+& DF  \hfill \phantom{LT} \hfill {LE} \hfill \phantom{.}\\
+  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?
+& DF\\
+
+\hline 
+\end{tabular} 
+\label{tbl:visual_stat} 
+\end{table*}
+
 \end{document}

### 2024-09-16 wiley_paper/wiley_paper.tex
Add files via upload

--- 2024-09-16 wiley_paper/wiley_paper.Rnw
@@
-\documentclass[10pt]{article}
+\documentclass[10pt]{article}\usepackage[]{graphicx}\usepackage[]{xcolor}
+% maxwidth is the original width if it is less than linewidth
+% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlsng}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hldef}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 \newcommand{\blindDoc}{1}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 Figure~\ref{practical_impact_und_und_graduate.pdf} illustrates this example showing fitted models for a US 18-25 female with either a high school education or an undergraduate degree.
-Similar calculations show the same negligible impact of age level 36-40 (\Sexpr{round(exp(-0.63+0.29)/(1+ exp(-0.63+0.29)) - exp(-0.63)/(1 + exp(-0.63)),3)} at most)  and country India (\Sexpr{round(exp(-0.63-0.19)/(1+ exp(-0.63-0.19)) - exp(-0.63)/(1 + exp(-0.63)), 3)} at most) on the probability of a correct response.
+Similar calculations show the same negligible impact of age level 36-40 (0.068 at most)  and country India (-0.042 at most) on the probability of a correct response.
 Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.

### 2024-09-16 wiley_paper/R/wiley_paper.R

@@ -1,0 +1,556 @@
+# The complete R code for Sociology Paper submitted to Wiley
+# Last updated by Mahbubul Majumder Sep 15, 202024
+
+# ==================================================================
+# Loading necessary libraries 
+# ------------------------------------------------------------------
+
+library(ggplot2)
+library(plyr)
+library(reshape2)
+library(lme4)
+library(xtable)
+library(grid)
+library(lubridate)
+library(dplyr)
+library(maps)
+
+# ==================================================================
+# Loading the data 
+# ------------------------------------------------------------------
+load("../data/experiment-data-all.RData")
+expDat <- allDat
+dat9 <- exp9Dat
+
+# getting the data table information
+# data.frame(sapply(expDat, class))
+# data.frame(sapply(dat9, class))
+
+# ==================================================================
+# generating statistics for section 4.1 "Overview of the Data"
+# ------------------------------------------------------------------
+
+# showing number of countries the subjects are coming from
+length(unique(expDat$country_code))
+# showing countries with more than 10 participants
+countryCount <- expDat %>% 
+  group_by(country_name) %>%
+  summarise(counts = length(unique(id))) %>%
+  arrange(desc(counts))
+# showing number of linesups evaluated
+length(unique(expDat$lineup_id))
+
+
+# ==================================================================
+# figure 3 map of turk participants and country-wise bar chart
+# ----------------------------------------------------------------
+# getting unique participants information
+# Since each participant has multiple responses in the actual data
+turker <- expDat[!duplicated(expDat$id),]
+turker$country <- factor(turker$country, 
+                         levels=names(table(turker$country))[order(table(turker$country))])
+
+map_theme <- list(theme(panel.grid.minor = element_blank(),
+                        panel.grid.major = element_blank(),
+                        panel.background = element_blank(),
+                        axis.line = element_blank(),
[…]

### 2024-09-16 wiley_paper/wiley_paper.Rnw
new edits

@@
 
-%\subsection{Location of Actual Data Plot in the Lineup}
-\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 panels is used.
+%\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 panels is used.
 A random number generator is used to determine the position where the actual data plot is placed in each lineup.
@@
 This leads to some concern that observers might be able to more easily identify the actual data plot, if it is placed at the top left of a lineup than when it the actual data is placed at the bottom right.
-%With the help of eye-tracking equipment \citet{zhao:2012} tracked the observers' eyes to see their approach in searching the plots in a lineup.
-The results suggest that many people read lineups from left to right, some read from upward to downward.
-%There is some concern that if the actual data plot is located in the top left of the lineup, observers may be able to identify it more easily than when the actual data plot is at the bottom right. 
-Ideally, this does not happen.
+
+%Ideally, this does not happen.
 Experiment~9  is designed to allow us to investigate the effects of location on detection.
@@
 
-\item{\bf Sample of Null Plots:} In classical inference, the test statistic is compared to the full null distribution, to decide if it is extreme or not.
+%\item{\bf Sample of Null Plots:} In classical inference, the test statistic is compared to the full null distribution, to decide if it is extreme or not.
 In visual inference the actual plot is compared to plots of a finite number of samples from the null distribution.
@@
 
-\subsection{Overview of the Data} A total of 2340 participants provided feedback data on the lineups in ten different experimental studies.
+ \subsection{Overview of the Data} In ten experimental studies, a total of 2,340 participants from 70 countries, primarily the United States and India, provided feedback data on lineups, ensuring a diverse pool of respondents in terms of geography, gender, age, and education.
+The demographic details, as summarized in Table~\ref{tbl:turker_location}, revealed a high number of female participants across most countries, which was an unexpected but positive outcome.
+Notably, countries like Canada, Romania, the United Kingdom, and Macedonia also had notable participant representation.
+
+The majority of participants were aged 18 to 35, with the United States showing a more balanced age distribution compared to India, where younger individuals were more prevalent.
+Educational backgrounds were diverse, with 38.42\% of participants reporting a graduate degree.
+However, there was a cultural misunderstanding as Indian participants classified any university degree as a graduate degree, while the North American definition refers to a Master’s degree or higher.
+Compared to the 2010 US Census \cite{census}, the participants skewed younger and more highly educated.
+
+In total, 1,912 lineups were evaluated across the studies, with each participant reviewing at least 10 lineups, except for experiment~9, where they reviewed three.
+To ensure quality, participants were given test plots to evaluate first, and incorrect responses led to their data being excluded from the analysis.
+Some demographic and geographical data were missing due to participants opting out or issues retrieving IP-based location data.
+
+%A total of 2340 participants provided feedback data on the lineups in ten different experimental studies.
 Figure~\ref{fig:turker_location} displays the locations of participants around the world.
@@
 The large number of female participants from all countries was a pleasant surprise to us.
-
-% \begin{figure}[htbp] 
-%    \centering
-%    \includegraphics[width=4.5in]{turker_country.pdf} 
-%    \includegraphics[width=4.5in]{turker_location.pdf}    
-%    \caption{Location of Amazon Mechanical Turk workers participating in the experiments. Most subjects came from India and the United States, but subjects from countries around the world took part.}
-%    \label{fig:turker_location}
-% \end{figure}
-
-Besides the United States and India, countries such as Canada, Romania, the United Kingdom and Macedonia have more than 10 participants each.
+% 
+% 
+% Besides the United States and India, countries such as Canada, Romania, the United Kingdom and Macedonia have more than 10 participants each.
 The remaining 64 countries have fewer than 10 participants each.
 The distribution of participants is similar in all ten experiments.
-
-% \begin{table}[hbtp]
-% \caption{Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
-% \centering
-% \scalebox{.85}{
-% \begin{tabular}{rlrrcr}
-%   \hline
-%   && \multicolumn{2}{c}{Participants}&Average& Number of\\
-%   \cline{3-4}
-%  Factor & Levels & Total &\% & Time & Responses \\ 
-%   \hline
-% Gender & Female & 989 & 42.26 & 43.69 & 10538 \\ 
-%    & Male & 1351 & 57.74 & 48.75 & 13542 \\  
-% \hline
-%   Education & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
-%    & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
-%    & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
-%    & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
-%    & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
-% \hline
-%   Age & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
-%    & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
-%    & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
-%    & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
-%    & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
-%    & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
-%    & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
-% \hline   
-%   Country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
-%    & India & 980 & 41.88 & 52.63 & 10238 \\ 
-%    & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\ 
-%   
-%    \hline
-% \end{tabular}
-% }
-% \label{tbl:demographics}
-% \end{table}
-%
-
-The largest number of participants falls within the age group of 18 to 25, with the majority being between 18 to 35.
-Many older participants also took part in the experiments.
-%Interestingly, there are many participants in the older age groups as well.
-The United States shows a more uniform participation beyond age 30,  participants from India are primarily younger.
-
-In terms of academic background, the majority of participants had a graduate degree (a total of 899 participants, about 38.42\%).
+% 
+% 
+% 
+% The largest number of participants falls within the age group of 18 to 25, with the majority being between 18 to 35. Many older participants also took part in the experiments. 
+% The United States shows a more uniform participation beyond age 30,  participants from India are primarily younger. 
+% 
+% In terms of academic background, the majority of participants had a graduate degree (a total of 899 participants, about 38.42\%).
 Examining the participants' location in conjunction with the academic background reveals that there is an above average  number of participants from India reporting to have a graduate degree.
@@
 report to have an undergraduate degree or at least have some undergraduate courses.
-
-Compared to data from the 2010 US Census \cite{census}, participants were both relatively younger, and more highly educated than the general US population. 
-Being turkers implicitly assures that the participants in these experiments have a relatively high level of technological fluency.
+% 
+% Compared to data from the 2010 US Census \cite{census}, participants were both relatively younger, and more highly educated than the general US population. 
+% Being turkers implicitly assures that the participants in these experiments have a relatively high level of technological fluency.
 In fact, the demographics of participants align somewhat better with demographics of internet users published by
-\citet{pew}.
-
-The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where the proportion of female participants is lower.
+% \citet{pew}.
+% 
+% The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where the proportion of female participants is lower.
 The distribution of education levels also differs slightly across countries for this age group.
-
-A total of 1912 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where each participant evaluated three lineups.
-As a measure to ensure a high quality in the evaluations,
- a test plot was shown to each observer, and the performance on this test plot was used for screening purposes.
-%A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
-Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.
-Data from participants who gave the wrong answer on the test plot % were considered as gaming the system, and their data were removed.
-led to the removal of all of their answers from the analysis. 
-Results from test plots were also not considered in any of the analyses.
- In some cases  participants chose to not provide their demographic information.
+% 
+% 
+% A total of 1912 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where each participant evaluated three lineups.
+% As a measure to ensure a high quality in the evaluations,
+%  a test plot was shown to each observer, and the performance on this test plot was used for screening purposes. 
+% Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.
+% Data from participants who gave the wrong answer on the test plot 
+% led to the removal of all of their answers from the analysis. 
+% Results from test plots were also not considered in any of the analyses.
+%  In some cases  participants chose to not provide their demographic information.
 Also, for some ip address, the actual geographical locations could not be retrieved.
@@
 With respect to time taken, all of the demographic factors are significant.
-
-% 
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 17:43:02 2017
-% \begin{table}[ht]
-% \caption{\label{tbl:anova_factor}Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
-Gender does not have a significant effect on detection rate, but does on time to respond.
-All factors significantly affect time to respond.}
-% \centering
-% \scalebox{0.9}{
-% \begin{tabular}{lrrrrlrrrr}
-% & \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
-% \cline{2-5} \cline{7-10} 
-%  & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
-% Full  & 52609.1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 24140.5 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
-%  - Age & 52081.3 & 527.8 & 6 & $<$0.001 & \hspace{0.5cm} & 24113.2 & 27.3 & 6 & $<$0.001 \\ 
-%  - Country & 52245.5 & 363.6 & 2 & $<$0.001 &  & 24114.8 & 25.7 & 2 & $<$0.001 \\ 
-%  - Degree & 52498.7 & 110.4 & 4 & $<$0.001 &  & 24116.0 & 24.5 & 4 & $<$0.001 \\ 
-%  - Gender & 52540.3 & 68.8 & 1 & $<$0.001 &  & 24138.9 & 1.6 & 1 & 0.200 \\ 
-% \cline{2-5} \cline{7-10}
-% \end{tabular}}
-% \end{table}
-
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 17:43:02 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{rrrrrlrrrr}
-%   \hline
-%  & deviance & Chisq & Chi.Df & Pr..Chisq. & g1 & deviance & Chisq & Chi.Df & Pr..Chisq. \\ 
-%   \hline
-% Full & 52609.06 & 0.00 & 6.00 & 0.00 &  & 24140.49 & 0.00 & 6.00 & 0.00 \\ 
-%   Age & 52081.29 & 527.77 & 6.00 & 0.00 &  & 24113.24 & 27.25 & 6.00 & 0.00 \\ 
-%   Country & 52245.46 & 363.60 & 2.00 & 0.00 &  & 24114.79 & 25.70 & 2.00 & 0.00 \\ 
-%   Degree & 52498.65 & 110.41 & 4.00 & 0.00 &  & 24116.02 & 24.47 & 4.00 & 0.00 \\ 
-%   Gender & 52540.30 & 68.76 & 1.00 & 0.00 &  & 24138.85 & 1.64 & 1.00 & 0.20 \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
-
-Parameter estimates from the model fits are shown in Table~\ref{tbl:model_result_demographics}.
-The first level of each factor serves as the baseline for the model.
-The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for other age groups except for 26-30.
-Participants from the rest of the world have a similar detection rate to those from the USA, but participants from India show a significantly lower detection rate.
-Subjects reporting to have  an undergraduate degree had a significantly higher detection rate.
-For time to respond, all age groups were significantly faster than  18-25 year olds.
-Participants from India as well as the rest of the world were faster than US Americans.
-A higher degree corresponded to longer time, and males were on average faster than females.
 
@@
 
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 17:22:24 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{rrrrllrrrl}
-%   \hline
-%  & Est & LB & UB & pvalue & g1 & Est & LB & UB & pvalue \\ 
-%   \hline
-% $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.63 & -0.81 & -0.45 & *** \\ 
-%   26-30 & 0.05 & 0.03 & 0.07 & *** &   & 0.03 & -0.06 & 0.13 &  \\ 
-%   31-35 & 0.06 & 0.03 & 0.09 & *** &   & 0.12 & 0.02 & 0.23 & * \\ 
-%   36-40 & 0.22 & 0.19 & 0.25 & *** &   & 0.29 & 0.17 & 0.41 & *** \\ 
-%   41-45 & 0.17 & 0.13 & 0.21 & *** &   & 0.19 & 0.03 & 0.34 & * \\ 
-%   46-50 & 0.28 & 0.23 & 0.32 & *** &   & 0.20 & 0.01 & 0.39 & * \\ 
-%   above 50 & 0.34 & 0.31 & 0.38 & *** &   & 0.16 & 0.02 & 0.29 & * \\ 
-%   India & 0.23 & 0.20 & 0.25 & *** &   & -0.19 & -0.28 & -0.10 & *** \\ 
-%   Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.06 & -0.05 & 0.18 &  \\ 
-%   Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & 0.00 & -0.14 & 0.14 &  \\ 
-%   Under grad degree & 0.14 & 0.10 & 0.17 & *** &   & 0.16 & 0.03 & 0.29 & * \\ 
-%   Graduate courses & 0.02 & -0.02 & 0.06 &  &   & -0.13 & -0.29 & 0.03 &  \\ 
-%   Graduate degree & 0.13 & 0.10 & 0.16 & *** &   & -0.04 & -0.17 & 0.09 &  \\ 
-%   Male & 0.08 & 0.06 & 0.10 & *** &   & 0.05 & -0.02 & 0.12 &  \\ 
-%   lineup & 0.29 &  &  &  &   & 2.27 &  &  &  \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
+Parameter estimates from the model fits are shown in Table~\ref{tbl:model_result_demographics}.
+The first level of each factor serves as the baseline for the model.
+The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for other age groups except for 26-30.
+Participants from the rest of the world have a similar detection rate to those from the USA, but participants from India show a significantly lower detection rate.
+Subjects reporting to have  an undergraduate degree had a significantly higher detection rate.
+For time to respond, all age groups were significantly faster than  18-25 year olds.
+Participants from India as well as the rest of the world were faster than US Americans.
+A higher degree corresponded to longer time, and males were on average faster than females.
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=4in]{demographic_effect.pdf} 
+   \caption{Boxplots of average log time taken and proportion correct responses (detection rate) of all the lineups plotted for each demographic factor levels.
+The dots inside the boxes represent means.
+Some differences in means of various demographic factors are observed.
+Variability in detection rate indicates large variability in lineup difficulties.
+}
+   \label{fig:demographic_effect}
+\end{figure}
 
@@
 
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=4in]{demographic_effect.pdf} 
-   \caption{Boxplots of average log time taken and proportion correct responses (detection rate) of all the lineups plotted for each demographic factor levels.
-The dots inside the boxes represent means.
-Some differences in means of various demographic factors are observed.
-Variability in detection rate indicates large variability in lineup difficulties.
-}
-   \label{fig:demographic_effect}
-\end{figure}
-
 While some of the demographic factors are strongly statistically significant, the main source of variation in detection rate is the lineup difficulty.
@@
 There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.
+
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 23:54:25 2017
+% \begin{table}[hbtp]
+% \centering
+% \caption{\label{tbl:model_result_response}Parameter estimates of models \eqref{eqn:trend_response} fitted to data from three different experiments using detection rate as the response to assess learning trend.
+Attempt number is fitted as a factor to enable modeling any sort of non-linear learning trend.
+Only experiment~6 shows some evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
+% \scalebox{.9}{
+% \begin{tabular}{rr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
+% \cline{2-5} \cline{7-10} \cline{12-15} 
+% & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
+% \cline{2-5} \cline{7-10} \cline{12-15} 
+% Effect & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  \\ 
+% \cline{2-5} \cline{7-10} \cline{12-15} 
+% Fixed \\
+% $\mu$ & -1.32 & -1.73 & -0.92 & *** &   & -0.16 & -0.46 & 0.13 &  &   & -1.73 & -2.81 & -0.66 & ** \\ [3pt]
+%   $\alpha2$ & 0.30 & -0.17 & 0.78 &  &   & 0.26 & -0.07 & 0.58 &  &   & -0.40 & -1.23 & 0.44 &  \\ 
+%   $\alpha3$ & -0.20 & -0.68 & 0.29 &  &   & 0.31 & -0.02 & 0.63 & . &   & -0.15 & -0.99 & 0.68 &  \\ 
+%   $\alpha4$ & 0.13 & -0.35 & 0.61 &  &   & 0.34 & 0.01 & 0.66 & * &   & -0.41 & -1.24 & 0.42 &  \\ 
+%   $\alpha5$ & 0.35 & -0.14 & 0.83 &  &   & 0.34 & 0.02 & 0.67 & * &   & -0.10 & -0.92 & 0.73 &  \\ 
+%   $\alpha6$ & 0.10 & -0.40 & 0.60 &  &   & 0.20 & -0.12 & 0.53 &  &   & 0.03 & -0.84 & 0.89 &  \\ 
+%   $\alpha7$ & 0.33 & -0.17 & 0.82 &  &   & 0.11 & -0.22 & 0.43 &  &   & -0.00 & -0.84 & 0.84 &  \\ 
+%   $\alpha8$ & -0.01 & -0.51 & 0.50 &  &   & 0.31 & -0.01 & 0.64 & . &   & -0.06 & -0.88 & 0.76 &  \\ 
+%   $\alpha9$ & -0.20 & -0.70 & 0.30 &  &   & 0.34 & 0.01 & 0.67 & * &   & 0.21 & -0.67 & 1.08 &  \\ 
+%   $\alpha10$ & 0.51 & 0.02 & 1.01 & * &   & 0.19 & -0.14 & 0.53 &  &   & -0.20 & -1.07 & 0.67 &  \\ 
+%   Random \\
+%   $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.91}  & \multicolumn{5}{l}{\phantom{-}0.89} & 0.82   \\ 
+%   $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.04}  & \multicolumn{5}{l}{\phantom{-}0.04} & 0.05   \\ 
+%   $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}1.47}  & \multicolumn{5}{l}{\phantom{-}1.43} & 3.38   \\ \cline{2-5} \cline{7-10} \cline{12-15} 
+% \\
+% \hline
+% \multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
+% \end{tabular}}
+% \end{table}
+
+% 
+% 
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 23:54:25 2017
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{rrrrllrrrllrrrl}
+%   \hline
+%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
+%   \hline
+% $\mu$ & -1.32 & -1.73 & -0.92 & *** &   & -0.16 & -0.46 & 0.13 &  &   & -1.73 & -2.81 & -0.66 & ** \\ 
+%   $\alpha2$ & 0.30 & -0.17 & 0.78 &  &   & 0.26 & -0.07 & 0.58 &  &   & -0.40 & -1.23 & 0.44 &  \\ 
+%   $\alpha3$ & -0.20 & -0.68 & 0.29 &  &   & 0.31 & -0.02 & 0.63 & . &   & -0.15 & -0.99 & 0.68 &  \\ 
+%   $\alpha4$ & 0.13 & -0.35 & 0.61 &  &   & 0.34 & 0.01 & 0.66 & * &   & -0.41 & -1.24 & 0.42 &  \\ 
+%   $\alpha5$ & 0.35 & -0.14 & 0.83 &  &   & 0.34 & 0.02 & 0.67 & * &   & -0.10 & -0.92 & 0.73 &  \\ 
+%   $\alpha6$ & 0.10 & -0.40 & 0.60 &  &   & 0.20 & -0.12 & 0.53 &  &   & 0.03 & -0.84 & 0.89 &  \\ 
+%   $\alpha7$ & 0.33 & -0.17 & 0.82 &  &   & 0.11 & -0.22 & 0.43 &  &   & -0.00 & -0.84 & 0.84 &  \\ 
+%   $\alpha8$ & -0.01 & -0.51 & 0.50 &  &   & 0.31 & -0.01 & 0.64 & . &   & -0.06 & -0.88 & 0.76 &  \\ 
+%   $\alpha9$ & -0.20 & -0.70 & 0.30 &  &   & 0.34 & 0.01 & 0.67 & * &   & 0.21 & -0.67 & 1.08 &  \\ 
+%   $\alpha10$ & 0.51 & 0.02 & 1.01 & * &   & 0.19 & -0.14 & 0.53 &  &   & -0.20 & -1.07 & 0.67 &  \\ 
+%   $\sigma^2_u$ & 0.91 &  &  &  &   & 0.89 &  &  &  &   & 0.82 &  &  &  \\ 
+%   $\sigma^2_a$ & 0.04 &  &  &  &   & 0.04 &  &  &  &   & 0.05 &  &  &  \\ 
+%   $\sigma^2_l$ & 1.47 &  &  &  &   & 1.43 &  &  &  &   & 3.38 &  &  &  \\ 
+%    \hline
+% \end{tabular}
+% \end{table}
+% 
+%
+
+To visualize how  detection rates change over successive attempts, we fitted model~\eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
+Least square regression lines were fitted through the subject specific residuals as shown in Figure~\ref{fig:learning_trend_response}.
+The averages of these residuals for each of the attempts are shown as dots.
+Two important features were observed; one is subject specific variability and the other is random slope with attempts which indicates subjects specific learning trend.
+Some subjects show improvement over time and some show the decrease in performance.
+Although, the model fit experiment~6 suggested a statistically significant learning curve the plots indicates that it is minimal.
+
+\begin{figure}[htbp] 
+   \centering
+    \includegraphics[width=6.3in]{learning_trend_subject.pdf} 
+   \caption{ Least square lines fitted through the subject specific residual proportion correct obtained from model~\eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
+Subject specific positive and negative slopes are observed.
+Mean residuals are shown as dots and least square regression lines fitted through the points show no overall learning trend in each of the three experiments.}
+   \label{fig:learning_trend_response}
+\end{figure}
+
+Table~\ref{tbl:model_result_time} presents the results of model~\eqref{eqn:trend_time} where response time is examined with respect to attempt.
+Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated.
+Interestingly, time to respond significantly decreases as number of attempts increases, for all three experiments.
+The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments.
+The negative estimates suggest that on average later attempts took less time.
+Even though observers did not improve in their performance in successive attempts, they became more efficient in their response.
+The parameter $\alpha_1$ for first attempt is also highly significant.
+The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts.
+One explanation might be that participants take the chance to  carefully read through instructions and familiarize themselves  with the types of plots used in the experiment before evaluating their first lineup, which would inflate the time taken on the first lineup.
+Each of the experiments asks observers to give a reason for their choice (out of a list of pre-defined answers.
+For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).
+Later pages of the web site were identical  except for the lineup.
+
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 23:37:55 2017
+% \begin{table}[bhtp]
+% \caption{\label{tbl:model_result_time}Parameter estimates of model~\eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
+Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments~5,~6 and~7.}
+% \centering
+% \scalebox{0.9}{
+% \begin{tabular}{lr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
+% \cline{2-5} \cline{7-10} \cline{12-15} 
+% & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
+% 
+% \cline{2-5} \cline{7-10} \cline{12-15} 
+% \multicolumn{2}{c}{Effect \hfill Est} & 2.5\% & 97.5\% &  &  & Est & 2.5\% & 97.5\% & &  & Est & 2.5\% & 97.5\% & \\ 
+%   \cline{2-5} \cline{7-10} \cline{12-15}
+% \multicolumn{2}{l}{Fixed}\\
+% $\mu$ & 3.80 & 3.72 & 3.88 & *** &   & 3.89 & 3.83 & 3.96 & *** &   & 3.75 & 3.64 & 3.85 & *** \\ 
+%   $\alpha_1$ & 0.32 & 0.25 & 0.39 & *** &   & 0.34 & 0.28 & 0.40 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
+%   $\alpha$ & -0.04 & -0.05 & -0.03 & *** &   & -0.04 & -0.05 & -0.03 & *** &   & -0.03 & -0.04 & -0.02 & *** \\ 
+% \multicolumn{2}{l}{Random}\\
+%   $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.52}   & \multicolumn{5}{l}{\phantom{-}0.49}  & 0.37   \\ 
+%   $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.03}  & \multicolumn{5}{l}{\phantom{-}0.05}   & 0.05   \\ 
+%   $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}0.09}   & \multicolumn{5}{l}{\phantom{-}0.20}   & 0.24   \\
+%   $\sigma^2$ & \multicolumn{5}{l}{\phantom{-}0.46}   & \multicolumn{5}{l}{\phantom{-}0.50}  & 0.45  \\ 
+%    \cline{2-5} \cline{7-10} \cline{12-15}
+% \end{tabular}}
+% \end{table}
+%
+
+% 
+% % latex table generated in R 3.4.1 by xtable 1.8-2 package
+% % Wed Oct 18 23:37:55 2017
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{rrrrllrrrllrrrl}
+%   \hline
+%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
+%   \hline
+% $\mu$ & 3.80 & 3.72 & 3.88 & *** &   & 3.89 & 3.83 & 3.96 & *** &   & 3.75 & 3.64 & 3.85 & *** \\ 
+%   $\alpha_1$ & 0.32 & 0.25 & 0.39 & *** &   & 0.34 & 0.28 & 0.40 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
+%   $\alpha$ & -0.04 & -0.05 & -0.03 & *** &   & -0.04 & -0.05 & -0.03 & *** &   & -0.03 & -0.04 & -0.02 & *** \\ 
+%   $\sigma^2_u$ & 0.52 &  &  &  &   & 0.49 &  &  &  &   & 0.37 &  &  &  \\ 
+%   $\sigma^2_a$ & 0.03 &  &  &  &   & 0.05 &  &  &  &   & 0.05 &  &  &  \\ 
+%   $\sigma^2_l$ & 0.09 &  &  &  &   & 0.20 &  &  &  &   & 0.24 &  &  &  \\ 
+%   $\sigma^2$ & 0.46 &  &  &  &   & 0.50 &  &  &  &   & 0.45 &  &  &  \\ 
+%    \hline
+% \end{tabular}
+% \end{table}
+
+To visualize how the time taken reduces over the successive attempts, we fitted model~\eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
+Least square regression lines are fitted through the subject specific residuals.
+Subject specific slopes are much different in each of the three experiments as we see in Figure~\ref{fig:learning_trend_time}.
+Some subjects improved over attempts by taking less time in the later attempts while others got worse.
+The averages of these residuals for each attempt are plotted as dots.
+Least square regression lines are fitted to these points excluding the first attempt since for first attempt we fitted an indicator covariate.
+The downward trends are evident in the plots.
+All the slopes are highly significant.
+As expected we observed large positive residuals for each of the experiments for first attempt.
+
+\begin{figure}[htbp] 
+   \centering
+%   \includegraphics[width=6.5in]{learning_trend_time.pdf}    
+ \includegraphics[width=6.3in]{learning_trend_time_subject.pdf}    
+   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting model~\eqref{eqn:trend_time} without covariate attempt.
+Differences in subject specific slopes are observed.
+Some of the subjects did worse over successive attempts while others did better.
+Averages of these residuals are plotted as dots and least square regression lines are fitted to obtain overall trends.
+For all the three experiments the overall downward slopes are statistically significant which indicates that MTurk workers become more efficient as they progress through their attempts.}
+   \label{fig:learning_trend_time}
+\end{figure}
+
+% \subsection{Location Effect}
+% 
+% A total of 111 subjects was recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
+Each subject evaluated two lineups;  one for Interaction effect and the other for Genotype.
+In total there were 222 responses for 50 lineups.
+The data on the test lineup were excluded from the analysis.
+% 
+% We are investigating location effect in two ways: (1) location effect of each individual plot location, and (2) location effect of inside/outside panel positions.
+% 
+% Figure~\ref{fig:location_effect} shows an overview of detection rates for both of the effects studied in experiment~9.
+Each dot corresponds to one lineup.
+The size of each dot gives the number of evaluations.
+Lineups with  the same set of null plots are connected by lines.
+% %The detection rate for each data plot location is shown in colored by null sets. 
+% The overall average detection rate for each location is shown using dashed lines.
+Size of the points represent the number of responses.
+For some locations we have as many as 10 responses.
+For location 1, we did not have any responses for null set 1 in one of the interaction lineups.
+% 
+% We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets.
+The lineup protocol is using a finite, small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set.
+If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate.
+From the plot we can see that null set 3 appeared to be more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot.
+This null set effect was tested, and a significant difference was found for null set 3 of the interaction lineup.
+The rest of the null sets do not show any differences at  a 1\% significance level.
+% 
+% 
+% \begin{figure}[htbp] 
+%    \centering
+%     \includegraphics[width=6.5in]{proportion_nulls_guide.pdf} 
+%    \caption{Location of data plot in the lineup and proportion correct (detection rate) for both Interaction and Genotype effect.
+Each colored line represents a null set and the size of the dots represents number of responses.
+The overall average proportions are shown by dashed lines.
+The actual data plot locations are shaded grey on the top panels to demonstrate their relative positions on a lineup.}
+%    \label{fig:location_effect}
+% \end{figure}
+% 
+% 
+% Model~\eqref{manova} was fit to the data to test, if  detection rates were significantly different for the different locations.
+For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R}.
+The results are shown in Table~\ref{tbl:manova}.
+The $p$-values for both the Interaction and the Genotype effect are much bigger than the conventional threshold of 0.05 suggesting that location of the actual data plot in the lineup does not significantly affect performance.
+% 
+% \begin{table}[hbtp]
+% \caption{The summary of the results obtained by fitting MANOVA model~\eqref{manova} using Wilks test.}
+% \begin{center}
+% \begin{tabular}{ccccccccc}
+%   \hline \hline
+%  Location & && & \multicolumn{3}{c} {Degrees of Freedom}  & F test \\
+%  \cline{5-7}
+%  Effect & DF & Wilks & Approx. F & Numerator& Denominator &Residual & $p$-value\\
+%   \hline
+%   Interaction& 3&0.07986 & 0.592 &15& \phantom{0}5.92 & 6 & 0.8082 \\ 
+%   Genotype & 4&0.01689 & 1.722 & 20&14.22 & 8 & 0.1488 \\ 
+%    \hline
+% \end{tabular}
+% \end{center}
+% \label{tbl:manova}
+% \end{table}
+% 
+% For the second approach in investigating the location effect, we distinguish between inner panels 7, 8, 9, 12, 13, 14 and outer panels (any panel with an outside edge).
+From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype.
+Model~\eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.
+
+% ===================================================================================
+\section{Conclusion}
+%------------------------------------------------------------------------------------
+
+Human demographics have the potential to influence performance when the lineup protocol is used for statistical inference.
+In our study of the demographic effects on performance across a set of different experiments we found some statistically significant effects.
+Age group 36-40, people who have a graduate degree have a significantly higher detection rate.
+India shows significantly lower detection rate compared to United States.
+Gender does not have any significant effect.
+However, the effects are minimal, on the order of a few percentage points different from the average.
+These results are very important for the power of visual tests as they demonstrate the robustness of the test against different human factors.
+
+Individual learning trend is observed in time taken but not so much in observer performance.
+Some individuals improved on their performances while others showed a decrease on successive attempts.
+This could be interpreted as good, in the sense that it means that observers could realistically be recruited from sources like MTurk, and that substantial training is not necessary in order to obtain useful evaluations of lineups in practice.
+However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea, at least for short term learning.
+
+The simulation experiment reveals that there is no significant effect of location on the actual data plot in the lineup.
+This is important as the visual statistical inference procedure prescribes that the data plot be placed at random in the lineup.
+This paper suggests that any random place in a lineup is as good as other places in the lineup.
+Even though there are variations on the performance depending on different null sets, their impact on probability to correctly evaluate a lineup is very negligible.
+
+There are many more ways that the lineup protocol might be tested to learn what we might expect about its performance for tackling real data mining problems.
+Possible changes in the lineup protocol in the pipeline are allowing observers to select more than one plot, and to vary the size of the lineup from 20 to a smaller number, and have observers all see different lineups with different null sets.
+This paper assessed the effect of human factors on the experiments conducted to date.
+As changes to the lineup protocol are suggested by other experiments the human factor effects may need to be examined again.
+
+\if1\blindDoc
+{
+%------------------------------------------------------------------------------------
+\paragraph{Acknowledgments}
+%------------------------------------------------------------------------------------
+This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
+} \fi
+
+\bibliographystyle{asa}
+\bibliography{references}
+
+% ---------------------------------------- Appendix ----------------------------------
+
+\appendix
+
+% ===================================================================================
+\section{Experimental Methods}\label{sec:exp_design}
+%------------------------------------------------------------------------------------
+
+Two of the factors, signal in the data and individual abilities, were studied in \cite{majumder:2013}.
+The choice of visual test statistic was examined in \cite{heike:2012}.
+In each of these analyses demographic factors were given a cursory glance, to ensure that they did not have large effects on the results.
+The design of experiments 5, 6 and 7 enables the examination of learning trend, which is studied in this paper.
+Experiment~9 was a real test case for visual statistical inference, and in order to understand the significance of the structure in the genomic data, multiple lineups were made in which location of the actual data plot, and the sample of nulls, were randomized.
+This enables the assessment of the effect of these factors on the results.
+This section describes the experimental methods used to examine the effects of demography, placement of the actual data plot, sample of null plots and the existence of a learning trend.
+
+\subsection{Demographic Factors}
+\noindent
+For all  ten experiments shown in Table~\ref{tbl:visual_stat}, the following demographic information was collected from subjects:
+
+\begin{enumerate} \leftmargin 5cm  \itemsep 0in
+\item {\it Age group}, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, above 50.
+\item {\it Gender}, male or female.
+\item {\it Education level}, with levels being high school or less, some undergraduate courses, undergraduate degree, some graduate courses, and graduate degree.
+\item {\it Geographical location}, collected from the IP address of the participants' computer, as latitude, longitude, city and country. 
+\end{enumerate}
+
+Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
+The factors are examined in association with the observer's response using a logistic regression with random effect terms:
+
+\begin{equation} \label{eqn:demographic_response}
+g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
+\end{equation}
+
+where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ picks the actual data plot from lineup $j$, $\mu$ is an overall population average, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively, for observer $i$.
+The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume independence and normality of the errors, i.e.\ $\ell_j \sim N(0, \sigma_\ell^2)$.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+
+Similarly, we model the  time  an observer takes to identify a panel from a lineup.
+Let $Z_{ij}$ denote the logarithm of time taken for  observer $i$ to evaluate  lineup $j$.
+Let $\mu_{ij}=  E(Z_{ij})$ be the average of the (log) time taken by  observer $i$ to pick a panel from lineup $j$.
+We model this in a mixed effects model of the same structure as model~(\ref{eqn:demographic_response}) given as:
+%\hh{
+\begin{equation} \label{eqn:demographic_time}
+Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
+\end{equation}
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model~\eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.
+
+\subsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive responses when multiple lineups are shown for evaluation.
+Experiments 5, 6 and 7 were used for this.
+Each subject was shown a total of 10 lineups randomly selected from a pool of lineups.
+The lineups are not necessarily of the same difficulty level, but the order of lineups was randomized.
+The responses of the lineups were recorded by attempt 1 through 10.
+Attempt 1 means that the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
+The goal is to estimate whether performance of the observer improves, or changes, from attempt 1 to attempt 10.
+
+It should be noted that we are examining the observer's performance, when we model response as detected or not, but this is not the goal of visual inference.
+Visual inference is constructed to measure the significance of structure discovered in data.
+It is expected that some observers will be more skilled at reading data plots, and hence, more readily detect the plot that is different.
+It is also expected that as observers gain experience in evaluating lineups that they become more proficient in reading data plots, particularly if feedback is given on whether the actual data plot was chosen or not.
+Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups.
+So in this context, detected, or not, is used as a response to examine individual differences.
+
+Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k$th evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
+Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the actual data plot from lineup $j$ in their $k$th attempt.
+Learning trend is assessed using a generalized mixed effects model of the form
+\begin{equation} \label{eqn:trend_response}
+g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
+\end{equation}
+where $\mu$ is an overall population average, $\alpha_k$ is the effect of the $k$th attempt on the probability, using the first attempt as reference, $\alpha_1 = 0$, and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
+The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
+The term $a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts, where $a_i \sim N(0, \sigma_a^2)$. 
+For $\ell_j$ a normal distribution, $N(0, \sigma_\ell^2)$, is assumed, and $\ell_j$ is a random intercept predicting lineup difficulty level.
+$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
+The inverse link function, $g^{-1}(.)$, from equation~\ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in the $k$th attempt by a single observer as
+\begin{equation} \label{eqn:trend_power}
+\hat p_{ijk} =  g^{-1}(\hat{\mu} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
+\end{equation}
+
+When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
+Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of the (log) of time taken by  observer $i$ to choosing a  panel from lineup $j$ in his/her $k$th attempt.
+We evaluate this in a mixed effects model of the form
+
+\begin{equation} \label{eqn:trend_time}
+Z_{ijk} = \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
+\end{equation}
+where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$ is the average change in log time taken for each additional attempt,  $\alpha_1$ is an offset in log time taken for the first attempt.
+All other effects are random effects: $u_i$ is a subject-specific intercept representing individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$, $a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$, with $a_i \sim N(0, \sigma_a^2)$, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
+Equation~\ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
+\begin{equation} \label{eqn:trend_time_est}
+\hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j.
+\end{equation}
+
+To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2015, lme4:paper}.
+We employ a normal approximation to obtain $p$-values corresponding to fixed effect parameters estimates.
+
+\subsection{Location Effect} \label{sec:location_design} Experiment~9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
+The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
+Two factors were of primary interest a main effect for genotype and an interaction effect between genotype and treatment condition.
+
+In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
+This overall significance is studied using visual inference  in \citet{tengfei:2013}.
+
+In order to study the effect that  location has on evaluating lineups, we used multiple lineups for each of the data plots.
+For each data plot, five sets of null plots were generated and the actual plot was randomly placed in one of five different locations in a lineup of size 20.
+For the genotype effect, the locations were  1, 8, 12, 17, 20 and locations 2, 9, 12, 16, 20 were used for the interaction effect.
+%Five different sets of null plots were used to produce 5 lineups for each location position, creating
+Overall this created a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
+Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to help clean the data.
+
+To examine if the difference in detection rate among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
+Let $\mathbf{Y}=(Y_1,Y_2, ...
+, Y_p)^\top$ be a vector of random variables with dimension $p = 5$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represent the $j$th vector response for  location $i$ with $i=1,2, ..., I$ for $U=5$.
+Because the same data plot is shown in each lineup, it is assumed that there could be some association between the responses for each null set, which suggests the MANOVA model rather than a univariate ANOVA.
+The MANOVA model
+\begin{equation}\label{manova}
+\mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
+\end{equation}
+where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$, tests for significant difference between the means.
+
+\subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through MTurk \citep{turk}.
+It is an online work place where people from around the world can sign up for so-called `HIT's, human intelligence tasks, generally short tasks that are humans are typically better at solving than computers.
+Usually tasks are very simple and no specialized training is required to do them.
+Tasks are designed for anyone to do but some tasks may require some skills depending on the recruiters' need.
+%Each task is usually planned to completed in a short time.
+ For completing a HIT workers are paid a small amount of money, on the order of minimum wage in the USA.
+
+We designed and developed a web application which enables the display of lineups to  observers as per experimental need.
+The MTurk workers were re-directed to this web application to complete their assigned tasks.
+Responses were collected, stored automatically into a local database server, along with demographic information, age group, gender and education level.
+The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
+Location of an observer is determined based on the ip address of the observer.
+
+\section{\\Some tables and figures}
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=4.5in]{turker_country.pdf} 
+   \includegraphics[width=4.5in]{turker_location.pdf}    
+   \caption{Location of Amazon Mechanical Turk workers participating in the experiments. Most subjects came from India and the United States, but subjects from countries around the world took part.}
+   \label{fig:turker_location}
+\end{figure}
+
+\begin{table}[hbtp]
+\caption{Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
+\centering
+\scalebox{.85}{
+\begin{tabular}{rlrrcr}
+  \hline
+  && \multicolumn{2}{c}{Participants}&Average& Number of\\
+  \cline{3-4}
+ Factor & Levels & Total &\% & Time & Responses \\ 
+  \hline
+Gender & Female & 989 & 42.26 & 43.69 & 10538 \\ 
+   & Male & 1351 & 57.74 & 48.75 & 13542 \\  
+\hline
+  Education & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
+   & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
+   & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
+   & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
+   & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
+\hline
+  Age & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
+   & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
+   & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
+   & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
+   & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
+   & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
+   & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
+\hline   
+  Country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
+   & India & 980 & 41.88 & 52.63 & 10238 \\ 
+   & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\
+
+   \hline
+\end{tabular}
+}
+\label{tbl:demographics}
+\end{table}
+
+% latex table generated in R 3.4.1 by xtable 1.8-2 package
+% Wed Oct 18 17:43:02 2017
+\begin{table}[ht]
+\caption{\label{tbl:anova_factor}Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
+Gender does not have a significant effect on detection rate, but does on time to respond.
+All factors significantly affect time to respond.}
+\centering
+\scalebox{0.9}{
+\begin{tabular}{lrrrrlrrrr}
+& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
+\cline{2-5} \cline{7-10} 
+ & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
+Full  & 52609.1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 24140.5 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
+ - Age & 52081.3 & 527.8 & 6 & $<$0.001 & \hspace{0.5cm} & 24113.2 & 27.3 & 6 & $<$0.001 \\ 
+ - Country & 52245.5 & 363.6 & 2 & $<$0.001 &  & 24114.8 & 25.7 & 2 & $<$0.001 \\ 
+ - Degree & 52498.7 & 110.4 & 4 & $<$0.001 &  & 24116.0 & 24.5 & 4 & $<$0.001 \\ 
+ - Gender & 52540.3 & 68.8 & 1 & $<$0.001 &  & 24138.9 & 1.6 & 1 & 0.200 \\ 
+\cline{2-5} \cline{7-10}
+\end{tabular}}
+\end{table}
 
@@
 
-% 
-% 
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 23:54:25 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{rrrrllrrrllrrrl}
-%   \hline
-%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
-%   \hline
-% $\mu$ & -1.32 & -1.73 & -0.92 & *** &   & -0.16 & -0.46 & 0.13 &  &   & -1.73 & -2.81 & -0.66 & ** \\ 
-%   $\alpha2$ & 0.30 & -0.17 & 0.78 &  &   & 0.26 & -0.07 & 0.58 &  &   & -0.40 & -1.23 & 0.44 &  \\ 
-%   $\alpha3$ & -0.20 & -0.68 & 0.29 &  &   & 0.31 & -0.02 & 0.63 & . &   & -0.15 & -0.99 & 0.68 &  \\ 
-%   $\alpha4$ & 0.13 & -0.35 & 0.61 &  &   & 0.34 & 0.01 & 0.66 & * &   & -0.41 & -1.24 & 0.42 &  \\ 
-%   $\alpha5$ & 0.35 & -0.14 & 0.83 &  &   & 0.34 & 0.02 & 0.67 & * &   & -0.10 & -0.92 & 0.73 &  \\ 
-%   $\alpha6$ & 0.10 & -0.40 & 0.60 &  &   & 0.20 & -0.12 & 0.53 &  &   & 0.03 & -0.84 & 0.89 &  \\ 
-%   $\alpha7$ & 0.33 & -0.17 & 0.82 &  &   & 0.11 & -0.22 & 0.43 &  &   & -0.00 & -0.84 & 0.84 &  \\ 
-%   $\alpha8$ & -0.01 & -0.51 & 0.50 &  &   & 0.31 & -0.01 & 0.64 & . &   & -0.06 & -0.88 & 0.76 &  \\ 
-%   $\alpha9$ & -0.20 & -0.70 & 0.30 &  &   & 0.34 & 0.01 & 0.67 & * &   & 0.21 & -0.67 & 1.08 &  \\ 
-%   $\alpha10$ & 0.51 & 0.02 & 1.01 & * &   & 0.19 & -0.14 & 0.53 &  &   & -0.20 & -1.07 & 0.67 &  \\ 
-%   $\sigma^2_u$ & 0.91 &  &  &  &   & 0.89 &  &  &  &   & 0.82 &  &  &  \\ 
-%   $\sigma^2_a$ & 0.04 &  &  &  &   & 0.04 &  &  &  &   & 0.05 &  &  &  \\ 
-%   $\sigma^2_l$ & 1.47 &  &  &  &   & 1.43 &  &  &  &   & 3.38 &  &  &  \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
-% 
-%
-
-To visualize how  detection rates change over successive attempts, we fitted model~\eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
-Least square regression lines were fitted through the subject specific residuals as shown in Figure~\ref{fig:learning_trend_response}.
-The averages of these residuals for each of the attempts are shown as dots.
-Two important features were observed; one is subject specific variability and the other is random slope with attempts which indicates subjects specific learning trend.
-Some subjects show improvement over time and some show the decrease in performance.
-Although, the model fit experiment~6 suggested a statistically significant learning curve the plots indicates that it is minimal.
-
-\begin{figure}[htbp] 
-   \centering
-    \includegraphics[width=6.3in]{learning_trend_subject.pdf} 
-   \caption{ Least square lines fitted through the subject specific residual proportion correct obtained from model~\eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
-Subject specific positive and negative slopes are observed.
-Mean residuals are shown as dots and least square regression lines fitted through the points show no overall learning trend in each of the three experiments.}
-   \label{fig:learning_trend_response}
-\end{figure}
-
-Table~\ref{tbl:model_result_time} presents the results of model~\eqref{eqn:trend_time} where response time is examined with respect to attempt.
-Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated.
-Interestingly, time to respond significantly decreases as number of attempts increases, for all three experiments.
-The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments.
-The negative estimates suggest that on average later attempts took less time.
-Even though observers did not improve in their performance in successive attempts, they became more efficient in their response.
-The parameter $\alpha_1$ for first attempt is also highly significant.
-The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts.
-One explanation might be that participants take the chance to  carefully read through instructions and familiarize themselves  with the types of plots used in the experiment before evaluating their first lineup, which would inflate the time taken on the first lineup.
-Each of the experiments asks observers to give a reason for their choice (out of a list of pre-defined answers.
-For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).
-Later pages of the web site were identical  except for the lineup.
-
 % latex table generated in R 3.4.1 by xtable 1.8-2 package
@@
    \cline{2-5} \cline{7-10} \cline{12-15}
-\end{tabular}}
-\end{table}
-
-% 
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 23:37:55 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{rrrrllrrrllrrrl}
-%   \hline
-%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
-%   \hline
-% $\mu$ & 3.80 & 3.72 & 3.88 & *** &   & 3.89 & 3.83 & 3.96 & *** &   & 3.75 & 3.64 & 3.85 & *** \\ 
-%   $\alpha_1$ & 0.32 & 0.25 & 0.39 & *** &   & 0.34 & 0.28 & 0.40 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
-%   $\alpha$ & -0.04 & -0.05 & -0.03 & *** &   & -0.04 & -0.05 & -0.03 & *** &   & -0.03 & -0.04 & -0.02 & *** \\ 
-%   $\sigma^2_u$ & 0.52 &  &  &  &   & 0.49 &  &  &  &   & 0.37 &  &  &  \\ 
-%   $\sigma^2_a$ & 0.03 &  &  &  &   & 0.05 &  &  &  &   & 0.05 &  &  &  \\ 
-%   $\sigma^2_l$ & 0.09 &  &  &  &   & 0.20 &  &  &  &   & 0.24 &  &  &  \\ 
-%   $\sigma^2$ & 0.46 &  &  &  &   & 0.50 &  &  &  &   & 0.45 &  &  &  \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
-
-To visualize how the time taken reduces over the successive attempts, we fitted model~\eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
-Least square regression lines are fitted through the subject specific residuals.
-Subject specific slopes are much different in each of the three experiments as we see in Figure~\ref{fig:learning_trend_time}.
-Some subjects improved over attempts by taking less time in the later attempts while others got worse.
-The averages of these residuals for each attempt are plotted as dots.
-Least square regression lines are fitted to these points excluding the first attempt since for first attempt we fitted an indicator covariate.
-The downward trends are evident in the plots.
-All the slopes are highly significant.
-As expected we observed large positive residuals for each of the experiments for first attempt.
-
-\begin{figure}[htbp] 
-   \centering
-%   \includegraphics[width=6.5in]{learning_trend_time.pdf}    
- \includegraphics[width=6.3in]{learning_trend_time_subject.pdf}    
-   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting model~\eqref{eqn:trend_time} without covariate attempt.
-Differences in subject specific slopes are observed.
-Some of the subjects did worse over successive attempts while others did better.
-Averages of these residuals are plotted as dots and least square regression lines are fitted to obtain overall trends.
-For all the three experiments the overall downward slopes are statistically significant which indicates that MTurk workers become more efficient as they progress through their attempts.}
-   \label{fig:learning_trend_time}
-\end{figure}
-
-% \subsection{Location Effect}
-% 
-% A total of 111 subjects was recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
-Each subject evaluated two lineups;  one for Interaction effect and the other for Genotype.
-In total there were 222 responses for 50 lineups.
-The data on the test lineup were excluded from the analysis.
-% 
-% We are investigating location effect in two ways: (1) location effect of each individual plot location, and (2) location effect of inside/outside panel positions.
-% 
-% Figure~\ref{fig:location_effect} shows an overview of detection rates for both of the effects studied in experiment~9.
-Each dot corresponds to one lineup.
-The size of each dot gives the number of evaluations.
-Lineups with  the same set of null plots are connected by lines.
-% %The detection rate for each data plot location is shown in colored by null sets. 
-% The overall average detection rate for each location is shown using dashed lines.
-Size of the points represent the number of responses.
-For some locations we have as many as 10 responses.
-For location 1, we did not have any responses for null set 1 in one of the interaction lineups.
-% 
-% We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets.
-The lineup protocol is using a finite, small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set.
-If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate.
-From the plot we can see that null set 3 appeared to be more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot.
-This null set effect was tested, and a significant difference was found for null set 3 of the interaction lineup.
-The rest of the null sets do not show any differences at  a 1\% significance level.
-% 
-% 
-% \begin{figure}[htbp] 
-%    \centering
-%     \includegraphics[width=6.5in]{proportion_nulls_guide.pdf} 
-%    \caption{Location of data plot in the lineup and proportion correct (detection rate) for both Interaction and Genotype effect.
-Each colored line represents a null set and the size of the dots represents number of responses.
-The overall average proportions are shown by dashed lines.
-The actual data plot locations are shaded grey on the top panels to demonstrate their relative positions on a lineup.}
-%    \label{fig:location_effect}
-% \end{figure}
-% 
-% 
-% Model~\eqref{manova} was fit to the data to test, if  detection rates were significantly different for the different locations.
-For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R}.
-The results are shown in Table~\ref{tbl:manova}.
-The $p$-values for both the Interaction and the Genotype effect are much bigger than the conventional threshold of 0.05 suggesting that location of the actual data plot in the lineup does not significantly affect performance.
-% 
-% \begin{table}[hbtp]
-% \caption{The summary of the results obtained by fitting MANOVA model~\eqref{manova} using Wilks test.}
-% \begin{center}
-% \begin{tabular}{ccccccccc}
-%   \hline \hline
-%  Location & && & \multicolumn{3}{c} {Degrees of Freedom}  & F test \\
-%  \cline{5-7}
-%  Effect & DF & Wilks & Approx. F & Numerator& Denominator &Residual & $p$-value\\
-%   \hline
-%   Interaction& 3&0.07986 & 0.592 &15& \phantom{0}5.92 & 6 & 0.8082 \\ 
-%   Genotype & 4&0.01689 & 1.722 & 20&14.22 & 8 & 0.1488 \\ 
-%    \hline
-% \end{tabular}
-% \end{center}
-% \label{tbl:manova}
-% \end{table}
-% 
-% For the second approach in investigating the location effect, we distinguish between inner panels 7, 8, 9, 12, 13, 14 and outer panels (any panel with an outside edge).
-From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype.
-Model~\eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.
-
-% ===================================================================================
-\section{Conclusion}
-%------------------------------------------------------------------------------------
-
-Human demographics have the potential to influence performance when the lineup protocol is used for statistical inference.
-In our study of the demographic effects on performance across a set of different experiments we found some statistically significant effects.
-Age group 36-40, people who have a graduate degree have a significantly higher detection rate.
-India shows significantly lower detection rate compared to United States.
-Gender does not have any significant effect.
-However, the effects are minimal, on the order of a few percentage points different from the average.
-These results are very important for the power of visual tests as they demonstrate the robustness of the test against different human factors.
-
-Individual learning trend is observed in time taken but not so much in observer performance.
-Some individuals improved on their performances while others showed a decrease on successive attempts.
-This could be interpreted as good, in the sense that it means that observers could realistically be recruited from sources like MTurk, and that substantial training is not necessary in order to obtain useful evaluations of lineups in practice.
-However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea, at least for short term learning.
-
-The simulation experiment reveals that there is no significant effect of location on the actual data plot in the lineup.
-This is important as the visual statistical inference procedure prescribes that the data plot be placed at random in the lineup.
-This paper suggests that any random place in a lineup is as good as other places in the lineup.
-Even though there are variations on the performance depending on different null sets, their impact on probability to correctly evaluate a lineup is very negligible.
-
-There are many more ways that the lineup protocol might be tested to learn what we might expect about its performance for tackling real data mining problems.
-Possible changes in the lineup protocol in the pipeline are allowing observers to select more than one plot, and to vary the size of the lineup from 20 to a smaller number, and have observers all see different lineups with different null sets.
-This paper assessed the effect of human factors on the experiments conducted to date.
-As changes to the lineup protocol are suggested by other experiments the human factor effects may need to be examined again.
-
-\if1\blindDoc
-{
-%------------------------------------------------------------------------------------
-\paragraph{Acknowledgments}
-%------------------------------------------------------------------------------------
-This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
-} \fi
-
-\bibliographystyle{asa}
-\bibliography{references}
-
-% ---------------------------------------- Appendix ----------------------------------
-
-\appendix
-
-% ===================================================================================
-\section{Experimental Methods}\label{sec:exp_design}
-%------------------------------------------------------------------------------------
-
-Two of the factors, signal in the data and individual abilities, were studied in \cite{majumder:2013}.
-The choice of visual test statistic was examined in \cite{heike:2012}.
-In each of these analyses demographic factors were given a cursory glance, to ensure that they did not have large effects on the results.
-The design of experiments 5, 6 and 7 enables the examination of learning trend, which is studied in this paper.
-Experiment~9 was a real test case for visual statistical inference, and in order to understand the significance of the structure in the genomic data, multiple lineups were made in which location of the actual data plot, and the sample of nulls, were randomized.
-This enables the assessment of the effect of these factors on the results.
-This section describes the experimental methods used to examine the effects of demography, placement of the actual data plot, sample of null plots and the existence of a learning trend.
-
-\subsection{Demographic Factors}
-\noindent
-For all  ten experiments shown in Table~\ref{tbl:visual_stat}, the following demographic information was collected from subjects:
-
-\begin{enumerate} \leftmargin 5cm  \itemsep 0in
-\item {\it Age group}, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, above 50.
-\item {\it Gender}, male or female.
-\item {\it Education level}, with levels being high school or less, some undergraduate courses, undergraduate degree, some graduate courses, and graduate degree.
-\item {\it Geographical location}, collected from the IP address of the participants' computer, as latitude, longitude, city and country. 
-\end{enumerate}
-
-Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
-The factors are examined in association with the observer's response using a logistic regression with random effect terms:
-
-\begin{equation} \label{eqn:demographic_response}
-g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
-\end{equation}
-
-where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ picks the actual data plot from lineup $j$, $\mu$ is an overall population average, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively, for observer $i$.
-The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume independence and normality of the errors, i.e.\ $\ell_j \sim N(0, \sigma_\ell^2)$.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-
-Similarly, we model the  time  an observer takes to identify a panel from a lineup.
-Let $Z_{ij}$ denote the logarithm of time taken for  observer $i$ to evaluate  lineup $j$.
-Let $\mu_{ij}=  E(Z_{ij})$ be the average of the (log) time taken by  observer $i$ to pick a panel from lineup $j$.
-We model this in a mixed effects model of the same structure as model~(\ref{eqn:demographic_response}) given as:
-%\hh{
-\begin{equation} \label{eqn:demographic_time}
-Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
-\end{equation}
-where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model~\eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.
-
-\subsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive responses when multiple lineups are shown for evaluation.
-Experiments 5, 6 and 7 were used for this.
-Each subject was shown a total of 10 lineups randomly selected from a pool of lineups.
-The lineups are not necessarily of the same difficulty level, but the order of lineups was randomized.
-The responses of the lineups were recorded by attempt 1 through 10.
-Attempt 1 means that the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
-The goal is to estimate whether performance of the observer improves, or changes, from attempt 1 to attempt 10.
-
-It should be noted that we are examining the observer's performance, when we model response as detected or not, but this is not the goal of visual inference.
-Visual inference is constructed to measure the significance of structure discovered in data.
-It is expected that some observers will be more skilled at reading data plots, and hence, more readily detect the plot that is different.
-It is also expected that as observers gain experience in evaluating lineups that they become more proficient in reading data plots, particularly if feedback is given on whether the actual data plot was chosen or not.
-Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups.
-So in this context, detected, or not, is used as a response to examine individual differences.
-
-Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k$th evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
-Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the actual data plot from lineup $j$ in their $k$th attempt.
-Learning trend is assessed using a generalized mixed effects model of the form
-\begin{equation} \label{eqn:trend_response}
-g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
-\end{equation}
-where $\mu$ is an overall population average, $\alpha_k$ is the effect of the $k$th attempt on the probability, using the first attempt as reference, $\alpha_1 = 0$, and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
-The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
-The term $a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts, where $a_i \sim N(0, \sigma_a^2)$. 
-For $\ell_j$ a normal distribution, $N(0, \sigma_\ell^2)$, is assumed, and $\ell_j$ is a random intercept predicting lineup difficulty level.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-The inverse link function, $g^{-1}(.)$, from equation~\ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in the $k$th attempt by a single observer as
-\begin{equation} \label{eqn:trend_power}
-\hat p_{ijk} =  g^{-1}(\hat{\mu} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
-\end{equation}
-
-When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
-Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of the (log) of time taken by  observer $i$ to choosing a  panel from lineup $j$ in his/her $k$th attempt.
-We evaluate this in a mixed effects model of the form
-
-\begin{equation} \label{eqn:trend_time}
-Z_{ijk} = \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
-\end{equation}
-where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$ is the average change in log time taken for each additional attempt,  $\alpha_1$ is an offset in log time taken for the first attempt.
-All other effects are random effects: $u_i$ is a subject-specific intercept representing individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$, $a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$, with $a_i \sim N(0, \sigma_a^2)$, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-Equation~\ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
-\begin{equation} \label{eqn:trend_time_est}
-\hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j.
-\end{equation}
-
-To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2015, lme4:paper}.
-We employ a normal approximation to obtain $p$-values corresponding to fixed effect parameters estimates.
-
-\subsection{Location Effect} \label{sec:location_design} Experiment~9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
-The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
-Two factors were of primary interest a main effect for genotype and an interaction effect between genotype and treatment condition.
-
-In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
-This overall significance is studied using visual inference  in \citet{tengfei:2013}.
-
-In order to study the effect that  location has on evaluating lineups, we used multiple lineups for each of the data plots.
-For each data plot, five sets of null plots were generated and the actual plot was randomly placed in one of five different locations in a lineup of size 20.
-For the genotype effect, the locations were  1, 8, 12, 17, 20 and locations 2, 9, 12, 16, 20 were used for the interaction effect.
-%Five different sets of null plots were used to produce 5 lineups for each location position, creating
-Overall this created a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
-Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to help clean the data.
-
-To examine if the difference in detection rate among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
-Let $\mathbf{Y}=(Y_1,Y_2, ...
-, Y_p)^\top$ be a vector of random variables with dimension $p = 5$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represent the $j$th vector response for  location $i$ with $i=1,2, ..., I$ for $U=5$.
-Because the same data plot is shown in each lineup, it is assumed that there could be some association between the responses for each null set, which suggests the MANOVA model rather than a univariate ANOVA.
-The MANOVA model
-\begin{equation}\label{manova}
-\mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
-\end{equation}
-where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$, tests for significant difference between the means.
-
-\subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through MTurk \citep{turk}.
-It is an online work place where people from around the world can sign up for so-called `HIT's, human intelligence tasks, generally short tasks that are humans are typically better at solving than computers.
-Usually tasks are very simple and no specialized training is required to do them.
-Tasks are designed for anyone to do but some tasks may require some skills depending on the recruiters' need.
-%Each task is usually planned to completed in a short time.
- For completing a HIT workers are paid a small amount of money, on the order of minimum wage in the USA.
-
-We designed and developed a web application which enables the display of lineups to  observers as per experimental need.
-The MTurk workers were re-directed to this web application to complete their assigned tasks.
-Responses were collected, stored automatically into a local database server, along with demographic information, age group, gender and education level.
-The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
-Location of an observer is determined based on the ip address of the observer.
-
-\section{\\Some tables and figures}
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=4.5in]{turker_country.pdf} 
-   \includegraphics[width=4.5in]{turker_location.pdf}    
-   \caption{Location of Amazon Mechanical Turk workers participating in the experiments. Most subjects came from India and the United States, but subjects from countries around the world took part.}
-   \label{fig:turker_location}
-\end{figure}
-
-\begin{table}[hbtp]
-\caption{Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
-\centering
-\scalebox{.85}{
-\begin{tabular}{rlrrcr}
-  \hline
-  && \multicolumn{2}{c}{Participants}&Average& Number of\\
-  \cline{3-4}
- Factor & Levels & Total &\% & Time & Responses \\ 
-  \hline
-Gender & Female & 989 & 42.26 & 43.69 & 10538 \\ 
-   & Male & 1351 & 57.74 & 48.75 & 13542 \\  
-\hline
-  Education & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
-   & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
-   & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
-   & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
-   & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
-\hline
-  Age & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
-   & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
-   & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
-   & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
-   & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
-   & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
-   & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
-\hline   
-  Country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
-   & India & 980 & 41.88 & 52.63 & 10238 \\ 
-   & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\
-
-   \hline
-\end{tabular}
-}
-\label{tbl:demographics}
-\end{table}
-
-% latex table generated in R 3.4.1 by xtable 1.8-2 package
-% Wed Oct 18 17:43:02 2017
-\begin{table}[ht]
-\caption{\label{tbl:anova_factor}Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable.
-Gender does not have a significant effect on detection rate, but does on time to respond.
-All factors significantly affect time to respond.}
-\centering
-\scalebox{0.9}{
-\begin{tabular}{lrrrrlrrrr}
-& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
-\cline{2-5} \cline{7-10} 
- & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
-Full  & 52609.1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 24140.5 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
- - Age & 52081.3 & 527.8 & 6 & $<$0.001 & \hspace{0.5cm} & 24113.2 & 27.3 & 6 & $<$0.001 \\ 
- - Country & 52245.5 & 363.6 & 2 & $<$0.001 &  & 24114.8 & 25.7 & 2 & $<$0.001 \\ 
- - Degree & 52498.7 & 110.4 & 4 & $<$0.001 &  & 24116.0 & 24.5 & 4 & $<$0.001 \\ 
- - Gender & 52540.3 & 68.8 & 1 & $<$0.001 &  & 24138.9 & 1.6 & 1 & 0.200 \\ 
-\cline{2-5} \cline{7-10}
 \end{tabular}}

### 2024-12-30 wiley_paper/wiley_paper.Rnw
Add files via upload

--- 2017-10-21 submitted-v1/sociology_paper.Rnw
@@
 
-The lineup protocol introduced in \citet{buja:2009} can be used to quantify the significance of graphical findings in the exploratory data analysis process.
-This methodology is  part of what is called visual statistical inference, which has been developed further  and validated by \citet{majumder:2013} in simulation studies of head to head comparisons with conventional inference.
-One of their major contributions   is to define the power of  visual tests and to provide a method for obtaining the power for a particular lineup.
-In some scenarios, the power of a visual test  was seen to outperform that of a conventional test.
+The lineup protocol introduced by \citet{buja:2009} quantifies the significance of graphical findings in exploratory data analysis, forming part of visual statistical inference.
+\citet{majumder:2013} extended and validated this approach through simulation studies, showing that visual inference can sometimes outperform conventional methods.
+They also defined the power of visual tests and proposed a way to calculate it for a given lineup.
+In visual inference, the test statistic is a plot of observed data, randomly placed among null plots generated under a null hypothesis, often assuming no structure.
+Observers evaluate the lineup, and if they consistently identify the data plot, this provides evidence against the null hypothesis, akin to a conventional hypothesis test.
 
-In visual inference, the test statistic is a plot of the observed data, called the {\it (actual) data plot}.
-To create a lineup the data plot is placed randomly in a grid of null plots.
-Null plots  are rendered from data generated as  specified by a null hypothesis.
-Often, this null hypothesis assumes independence or no structure in the data.
-An observer is then asked to evaluate the lineup.
-If the  actual data plot is detected by the observer, this counts as evidence against the null hypothesis.
-Based on the number of evaluations by independent observers and the number of times that the data plot is identified, we are  able to evaluate the significance of the graphical relationship in the same way and the same rigor as a conventional hypothesis test.
-
-\begin{figure}[hbtp] 
-   \centering
-   \includegraphics[width=0.99\textwidth]{electoral-5-13.pdf} 
-   \caption{ \label{fig:elect} Which one of the plots, labelled 1 through 20 from top left to bottom right, is the most different from the others?}
-\end{figure}
-
-Figure~\ref{fig:elect} displays an example lineup, one of the plots is based on the observed data, while the remaining 19 plots show  data generated from a null model.
-Which one of the 20 plots is the most different from the others?
-When asked this question, 12 out of 72 observers picked the  data plot located in panel number $4^2-3$ in the lineup\footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
-The corresponding $p$-value is 0.0077, indicating sufficient evidence to reject the null hypothesis.
-
-How do we interpret this finding, though? For that, we need to know the context of the data and we need to have more information about how the null plots were generated. 
-This particular example investigates the results from the 2012 US presidential election in comparison to  poll results just prior to the election. 
-Although this example is more simplistic than most of the tests conducted to date, it serves the purpose of illustrating the lineup protocol.
-The data consists of the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
-Each panel in Figure~\ref{fig:elect} shows an `electoral building' \citep{mosley2010} where each state in the union is represented by a rectangle.
-The difference in poll results is plotted horizontally, while the height of each box corresponds to the number of that state's electoral votes.
-Color indicates party affiliation.
-The null hypothesis is that ``the election results are consistent with the polls''.
-Therefore the polling results provide the null model from which data is simulated.
-A normal model with mean and standard deviation based on a poll and its margin of error is used to simulate different scenarios that might have resulted on election day, if the polls were on target.
-Each null data set is generated as a set of draws from this model.
-% normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
-These samples are plotted as electoral buildings, and the plot based on the actual election results is placed randomly among the null plots in a lineup of size $5\times 4$.
-If the null hypothesis is true, the  data plot should look  like any of the other plots, and not be identifiable by an observer.
-Figure~\ref{fig:tower} shows the plot of the electoral building with the additional context information and labels.
-
-\begin{figure}[hbtp] 
-   \centering
-   \begin{minipage}[c]{0.49\textwidth}
-   \includegraphics[width=\textwidth]{tower.pdf}
-   \end{minipage}
-    \hfill
-   \begin{minipage}[c]{0.45\textwidth}
-   Poll aggregator 1 \\
-      \includegraphics[width=\textwidth]{freedom-sim.pdf} 
-   Poll aggregator 2\\
-      \includegraphics[width=\textwidth]{rcp-sim.pdf} 
-   \end{minipage}
-   \caption{ \label{fig:tower} Electoral building plot of the results of the 2012 U.S.
-Presidential Election (left).
-On the right two histograms showing simulated poll results for the Democratic presidential candidate using  polling averages collected from two different sources, Freedom's Lighthouse Averages (top), RealClearPolitics (bottom).
-If the polls matched the electoral result the simulated results would be centered at the electoral result.
-This happens for the RealClearPolitics polls (bottom) but not for Freedom's Lighthouse.
-For the latter, only 5.33\% of simulated results were as high as the actual election result, which converts to a $p$-value of 0.0533, indicating an almost statistically significant difference.
-It hints at the polls from this source under-estimating the vote for Obama.
-This source was used to generate the null plots in the lineup in Figure~\ref{fig:elect}.
-The $p$-value from the lineup protocol, was 0.0077, which more strongly concludes that these polls mismatched the actual result.}
-\end{figure}
-The lineup of Figure~\ref{fig:elect}, as well as the other lineups in the manuscript are created using the nullabor package by \citep{nullabor}.
-
-A lineup can be evaluated by a single person or multiple observers.
-A distribution similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario, is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis \citep{vinference}.
-To avoid expectation errors \citep{meilgaard}, particularly in emotionally charged areas such as election results, observers should not be aware of the data that constitutes a lineup, and should therefore not have seen the actual data plot before inspecting the lineup.
-This is the reason that we revealed the context of the election problem only after the interested reader had a chance to evaluate the lineup.
-
-When presenting a lineup to an observer the accompanying question should be phrased in as general a manner as possible, effectively asking the observer to pick the plot that is most different, and allowing observers to provide their own reasons for their choice.
-This ensures all possible deviations from the null hypothesis to be detected.
-However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments in \cite{majumder:2013} or \cite{tengfei:2013}, these questions have to be phrased much more specifically, in order to avoid  type III errors \citep{mosteller:48} consisting of correctly rejecting the null hypothesis for a wrong reason.
-In those experiments, the structure in the data is strictly controlled in the simulation process, which allows for specific questions to be asked.
-
-In the election example, observers were asked ``which plot is the most different?''.
-The type of plot, showing two (modified) stacked bar charts in different colors should suggest to the observer that the interesting feature is the difference between the two heights.
-Most observers based their decision on this -- besides  `asymmetry' between the towers (a reason suggested by us), most free-form responses for the reason of choice describe a comparison between the red and blue towers.
-In the responses for this lineup we see something that is very typical of responses in lineups: it is only a few panels from the lineup that observers pick, while most panels do not get picked or get picked only a few times.
-Out of a total of 72~responses, 23 observers picked panel~\#8, 12 observers picked the (data) panel \#13, 10 observers picked panel~\#4, 6 observers picked panel \#7, and all of the remaining panels were picked at most three times.
-From the panels picked, we see that the top three choices  are all extreme with respect to heights of blue versus red towers:  both panels \#8 and \#13 show a large difference between the height of the towers, while panel \#4 is the only panel, in which the red tower is higher than the blue one.
+A lineup can be evaluated by one or more observers, and a distribution similar to the binomial, adjusted for dependencies from the lineup scenario, is used to calculate the $p$-value based on how often the actual data plot is identified \citep{vinference}.
+To avoid expectation bias \citep{meilgaard}, observers must not know the data constituting the lineup or have seen the actual data plot beforehand.
+The accompanying question in a lineup should be broad, asking observers to pick the most distinct plot, allowing detection of any deviations from the null hypothesis.
+However, in head-to-head tests with conventional methods \citep{majumder:2013, tengfei:2013}, specific questions are needed to avoid type III errors \citep{mosteller:48} consisting of correctly rejecting the null hypothesis for a wrong reason.
 
@@
 
-\begin{table*}[hbtp] 
-\centering 
-\caption{Overview of 10 different Turk experiments, from which data was collected to study human factor effects.
-All of the experimental data were used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.}
-\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
-\hline\hline 
- & Experiment &  Test Statistic  & Lineup question & Used in study of\\ [0.5ex] % inserts table %heading 
-\hline 
-1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
-between group A and B? & DF\\
-2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
-& DF\\
-  3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
-& DF\\
- 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
-  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
-  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?
-&  DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
-  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?
-& DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
-  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?
-& DF\\
-  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?
-& DF  \hfill \phantom{LT} \hfill {LE} \hfill \phantom{.}\\
-  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?
-& DF\\
+We conducted several experiments (see Table~\ref{tbl:visual_stat} in the Appendix~\ref{sec:tbl_fig}) using Amazon's Mechanical Turk~\citep{turk} (MTurk) for validating the lineup protocol, comparing plot designs, and evaluating data analysis structures.
+Demographic data such as gender, age, and educational background were collected to assess the influence of these covariates on lineup evaluations.
+Additionally, experiments investigated the short-term learning trends.
 
-\hline 
-\end{tabular} 
-\label{tbl:visual_stat} 
-\end{table*}
-
-We have conducted a series of experiments (see Table~\ref{tbl:visual_stat} for an overview)
-%There have been ten experiments conducted 
-using Amazon's Mechanical Turk~\citep{turk} (MTurk) for a variety to purposes: validating the protocol against existing tests, comparing plot designs, and evaluating structure in data analysis problems.
-In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.
-Mining this data will provide a basis for evaluating
-%are being used to evaluate 
-the effects of different covariates on an individual's ability to evaluate lineups.
-The design of some of the experiments allow a further investigation into aspects beyond participants' demographics: experiments~5, 6 and 7 can be used to investigate short term learning trends, while experiment~9 allows insight on whether the  positioning of the  data plot in the lineup has an effect on detection rate.
- %Each of these experiments collected demographic details of the subjects. 
-Section~\ref{sec:factor_performance} discusses human factors that we expect to be influential on the performance of the observer.
-Section~\ref{sec:exp_design} describes the methods used to assess the effects, and Section~\ref{sec:result_socio} summarizes the results of what we have learned about human factors affecting lineup evaluation.
+Section~\ref{sec:factor_performance} discusses expected human factors influencing observer performance and Section~\ref{sec:result_socio} presents findings on factors affecting lineup evaluations.
+The methods used to assess the effects are oulined in Appendix~\ref{sec:exp_design}.
 
@@
 \begin{itemize} \itemsep 0in \parsep-10pt %\parskip0pt
-%\subsection{Choice of Visual Test Statistic}  
-\item{\bf Choice of Visual Test Statistic:}  
-%There are a variety of ways that
-Data can be plotted in a variety of ways and we can enhance plots by axes, legends or representations of models.
-The choice of plot primarily should be appropriate for the problem being investigated, as the data plot is the visual test statistic.
-When for example studying the association between two quantitative variables a scatterplot is usually considered best practice.
-Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.
-Some choices of visual test statistics provide better results than others when studying the same problem \citep{heike:2012}, because some plot types make it easier for certain structures to be seen.
 
-Table~\ref{tbl:visual_stat} shows the large variety of visual test statistics used in the Turk studies completed so far.
-These experiments were conducted to examine many different visual tasks.
-Side-by-side boxplots were used in experiment~1 to study the importance of a categorical variable in a regression model.
-For both experiments~2 and~3, a scatterplot is the visual test statistic, but a regression line fitted through the points is overlaid for experiment~2.
-The experiments studied slightly different problems, for which the difference in plot type was important.
-Experiment~4 compared polar coordinates to euclidean coordinates on the reading of structure in a barchart.
-Experiment~5 examined the style of plot for a bivariate data problem, by comparing results when using a scatterplot, side-by-side dot plots, overlaid densities or side-by-side boxplots.
-Experiment~6 studied the difference between variations of boxplots.
-Experiment~7 studied the effects of high dimension low sample size on the separation of groups in low-dimensional projections.
-The sine illusion, where people read the strength of differences between curves as different when there is a strong sinusoidal structure, is examined in experiment~8.
-Experiment~9 was conducted to test if there was any structure in an RNA-seq data set measured on soybeans, and used interaction plots classically constructed to examine interaction effects in $2\times 2$ factorial designs.
-Experiment~10 used different sorts of residual plots for hierarchical linear models as the visual test statistics.
-
-In experiments~4,~5 and~6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices, the use of lineup tests for comparisons and  examples are discussed in \citet{heike:2012, qqplots:2015}
-
-%\subsection{Signal in the Data}
-\item{\bf Signal in the Data:}
-A visual test statistic is chosen for a particular task.
-If it is well-designed  departures from the null hypothesis will be visible in the plot, especially if the signal is strong.
-An important factor that helps an observer to identify the actual data plot in a lineup is the strength of the signal in the data.
-Simulated data was generated for all of the experiments, except exps.~4 and~9, enabling the study of signal strength and actual data plot detection.
-In experiment~4 we took a different approach by varying the sampling size, which also allowed us to investigate the effect a design had on detection rate while varying signal/noise ratio.
-Primarily the results were as expected, that as the signal strength increased observers more frequently picked the actual data plot as different from the null plots.
-
-\item{\bf Question Design:} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
-Visual inference supports discovery of structure, enabling the unexpected to surprise us.
-This is tightly connected to the task that the human observer is asked to perform.
-For most purposes the observer is asked a very general question, such as to pick the plot that is most different from the others, and explain the reasons why they see it as different.
-By being general, all possible discoverable structures are included in the significance calculations \citep[see also ][]{buja:2009}.
-
-In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted.
-This was important to allow a direct comparison of the results to those from conventional testing.
-%in order to make a comparison of the results.
-For latter studies (4, 7, 10), questions posed to the observer were more general. Table~\ref{tbl:visual_stat} lists the questions used for each experiment.
-
-%\subsection{Demographics} 
 \item{\bf Demographics:}  During each of the experiments, data on age, gender, education level and geographic location was collected.
@@
 The IP address of the computer afforded the geographical location of the subject.
-The purpose of collecting this information with each experiment is to examine the effect that they have on the results of visual inference -- ideally very little.
 
@@
 
-%\subsection{Location of Actual Data Plot in the Lineup}
-\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 panels is used.
-A random number generator is used to determine the position where the actual data plot is placed in each lineup.
-Eye tracking experiments \citep{zhao:2012} suggest that some observers traverse lineups in horizontal direction, while others have a vertical up-and-down approach to their search.
-Almost universally, observers start at the top left of a lineup.
-This leads to some concern that observers might be able to more easily identify the actual data plot, if it is placed at the top left of a lineup than when it the actual data is placed at the bottom right.
-%With the help of eye-tracking equipment \citet{zhao:2012} tracked the observers' eyes to see their approach in searching the plots in a lineup.
-The results suggest that many people read lineups from left to right, some read from upward to downward.
-%There is some concern that if the actual data plot is located in the top left of the lineup, observers may be able to identify it more easily than when the actual data plot is at the bottom right. 
-Ideally, this does not happen.
-Experiment~9  is designed to allow us to investigate the effects of location on detection.
-Five different locations in the lineup were used to assess how fast and accurately observers identified the actual data plot.
-
-\item{\bf Sample of Null Plots:} In classical inference, the test statistic is compared to the full null distribution, to decide if it is extreme or not.
-In visual inference the actual plot is compared to plots of a finite number of samples from the null distribution.
-In the lineups used in the Turk experiments, the actual data plot is compared on 19 null plots.
-These null samples are random draws, and there is a chance that one or more null plots in the lineup may be similar or even more extreme than the actual data plot.
-The sample of null plots can affect the observer's decision.
-This is tested with the data from experiment~9, where the experiment was set up with lineups made from different null plots.
-
 %\subsection{Individual Skill or Ability} 
@@
 We can examine the effect of individuals' skills and abilities because multiple subjects evaluated the same lineups, which allows us to estimate, if some subjects consistently detect the actual data plot more often than others.
- %For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.
-This influences the response of the observer.
-Also subject specific variation in the power of visual test is observed in \cite{majumder:2013}.
+
 \end{itemize}
+
+% ===================================================================================
+\section{Results}\label{sec:result_socio}
+%------------------------------------------------------------------------------------
+
+ \subsection{Overview of the Data} In ten experimental studies, a total of 2,340 participants from 70 countries, primarily the United States and India, provided feedback data on lineups, ensuring a diverse pool of respondents in terms of geography, gender, age, and education.
+There were a high number of female participants(42\%) across most countries, which was an unexpected but positive outcome(see Table~\ref{tbl:turker_location} in Appendix~\ref{sec:tbl_fig}).
+Notably, countries like Canada, Romania, the United Kingdom, and Macedonia also had notable participant representation.
+
+The majority of participants were aged 18 to 35, with the United States showing a more balanced age distribution compared to India, where younger individuals were more prevalent.
+Educational backgrounds were diverse, with 38.42\% of participants reporting a graduate degree.
+However, there was a cultural misunderstanding as Indian participants classified any university degree as a graduate degree, while the North American definition refers to a Master’s degree or higher.
+Compared to the 2010 US Census \cite{census}, the participants skewed younger and more highly educated.
+
+A total of 1,912 lineups were evaluated across the studies, with each participant reviewing at least 10 lineups, except for experiment~9, where they reviewed three.
+To ensure quality, participants were given test plots to evaluate first, and incorrect responses led to their data being excluded from the analysis.
+Some demographic and geographical data were missing due to participants opting out or issues retrieving IP-based location data.
+
+\subsection{Demographic Factors}
+
+To explore the significance of demographic factors in detecting the data plot in the lineup, a mixed-effects model~\eqref{eqn:demographic_response} is fitted to the data, with age, country, education, and gender as fixed effects.
+To estimate the significance of each of the factors, reduced models are fitted with that factor removed from the model.
+Table~\ref{tbl:model_result_demographics} summarizes the results.
+Except for gender, all of the demographic factors show significant differences in  detection rates.
+With respect to time taken, all of the demographic factors are significant.
+
+% latex table generated in R 3.4.1 by xtable 1.8-2 package
+% Wed Oct 18 17:22:24 2017
+\begin{table}[hbtp]
+\caption{\label{tbl:model_result_demographics}Parameter estimates of models~\eqref{eqn:demographic_time} and~\eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively.
+For time taken all the demographic factors are significant.
+For detection rate, age group 36-40, rest of the world and graduate degree are significantly different.
+For gender no difference in performance is observed.
+Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
+\centering
+\scalebox{.9}{
+\begin{tabular}{rrr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
+ && \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
+
+\cline{3-6} \cline{8-11} 
+\multicolumn{2}{c}{{Demographic Factor} \hfill Level} & Est & 2.5\% & 97.5\% &   && Est & 2.5\% & 97.5\% &   \\ 
+\cline{3-6} \cline{8-11}
+\bf Fixed Effects & Average $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.63 & -0.81 & -0.45 & *** \\ [3pt]
+\bf Age Category $(\alpha)$ & 18-25 & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
+&  26-30 & 0.05 & 0.03 & 0.07 & *** &   & 0.03 & -0.06 & 0.13 &  \\ 
+&  31-35 & 0.06 & 0.03 & 0.09 & *** &   & 0.12 & 0.02 & 0.23 & * \\ 
+&  36-40 & 0.22 & 0.19 & 0.25 & *** &   & 0.29 & 0.17 & 0.41 & *** \\ 
+&  41-45 & 0.17 & 0.13 & 0.21 & *** &   & 0.19 & 0.03 & 0.34 & * \\ 
+&  46-50 & 0.28 & 0.23 & 0.32 & *** &   & 0.20 & 0.01 & 0.39 & * \\ 
+&  above 50 & 0.34 & 0.31 & 0.38 & *** &   & 0.16 & 0.02 & 0.29 & * \\ [3pt]
+\bf Country $(\kappa)$& U.S. & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
+&  India & 0.23 & 0.20 & 0.25 & *** &   & -0.19 & -0.28 & -0.10 & *** \\ 
+&  Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.06 & -0.05 & 0.18 &  \\ [3pt]
+\bf Education $(\tau)$& High School & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
+&  Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & 0.00 & -0.14 & 0.14 &  \\ 
+&  Under grad degree & 0.14 & 0.10 & 0.17 & *** &   & 0.16 & 0.03 & 0.29 & * \\ 
+&  Graduate courses & 0.02 & -0.02 & 0.06 &  &   & -0.13 & -0.29 & 0.03 &  \\ 
+&  Graduate degree & 0.13 & 0.10 & 0.16 & *** &   & -0.04 & -0.17 & 0.09 &  \\ [3pt]
+\bf Gender $(\gamma)$& Female & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
+&  Male & 0.08 & 0.06 & 0.10 & *** &   & 0.05 & -0.02 & 0.12 &  \\
+\cline{3-6} \cline{8-11}
+\bf Random Effects &  lineup $(\sigma_\ell)$& \multicolumn{5}{l}{0.29} & \multicolumn{4}{l}{\phantom{-}2.27} \\ 
+                   &  Error $(\sigma)$& \multicolumn{5}{l}{0.69} & \multicolumn{4}{l}{\phantom{-}0.89}  \\ 
+\cline{3-6} \cline{8-11}\\
+\hline
+\multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
+\end{tabular}}
+\end{table}
+
+The first level of each factor serves as the baseline for the model.
+The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for other age groups except for 26-30.
+Participants from the rest of the world have a similar detection rate to those from the USA, but participants from India show a significantly lower detection rate.
+Subjects reporting to have  an undergraduate degree had a significantly higher detection rate.
+For time to respond, all age groups were significantly faster than  18-25 year olds.
+Participants from India as well as the rest of the world were faster than US Americans.
+A higher degree corresponded to longer time, and males were on average faster than females.
+
+Figure~\ref{fig:demographic_effect} supports the model results.
+The average time taken to respond (natural log) and detection rate are computed by lineup for the different demographic factor levels, and displayed using boxplots.
+Mean values are shown as dots inside the boxplot, as this statistic rather than the median is used in the models.
+The means and medians for response time (top row) are essentially equal.
+There are some differences between the two in the detection rate plots (bottom row).
+The differences in these values for different factors, that was tested by the models, can be seen in the plots.
+However, particularly for detection rate, the variation is huge.
+A large component of the variation is difficulty of the lineup.
+In some lineups the actual data plot was distinctively different from the null plots, and we would expect that the detection rate would be close to 100\%.
+In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond.
+
+\begin{figure}[htbp] 
+   \centering
+   \includegraphics[width=4in]{demographic_effect.pdf} 
+   \caption{Boxplots of average log time taken and proportion correct responses (detection rate) of all the lineups plotted for each demographic factor levels.
+The dots inside the boxes represent means.
+Some differences in means of various demographic factors are observed.
+Variability in detection rate indicates large variability in lineup difficulties.
+}
+   \label{fig:demographic_effect}
+\end{figure}
+
+In Table~\ref{tbl:model_result_demographics} notice that lineup specific error is estimated as 2.27 which is much higher than any of the other parameter estimates in the model.
+This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
+So although the demographic factors emerge from the model to be statistically significant, the practical significance is minimal.
+We illustrate this with the following example of graduate degree.
+
+While some of the demographic factors are strongly statistically significant, the main source of variation in detection rate is the lineup difficulty.
+For example, let's examine the effect of an undergraduate degree.
+To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with some graduate course work as compared to an undergraduate degree, for an average difficulty lineup (random effect of zero).
+Plugging in the relevant quantities to the fitted model gives a difference equal to:
+$$ \frac {\exp(-0.63-0.13)}{1+\exp(-0.63-0.13)}-  \frac {\exp(-0.63+0.16)}{1+\exp(-0.63+0.16)} = 0.319- 0.385 =-0.066.$$
+
+The person with some graduate course work picks the data plot  on average in 31.9\% of lineups of average difficulty, as compared to 38.5\% if they have an undergraduate degree.
+This difference is reduced to 4\% for a lineup with one standard deviation order of magnitude difference in difficulty.
+For two standard deviations it further reduces to 0.6\%.
+Thus, although there is a statistically significant difference in the proportion at which participants identify the  data plot for some demographic factors, these are not  significant in any practical sense.
+Figure~\ref{practical_impact_und_und_graduate.pdf} illustrates this example showing fitted models for a US 18-25 female with either a high school education or an undergraduate degree.
+Similar calculations show the same negligible impact of age level 36-40 (\Sexpr{round(exp(-0.63+0.29)/(1+ exp(-0.63+0.29)) - exp(-0.63)/(1 + exp(-0.63)),3)} at most)  and country India (\Sexpr{round(exp(-0.63-0.19)/(1+ exp(-0.63-0.19)) - exp(-0.63)/(1 + exp(-0.63)), 3)} at most) on the probability of a correct response.
+Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.
+
+\begin{figure}[htbp] 
+   \centering
+    \includegraphics[width=3.2in]{practical_impact_und_graduate.pdf} 
+    \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
+   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with an under graduate degree compared to a high school degree.
+Even though having an under graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.04 which is  negligible from a practical point of view.
+The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} = 2.27$)
+   of lineup variability.}
+   \label{practical_impact_und_und_graduate.pdf}
+\end{figure}
+
+\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments~5,~6, and~7 separately.
+Each participant evaluated 10 lineups in 10 attempts.
+In this model, attempt is fitted as a factor variable to allow for all possible non-linear learning trends.
+It should be noted that we also examined an alternative to model~\eqref{eqn:trend_response} where attempt was linearly fitted as a continuous covariate, but this also was not significant.
+
+The parameter estimates and $p$-values of the fixed effects from model~\eqref{eqn:trend_response} are presented in Table~\ref{tbl:model_result_response} in the Appendix~\ref{sec:exp_design}.
+Only for experiment~6 there is some evidence of a learning curve.
+There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.
+
+To visualize how  detection rates change over successive attempts, we fitted model~\eqref{eqn:trend_response} excluding the covariates related to attempt from the model and computed the residuals.
+Least square regression lines were fitted through the subject specific residuals as shown in Figure~\ref{fig:learning_trend_response}.
+The averages of these residuals for each of the attempts are shown as dots.
+Two important features were observed; one is subject specific variability and the other is random slope with attempts which indicates subjects specific learning trend.
+Some subjects show improvement over time and some show the decrease in performance.
+Although, the model fit experiment~6 suggested a statistically significant learning curve the plots indicates that it is minimal.
+
+\begin{figure}[htbp] 
+   \centering
+    \includegraphics[width=6.3in]{learning_trend_subject.pdf} 
+   \caption{ Least square lines fitted through the subject specific residual proportion correct obtained from model~\eqref{eqn:trend_response} fitted without attempt are plotted against attempt.
+Subject specific positive and negative slopes are observed.
+Mean residuals are shown as dots and least square regression lines fitted through the points show no overall learning trend in each of the three experiments.}
+   \label{fig:learning_trend_response}
+\end{figure}
+
+Table~\ref{tbl:model_result_time} in the Appendix~\ref{sec:exp_design} presents the results of model~\eqref{eqn:trend_time} where response time is examined with respect to attempt.
+Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated.
+Interestingly, time to respond significantly decreases as number of attempts increases, for all three experiments.
+The parameter $\alpha$ for fixed effect covariate attempt is highly significant in all the experiments.
+The negative estimates suggest that on average later attempts took less time.
+Even though observers did not improve in their performance in successive attempts, they became more efficient in their response.
+The parameter $\alpha_1$ for first attempt is also highly significant.
+The positive estimates of $\alpha_1$ indicates that first attempt made by an observer required much more time than any of the other attempts.
+One explanation might be that participants take the chance to  carefully read through instructions and familiarize themselves  with the types of plots used in the experiment before evaluating their first lineup, which would inflate the time taken on the first lineup.
+Each of the experiments asks observers to give a reason for their choice (out of a list of pre-defined answers.
+For each lineup evaluation, participants are also asked to state how confident in their choice they are (on a scale of 1 to 5).
+Later pages of the web site were identical  except for the lineup.
+
+To visualize how the time taken reduces over the successive attempts, we fitted model~\eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
+Least square regression lines are fitted through the subject specific residuals.
+Subject specific slopes are much different in each of the three experiments as we see in Figure~\ref{fig:learning_trend_time}.
+Some subjects improved over attempts by taking less time in the later attempts while others got worse.
+The averages of these residuals for each attempt are plotted as dots.
+Least square regression lines are fitted to these points excluding the first attempt since for first attempt we fitted an indicator covariate.
+The downward trends are evident in the plots.
+All the slopes are highly significant.
+As expected we observed large positive residuals for each of the experiments for first attempt.
+
+\begin{figure}[htbp] 
+   \centering
+%   \includegraphics[width=6.5in]{learning_trend_time.pdf}    
+ \includegraphics[width=6.3in]{learning_trend_time_subject.pdf}    
+   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting model~\eqref{eqn:trend_time} without covariate attempt.
+Differences in subject specific slopes are observed.
+Some of the subjects did worse over successive attempts while others did better.
+Averages of these residuals are plotted as dots and least square regression lines are fitted to obtain overall trends.
+For all the three experiments the overall downward slopes are statistically significant which indicates that MTurk workers become more efficient as they progress through their attempts.}
+   \label{fig:learning_trend_time}
+\end{figure}
+
+% ===================================================================================
+\section{Conclusion}
+%------------------------------------------------------------------------------------
+
+Human demographics have the potential to influence performance when the lineup protocol is used for statistical inference.
+In our study of the demographic effects on performance across a set of different experiments we found some statistically significant effects.
+Age group 36-40, people who have a graduate degree have a significantly higher detection rate.
+India shows significantly lower detection rate compared to United States.
+Gender does not have any significant effect.
+However, the effects are minimal, on the order of a few percentage points different from the average.
+These results are very important for the power of visual tests as they demonstrate the robustness of the test against different human factors.
+
+Individual learning trend is observed in time taken but not so much in observer performance.
+Some individuals improved on their performances while others showed a decrease on successive attempts.
+This could be interpreted as good, in the sense that it means that observers could realistically be recruited from sources like MTurk, and that substantial training is not necessary in order to obtain useful evaluations of lineups in practice.
+However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea, at least for short term learning.
+
+There are many more ways that the lineup protocol might be tested to learn what we might expect about its performance for tackling real data mining problems.
+Possible changes in the lineup protocol in the pipeline are allowing observers to select more than one plot, and to vary the size of the lineup from 20 to a smaller number, and have observers all see different lineups with different null sets.
+This paper assessed the effect of human factors on the experiments conducted to date.
+As changes to the lineup protocol are suggested by other experiments the human factor effects may need to be examined again.
+
+\if1\blindDoc
+{
+%------------------------------------------------------------------------------------
+\paragraph{Acknowledgments}
+%------------------------------------------------------------------------------------
+This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
+} \fi
+
+\bibliographystyle{asa}
+\bibliography{references}
+
+% ---------------------------------------- Appendix ----------------------------------
+
+\appendix
 
@@
 This enables the assessment of the effect of these factors on the results.
-This section describes the experimental methods used to examine the effects of demography, placement of the actual data plot, sample of null plots and the existence of a learning trend.
+This section describes the experimental methods used to examine the effects of demography, sample of null plots and the existence of a learning trend.
 
@@
 
-\subsection{Location Effect} \label{sec:location_design} Experiment~9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
-The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
-Two factors were of primary interest a main effect for genotype and an interaction effect between genotype and treatment condition.
-
-In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
-This overall significance is studied using visual inference  in \citet{tengfei:2013}.
-
-In order to study the effect that  location has on evaluating lineups, we used multiple lineups for each of the data plots.
-For each data plot, five sets of null plots were generated and the actual plot was randomly placed in one of five different locations in a lineup of size 20.
-For the genotype effect, the locations were  1, 8, 12, 17, 20 and locations 2, 9, 12, 16, 20 were used for the interaction effect.
-%Five different sets of null plots were used to produce 5 lineups for each location position, creating
-Overall this created a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
-Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to help clean the data.
-
-To examine if the difference in detection rate among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
-Let $\mathbf{Y}=(Y_1,Y_2, ...
-, Y_p)^\top$ be a vector of random variables with dimension $p = 5$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represent the $j$th vector response for  location $i$ with $i=1,2, ..., I$ for $U=5$.
-Because the same data plot is shown in each lineup, it is assumed that there could be some association between the responses for each null set, which suggests the MANOVA model rather than a univariate ANOVA.
-The MANOVA model
-\begin{equation}\label{manova}
-\mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
-\end{equation}
-where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$, tests for significant difference between the means.
-
 \subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through MTurk \citep{turk}.
@@
 % ===================================================================================
-\section{Results}\label{sec:result_socio}
+\section{Some tables and figures}\label{sec:tbl_fig}
 %------------------------------------------------------------------------------------
-
-\subsection{Overview of the Data} A total of 2340 participants provided feedback data on the lineups in ten different experimental studies.
-Figure~\ref{fig:turker_location} displays the locations of participants around the world.
-Most of the participants were from the United States and India.
-There were 70 other different countries represented.
-This provides a diverse pool of participants.
-The diversity in not only  geographic but also in gender, age group and education level as can be seen in Table~\ref{tbl:demographics}.
-The large number of female participants from all countries was a pleasant surprise to us.
 
@@
 
-Besides the United States and India, countries such as Canada, Romania, the United Kingdom and Macedonia have more than 10 participants each.
-The remaining 64 countries have fewer than 10 participants each.
-The distribution of participants is similar in all ten experiments.
-
 \begin{table}[hbtp]
-\caption{Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
+\caption{\label{tbl:turker_location}Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
 \centering
@@
 
-% 
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 16:56:32 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{llrrrr}
-%   \hline
-% var & lbls & subsjects & percents & avg\_time & response \\ 
-%   \hline
-% gender\_level & Female & 989 & 42.26 & 43.69 & 10538 \\ 
-%   gender\_level & Male & 1351 & 57.74 & 48.75 & 13542 \\ 
-%   degree & High school or less & 194 & 8.29 & 37.13 & 2253 \\ 
-%   degree & Under grad courses & 418 & 17.86 & 42.85 & 4068 \\ 
-%   degree & Under grad degree & 585 & 25.00 & 44.53 & 5792 \\ 
-%   degree & Graduate courses & 247 & 10.56 & 44.63 & 2471 \\ 
-%   degree & Graduate degree & 899 & 38.42 & 52.07 & 9496 \\ 
-%   age\_level & 18-25 & 743 & 31.75 & 43.50 & 7340 \\ 
-%   age\_level & 26-30 & 549 & 23.46 & 46.19 & 5610 \\ 
-%   age\_level & 31-35 & 375 & 16.03 & 44.19 & 3912 \\ 
-%   age\_level & 36-40 & 257 & 10.98 & 54.97 & 2714 \\ 
-%   age\_level & 41-45 & 140 & 5.98 & 43.89 & 1510 \\ 
-%   age\_level & 46-50 &  94 & 4.02 & 49.54 & 994 \\ 
-%   age\_level & above 50 & 184 & 7.86 & 52.32 & 2000 \\ 
-%   country & United States & 1086 & 46.41 & 39.65 & 10763 \\ 
-%   country & India & 980 & 41.88 & 52.63 & 10238 \\ 
-%   country & Rest of the world & 279 & 11.92 & 50.38 & 3079 \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
-
-The largest number of participants falls within the age group of 18 to 25, with the majority being between 18 to 35.
-Many older participants also took part in the experiments.
-%Interestingly, there are many participants in the older age groups as well.
-The United States shows a more uniform participation beyond age 30,  participants from India are primarily younger.
-
-In terms of academic background, the majority of participants had a graduate degree (a total of 899 participants, about 38.42\%).
-Examining the participants' location in conjunction with the academic background reveals that there is an above average  number of participants from India reporting to have a graduate degree.
-In retrospect this turned out to be a culturally based misunderstanding, as any university degree in India is considered to be a graduate degree, while the North American definition of a graduate degree refers to a Masters level education or above.
-Most of the  participants from the U.S.
-report to have an undergraduate degree or at least have some undergraduate courses.
-
-Compared to data from the 2010 US Census \cite{census}, participants were both relatively younger, and more highly educated than the general US population. 
-Being turkers implicitly assures that the participants in these experiments have a relatively high level of technological fluency.
-In fact, the demographics of participants align somewhat better with demographics of internet users published by
-\citet{pew}.
-
-The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where the proportion of female participants is lower.
-The distribution of education levels also differs slightly across countries for this age group.
-
-A total of 1912 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where each participant evaluated three lineups.
-As a measure to ensure a high quality in the evaluations,
- a test plot was shown to each observer, and the performance on this test plot was used for screening purposes.
-%A test plot was shown to each observer and the feedback received from that plot is used to ensure data quality.
-Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.
-Data from participants who gave the wrong answer on the test plot % were considered as gaming the system, and their data were removed.
-led to the removal of all of their answers from the analysis. 
-Results from test plots were also not considered in any of the analyses.
- In some cases  participants chose to not provide their demographic information.
-Also, for some ip address, the actual geographical locations could not be retrieved.
-This resulted in some missing geographic information.
-
-\subsection{Demographic Factors}
-
-To explore the significance of the demographic factors model~\eqref{eqn:demographic_response} is fit to the data, with age, country, education and gender as fixed effects.
-To estimate the significance of each of the factors, reduced models are fit with that factor removed from the model.
-Table~\ref{tbl:anova_factor} summarizes the results.
-Except for gender, all of the demographic factors show significant differences in  detection rates.
-With respect to time taken, all of the demographic factors are significant.
-
 % latex table generated in R 3.4.1 by xtable 1.8-2 package
@@
 \end{table}
-
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 17:43:02 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{rrrrrlrrrr}
-%   \hline
-%  & deviance & Chisq & Chi.Df & Pr..Chisq. & g1 & deviance & Chisq & Chi.Df & Pr..Chisq. \\ 
-%   \hline
-% Full & 52609.06 & 0.00 & 6.00 & 0.00 &  & 24140.49 & 0.00 & 6.00 & 0.00 \\ 
-%   Age & 52081.29 & 527.77 & 6.00 & 0.00 &  & 24113.24 & 27.25 & 6.00 & 0.00 \\ 
-%   Country & 52245.46 & 363.60 & 2.00 & 0.00 &  & 24114.79 & 25.70 & 2.00 & 0.00 \\ 
-%   Degree & 52498.65 & 110.41 & 4.00 & 0.00 &  & 24116.02 & 24.47 & 4.00 & 0.00 \\ 
-%   Gender & 52540.30 & 68.76 & 1.00 & 0.00 &  & 24138.85 & 1.64 & 1.00 & 0.20 \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
-
-Parameter estimates from the model fits are shown in Table~\ref{tbl:model_result_demographics}.
-The first level of each factor serves as the baseline for the model.
-The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for other age groups except for 26-30.
-Participants from the rest of the world have a similar detection rate to those from the USA, but participants from India show a significantly lower detection rate.
-Subjects reporting to have  an undergraduate degree had a significantly higher detection rate.
-For time to respond, all age groups were significantly faster than  18-25 year olds.
-Participants from India as well as the rest of the world were faster than US Americans.
-A higher degree corresponded to longer time, and males were on average faster than females.
-
-% latex table generated in R 3.4.1 by xtable 1.8-2 package
-% Wed Oct 18 17:22:24 2017
-\begin{table}[hbtp]
-\caption{\label{tbl:model_result_demographics}Parameter estimates of models~\eqref{eqn:demographic_time} and~\eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively.
-For time taken all the demographic factors are significant.
-For detection rate, age group 36-40, rest of the world and graduate degree are significantly different.
-For gender no difference in performance is observed.
-Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
-\centering
-\scalebox{.9}{
-\begin{tabular}{rrr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
- && \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
-
-\cline{3-6} \cline{8-11} 
-\multicolumn{2}{c}{{Demographic Factor} \hfill Level} & Est & 2.5\% & 97.5\% &   && Est & 2.5\% & 97.5\% &   \\ 
-\cline{3-6} \cline{8-11}
-\bf Fixed Effects & Average $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.63 & -0.81 & -0.45 & *** \\ [3pt]
-\bf Age Category $(\alpha)$ & 18-25 & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
-&  26-30 & 0.05 & 0.03 & 0.07 & *** &   & 0.03 & -0.06 & 0.13 &  \\ 
-&  31-35 & 0.06 & 0.03 & 0.09 & *** &   & 0.12 & 0.02 & 0.23 & * \\ 
-&  36-40 & 0.22 & 0.19 & 0.25 & *** &   & 0.29 & 0.17 & 0.41 & *** \\ 
-&  41-45 & 0.17 & 0.13 & 0.21 & *** &   & 0.19 & 0.03 & 0.34 & * \\ 
-&  46-50 & 0.28 & 0.23 & 0.32 & *** &   & 0.20 & 0.01 & 0.39 & * \\ 
-&  above 50 & 0.34 & 0.31 & 0.38 & *** &   & 0.16 & 0.02 & 0.29 & * \\ [3pt]
-\bf Country $(\kappa)$& U.S. & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
-&  India & 0.23 & 0.20 & 0.25 & *** &   & -0.19 & -0.28 & -0.10 & *** \\ 
-&  Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.06 & -0.05 & 0.18 &  \\ [3pt]
-\bf Education $(\tau)$& High School & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
-&  Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & 0.00 & -0.14 & 0.14 &  \\ 
-&  Under grad degree & 0.14 & 0.10 & 0.17 & *** &   & 0.16 & 0.03 & 0.29 & * \\ 
-&  Graduate courses & 0.02 & -0.02 & 0.06 &  &   & -0.13 & -0.29 & 0.03 &  \\ 
-&  Graduate degree & 0.13 & 0.10 & 0.16 & *** &   & -0.04 & -0.17 & 0.09 &  \\ [3pt]
-\bf Gender $(\gamma)$& Female & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
-&  Male & 0.08 & 0.06 & 0.10 & *** &   & 0.05 & -0.02 & 0.12 &  \\
-\cline{3-6} \cline{8-11}
-\bf Random Effects &  lineup $(\sigma_\ell)$& \multicolumn{5}{l}{0.29} & \multicolumn{4}{l}{\phantom{-}2.27} \\ 
-                   &  Error $(\sigma)$& \multicolumn{5}{l}{0.69} & \multicolumn{4}{l}{\phantom{-}0.89}  \\ 
-\cline{3-6} \cline{8-11}\\
-\hline
-\multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
-\end{tabular}}
-\end{table}
-
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 17:22:24 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{rrrrllrrrl}
-%   \hline
-%  & Est & LB & UB & pvalue & g1 & Est & LB & UB & pvalue \\ 
-%   \hline
-% $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.63 & -0.81 & -0.45 & *** \\ 
-%   26-30 & 0.05 & 0.03 & 0.07 & *** &   & 0.03 & -0.06 & 0.13 &  \\ 
-%   31-35 & 0.06 & 0.03 & 0.09 & *** &   & 0.12 & 0.02 & 0.23 & * \\ 
-%   36-40 & 0.22 & 0.19 & 0.25 & *** &   & 0.29 & 0.17 & 0.41 & *** \\ 
-%   41-45 & 0.17 & 0.13 & 0.21 & *** &   & 0.19 & 0.03 & 0.34 & * \\ 
-%   46-50 & 0.28 & 0.23 & 0.32 & *** &   & 0.20 & 0.01 & 0.39 & * \\ 
-%   above 50 & 0.34 & 0.31 & 0.38 & *** &   & 0.16 & 0.02 & 0.29 & * \\ 
-%   India & 0.23 & 0.20 & 0.25 & *** &   & -0.19 & -0.28 & -0.10 & *** \\ 
-%   Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.06 & -0.05 & 0.18 &  \\ 
-%   Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & 0.00 & -0.14 & 0.14 &  \\ 
-%   Under grad degree & 0.14 & 0.10 & 0.17 & *** &   & 0.16 & 0.03 & 0.29 & * \\ 
-%   Graduate courses & 0.02 & -0.02 & 0.06 &  &   & -0.13 & -0.29 & 0.03 &  \\ 
-%   Graduate degree & 0.13 & 0.10 & 0.16 & *** &   & -0.04 & -0.17 & 0.09 &  \\ 
-%   Male & 0.08 & 0.06 & 0.10 & *** &   & 0.05 & -0.02 & 0.12 &  \\ 
-%   lineup & 0.29 &  &  &  &   & 2.27 &  &  &  \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
-
-Figure~\ref{fig:demographic_effect} supports the model results.
-The average time taken to respond (natural log) and detection rate are computed by lineup for the different demographic factor levels, and displayed using boxplots.
-Mean values are shown as dots inside the boxplot, as this statistic rather than the median is used in the models.
-The means and medians for response time (top row) are essentially equal.
-There are some differences between the two in the detection rate plots (bottom row).
-The differences in these values for different factors, that was tested by the models, can be seen in the plots.
-However, particularly for detection rate, the variation is huge.
-A large component of the variation is difficulty of the lineup.
-In some lineups the actual data plot was distinctively different from the null plots, and we would expect that the detection rate would be close to 100\%.
-In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond.
-In Table~\ref{tbl:model_result_demographics} notice that lineup specific error is estimated as 2.27 which is much higher than any of the other parameter estimates in the model.
-This indicates that the major and most important factor affecting the detection rate is  lineup difficulty.
-So although the demographic factors emerge from the model to be statistically significant, the practical significance is minimal.
-We illustrate this with the following example of graduate degree.
-
-\begin{figure}[htbp] 
-   \centering
-   \includegraphics[width=4in]{demographic_effect.pdf} 
-   \caption{Boxplots of average log time taken and proportion correct responses (detection rate) of all the lineups plotted for each demographic factor levels.
-The dots inside the boxes represent means.
-Some differences in means of various demographic factors are observed.
-Variability in detection rate indicates large variability in lineup difficulties.
-}
-   \label{fig:demographic_effect}
-\end{figure}
-
-While some of the demographic factors are strongly statistically significant, the main source of variation in detection rate is the lineup difficulty.
-For example, let's examine the effect of an undergraduate degree.
-To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with some graduate course work as compared to an undergraduate degree, for an average difficulty lineup (random effect of zero).
-Plugging in the relevant quantities to the fitted model gives a difference equal to:
-$$ \frac {\exp(-0.63-0.13)}{1+\exp(-0.63-0.13)}-  \frac {\exp(-0.63+0.16)}{1+\exp(-0.63+0.16)} = 0.319- 0.385 =-0.066.$$ 
-The person with some graduate course work picks the data plot  on average in 31.9\% of lineups of average difficulty, as compared to 38.5\% if they have an undergraduate degree.
-This difference is reduced to 4\% for a lineup with one standard deviation order of magnitude difference in difficulty.
-For two standard deviations it further reduces to 0.6\%.
-Thus, although there is a statistically significant difference in the proportion at which participants identify the  data plot for some demographic factors, these are not  significant in any practical sense.
-Figure~\ref{practical_impact_und_und_graduate.pdf} illustrates this example showing fitted models for a US 18-25 female with either a high school education or an undergraduate degree.
-Similar calculations show the same negligible impact of age level 36-40 (\Sexpr{round(exp(-0.63+0.29)/(1+ exp(-0.63+0.29)) - exp(-0.63)/(1 + exp(-0.63)),3)} at most)  and country India (\Sexpr{round(exp(-0.63-0.19)/(1+ exp(-0.63-0.19)) - exp(-0.63)/(1 + exp(-0.63)), 3)} at most) on the probability of a correct response.
-Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.
-
-\begin{figure}[htbp] 
-   \centering
-    \includegraphics[width=3.2in]{practical_impact_und_graduate.pdf} 
-    \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
-   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with an under graduate degree compared to a high school degree.
-Even though having an under graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.04 which is  negligible from a practical point of view.
-The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} = 2.27$)
-   of lineup variability.}
-   \label{practical_impact_und_und_graduate.pdf}
-\end{figure}
-
-\subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments~5,~6, and~7 separately.
-In this model, attempt is fitted as a factor variable to allow for all possible non-linear learning trends.
-It should be noted that we also examined an alternative to model~\eqref{eqn:trend_response} where attempt was linearly fitted as a continuous covariate, but this also was not significant.
-
-Table~\ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model~\eqref{eqn:trend_response} examining detection rate.
-Only for experiment~6 there is some evidence of a learning curve.
-There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.
 
@@
 
-% 
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 23:37:55 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{rrrrllrrrllrrrl}
-%   \hline
-%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
-%   \hline
-% $\mu$ & 3.80 & 3.72 & 3.88 & *** &   & 3.89 & 3.83 & 3.96 & *** &   & 3.75 & 3.64 & 3.85 & *** \\ 
-%   $\alpha_1$ & 0.32 & 0.25 & 0.39 & *** &   & 0.34 & 0.28 & 0.40 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
-%   $\alpha$ & -0.04 & -0.05 & -0.03 & *** &   & -0.04 & -0.05 & -0.03 & *** &   & -0.03 & -0.04 & -0.02 & *** \\ 
-%   $\sigma^2_u$ & 0.52 &  &  &  &   & 0.49 &  &  &  &   & 0.37 &  &  &  \\ 
-%   $\sigma^2_a$ & 0.03 &  &  &  &   & 0.05 &  &  &  &   & 0.05 &  &  &  \\ 
-%   $\sigma^2_l$ & 0.09 &  &  &  &   & 0.20 &  &  &  &   & 0.24 &  &  &  \\ 
-%   $\sigma^2$ & 0.46 &  &  &  &   & 0.50 &  &  &  &   & 0.45 &  &  &  \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
+\begin{table*}[hbtp] 
+\centering 
+\caption{Overview of 10 different Turk experiments, from which data was collected to study human factor effects.
+All of the experimental data were used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.}
+\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
+\hline\hline 
+ & Experiment &  Test Statistic  & Lineup question & Used in study of\\ [0.5ex] % inserts table %heading 
+\hline 
+1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
+between group A and B? & DF\\
+2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+& DF\\
+  3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope?
+& DF\\
+ 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
+  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
+  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group?
+&  DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
+  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups?
+& DF  \hfill LT \hfill \phantom{.} \hfill \phantom{.} \\
+  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent?
+& DF\\
+  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small?
+& DF  \hfill \phantom{LT} \hfill {LE} \hfill \phantom{.}\\
+  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others?
+& DF\\
 
-To visualize how the time taken reduces over the successive attempts, we fitted model~\eqref{eqn:trend_time} excluding the covariate attempt from the model and computed the residuals.
-Least square regression lines are fitted through the subject specific residuals.
-Subject specific slopes are much different in each of the three experiments as we see in Figure~\ref{fig:learning_trend_time}.
-Some subjects improved over attempts by taking less time in the later attempts while others got worse.
-The averages of these residuals for each attempt are plotted as dots.
-Least square regression lines are fitted to these points excluding the first attempt since for first attempt we fitted an indicator covariate.
-The downward trends are evident in the plots.
-All the slopes are highly significant.
-As expected we observed large positive residuals for each of the experiments for first attempt.
-
-\begin{figure}[htbp] 
-   \centering
-%   \includegraphics[width=6.5in]{learning_trend_time.pdf}    
- \includegraphics[width=6.3in]{learning_trend_time_subject.pdf}    
-   \caption{Least square regression lines fitted through the subject specific residuals obtained by  fitting model~\eqref{eqn:trend_time} without covariate attempt.
-Differences in subject specific slopes are observed.
-Some of the subjects did worse over successive attempts while others did better.
-Averages of these residuals are plotted as dots and least square regression lines are fitted to obtain overall trends.
-For all the three experiments the overall downward slopes are statistically significant which indicates that MTurk workers become more efficient as they progress through their attempts.}
-   \label{fig:learning_trend_time}
-\end{figure}
-
-\subsection{Location Effect}
-
-A total of 111 subjects was recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
-Each subject evaluated two lineups;  one for Interaction effect and the other for Genotype.
-In total there were 222 responses for 50 lineups.
-The data on the test lineup were excluded from the analysis.
-
-We are investigating location effect in two ways: (1) location effect of each individual plot location, and (2) location effect of inside/outside panel positions.
-
-Figure~\ref{fig:location_effect} shows an overview of detection rates for both of the effects studied in experiment~9.
-Each dot corresponds to one lineup.
-The size of each dot gives the number of evaluations.
-Lineups with  the same set of null plots are connected by lines.
-%The detection rate for each data plot location is shown in colored by null sets. 
-The overall average detection rate for each location is shown using dashed lines.
-Size of the points represent the number of responses.
-For some locations we have as many as 10 responses.
-For location 1, we did not have any responses for null set 1 in one of the interaction lineups.
-
-We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets.
-The lineup protocol is using a finite, small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set.
-If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate.
-From the plot we can see that null set 3 appeared to be more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot.
-This null set effect was tested, and a significant difference was found for null set 3 of the interaction lineup.
-The rest of the null sets do not show any differences at  a 1\% significance level.
-
-\begin{figure}[htbp] 
-   \centering
-    \includegraphics[width=6.5in]{proportion_nulls_guide.pdf} 
-   \caption{Location of data plot in the lineup and proportion correct (detection rate) for both Interaction and Genotype effect.
-Each colored line represents a null set and the size of the dots represents number of responses.
-The overall average proportions are shown by dashed lines.
-The actual data plot locations are shaded grey on the top panels to demonstrate their relative positions on a lineup.}
-   \label{fig:location_effect}
-\end{figure}
-
-Model~\eqref{manova} was fit to the data to test, if  detection rates were significantly different for the different locations.
-For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R}.
-The results are shown in Table~\ref{tbl:manova}.
-The $p$-values for both the Interaction and the Genotype effect are much bigger than the conventional threshold of 0.05 suggesting that location of the actual data plot in the lineup does not significantly affect performance.
-
-\begin{table}[hbtp]
-\caption{The summary of the results obtained by fitting MANOVA model~\eqref{manova} using Wilks test.}
-\begin{center}
-\begin{tabular}{ccccccccc}
-  \hline \hline
- Location & && & \multicolumn{3}{c} {Degrees of Freedom}  & F test \\
- \cline{5-7}
- Effect & DF & Wilks & Approx. F & Numerator& Denominator &Residual & $p$-value\\
-  \hline
-  Interaction& 3&0.07986 & 0.592 &15& \phantom{0}5.92 & 6 & 0.8082 \\ 
-  Genotype & 4&0.01689 & 1.722 & 20&14.22 & 8 & 0.1488 \\ 
 \hline 
 \end{tabular} 
-\end{center}
-\label{tbl:manova}
-\end{table}
-
-For the second approach in investigating the location effect, we distinguish between inner panels 7, 8, 9, 12, 13, 14 and outer panels (any panel with an outside edge).
-From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype.
-Model~\eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.
-
-% ===================================================================================
-\section{Conclusion}
-%------------------------------------------------------------------------------------
-
-Human demographics have the potential to influence performance when the lineup protocol is used for statistical inference.
-In our study of the demographic effects on performance across a set of different experiments we found some statistically significant effects.
-Age group 36-40, people who have a graduate degree have a significantly higher detection rate.
-India shows significantly lower detection rate compared to United States.
-Gender does not have any significant effect.
-However, the effects are minimal, on the order of a few percentage points different from the average.
-These results are very important for the power of visual tests as they demonstrate the robustness of the test against different human factors.
-
-Individual learning trend is observed in time taken but not so much in observer performance.
-Some individuals improved on their performances while others showed a decrease on successive attempts.
-This could be interpreted as good, in the sense that it means that observers could realistically be recruited from sources like MTurk, and that substantial training is not necessary in order to obtain useful evaluations of lineups in practice.
-However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea, at least for short term learning.
-
-The simulation experiment reveals that there is no significant effect of location on the actual data plot in the lineup.
-This is important as the visual statistical inference procedure prescribes that the data plot be placed at random in the lineup.
-This paper suggests that any random place in a lineup is as good as other places in the lineup.
-Even though there are variations on the performance depending on different null sets, their impact on probability to correctly evaluate a lineup is very negligible.
-
-There are many more ways that the lineup protocol might be tested to learn what we might expect about its performance for tackling real data mining problems.
-Possible changes in the lineup protocol in the pipeline are allowing observers to select more than one plot, and to vary the size of the lineup from 20 to a smaller number, and have observers all see different lineups with different null sets.
-This paper assessed the effect of human factors on the experiments conducted to date.
-As changes to the lineup protocol are suggested by other experiments the human factor effects may need to be examined again.
-
-\if1\blindDoc
-{
-%------------------------------------------------------------------------------------
-\paragraph{Acknowledgments}
-%------------------------------------------------------------------------------------
-This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
-} \fi
-
-\bibliographystyle{asa}
-\bibliography{references}
+\label{tbl:visual_stat} 
+\end{table*}
 

### 2024-12-30 wiley_paper/wiley_paper.tex
Add files via upload

@@
 
-%The lineup protocol introduced in \citet{buja:2009} can be used to quantify the significance of graphical findings in the exploratory data analysis process.
-This methodology is  part of what is called visual statistical inference, which has been developed further  and validated by \citet{majumder:2013} in simulation studies of head to head comparisons with conventional inference.
-One of their major contributions is to define the power of  visual tests and to provide a method for obtaining the power for a particular lineup.
-In some scenarios, the power of a visual test  was seen to outperform that of a conventional test.
-
-%In visual inference, the test statistic is a plot of the observed data, called the {\it (actual) data plot}.
-To create a lineup the data plot is placed randomly in a grid of null plots.
-Null plots  are rendered from data generated as  specified by a null hypothesis.
-Often, this null hypothesis assumes independence or no structure in the data.
-An observer is then asked to evaluate the lineup.
-If the  actual data plot is detected by the observer, this counts as evidence against the null hypothesis.
-Based on the number of evaluations by independent observers and the number of times that the data plot is identified, we are  able to evaluate the significance of the graphical relationship in the same way and the same rigor as a conventional hypothesis test.
-
 The lineup protocol introduced by \citet{buja:2009} quantifies the significance of graphical findings in exploratory data analysis, forming part of visual statistical inference.
@@
 
-% \begin{figure}[hbtp] 
-%    \centering
-%    \includegraphics[width=0.99\textwidth]{electoral-5-13.pdf} 
-%    \caption{ \label{fig:elect} Which one of the plots, labelled 1 through 20 from top left to bottom right, is the most different from the others?}
-% \end{figure}
-
-%Figure~\ref{fig:elect} displays an example lineup, one of the plots is based on the observed data, while the remaining 19 plots show  data generated from a null model.
-Which one of the 20 plots is the most different from the others?
-%When asked this question, 12 out of 72 observers picked the  data plot located in panel number $4^2-3$ in the lineup\footnote{The little piece of calculus used for describing the location of the data plot in the lineup provides a small cognitive hurdle that is supposed to enable the interested reader to inspect the lineup without being biased by knowing the location of the data plot.}.
-The corresponding $p$-value is 0.0077, indicating sufficient evidence to reject the null hypothesis.
-
-%How do we interpret this finding, though? For that, we need to know the context of the data and we need to have more information about how the null plots were generated. 
-%This particular example investigates the results from the 2012 US presidential election in comparison to  poll results just prior to the election. 
-%Although this example is more simplistic than most of the tests conducted to date, it serves the purpose of illustrating the lineup protocol.
-The data consists of the difference in poll results between the two (major) presidential candidates, Obama and Romney, for all states.
-Each panel in Figure~\ref{fig:elect} shows an `electoral building' \citep{mosley2010} where each state in the union is represented by a rectangle.
-The difference in poll results is plotted horizontally, while the height of each box corresponds to the number of that state's electoral votes.
-Color indicates party affiliation.
-The null hypothesis is that ``the election results are consistent with the polls''.
-Therefore the polling results provide the null model from which data is simulated.
-A normal model with mean and standard deviation based on a poll and its margin of error is used to simulate different scenarios that might have resulted on election day, if the polls were on target.
-Each null data set is generated as a set of draws from this model.
-% normal distribution, with mean equal to the difference in poll percentage of the latest state poll results, and standard deviation equal to 2.5, approximating a margin of error of 5\%.
-%These samples are plotted as electoral buildings, and the plot based on the actual election results is placed randomly among the null plots in a lineup of size $5\times 4$.
-If the null hypothesis is true, the  data plot should look  like any of the other plots, and not be identifiable by an observer.
-Figure~\ref{fig:tower} shows the plot of the electoral building with the additional context information and labels.
-
-% \begin{figure}[hbtp] 
-%    \centering
-%    \begin{minipage}[c]{0.49\textwidth}
-%    \includegraphics[width=\textwidth]{tower.pdf}
-%    \end{minipage}
-%     \hfill
-%    \begin{minipage}[c]{0.45\textwidth}
-%    Poll aggregator 1 \\
-%       \includegraphics[width=\textwidth]{freedom-sim.pdf} 
-%    Poll aggregator 2\\
-%       \includegraphics[width=\textwidth]{rcp-sim.pdf} 
-%    \end{minipage}
-%    \caption{ \label{fig:tower} Electoral building plot of the results of the 2012 U.S.
-Presidential Election (left).
-On the right two histograms showing simulated poll results for the Democratic presidential candidate using  polling averages collected from two different sources, Freedom's Lighthouse Averages (top), RealClearPolitics (bottom).
-If the polls matched the electoral result the simulated results would be centered at the electoral result.
-This happens for the RealClearPolitics polls (bottom) but not for Freedom's Lighthouse.
-For the latter, only 5.33\% of simulated results were as high as the actual election result, which converts to a $p$-value of 0.0533, indicating an almost statistically significant difference.
-It hints at the polls from this source under-estimating the vote for Obama.
-This source was used to generate the null plots in the lineup in Figure~\ref{fig:elect}.
-The $p$-value from the lineup protocol, was 0.0077, which more strongly concludes that these polls mismatched the actual result.}
-%\end{figure}
-%The lineup of Figure~\ref{fig:elect}, as well as the other lineups in the manuscript are created using the nullabor package by \citep{nullabor}.
-
-%A lineup can be evaluated by a single person or multiple observers.
-A distribution similar to the binomial distribution, but adjusted for dependencies introduced by the lineup scenario, is used to calculate the $p$-value based on the number of times observers identify the actual data plot, which provides the information needed to make a decision on rejecting or failing to reject the null hypothesis \citep{vinference}.
-To avoid expectation errors \citep{meilgaard}, particularly in emotionally charged areas such as election results, observers should not be aware of the data that constitutes a lineup, and should therefore not have seen the actual data plot before inspecting the lineup.
-This is the reason that we revealed the context of the election problem only after the interested reader had a chance to evaluate the lineup.
-
-%When presenting a lineup to an observer the accompanying question should be phrased in as general a manner as possible, effectively asking the observer to pick the plot that is most different, and allowing observers to provide their own reasons for their choice.
-This ensures all possible deviations from the null hypothesis to be detected.
-However, if lineup tests are run in head-to-head comparisons with conventional statistical tests, such as the experiments in \cite{majumder:2013} or \cite{tengfei:2013}, these questions have to be phrased much more specifically, in order to avoid  type III errors \citep{mosteller:48} consisting of correctly rejecting the null hypothesis for a wrong reason.
-In those experiments, the structure in the data is strictly controlled in the simulation process, which allows for specific questions to be asked.
-
 A lineup can be evaluated by one or more observers, and a distribution similar to the binomial, adjusted for dependencies from the lineup scenario, is used to calculate the $p$-value based on how often the actual data plot is identified \citep{vinference}.
-To avoid expectation bias \citep{meilgaard}, especially in sensitive areas like election results, observers must not know the data constituting the lineup or have seen the actual data plot beforehand.
-The context of the election problem was revealed only after lineup evaluation to prevent bias.
+To avoid expectation bias \citep{meilgaard}, observers must not know the data constituting the lineup or have seen the actual data plot beforehand.
 The accompanying question in a lineup should be broad, asking observers to pick the most distinct plot, allowing detection of any deviations from the null hypothesis.
 However, in head-to-head tests with conventional methods \citep{majumder:2013, tengfei:2013}, specific questions are needed to avoid type III errors \citep{mosteller:48} consisting of correctly rejecting the null hypothesis for a wrong reason.
-
-%In the election example, observers were asked ``which plot is the most different?''.
-The type of plot, showing two (modified) stacked bar charts in different colors should suggest to the observer that the interesting feature is the difference between the two heights.
-Most observers based their decision on this -- besides  `asymmetry' between the towers (a reason suggested by us), most free-form responses for the reason of choice describe a comparison between the red and blue towers.
-In the responses for this lineup we see something that is very typical of responses in lineups: it is only a few panels from the lineup that observers pick, while most panels do not get picked or get picked only a few times.
-Out of a total of 72~responses, 23 observers picked panel~\#8, 12 observers picked the (data) panel \#13, 10 observers picked panel~\#4, 6 observers picked panel \#7, and all of the remaining panels were picked at most three times.
-From the panels picked, we see that the top three choices  are all extreme with respect to heights of blue versus red towers:  both panels \#8 and \#13 show a large difference between the height of the towers, while panel \#4 is the only panel, in which the red tower is higher than the blue one.
 
@@
 
-%We have conducted a series of experiments (see Table~\ref{tbl:visual_stat} for an overview) using Amazon's Mechanical Turk~\citep{turk} (MTurk) for a variety to purposes: validating the protocol against existing tests, comparing plot designs, and evaluating structure in data analysis problems.
-In all of the experiments we collected demographic details, such as gender, age bracket, and educational background.
-Mining this data will provide a basis for evaluating the effects of different covariates on an individual's ability to evaluate lineups.
-The design of some of the experiments allow a further investigation into aspects beyond participants' demographics: experiments~5, 6 and 7 can be used to investigate short term learning trends, while experiment~9 allows insight on whether the  positioning of the  data plot in the lineup has an effect on detection rate.
-
-We conducted several experiments (see Table~\ref{tbl:visual_stat}) using Amazon's Mechanical Turk~\citep{turk} (MTurk) for validating the lineup protocol, comparing plot designs, and evaluating data analysis structures.
+We conducted several experiments (see Table~\ref{tbl:visual_stat} in the Appendix~\ref{sec:tbl_fig}) using Amazon's Mechanical Turk~\citep{turk} (MTurk) for validating the lineup protocol, comparing plot designs, and evaluating data analysis structures.
 Demographic data such as gender, age, and educational background were collected to assess the influence of these covariates on lineup evaluations.
-Additionally, experiments investigated short-term learning trends as well as explored the effect of data plot positioning on detection rates.
-
-%Section~\ref{sec:factor_performance} discusses human factors that we expect to be influential on the performance of the observer.
-Section~\ref{sec:exp_design} describes the methods used to assess the effects, and Section~\ref{sec:result_socio} summarizes the results of what we have learned about human factors affecting lineup evaluation.
+Additionally, experiments investigated the short-term learning trends.
 
@@
 
-%\item{\bf Choice of Visual Test Statistic:}  
-%Data can be plotted in a variety of ways and we can enhance plots by axes, legends or representations of models.
-The choice of plot primarily should be appropriate for the problem being investigated, as the data plot is the visual test statistic.
-%When for example studying the association between two quantitative variables a scatterplot is usually considered best practice.
-Enhancements, such as a line overlaying a scatterplot to represent a fitted model, might be appropriate for studying models for the data.
-Some choices of visual test statistics provide better results than others when studying the same problem \citep{heike:2012}, because some plot types make it easier for certain structures to be seen.
-
-%Table~\ref{tbl:visual_stat} shows the large variety of visual test statistics used in the Turk studies completed so far.
-These experiments were conducted to examine many different visual tasks.
-Side-by-side boxplots were used in experiment~1 to study the importance of a categorical variable in a regression model.
-For both experiments~2 and~3, a scatterplot is the visual test statistic, but a regression line fitted through the points is overlaid for experiment~2.
-The experiments studied slightly different problems, for which the difference in plot type was important.
-Experiment~4 compared polar coordinates to euclidean coordinates on the reading of structure in a barchart.
-Experiment~5 examined the style of plot for a bivariate data problem, by comparing results when using a scatterplot, side-by-side dot plots, overlaid densities or side-by-side boxplots.
-Experiment~6 studied the difference between variations of boxplots.
-Experiment~7 studied the effects of high dimension low sample size on the separation of groups in low-dimensional projections.
-The sine illusion, where people read the strength of differences between curves as different when there is a strong sinusoidal structure, is examined in experiment~8.
-Experiment~9 was conducted to test if there was any structure in an RNA-seq data set measured on soybeans, and used interaction plots classically constructed to examine interaction effects in $2\times 2$ factorial designs.
-Experiment~10 used different sorts of residual plots for hierarchical linear models as the visual test statistics.
-
-%In experiments~4,~5 and~6, different visual test statistics were used on the same data, providing the ability to examine the effect of these choices, the use of lineup tests for comparisons and  examples are discussed in \citet{heike:2012, qqplots:2015}
-
-%\item{\bf Signal in the Data:}
-%A visual test statistic is chosen for a particular task.
-If it is well-designed  departures from the null hypothesis will be visible in the plot, especially if the signal is strong.
-An important factor that helps an observer to identify the actual data plot in a lineup is the strength of the signal in the data.
-Simulated data was generated for all of the experiments, except exps.~4 and~9, enabling the study of signal strength and actual data plot detection.
-In experiment~4 we took a different approach by varying the sampling size, which also allowed us to investigate the effect a design had on detection rate while varying signal/noise ratio.
-Primarily the results were as expected, that as the signal strength increased observers more frequently picked the actual data plot as different from the null plots.
-
-%\item{\bf Question Design:} A primary purpose of visual inference is to preclude the necessity to pre-specify discoverable features in data that is required by classical statistical inference.
-Visual inference supports discovery of structure, enabling the unexpected to surprise us.
-This is tightly connected to the task that the human observer is asked to perform.
-For most purposes the observer is asked a very general question, such as to pick the plot that is most different from the others, and explain the reasons why they see it as different.
-By being general, all possible discoverable structures are included in the significance calculations \citep[see also ][]{buja:2009}.
-
-%In the initial Turk experiments (1,2,3) conducted with simulated data to compare the results from visual inference with those obtained by classical inference, the questions posed to the observers were more targeted.
-This was important to allow a direct comparison of the results to those from conventional testing.
-%in order to make a comparison of the results.
-%For latter studies (4, 7, 10), questions posed to the observer were more general. Table~\ref{tbl:visual_stat} lists the questions used for each experiment.
-
 \item{\bf Demographics:}  During each of the experiments, data on age, gender, education level and geographic location was collected.
@@
 
-%\item{\bf Location of Actual Data Plot in the Lineup} For all of the lineups used in the experiments a $5\times 4$ grid of 20 panels is used.
-A random number generator is used to determine the position where the actual data plot is placed in each lineup.
-Eye tracking experiments \citep{zhao:2012} suggest that some observers traverse lineups in horizontal direction, while others have a vertical up-and-down approach to their search.
-Almost universally, observers start at the top left of a lineup.
-This leads to some concern that observers might be able to more easily identify the actual data plot, if it is placed at the top left of a lineup than when it the actual data is placed at the bottom right.
-
-%Ideally, this does not happen.
-Experiment~9  is designed to allow us to investigate the effects of location on detection.
-Five different locations in the lineup were used to assess how fast and accurately observers identified the actual data plot.
-
-%\item{\bf Sample of Null Plots:} In classical inference, the test statistic is compared to the full null distribution, to decide if it is extreme or not.
-In visual inference the actual plot is compared to plots of a finite number of samples from the null distribution.
-In the lineups used in the Turk experiments, the actual data plot is compared on 19 null plots.
-These null samples are random draws, and there is a chance that one or more null plots in the lineup may be similar or even more extreme than the actual data plot.
-The sample of null plots can affect the observer's decision.
-This is tested with the data from experiment~9, where the experiment was set up with lineups made from different null plots.
-
 %\subsection{Individual Skill or Ability} 
@@
 We can examine the effect of individuals' skills and abilities because multiple subjects evaluated the same lineups, which allows us to estimate, if some subjects consistently detect the actual data plot more often than others.
- %For example, in a controlled experimental study \citet{zhao:2012}  noticed that some people spent a lot of time to decide no matter whether the lineup is difficult or easy while some simply glanced at lineups to make a decision.
-This influences the response of the observer.
-Also subject specific variation in the power of visual test is observed in \cite{majumder:2013}.
+
 \end{itemize}
-
-% 
-% % ===================================================================================
-% \section{Experimental Methods}\label{sec:exp_design}
-% %------------------------------------------------------------------------------------
-% 
-% Two of the factors, signal in the data and individual abilities, were studied in \cite{majumder:2013}.
-The choice of visual test statistic was examined in \cite{heike:2012}.
-In each of these analyses demographic factors were given a cursory glance, to ensure that they did not have large effects on the results.
-The design of experiments 5, 6 and 7 enables the examination of learning trend, which is studied in this paper.
-Experiment~9 was a real test case for visual statistical inference, and in order to understand the significance of the structure in the genomic data, multiple lineups were made in which location of the actual data plot, and the sample of nulls, were randomized.
-This enables the assessment of the effect of these factors on the results.
-This section describes the experimental methods used to examine the effects of demography, placement of the actual data plot, sample of null plots and the existence of a learning trend.
-% 
-% \subsection{Demographic Factors}
-% \noindent
-% For all  ten experiments shown in Table~\ref{tbl:visual_stat}, the following demographic information was collected from subjects:
-% 
-% \begin{enumerate} \leftmargin 5cm  \itemsep 0in
-% \item {\it Age group}, with categories set to be 18-25, 26-30, 31-35, 36-40, 41-45, 46-50, above 50.
-% \item {\it Gender}, male or female.
-% \item {\it Education level}, with levels being high school or less, some undergraduate courses, undergraduate degree, some graduate courses, and graduate degree.
-% \item {\it Geographical location}, collected from the IP address of the participants' computer, as latitude, longitude, city and country. 
-% \end{enumerate}
-% 
-% 
-% Let $Y_{ij}$ denote the response from observer $i$ on a lineup $j$, with $Y_{ij} =1$ if the actual data plot is chosen, otherwise $Y_{ij} =0$.
-The factors are examined in association with the observer's response using a logistic regression with random effect terms:
-% 
-% \begin{equation} \label{eqn:demographic_response}
-% g( \pi_{ij} )= \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j,  
-% \end{equation}
-% 
-% where $\pi_{ij}=  E(Y_{ij})$ is the probability that  observer $i$ picks the actual data plot from lineup $j$, $\mu$ is an overall population average, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are the effects of age group $k(i)$, gender $l(i)$, education level $m(i)$ and country name $s(i)$, respectively, for observer $i$.
-The term $\ell_j$ is a random intercept predicting lineup difficulty level and we assume independence and normality of the errors, i.e.\ $\ell_j \sim N(0, \sigma_\ell^2)$.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-% 
-% 
-% Similarly, we model the  time  an observer takes to identify a panel from a lineup.
-Let $Z_{ij}$ denote the logarithm of time taken for  observer $i$ to evaluate  lineup $j$.
-Let $\mu_{ij}=  E(Z_{ij})$ be the average of the (log) time taken by  observer $i$ to pick a panel from lineup $j$.
-We model this in a mixed effects model of the same structure as model~(\ref{eqn:demographic_response}) given as:
-% %\hh{
-% \begin{equation} \label{eqn:demographic_time}
-% Z_{ij} = \mu + \alpha_{k(i)} + \gamma_{l(i)} + \tau_{m(i)}+ \kappa_{s(i)} + \ell_j+ \epsilon_{ij},  
-% \end{equation}
-% where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$, $\gamma$, $\tau$ and  $\kappa$ are as described in model~\eqref{eqn:demographic_response}, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, with $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ij} \sim N(0, \sigma^2)$.
-% 
-% \subsection{Learning Trend} Learning trend of a subject can be observed in terms of performance over successive responses when multiple lineups are shown for evaluation.
-Experiments 5, 6 and 7 were used for this.
-Each subject was shown a total of 10 lineups randomly selected from a pool of lineups.
-The lineups are not necessarily of the same difficulty level, but the order of lineups was randomized.
-The responses of the lineups were recorded by attempt 1 through 10.
-Attempt 1 means that the response is for the first lineup the observer evaluates and attempt 10 refers to the response for the 10th lineup.
-The goal is to estimate whether performance of the observer improves, or changes, from attempt 1 to attempt 10.
-% 
-% It should be noted that we are examining the observer's performance, when we model response as detected or not, but this is not the goal of visual inference.
-Visual inference is constructed to measure the significance of structure discovered in data.
-It is expected that some observers will be more skilled at reading data plots, and hence, more readily detect the plot that is different.
-It is also expected that as observers gain experience in evaluating lineups that they become more proficient in reading data plots, particularly if feedback is given on whether the actual data plot was chosen or not.
-Choosing the actual data plot will be more difficult in some lineups than others, and indeed should happen purely by chance in some lineups.
-So in this context, detected, or not, is used as a response to examine individual differences.
-% 
-% Let $Y_{ijk}$ denote the response from observer $i$ on lineup $j$ at their $k$th evaluation attempt, where $Y_{ijk}=1$ if the observer detected the actual data plot otherwise  $Y_{ijk}=0$.
-Let $\pi_{ijk}=  E(Y_{ijk})$ be the probability that  observer $i$ picks the actual data plot from lineup $j$ in their $k$th attempt.
-Learning trend is assessed using a generalized mixed effects model of the form
-% \begin{equation} \label{eqn:trend_response}
-% g( \pi_{ijk} )= \mu + \alpha_k + u_i +  a_{i} k + \ell_j,  
-% \end{equation}
-% where $\mu$ is an overall population average, $\alpha_k$ is the effect of the $k$th attempt on the probability, using the first attempt as reference, $\alpha_1 = 0$, and $k = 1, ..., K$, $u_i$ and $a_i$ are observer specific random effects, $i = 1, ..., I$.
-The term, $u_i$ is a random intercept, describing a basic subject-specific ability, with $u_i \sim N(0, \sigma_u^2)$.
-% The term $a_i$ is a random slope capturing an individual's specific learning effect over the course of $K$ attempts, where $a_i \sim N(0, \sigma_a^2)$. 
-% For $\ell_j$ a normal distribution, $N(0, \sigma_\ell^2)$, is assumed, and $\ell_j$ is a random intercept predicting lineup difficulty level.
-$g(.)$ denotes the {\it logit} link function $g(\pi)=\log(\pi) - \log(1-\pi); 0 \le \pi \le 1$.
-The inverse link function, $g^{-1}(.)$, from equation~\ref{eqn:trend_response} leads to the estimate of the subject and the lineup specific probability of successful evaluation in the $k$th attempt by a single observer as
-% \begin{equation} \label{eqn:trend_power}
-% \hat p_{ijk} =  g^{-1}(\hat{\mu} + \hat{\alpha}_k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j).
-% \end{equation}
-% 
-% When time taken to evaluate a lineup is used as the response, let $Z_{ijk}$ denote the logarithm of time taken for an observer $i$ to evaluate a  lineup $j$ in his/her $k$th attempt.
-Let $\mu_{ijk}=  E(Z_{ijk})$ be the average of the (log) of time taken by  observer $i$ to choosing a  panel from lineup $j$ in his/her $k$th attempt.
-We evaluate this in a mixed effects model of the form
-% 
-% \begin{equation} \label{eqn:trend_time}
-% Z_{ijk} = \mu + \alpha_1 + \alpha k + u_i +  a_{i} k + \ell_j + \epsilon_{ijk},  
-% \end{equation}
-% where $\mu$ represents overall average of log time taken by an observer to evaluate a lineup, $\alpha$ is the average change in log time taken for each additional attempt,  $\alpha_1$ is an offset in log time taken for the first attempt.
-All other effects are random effects: $u_i$ is a subject-specific intercept representing individual speed of an observer with $u_i \sim N(0, \sigma_u^2)$, $a_i$ is a subject-specific slope representing the deviation of the speed-up (or -down) by attempt $k$, with $a_i \sim N(0, \sigma_a^2)$, $\ell_j$ is a lineup-specific random effect for the time needed to evaluate a lineup, $\ell_j \sim N(0, \sigma_\ell^2)$ and the overall error $\epsilon_{ijk} \sim N(0, \sigma^2)$.
-% Equation~\ref{eqn:trend_time} leads to the estimate of the subject and the lineup specific time taken for an evaluation in $k$th attempt by a single observer as 
-% \begin{equation} \label{eqn:trend_time_est}
-% \hat \mu_{ijk} =  \hat{\mu} + \hat{\alpha_1}+ \hat{\alpha}k + \hat{u}_i +  \hat{a}_i k + \hat{\ell}_j.
-% \end{equation}
-% 
-% To fit all these mixed effect models the function {\tt lmer()} is used from R package {\tt lme4} by \cite{lme4:2015, lme4:paper}.
-We employ a normal approximation to obtain $p$-values corresponding to fixed effect parameters estimates.
-% 
-% \subsection{Location Effect} \label{sec:location_design} Experiment~9 studied significant expression in an RNA-seq study, and was designed so that location effect of the actual data plot in a lineup could also be assessed.
-The data used, documented in \cite{atwood:2013}, measures gene expression of soybean by RNA-seq methods.
-Two factors were of primary interest a main effect for genotype and an interaction effect between genotype and treatment condition.
-% 
-% In large studies such as this there is a valid question whether the data exhibits any structure at all, or if the small $p$-values are simply occurring by chance, from the massive multiple testing.
-This overall significance is studied using visual inference  in \citet{tengfei:2013}.
-% 
-% In order to study the effect that  location has on evaluating lineups, we used multiple lineups for each of the data plots.
-For each data plot, five sets of null plots were generated and the actual plot was randomly placed in one of five different locations in a lineup of size 20.
-% For the genotype effect, the locations were  1, 8, 12, 17, 20 and locations 2, 9, 12, 16, 20 were used for the interaction effect.
-%Five different sets of null plots were used to produce 5 lineups for each location position, creating
-% Overall this created a total of 25 lineups for studying the genotype effect, and another 25 lineups for studying the interaction effect.
-Each observer saw three lineups, one for genotype, one for interaction, and one easy lineup that was used to help clean the data.
-% 
-% To examine if the difference in detection rate among the locations is statistically significant a one-way multivariate analysis of variance (MANOVA) model is fit to the data.
-% Let $\mathbf{Y}=(Y_1,Y_2, ...
-, Y_p)^\top$ be a vector of random variables with dimension $p = 5$, the total number of null sets, and let $\mathbf{Y}_{ij}$ represent the $j$th vector response for  location $i$ with $i=1,2, ..., I$ for $U=5$.
-Because the same data plot is shown in each lineup, it is assumed that there could be some association between the responses for each null set, which suggests the MANOVA model rather than a univariate ANOVA.
-The MANOVA model
-% \begin{equation}\label{manova}
-% \mathbf{Y}_{ij} = \mu_{i} + \epsilon_{ij}
-% \end{equation}
-% where $\mu_{i}= (\mu_{1i},\mu_{2i}, ..., \mu_{pi})^\top$ is the mean vector for location $i$ and $Var(\epsilon_{ij})=\Sigma$, tests for significant difference between the means. 
-% 
-% \subsection{Data Collection Methods}  Human subjects were recruited to evaluate the experimental lineups through MTurk \citep{turk}.
-It is an online work place where people from around the world can sign up for so-called `HIT's, human intelligence tasks, generally short tasks that are humans are typically better at solving than computers.
-Usually tasks are very simple and no specialized training is required to do them.
-Tasks are designed for anyone to do but some tasks may require some skills depending on the recruiters' need.
-%Each task is usually planned to completed in a short time.
-%  For completing a HIT workers are paid a small amount of money, on the order of minimum wage in the USA. 
-% 
-% 
-% We designed and developed a web application which enables the display of lineups to  observers as per experimental need.
-The MTurk workers were re-directed to this web application to complete their assigned tasks.
-Responses were collected, stored automatically into a local database server, along with demographic information, age group, gender and education level.
-The time taken for each evaluation is computed based on the time the plot was shown and the time the feedback was received.
-Location of an observer is determined based on the ip address of the observer.
-%
 
@@
  \subsection{Overview of the Data} In ten experimental studies, a total of 2,340 participants from 70 countries, primarily the United States and India, provided feedback data on lineups, ensuring a diverse pool of respondents in terms of geography, gender, age, and education.
-The demographic details, as summarized in Table~\ref{tbl:turker_location}, revealed a high number of female participants across most countries, which was an unexpected but positive outcome.
+There were a high number of female participants(42\%) across most countries, which was an unexpected but positive outcome(see Table~\ref{tbl:turker_location} in Appendix~\ref{sec:tbl_fig}).
 Notably, countries like Canada, Romania, the United Kingdom, and Macedonia also had notable participant representation.
@@
 
-In total, 1,912 lineups were evaluated across the studies, with each participant reviewing at least 10 lineups, except for experiment~9, where they reviewed three.
+A total of 1,912 lineups were evaluated across the studies, with each participant reviewing at least 10 lineups, except for experiment~9, where they reviewed three.
 To ensure quality, participants were given test plots to evaluate first, and incorrect responses led to their data being excluded from the analysis.
@@
 
-%A total of 2340 participants provided feedback data on the lineups in ten different experimental studies.
-Figure~\ref{fig:turker_location} displays the locations of participants around the world.
-Most of the participants were from the United States and India.
-There were 70 other different countries represented.
-This provides a diverse pool of participants.
-The diversity in not only  geographic but also in gender, age group and education level as can be seen in Table~\ref{tbl:demographics}.
-The large number of female participants from all countries was a pleasant surprise to us.
-% 
-% 
-% Besides the United States and India, countries such as Canada, Romania, the United Kingdom and Macedonia have more than 10 participants each.
-The remaining 64 countries have fewer than 10 participants each.
-The distribution of participants is similar in all ten experiments.
-% 
-% 
-% 
-% The largest number of participants falls within the age group of 18 to 25, with the majority being between 18 to 35. Many older participants also took part in the experiments. 
-% The United States shows a more uniform participation beyond age 30,  participants from India are primarily younger. 
-% 
-% In terms of academic background, the majority of participants had a graduate degree (a total of 899 participants, about 38.42\%).
-Examining the participants' location in conjunction with the academic background reveals that there is an above average  number of participants from India reporting to have a graduate degree.
-In retrospect this turned out to be a culturally based misunderstanding, as any university degree in India is considered to be a graduate degree, while the North American definition of a graduate degree refers to a Masters level education or above.
-Most of the  participants from the U.S.
-report to have an undergraduate degree or at least have some undergraduate courses.
-% 
-% Compared to data from the 2010 US Census \cite{census}, participants were both relatively younger, and more highly educated than the general US population. 
-% Being turkers implicitly assures that the participants in these experiments have a relatively high level of technological fluency.
-In fact, the demographics of participants align somewhat better with demographics of internet users published by
-% \citet{pew}.
-% 
-% The distribution of male and female participants are similar among all age groups except for age group 18-25 in India where the proportion of female participants is lower.
-The distribution of education levels also differs slightly across countries for this age group.
-% 
-% 
-% A total of 1912 lineups were evaluated in the ten experiments. Each person evaluated at least 10 lineups except for experiment~9 where each participant evaluated three lineups.
-% As a measure to ensure a high quality in the evaluations,
-%  a test plot was shown to each observer, and the performance on this test plot was used for screening purposes. 
-% Test plots were chosen in a way that any serious attempt at answering the lineup would result in a correct answer.
-% Data from participants who gave the wrong answer on the test plot 
-% led to the removal of all of their answers from the analysis. 
-% Results from test plots were also not considered in any of the analyses.
-%  In some cases  participants chose to not provide their demographic information.
-Also, for some ip address, the actual geographical locations could not be retrieved.
-This resulted in some missing geographic information.
-
 \subsection{Demographic Factors}
 
-To explore the significance of the demographic factors model~\eqref{eqn:demographic_response} is fit to the data, with age, country, education and gender as fixed effects.
-To estimate the significance of each of the factors, reduced models are fit with that factor removed from the model.
-Table~\ref{tbl:anova_factor} summarizes the results.
+To explore the significance of demographic factors in detecting the data plot in the lineup, a mixed-effects model~\eqref{eqn:demographic_response} is fitted to the data, with age, country, education, and gender as fixed effects.
+To estimate the significance of each of the factors, reduced models are fitted with that factor removed from the model.
+Table~\ref{tbl:model_result_demographics} summarizes the results.
 Except for gender, all of the demographic factors show significant differences in  detection rates.
@@
 
-Parameter estimates from the model fits are shown in Table~\ref{tbl:model_result_demographics}.
 The first level of each factor serves as the baseline for the model.
@@
 A higher degree corresponded to longer time, and males were on average faster than females.
+
+Figure~\ref{fig:demographic_effect} supports the model results.
+The average time taken to respond (natural log) and detection rate are computed by lineup for the different demographic factor levels, and displayed using boxplots.
+Mean values are shown as dots inside the boxplot, as this statistic rather than the median is used in the models.
+The means and medians for response time (top row) are essentially equal.
+There are some differences between the two in the detection rate plots (bottom row).
+The differences in these values for different factors, that was tested by the models, can be seen in the plots.
+However, particularly for detection rate, the variation is huge.
+A large component of the variation is difficulty of the lineup.
+In some lineups the actual data plot was distinctively different from the null plots, and we would expect that the detection rate would be close to 100\%.
+In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond.
 
@@
 
-Figure~\ref{fig:demographic_effect} supports the model results.
-The average time taken to respond (natural log) and detection rate are computed by lineup for the different demographic factor levels, and displayed using boxplots.
-Mean values are shown as dots inside the boxplot, as this statistic rather than the median is used in the models.
-The means and medians for response time (top row) are essentially equal.
-There are some differences between the two in the detection rate plots (bottom row).
-The differences in these values for different factors, that was tested by the models, can be seen in the plots.
-However, particularly for detection rate, the variation is huge.
-A large component of the variation is difficulty of the lineup.
-In some lineups the actual data plot was distinctively different from the null plots, and we would expect that the detection rate would be close to 100\%.
-In other lineups there was no difference, making it very difficult to evaluate, with low detection rate and longer time to respond.
 In Table~\ref{tbl:model_result_demographics} notice that lineup specific error is estimated as 2.27 which is much higher than any of the other parameter estimates in the model.
@@
 $$ \frac {\exp(-0.63-0.13)}{1+\exp(-0.63-0.13)}-  \frac {\exp(-0.63+0.16)}{1+\exp(-0.63+0.16)} = 0.319- 0.385 =-0.066.$$
+
 The person with some graduate course work picks the data plot  on average in 31.9\% of lineups of average difficulty, as compared to 38.5\% if they have an undergraduate degree.
@@
 \subsection{Learning Trend} Models \eqref{eqn:trend_response} and \eqref{eqn:trend_time} are fitted to the data from experiments~5,~6, and~7 separately.
+Each participant evaluated 10 lineups in 10 attempts.
 In this model, attempt is fitted as a factor variable to allow for all possible non-linear learning trends.
@@
 
-Table~\ref{tbl:model_result_response} displays the parameter estimates and $p$-values of fixed effect estimates of model~\eqref{eqn:trend_response} examining detection rate.
+The parameter estimates and $p$-values of the fixed effects from model~\eqref{eqn:trend_response} are presented in Table~\ref{tbl:model_result_response} in the Appendix~\ref{sec:exp_design}.
 Only for experiment~6 there is some evidence of a learning curve.
 There are marginally small $p$-values for most of the attempt levels, and positive parameter estimates, suggesting that more attempts increases detection rate.
-
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 23:54:25 2017
-% \begin{table}[hbtp]
-% \centering
-% \caption{\label{tbl:model_result_response}Parameter estimates of models \eqref{eqn:trend_response} fitted to data from three different experiments using detection rate as the response to assess learning trend.
-Attempt number is fitted as a factor to enable modeling any sort of non-linear learning trend.
-Only experiment~6 shows some evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
-% \scalebox{.9}{
-% \begin{tabular}{rr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
-% \cline{2-5} \cline{7-10} \cline{12-15} 
-% & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
-% \cline{2-5} \cline{7-10} \cline{12-15} 
-% Effect & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  \\ 
-% \cline{2-5} \cline{7-10} \cline{12-15} 
-% Fixed \\
-% $\mu$ & -1.32 & -1.73 & -0.92 & *** &   & -0.16 & -0.46 & 0.13 &  &   & -1.73 & -2.81 & -0.66 & ** \\ [3pt]
-%   $\alpha2$ & 0.30 & -0.17 & 0.78 &  &   & 0.26 & -0.07 & 0.58 &  &   & -0.40 & -1.23 & 0.44 &  \\ 
-%   $\alpha3$ & -0.20 & -0.68 & 0.29 &  &   & 0.31 & -0.02 & 0.63 & . &   & -0.15 & -0.99 & 0.68 &  \\ 
-%   $\alpha4$ & 0.13 & -0.35 & 0.61 &  &   & 0.34 & 0.01 & 0.66 & * &   & -0.41 & -1.24 & 0.42 &  \\ 
-%   $\alpha5$ & 0.35 & -0.14 & 0.83 &  &   & 0.34 & 0.02 & 0.67 & * &   & -0.10 & -0.92 & 0.73 &  \\ 
-%   $\alpha6$ & 0.10 & -0.40 & 0.60 &  &   & 0.20 & -0.12 & 0.53 &  &   & 0.03 & -0.84 & 0.89 &  \\ 
-%   $\alpha7$ & 0.33 & -0.17 & 0.82 &  &   & 0.11 & -0.22 & 0.43 &  &   & -0.00 & -0.84 & 0.84 &  \\ 
-%   $\alpha8$ & -0.01 & -0.51 & 0.50 &  &   & 0.31 & -0.01 & 0.64 & . &   & -0.06 & -0.88 & 0.76 &  \\ 
-%   $\alpha9$ & -0.20 & -0.70 & 0.30 &  &   & 0.34 & 0.01 & 0.67 & * &   & 0.21 & -0.67 & 1.08 &  \\ 
-%   $\alpha10$ & 0.51 & 0.02 & 1.01 & * &   & 0.19 & -0.14 & 0.53 &  &   & -0.20 & -1.07 & 0.67 &  \\ 
-%   Random \\
-%   $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.91}  & \multicolumn{5}{l}{\phantom{-}0.89} & 0.82   \\ 
-%   $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.04}  & \multicolumn{5}{l}{\phantom{-}0.04} & 0.05   \\ 
-%   $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}1.47}  & \multicolumn{5}{l}{\phantom{-}1.43} & 3.38   \\ \cline{2-5} \cline{7-10} \cline{12-15} 
-% \\
-% \hline
-% \multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
-% \end{tabular}}
-% \end{table}
-
-% 
-% 
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 23:54:25 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{rrrrllrrrllrrrl}
-%   \hline
-%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
-%   \hline
-% $\mu$ & -1.32 & -1.73 & -0.92 & *** &   & -0.16 & -0.46 & 0.13 &  &   & -1.73 & -2.81 & -0.66 & ** \\ 
-%   $\alpha2$ & 0.30 & -0.17 & 0.78 &  &   & 0.26 & -0.07 & 0.58 &  &   & -0.40 & -1.23 & 0.44 &  \\ 
-%   $\alpha3$ & -0.20 & -0.68 & 0.29 &  &   & 0.31 & -0.02 & 0.63 & . &   & -0.15 & -0.99 & 0.68 &  \\ 
-%   $\alpha4$ & 0.13 & -0.35 & 0.61 &  &   & 0.34 & 0.01 & 0.66 & * &   & -0.41 & -1.24 & 0.42 &  \\ 
-%   $\alpha5$ & 0.35 & -0.14 & 0.83 &  &   & 0.34 & 0.02 & 0.67 & * &   & -0.10 & -0.92 & 0.73 &  \\ 
-%   $\alpha6$ & 0.10 & -0.40 & 0.60 &  &   & 0.20 & -0.12 & 0.53 &  &   & 0.03 & -0.84 & 0.89 &  \\ 
-%   $\alpha7$ & 0.33 & -0.17 & 0.82 &  &   & 0.11 & -0.22 & 0.43 &  &   & -0.00 & -0.84 & 0.84 &  \\ 
-%   $\alpha8$ & -0.01 & -0.51 & 0.50 &  &   & 0.31 & -0.01 & 0.64 & . &   & -0.06 & -0.88 & 0.76 &  \\ 
-%   $\alpha9$ & -0.20 & -0.70 & 0.30 &  &   & 0.34 & 0.01 & 0.67 & * &   & 0.21 & -0.67 & 1.08 &  \\ 
-%   $\alpha10$ & 0.51 & 0.02 & 1.01 & * &   & 0.19 & -0.14 & 0.53 &  &   & -0.20 & -1.07 & 0.67 &  \\ 
-%   $\sigma^2_u$ & 0.91 &  &  &  &   & 0.89 &  &  &  &   & 0.82 &  &  &  \\ 
-%   $\sigma^2_a$ & 0.04 &  &  &  &   & 0.04 &  &  &  &   & 0.05 &  &  &  \\ 
-%   $\sigma^2_l$ & 1.47 &  &  &  &   & 1.43 &  &  &  &   & 3.38 &  &  &  \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
-% 
-%
 
@@
 
-Table~\ref{tbl:model_result_time} presents the results of model~\eqref{eqn:trend_time} where response time is examined with respect to attempt.
+Table~\ref{tbl:model_result_time} in the Appendix~\ref{sec:exp_design} presents the results of model~\eqref{eqn:trend_time} where response time is examined with respect to attempt.
 Attempt is modeled as a linear covariate here, with a shift factor used to adjust for additional time on the first lineup evaluated.
@@
 Later pages of the web site were identical  except for the lineup.
-
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 23:37:55 2017
-% \begin{table}[bhtp]
-% \caption{\label{tbl:model_result_time}Parameter estimates of model~\eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup.
-Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments~5,~6 and~7.}
-% \centering
-% \scalebox{0.9}{
-% \begin{tabular}{lr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
-% \cline{2-5} \cline{7-10} \cline{12-15} 
-% & \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
-% 
-% \cline{2-5} \cline{7-10} \cline{12-15} 
-% \multicolumn{2}{c}{Effect \hfill Est} & 2.5\% & 97.5\% &  &  & Est & 2.5\% & 97.5\% & &  & Est & 2.5\% & 97.5\% & \\ 
-%   \cline{2-5} \cline{7-10} \cline{12-15}
-% \multicolumn{2}{l}{Fixed}\\
-% $\mu$ & 3.80 & 3.72 & 3.88 & *** &   & 3.89 & 3.83 & 3.96 & *** &   & 3.75 & 3.64 & 3.85 & *** \\ 
-%   $\alpha_1$ & 0.32 & 0.25 & 0.39 & *** &   & 0.34 & 0.28 & 0.40 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
-%   $\alpha$ & -0.04 & -0.05 & -0.03 & *** &   & -0.04 & -0.05 & -0.03 & *** &   & -0.03 & -0.04 & -0.02 & *** \\ 
-% \multicolumn{2}{l}{Random}\\
-%   $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.52}   & \multicolumn{5}{l}{\phantom{-}0.49}  & 0.37   \\ 
-%   $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.03}  & \multicolumn{5}{l}{\phantom{-}0.05}   & 0.05   \\ 
-%   $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}0.09}   & \multicolumn{5}{l}{\phantom{-}0.20}   & 0.24   \\
-%   $\sigma^2$ & \multicolumn{5}{l}{\phantom{-}0.46}   & \multicolumn{5}{l}{\phantom{-}0.50}  & 0.45  \\ 
-%    \cline{2-5} \cline{7-10} \cline{12-15}
-% \end{tabular}}
-% \end{table}
-%
-
-% 
-% % latex table generated in R 3.4.1 by xtable 1.8-2 package
-% % Wed Oct 18 23:37:55 2017
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{rrrrllrrrllrrrl}
-%   \hline
-%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
-%   \hline
-% $\mu$ & 3.80 & 3.72 & 3.88 & *** &   & 3.89 & 3.83 & 3.96 & *** &   & 3.75 & 3.64 & 3.85 & *** \\ 
-%   $\alpha_1$ & 0.32 & 0.25 & 0.39 & *** &   & 0.34 & 0.28 & 0.40 & *** &   & 0.28 & 0.18 & 0.38 & *** \\ 
-%   $\alpha$ & -0.04 & -0.05 & -0.03 & *** &   & -0.04 & -0.05 & -0.03 & *** &   & -0.03 & -0.04 & -0.02 & *** \\ 
-%   $\sigma^2_u$ & 0.52 &  &  &  &   & 0.49 &  &  &  &   & 0.37 &  &  &  \\ 
-%   $\sigma^2_a$ & 0.03 &  &  &  &   & 0.05 &  &  &  &   & 0.05 &  &  &  \\ 
-%   $\sigma^2_l$ & 0.09 &  &  &  &   & 0.20 &  &  &  &   & 0.24 &  &  &  \\ 
-%   $\sigma^2$ & 0.46 &  &  &  &   & 0.50 &  &  &  &   & 0.45 &  &  &  \\ 
-%    \hline
-% \end{tabular}
-% \end{table}
 
@@
 
-% \subsection{Location Effect}
-% 
-% A total of 111 subjects was recruited to evaluate lineups designed to investigate the location effect of the actual data plot in the lineup as described in Section~\ref{sec:location_design}.
-Each subject evaluated two lineups;  one for Interaction effect and the other for Genotype.
-In total there were 222 responses for 50 lineups.
-The data on the test lineup were excluded from the analysis.
-% 
-% We are investigating location effect in two ways: (1) location effect of each individual plot location, and (2) location effect of inside/outside panel positions.
-% 
-% Figure~\ref{fig:location_effect} shows an overview of detection rates for both of the effects studied in experiment~9.
-Each dot corresponds to one lineup.
-The size of each dot gives the number of evaluations.
-Lineups with  the same set of null plots are connected by lines.
-% %The detection rate for each data plot location is shown in colored by null sets. 
-% The overall average detection rate for each location is shown using dashed lines.
-Size of the points represent the number of responses.
-For some locations we have as many as 10 responses.
-For location 1, we did not have any responses for null set 1 in one of the interaction lineups.
-% 
-% We observe some variability in the performance due to different null sets even though the same actual data plot was used for all of  these null sets.
-The lineup protocol is using a finite, small sample (19 in these experiments) of all possible null sets, so it is possible to see some variability in performance depending on the null set.
-If null sets come from the ``extremes'' of the sampling distribution they may have structure that is more extreme than the actual data plot, making the lineup more difficult to evaluate.
-From the plot we can see that null set 3 appeared to be more difficult in the Interaction lineup, because the detection rate was lower regardless of the position of the actual data plot.
-This null set effect was tested, and a significant difference was found for null set 3 of the interaction lineup.
-The rest of the null sets do not show any differences at  a 1\% significance level.
-% 
-% 
-% \begin{figure}[htbp] 
-%    \centering
-%     \includegraphics[width=6.5in]{proportion_nulls_guide.pdf} 
-%    \caption{Location of data plot in the lineup and proportion correct (detection rate) for both Interaction and Genotype effect.
-Each colored line represents a null set and the size of the dots represents number of responses.
-The overall average proportions are shown by dashed lines.
-The actual data plot locations are shaded grey on the top panels to demonstrate their relative positions on a lineup.}
-%    \label{fig:location_effect}
-% \end{figure}
-% 
-% 
-% Model~\eqref{manova} was fit to the data to test, if  detection rates were significantly different for the different locations.
-For this we use the {\it manova()} function provided in the {\it stats} package of the software R \citep{R}.
-The results are shown in Table~\ref{tbl:manova}.
-The $p$-values for both the Interaction and the Genotype effect are much bigger than the conventional threshold of 0.05 suggesting that location of the actual data plot in the lineup does not significantly affect performance.
-% 
-% \begin{table}[hbtp]
-% \caption{The summary of the results obtained by fitting MANOVA model~\eqref{manova} using Wilks test.}
-% \begin{center}
-% \begin{tabular}{ccccccccc}
-%   \hline \hline
-%  Location & && & \multicolumn{3}{c} {Degrees of Freedom}  & F test \\
-%  \cline{5-7}
-%  Effect & DF & Wilks & Approx. F & Numerator& Denominator &Residual & $p$-value\\
-%   \hline
-%   Interaction& 3&0.07986 & 0.592 &15& \phantom{0}5.92 & 6 & 0.8082 \\ 
-%   Genotype & 4&0.01689 & 1.722 & 20&14.22 & 8 & 0.1488 \\ 
-%    \hline
-% \end{tabular}
-% \end{center}
-% \label{tbl:manova}
-% \end{table}
-% 
-% For the second approach in investigating the location effect, we distinguish between inner panels 7, 8, 9, 12, 13, 14 and outer panels (any panel with an outside edge).
-From Figure~\ref{fig:location_effect} we see that location 9, 12 are inner panels for Interaction effect and locations 8, 12 are inner panels for Genotype.
-Model~\eqref{manova} fitted with two locations effects, inner and outer, did not result in any significant differences.
-
 % ===================================================================================
@@
 However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea, at least for short term learning.
-
-The simulation experiment reveals that there is no significant effect of location on the actual data plot in the lineup.
-This is important as the visual statistical inference procedure prescribes that the data plot be placed at random in the lineup.
-This paper suggests that any random place in a lineup is as good as other places in the lineup.
-Even though there are variations on the performance depending on different null sets, their impact on probability to correctly evaluate a lineup is very negligible.
 
@@
 
-\section{\\Some tables and figures}
+% ===================================================================================
+\section{Some tables and figures}\label{sec:tbl_fig}
+%------------------------------------------------------------------------------------
 
@@
 \begin{table}[hbtp]
-\caption{Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
+\caption{\label{tbl:turker_location}Demographic information of the subjects participating in the MTurk experiments. Average time taken for evaluating a lineup is shown in seconds.}
 \centering

### 2024-12-31 wiley_paper/wiley_paper.Rnw
Add files via upload

@@
   \begin{center}
-    {\bf \LARGE Human Factors Influencing Visual Statistical Inference}
+    {\bf \LARGE Effect of Sociological Factors on Visual Statistical Inference}
 \end{center}
@@
 The effectiveness of this method, measured by power, depends on combining evaluations from multiple observers.
-Factors influencing power include observer demographics, visual skills, experience, the sample of null plots, plot position, and signal strength.
+Factors influencing power include observer demographics, visual skills, experience, the sample of null plots, and signal strength.
 This paper examines these factors through studies using Amazon's Mechanical Turk, finding individual skills vary but demographics have little impact.
-Learning increases speed but not accuracy, and plot position doesn’t affect results.
+Learning increases speed but not accuracy.
 
@@
 
-\if1\blindDoc
-{
-%------------------------------------------------------------------------------------
-\paragraph{Acknowledgments}
-%------------------------------------------------------------------------------------
+\section{Funding Information}
 This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
-} \fi
+
+% \if1\blindDoc
+% {
+% %------------------------------------------------------------------------------------
+% \paragraph{Acknowledgments}
+% %------------------------------------------------------------------------------------
+% This work was funded in part by National Science Foundation grant DMS 1007697. All studies were conducted with approval from the Institutional Review Board IRB 10-347.
+% } \fi
 

### 2025-04-17 wiley_paper/response-to-review.qmd
made response into quarto doc

---
title: "Response to Review"
author: "Majumder, Cook, Hofmann"
format: pdf
---

Thank you for the careful and positive reviews. *Our response to comments is written in italics*.

## From the editor

Good set of comments. FYI, feel free to change the review title (if you wish)
and add more material (as an on-line journal, page limits not really active).
Look forward to seeing your revision.   D Scott co-editor  scottdw@rice.edu

Response: *Thank you*

I am writing in regards to Manuscript ID EOCS-756 entitled "Effect of Sociological Factors on Visual Statistical Inference," which you submitted to WIREs Computational Statistics. The reviewer(s) have suggested major revisions to improve the manuscript. Their comments are included at the bottom of this letter; I invite you to respond to them and revise your manuscript.

Please submit your revision within 6 weeks. If you need extra time, please let us know so that we can keep our systems updated for editorial planning purposes.

A revised version of your manuscript that takes into account the comments of the reviewer(s) will be reconsidered for publication. Please note that revising your manuscript does not guarantee eventual acceptance, and that your revision will be subject to re-review before a decision is rendered.

If you are updating your Main Document in Word, make sure that "Track Changes" is enabled so that the editors can easily find your amendments. If you are using LaTeX, make sure that revisions are made in red font.

You can upload your revised manuscript using this direct link: *** PLEASE NOTE: This is a two-step process. After clicking on the link, you will be directed to a webpage to confirm. ***

https://urldefense.com/v3/__https://mc.manuscriptcentral.com/compstats?URL_MASK=438fdf874ee14f659af35cd0e26d4ced__;!!PvXuogZ4sRB2p-tU!GOtJWWHikD4k_JosJN5bkg2Kphby7E2KmkmqHJ4dZ95gyJOEY62FgPhrkDJhfcMHLuoagvllP182luVdjkcYTREEH-IXMjM$ .

When you submit your revision, you must provide a point-by-point summary of how each of the issues raised by the reviewer(s) were addressed. Please note that the space provided is for plain text only, as the system will not accept any formatting. For each criticism raised, please provide:
1. The criticism from the reviewer.
2. The response and explanation of how the manuscript has been modified in response to the criticism.

IMPORTANT: We have your original files; therefore, please delete the file(s) that you wish to replace and then upload the revised file(s). You will be able to upload both marked and clean versions of your Main Document. Please note that figures must be in PRODUCTION-READY format.

Please do not hesitate to get in touch with us if you have any queries or concerns.

Once again, thank you for submitting your article to WIREs Computational Statistics. I  look forward to receiving your revision.

Sincerely,

Hashini Selvam
On behalf of
Dr. David Scott
Co-Editor-in-Chief and Review Editor, WIREs Computational Statistics
scottdw@rice.edu

## Reviewer: 1

### Comments to the Author(s)

This is an interesting study where the authors attempt to examine the impact of demographics on evaluating a "line up" of statistical graphics to visually pick out a particular data set compared to several null graphs. The main results seem to be that while demographics do seem to vary the results from a statistical perspective (i.e. the p-values are significant, given the large sample size), the impact from a scientifically meaningful perspective is minimal. I have a several specific comments to highlight below.

Author response:  *The authors agree with this comment. The results support our conclusion that the visual test is robust to variations in human factors.*

### Specific comments:

1. The authors choose to examine two outcomes, log time (how long it takes the user to complete the task) and detection rate. It is not clear to me why a joint model is not shown. It seems relevant, particularly because there are clear demographic differences seen (albeit quantitatively small) that it would make sense to examine detection rate with the demographics *and* time in model A.1 to see whether, for example, the demographic differences seen in A.1 are fully explained by time as there are definitely impacts of demographics on time and it seems plausible that time would impact accuracy. Likewise, it is possible that after adjusting for time a learning trend may be more visible.

Author response: *A joint model is fitted to simultaneously analyze detection accuracy and log time taken. The estimated covariate effects in the joint model are largely consistent with those obtained from the separate models, with one notable difference: the estimated variance of the lineup-level random intercept for detection rate (lieup difficulty) is slightly higher in the joint model (2.37) compared to the separate model (2.27), suggesting that modeling the outcomes jointly may better account for heterogeneity across lineups. Importantly, the joint model also provides an estimate of the latent correlation between the lineup-level random intercepts for time taken and detection rate. This correlation is estimated to be -0.55, indicating that lineups which take longer to evaluate tend to be associated with lower detection rates. This result aligns with intuition: more difficult lineups likely require more evaluation time, yet are also more prone to non-detection. These findings and interpretations have been incorporated into the revised manuscript at the end of section 3.2.*

2. The authors state that the largest effect is by far the lineup difficulty. This mixture of difficulties makes it hard to learn much from the overall results (e.g. it makes the bottom portion of Figure 1 very hard to parse). Figure 2 (B) seems to show a the impact of a variable (education) across three different classifications of "difficulty" -- from this plot it seems that the easy line ups and difficult line ups are providing very little information, as the proportion classified correctly is very close to 1 or 0 (at least for the 18-25 year old females this plot was drawn for). It may be more compelling to show the results excluding these and focusing only on the medium lineups (at least as a subgroup analysis).

Author response: *Lineup difficulties were predetermined as part of the experimental design, and each participant was randomly assigned 10 lineups spanning a range of difficulty levels. The fitted model accounts for lineup-level variation through random effects, effectively adjusting for lineup difficulty when estimating the effects of covariates on detection rates. For this reason, separate analyses by lineup difficulty are not the focus of this study?* XXX DC suggests changing the caption further: "Some differences in means of various demographic factors are observed. Variability in detection rate indicates large variability in lineup difficulties." to "Some differences in means of various demographic factors are observed, but the variation is large, which we will see is primarily due to difference in the controlled factor, lineup difficulty."

3. Building on the previous comment, it is possible (likely?) that the reason there is no "learning trend" seen is due to these very easy or very hard lineups. I would like to see a model for this that only includes the medium lineups.

Author response: *The model accounts for lineup difficulty when estimating the learning trend. It would be interesting to examine learning trend more deeply, in particular in relation to using lineups for teaching people how to read data plots. However, learning trend in itself is not the focus of this paper. Here we are assessing whether the human evaluation of lineups is consistent across demographic factors and trials. Each subject evaluated 10 lineups, three of each difficulty level. It is insufficient to only use one of these difficulty levels to assess the learning trend.* XXX **We have added additional discussion on this point at the end of the paper.** Change this sentence in conclusion: "However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea, at least for short term learning." to "Although this study showed that there was not a learning effect in the 10 lineups presented to each subject, we do hope that there is a possibility that lineups might be used for helping people to learn to read data plots. An entirely different experiment would need to be designed to assess this."

4. It is interesting that while there are many significant (from a statistical standpoint) results, both the abstract and conclusion state that "demographics have very little impact" based on the coefficient size, which makes me wonder how the sample size of 2,340 participants was arrived upon. If a lower "effect" would be deemed meaningful, why was there such a large sample? It is a bit unsatisfying to have the results show statistical significance and then post-hoc have the authors state that the point estimate was too small to be meaningful -- perhaps choosing a test that takes into account both the scientific and statistical significance would make this more compelling. For example, something like using an equivalence zone that is deemed scientifically unimportant.

Author response: *Agreed, it could be done differently, but this approach is simple and familiar. The statistical significance is meaningful and it encourages a discussion of effect size and practical significance.* XXX DC thinks this needed re-writing **When you say the variability is in lineup difficulty, it might be better to state it as in the random effects, in various places in the paper.**

5. The authors state "Notably, countries like Canada, Romania, the United Kingdom, and Macedonia also had notable participant representation." -- it is not clear what this means, does this mean there were many individuals from these countries? If this is important information for the reader, it is not easy to glean it or read from Figure 5 as shown (or even just adding the N next to each of these countries in the initial sentence, it just was unclear why this was included).

Author response: *This is simply to illustrate examples of countries from which we did not expect to receive much participants via Mechanical Turk. The first paragraph of section 3.1 has been revised to address this concern and clarify the discussion.*

6. I think it is instructive to layout a specific example (the 18-25 year old female undergraduate vs someone with some graduate course work), although I'm not sure the basic equation for the difference is necessary to include in the paper.

Author response: *We have removed the detailed of the calculation on page 6 and report only the final number.*

7. The authors state that participants who performed poorly on an initial test were excluded from the analysis - it would be interesting to see a table of the demographics of this excluded population compared to the included population, particularly as it is possible that researchers using "visual inference" may not always include this step (and perhaps it is crucial for the assumption that the demographic impact is small?)

Author response: *Online experiments can sometimes attract insincere participants who do not engage seriously with the task. To address this, an initial screening test was used to identify and exclude such individuals from the analysis. There were insufficient numbers to warrant additional analysis of this group.*

## Reviewer: 2

### Comments to the Author(s)

Summary:
The paper investigates the effect of demographic variables on visual statistical inference and finds that demographic variables do not explain the variability of the main outcome, detection rate, well. Moreover, learning over time increases speed but not accuracy. I really like the paper. It is well written, I love the visual abstract, and the findings are strong, and
for some findings effect sizes are given. In my opinion, the paper could be improved by changing the title: “sociological factors” is not a term I have heard before and it paints an entire discipline with an overly broad brush. Likewise, I would not argue anything about “cognitive skills” (abstract) given that you have no measures of cognitive ability. I enjoyed
reviewing this paper.

### Individual points:

1)Title: The title strikes me as odd. What are sociological factors? As far as I can tell none of the authors is a sociologist. The word “sociological” only appears in the abstract. I argue for a different title.

Author response: *The title has been changed in the revised manuscript. The new title is 'Effect of Human Factors on Visual Statistical Inference,' where the word ‘Sociological' has been replaced with ‘Human.'*

2) Abstract: “The effectiveness of this method, measured by power, depends on combining evaluations from multiple observers. ”
I do not understand what this means. My association with “power” is sample size calculations; I am not sure what this means here. I see you are alluding to a definition of power in line 37 in a different paper (without explanation) and then define it on page 2 as the detection rate. I think the abstract would be easier to understand if you used “detection
rate” or you explained what power means.

Author response: *The abstract has been revised to explain the concept of power and its relationship to the detection or identification of the observed plot in a lineup. Additionally, the introduction now includes further discussion to clarify why detection rate is used to assess the impact of human factors on visual statistical inference.*

3) Abstract: “Factors influencing power include observer demographics, visual skills, experience, the sample of null plots, and signal strength. ” This sentence does not move the narrative much. Also, it is not clear to me whether “visual skill “and “individual skill” in the next sentence are the same. The number of factors you list only partially matches the
factors listed in section 2.

Author response: *The abstract has been revised to focus on the human factors examined in this paper.*

4) >There were a high number of female participants (42%) across most countries This sounds a bit funny – do you mean in the sample there were at least 42% females for most countries?

Author response: *Overall, female participants make up 42%. The manuscript has been revised to clarify this.*

5) >Notably, countries like Canada, Romania, the United Kingdom, and Macedonia also had notable participant representation
I don’t know what “notable participant representation” means.

Author response: *The manuscript has been revised to include a list of countries, other than the United States and India, where participants were located, highlighting the geographical diversity of the sample.*

6) Figure 1: I did not understand how you get a box plot out of an indicator variable for Detection (yes/no). May be it is not an indicator variable but an average of multiple detections? (there are 10 tasks listed in Table 6)

Author response: *The detection rate in Figure 1 refers to the proportion of times the observed plot is correctly identified in a lineup by each demographic factor group. It is a continuous variable ranging from 0 to 1. This measure is not derived from the 10 experiments shown in Table 6; rather, detection rates are calculated for each combination of demographic factors and lineups. For example, participants identified as male evaluate multiple lineups, resulting in multiple detection rate values corresponding to the lineups they assess.*

*Figure 1 caption has been revised to clarify how the detection rate is calculated. The revised caption includes “proportion of correct responses (Detection rate) across all lineups, aggregated for each level of the demographic factors."*

7) Figure1: “Variability in detection rate indicates large variability in lineup difficulties. ” This is a very important finding. It took me a while to understand your argument: Maybe you argue because the variability is huge, it must be due to a factor other than the ones shown (demographics), and that factor must be lineup difficulty. I am not confident that this
conclusion necessarily follows. Would it not be easier to extend the Figure by 3 columns with the 3 levels of difficulty? If you are right -- and I believe you are – then I would expect three much narrower box plots and the difference to the demographic variables would be more obvious.

Author response: *Lineup difficulties were predetermined as part of the experimental design, and each participant was randomly assigned 10 lineups spanning a range of difficulty levels. Figure 1 demonstrates that participants from each demographic group were exposed to lineups with varying difficulty levels. This is further illustrated in Figure 2(B), where  three levels of lineup difficulty are shown more distinctly.*

The fitted model accounts for lineup-level variation through random effects, effectively adjusting for lineup difficulty when estimating the effects of covariates on detection rates. That is why In Table 1 we notice that lineup specific variability is estimated as 2.27 which is much higher than any of the other parameter estimates in model A.1.

8) Conclusion: page 10 line 7-10 You give significant effects. For India you mention the comparison group (USA), for age and
for people with graduate degrees you do not mention it. It would also be helpful to give the percentage differences in parenthesis, e.g. “India shows significantly lower detection rate compared to United States (xx% vs xx%)”.

Author response: *The Table 1 shows the baseline levels for each category and the other levels are compared with the baseline. Significance of detection rates are discussed in details in section 3.2 based on the estimated model parameters in Table 1. The conclusion has been revised as suggested and clarifies the reported results. Should we report detection rate compared to United States (xx% vs xx%)? Ok, just report some of these in terms of log odds.*

9) Discussion: I am in particularly intrigued by the finding that experience improves speed but not accuracy. I wonder whether other researchers have found something similar in a different context. This may or may not be easy to answer in a couple of sentences – so please use your judgment.

Author response: *The ten experiments included in this research cover a variety of scenarios and contexts, as shown in Table 6. These diverse visual inferential tasks provide strong evidence that individuals’ experience with lineup evaluations leads to improved speed, but not necessarily improved accuracy in identifying the data plot. This suggests that familiarity with the task may help streamline the evaluation process without enhancing the ability to detect the true signal. While this finding is specific to visual inference tasks, it aligns with broader cognitive research where practice often improves task efficiency more than accuracy.*

*This suggests that familiarity with the task may help streamline the evaluation process without enhancing the ability to detect the true signal. Similar patterns have been observed in cognitive psychology, where repeated task exposure often improves response time without significantly affecting accuracy (Logan, 1988; Schneider & Shiffrin, 1977). We acknowledge that further exploration and comparison with findings from other domains would be valuable and encourage future work in this direction. Should we include this discussion in the revised manuscript?* XXX To make additional discussion.

### 2025-04-17 wiley_paper/response-to-review.qmd
Update response-to-review.qmd

@@
 
-Author response: *A joint model is fitted to simultaneously analyze detection accuracy and log time taken.
+Author response: *A joint model is fitted to simultaneously analyze detection rate and log time taken.
 The estimated covariate effects in the joint model are largely consistent with those obtained from the separate models, with one notable difference: the estimated variance of the lineup-level random intercept for detection rate (lieup difficulty) is slightly higher in the joint model (2.37) compared to the separate model (2.27), suggesting that modeling the outcomes jointly may better account for heterogeneity across lineups.
@@
 For this reason, separate analyses by lineup difficulty are not the focus of this study?* XXX DC suggests changing the caption further: "Some differences in means of various demographic factors are observed.
-Variability in detection rate indicates large variability in lineup difficulties." to "Some differences in means of various demographic factors are observed, but the variation is large, which we will see is primarily due to difference in the controlled factor, lineup difficulty."
+Variability in detection rate indicates large variability in lineup difficulties." to "Some differences in means of various demographic factors are observed, but the variation is large, which we will see is primarily due to difference in the controlled factor, lineup difficulty." DONE.
 

### 2025-04-17 wiley_paper/response-to-review.qmd
Update response-to-review.qmd

@@
 It is insufficient to only use one of these difficulty levels to assess the learning trend.* XXX **We have added additional discussion on this point at the end of the paper.** Change this sentence in conclusion: "However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea, at least for short term learning." to "Although this study showed that there was not a learning effect in the 10 lineups presented to each subject, we do hope that there is a possibility that lineups might be used for helping people to learn to read data plots.
-An entirely different experiment would need to be designed to assess this."
+An entirely different experiment would need to be designed to assess this." DONE
 
@@
 Author response: *Agreed, it could be done differently, but this approach is simple and familiar.
-The statistical significance is meaningful and it encourages a discussion of effect size and practical significance.* XXX DC thinks this needed re-writing **When you say the variability is in lineup difficulty, it might be better to state it as in the random effects, in various places in the paper.**
+The statistical significance is meaningful and it encourages a discussion of effect size and practical significance.* XXX DC thinks this needed re-writing **When you say the variability is in lineup difficulty, it might be better to state it as in the random effects, in various places in the paper.** DONE
 

### 2025-04-17 wiley_paper/response-to-review.qmd
Update response-to-review.qmd

@@
 
-Author response: *We have removed the detailed of the calculation on page 6 and report only the final number.*
+Author response: *We have removed the detailed calculation on page 6 and report only the final number.*
 
@@
 *Figure 1 caption has been revised to clarify how the detection rate is calculated.
-The revised caption includes “proportion of correct responses (Detection rate) across all lineups, aggregated for each level of the demographic factors."*
+The revised caption includes “proportion of data identifications (detection rate) across all lineups, aggregated for each level of the demographic factors."*
 

### 2025-04-21 wiley_paper/response-to-review.qmd
Update response-to-review.qmd

@@
 Significance of detection rates are discussed in details in section 3.2 based on the estimated model parameters in Table 1.
-The conclusion has been revised as suggested and clarifies the reported results.
-Should we report detection rate compared to United States (xx% vs xx%)?
-Ok, just report some of these in terms of log odds.*
+The conclusion has been revised as suggested and now clarifies the reported results by including detection rates for comparison.
 

### 2025-04-21 wiley_paper/response-to-review.qmd
Update response-to-review.qmd

@@
 
-Author response: *The ten experiments included in this research cover a variety of scenarios and contexts, as shown in Table 6.
-These diverse visual inferential tasks provide strong evidence that individuals’ experience with lineup evaluations leads to improved speed, but not necessarily improved accuracy in identifying the data plot.
-This suggests that familiarity with the task may help streamline the evaluation process without enhancing the ability to detect the true signal.
-While this finding is specific to visual inference tasks, it aligns with broader cognitive research where practice often improves task efficiency more than accuracy.*
-
-*This suggests that familiarity with the task may help streamline the evaluation process without enhancing the ability to detect the true signal.
-Similar patterns have been observed in cognitive psychology, where repeated task exposure often improves response time without significantly affecting accuracy (Logan, 1988; Schneider & Shiffrin, 1977).
-We acknowledge that further exploration and comparison with findings from other domains would be valuable and encourage future work in this direction.
-Should we include this discussion in the revised manuscript?* XXX To make additional discussion.
+Author response: *The finding that experience improves speed but not accuracy in lineup evaluations is consistent with prior research in different contexts.
+For example, Hong and Williamson (2008) showed that participants trained on a simple psychomotor task exhibited significant improvements in all speed-related measures, while their error rates remained largely unchanged.
+This discussion has now been added to the conclusion section.

### 2025-04-21 wiley_paper/response-to-review.qmd
Add files via upload

@@
 The fitted model accounts for lineup-level variation through random effects, effectively adjusting for lineup difficulty when estimating the effects of covariates on detection rates.
-For this reason, separate analyses by lineup difficulty are not the focus of this study?* XXX DC suggests changing the caption further: "Some differences in means of various demographic factors are observed.
-Variability in detection rate indicates large variability in lineup difficulties." to "Some differences in means of various demographic factors are observed, but the variation is large, which we will see is primarily due to difference in the controlled factor, lineup difficulty." DONE.
+For this reason, separate analyses by lineup difficulty are not the focus of this study.*
 
@@
 Here we are assessing whether the human evaluation of lineups is consistent across demographic factors and trials.
-Each subject evaluated 10 lineups, three of each difficulty level.
-It is insufficient to only use one of these difficulty levels to assess the learning trend.* XXX **We have added additional discussion on this point at the end of the paper.** Change this sentence in conclusion: "However, we had hoped that lineups may be a useful teaching tool, to improve students ability to read data plots, and the lack of a learning trend diminishes this idea, at least for short term learning." to "Although this study showed that there was not a learning effect in the 10 lineups presented to each subject, we do hope that there is a possibility that lineups might be used for helping people to learn to read data plots.
-An entirely different experiment would need to be designed to assess this." DONE
+Each subject evaluated 10 lineups, at least three of each difficulty level.
+It is insufficient to only use one of these difficulty levels to assess the learning trend.*
 
@@
 Author response: *Agreed, it could be done differently, but this approach is simple and familiar.
-The statistical significance is meaningful and it encourages a discussion of effect size and practical significance.* XXX DC thinks this needed re-writing **When you say the variability is in lineup difficulty, it might be better to state it as in the random effects, in various places in the paper.** DONE
+The statistical significance is meaningful and it encourages a discussion of effect size and practical significance.*
 
@@
 I see you are alluding to a definition of power in line 37 in a different paper (without explanation) and then define it on page 2 as the detection rate.
-I think the abstract would be easier to understand if you used “detection
-rate” or you explained what power means.
+I think the abstract would be easier to understand if you used “detection rate” or you explained what power means.
 
@@
 
-The fitted model accounts for lineup-level variation through random effects, effectively adjusting for lineup difficulty when estimating the effects of covariates on detection rates.
-That is why In Table 1 we notice that lineup specific variability is estimated as 2.27 which is much higher than any of the other parameter estimates in model A.1.
+*The fitted model accounts for lineup-level variation through random effects, effectively adjusting for lineup difficulty when estimating the effects of covariates on detection rates.
+That is why In Table 1 we notice that lineup specific variability is estimated as 2.27 which is much higher than any of the other parameter estimates in model A.1.*
 
@@
 Significance of detection rates are discussed in details in section 3.2 based on the estimated model parameters in Table 1.
-The conclusion has been revised as suggested and now clarifies the reported results by including detection rates for comparison.
+The conclusion has been revised as suggested and now clarifies the reported results by including detection rates for comparison.*
 
@@
 For example, Hong and Williamson (2008) showed that participants trained on a simple psychomotor task exhibited significant improvements in all speed-related measures, while their error rates remained largely unchanged.
-This discussion has now been added to the conclusion section.
+This discussion has now been added to the conclusion section.*

### 2025-06-04 wiley_paper/review2/response-to-review.qmd
Add files via upload

--- 2025-04-21 wiley_paper/response-to-review.qmd
@@
 title: "Response to Review"
-author: "Majumder, Cook, Hofmann"
+author: "Majumder, Hofmann, Cook"
 format: pdf
@@
 
-Thank you for the careful and positive reviews. *Our response to comments is written in italics*.
+Thank you for the careful and positive reviews.
+*Our response to comments is written in italics*.
+No changes were made to the main text in this revision.
+However, we have revised Figures 2, 3, and 4 by enlarging the axis labels for improved readability, as suggested by the reviewer.
 
@@
 
-Good set of comments. FYI, feel free to change the review title (if you wish)
-and add more material (as an on-line journal, page limits not really active).
-Look forward to seeing your revision.   D Scott co-editor  scottdw@rice.edu
+FYI, your revision is acceptable, but Wiley rules do not allow for
+any changes after full acceptance. So look at the 2 comments, and
+make appropriate changes, submit, and I will handle the final details.
 
-Response: *Thank you*
+Cheers, David Scott, co-editor
 
-I am writing in regards to Manuscript ID EOCS-756 entitled "Effect of Sociological Factors on Visual Statistical Inference," which you submitted to WIREs Computational Statistics.
-The reviewer(s) have suggested major revisions to improve the manuscript.
+Response: *Thank you for the clarification.*
+
+I am writing in regards to Manuscript ID EOCS-756.R1 entitled "Effect of Human Factors on Visual Statistical Inference," which you submitted to WIREs Computational Statistics.
+The reviewer(s) have recommended publication, but also suggest some minor revisions.
 Their comments are included at the bottom of this letter; I invite you to respond to them and revise your manuscript.
 
-Please submit your revision within 6 weeks. If you need extra time, please let us know so that we can keep our systems updated for editorial planning purposes.
+Please submit your revision within 3 weeks. If you need extra time, please let us know so that we can keep our systems updated for editorial planning purposes.
 
 A revised version of your manuscript that takes into account the comments of the reviewer(s) will be reconsidered for publication.
-Please note that revising your manuscript does not guarantee eventual acceptance, and that your revision will be subject to re-review before a decision is rendered.
 
@@
 
-https://urldefense.com/v3/__https://mc.manuscriptcentral.com/compstats?URL_MASK=438fdf874ee14f659af35cd0e26d4ced__;!!PvXuogZ4sRB2p-tU!GOtJWWHikD4k_JosJN5bkg2Kphby7E2KmkmqHJ4dZ95gyJOEY62FgPhrkDJhfcMHLuoagvllP182luVdjkcYTREEH-IXMjM$ .
+https://urldefense.com/v3/__https://mc.manuscriptcentral.com/compstats?URL_MASK=3014fa56294b4d20905fcb2c8e0b8338__;!!PvXuogZ4sRB2p-tU!Htx4nGnItiXf30FBpb2ENpauRGwVrxiScnsEX1RPxAuEH-6s4AYg-nHnv9J_VYI3zmALTmoZm_xxcrm8BE4Te2ZCzDcZQEc$ .
 
@@
 ### Comments to the Author(s)
+Good revision. 2 minor comments below
 
-This is an interesting study where the authors attempt to examine the impact of demographics on evaluating a "line up" of statistical graphics to visually pick out a particular data set compared to several null graphs.
-The main results seem to be that while demographics do seem to vary the results from a statistical perspective (i.e.
-the p-values are significant, given the large sample size), the impact from a scientifically meaningful perspective is minimal.
-I have a several specific comments to highlight below.
-
-Author response:  *The authors agree with this comment. The results support our conclusion that the visual test is robust to variations in human factors.*
+Author response:  *We sincerely thank the reviewer for the positive feedback and for acknowledging the revision.
+We appreciate the two minor comments provided and have addressed them carefully in the revised manuscript.*
 
@@
 
-1.
-The authors choose to examine two outcomes, log time (how long it takes the user to complete the task) and detection rate.
-It is not clear to me why a joint model is not shown.
-It seems relevant, particularly because there are clear demographic differences seen (albeit quantitatively small) that it would make sense to examine detection rate with the demographics *and* time in model A.1 to see whether, for example, the demographic differences seen in A.1 are fully explained by time as there are definitely impacts of demographics on time and it seems plausible that time would impact accuracy.
-Likewise, it is possible that after adjusting for time a learning trend may be more visible.
+1) Figure 2, Figure 3 : It would help if you could make the axes labels bigger
 
-Author response: *A joint model is fitted to simultaneously analyze detection rate and log time taken.
-The estimated covariate effects in the joint model are largely consistent with those obtained from the separate models, with one notable difference: the estimated variance of the lineup-level random intercept for detection rate (lieup difficulty) is slightly higher in the joint model (2.37) compared to the separate model (2.27), suggesting that modeling the outcomes jointly may better account for heterogeneity across lineups.
-Importantly, the joint model also provides an estimate of the latent correlation between the lineup-level random intercepts for time taken and detection rate.
-This correlation is estimated to be -0.55, indicating that lineups which take longer to evaluate tend to be associated with lower detection rates.
-This result aligns with intuition: more difficult lineups likely require more evaluation time, yet are also more prone to non-detection.
-These findings and interpretations have been incorporated into the revised manuscript at the end of section 3.2.*
+Author response: *Thank you for the helpful suggestion. We have revised Figures 2, 3, and 4 to enlarge the axis labels and improve overall readability.*
 
-2.
-The authors state that the largest effect is by far the lineup difficulty.
-This mixture of difficulties makes it hard to learn much from the overall results (e.g.
-it makes the bottom portion of Figure 1 very hard to parse).
-Figure 2 (B) seems to show a the impact of a variable (education) across three different classifications of "difficulty" -- from this plot it seems that the easy line ups and difficult line ups are providing very little information, as the proportion classified correctly is very close to 1 or 0 (at least for the 18-25 year old females this plot was drawn for).
-It may be more compelling to show the results excluding these and focusing only on the medium lineups (at least as a subgroup analysis).
+2) Title: Thank you for changing the title.
+Although better, I still find "Human factors" slightly confusing.
+An alternative would be "Socio-demographic correlates of visual inference".
+You can leave your title as is ; reviewers should not dictate the exact title and maybe your reading of the literature is different than mine.
 
-Author response: *Lineup difficulties were predetermined as part of the experimental design, and each participant was randomly assigned 10 lineups spanning a range of difficulty levels.
-The fitted model accounts for lineup-level variation through random effects, effectively adjusting for lineup difficulty when estimating the effects of covariates on detection rates.
-For this reason, separate analyses by lineup difficulty are not the focus of this study.*
-
-3.
-Building on the previous comment, it is possible (likely?) that the reason there is no "learning trend" seen is due to these very easy or very hard lineups.
-I would like to see a model for this that only includes the medium lineups.
-
-Author response: *The model accounts for lineup difficulty when estimating the learning trend.
-It would be interesting to examine learning trend more deeply, in particular in relation to using lineups for teaching people how to read data plots.
-However, learning trend in itself is not the focus of this paper.
-Here we are assessing whether the human evaluation of lineups is consistent across demographic factors and trials.
-Each subject evaluated 10 lineups, at least three of each difficulty level.
-It is insufficient to only use one of these difficulty levels to assess the learning trend.*
-
-4.
-It is interesting that while there are many significant (from a statistical standpoint) results, both the abstract and conclusion state that "demographics have very little impact" based on the coefficient size, which makes me wonder how the sample size of 2,340 participants was arrived upon.
-If a lower "effect" would be deemed meaningful, why was there such a large sample?
-It is a bit unsatisfying to have the results show statistical significance and then post-hoc have the authors state that the point estimate was too small to be meaningful -- perhaps choosing a test that takes into account both the scientific and statistical significance would make this more compelling.
-For example, something like using an equivalence zone that is deemed scientifically unimportant.
-
-Author response: *Agreed, it could be done differently, but this approach is simple and familiar.
-The statistical significance is meaningful and it encourages a discussion of effect size and practical significance.*
-
-5.
-The authors state "Notably, countries like Canada, Romania, the United Kingdom, and Macedonia also had notable participant representation." -- it is not clear what this means, does this mean there were many individuals from these countries?
-If this is important information for the reader, it is not easy to glean it or read from Figure 5 as shown (or even just adding the N next to each of these countries in the initial sentence, it just was unclear why this was included).
-
-Author response: *This is simply to illustrate examples of countries from which we did not expect to receive much participants via Mechanical Turk.
-The first paragraph of section 3.1 has been revised to address this concern and clarify the discussion.*
-
-6.
-I think it is instructive to layout a specific example (the 18-25 year old female undergraduate vs someone with some graduate course work), although I'm not sure the basic equation for the difference is necessary to include in the paper.
-
-Author response: *We have removed the detailed calculation on page 6 and report only the final number.*
-
-7.
-The authors state that participants who performed poorly on an initial test were excluded from the analysis - it would be interesting to see a table of the demographics of this excluded population compared to the included population, particularly as it is possible that researchers using "visual inference" may not always include this step (and perhaps it is crucial for the assumption that the demographic impact is small?)
-
-Author response: *Online experiments can sometimes attract insincere participants who do not engage seriously with the task.
-To address this, an initial screening test was used to identify and exclude such individuals from the analysis.
-There were insufficient numbers to warrant additional analysis of this group.*
-
-## Reviewer: 2
-
-### Comments to the Author(s)
-
-Summary:
-The paper investigates the effect of demographic variables on visual statistical inference and finds that demographic variables do not explain the variability of the main outcome, detection rate, well.
-Moreover, learning over time increases speed but not accuracy.
-I really like the paper.
-It is well written, I love the visual abstract, and the findings are strong, and
-for some findings effect sizes are given.
-In my opinion, the paper could be improved by changing the title: “sociological factors” is not a term I have heard before and it paints an entire discipline with an overly broad brush.
-Likewise, I would not argue anything about “cognitive skills” (abstract) given that you have no measures of cognitive ability.
-I enjoyed
-reviewing this paper.
-
-### Individual points:
-
-1)Title: The title strikes me as odd.
-What are sociological factors?
-As far as I can tell none of the authors is a sociologist.
-The word “sociological” only appears in the abstract.
-I argue for a different title.
-
-Author response: *The title has been changed in the revised manuscript.
-The new title is 'Effect of Human Factors on Visual Statistical Inference,' where the word ‘Sociological' has been replaced with ‘Human.'*
-
-2) Abstract: “The effectiveness of this method, measured by power, depends on combining evaluations from multiple observers. ”
-I do not understand what this means.
-My association with “power” is sample size calculations; I am not sure what this means here.
-I see you are alluding to a definition of power in line 37 in a different paper (without explanation) and then define it on page 2 as the detection rate.
-I think the abstract would be easier to understand if you used “detection rate” or you explained what power means.
-
-Author response: *The abstract has been revised to explain the concept of power and its relationship to the detection or identification of the observed plot in a lineup.
-Additionally, the introduction now includes further discussion to clarify why detection rate is used to assess the impact of human factors on visual statistical inference.*
-
-3) Abstract: “Factors influencing power include observer demographics, visual skills, experience, the sample of null plots, and signal strength.
-” This sentence does not move the narrative much.
-Also, it is not clear to me whether “visual skill “and “individual skill” in the next sentence are the same.
-The number of factors you list only partially matches the
-factors listed in section 2.
-
-Author response: *The abstract has been revised to focus on the human factors examined in this paper.*
-
-4) >There were a high number of female participants (42%) across most countries This sounds a bit funny – do you mean in the sample there were at least 42% females for most countries?
-
-Author response: *Overall, female participants make up 42%. The manuscript has been revised to clarify this.*
-
-5) >Notably, countries like Canada, Romania, the United Kingdom, and Macedonia also had notable participant representation
-I don’t know what “notable participant representation” means.
-
-Author response: *The manuscript has been revised to include a list of countries, other than the United States and India, where participants were located, highlighting the geographical diversity of the sample.*
-
-6) Figure 1: I did not understand how you get a box plot out of an indicator variable for Detection (yes/no).
-May be it is not an indicator variable but an average of multiple detections?
-(there are 10 tasks listed in Table 6)
-
-Author response: *The detection rate in Figure 1 refers to the proportion of times the observed plot is correctly identified in a lineup by each demographic factor group.
-It is a continuous variable ranging from 0 to 1.
-This measure is not derived from the 10 experiments shown in Table 6; rather, detection rates are calculated for each combination of demographic factors and lineups.
-For example, participants identified as male evaluate multiple lineups, resulting in multiple detection rate values corresponding to the lineups they assess.*
-
-*Figure 1 caption has been revised to clarify how the detection rate is calculated.
-The revised caption includes “proportion of data identifications (detection rate) across all lineups, aggregated for each level of the demographic factors."*
-
-7) Figure1: “Variability in detection rate indicates large variability in lineup difficulties.
-” This is a very important finding.
-It took me a while to understand your argument: Maybe you argue because the variability is huge, it must be due to a factor other than the ones shown (demographics), and that factor must be lineup difficulty.
-I am not confident that this
-conclusion necessarily follows.
-Would it not be easier to extend the Figure by 3 columns with the 3 levels of difficulty?
-If you are right -- and I believe you are – then I would expect three much narrower box plots and the difference to the demographic variables would be more obvious.
-
-Author response: *Lineup difficulties were predetermined as part of the experimental design, and each participant was randomly assigned 10 lineups spanning a range of difficulty levels.
-Figure 1 demonstrates that participants from each demographic group were exposed to lineups with varying difficulty levels.
-This is further illustrated in Figure 2(B), where  three levels of lineup difficulty are shown more distinctly.*
-
-*The fitted model accounts for lineup-level variation through random effects, effectively adjusting for lineup difficulty when estimating the effects of covariates on detection rates.
-That is why In Table 1 we notice that lineup specific variability is estimated as 2.27 which is much higher than any of the other parameter estimates in model A.1.*
-
-8) Conclusion: page 10 line 7-10 You give significant effects. For India you mention the comparison group (USA), for age and
-for people with graduate degrees you do not mention it.
-It would also be helpful to give the percentage differences in parenthesis, e.g.
-“India shows significantly lower detection rate compared to United States (xx% vs xx%)”.
-
-Author response: *The Table 1 shows the baseline levels for each category and the other levels are compared with the baseline.
-Significance of detection rates are discussed in details in section 3.2 based on the estimated model parameters in Table 1.
-The conclusion has been revised as suggested and now clarifies the reported results by including detection rates for comparison.*
-
-9) Discussion: I am in particularly intrigued by the finding that experience improves speed but not accuracy.
-I wonder whether other researchers have found something similar in a different context.
-This may or may not be easy to answer in a couple of sentences – so please use your judgment.
-
-Author response: *The finding that experience improves speed but not accuracy in lineup evaluations is consistent with prior research in different contexts.
-For example, Hong and Williamson (2008) showed that participants trained on a simple psychomotor task exhibited significant improvements in all speed-related measures, while their error rates remained largely unchanged.
-This discussion has now been added to the conclusion section.*
+Author response: *We appreciate your suggestion and understand your perspective.
+We have updated the title in the last revised manuscript to "Effect of Human Factors on Visual Statistical Inference." We chose the term "Human Factors" to encompass a broader range of individual-level influences, including but not limited to socio-demographic variables.
+We acknowledge that interpretations may vary, and we appreciate your flexibility in allowing us to retain our chosen framing.*
